### 2014-07-03 low-or-no-cost-abstract.tex
First draft to be completed

\documentclass{svmult}

\usepackage{mathptmx,helvet,courier,graphicx,multicol}

\begin{document}

\title*{Low or no cost distributed evolutionary computation}
\author{Juan J. Merelo-Guervós}
\institute{GeNeura Group (\url{http://geneura.wordpress.com}), Department of Computer Architecture (\url{http://atc.ugr.es}) and CITIC (\url{http://citic.ugr.es}), University of Granada (Spain) (\url{http://www.ugr.es})}

\maketitle

\abstract{From the era of big science we are back to the "do it yourself", where you don't have any money to buy clusters or subscribe to grids but still have algorithms that crave many computing nodes and need them to measure scalability. Fortunately, this coincides with the era of big data, cloud computing, and browsers that include JavaScript virtual machines. Those are the reason why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilites and operate in complete asynchrony of each other.}

\end{document}

### 2014-07-08 low-or-no-cost-abstract.tex
Version sent to the conference

@@
 
-\abstract{From the era of big science we are back to the "do it yourself", where you don't have any money to buy clusters or subscribe to grids but still have algorithms that crave many computing nodes and need them to measure scalability.
+\abstract{From the era of big science we are back to the "do it yourself", where you do not have any money to buy clusters or subscribe to grids but still have algorithms that crave many computing nodes and need them to measure scalability.
 Fortunately, this coincides with the era of big data, cloud computing, and browsers that include JavaScript virtual machines.
@@
 
+\section{What is the point of low or no cost evolutionary algorithms}
+
+The world has computational resources in spades. Most of them do not belong to you or your lab. That does not mean you cannot use it. The problem is how.
+
+Most theory in parallel computing has been devoted to predict and optimize the performance where the number of nodes, their connections, and the time every one is devoting to the computation is known in advance.
+However, even if Big Science is not over, the era of Citizen science has started (with SETI@home and then BOINC) and it offers a vast amount of computational resources to attract, if only you know how.
+But there is a challenge: knowing, or at least having a ballpark, of how your algorithm is going to perform in this uncertain environment, where none of the factors is known: neither the number of nodes, through how they are connected, to how long are they going to be focused on doing what you want them to.
+
+Besides, since Amazon started selling EC2 several years ago, reliable and scalable computing resources are also available for a low price and on demand.
+Recently, Google has also refurbished its offering lowering their prices.
+This means that the conjunction of free or low-cost cloud computing engines, volunteer computing systems and and untapped capability of desktop systems can be used for creating massive, or at least potentially massive, distributed computing experiments.
+
+\section{Conclussion}
+
+In this talk we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly using evolutionary algorithms.
+
+There are many challenges involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}.
+But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.
+This, among other things, means that popularity will become directly performance of the metacomputer you create by attrating more users.
+
+\bibliographystyle{unsrt}
+\bibliography{geneura}
+
 \end{document}

### 2014-08-31 low-or-no-cost-abstract.tex
Some error

@@
 Fortunately, this coincides with the era of big data, cloud computing, and browsers that include JavaScript virtual machines.
-Those are the reason why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilites and operate in complete asynchrony of each other.}
+Those are the reason why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilities and operate in complete asynchrony of each other.}
 
@@
 
-\section{Conclussion}
+\section{Conclusion}
 
@@
 But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.
-This, among other things, means that popularity will become directly performance of the metacomputer you create by attrating more users.
+This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.
 

### 2014-08-31 low-or-no-cost-abstract.tex
Post-publication EAs

@@
 Fortunately, this coincides with the era of big data, cloud computing, and browsers that include JavaScript virtual machines.
-Those are the reason why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilities and operate in complete asynchrony of each other.}
+Those are the reasons why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilities and operate in complete asynchrony of each other.}
 
-\section{What is the point of low or no cost evolutionary algorithms}
+\section{What is the point of low or no cost evolutionary algorithms?}
 
@@
 
-Most theory in parallel computing has been devoted to predict and optimize the performance where the number of nodes, their connections, and the time every one is devoting to the computation is known in advance.
+Most theory in parallel computing has been devoted to predict and optimize the performance in systems where the number of nodes, their connections, and the time every one is devoting to the computation is known in advance.
 However, even if Big Science is not over, the era of Citizen science has started (with SETI@home and then BOINC) and it offers a vast amount of computational resources to attract, if only you know how.
@@
 
-In this talk we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly using evolutionary algorithms.
+In this talk we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly applied to the design of evolutionary algorithms.
 
-There are many challenges involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}.
+There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}.
 But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.

### 2015-02-05 modeling-volunteer-computing-evopar2015.tex
Paper I didn't start or finish

--- 2014-08-31 low-or-no-cost-abstract.tex
@@
-\documentclass{svmult}
+\documentclass[a4paper]{article}
+\usepackage{amssymb}
+\setcounter{tocdepth}{3}
+\usepackage{graphicx,epsfig}
+\usepackage{listings}
+\usepackage{rotating}
+\usepackage{subfigure}
+\usepackage[affil-it]{authblk}
 
-\usepackage{mathptmx,helvet,courier,graphicx,multicol}
+%%%%
+
+\usepackage{color}
+\usepackage{alltt}
+\usepackage{verbatim}
+\usepackage{url}
+
+\usepackage{url}
 
@@
 
-\title*{Low or no cost distributed evolutionary computation}
-\author{Juan J. Merelo-Guervós}
-\institute{GeNeura Group (\url{http://geneura.wordpress.com}), Department of Computer Architecture (\url{http://atc.ugr.es}) and CITIC (\url{http://citic.ugr.es}), University of Granada (Spain) (\url{http://www.ugr.es})}
+\title{Modeling browser-based evolutionary computation systems}
+%\author{Juan J. Merelo-Guervós}
+\author{Som E. One}
+\affil{Somewhere Institute}
 
@@
 
-\abstract{From the era of big science we are back to the "do it yourself", where you do not have any money to buy clusters or subscribe to grids but still have algorithms that crave many computing nodes and need them to measure scalability.
-Fortunately, this coincides with the era of big data, cloud computing, and browsers that include JavaScript virtual machines.
-Those are the reasons why this talk will focus on two different aspects of volunteer or freeriding computing: first, the pragmatic: where to find those resources, which ones can be used, what kind of support you have to give them; and then, the theoretical: how evolutionary algorithms can be adapted to a environment in which nodes come and go, have different computing capabilities and operate in complete asynchrony of each other.}
+\begin{abstract}
+From the era of big science we are back to the "do it yourself", where
+you do not have any money to buy clusters or subscribe to grids but
+still have algorithms that crave many computing nodes and need them to
+measure scalability. Fortunately, this coincides with the era of big
+data, cloud computing, and browsers that include JavaScript virtual
+machines. Those are the reasons why this talk will focus on two
+different aspects of volunteer or freeriding computing: first, the
+pragmatic: where to find those resources, which ones can be used, what
+kind of support you have to give them; and then, the theoretical: how
+evolutionary algorithms can be adapted to a environment in which nodes
+come and go, have different computing capabilities and operate in
+complete asynchrony of each other. 
+\end{abstract}
 
@@
 
-\bibliographystyle{unsrt}
-\bibliography{geneura}
+\bibliographystyle{splncs}
+\bibliography{geneura,volunteer}
 

### 2015-02-06 modeling-volunteer-computing-gecco.tex
Having a shot at it

--- 2015-02-05 modeling-volunteer-computing-evopar2015.tex
@@
-\documentclass[a4paper]{article}
+\documentclass{sig-alternate}
+
+\usepackage[latin1]{inputenc}
 \usepackage{amssymb}
 \setcounter{tocdepth}{3}
-\usepackage{graphicx,epsfig}
-\usepackage{listings}
-\usepackage{rotating}
+\usepackage{graphicx}
 \usepackage{subfigure}
-\usepackage[affil-it]{authblk}
-
-%%%%
-
-\usepackage{color}
-\usepackage{alltt}
-\usepackage{verbatim}
-\usepackage{url}
 
@@
 
-\title{Modeling browser-based evolutionary computation systems}
-%\author{Juan J. Merelo-Guervós}
-\author{Som E. One}
-\affil{Somewhere Institute}
+\title{Modeling browser-based distributed evolutionary computation systems}
+
+\numberofauthors{1} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Som E. One\\
+       \affaddr{Some Institute}\\
+       \email{em@a.il}
+}
 
@@
 
-\section{What is the point of low or no cost evolutionary algorithms?}
+\keywords{Evolutionary computation, volunteer computing, distributed
+  computing, node.js, javascript}
+
+\section{Introduction}
 
@@
 
-In this talk we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly applied to the design of evolutionary algorithms.
+In this paper we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly applied to the design of evolutionary algorithms.
 
@@
 
-\bibliographystyle{splncs}
+\bibliographystyle{abbrv}
 \bibliography{geneura,volunteer}

### 2015-02-06 modeling-volunteer-computing-gecco.tex
Here goes a compiling version

@@
 data, cloud computing, and browsers that include JavaScript virtual
-machines. Those are the reasons why this talk will focus on two
+machines. Those are the reasons why this paper will focus on two
 different aspects of volunteer or freeriding computing: first, the
@@
 come and go, have different computing capabilities and operate in
-complete asynchrony of each other. 
+complete asynchrony of each other. We will examine the setup needed to
+create a very simple distributed evolutionary algorithm using
+JavaScript and then find a model of how users react to it by
+collecting data from several experiments featuring different classical
+benchmark functions. 
 \end{abstract}
@@
 
-The world has computational resources in spades. Most of them do not belong to you or your lab. That does not mean you cannot use it. The problem is how.
+The world has computational resources in spades. Most of them do not
+belong to you or your lab. That does not mean you cannot use it. The
+problem is how.
 
-Most theory in parallel computing has been devoted to predict and optimize the performance in systems where the number of nodes, their connections, and the time every one is devoting to the computation is known in advance.
-However, even if Big Science is not over, the era of Citizen science has started (with SETI@home and then BOINC) and it offers a vast amount of computational resources to attract, if only you know how.
-But there is a challenge: knowing, or at least having a ballpark, of how your algorithm is going to perform in this uncertain environment, where none of the factors is known: neither the number of nodes, through how they are connected, to how long are they going to be focused on doing what you want them to.
+Most theory in parallel computing has been devoted to predict and
+optimize the performance in systems where the number of nodes, their
+connections, and the time every one is devoting to the computation is
+known in advance. However, even if Big Science is not really over and
+it is slated to come back, the era of
+Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
+offers a vast amount of computational resources to attract, if only
+you know how. But there is a challenge: knowing, or at least having a
+ballpark, of how your algorithm is going to perform in this uncertain
+environment, where none of the factors is known: neither the number of
+nodes, through how they are connected, to how long are they going to
+be focused on doing what you want them to. That is why some effort is
+needed to first understand the dynamics of the decision to participate
+in an experiment that requires you to click on a link and then stay
+for a while looking at the screen (or just leave it there running).
 
@@
 This means that the conjunction of free or low-cost cloud computing engines, volunteer computing systems and and untapped capability of desktop systems can be used for creating massive, or at least potentially massive, distributed computing experiments.
+
+\begin{figure*}[h!tb]
+  \centering
+  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
+  \label{fig:sshot}
+  \caption{Screenshot of the talk that includes at the graph
+    that shows the progression of the evolutionary
+    algorithm at the top right corner}
+\end{figure*}
 

### 2015-02-07 modeling-volunteer-computing-gecco.tex
Well, something done...2

@@
 
-Besides, since Amazon started selling EC2 several years ago, reliable and scalable computing resources are also available for a low price and on demand.
-Recently, Google has also refurbished its offering lowering their prices.
-This means that the conjunction of free or low-cost cloud computing engines, volunteer computing systems and and untapped capability of desktop systems can be used for creating massive, or at least potentially massive, distributed computing experiments.
+Besides, since Amazon started selling EC2 several years ago, reliable
+and scalable computing resources are also available for a low price
+and on demand. Recently, Google has also refurbished its offering
+lowering their prices. This means that the conjunction of free or
+low-cost cloud computing engines, volunteer computing systems and and
+untapped capability of desktop systems can be used for creating
+massive, or at least potentially massive, distributed computing
+experiments.
+
+The rest of the paper is organized as follows: next we will present
+the state of the art in volunteer computing and its modelization. We
+will proceed to describe the experimental setup in Section
+\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
+finally we will present our conclusions in Section \ref{sec:conc}
+along with future lines of work.
+
+\section{State of the art}
+\label{sec:soa}
+
+Using the web as a resource for distributed, or plainly user based,
+evolutionary computation has a long history since JavaScript was
+introduced as a browser-based language in 1997 and even before, when
+other procedures, including Flash animations, VBScript, ActiveX or Java applets were
+used. Java was pointed out in \cite{soares1998get} as a ``language for
+internet'' and
+\begin{quote}
+Java can bring some important advantages... solves architectural
+heterogeneity [...] easy to install and [...] built-in security
+mechanisms
+\end{quote}
+The same paper by Soares et al. describes JET, a system that supports
+the execution of parallel applications over the Internet. In fact the
+same paper describes some other systems working at the same time,
+except they cannot be used for science since they do not have a
+comprehensive statistics support like JET.
+
+However, Java (and all the other technologies) is not universal in the
+sense than an extra component, namely, the Java virtual machine, has
+to be installed in the browser. JavaScript
+\cite{flanagan2006javascript} was introduced in 1997 as a
+browser-embedded language and has, since them, become a set of standards
+\cite{ECMA-262} for the language, its components and future versions
+that have been widely adopted by the industry and also by scientists,
+who used them for creating a non-distributed evolutionary algorithm on
+the browser as early as 1998 \cite{jj-ppsn98}.
+
+The potential for volunteer computing using browsers was realized
+later on \cite{sarmenta-bayanihan} as well as the potential for
+mischief of users with code in their hands
+\cite{sarmenta-sabotagetolerance} however, these early efforts by
+Sarmenta once again used Java and not JavaScript, making this effort
+less universal. JavaScript can be used either for unwitting
+\cite{unwitting-ec} or volunteer
+\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
+evolutionary computation and has been used ever since by several
+authors, including recent efforts \cite{duda2013distributed} that even
+used the client's GPU \cite{duda2013gpu}.
+
+There have been relatively few efforts to analyze what is the
+performance that can be obtained from these volunteer computing
+effort. There was some effort initially to dodge the issue by making
+the algorithm adaptive to the kind of resources allotted to it
+\cite{sarmenta-bayanihan}, which is actually not such a big problem in
+algorithms such as the evolutionary algorithm that can easily be
+paralelized via population splitting or farming out evaluations to all
+the nodes available.
+
+\section{Methodology}
+\label{sec:exp}
 
@@
 \bibliographystyle{abbrv}
-\bibliography{geneura,volunteer}
+\bibliography{geneura,volunteer,javascript,ror-js}
 

### 2015-02-07 modeling-volunteer-computing-gecco.tex
Starting with methodologyÇ

@@
 massive, or at least potentially massive, distributed computing
-experiments.
+experiments. These experiments can be easily created using open-source
+repository sites like GitHub and deployed automatically to Platform as
+a Service products such as Heroku or OpenShift.
+
+In general, any volunteer computing experiment will have to be made
+``in the open''. The fact that somebody is giving you, basically for
+free, computing resources, means that the scientist using them has to
+give back. The baseline is releasing the source code used: all
+volunteer computing platforms, from SETI@home on, have done it. The
+opposite is probably the reason why many companies like PopularPower \cite{buyya2001compute}
+have folded or others like CrowdProcess have decided to turn their
+product to in-house computation: there must be a mutual relationship
+of trust among the scientist and the person that is running his/her
+code in the browser. As has been mentioned in the early stages of what
+was then called {\em desktop grid computing}, \cite{gc:bausch}, in
+fact CPU cycle selling might not make economic sense since there are
+not so much demand for it and supply is quite high. However, while
+potential supply is in fact huge, {\em actual} supply will depend on
+the person holding it willing to actually allocate it to a particular
+company selling it or a particular experiment needing it, not to
+mention the fact that the experiment {\em actually} has to draw the
+attention of the supplier. That is why trust is essential, and using
+free software might not be enough: Openness
+has to progress from open source code to open science: releasing all
+data obtained in the experiment in a repository such as GitHub,
+mentioned above, and even allowing real-time access to experiment data
+to users.
+
+Another possible reason of the failure of former companies to create a
+for-profit desktop grid might be the lack of a way to predict what is
+going to be that supply in a particular moment. It is impossible to
+predict, in advance, to know how many persons  are going to visit a
+particular website. Even if you pull all the resources you have and
+they lie across continents and time zones, the number of cycles
+apportioned to a particular experiment will depend not only on the
+users lending their web real state to the experiments (which is
+usually the case for cycle brokerage companies) but especially to
+users going to a web site and spending some time on them. Even if it
+is theoretically impossible to predict to a high precision what
+happens, it is in practice possible to approximate the number of
+visits in a site, at least in a particular one, using time series. But
+in the short term and using a more general model, there is still need
+to model the behavior of users so that more factors can be added to
+the model other than the time series. This user behavior, of course,
+presupposes that you are respecting their anonymity and privacy (not using
+cookies, for instance) and that you are respecting the open approach
+mentioned above. All computation can be done without the knowledge of
+the user \cite{unwitting-ec}, but this would work against openness
+that would, curiously enough, result in a huge decrease of the future
+performance of any experiment you might want to perform.
+
+These are best practices that have been followed in the experiments in
+this paper, that first presents the first versions of a platform for
+distributed evolutionary algorithms using the browser and a free (as
+in free beer) backend, and second, shows the result from an
+statistical point of view in order to put the basis for a model of the
+metacomputer obtained by joining all volunteers and the free backend
+used for the experiment. This is not an exhaustive or complete
+exploration of the possibilities of this kind of computation, but it
+is enough to present the free software used to perform the experiments
+and will allow us to describe, in general, the behavior of the users
+as well as the performance achieved on the experiments done so far,
+which should show some advantage over doing the same kind of
+experiments locally using available resources.
 
@@
 evolutionary computation and has been used ever since by several
-authors, including recent efforts \cite{duda2013distributed} that even
-used the client's GPU \cite{duda2013gpu}.
+authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
+used the client's GPU \cite{duda2013gpu}. Many other researchers have
+used Java \cite{chong:1999:jDGPi} and others have gone away from the
+server-based paradigm to embrace peer to peer systems
+\cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
+platform avoid single points of failure (the server) but, since they
+need a certain amount of infrastructure installed to start, the
+threshold to join them is much lower.
 
@@
 the algorithm adaptive to the kind of resources allotted to it
-\cite{sarmenta-bayanihan}, which is actually not such a big problem in
+\cite{milani2004online}, which is actually not such a big problem in
 algorithms such as the evolutionary algorithm that can easily be
 paralelized via population splitting or farming out evaluations to all
-the nodes available.
+the nodes available. Lately, several approaches have focused on the
+fault-tolerance of volunteer algorithms
+\cite{gonzalez2010characterizing} which can, of course, be studied in
+a more general distributed computing context
+\cite{nogueras2015studying} or including it in a more general study of
+performance of the evolutionary algorithm itself
+\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
+volunteer computing, number of users and the time spent in the
+computation in browser-based volunteer computing experiments, have only been studied in a limited way in
+\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
+run. Studies using volunteer computing platforms such as SETI@home
+\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
+modeled quite well the availability of resources in several clusters
+of that framework; the shape of those distributions is a skewed bell
+with more resources in the {\em low} areas than in the high areas:
+there are many users that give a small amount of cycles, while there
+are just a few that give many cycles. This is in concordance with the
+results obtained in \cite{agajaj}.
+
+As far as we know, this is the only experiment that uses computational
+resources that are as dissimilar as smartphones and powerful laptops
+or desktop computers in a research center. The methodology used to
+gather resources and the algorithms used will be described next.
 
@@
 
+In order to test the volunteer computing environment, a presentation
+describing a low cost volunteer computing environment was
+created. This presentation was actually delivered in several
+conferences\footnote{Names withheld for the double blind review} and
+it has the appearance shown in Figure \ref{fig:sshot}. During the
+conference, it was revealed to the users that they were participating
+in the experiment. The same procedure was used when trying to gather
+users in social networks: a description of the container (the
+presentation) and disclosure of its purpose. As it can be seen in the
+image, the upper right corner shows the progress of the evolution. It
+stops when the solution has been reached.
+
+Above the experiment, from the point of view of the user, has been
+described. What it is actually running is an island model
+\cite{muhlenbein1991parallel} in every browser that uses the server as
+a {\em shuttle} to transfer individuals from one browser to another in
+what can be eventually a fully connected topology. This deals with
+several problems. The connection is stateless: islands connect to the
+server to send and receive a single individual but there is no task
+balancing: all islands start clicking on an URL and finish when they
+browse off to the next page. Fault tolerance is implicit, in the sense
+that no island has information that cannot be, or generated, somewhere
+else, although if the server fails the experiment results might be
+lost. All operation is asynchronous, with islands entering and leaving
+the experiment all of their own.
+
+We will next explain its different parts from the point of
+view of the architecture itself and the algorithm that it is actually
+running. First we will describe the client code and next, in
+subsection \ref{ss:server} the server architecture and where it is
+hosted.
+
+\subsection{Volunteer distributed evolutionary computation in the browser}
+
+\section{Results}
+\label{sec:res}
+
 \section{Conclusion}
+\label{sec:conc}
 
-In this paper we will offer our experience on using browser-based computing since 1999 \cite{jesusIWANN99} and other emerging paradigms, such as peer to peer based computing \cite{evag:gpem}, mainly applied to the design of evolutionary algorithms.
+In this paper we have presented our experience on using browser-based
+computing applied
+to the design of evolutionary algorithms.
 
@@
 \bibliographystyle{abbrv}
-\bibliography{geneura,volunteer,javascript,ror-js}
+\bibliography{geneura,volunteer,javascript,ror-js,GA-general}
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Well, first paragraph written

@@
 
-\subsection{Volunteer distributed evolutionary computation in the browser}
+\subsection{Volunteer distributed evolutionary computation in the
+  browser}
+
+As stated above, the evolutionary algorithms runs mainly in the
+browser. The problem run is a multimodal problem called {\em l-trap},
+which has been used extensively as a benchmark for evolutionary
+algorithms \cite{fernandes2009using,nijssen2003analysis}. This
+function counts the number of bits in a sequence of $l$ and assigns
+the local maximum if it has got 0 bits and the global maximum to $l$
+bits. The fitness goes down to a {\em trap} as you increase the number
+of bits, that is why it creates a rugged fitness landscape that is
+difficult to surmount for evolutionary algorithms. Its difficulty is
+increased as the number of traps concatenated grows, so that in some
+cases it might need millions of evaluations to find the solution. In
+our case we have used several values, from 30 to 50 traps, although we
+were not so much focused in finding the solution (or the fastest way
+of finding it) but on creating a experiment that could last for a
+certain amount of time, around one hour, so that there would be a
+chance for having many users checkout out the page and contributing
+cycles to it.
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Finishing subsection

@@
 
+This fitness is implemented as part of a (name hidden) evolutionary
+algorithm library written in node.js, a server-based implementation of
+JavaScript. However, since it is going to be run on the browser, a
+procedure called ``browserification'' must be applied to it so that it
+can run, without modifications, in any browser. The main difference
+between JavaScript on the browser and node is the way they load the
+external libraries. The algorithm used a population of 256
+individuals, random initialization, a canonical evolutionary algorithm
+with an elite of 2 individuals and crossover and then mutation applied
+to all the population pool, that was generated using 2-individual
+tournament selection. Once again, the particular evolutionary
+algorithm used is not so important, since we were rather focused on
+the user behavior, although it is quite clear that the selective
+pressure of the algorithm will have an impact in its performance and
+will play for or against its asynchrony in many different and complex
+ways \cite{jj:2008:PPSN}.
+
+The initial code was only modified to represent the fitness in a chart
+(that uses the {\tt chart.js} JavaScript library) so that the user has
+a visual feedback of what is going on. The {\em migration} part is
+also added: after every 100 generations, a single individual is sent
+to the server and another request is made to the server of a single
+and random individual. This makes the procedure less greedy, and also
+decreases the probability of getting the same individual that was sent
+some generations back. These two operations are made in {\em fire and
+  forget} style without any error checking, so that if the server is
+down, or indeed if the algorithm is run locally, it is not interrupted
+and continues operation until the solution is found, a string with all
+ones.
+
+As a summary, the only thing needed to run an evolutionary algorithm
+in the browser is to add {\em migration} functions. To make the
+operation of the algorithm independent of them is a plus, because it
+makes the local algorithm more fault-tolerant. The algorithm running
+on every browser has the same parameters and starts from a random
+population when the page is loaded.
+
 \section{Results}

### 2015-02-08 modeling-volunteer-computing-gecco.tex
New data

@@
 
+\subsection{Server side}
+\label{ss:server}
+
+The server was also written using node.js and the {\tt express.js}
+module and has three REST \cite{DBLP:journals/corr/abs-1105-4971}
+functions, that is, three functions that can be accessed from any
+client, including the browser, via a request encoded in an URL and a
+standard HTTP command. These functions are\begin{itemize}
+\item {\tt GET random} returns a random, non-evaluated individual,
+  from the pool. The pool is just an array containing chromosomes,
+  that is initialized at the beginning of the experiment.
+\item {\tt PUT one} puts a single individual in the pool. In REST
+  conventions, {\tt PUT} is used to create a resource, that is why it
+  is used, in this case, to add a new individual to the pool.
+\item {\tt GET log} returns the experiment log so far. It is used by
+  the researcher to gather data, and as a transparency measure so that
+  anyone can use it. 
+\end{itemize}
+
+In fact, the program includes more logic to serve the static pages
+that contain the talk in which it is included and the JavaScript/CSS
+code also, but it all amounts to a small amount of code. This code can
+be deployed, in principle, in any server running node.js, but for the
+purposes of this experiment Heroku, a Platform as a Service with a
+free tier, has been chosen. It could be also be deployed, with small
+or no modification, to OpenShift or other similar PaaS. Far from being
+a freeriding way, using free or very low cost servers can be a
+sustainable way of performing massive distributed evolutionary
+computation experiments \cite{Merelo2014}
+
+Having the pages and the server in the same domain allows to work in
+default Ajax mode, that does not allow cross-site requests. However,
+these can also be disabled if needed. In this case, every domain holds
+a single experiment, there is no multi-tenancy of experiments, so its
+management is quite simple.
+
+When the experiment is finished, that is, the solution has been found,
+the server is restarted using  command line tools provided by
+Heroku. Every experiment was first done during the talk and then
+announced in Twitter several times until enough data was gathered.
+
+As indicated in the introduction, every part of the experiment and
+data gathered have been released as free software in a GitHub
+repository (address hidden for double blind review).
+
 \section{Results}

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Adding graphs with results

@@
 sustainable way of performing massive distributed evolutionary
-computation experiments \cite{Merelo2014}
+computation experiments \cite{Merelo2014,jj:idc:lowcost}.
 
@@
 data gathered have been released as free software in a GitHub
-repository (address hidden for double blind review).
+repository (address hidden for double blind review). These results
+will be analyzed in the next section.
 
@@
 \label{sec:res}
+
+The litmus test that all experiments should pass, that is, that they
+work and found the solution, passed. In fact, the first version of the
+algorithm (until January 2015) used too much memory on the browser,
+which crashed after a while running. A total of 7 experiments were
+done with this version. After January, a new version of the
+evolutionary algorithm with non-cached fitness was used which did not
+crash the browser and was thus allowed to run as much time as
+needed. That was the version used in a lightning talk at the beginning
+of February. A total of 10 experiments were made with this second
+version. In the first version, finding the solution took a few
+minutes. The second needed around 20 minutes. In the logs used, it was
+not checked that the algorithm had found the solution, but in most
+cases it had. More systematic experimentation would have to wait until
+the next version, since in this first one we were looking for some
+kind of systematicity in the behavior of users and the experimental
+apparatus is the minimum possible to have it running.
+%
+\begin{figure*}[htb]
+        \centering
+        \subfigure[First experiment]{
+                \includegraphics[width=0.45\linewidth]{data/diffs-ips-1.png}
+                \label{fig:e1_p}
+        }
+        \subfigure[Second experiment]{
+                \includegraphics[width=0.45\linewidth]{data/diffs-ips-2.png}
+                \label{fig:e2_p}
+        }
+        \caption{Histogram of number of different IPs in an experiment}\label{fig:histo:ips}
+\end{figure*}
+
+The fact that it run correctly during a massive lightning talk means that this setup is a
+valid, no-cost, platform for massive evolutionary algorithms. There
+were several ways of introducing it: during a talk, telling the
+audience to visit the web that was posted in the first slide, and
+after the talk, inserting the URL in a tweet with a brief explanation
+and several tweets more indicating what it was about and what it
+did. If any question arose, it was answered as fast as possible.
 

### 2015-02-08 data/IPs.R

@@ -1,0 +1,2 @@
+
+number.ips.2 <- c(length(traps.ips.1$IP), length(traps.ips.2$IP), length(traps.ips.3$IP), length(traps.ips.4$IP),length(traps.ips.5$IP), length(traps.ips.6$IP), length(traps.ips.7$IP),length(traps.ips.8$IP), length(traps.ips.8$IP), length(traps.ips.10$IP))

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Explaining the experiment

@@
 and several tweets more indicating what it was about and what it
-did. If any question arose, it was answered as fast as possible.
+did. If any question arose, it was answered as fast as possible. In
+the first experiment, the median number of different IPs was 18, with
+a maximum of 28 and a minimum of 6. In the second, the median was 12.5
+with a maximum of 19 and minimum of 8. The histogram of different IPs
+in each experiment is shown in figure \ref{fig:histo:ips}. Probably
+the most remarkable thing is that a cluster of 6 computers can, as a
+minimum, be gathered for a distributed computing experiment by just
+announcing it as a tweet, but the fact that we can obtain more than a
+dozen computers in more than half the cases is also remarkable, and
+goes to prove the excess of computing resources that people,
+willingly, lend to a simple experiment.
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Getting closer to the end

@@
 willingly, lend to a simple experiment.
+%
+\begin{figure*}[htb]
+        \centering
+        \subfigure[First experiment, run 3]{
+                \includegraphics[width=0.45\linewidth]{data/log-puts-1-exp7.png}
+                \label{fig:d1}
+        }
+        \subfigure[Second experiment, run 7]{
+                \includegraphics[width=0.45\linewidth]{data/log-puts-2-traps-eps-3.png}
+                \label{fig:d2}
+        }
+        \caption{IPs and the number of {\tt PUT}s (one every $n$
+          generations, 50 in the first experiment, 100 in the second
+          experiment) they did, sorted. These correspond to the runs
+          in each batch of experiments with the max number of
+          different IPs. Please note the algorithmic scale in the $x$ axis.}\label{fig:dist}
+\end{figure*}
+
+Not all users contribute in the same way. The distribution for the
+runs in each experiment with the most IPs is shown in Figure
+\ref{fig:dist}, which chooses the experiment with the most number of
+uses and ranks, tops to down, every IP with the number of PUTs (which
+should be multiplied by the number of generations per PUT, 50 in the
+first case, 100 in the second, although at this point it is not
+important, since all nodes in the experiment do it after the same
+number of generations) contributed to the experiment. We can observe
+in both cases a power law, something already observed by \cite{agajaj}
+and which implies that there is a fixed proportion between the number
+of cycles (generations) contributed by the first and that contributed
+by the second which is roughly the same as the one between the second
+and third. Note also that, since the experiment is hosted in a PaaS,
+the IPs listed are anonymous, with a 10 always as the first number.
+%
+\begin{figure*}[htb]
+        \centering
+        \includegraphics[width=0.8\linewidth]{data/time-vs-ips.png}
+        \caption{Total experiment time (time of the last PUT - time of
+        the first one) vs number of IPs participating in it.}\label{fig:t}
+\end{figure*}
+%
+\begin{figure*}[htb]
+        \centering
+        \includegraphics[width=0.8\linewidth]{data/intervals-histo.png}
+        \caption{Histogram of the time between two PUTs for all the
+          experiments in the second batch. The $x$ axis is logarithmic
+        and bin size is 0.5, which means that the first bin contains
+        all intervals of less than 3 seconds.}\label{fig:intervals}
+\end{figure*}
+
+At any rate, this exponential trend in the first posts breaks down
+after a number of IPs. That could be due to several possible factors:
+persons checking the web for just an instant, late comers, or very
+slow computers that are not able to lend more cycles to the
+simulation. The first IP in both cases contributes on the order of a
+hundred PUTs, that is, several thousands generations and several
+hundred thousands evaluations, which, once again, is quite remarkable
+for such a simple experiment announced by an ephemeral tweet. It
+should be noted that, since the speed in the browser virtual machines
+(Chrome, mainly, but also Firefox) is as fast as the node.js version,
+a parallel version does not add overhead to the single-browser
+version, which, as mentioned above, can run in an independent
+way: a second browser always adds to the first, even if the total
+number of evaluations and the time needed will not speed up in the
+same way. However, even taking into account the stochastic nature of a
+single evolutionary algorithm, what we are going to find is a high
+dispersion of the times needed to reach a solution, which are plotted
+in Figure \ref{fig:t} against the number of IPs. There is a wide
+range of durations, although there is no clear decrease of the time
+needed as the number of IPs used increase. However, we will have to
+investigate further the speed ups achieved and if these are
+algorithmic or due to the particular asynchronous and web-based
+implementation. Figure \ref{fig:intervals} might help understand the
+kind of raw material we are using: It plots an histogram of the time
+needed by all experiments to process 100 generations, 25600
+evaluations and the associated evolutionary geneticry. The vast
+majority of browsers processes it in less than 3 seconds (logarithm(
+time ) < 0.5, which is equivalent to 3.16). However, it is interesting
+to note that there are some intervals that go up to the thousands of
+seconds, for which we cannot really offer an explanation, other than
+the user leaving the page and returning after a while. However, 75\%
+of the cases take less than 4 seconds, which is thus a statistical
+measure of the kind of performance we can expect from clients for this
+type of problem. We can also conclude that 75\%, or a majority, will
+be in the fastest tier, although there will be a small amount of them
+that will be quite slow and will probably be impossible to accomodate
+in the distributed computing experiment.
 

### 2015-02-08 data/process-intervals.pl

@@ -1,0 +1,30 @@
+#!/usr/bin/env perl
+
+use File::Slurp::Tiny qw(read_file);
+use JSON;
+use v5.14;
+
+my $file_name = shift || "log";
+
+my $log_json = read_file( $file_name ) || die "Can't read file $file_name\n";
+
+my $log = decode_json $log_json;
+
+my %times;
+for my $l (@$log ) {
+    my $IP = $l->{'IP'};
+    if ( $l->{'put'} ) {
+	push @{$times{$IP}}, $l->{'put'}[0];
+    };
+}
+
+for my $ip ( keys %times ) {
+  if ( @{$times{$ip}} > 1 ) {
+    my $time = shift @{$times{$ip}};
+    for my $t (  @{$times{$ip}} ) {
+      say "$ip;", $t-$time;
+      $time = $t;
+    }
+  }
+}
+

### 2015-02-08 data/process-time.pl

@@ -1,0 +1,32 @@
+#!/usr/bin/env perl
+
+use File::Slurp::Tiny qw(read_file);
+use JSON;
+use v5.14;
+
+my $file_name = shift || "log";
+
+my $log_json = read_file( $file_name ) || die "Can't read file $file_name\n";
+
+my $log = decode_json $log_json;
+
+my %IPs;
+
+my @times;
+for my $l (@$log ) {
+  if ( $l->{'put'} ) {
+    push @times, $l->{'put'}[0];
+  } else {
+    push @times, $l->{'get'}[0];
+  }
+  if ( $l->{'put'} ){
+    $IPs{$l->{'IP'}}++
+  };
+}
+
+my $total_time = $times[$#times]-$times[0];
+my $total_ips = keys(%IPs);
+
+
+say "$total_time;$total_ips";
+

### 2015-02-08 modeling-volunteer-computing-gecco.tex
First look to the paper and a possible reference to use added
Some comments have been added and also some typos fixed.

--- 2015-02-08 modeling-volunteer-computing-gecco.tex
@@
 \begin{abstract}
-From the era of big science we are back to the "do it yourself", where
+From the era of big science we are back to the ``do it yourself'', where
 you do not have any money to buy clusters or subscribe to grids but
@@
 optimize the performance in systems where the number of nodes, their
-connections, and the time every one is devoting to the computation is
+connections, and the time every one is dedicating to the computation is
+%FERGU: cambiado el devoting que se repite, por dedicating 
 known in advance. However, even if Big Science is not really over and
@@
 experiments. These experiments can be easily created using open-source
-repository sites like GitHub and deployed automatically to Platform as
-a Service products such as Heroku or OpenShift.
+repository sites like GitHub\footnote{\url{http://www.github.com}} and deployed automatically to Platform as
+a Service (PaaS) products such as Heroku\footnote{\url{https://www.heroku.com/}} or OpenShift\footnote{\url{https://www.openshift.com/}}. 
+%FERGU: luego usas el acrónimo PaaS
 
@@
 of trust among the scientist and the person that is running his/her
-code in the browser. As has been mentioned in the early stages of what
-was then called {\em desktop grid computing}, \cite{gc:bausch}, in
+code in the browser. As it has been mentioned in the early stages of what
+was then called {\em desktop grid computing} \cite{gc:bausch}, in
 fact CPU cycle selling might not make economic sense since there are
@@
 to users.
+
+% Había metido en el .bib el artículo de detectar vulnerabilidades 
+% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
+% hablar en este paper de él (Muszynski14vulnerabilities)
 
@@
 this paper, that first presents the first versions of a platform for
-distributed evolutionary algorithms using the browser and a free (as
+distributed evolutionary algorithms (EAs) using the browser and a free (as
 in free beer) backend, and second, shows the result from an
@@
 the state of the art in volunteer computing and its modelization. We 
+%FERGU: aunque hay artículos con el término "modelization", no sé si el correcto será "modelling"
 will proceed to describe the experimental setup in Section
@@
 used. Java was pointed out in \cite{soares1998get} as a ``language for
-internet'' and
-\begin{quote}
-Java can bring some important advantages... solves architectural
-heterogeneity [...] easy to install and [...] built-in security
-mechanisms
-\end{quote}
-The same paper by Soares et al. describes JET, a system that supports
-the execution of parallel applications over the Internet. In fact the
-same paper describes some other systems working at the same time,
-except they cannot be used for science since they do not have a
-comprehensive statistics support like JET.
+internet'', providing some advantages such as multi-architecture compatibility or 
+security mechanisms.
+%\begin{quote}
+%Java can bring some important advantages... solves architectural
+%heterogeneity [...] easy to install and [...] built-in security
+%mechanisms
+%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
+In that work, Soares et al. describe JET, a system that supports
+the execution of parallel applications over the Internet. This system has 
+a comprehensive statistics support, allowing its use for science, 
+unlike the other systems compared in that paper.
 
-However, Java (and all the other technologies) is not universal in the
+However, Java (and all the other technologies)
+is not universal in the
 sense than an extra component, namely, the Java virtual machine, has
@@
 that have been widely adopted by the industry and also by scientists,
-who used them for creating a non-distributed evolutionary algorithm on
+who used them for creating a non-distributed EA on
 the browser as early as 1998 \cite{jj-ppsn98}.
@@
 mischief of users with code in their hands
-\cite{sarmenta-sabotagetolerance} however, these early efforts by
+\cite{sarmenta-sabotagetolerance}. However, these early efforts by
 Sarmenta once again used Java and not JavaScript, making this effort
@@
 \cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
-evolutionary computation and has been used ever since by several
+evolutionary computation and it has been used ever since by several
 authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
@@
 performance that can be obtained from these volunteer computing
-effort. There was some effort initially to dodge the issue by making
+effort. There was some attempt initially to dodge the issue by making
 the algorithm adaptive to the kind of resources allotted to it
 \cite{milani2004online}, which is actually not such a big problem in
-algorithms such as the evolutionary algorithm that can easily be
-paralelized via population splitting or farming out evaluations to all
+algorithms such as the EA that can easily be
+parallelized via population splitting or farming out evaluations to all
 the nodes available. Lately, several approaches have focused on the
@@
 \cite{nogueras2015studying} or including it in a more general study of
-performance of the evolutionary algorithm itself
+performance of the EA itself %Con tu permiso lo paso a acrónimos para que no sea tan largo el párrafo. FERGU
 \cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
@@
 
-As far as we know, this is the only experiment that uses computational
+As far as we know, this paper presents the only experiment that uses computational
 resources that are as dissimilar as smartphones and powerful laptops
@@
 gather resources and the algorithms used will be described next.
+
+%FERGU: volver a indicar que además se dan resultados de uso, para que quede clara la contribución?
 
@@
 \end{figure*}
-
+%FERGU: No puedes poner el gráfico un poco más grande? (y de paso 
+% recortar mejor la imagen para que no se vea el texto de la derecha)
 In order to test the volunteer computing environment, a presentation
@@
 a {\em shuttle} to transfer individuals from one browser to another in
-what can be eventually a fully connected topology. This deals with
+what can be eventually a fully connected topology. 
+%FERGU: explica esto: los navegadores se conectan entre sí o sólo a tu servidor?
+This deals with
 several problems. The connection is stateless: islands connect to the
@@
 This fitness is implemented as part of a (name hidden) evolutionary
+%FERGU: creo que si usas un framework libre puedes dejarlo visible
 algorithm library written in node.js, a server-based implementation of
@@
 can run, without modifications, in any browser. The main difference
-between JavaScript on the browser and node is the way they load the
+between JavaScript on the browser and node.js is the way they load the
+%FERGU: puesto .js para distinguir la palabra node del nombre de la librería
 external libraries. The algorithm used a population of 256
@@
 will play for or against its asynchrony in many different and complex
-ways \cite{jj:2008:PPSN}.
+ways \cite{jj:2008:PPSN}. %FERGU: este párrafo me lo copio para el de la ASOCO ;)
 
@@
   forget} style without any error checking, so that if the server is
+  %FERGU: Qué es fire and forget? (se puede intuir, pero dejarlo claro)
 down, or indeed if the algorithm is run locally, it is not interrupted
@@
 sustainable way of performing massive distributed evolutionary
-computation experiments \cite{Merelo2014,jj:idc:lowcost}.
+computation experiments \cite{Merelo2014}
 
@@
 data gathered have been released as free software in a GitHub
-repository (address hidden for double blind review). These results
-will be analyzed in the next section.
+repository (address hidden for double blind review).
 
@@
 \label{sec:res}
-
-The litmus test that all experiments should pass, that is, that they
-work and found the solution, passed. In fact, the first version of the
-algorithm (until January 2015) used too much memory on the browser,
-which crashed after a while running. A total of 7 experiments were
-done with this version. After January, a new version of the
-evolutionary algorithm with non-cached fitness was used which did not
-crash the browser and was thus allowed to run as much time as
-needed. That was the version used in a lightning talk at the beginning
-of February. A total of 10 experiments were made with this second
-version. In the first version, finding the solution took a few
-minutes. The second needed around 20 minutes. In the logs used, it was
-not checked that the algorithm had found the solution, but in most
-cases it had. More systematic experimentation would have to wait until
-the next version, since in this first one we were looking for some
-kind of systematicity in the behavior of users and the experimental
-apparatus is the minimum possible to have it running.
-%
-\begin{figure*}[htb]
-        \centering
-        \subfigure[First experiment]{
-                \includegraphics[width=0.45\linewidth]{data/diffs-ips-1.png}
-                \label{fig:e1_p}
-        }
-        \subfigure[Second experiment]{
-                \includegraphics[width=0.45\linewidth]{data/diffs-ips-2.png}
-                \label{fig:e2_p}
-        }
-        \caption{Histogram of number of different IPs in an experiment}\label{fig:histo:ips}
-\end{figure*}
-
-The fact that it run correctly during a massive lightning talk means that this setup is a
-valid, no-cost, platform for massive evolutionary algorithms. There
-were several ways of introducing it: during a talk, telling the
-audience to visit the web that was posted in the first slide, and
-after the talk, inserting the URL in a tweet with a brief explanation
-and several tweets more indicating what it was about and what it
-did. If any question arose, it was answered as fast as possible.
 
@@
 
-There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}.
+There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}, or even security issues.
 But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.

### 2015-02-08 modeling-volunteer-computing-gecco.tex

@@
 \label{sec:res}
+
+The litmus test that all experiments should pass, that is, that they
+work and found the solution, passed. In fact, the first version of the
+algorithm (until January 2015) used too much memory on the browser,
+which crashed after a while running. A total of 7 experiments were
+done with this version. After January, a new version of the
+evolutionary algorithm with non-cached fitness was used which did not
+crash the browser and was thus allowed to run as much time as
+needed. That was the version used in a lightning talk at the beginning
+of February. A total of 10 experiments were made with this second
+version. In the first version, finding the solution took a few
+minutes. The second needed around 20 minutes. In the logs used, it was
+not checked that the algorithm had found the solution, but in most
+cases it had. More systematic experimentation would have to wait until
+the next version, since in this first one we were looking for some
+kind of systematicity in the behavior of users and the experimental
+apparatus is the minimum possible to have it running.
+%
+\begin{figure*}[htb]
+        \centering
+        \subfigure[First experiment]{
+                \includegraphics[width=0.45\linewidth]{data/diffs-ips-1.png}
+                \label{fig:e1_p}
+        }
+        \subfigure[Second experiment]{
+                \includegraphics[width=0.45\linewidth]{data/diffs-ips-2.png}
+                \label{fig:e2_p}
+        }
+        \caption{Histogram of number of different IPs in an experiment}\label{fig:histo:ips}
+\end{figure*}
+
+The fact that it run correctly during a massive lightning talk means that this setup is a
+valid, no-cost, platform for massive evolutionary algorithms. There
+were several ways of introducing it: during a talk, telling the
+audience to visit the web that was posted in the first slide, and
+after the talk, inserting the URL in a tweet with a brief explanation
+and several tweets more indicating what it was about and what it
+did. If any question arose, it was answered as fast as possible. In
+the first experiment, the median number of different IPs was 18, with
+a maximum of 28 and a minimum of 6. In the second, the median was 12.5
+with a maximum of 19 and minimum of 8. The histogram of different IPs
+in each experiment is shown in figure \ref{fig:histo:ips}. Probably
+the most remarkable thing is that a cluster of 6 computers can, as a
+minimum, be gathered for a distributed computing experiment by just
+announcing it as a tweet, but the fact that we can obtain more than a
+dozen computers in more than half the cases is also remarkable, and
+goes to prove the excess of computing resources that people,
+willingly, lend to a simple experiment.
+%
+\begin{figure*}[htb]
+        \centering
+        \subfigure[First experiment, run 3]{
+                \includegraphics[width=0.45\linewidth]{data/log-puts-1-exp7.png}
+                \label{fig:d1}
+        }
+        \subfigure[Second experiment, run 7]{
+                \includegraphics[width=0.45\linewidth]{data/log-puts-2-traps-eps-3.png}
+                \label{fig:d2}
+        }
+        \caption{IPs and the number of {\tt PUT}s (one every $n$
+          generations, 50 in the first experiment, 100 in the second
+          experiment) they did, sorted. These correspond to the runs
+          in each batch of experiments with the max number of
+          different IPs. Please note the algorithmic scale in the $x$ axis.}\label{fig:dist}
+\end{figure*}
+
+Not all users contribute in the same way. The distribution for the
+runs in each experiment with the most IPs is shown in Figure
+\ref{fig:dist}, which chooses the experiment with the most number of
+uses and ranks, tops to down, every IP with the number of PUTs (which
+should be multiplied by the number of generations per PUT, 50 in the
+first case, 100 in the second, although at this point it is not
+important, since all nodes in the experiment do it after the same
+number of generations) contributed to the experiment. We can observe
+in both cases a power law, something already observed by \cite{agajaj}
+and which implies that there is a fixed proportion between the number
+of cycles (generations) contributed by the first and that contributed
+by the second which is roughly the same as the one between the second
+and third. Note also that, since the experiment is hosted in a PaaS,
+the IPs listed are anonymous, with a 10 always as the first number.
+%
+\begin{figure*}[htb]
+        \centering
+        \includegraphics[width=0.8\linewidth]{data/time-vs-ips.png}
+        \caption{Total experiment time (time of the last PUT - time of
+        the first one) vs number of IPs participating in it.}\label{fig:t}
+\end{figure*}
+%
+\begin{figure*}[htb]
+        \centering
+        \includegraphics[width=0.8\linewidth]{data/intervals-histo.png}
+        \caption{Histogram of the time between two PUTs for all the
+          experiments in the second batch. The $x$ axis is logarithmic
+        and bin size is 0.5, which means that the first bin contains
+        all intervals of less than 3 seconds.}\label{fig:intervals}
+\end{figure*}
+
+At any rate, this exponential trend in the first posts breaks down
+after a number of IPs. That could be due to several possible factors:
+persons checking the web for just an instant, late comers, or very
+slow computers that are not able to lend more cycles to the
+simulation. The first IP in both cases contributes on the order of a
+hundred PUTs, that is, several thousands generations and several
+hundred thousands evaluations, which, once again, is quite remarkable
+for such a simple experiment announced by an ephemeral tweet. It
+should be noted that, since the speed in the browser virtual machines
+(Chrome, mainly, but also Firefox) is as fast as the node.js version,
+a parallel version does not add overhead to the single-browser
+version, which, as mentioned above, can run in an independent
+way: a second browser always adds to the first, even if the total
+number of evaluations and the time needed will not speed up in the
+same way. However, even taking into account the stochastic nature of a
+single evolutionary algorithm, what we are going to find is a high
+dispersion of the times needed to reach a solution, which are plotted
+in Figure \ref{fig:t} against the number of IPs. There is a wide
+range of durations, although there is no clear decrease of the time
+needed as the number of IPs used increase. However, we will have to
+investigate further the speed ups achieved and if these are
+algorithmic or due to the particular asynchronous and web-based
+implementation. Figure \ref{fig:intervals} might help understand the
+kind of raw material we are using: It plots an histogram of the time
+needed by all experiments to process 100 generations, 25600
+evaluations and the associated evolutionary geneticry. The vast
+majority of browsers processes it in less than 3 seconds (logarithm(
+time ) < 0.5, which is equivalent to 3.16). However, it is interesting
+to note that there are some intervals that go up to the thousands of
+seconds, for which we cannot really offer an explanation, other than
+the user leaving the page and returning after a while. However, 75\%
+of the cases take less than 4 seconds, which is thus a statistical
+measure of the kind of performance we can expect from clients for this
+type of problem. We can also conclude that 75\%, or a majority, will
+be in the fastest tier, although there will be a small amount of them
+that will be quite slow and will probably be impossible to accomodate
+in the distributed computing experiment.
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Here it goes

--- 2015-02-08 modeling-volunteer-computing-gecco.tex
@@
 In this paper we have presented our experience on using browser-based
-computing applied
-to the design of evolutionary algorithms.
+computing applied to the design of evolutionary algorithms. In the
+spirit of Open Science, we have released all material related to the
+experiment, including this paper. Experiments have been performed at
+several times of the day and announced in Twitter or during a live
+presentation about distributed computation using the browser.
 
-There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}.
-But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.
-This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.
+What we conclude from these computers is that the nature of the
+performance, which is due to the number of persons deciding to
+participate in the experiment, is not totally random. There are at
+least lower bounds we can count on: several computers (as few as six)
+can almost always be relied on, and in some cases up to 30 can
+participate in a single experiment. However, it is not clear how all
+these computers contribute to the overall performance in terms of
+time, although all experiments were performed until a solution was
+found. Most computers participating in the experiment are in the same
+performance tier, although a small percentage of them (which will be
+around 25\%) are quite slow.
+
+The evolutionary algorithm programmed in this way can easily
+accommodate all kinds of browsers. However, we are using an homogeneous
+configuration for all of them, which might not be the best from an
+algorithmic point of view; a random parameter setting like the one
+used by \cite{LNCS86720702} could offer better results, since we do
+not know in advance what is the performance of the nodes. More
+systematic experimentation is also needed, specially using different
+kind of problems, including more complex problems in which the fitness
+function is {\em heavier} and takes longer to be processed. However,
+the main intention of this paper, which was to prove that distributed
+evolutionary computation could be done efficiently in a volunteer/free
+environment, has been sufficiently proved.
+
+There are many issues involved in using these resources: from adapting
+current algorithms so that they match this environment 
+to check which EA configuration works the best in it, or creating a
+framework that can use it easily. But the main
+challenge is that asking people to contribute resources implies that
+you are opening your science to society and you have to give something
+in return: you have to adopt a set of best practices that have come to
+be known as Open Science, because ``Give, and it shall be given unto
+you'', you will get as much back from society as you give to it
+opening your science and explaining it to the public. This, among
+other things, means that popularity will become directly performance
+of the metacomputer you create by attracting more users. And this
+interplay between the social network itself (popularity in Twitter,
+people attending to the conference and interested in following the
+slides in their computer) is another, very interesting, venue to
+explore in the future.
+
+\section{Acknowledgements}
+
+This paper is supported in part by several grants hidden for
+anonymity.
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Solving conflict

--- 2015-02-08 modeling-volunteer-computing-gecco.tex
@@
 In this paper we have presented our experience on using browser-based
-computing applied
-to the design of evolutionary algorithms.
+computing applied to the design of evolutionary algorithms. In the
+spirit of Open Science, we have released all material related to the
+experiment, including this paper. Experiments have been performed at
+several times of the day and announced in Twitter or during a live
+presentation about distributed computation using the browser.
 
-There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}, or even security issues.
-But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public.
-This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.
+What we conclude from these computers is that the nature of the
+performance, which is due to the number of persons deciding to
+participate in the experiment, is not totally random. There are at
+least lower bounds we can count on: several computers (as few as six)
+can almost always be relied on, and in some cases up to 30 can
+participate in a single experiment. However, it is not clear how all
+these computers contribute to the overall performance in terms of
+time, although all experiments were performed until a solution was
+found. Most computers participating in the experiment are in the same
+performance tier, although a small percentage of them (which will be
+around 25\%) are quite slow.
+
+The evolutionary algorithm programmed in this way can easily
+accommodate all kinds of browsers. However, we are using an homogeneous
+configuration for all of them, which might not be the best from an
+algorithmic point of view; a random parameter setting like the one
+used by \cite{LNCS86720702} could offer better results, since we do
+not know in advance what is the performance of the nodes. More
+systematic experimentation is also needed, specially using different
+kind of problems, including more complex problems in which the fitness
+function is {\em heavier} and takes longer to be processed. However,
+the main intention of this paper, which was to prove that distributed
+evolutionary computation could be done efficiently in a volunteer/free
+environment, has been sufficiently proved.
+
+There are many issues involved in using these resources: from adapting
+current algorithms so that they match this environment 
+to check which EA configuration works the best in it, or creating a
+framework that can use it easily. But the main
+challenge is that asking people to contribute resources implies that
+you are opening your science to society and you have to give something
+in return: you have to adopt a set of best practices that have come to
+be known as Open Science, because ``Give, and it shall be given unto
+you'', you will get as much back from society as you give to it
+opening your science and explaining it to the public. This, among
+other things, means that popularity will become directly performance
+of the metacomputer you create by attracting more users. And this
+interplay between the social network itself (popularity in Twitter,
+people attending to the conference and interested in following the
+slides in their computer) is another, very interesting, venue to
+explore in the future.
+
+\section{Acknowledgements}
+
+This paper is supported in part by several grants hidden for
+anonymity.
 

### 2015-02-08 modeling-volunteer-computing-gecco.tex
Commented results and modified a reference
Using a book chapter instead of a arxiv pre-print. Also, commented some parts in the results and conclusions sections.

--- 2015-02-08 modeling-volunteer-computing-gecco.tex
@@
 The server was also written using node.js and the {\tt express.js}
-module and has three REST \cite{DBLP:journals/corr/abs-1105-4971}
+module and has three REST \cite{Castillo12REST} %FERGU: cito a este paper que es un capítulo de libro en vez de Arxiv
 functions, that is, three functions that can be accessed from any
@@
 The litmus test that all experiments should pass, that is, that they
-work and found the solution, passed. In fact, the first version of the
+work and found the solution, passed. %FERGU: que quiere decir esta primera frase?
+In fact, the first version of the
 algorithm (until January 2015) used too much memory on the browser,
@@
 kind of systematicity in the behavior of users and the experimental
+%FERGU: systematicity? systematization mejor?
 apparatus is the minimum possible to have it running.
@@
 willingly, lend to a simple experiment.
+%FERGU: dí que la persona que lo anunció en twitter tiene X seguidores
 %
@@
 needed as the number of IPs used increase. However, we will have to
-investigate further the speed ups achieved and if these are
+investigate further the speedups achieved and if these are %FERGU: speedup va junto
 algorithmic or due to the particular asynchronous and web-based
@@
 needed by all experiments to process 100 generations, 25600
-evaluations and the associated evolutionary geneticry. The vast
+evaluations and the associated evolutionary geneticry. The vast %FERGU: geneticry?
 majority of browsers processes it in less than 3 seconds (logarithm( 
@@
 be in the fastest tier, although there will be a small amount of them
-that will be quite slow and will probably be impossible to accomodate
+that will be quite slow and will probably be impossible to accommodate
 in the distributed computing experiment.
@@
 
+%FERGU: Creo que te falta terminar las conclusiones, pero por si acaso habría que decir: Este tipo de computación se basa en tal y tal y tiene tales problemas.
+Nuestra metodología ha propuesto usar tal tecnología por tal y tal.
+Los resultados demuestran que los usuarios hacen tal y tal y que por lo tanto hay que tener en cuenta el tal.
+Bueno, qué te voy a contar...
+
 There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}, or even security issues.

### 2015-02-08 modeling-volunteer-computing-gecco.tex

--- 2015-02-08 modeling-volunteer-computing-gecco.tex
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 

### 2015-03-31 modeling-volunteer-computing-gecco.tex
Answering @fergunet reviews

@@
 
-\title{Modeling browser-based distributed evolutionary computation systems}
+\title{Designing and modeling a browser-based distributed evolutionary computation systems}
 
-\numberofauthors{1} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
+\numberofauthors{1}
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
 \alignauthor 
-Som E. One\\
-       \affaddr{Some Institute}\\
-       \email{em@a.il}
+J. J. Merelo GuervÛs and Pablo GarcÌa-S·nchez\\
+  \affaddr{GeNeura \url{http://geneura.wordpress.com}, Dept. of Computer Architecture and Technology \url{http://atc.ugr.es} and CITIC \url{http://citic.ugr.es}}\\
+  \affaddr{University of Granada \url{http://www.ugr.university}}\\
+  \email{(jmerelo,pablogarcia)@ugr.es}
 }
@@
 \begin{abstract}
-From the era of big science we are back to the ``do it yourself'', where
-you do not have any money to buy clusters or subscribe to grids but
-still have algorithms that crave many computing nodes and need them to
-measure scalability. Fortunately, this coincides with the era of big
-data, cloud computing, and browsers that include JavaScript virtual
-machines. Those are the reasons why this paper will focus on two
-different aspects of volunteer or freeriding computing: first, the
+Web browsers have scaled from simple page-rendering engines to operating
+systems that include most services the lower OS layer has, with the
+added facility that applications can be run by just visiting a web
+page. In this paper we will describe the front and back end of a
+distributed evolutionary computation system that uses the browser's
+capabilities of running programs written in JavaScript. We will focus on two
+different aspects of volunteer computing: first, the
 pragmatic: where to find those resources, which ones can be used, what
 kind of support you have to give them; and then, the theoretical: how
-evolutionary algorithms can be adapted to a environment in which nodes
+evolutionary algorithms can be adapted to an environment in which nodes
 come and go, have different computing capabilities and operate in
 complete asynchrony of each other. We will examine the setup needed to
-create a very simple distributed evolutionary algorithm using
+create a  simple distributed evolutionary algorithm using
 JavaScript and then find a model of how users react to it by
@@
 
-\keywords{Evolutionary computation, volunteer computing, distributed
-  computing, node.js, javascript}
+\keywords{Desktop grid, distributed computing, volunteer computing,
+  Internet computing, socio-technical systems}
 
@@
 
-The world has computational resources in spades. Most of them do not
-belong to you or your lab. That does not mean you cannot use it. The
-problem is how.
+The world has computational resources in spades, with every person
+wielding one or more computational devices. The issue is how to make
+them allocate even a small part of those resources to your particular
+experiment.
 
-Most theory in parallel computing has been devoted to predict and
+Most theory and experimentation in parallel computing has been devoted to predict and
 optimize the performance in systems where the number of nodes, their
 connections, and the time every one is dedicating to the computation is
-%FERGU: cambiado el devoting que se repite, por dedicating 
-known in advance. However, even if Big Science is not really over and
-it is slated to come back, the era of
+known in advance as it often did in the Big Science era, where labs
+could easily buy 16-node clusters or subscribe to the resources of a
+big grid. Even today, you can rent instances of virtual machines in
+the cloud relatively cheaply and use them for evolutionary computation
+experiments \cite{sherry2012flex}. Nothing beats free, which is one of
+the reasons why the era of
 Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
-offers a vast amount of computational resources to attract, if only
-you know how. But there is a challenge: knowing, or at least having a
+offers a vast amount of computational resources to be used, if only
+you know how to attract them. But there is a challenge: knowing, or at least having a
 ballpark, of how your algorithm is going to perform in this uncertain
@@
 lowering their prices. This means that the conjunction of free or
-low-cost cloud computing engines, volunteer computing systems and and
+low-cost cloud computing engines, volunteer computing systems and
 untapped capability of desktop systems can be used for creating
@@
 repository sites like GitHub\footnote{\url{http://www.github.com}} and deployed automatically to Platform as
-a Service (PaaS) products such as Heroku\footnote{\url{https://www.heroku.com/}} or OpenShift\footnote{\url{https://www.openshift.com/}}. 
-%FERGU: luego usas el acrónimo PaaS
+a Service (PaaS) sites such as
+Heroku\footnote{\url{https://www.heroku.com/}} or
+OpenShift\footnote{\url{https://www.openshift.com/}} or even
+Infraestructure as a Service (IaaS) sites such as Microsoft's Azure.
 
-In general, any volunteer computing experiment will have to be made
-``in the open''. The fact that somebody is giving you, basically for
+In general, it should be a requisite for volunteer computing
+experiments  to be performed 
+``in the open''. The fact that somebody is giving to you, basically for
 free, computing resources, means that the scientist using them has to
-give back. The baseline is releasing the source code used: all
+give back; besides, people do not usually visit a resource unless they
+can trust it; trust is an essential factor in the sharing of resources
+such as this one, computing power \cite{van2009would}. The point of
+departure for establishing this trust is releasing the source code used: all
 volunteer computing platforms, from SETI@home on, have done it. The
@@
 have folded or others like CrowdProcess have decided to turn their
-product to in-house computation: there must be a mutual relationship
+product to in-house computation \cite{ycombinator}: there must be a mutual relationship
 of trust among the scientist and the person that is running his/her
@@
 mention the fact that the experiment {\em actually} has to draw the
-attention of the supplier. That is why trust is essential, and using
-free software might not be enough: Openness
+attention of the supplier, and thus it is better analyzed using the
+Uses and Gratifications Theory \cite{severin2010communication}, that
+is, taking into account {\em why} an user would do it, than using a
+purely computational approach. That is why considering trust is
+essential, and using free software might not be enough: Openness
 has to progress from open source code to open science: releasing all
 data obtained in the experiment in a repository such as GitHub,
-mentioned above, and even allowing real-time access to experiment data
+mentioned above, and even allowing real-time access to experimental data
 to users.
 
-% Había metido en el .bib el artículo de detectar vulnerabilidades 
+% Hab√≠a metido en el .bib el art√≠culo de detectar vulnerabilidades 
 % en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
-% hablar en este paper de él (Muszynski14vulnerabilities)
+% hablar en este paper de √©l (Muszynski14vulnerabilities)
+%Lo voy a meter
+In fact, this relationship of trust is in many cases the only
+mechanism preventing the malicious use of the system by creating bogus
+solutions or simply attempting to exploit vulnerabilities by injecting
+code. These problems have been analyzed in
+\cite{Muszynski14vulnerabilities} but in principle our working
+hypothesis will be that the clients using the browser are not
+manipulating the code make the experiment finish before it should,
+which is possibly the only way in which the experiment would be
+affected.
 
@@
 in the short term and using a more general model, there is still need
-to model the behavior of users so that more factors can be added to
+to model the behavior of users, so that more factors can be added to
 the model other than the time series. This user behavior, of course,
@@
 distributed evolutionary algorithms (EAs) using the browser and a free (as
-in free beer) backend, and second, shows the result from an
+in free beer) backend, and second, shows the result from a
 statistical point of view in order to put the basis for a model of the
@@
 The rest of the paper is organized as follows: next we will present
-the state of the art in volunteer computing and its modelization. We 
-%FERGU: aunque hay artículos con el término "modelization", no sé si el correcto será "modelling"
+the state of the art in volunteer computing and its modelling. We 
+%FERGU: aunque hay art√≠culos con el t√©rmino "modelization", no s√©
+%si el correcto ser√° "modelling"
+% OK- JJ
 will proceed to describe the experimental setup in Section
@@
 %\end{quote} FERGU: quito esta cita y lo pongo sin el quote
+% OK, pero øpor? - JJ
 In that work, Soares et al. describe JET, a system that supports
@@
 \cite{flanagan2006javascript} was introduced in 1997 as a
-browser-embedded language and has, since them, become a set of standards
+browser-embedded language and has, since then, become a set of standards
 \cite{ECMA-262} for the language, its components and future versions
@@
 \cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
-platform avoid single points of failure (the server) but, since they
+platforms avoid single points of failure (the server) but, since they
 need a certain amount of infrastructure installed to start, the
@@
 \cite{nogueras2015studying} or including it in a more general study of
-performance of the EA itself %Con tu permiso lo paso a acrónimos para que no sea tan largo el párrafo. FERGU
+performance of the EA itself %Con tu permiso lo paso a acr√≥nimos para
+                             %que no sea tan largo el p√°rrafo. FERGU
+%OK - JJ
 \cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
@@
 or desktop computers in a research center. The methodology used to
-gather resources and the algorithms used will be described next.
+gather resources and the algorithms used will be described next, as
+well as the results obtained with the setup.
 
-%FERGU: volver a indicar que además se dan resultados de uso, para que quede clara la contribución?
+%FERGU: volver a indicar que adem√°s se dan resultados de uso, para
+%que quede clara la contribuci√≥n?
+%JJ: Hecho.
 
-\section{Methodology}
+\section{Materials: running an island model in the browser}
 \label{sec:exp}
 
-\begin{figure*}[h!tb]
-  \centering
-  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
-  \label{fig:sshot}
-  \caption{Screenshot of the talk that includes at the graph
-    that shows the progression of the evolutionary
-    algorithm at the top right corner}
-\end{figure*}
-%FERGU: No puedes poner el gráfico un poco más grande? (y de paso 
-% recortar mejor la imagen para que no se vea el texto de la derecha)
 In order to test the volunteer computing environment, a presentation
@@
 created. This presentation was actually delivered in several
-conferences\footnote{Names withheld for the double blind review} and
-it has the appearance shown in Figure \ref{fig:sshot}. During the
-conference, it was revealed to the users that they were participating
+conferences (such as IDC 2014). During the %FERGU: TODO JJ, no te
+                                %olvides de decirme esto. Es la IDC
+                                %2014?
+% Y el PPSN. Pero eso lo voy a dejar para el trabajo anterior, voy a
+% intentar hacer experimentos nuevos. - JJ
+conference, it was revealed to the persons attending it that they were participating
 in the experiment. The same procedure was used when trying to gather
 users in social networks: a description of the container (the
-presentation) and disclosure of its purpose. As it can be seen in the
-image, the upper right corner shows the progress of the evolution. It
+presentation) and disclosure of its purpose. The upper right corner of
+the presentation shows the progress of the evolution. It
 stops when the solution has been reached.
@@
 a {\em shuttle} to transfer individuals from one browser to another in
-what can be eventually a fully connected topology. 
-%FERGU: explica esto: los navegadores se conectan entre sí o sólo a tu servidor?
+what can be eventually a fully connected topology; however, all
+connections take place between the browser and the server.
 This deals with
@@
 view of the architecture itself and the algorithm that it is actually
-running. First we will describe the client code and next, in
+running. First, we will describe the client code and next, in
 subsection \ref{ss:server} the server architecture and where it is
@@
 
-As stated above, the evolutionary algorithms runs mainly in the
+As stated above, the evolutionary algorithms run mainly in the
 browser. The problem run is a multimodal problem called {\em l-trap},
@@
 our case we have used several values, from 30 to 50 traps, although we
-were not so much focused in finding the solution (or the fastest way
+were not so much focused on finding the solution (or the fastest way
 of finding it) but on creating a experiment that could last for a
@@
 
-This fitness is implemented as part of a (name hidden) evolutionary
-%FERGU: creo que si usas un framework libre puedes dejarlo visible
+This fitness is implemented as part of the Nodeo evolutionary
 algorithm library written in node.js, a server-based implementation of
@@
 between JavaScript on the browser and node.js is the way they load the
-%FERGU: puesto .js para distinguir la palabra node del nombre de la librería
 external libraries. The algorithm used a population of 256
@@
 will play for or against its asynchrony in many different and complex
-ways \cite{jj:2008:PPSN}. %FERGU: este párrafo me lo copio para el de la ASOCO ;)
+ways \cite{jj:2008:PPSN}.
 
@@
 some generations back. These two operations are made in {\em fire and
-  forget} style without any error checking, so that if the server is
-  %FERGU: Qué es fire and forget? (se puede intuir, pero dejarlo claro)
+  forget} style, that is, without any error checking, so that if the server is
 down, or indeed if the algorithm is run locally, it is not interrupted
@@
 The server was also written using node.js and the {\tt express.js}
-module and has three REST \cite{Castillo12REST} %FERGU: cito a este paper que es un capítulo de libro en vez de Arxiv
+module and has three REST \cite{Castillo12REST} %FERGU: cito a este
+                                %paper que es un cap√≠tulo de libro en
+                                %vez de Arxiv
+% OK - JJ
 functions, that is, three functions that can be accessed from any
@@
   that is initialized at the beginning of the experiment.
-\item {\tt PUT one} puts a single individual in the pool. In REST
+\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
   conventions, {\tt PUT} is used to create a resource, that is why it
   is used, in this case, to add a new individual to the pool.
-\item {\tt GET log} returns the experiment log so far. It is used by
-  the researcher to gather data, and as a transparency measure so that
-  anyone can use it. 
 \end{itemize}
@@
 As indicated in the introduction, every part of the experiment and
-data gathered have been released as free software in a GitHub
-repository (address hidden for double blind review). These results
+data gathered have been released as free software in the GitHub
+repository \url{https://github.com/JJ/idc-keynote}. %FERGU: JJ, es
+                                %este el repo correcto, no?
+% Todo esto lo voy a cambiar, no lo mires mucho m·s que por
+% metodologÌa. Ahora uso http://github.com/JJ/splash-volunteer
+These results
 will be analyzed in the next section.
@@
 The litmus test that all experiments should pass, that is, that they
-work and found the solution, passed. %FERGU: que quiere decir esta primera frase?
+work and found the solution, passed. %FERGU: que quiere decir esta
+                                %primera frase?
+% que funcionan. - JJ
 In fact, the first version of the
 algorithm (until January 2015) used too much memory on the browser,
-which crashed after a while running. A total of 7 experiments were
+which crashed after a while running. A total of seven experiments was
+%el "were" es para los experimento o para el "total" A total was? 
+%Correcto - JJ
 done with this version. After January, a new version of the
@@
 the next version, since in this first one we were looking for some
-kind of systematicity in the behavior of users and the experimental
+kind of systematic patterns in the behavior of users and the experimental
 %FERGU: systematicity? systematization mejor?
+% patterns mejor - JJ
 apparatus is the minimum possible to have it running.
@@
         }
-        \caption{Histogram of number of different IPs in an experiment}\label{fig:histo:ips}
+        \caption{Histogram of number of different IPs in a run (7 runs for first experiment and 10 for the second).}\label{fig:histo:ips}
 \end{figure*}
-
+%FERGU: uhm, yo en vez de poenr 2 veces 7 ips distintas, simplemente
+%pondrÌa las barras de Ips distintas (del 1 al 7 y del 1 al 10).
+%jj: no te entiendo, pero los datos est·n por ahÌ - JJ
 The fact that it run correctly during a massive lightning talk means that this setup is a
@@
 a maximum of 28 and a minimum of 6. In the second, the median was 12.5
-with a maximum of 19 and minimum of 8. The histogram of different IPs
+with a maximum of 19 and a minimum of 8. The histogram of different IPs
 in each experiment is shown in figure \ref{fig:histo:ips}. Probably
@@
 willingly, lend to a simple experiment.
-%FERGU: dí que la persona que lo anunció en twitter tiene X seguidores
-%
+%FERGU: d√≠ que la persona que lo anunci√≥ en twitter tiene X seguidores
+% OK, con los nuevos experimentos lo harÈ. De hecho, deberÌamos hacer
+% experimentos anunci·ndolo t˙ y anunci·ndolo yo. - JJ
 \begin{figure*}[htb]
@@
           in each batch of experiments with the max number of
-          different IPs. Please note the algorithmic scale in the $x$ axis.}\label{fig:dist}
+          different IPs. Please note the logarithmic scale in the $x$ axis, and the IPs are anonymous (starting with 10).}\label{fig:dist}
 \end{figure*}
@@
 needed as the number of IPs used increase. However, we will have to
-investigate further the speedups achieved and if these are %FERGU: speedup va junto
+investigate further the speedups achieved and if these are %FERGU:
+                                %speedup va junto
+%si son correcciones de estas no hace falta que las comentes, hazlas y
+%ya est· - JJ
 algorithmic or due to the particular asynchronous and web-based
 implementation. Figure \ref{fig:intervals} might help understand the
-kind of raw material we are using: It plots an histogram of the time
+kind of raw material we are using: It plots a histogram of the time
 needed by all experiments to process 100 generations, 25600
-evaluations and the associated evolutionary geneticry. The vast %FERGU: geneticry?
+evaluations and the associated evolutionary geneticry. The vast
+%FERGU: geneticry?
+%JJ un palabro inventao 
 majority of browsers processes it in less than 3 seconds (logarithm( 
@@
 
-This paper is supported in part by several grants hidden for
-anonymity.
+This work has been supported in part by TIN2011-28627-C04-02 and
+TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
+SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
+GENIL project (CEI-BIOTIC Granada).
 
 \bibliographystyle{abbrv}
-\bibliography{geneura,volunteer,javascript,ror-js,GA-general}
+\bibliography{geneura-latin1,volunteer,javascript,ror-js,GA-general}
 

### 2015-03-31 modeling-volunteer-computing-gecco.tex
Ack'ing previous paper

@@
 
-\section{Materials: running an island model in the browser}
+\section{An island model in the browser}
 \label{sec:exp}
@@
 
-The server was also written using node.js and the {\tt express.js}
-module and has three REST \cite{Castillo12REST} %FERGU: cito a este
-                                %paper que es un cap√≠tulo de libro en
-                                %vez de Arxiv
-% OK - JJ
-functions, that is, three functions that can be accessed from any
-client, including the browser, via a request encoded in an URL and a
-standard HTTP command. These functions are\begin{itemize}
+The server, that was also introduced during \cite{arxiv:modeling}, has
+been rewritten in part to add new features that make experimentation
+easier. I has been written using node.js and the {\tt express.js}
+module and includes the following REST \cite{Castillo12REST} 
+routes:
 \item {\tt GET random} returns a random, non-evaluated individual,
@@
 
-In fact, the program includes more logic to serve the static pages
+These two routes are, in fact, an universal interface that can be
+accessed from any program, written in any language that can connect
+via HTTP.
+
+The server program includes more logic to serve the static pages
 that contain the talk in which it is included and the JavaScript/CSS

### 2015-04-01 modeling-volunteer-computing-gecco.tex
Completing biblio
Expanding server description

@@
 been rewritten in part to add new features that make experimentation
-easier. I has been written using node.js and the {\tt express.js}
-module and includes the following REST \cite{Castillo12REST} 
+easier. I has been written using node.js and the {\tt express.js} and {\tt winston}
+modules for REST and logging and includes the following REST \cite{Castillo12REST} 
 routes:
+\begin{itemize}
 \item {\tt GET random} returns a random, non-evaluated individual,
@@
 
-The server program includes more logic to serve the static pages
-that contain the talk in which it is included and the JavaScript/CSS
-code also, but it all amounts to a small amount of code. This code can
-be deployed, in principle, in any server running node.js, but for the
+The server program includes also additional logic for logging to
+different devices, routes to obtain information from the server and code to serve the static pages
+and the JavaScript/CSS
+code also. Additionally to the code in the previous paper
+\cite{arxiv:modeling} this version restarts the chromosome pool when
+the experiment has finished; it includes a configurable termination
+condition that checks for the solution and empties the chromosome pool
+and logs the end of the experiment. This allows us to run continuously
+experiments without restarting the server by hand as we did in the
+previous version.
+
+All in all, it amounts to a few hundred lines of code
+including all languages. This code can
+be deployed, in principle, in any server that is able to run node.js, but for the
 purposes of this experiment Heroku, a Platform as a Service with a
@@
 or no modification, to OpenShift or other similar PaaS. Far from being
-a freeriding way, using free or very low cost servers can be a
+a cheap way of performing experiments, using free or very low cost servers can be a
 sustainable way of performing massive distributed evolutionary
@@
 
-Having the pages and the server in the same domain allows to work in
+Having the pages and the server in the same domain allows to work in the
 default Ajax mode, that does not allow cross-site requests. However,
-these can also be disabled if needed. In this case, every domain holds
+these can also be enabled if needed so that the static pages and the
+REST server can be hosted in different servers. In this case, every domain holds
 a single experiment, there is no multi-tenancy of experiments, so its

### 2015-04-01 modeling-volunteer-computing-gecco.tex
Finishing description

@@
 
-When the experiment is finished, that is, the solution has been found,
-the server is restarted using  command line tools provided by
-Heroku. Every experiment was first done during the talk and then
-announced in Twitter several times until enough data was gathered.
+During and after the experiment, the data can be collected from one of
+the logs: the log file (if the PaaS allows writing to the filesystem,
+which is not the case in Heroku) or to online logging services such as
+Loggly, which is the one we are using.
 
@@
 data gathered have been released as free software in the GitHub
-repository \url{https://github.com/JJ/idc-keynote}. %FERGU: JJ, es
-                                %este el repo correcto, no?
-% Todo esto lo voy a cambiar, no lo mires mucho m·s que por
-% metodologÌa. Ahora uso http://github.com/JJ/splash-volunteer
-These results
+repository \url{https://github.com/JJ/splash-volunteer}. 
 will be analyzed in the next section.

### 2015-04-01 modeling-volunteer-computing-gecco.tex
Adds loggly description

@@
 which is not the case in Heroku) or to online logging services such as
-Loggly, which is the one we are using.
-
+Loggly, which is the one we are using. Loggly was chosen for its
+easiness and support from the {\tt winston} logging library; however,
+it is only free for a month, so a more sustainable alternative (read:
+free) will have to be sought in the future. Other online logging
+services have a free tier that can be used for ever; however, its
+support and facilities are not so good. Choosing a good logging
+service for experiments in open science is an open issue but it is
+outside the scope of this paper. It will have to be left for future
+work. 
+% Include loggly graphs here.
 As indicated in the introduction, every part of the experiment and

### 2015-04-10 volunteer-clouds-ieee-ci-si.tex
Paper in the new format
Also proposes a title refs #3

%% bare_jrnl.tex
%% V1.4a
%% 2014/09/17
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices and paper sizes can       ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal,onecolumn]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Volunteers in the Clouds: an Architecture for Low Cost Massively Distributed Evolutionary Computation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
        John~Doe,~\IEEEmembership{Fellow,~OSA,}
        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
for IEEE journal papers produced under \LaTeX\ using
IEEEtran.cls version 1.8a and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill September 17, 2014

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2015-07-10 volunteer-clouds-ieee-ci-si.tex
Starting to write this

@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{op-tical net-works}
 
@@
 
-\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
-        John~Doe,~\IEEEmembership{Fellow,~OSA,}
-        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
-\thanks{M. Shell is with the Department
-of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
-GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
-\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
-\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}
+\author{Juan-J.~Merelo, %Add your name here
+}% <-this % stops a space
+\thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
+e-mail: (jmerelo@ugr.es).}% <-this % stops a space
+%Add the rest here
 
@@
 \begin{IEEEkeywords}
-IEEEtran, journal, \LaTeX, paper, template.
+Evolutionary algorithm, distributed computing, internet computing, social networks, 
 \end{IEEEkeywords}

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex
Proposed new title refs #3
Changes format, adds acks and fixes biblio

--- 2015-04-10 volunteer-clouds-ieee-ci-si.tex
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-\appendices
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
-
 % use section* for acknowledgment
@@
 
-The authors would like to thank...
+This work has been supported in part by TIN2011-28627-C04-02 and
+TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
+SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
+GENIL project (CEI-BIOTIC Granada). We would also like to thank the
+anonymous reviewers of previous versions of this paper who have really helped us to improve
+this paper (and our work) with their suggestions.
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
+\bibliographystyle{IEEEtran}
+\bibliography{geneura-latin1,volunteer,javascript,ror-js,GA-general}
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
-\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Jane Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex

@@
 % Do not put math or special symbols in the title.
-\title{Volunteers in the Clouds: an Architecture for Low Cost Massively Distributed Evolutionary Computation}
+\title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation}
 %

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex
First abstract draft
Please @ccotap @mariosky @fergunet take a look at it, even if only to know the general paper layout. Remember that it's a special issue on metaheuristics software.

@@
 \begin{abstract}
-The abstract goes here.
+Volunteer computing is used routinely in science and enterprise but it
+still has a long way to go if it is going to become mainstream in the
+same way the cloud in general has. The lack of general purpose tools
+with analytics built in could be one of the reasons why this is so, as
+well ad the lack of proofs of concept. That is why in this paper we
+present NodIO, a client-server architecture for distributed
+evolutionary algorithms, written entirely in JavaScript, which needs
+minimal server support to work and is able to support persistent,
+asynchronous, distributed evolutionary algorithms running in the
+browser. The architecture, along with some initial results on the expected performance and scaling
+behaviour is presented. We also use the algorithm analytics to try and
+create a model of user behavior which, as is usual in this kind of
+computing models, follow a Weibull distribution. 
 \end{abstract}

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex
Advancing the state of the art
Mainly for @mariosky to check out.

@@
 \usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
 
@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+
 \title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation}
@@
 
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
-%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
@@
 \begin{IEEEkeywords}
-Evolutionary algorithm, distributed computing, internet computing, social networks, 
+Evolutionary algorithm, distributed computing, internet computing,
+social networks, volunteer computing, metacomputing. 
 \end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
 
@@
 Subsubsection text here.
+
+\section{State of the art}
+\label{sec:soa}
+
+The web used as a resource for distributed computing has a
+long history, and almost since the beginning it was used for 
+evolutionary computation even before JavaScript was
+introduced as a browser-based scripting language in 1997. Before that
+time other procedures, including Flash animations and VBScript, ActiveX or Java applets were
+used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
+internet'', providing some advantages such as multi-architecture compatibility or 
+security mechanisms and proposed in \cite{chandy1996world} as the
+basis for the creation of a world-wide peer to peer system using Java
+compiled programs.
+
+Using any of these languages allowed the creation
+of client-server applications in the browser, provided the user had
+installed the virtual machine plugin in it, but with the advantage of
+providing a single and universal way of accessing resources via an
+URI, something that is not guaranteed using desktop applications. However, this style of
+computing presents several issues and challenges which were addressed
+as early as 1996 by Baratloo et al. in their system Charlotte:
+\cite{baratloo1996charlott}. This system, which implemented a
+distributed shared memory, was inspired by previous work
+on parallel computing systems PVM and MPI, but solved most issues by
+using the Java virtual machine embedded in browser, which provides
+security to the user, uniformity of computing environment
+independently of the operating system and  access to a common data store in the
+server. Other issues such as scheduling are solved on the server. This
+paper was later extended and published in a journal in 1999
+\cite{baratloo1999charlotte}, but it is interesting to note that
+already in the early stages of the web the possibility of
+browser-based distributed computing was already researched.
+
+One of the first papers by Soares et al. used Java, and in their paper \cite{soares1998get} they
+describe JET, a system that supports 
+the execution of parallel applications over the Internet. This system has 
+a comprehensive statistics support, allowing its use for science,
+since it allows the collection of algorithm analytics by the
+experiment designer, unlike the other systems compared in that paper.
+
+However, Java (and all the other technologies)
+is not universal in the
+sense than an extra component, namely, the Java virtual machine, has
+to be installed in the browser. JavaScript
+\cite{flanagan2006javascript} was introduced in 1997 as a
+browser-embedded language and has, since then, become a set of standards
+\cite{ECMA-262} for the language, its components and future versions
+that have been widely adopted by the industry and also by scientists,
+who used them for creating a non-distributed EA on
+the browser as early as 1998 \cite{jj-ppsn98}.
+
+The potential for volunteer computing using browsers was realized
+later on \cite{sarmenta-bayanihan} as well as the potential for
+mischief of users with code in their hands
+\cite{sarmenta-sabotagetolerance}. However, these early efforts by
+Sarmenta once again used Java and not JavaScript, making this effort
+less universal. JavaScript can be used either for unwitting
+\cite{unwitting-ec} or volunteer
+\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
+evolutionary computation and it has been used ever since by several
+authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
+used the client's GPU \cite{duda2013gpu}. Many other researchers have
+used Java \cite{chong:1999:jDGPi} and others have gone away from the
+server-based paradigm to embrace peer to peer systems
+\cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
+platforms avoid single points of failure (the server) but, since they
+need a certain amount of infrastructure installed to start, the
+threshold to join them is much lower.
+
+There have been relatively few efforts to analyze what is the
+performance that can be obtained from these volunteer computing
+effort. There was some attempt initially to dodge the issue by making
+the algorithm adaptive to the kind of resources allotted to it
+\cite{milani2004online}, which is actually not such a big problem in
+algorithms such as the EA that can easily be
+parallelized via population splitting or farming out evaluations to all
+the nodes available. Lately, several approaches have focused on the
+fault-tolerance of volunteer algorithms
+\cite{gonzalez2010characterizing} which can, of course, be studied in
+a more general distributed computing context
+\cite{nogueras2015studying} or including it in a more general study of
+performance of the EA itself %Con tu permiso lo paso a acrónimos para
+                             %que no sea tan largo el párrafo. FERGU
+%OK - JJ
+\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
+volunteer computing, number of users and the time spent in the
+computation in browser-based volunteer computing experiments, have only been studied in a limited way in
+\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
+run. Studies using volunteer computing platforms such as SETI@home
+\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
+modeled quite well the availability of resources in several clusters
+of that framework; the shape of those distributions is a skewed bell
+with more resources in the {\em low} areas than in the high areas:
+there are many users that give a small amount of cycles, while there
+are just a few that give many cycles. This is in concordance with the
+results obtained in \cite{agajaj}.
+
+As far as we know, this paper presents the only experiment that uses computational
+resources that are as dissimilar as smartphones and powerful laptops
+or desktop computers in a research center. The methodology used to
+gather resources and the algorithms used will be described next, as
+well as the results obtained with the setup.
 
@@
 \end{document}
+%%% Local Variables:
+%%% ispell-local-dictionary: "english"
+%%% End:

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex
Fixed first page problem

@@
-%% bare_jrnl.tex
-%% V1.4a
-%% 2014/09/17
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
-%% journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
-%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices and paper sizes can       ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[journal,onecolumn]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
 
@@
 
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
+\usepackage{cite}
 
@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
   % declare the path(s) where your graphic files are
@@
   % driver is specified.
-  % \usepackage[dvips]{graphicx}
+  \usepackage[dvips]{graphicx}
   % declare the path(s) where your graphic files are
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/tex-archive/info/epslatex/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
 
@@
 
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Do not use the stfloats baselinefloat ability as IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
-% Axel Sommerfeldt. This package may be useful when used in conjunction with 
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
 % *** PDF, URL AND HYPERLINK PACKAGES ***
@@
 \author{Juan-J.~Merelo, %Add your name here
+  \thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
+e-mail: (jmerelo@ugr.es).}
 }% <-this % stops a space
-\thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
-e-mail: (jmerelo@ugr.es).}% <-this % stops a space
+
 %Add the rest here
-
+% Thanks to http://www.michaelshell.org/tex/ieeetran/ for the fix 
 \maketitle

### 2015-07-26 volunteer-clouds-ieee-ci-si.tex
continuing with historical metacomputing systems

@@
 basis for the creation of a world-wide peer to peer system using Java
-compiled programs.
+compiled programs. The ATLAS system
+\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
+with another language called Cilk; all these papers indicate that
+the introduction of an architecture-independent system such as the
+Java virtual machine opened the doors for ultrascale, Internet-wide,
+metacomputers.
 
@@
 as early as 1996 by Baratloo et al. in their system Charlotte:
-\cite{baratloo1996charlott}. This system, which implemented a
+\cite{baratloo1996charlotte}. This system, which implemented a
 distributed shared memory, was inspired by previous work
@@
 since it allows the collection of algorithm analytics by the
-experiment designer, unlike the other systems compared in that paper.
-
-However, Java (and all the other technologies)
-is not universal in the
-sense than an extra component, namely, the Java virtual machine, has
-to be installed in the browser. JavaScript
-\cite{flanagan2006javascript} was introduced in 1997 as a
-browser-embedded language and has, since then, become a set of standards
-\cite{ECMA-262} for the language, its components and future versions
-that have been widely adopted by the industry and also by scientists,
-who used them for creating a non-distributed EA on
-the browser as early as 1998 \cite{jj-ppsn98}.
-
-The potential for volunteer computing using browsers was realized
+experiment designer, unlike the other systems compared in that
+paper. Contemporary proposals, such as SuperWeb
+\cite{alexandrov1997superweb} also propose a translation of classical
+distribution techniques, with complicated architectures that include
+brokers for resource discovery. It is focused, however, in solving the
+issues involved in browser-based computing and also in making a system
+that is able to work across different classes of computers, mentioning
+SPARCstations and Windows-95 computers. Interestingly enough, this
+paper incorporates a discussion on the economic model, paving the way
+for the for-profit distributed computing systems that later on
+arose. On the other hand, the potential for volunteer computing using
+browsers was realized 
 later on \cite{sarmenta-bayanihan} as well as the potential for
@@
 Sarmenta once again used Java and not JavaScript, making this effort
-less universal. JavaScript can be used either for unwitting
+less universal.
+
+The main problem with Java and all the other browser-embedded
+technologies is that they are not {\em universal} in the
+sense than an extra component, namely, the Java virtual machine, has
+to be installed in the browser. JavaScript, however,
+\cite{flanagan2006javascript} was introduced in 1997 as a
+browser-embedded language and has, since then, become a set of standards
+\cite{ECMA-262} that include the language and its components. Since
+early on it was adopted by the industry and also by scientists,
+who used them for creating a  EA on
+the browser as early as 1998 \cite{jj-ppsn98}.
+
+JavaScript can be used either for unwitting
 \cite{unwitting-ec} or volunteer

### 2015-07-27 volunteer-clouds-ieee-ci-si.tex
Answering @mariosky and accepting his changes

@@
 
+\author{Mario Garcia-Valdez, %Add your name here
+  \thanks{M Garcia-Valdez is with the Dept. of Graduate Studies at Instituto Tecnologico de Tijuana
+e-mail: (mario@tectijuana.edu.mx).}
+}% <-this % stops a space
+
 %Add the rest here
@@
 with analytics built in could be one of the reasons why this is so, as
-well ad the lack of proofs of concept. That is why in this paper we
+well as the lack of proofs of concept. That is why in this paper we
 present NodIO, a client-server architecture for distributed
@@
 introduced as a browser-based scripting language in 1997. Before that
-time other procedures, including Flash animations and VBScript, ActiveX or Java applets were
+time other technologies,
+%other technologies?
+% changed - JJ
+ including Flash animations and VBScript, ActiveX or Java applets were
 used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
 internet'', providing some advantages such as multi-architecture compatibility or 
-security mechanisms and proposed in \cite{chandy1996world} as the
+security mechanisms and was proposed in \cite{chandy1996world} as the
 basis for the creation of a world-wide peer to peer system using Java
@@
 \cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
-with another language called Cilk; all these papers indicate that
+with another language called Cilk; all these papers indicated that
 the introduction of an architecture-independent system such as the
@@
 Using any of these languages allowed the creation
-of client-server applications in the browser, provided the user had
-installed the virtual machine plugin in it, but with the advantage of
+of client-server applications in the browser, provided that the user had
+installed the required virtual machine plugin first, but with the advantage of
 providing a single and universal way of accessing resources via an
@@
 on parallel computing systems PVM and MPI, but solved most issues by
-using the Java virtual machine embedded in browser, which provides
+using the Java virtual machine embedded in the browser,
+which provides
 security to the user, uniformity of computing environment
@@
 early on it was adopted by the industry and also by scientists,
-who used them for creating a  EA on
+who used them for creating an EA on
 the browser as early as 1998 \cite{jj-ppsn98}.
@@
 
-There have been relatively few efforts to analyze what is the
+There have been relatively few efforts to analyze the
 performance that can be obtained from these volunteer computing
-effort. There was some attempt initially to dodge the issue by making
+efforts. 
+%Too many efforts?
+% No, too few - JJ
+There were some initial attempts  to dodge the issue by making
 the algorithm adaptive to the kind of resources allotted to it
@@
 \cite{nogueras2015studying} or including it in a more general study of
-performance of the EA itself %Con tu permiso lo paso a acrónimos para
-                             %que no sea tan largo el párrafo. FERGU
-%OK - JJ
+performance of the EA itself
 \cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
@@
 well as the results obtained with the setup.
+
+%El último párrafo podría se más preciso. Es el único experimento que utiliza recursos heterogeneos
+%empleando las tecnologías estándar (JavaScript, JSON, Cookies, etc.) mencionadas previamente.
+
+%Si no se extiende mucho se podría también hablar de la conexión con Cloud Computing,
+%se discuten las ventajas de utilizar Javascript para acceder a recursos donados por voluntarios
+%pero falta hablar de como se explotaría el recurso Cloud. Lo digo por el titulo de Volunteers in the Clouds. Tal vez
+%se deba hablar de la posibilidad de escalar los sistemas a bajo costo
+%o según la demanda.
+
+% Si te animas, propón un párrafo - JJ
+
+%Algo novedoso es el uso de Javascript no solo en el cliente si no también en el Servidor.
+%Como el tema es el Software creo que esto se tiene que resaltar. En los sistemas anteriores 
+%se tenía lo de Java en el cliente, pero en el server estaba lo grueso del trabajo:
+% Administrar el server - Ahora más fácil con cloud
+% Programar el server side - Ahora se puede usar JavaScript también para eso.
 

### 2015-07-27 volunteer-clouds-ieee-ci-si.tex
Adding global computing desirable features

@@
 Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers.
+metacomputers. This paper also establishes the desirable features of
+any {\em global computing} infrastructure, that can also be applied to
+distributed evolutionary computation experiments that rely in
+volunteer computing: Scalability, heterogeneity, fault tolerance,
+adaptive parallelism, safety, anonymity, hierarchy, ease of use and
+reasonable performance. Of all these, probably {\em hierarchy}, that
+is, the capability of using preferentially resources in the same
+organization could be either disregarded or included in adaptivity,
+leaving us with eight desirable properties.
 
-Using any of these languages allowed the creation
-of client-server applications in the browser, provided that the user had
+Fulfilling these properties could be achieved using any of the
+languages, and associated technologies, mentioned above, provided that the user had
 installed the required virtual machine plugin first, but with the advantage of
@@
 
-One of the first papers by Soares et al. used Java, and in their paper \cite{soares1998get} they
+One of the first papers by Soares et al. used Java, and in their paper
+\cite{soares1998get} they 
 describe JET, a system that supports 

### 2015-07-27 volunteer-clouds-ieee-ci-si.tex
Starting with churn2

@@
 providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed using desktop applications. However, this style of
-computing presents several issues and challenges which were addressed
-as early as 1996 by Baratloo et al. in their system Charlotte:
+URI, something that is not guaranteed using desktop
+applications. Baratloo et al. tried in 1996 to create a system with
+the aforementioned qualities called {\em Charlotte}
 \cite{baratloo1996charlotte}. This system, which implemented a
 distributed shared memory, was inspired by previous work
-on parallel computing systems PVM and MPI, but solved most issues by
+on parallel computing systems PVM and MPI and solved most issues by
 using the Java virtual machine embedded in the browser,
@@
 independently of the operating system and access to a common data store in the
-server. Other issues such as scheduling are solved on the server. This
+server. Since the virtual machine is a sandbox, it is safe to the user
+and running a compiled bytecode guarantees a certain level of
+performance. 
+ Other issues such as scheduling, which are related to adaptive
+ parallelism, are solved on the server. This 
 paper was later extended and published in a journal in 1999
@@
 
-One of the first papers by Soares et al. used Java, and in their paper
-\cite{soares1998get} they 
-describe JET, a system that supports 
-the execution of parallel applications over the Internet. This system has 
-a comprehensive statistics support, allowing its use for science,
+Other systems, like JET, proposed by Soares et
+al. \cite{soares1998get}, also used Java. JET supports 
+the execution of parallel applications over the Internet and features
+a comprehensive statistics collection subsystem, allowing its use for science,
 since it allows the collection of algorithm analytics by the
-experiment designer, unlike the other systems compared in that
-paper. Contemporary proposals, such as SuperWeb
-\cite{alexandrov1997superweb} also propose a translation of classical
+experiment designer, unlike the other systems mentioned
+above. Proposals contemporary to Jet, such as SuperWeb
+\cite{alexandrov1997superweb} are roughly a translation of classical
 distribution techniques, with complicated architectures that include
@@
 %Too many efforts?
-% No, too few - JJ
+% No, too little - JJ
 There were some initial attempts  to dodge the issue by making
@@
 
+On the other hand, measuring performance of the resulting metacomputer
+involves understanding the dynamics of this kind of systems. Initial
+work was done for peer to peer systems by Stutzbach et
+al. \cite{stutzbach2006understanding} and extended to volunteer
+computing by Laredo et al. \cite{churn08,laredo2008rcp}.
+
+%------------------------------------------------------------------------------------
 % An example of a floating figure using the graphicx package.
@@
 
+\section{Description of the system}
+
 \section{Conclusion}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{geneura-latin1,volunteer,javascript,ror-js,GA-general}
+\bibliography{geneura,volunteer,javascript,ror-js,GA-general}
 

### 2015-07-28 volunteer-clouds-ieee-ci-si.tex
Some minor corrections and a paragraph and reference to justify JS usage. Also, comment for @JJ

@@
 \begin{abstract}
-Volunteer computing is used routinely in science and enterprise but it
+Volunteer computing is used routinely in science and enterprise, but it
 still has a long way to go if it is going to become mainstream in the
@@
 browser. The architecture, along with some initial results on the expected performance and scaling
-behaviour is presented. We also use the algorithm analytics to try and
+behavior is presented. We also use the algorithm analytics to try and
 create a model of user behavior which, as is usual in this kind of
@@
 any {\em global computing} infrastructure, that can also be applied to
-distributed evolutionary computation experiments that rely in
+distributed evolutionary computation experiments that rely on
 volunteer computing: Scalability, heterogeneity, fault tolerance,
@@
 \cite{ECMA-262} that include the language and its components. Since
-early on it was adopted by the industry and also by scientists,
+early on, it was adopted by the industry and also by scientists,
 who used them for creating an EA on
@@
 a more general distributed computing context
-\cite{nogueras2015studying} or including it in a more general study of
+\cite{nogueras2015studying} or including it in a more general study of the
 performance of the EA itself
@@
 
-As far as we know, this paper presents the only experiment that uses computational
+%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
+Javascript is also compatible with the functional programming paradigm. According to 
+\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
+metaheuristics optimization allows an improvement in communicability, 
+reproducibility, interoperability, automated assembly, knowledge engineering 
+and efficiency.
+
+As far as we know, this paper presents the only experiment that uses computational %FERGU: yo no estaría tan seguro de esto
 resources that are as dissimilar as smartphones and powerful laptops

### 2015-07-29 volunteer-clouds-ieee-ci-si.tex
Comment
Cloud and server side JS

@@
 
+Recent works propose the use of cloud computing services for the components of 
+client-server distributed EAs. Cloud resources 
+can reduce operational costs by outsourcing both hardware and software maintenance 
+to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what 
+is available in most laboratories, allowing researches to 
+scale algorithms at reasonable costs. Researchers have proposed the use of 
+public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
+and Drop Box\cite{garcia2011}. The use of JavaScript to implement server side components, 
+gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
+in most Platforms as a Service (PaaS) as many support the nodejs framework. Also many web services
+have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
+Developers can use the same software tools for both client and server side development.
+%Ayuda con la mini conclusion
+%In this work JavaScript was used in the full stack, some percieved advantages are discussed?
+%The proposed software has the advantage of full stack JavaScript development..
+
 There have been relatively few efforts to analyze the
@@
 % No, too little - JJ 
+% Digo efforts en la oración anterior :) - Mario
 There were some initial attempts  to dodge the issue by making
@@
 
+%Mario: Estaba pensando que otra justificación muy fuerte para el uso de JS sobre plug-ins 
+%ha sido la seguridad ya que estos presentan otra fuente de vulnerabilidad. Para complicar las cosas a veces (Flash, ActiveX) 
+%son tecnologías cerradas y propietarias. Por lo menos creo que esas fueron las razones que se dieron para eliminar a Flash del iPhone.
+
 As far as we know, this paper presents the only experiment that uses computational %FERGU: yo no estaría tan seguro de esto
@@
 
-%El último párrafo podría se más preciso. Es el único experimento que utiliza recursos heterogeneos
+%El último párrafo podría ser más preciso. Es el único experimento que utiliza recursos heterogeneos
 %empleando las tecnologías estándar (JavaScript, JSON, Cookies, etc.) mencionadas previamente.
@@
 % Si te animas, propón un párrafo - JJ
+% Ok, lo escribo. - Mario
+% Voy escribiendo algo de HTML5 y Web Workers?, no necesariemente para esta sección.
 

### 2015-07-29 volunteer-clouds-ieee-ci-si.tex
OK, perdón @fergunet
Ha sido @mariosky, que no sabe lo de la bibliografía, quien lo ha incluido. En todo caso, Mario, si quieres incluir algún item de biblio que sea de nuestro grupo usamos directamente un fichero que tenemos también en GitHub.

Perdón a ambos. Ya está corregido.

@@
 public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
-and Drop Box\cite{garcia2011}. The use of JavaScript to implement server side components, 
+and DropBox~\cite{DBLP:conf/cec/ArenasGMCRL11}. The use of JavaScript to implement server side components, 
 gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen

### 2015-07-29 volunteer-clouds-ieee-ci-si.tex
Made a first attempt at introduction
It would really help if our expert in intros, @ccottap, would help with it.

@@
 \begin{abstract}
-Volunteer computing is used routinely in science and enterprise, but it
+Volunteer computing is used routinely in science and industry, but it
 still has a long way to go if it is going to become mainstream in the
@@
 \section{Introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-for IEEE journal papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8a and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
+\IEEEPARstart{N}{owadays}, there is a big and increasing  number of
+distributed evolutionary computation software frameworks available on
+the market, usually as open source systems. Most of these systems have
+two features: they are {\em desktop} applications, that is, they must
+be compiled or run directly on the operating system layer and, second,
+they assume that the nodes that are going to run the system are under
+your control or at least you have access to them.
 
-\hfill September 17, 2014
+There is, then, space for systems that do not have any of those
+assumptions, such as the one presented in this paper, which will be
+provisionally named {\sf NodIO}. NodIO is first an application that
+runs in two tiers, a server and a client that can be possibly run on a
+browser. As such, it can be run in any computer that loads the web
+page where the client is embedded, including possibly many over which
+the scientist has no control.
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+When faced with the problem of diminishing funds for buying new
+hardware and continuous hardware failures of existing clusters and
+taking into account our experience of working with web applications,
+including JavaScript, from the beginning and other volunteer and
+unconventional distributed evolutionary computing systems, our group
+decided to create a new free framework that would allow anyone to
+create a volunteer distributed EC experiment using cloud resources as
+servers and browsers as clients.
 
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
+These systems must solve a series of issues before even starting to
+measure performance; these issues are related to security as well as
+safety, which introduces a {\em social} aspect to the design of a
+computing system that makes it a {\em techno-social system}
+\cite{vespignani2009predicting}. In this paper our intention is to
+present the {\sf NodIO} framework and analyze it from two different
+angles: the purely technical angle, including the decisions that went
+into its features, and the social angle, to account on how the persons
+that voluntarily choose to enter the web page that runs the system
+determine its performance and how it can be modeled and, eventually,
+predicted.
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+The rest of the paper is organized as follows. Next we present the
+state of the art in web-based distributed computation systems along
+with attempts to predict and model its behavior.
 

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Completed sentence by @mariosky

@@
 %Ayuda con la mini conclusion
-%In this work JavaScript was used in the full stack, some percieved advantages are discussed?
-%The proposed software has the advantage of full stack JavaScript development..
+
+In the work presented in this paper, JavaScript was used in the full
+stack, which has the advantage of using the same code for writing
+desktop or command line and browser clients, both based in the EC
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages
+expertise acquired in the library itself. 
+%The proposed software has the advantage of full stack JavaScript
+%development..
 

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Rearranging SoA by @fergunet and @mariosky

@@
 %development..
+%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
+Javascript is also compatible with the functional programming paradigm. According to 
+\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
+metaheuristics optimization allows an improvement in communicability, 
+reproducibility, interoperability, automated assembly, knowledge engineering 
+and efficiency.
 
@@
 
-%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
-Javascript is also compatible with the functional programming paradigm. According to 
-\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
-metaheuristics optimization allows an improvement in communicability, 
-reproducibility, interoperability, automated assembly, knowledge engineering 
-and efficiency.
-
 %Mario: Estaba pensando que otra justificación muy fuerte para el uso de JS sobre plug-ins 
@@
 
-As far as we know, this paper presents the only experiment that uses computational %FERGU: yo no estaría tan seguro de esto
+As far as we know, this paper presents one of the few experiments that
+uses computational 
+%FERGU: yo no estaría tan seguro de esto
+% Changed - JJ
 resources that are as dissimilar as smartphones and powerful laptops
@@
 % Ok, lo escribo. - Mario
-% Voy escribiendo algo de HTML5 y Web Workers?, no necesariemente para esta sección.
+% Voy escribiendo algo de HTML5 y Web Workers?, no necesariemente para
+% esta sección. 
+% Para la intro o esta sección - JJ
 

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
No aparecía el primer autor (arreglado)

@@
 
-\author{Juan-J.~Merelo, %Add your name here
-  \thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
-e-mail: (jmerelo@ugr.es).}
-}% <-this % stops a space
+%  \author{Juan-J.~Merelo, %Add your name here
+%  \thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
+%  e-mail: (jmerelo@ugr.es).}
+%  }% <-this % stops a space
 
-\author{Mario Garcia-Valdez, %Add your name here
-  \thanks{M Garcia-Valdez is with the Dept. of Graduate Studies at Instituto Tecnologico de Tijuana
-e-mail: (mario@tectijuana.edu.mx).}
-}% <-this % stops a space
+%  \author{Mario Garcia-Valdez, %Add your name here
+%    \thanks{M Garcia-Valdez is with the Dept. of Graduate Studies at Instituto Tecnologico de Tijuana
+%  e-mail: (mario@tectijuana.edu.mx).}
+%  }% <-this % stops a space
+
+\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$
+\thanks{Manuscript submitted for review on \today.}%
+\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
+\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
+\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es}%
+\thanks{*Corresponding author.}%
+}
 

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Finishing state of the art with some changes please @fergunet and @mariosky take a look

@@
 performance of the EA itself
-\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
-volunteer computing, number of users and the time spent in the
-computation in browser-based volunteer computing experiments, have only been studied in a limited way in
-\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
-run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
-modeled quite well the availability of resources in several clusters
-of that framework; the shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas:
-there are many users that give a small amount of cycles, while there
-are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}.
+\cite{DBLP:journals/gpem/LaredoBGVAGF14}.
 
@@
 al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}.
-
-%------------------------------------------------------------------------------------
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
+computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
+volunteer computing, number of users and the time spent in the
+computation in browser-based volunteer computing experiments, have only been studied in a limited way in
+\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
+run. Studies using volunteer computing platforms such as SETI@home
+\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
+modeled quite well the availability of resources in several clusters
+of that framework; the shape of those distributions is a skewed bell
+with more resources in the {\em low} areas than in the high areas:
+there are many users that give a small amount of cycles, while there
+are just a few that give many cycles. This is in concordance with the
+results obtained in \cite{agajaj}. Our initial tests indicate that
+this kind of distribution might be the most appropriate to describe
+user participation in several different metrics. However, in this
+paper we will perform different experiments in different setups and
+will try to come up with a more precise model. Next we describe the
+system itself.
 

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
some comments

@@
 
+% [Pedro] no aparecía el primer autor. En cualquier caso, lo dejo comentado a continuación.
+
 %  \author{Juan-J.~Merelo, %Add your name here
@@
 %  }% <-this % stops a space
+
+% [Pedro] Alternativa para poner los autores. Ahora sí los muestra correctamente.
 
@@
 still has a long way to go if it is going to become mainstream in the
-same way the cloud in general has. The lack of general purpose tools
+same way the cloud technologies in general have. The lack of general purpose tools
 with analytics built in could be one of the reasons why this is so, as
@@
 browser. The architecture, along with some initial results on the expected performance and scaling
-behavior is presented. We also use the algorithm analytics to try and
+behavior is presented. We also use the algorithm analytics to try and  %[Pedro] I don't understand what "the algorithm analytics" stand for???
 create a model of user behavior which, as is usual in this kind of

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
some comments

@@
 provisionally named {\sf NodIO}. NodIO is first an application that
-runs in two tiers, a server and a client that can be possibly run on a
+runs in two tiers, a server and a client that can be possibly run on a  % [Pedro] Both the server and the client in browsers? different browsers as I understand?
 browser. As such, it can be run in any computer that loads the web
@@
 
-When faced with the problem of diminishing funds for buying new
+When faced with the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
 hardware and continuous hardware failures of existing clusters and

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Minor changes and comments for @JJ for issue #11

--- 2015-07-30 volunteer-clouds-ieee-ci-si.tex
@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$
+\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'a-S\'anchez$^1$
 \thanks{Manuscript submitted for review on \today.}%
@@
 \thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
-\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es}%
+\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es, pablogarcia@ugr.es}%
 \thanks{*Corresponding author.}%
@@
 runs in two tiers, a server and a client that can be possibly run on a
-browser. As such, it can be run in any computer that loads the web
+browser. As such, it can be run on any computer that loads the web
 page where the client is embedded, including possibly many over which
@@
 distributed evolutionary computation experiments that rely on
-volunteer computing: Scalability, heterogeneity, fault tolerance,
+volunteer computing: scalability, heterogeneity, fault tolerance,
 adaptive parallelism, safety, anonymity, hierarchy, ease of use and
@@
 public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
-and DropBox~\cite{DBLP:conf/cec/ArenasGMCRL11}. The use of JavaScript to implement server side components, 
+and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
 gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
-in most Platforms as a Service (PaaS) as many support the nodejs framework. Also many web services
+in most Platforms as a Service (PaaS) as many support the Node.js framework 
+\footnote{\url{https://strongloop.com/strongblog/a-comparison-of-node-js-paas-hosting-providers/}}.
+Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias
 have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
-Developers can use the same software tools for both client and server side development.
-%Ayuda con la mini conclusion
+developers can use the same software tools for both client and server side development. 
+Besides the advantages to deal with the same language in the full stack as previously mentioned, 
+this also facilitates the interoperability of the exchanged data types.
+%Ayuda con la mini conclusion FERGU: mira la última frase de arriba
 
@@
 desktop or command line and browser clients, both based in the EC
-library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages
-expertise acquired in the library itself. 
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages %FERGU: párrafo en pasado "was used" y presente "saves"
+expertise acquired in the library itself. %FERGU: tienes info de LoC usadas para la sección experimental y justificar esta frase?
 %The proposed software has the advantage of full stack JavaScript

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex

@@
 provisionally named {\sf NodIO}. NodIO is first an application that
-runs in two tiers, a server and a client that can be possibly run on a
-browser. As such, it can be run on any computer that loads the web
+runs in two tiers, a server and a client that can be possibly run on a  % [Pedro] Both the server and the client in browsers? different browsers as I understand?
+browser. As such, it can be run in any computer that loads the web
 page where the client is embedded, including possibly many over which

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
some comments

--- 2015-07-30 volunteer-clouds-ieee-ci-si.tex
@@
 
+%---------------------------------------------------------------
 \section{Introduction}
@@
 
-The rest of the paper is organized as follows. Next we present the
-state of the art in web-based distributed computation systems along
-with attempts to predict and model its behavior.
+The rest of the paper is organized as follows: Next we present the
+state of the art (Section \ref{sec:soa}) in web-based distributed 
+computation systems along with attempts to predict and model its behavior. 
+Then, a detailed description of the proposed system is presented 
+in Section \ref{sec:description}, followed by the experiments and 
+obtained results in Section \ref{sec:experiments}.
+Finally, conclusions and future work are presented in Section \ref{sec:conclusion}.
 
+%---------------------------------------------------------------
 \section{State of the art}
@@
 
+%---------------------------------------------------------------
 \section{Description of the system}
+\label{sec:description}
 
+%---------------------------------------------------------------
+\section{Experiments and obtained results}
+\label{sec:experiments}
+
+%---------------------------------------------------------------
 \section{Conclusion}
+\label{sec:conclusion}
+
 The conclusion goes here.
 
+%---------------------------------------------------------------
 % use section* for acknowledgment

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
some comments
Partial improvement

--- 2015-07-30 volunteer-clouds-ieee-ci-si.tex
@@
 
-% [Pedro] no aparecía el primer autor. En cualquier caso, lo dejo comentado a continuación.
-
 %  \author{Juan-J.~Merelo, %Add your name here
@@
 %  }% <-this % stops a space
-
-% [Pedro] Alternativa para poner los autores. Ahora sí los muestra correctamente.
 
@@
 still has a long way to go if it is going to become mainstream in the
-same way the cloud technologies in general have. The lack of general purpose tools
+same way the cloud in general has. The lack of general purpose tools
 with analytics built in could be one of the reasons why this is so, as
@@
 browser. The architecture, along with some initial results on the expected performance and scaling
-behavior is presented. We also use the algorithm analytics to try and  %[Pedro] I don't understand what "the algorithm analytics" stand for???
+behavior is presented. We also use the algorithm analytics to try and
 create a model of user behavior which, as is usual in this kind of
@@
 
-%---------------------------------------------------------------
 \section{Introduction}
@@
 provisionally named {\sf NodIO}. NodIO is first an application that
-runs in two tiers, a server and a client that can be possibly run on a  % [Pedro] Both the server and the client in browsers? different browsers as I understand?
+runs in two tiers, a server and a client that can be possibly run on a
 browser. As such, it can be run in any computer that loads the web
@@
 
-When faced with the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
+When faced with the problem of diminishing funds for buying new
 hardware and continuous hardware failures of existing clusters and
@@
 
-The rest of the paper is organized as follows: Next we present the
-state of the art (Section \ref{sec:soa}) in web-based distributed 
-computation systems along with attempts to predict and model its behavior. 
-Then, a detailed description of the proposed system is presented 
-in Section \ref{sec:description}, followed by the experiments and 
-obtained results in Section \ref{sec:experiments}.
-Finally, conclusions and future work are presented in Section \ref{sec:conclusion}.
+The rest of the paper is organized as follows. Next we present the
+state of the art in web-based distributed computation systems along
+with attempts to predict and model its behavior.
 
-%---------------------------------------------------------------
 \section{State of the art}
@@
 
-%---------------------------------------------------------------
 \section{Description of the system}
-\label{sec:description}
 
-%---------------------------------------------------------------
-\section{Experiments and obtained results}
-\label{sec:experiments}
+In general, a volunteer-based distributed evolutionary computation
+system based on the browser is simply a client-server architecture
+whose client is, or can be, embedded in the browser via
+JavaScript. All parts of the system are free and available with a free
+license from \url{https://github.com/JJ/splash-volunteer}.
+Thus, {\sf NodIO} has two parts:\begin{enumerate}
+\item A REST server, that is, a server that includes several remote
+  requests for storing and retrieving information (the CRUD cycle:
+  create, request, update and delete) from the server. The information
+  is of two kinds: {\em problem} based, that is, related to the
+  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
+    GET}ing a random chromosome from it. It also peforms logging
+  duties, but is basically a very lightweight and hign performance
+  data store \cite{jj:idc:lowcost}.
+\item A client that includes the evolutionary algorithm as 
+  JavaScript code embedded in a web page that displays graphs and some
+  additional links and information on the experiment. This code runs
+  an evolutionary algorithm {\em island} starting with random
+  population and sending, every 100 generations, the best individual
+  back to the server (via a {\tt PUT} request) and requesting a random
+  individual from the server (via a {\tt GET} request).
+\end{enumerate}
 
-%---------------------------------------------------------------
+The researcher only has to change a function to solve a different
+problem, in this case the classical Trap function has been
+used. JavaScript is a functional language and
+
 \section{Conclusion}
-\label{sec:conclusion}
-
 The conclusion goes here.
 
-%---------------------------------------------------------------
 % use section* for acknowledgment

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex

--- 2015-07-30 volunteer-clouds-ieee-ci-si.tex
@@
 
+In general, a volunteer-based distributed evolutionary computation
+system based on the browser is simply a client-server architecture
+whose client is, or can be, embedded in the browser via
+JavaScript. All parts of the system are free and available with a free
+license from \url{https://github.com/JJ/splash-volunteer}.
+Thus, {\sf NodIO} has two parts:\begin{enumerate}
+\item A REST server, that is, a server that includes several remote
+  requests for storing and retrieving information (the CRUD cycle:
+  create, request, update and delete) from the server. The information
+  is of two kinds: {\em problem} based, that is, related to the
+  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
+    GET}ing a random chromosome from it. It also peforms logging
+  duties, but is basically a very lightweight and hign performance
+  data store \cite{jj:idc:lowcost}.
+\item A client that includes the evolutionary algorithm as 
+  JavaScript code embedded in a web page that displays graphs and some
+  additional links and information on the experiment. This code runs
+  an evolutionary algorithm {\em island} starting with random
+  population and sending, every 100 generations, the best individual
+  back to the server (via a {\tt PUT} request) and requesting a random
+  individual from the server (via a {\tt GET} request).
+\end{enumerate}
+
+The researcher only has to change a function to solve a different
+problem, in this case the classical Trap function has been
+used. JavaScript is a functional language and
+
 %---------------------------------------------------------------

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Changed title to final form closes #3
System explained

@@
 
-\title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation}
+\title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation}
 %
@@
 
-\IEEEPARstart{N}{owadays}, there is a big and increasing number of
+\IEEEPARstart{N}{owadays} there is a big and increasing number of
 distributed evolutionary computation software frameworks available on
@@
 that is able to work across different classes of computers, mentioning
-SPARCstations and Windows-95 computers. Interestingly enough, this
+SPARCstations and PCs with Windows-95. Interestingly enough, this
 paper incorporates a discussion on the economic model, paving the way
@@
 uses computational 
-%FERGU: yo no estaría tan seguro de esto
-% Changed - JJ
 resources that are as dissimilar as smartphones and powerful laptops
@@
   duties, but is basically a very lightweight and hign performance
-  data store \cite{jj:idc:lowcost}.
+  data store \cite{jj:idc:lowcost}. The server has the capability to
+  run a single experiment, storing the chromosomes in a data structure
+  that is reset when the solution is found. 
 \item A client that includes the evolutionary algorithm as 
@@
   back to the server (via a {\tt PUT} request) and requesting a random
-  individual from the server (via a {\tt GET} request).
+  individual from the server (via a {\tt GET} request). Apart from
+  these constants, the algorithm can be run in many different ways:
+  stopping when solution is found or, as we have done in this paper,
+  using WebWorkers, which run in the background even if the webpage
+  with the experiment is not currently on focus. 
 \end{enumerate}
@@
 problem, in this case the classical Trap function has been
-used. JavaScript is a functional language and
+used. JavaScript is a functional language and declaring a different
+function and handling at the creation of the algorithm object, called
+{\tt Classic} is the only change needed to work with a different
+problem. Let us see how this system addresses the different challenges
+outlined in the State of the Art:\begin{itemize}
+\item {\em Scalability} is provided via the use of a lightweight and
+  high-peformance, single-threaded, server based in NodeJS and
+  Express.js. Although this single server is a bottleneck since it
+  will eventually saturate, the fact that it runs as a single thread
+  allows the service of many requests. In fact, a limit with the
+  number of simultaneous requests will be reached but so far it has
+  not been found, unlike what we found in our previous systems, DCoR,
+  \cite{gecco07:workshop:dcor}, which had a low scalability.
+\item The system is {\em heterogeneous} since it does not need any
+  performance, operating system or even browser requirement: anyone
+  visiting the page can load the algorithm.
+\item {\em Fault tolerance} is always an issue, and in this case, the
+  single point of failure would be the server: the system, as such,
+  would break down if the server fails. However, the individual
+  islands in every browser would continue running, and having access
+  to just one of them would allow the local algorithm to proceed. In
+  fact, the island does not need the server to run: it runs locally if
+  needed, with the only exception that it is obviously unable to
+  communicate with the rest of the islands.
+\item {\em Adaptiveness} is achieved simply through autonomous
+  operation of every individual island and no synchronization
+  mechanism. The islands in the system are, in fact, unaware of each
+  other, communicating only through the server.
+\item Since the algorithm is run on the browser, {\em safety} is
+  achieved through its sandbox mechanisms. The user is thus assured
+  that there is no unsafe access either to their local files or even
+  to more resources that the browser should be allotted.
+\item Running the algorithm is just a matter of loading the page,
+  which means that operation is totally {\em anonymous}. For the same
+  reason, {\em ease of use} is optimal, being as easy as simply
+  clicking on an URL and available to anyone with access to a browser.
+\item {\em Reasonable performance} is not ensured. In fact, we should
+  make sure that there is a reasonable amount of clients over which
+  the performance achieved is bigger than what you would obtain in
+  your own desktop system. If that is not done, it is a pointless
+  academic exercise.
+\end{itemize}
+
+This last point is what we will try to prove next.
 

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
adding some references to Functional Programming Paradigm

@@
 %FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
-Javascript is also compatible with the functional programming paradigm. According to 
-\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
+Javascript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
+According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
 metaheuristics optimization allows an improvement in communicability, 

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
Provisional results

--- 2015-07-30 volunteer-clouds-ieee-ci-si.tex
@@
   stopping when solution is found or, as we have done in this paper,
-  using WebWorkers, which run in the background even if the webpage
-  with the experiment is not currently on focus. 
+  using WebWorkers, which run in the background even if the webpage or
+  browser tab with the experiment is not currently on focus.  
 \end{enumerate}
@@
 
+First we establish the baseline performance by running the
+evolutionary algorithm in a desktop client written using {\sf NodEO},
+the node.js evolutionary algorithm library. This experiment tries to
+find the solution to the 40-trap function with parameters $l=4, a=1,
+b=2, z=3$, a population size equal 512, which was found in previous
+experiments to be the correct one for finding the solution 95\% of the
+time. The algorithm was run until the solution (a string with all
+ones) was found or five millions evaluations had been performed. It
+took around a minute, on average, that is $68.9694$ seconds, to
+perform the fifty runs.
+
 %---------------------------------------------------------------

### 2015-07-31 data/zipf-ips-worker.R

@@ -1,0 +1,8 @@
+ips.worker.cartones <- read.csv('ips-worker.csv')
+p.ips.worker.cartones <- ips.worker.cartones$puts/sum(ips.worker.cartones$puts )
+
+lzipf <- function(s,N) -s*log(1:N)-log(sum(1/(1:N)^s))
+
+opt.f <- function(s) sum((log(p.ips.worker.cartones)-lzipf(s,length(p.ips.worker.cartones)))^2)
+
+opt <- optimize(opt.f,c(0.5,10))

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
Plotting duration

@@
 took around a minute, on average, that is $68.9694$ seconds, to
-perform the fifty runs.
+perform the fifty runs. In this experiment only 33, that is, 66\% were
+successful.
 

### 2015-07-31 data/zipf-ips-worker-v2.R
Advancing experimental data

@@ -1,0 +1,8 @@
+ips.worker.cartones.v2 <- read.csv('ips-worker-v2.csv')
+p.ips.worker.cartones.v2 <- ips.worker.cartones.v2$puts/sum(ips.worker.cartones.v2$puts )
+
+lzipf <- function(s,N) -s*log(1:N)-log(sum(1/(1:N)^s))
+
+opt.f <- function(s) sum((log(p.ips.worker.cartones.v2)-lzipf(s,length(p.ips.worker.cartones.v2)))^2)
+
+opt <- optimize(opt.f,c(0.5,10))

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
Completes references
Including one to jsEO requested vy @vrivas. Closes #14

@@
 authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
-used the client's GPU \cite{duda2013gpu}. Many other researchers have
+used the client's GPU \cite{duda2013gpu} or create an object-oriented
+framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
 used Java \cite{chong:1999:jDGPi} and others have gone away from the
 server-based paradigm to embrace peer to peer systems
-\cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
+\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
 platforms avoid single points of failure (the server) but, since they

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
We're all in
Names checked and closes #12
Adds plots

@@
 \thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
-\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es, pablogarcia@ugr.es}%
+\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
+    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
 \thanks{*Corresponding author.}%
@@
 
-%Add the rest here
-% Thanks to http://www.michaelshell.org/tex/ieeetran/ for the fix 
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
 \begin{abstract}
@@
 time other technologies,
-%other technologies?
-% changed - JJ
  including Flash animations and VBScript, ActiveX or Java applets were
@@
 
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{myfigure}
+\caption{Comparison of time to solution for the baseline system, using a {\tt
+    node.js} client, for different population sizes. They only compare
+runs in which solution has been found.}
+%\label{fig_sim}
+%\end{figure}
 First we establish the baseline performance by running the
@@
 perform the fifty runs. In this experiment only 33, that is, 66\% were
-successful.
+successful. The experiments were repeated for population $p=1024$,
+with success rate upgraded to 100\%. Results for these two experiments
+are shown in Figure \ref{fig:baseline}.
 

### 2015-07-31 data/plot-zipf.R

@@ -1,0 +1,2 @@
+library(ggplot2)
+ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4,aes(x=x/length(dt.puts.4$x),y=y,color='~1400'),stat='Identity')+geom_point(data=dt.puts.3,aes(x=x/length(dt.puts.3$x),y=y,color='~ 800'),stat='Identity')+geom_point(data=dt.puts.1,aes(x=x/length(dt.puts.1$x),y=y,color='~ 500'),stat='Identity')

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex

@@
 
+Figure \ref{fig:system} describes the system.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=4in]{img/system.pdf}
+\caption{Description of the used system. Clients execute a JS EA in the browser. Every 100 generations it sends the best individual and receives a random one from the server.}
+\label{fig:system}
+\end{figure}
+
 The researcher only has to change a function to solve a different
@@
 \centering
-\includegraphics[width=2.5in]{myfigure}
+%\includegraphics[width=2.5in]{myfigure}
 \caption{Comparison of time to solution for the baseline system, using a {\tt
@@
 runs in which solution has been found.}
-%\label{fig_sim}
-%\end{figure}
+\label{fig_sim}
+\end{figure}
 First we establish the baseline performance by running the

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
Footnote converted to reference
@fergunet you shouldn't use footnotes to point to documents, create a bib entry and cite it instead.

@@
 gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
-in most Platforms as a Service (PaaS) as many support the Node.js framework 
-\footnote{\url{https://strongloop.com/strongblog/a-comparison-of-node-js-paas-hosting-providers/}}.
+in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}.
 Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
First draft with actual plots
Closes #10
Revised figure by @fergunet looks OK closes #13

@@
 
-Figure \ref{fig:system} describes the system.
-
+Figure \ref{fig:system} describes the general system architecture.
+%
 \begin{figure}[!t]
@@
 \includegraphics[width=4in]{img/system.pdf}
-\caption{Description of the used system. Clients execute a JS EA in the browser. Every 100 generations it sends the best individual and receives a random one from the server.}
+\caption{Description of the used system. Clients execute a JS EA in the browser, which, every 100 generations, sends the best individual and receives a random one from the server. } 
 \label{fig:system}
@@
 \centering
-%\includegraphics[width=2.5in]{myfigure}
+\includegraphics[width=12cm]{img/baseline-times.png}
 \caption{Comparison of time to solution for the baseline system, using a {\tt
@@
 runs in which solution has been found.}
-\label{fig_sim}
+\label{fig:baseline}
 \end{figure}

### 2015-08-01 data/time-vs-IPs.R
Adding interesting graphics

@@ -1,0 +1,7 @@
+library(ggplot2)
+
+ips.time <- read.csv("ips-time.csv")
+for ( i in 1:nrow(ips.time) ) {
+    ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
+}
+ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")

### 2015-08-01 data/time-vs-IPs.R
New beautiful graphs

@@ -1,7 +1,11 @@
 library(ggplot2)
 
 ips.time <- read.csv("ips-time.csv")
+ips.time$howmany <- 1:nrow(ips.time)
 for ( i in 1:nrow(ips.time) ) {
     ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
 }
+ips.time$i <- 1:nrow(ips.time)
 ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
+ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
+

### 2015-08-03 volunteer-clouds-ieee-ci-si.tex
Minor changes and a couple of comments for @JJ

@@
 The researcher only has to change a function to solve a different
-problem, in this case the classical Trap function has been
+problem, in this case the classical Trap function has been %FERGU: cita a Trap function
 used. JavaScript is a functional language and declaring a different
-function and handling at the creation of the algorithm object, called
+function and handling at the creation of the algorithm object, called %FERGU: poner algo tras el handling: handling it?
 {\tt Classic} is the only change needed to work with a different
@@
 \item {\em Scalability} is provided via the use of a lightweight and
-  high-peformance, single-threaded, server based in NodeJS and
+  high-performance, single-threaded, server based in NodeJS and
   Express.js. Although this single server is a bottleneck since it
@@
   allows the service of many requests. In fact, a limit with the
-  number of simultaneous requests will be reached but so far it has
+  number of simultaneous requests will be reached, but so far it has
   not been found, unlike what we found in our previous systems, DCoR,
@@
   performance, operating system or even browser requirement: anyone
-  visiting the page can load the algorithm.
+  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
 \item {\em Fault tolerance} is always an issue, and in this case, the

### 2015-08-03 data/time-vs-IPs-openshift.R
Plotting for old experiments

@@ -1,0 +1,11 @@
+library(ggplot2)
+
+ips.time <- read.csv("ips-duration-openshift-7-31.dat")
+ips.time$howmany <- 1:nrow(ips.time)
+for ( i in 1:nrow(ips.time) ) {
+    ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
+}
+ips.time$i <- 1:nrow(ips.time)
+ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
+ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
+

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
una falta
changing some words
adding some sentences
adding Trap-function citation

@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'a-S\'anchez$^1$
+\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
 \thanks{Manuscript submitted for review on \today.}%
@@
 system based on the browser is simply a client-server architecture
-whose client is, or can be, embedded in the browser via
-JavaScript. All parts of the system are free and available with a free
+whose client is, or can be, embedded in the browser via JavaScript.
+
+% [Pedro] desde el abstract no se había nombrado "NodIO". 
+% Creo que en este párrafo debería introducirse, principalmente el nombre.
+% Añado una frase presentándolo:
+In this sense, in this paper we propose {\sf NodIO}, a PaaS based 
+volunteer computing system, whose architecture
+has been developed using JavaScript on both the client and server sides.
+All parts of the system are free and available with a free
 license from \url{https://github.com/JJ/splash-volunteer}.
@@
 \includegraphics[width=4in]{img/system.pdf}
-\caption{Description of the used system. Clients execute a JS EA in the browser, which, every 100 generations, sends the best individual and receives a random one from the server. } 
+\caption{Description of the proposed system. Clients execute a JS EA in the browser, which, every 100 generations, sends the best individual and receives a random one from the server. } 
 \label{fig:system}
@@
 The researcher only has to change a function to solve a different
-problem, in this case the classical Trap function has been %FERGU: cita a Trap function
+problem, in this case the classical Trap function \cite{Ackley1987} has been %FERGU: cita a Trap function
 used. JavaScript is a functional language and declaring a different

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
adding Trap-function citation

@@
 The researcher only has to change a function to solve a different
-problem, in this case the classical Trap function \cite{Ackley1987} has been %FERGU: cita a Trap function
+problem, in this case the classical Trap function \cite{Ackley1987} has been 
 used. JavaScript is a functional language and declaring a different

### 2015-08-04 data/time-vs-IPs-openshift.R
Updated data and representation

@@ -7,5 +7,6 @@ for ( i in 1:nrow(ips.time) ) {
 }
 ips.time$i <- 1:nrow(ips.time)
 ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
+ggplot(data=ips.time, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
 ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
 

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Answering @pacastillo questions

@@
 provisionally named {\sf NodIO}. NodIO is first an application that
-runs in two tiers, a server and a client that can be possibly run on a  % [Pedro] Both the server and the client in browsers? different browsers as I understand?
+runs in two tiers, a server and a client, the latter  can be possibly
+run on a  % [Pedro] Both the server and the client in browsers?
+          % different browsers as I understand?  No, server in
+          % server, client in browser, although there's nothing that
+          % can't be done in a browser. 
 browser. As such, it can be run in any computer that loads the web
@@
 
-When faced with the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
-hardware and continuous hardware failures of existing clusters and
-taking into account our experience of working with web applications,
-including JavaScript, from the beginning and other volunteer and
-unconventional distributed evolutionary computing systems, our group
-decided to create a new free framework that would allow anyone to
+Some time ago, our group faced  the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
+hardware and continuous hardware failures of existing
+clusters. Considering this, we leveraged our experience of working
+with web applications, including JavaScript, since their inception and other volunteer and
+unconventional distributed evolutionary computing systems, and we 
+decided to designd and relase a new free framework that would allow anyone to
 create a volunteer distributed EC experiment using cloud resources as
@@
 problem. Let us see how this system addresses the different challenges
-outlined in the State of the Art:\begin{itemize}
+outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
 \item {\em Scalability} is provided via the use of a lightweight and

### 2015-08-04 data/time-vs-IPs-openshift.R

@@ -9,4 +9,5 @@ ips.time$i <- 1:nrow(ips.time)
 ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
 ggplot(data=ips.time, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
 ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
+ggplot(data=ips.time, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.x=element_text("Ordered experiments"))
 

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Clearing unneeded comments
Some corrections added

@@
 
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
+\usepackage{array}
 
@@
 
-\title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-
-% [Pedro] no aparecía el primer autor. En cualquier caso, lo dejo comentado a continuación.
-
-%  \author{Juan-J.~Merelo, %Add your name here
-%  \thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
-%  e-mail: (jmerelo@ugr.es).}
-%  }% <-this % stops a space
-
-%  \author{Mario Garcia-Valdez, %Add your name here
-%    \thanks{M Garcia-Valdez is with the Dept. of Graduate Studies at Instituto Tecnologico de Tijuana
-%  e-mail: (mario@tectijuana.edu.mx).}
-%  }% <-this % stops a space
+\title{Volunteers in the Clouds: an Architecture for Low Cost and
+  Potentially Massive Distributed Evolutionary Computation}
 
@@
 run on a  % [Pedro] Both the server and the client in browsers?
-          % different browsers as I understand?  No, server in
+          % different browsers as I understand? [JJ]  No, server in
           % server, client in browser, although there's nothing that
@@
 
-Some time ago, our group faced  the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
+Some time ago, our group faced  the problem of diminishing funds for
+buying new     % [Pedro] This sentence is so so long. I would split it
+               % in two sentences. [JJ] Done.
 hardware and continuous hardware failures of existing
@@
 and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
-gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
+gives software developers certain advantages in the cloud paradigm. JavaScript is a first-classe citizen
 in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}.
@@
 
-In the work presented in this paper, JavaScript was used in the full 
-stack, which has the advantage of using the same code for writing
+In the work presented in this paper, JavaScript has been used throughout the full 
+stack, which has the advantage of using the same lines of code for
+writing an evolutionary algorithm for the
 desktop or command line and browser clients, both based in the EC
-library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages %FERGU: párrafo en pasado "was used" y presente "saves"
-expertise acquired in the library itself. %FERGU: tienes info de LoC usadas para la sección experimental y justificar esta frase?
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time
+when writing code and leverages %FERGU: párrafo en pasado "was used" y
+                                %presente "saves" 
+% done - JJ
+expertise acquired in the library itself. %FERGU: tienes info de LoC
+                                %usadas para la sección experimental y
+                                %justificar esta frase?
+% Es el mismo código - JJ
 %The proposed software has the advantage of full stack JavaScript
@@
 performance that can be obtained from these volunteer computing
-efforts. 
+experiments. 
 %Too many efforts?
@@
 % Digo efforts en la oración anterior :) - Mario
-There were some initial attempts  to dodge the issue by making
+% Visto :-) Gracias- JJ
+There were some initial attempts to simply dodge the issue by making
 the algorithm adaptive to the kind of resources allotted to it
@@
 % Añado una frase presentándolo:
-In this sense, in this paper we propose {\sf NodIO}, a PaaS based 
+In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
 volunteer computing system, whose architecture

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Finishing intro to experiments

@@
 \caption{Comparison of time to solution for the baseline system, using a {\tt
-    node.js} client, for different population sizes. They only compare
+    node.js} client, for different population sizes. They only compaonre
 runs in which solution has been found.}
@@
 successful. The experiments were repeated for population $p=1024$,
-with success rate upgraded to 100\%. Results for these two experiments
+with success rate upgraded to 100\% and average duration $3.46
+seconds$. Results for these two experiments 
 are shown in Figure \ref{fig:baseline}.
+
+These two experiments establish first a baseline result, which will
+depend on the population. The volunteer computing experiments that we
+will describe next do not and can not have the same conditions, but
+the baseline is that if they eventually take longer than a basic
+destktop, their interest will be purely academic. We will try and
+prove next that that target performance can be achieved by carrying
+out several experiments in different conditions.
+
+\subsection{Volunteer evolutionary computation: first experiments and results}
 

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Comentarios de la Introducción

@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
+\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
 \thanks{Manuscript submitted for review on \today.}%
 \thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
-\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
+\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
 \thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
@@
 
-There is, then, space for systems that do not have any of those
+There is, then, space for systems that do not have any of those %[Mario] these assumptions? 
 assumptions, such as the one presented in this paper, which will be
@@
 browser. As such, it can be run in any computer that loads the web
-page where the client is embedded, including possibly many over which
+page where the client is embedded, including possibly many over which%[Mario]Quitar el possibly?   
 the scientist has no control.
@@
 hardware and continuous hardware failures of existing
-clusters. Considering this, we leveraged our experience of working
+clusters. 
+%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
+%for buying new hardware. This was aggravated by the increasing maintance costs 
+%resulting from the continuous failures of existing clusters.
+Considering this, we leveraged our experience of working
 with web applications, including JavaScript, since their inception and other volunteer and
@@
 These systems must solve a series of issues before even starting to
-measure performance; these issues are related to security as well as
+measure their performance; these issues are related to security as well as
 safety, which introduces a {\em social} aspect to the design of a
@@
 % Añado una frase presentándolo:
+% [Mario] Nota: se menciona también en el segundo párrafo de la Introducción.
 In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
@@
 Thus, {\sf NodIO} has two parts:\begin{enumerate}
-\item A REST server, that is, a server that includes several remote
+\item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces   
   requests for storing and retrieving information (the CRUD cycle:
@@
   duties, but is basically a very lightweight and hign performance
-  data store \cite{jj:idc:lowcost}. The server has the capability to
+  data store \cite{jj:idc:lowcost}. 
+  %The information is of two kinds: Problem Based y ¿Cual otra? ¿Logging? No se entiende bien. ¿Podría ser problem domain?   
+  The server has the capability to
   run a single experiment, storing the chromosomes in a data structure
@@
 \caption{Comparison of time to solution for the baseline system, using a {\tt
-    node.js} client, for different population sizes. They only compaonre
+    node.js} client, for different population sizes. They only compare on
 runs in which solution has been found.}

### 2015-08-05 volunteer-clouds-ieee-ci-si.tex
Revisando comentarios
De @pacastillo, @mariosky y @fergunet. En general, si queréis reescribir una frase hacedlo y si acaso añadís un comentario diciendo por qué.

@@
 Volunteer computing is used routinely in science and industry, but it
-still has a long way to go if it is going to become mainstream in the
-same way the cloud technologies in general have. The lack of general purpose tools
+still has a long way to go if it is going to become a mainstream technology in the
+same way the cloud technologies in general have. The lack of general purpose and open source tools
 with analytics built in could be one of the reasons why this is so, as
@@
 browser. The architecture, along with some initial results on the expected performance and scaling
-behavior is presented. We also use the algorithm analytics to try and  %[Pedro] I don't understand what "the algorithm analytics" stand for???
-create a model of user behavior which, as is usual in this kind of
-computing models, follow a Weibull distribution. 
+behavior is presented. We also use the algorithm data to  %[Pedro] I don't understand what "the algorithm analytics" stand for???
+% data - JJ
+find out what kind of patterns the user behaviour follow and try to fit them to some statistical distribution.
 \end{abstract}
@@
 
-There is, then, space for systems that do not have any of those %[Mario] these assumptions? 
+There is, then, space for systems that do not have any of these %[Mario] these assumptions? [JJ] OK
 assumptions, such as the one presented in this paper, which will be
@@
 browser. As such, it can be run in any computer that loads the web
-page where the client is embedded, including possibly many over which%[Mario]Quitar el possibly?   
+page where the client is embedded, including many over which%[Mario]Quitar el possibly?   [JJ] Done
 the scientist has no control.
@@
                % in two sentences. [JJ] Done.
-hardware and continuous hardware failures of existing
-clusters. 
+hardware. This was aggravated by the increasing maintance costs and extended downtime resulting from the continuous failures of existing clusters.
 %[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
@@
 %resulting from the continuous failures of existing clusters.
-Considering this, we leveraged our experience of working
-with web applications, including JavaScript, since their inception and other volunteer and
+% [JJ] Done 
+Considering this, we leveraged our experience in the design of web
+applications, including JavaScript, since their inception and other
+volunteer and 
 unconventional distributed evolutionary computing systems, and we 
@@
 experiments. 
-%Too many efforts?
-% No, too little - JJ 
-% Digo efforts en la oración anterior :) - Mario
-% Visto :-) Gracias- JJ
 There were some initial attempts to simply dodge the issue by making
@@
 %ha sido la seguridad ya que estos presentan otra fuente de vulnerabilidad. Para complicar las cosas a veces (Flash, ActiveX) 
-%son tecnologías cerradas y propietarias. Por lo menos creo que esas fueron las razones que se dieron para eliminar a Flash del iPhone.
+%son tecnologías cerradas y propietarias. Por lo menos creo que esas
+%fueron las razones que se dieron para eliminar a Flash del iPhone.
+% [JJ] Escribe algo...
 
@@
 % Añado una frase presentándolo:
-% [Mario] Nota: se menciona también en el segundo párrafo de la Introducción.
+% [Mario] Nota: se menciona también en el segundo párrafo de la
+% Introducción.
+% OK [JJ]
 In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
@@
 license from \url{https://github.com/JJ/splash-volunteer}.
-Thus, {\sf NodIO} has two parts:\begin{enumerate}
+Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
 \item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces   
@@
   evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
-    GET}ing a random chromosome from it. It also peforms logging
+    GET}ing a random chromosome from it, and {\em information} related
+  to performance and state of the experiments. It also peforms logging
   duties, but is basically a very lightweight and hign performance
   data store \cite{jj:idc:lowcost}. 
-  %The information is of two kinds: Problem Based y ¿Cual otra? ¿Logging? No se entiende bien. ¿Podría ser problem domain?   
+  %The information is of two kinds: Problem Based y ¿Cual otra?
+  %¿Logging? No se entiende bien. ¿Podría ser problem domain?   
+% info. Añadido [JJ]
   The server has the capability to
@@
 \includegraphics[width=4in]{img/system.pdf}
-\caption{Description of the proposed system. Clients execute a JS EA in the browser, which, every 100 generations, sends the best individual and receives a random one from the server. } 
+\caption{Description of the proposed system. Clients execute a JS EA
+  in the browser, which, every 100 generations, sends the best
+  individual and receives a random one from the server. }  
 \label{fig:system}
@@
 used. JavaScript is a functional language and declaring a different
-function and handling at the creation of the algorithm object, called %FERGU: poner algo tras el handling: handling it?
+function and handing at the creation of the algorithm object, called
+%FERGU: poner algo tras el handling: handling it?
+%handing - JJ
 {\tt Classic} is the only change needed to work with a different

### 2015-08-05 volunteer-clouds-ieee-ci-si.tex
A bit more intro

@@
 
+Initial experiments were set up using the OpenShift
+Platform-as-a-Service, which provides a free tier within which this
+experiment could be performed. That way, and as required, the hardware
+and cloud cost for this experiment were zero. Experiments were
+announced through a post in Twitter and other social networks, and
+results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
+purpose of this paper, we repeated the announcement several times
+through the months of April and then by the beginning of August.
+
 %---------------------------------------------------------------

### 2015-08-05 data/plot-zipf-openshift.R
New data and plots added

@@ -1,0 +1,9 @@
+library(ggplot2)
+puts.4.4 <- read.csv("ips-puts-openshift-4-4.dat")
+puts.4.24 <- read.csv("ips-puts-openshift-4-24.dat")
+puts.7.31 <- read.csv("ips-puts-openshift-7-31.dat")
+
+dt.puts.4.4 <- data.frame( x=1:nrow(puts.4.4),puts=sort(puts.4.4$puts,decreasing=T))
+dt.puts.4.24 <- data.frame( x=1:nrow(puts.4.24),puts=sort(puts.4.24$puts,decreasing=T))
+dt.puts.7.31 <- data.frame( x=1:nrow(puts.7.31),puts=sort(puts.7.31$puts,decreasing=T))
+ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4.4,aes(x=x/length(dt.puts.4.4$x),y=puts,color='4/4'),stat='Identity')+geom_point(data=dt.puts.4.24,aes(x=x/length(dt.puts.4.24$x),y=puts,color='4/24'),stat='Identity')+geom_point(data=dt.puts.7.31,aes(x=x/length(dt.puts.7.31$x),y=puts,color='7/31'),stat='Identity')

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
New data
Some fixes to earlier writing
There's still some pending work by @mariosky and @fergunet and reviewing answers to their comments.

@@
 \documentclass[journal,onecolumn]{IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
 
@@
 \usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
 
@@
 \begin{IEEEkeywords}
-Evolutionary algorithm, distributed computing, internet computing,
+Evolutionary algorithms, evolutionary computation, distributed computing, internet computing,
 social networks, volunteer computing, metacomputing. 
@@
 experiment designer, unlike the other systems mentioned
-above. Proposals contemporary to Jet, such as SuperWeb
+above. Other proposals contemporary to Jet, such as SuperWeb
 \cite{alexandrov1997superweb} are roughly a translation of classical
@@
 Sarmenta once again used Java and not JavaScript, making this effort
-less universal.
+less widespread.
 
@@
 experiments. 
-There were some initial attempts to simply dodge the issue by making
+There were some initial attempts to avoid the differences in performance
+that could be obtained from volunteers  by making
 the algorithm adaptive to the kind of resources allotted to it

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Adds tables with initial results

@@
 purpose of this paper, we repeated the announcement several times
-through the months of April and then by the beginning of August.
+through the months of April and then by the beginning of August. All
+in all, we have the set of runs with the characteristics shown in
+Table \ref{tab:runs}.
+%
+\begin{table*}
+\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
+\begin{center}
+\begin{tabular}{l|r}
+\hline
+Date & Number of experiments & Different IPs \\
+\hline
+April 4th 4/4& & 57 & 191\\
+April 24th 4/24 &  231 & 559 \\
+July 31th 7/31 & 97 & 179\\
+\hline
+\end{tabular}
+\end{table}
+%
+
+\begin{table*}
+\caption{Summary of time per run, number of IPS and number of {\tt
+    PUT}s per IP in the initial runs. \label{tab:runs}}
+\begin{center}
+\begin{tabular}{l|rrr}
+\hline
+Date & Median IPs & Median time (s) & Median #{\tt PUT}s  \\
+\hline
+4/4& & 5 & 2040 & 18\\
+April 24th 4/24 &  5 & 732 & 11\\
+July 31th 7/31 & 5 & 260 & 23 \\
+\hline
+\end{tabular}
+\end{table}
+%
+A summary of the characteristics of each run is also shown in Table
+\ref{tab:summary}.
 

### 2015-08-06 data/time-vs-IPs-openshift.R

@@ -1,13 +1,27 @@
 library(ggplot2)
 
-ips.time <- read.csv("ips-duration-openshift-7-31.dat")
-ips.time$howmany <- 1:nrow(ips.time)
-for ( i in 1:nrow(ips.time) ) {
-    ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
+                        
+time.vs.IPs <- function( data, preffix ) {
+    data$howmany <- 1:nrow(data)
+    for ( i in 1:nrow(data) ) {
+        data[i,]$howmany <- length(data[data$IPs==data[i,]$IPs,]$milliseconds)
+    }
+    data$i <- 1:nrow(data)
+    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
+    ggsave(paste0("time-vs-ips-OS-",preffix,".png"))
+    ggplot(data=data, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
+    ggsave(paste0("duration-vs-time-OS-",preffix,".png"))
+    ggplot(data=data, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
+    ggsave(paste0("duration-vs-time-OS-bar-",preffix,".png"))
+    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.x=element_text("Ordered experiments"))
+    ggsave(paste0("time-vs-rank-OS-",preffix,".png"))
 }
-ips.time$i <- 1:nrow(ips.time)
-ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
-ggplot(data=ips.time, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
[…]

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Updating tables

@@
 successful. The experiments were repeated for population $p=1024$,
-with success rate upgraded to 100\% and average duration $3.46
-seconds$. Results for these two experiments 
+with success rate upgraded to 100\% and average duration $3.46$
+seconds. Results for these two experiments 
 are shown in Figure \ref{fig:baseline}.
@@
 \begin{center}
-\begin{tabular}{l|r}
+\begin{tabular}{l|rr}
 \hline
-Date & Number of experiments & Different IPs \\
+Date & \# experiments & Different IPs \\
 \hline 
-April 4th 4/4& & 57 & 191\\
+April 4th 4/4 & 57 & 191 \\
 April 24th 4/24 &  231 & 559 \\
-July 31th 7/31 & 97 & 179\\
+July 31th 7/31 & 97 & 179 \\
 \hline 
 \end{tabular}
-\end{table}
+\end{center}
+\end{table*}
 %
-
 \begin{table*}
 \caption{Summary of time per run, number of IPS and number of {\tt
-    PUT}s per IP in the initial runs. \label{tab:runs}}
+    PUT}s per IP in the initial runs. \label{tab:summary}}
 \begin{center}
-\begin{tabular}{l|rrr}
+\begin{tabular}{l|ccccc}
 \hline
-Date & Median IPs & Median time (s) & Median #{\tt PUT}s  \\
+Date & Median \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
 \hline 
-4/4& & 5 & 2040 & 18\\
-April 24th 4/24 &  5 & 732 & 11\\
-July 31th 7/31 & 5 & 260 & 23 \\
+4/4 & 5 & 2040 & 18 & 14.29\% & 5.36\% \\
+4/24 &  5 & 732 & 11 & 47.39\% & 3.91\% \\
+7/31 & 5 & 260 & 23 & 27.08\% & 1.04\%  \\
 \hline 
 \end{tabular}
-\end{table}
+\end{center}
+\end{table*}
 %

### 2015-08-06 data/time-vs-IPs-openshift.R

@@ -23,5 +23,7 @@ for ( i in suffix ) {
     ips.time <- read.csv(paste0("ips-duration-openshift-",i,".dat"))
     time.vs.IPs(ips.time, i)
     print(paste(i, summary(ips.time)))
+    print(paste("< 69s", length(ips.time[ips.time$milliseconds < 69000,]$milliseconds)/length(ips.time$milliseconds)))
+    print(paste("< 3.46s", length(ips.time[ips.time$milliseconds < 3460,]$milliseconds)/length(ips.time$milliseconds)))
 }
  

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Fixing typos, added new reference, changed figure,  and comments for @JJ and @mariosky

--- 2015-08-06 volunteer-clouds-ieee-ci-si.tex
@@
 % data - JJ
-find out what kind of patterns the user behaviour follow and try to fit them to some statistical distribution.
+find out what kind of patterns the user behavior follow and try to fit them to some statistical distribution.
 \end{abstract}
@@
 distributed evolutionary computation software frameworks available on
-the market, usually as open source systems. Most of these systems have
+the market, usually as open source systems \cite{Parejo12Survey}. Most of these systems have %FERGU: añadida referencia
 two features: they are {\em desktop} applications, that is, they must
@@
                % in two sentences. [JJ] Done.
-hardware. This was aggravated by the increasing maintance costs and extended downtime resulting from the continuous failures of existing clusters.
+hardware. This was aggravated by the increasing maintenance costs and extended downtime resulting from the continuous failures of existing clusters.
 %[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
@@
 unconventional distributed evolutionary computing systems, and we 
-decided to designd and relase a new free framework that would allow anyone to
+decided to design and release a new free framework that would allow anyone to
 create a volunteer distributed EC experiment using cloud resources as
@@
 
-These systems must solve a series of issues before even starting to
+This kind of systems must solve a series of issues before even starting to %FERGU: cambiado el "these systems" porque parecía que hacía referencia a "los anteriores" 
 measure their performance; these issues are related to security as well as
@@
 state of the art (Section \ref{sec:soa}) in web-based distributed 
-computation systems along with attempts to predict and model its behavior. 
+computation systems along with attempts to predict and model its behavior. %FERGU: El IEEE es americano
 Then, a detailed description of the proposed system is presented 
@@
 gives software developers certain advantages in the cloud paradigm. JavaScript is a first-classe citizen
-in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}.
-Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias
+in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}. Also many web services 
 have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
@@
 library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time
-when writing code and leverages %FERGU: párrafo en pasado "was used" y
-                                %presente "saves" 
-% done - JJ
+when writing code and leverages %FERGU: qué diferencias hay entre NodIO y NodEO??
 expertise acquired in the library itself. %FERGU: tienes info de LoC
@@
 %development..
-%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
+
 Javascript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
@@
     GET}ing a random chromosome from it, and {\em information} related
-  to performance and state of the experiments. It also peforms logging
-  duties, but is basically a very lightweight and hign performance
+  to performance and state of the experiments. It also performs logging
+  duties, but is basically a very lightweight and high performance
   data store \cite{jj:idc:lowcost}. 
@@
 
-Figure \ref{fig:system} describes the general system architecture.
+Figure \ref{fig:system} describes the general system architecture and
+algorithm behavior. Different web technologies, such as JQuery or chart.js have
+been used to build the elements of the system. 
+%FERGU: Mario, explica en el texto de abajo qué es winston en la figura también. 
+%Vamos, que todo lo que salga en la figura tiene que ir en el texto en alguna parte :)
 %
@@
 \centering
-\includegraphics[width=4in]{img/system.pdf}
+\includegraphics[width=4in]{img/Algorithm.pdf}
 \caption{Description of the proposed system. Clients execute a JS EA
@@
 function and handing at the creation of the algorithm object, called
-%FERGU: poner algo tras el handling: handling it?
-%handing - JJ
 {\tt Classic} is the only change needed to work with a different
@@
   performance, operating system or even browser requirement: anyone
-  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
+  visiting the page, even from mobile devices, can load the algorithm. 
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
 First we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using {\sf NodEO},
+evolutionary algorithm in a desktop client written using NodEO, 
+%FERGU: es NodEO o NodIO? Quito el \sf, por cierto, que queda feo
+%Por qué usamos NodEO aquí y no NodIO, vamos a compararlos? 
+%Qué diferencias hay entre los dos? (ponerlo en el SOA)
 the node.js evolutionary algorithm library. This experiment tries to 
+%FERGU: quitar el "the node.js evolutionary algorithm library", si ya lo hemos dicho
+%o explicarlo mejor: "the basic js library we have used as base to built NodIO", por ejemplo
 find the solution to the 40-trap function with parameters $l=4, a=1,
@@
 experiments to be the correct one for finding the solution 95\% of the
+%Fergu: cita a esos experimentos?
 time. The algorithm was run until the solution (a string with all
@@
 the baseline is that if they eventually take longer than a basic
-destktop, their interest will be purely academic. We will try and
+desktop, their interest will be purely academic. We will try and
 prove next that that target performance can be achieved by carrying

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex

@@
 purpose of this paper, we repeated the announcement several times
-through the months of April and then by the beginning of August.
+through the months of April and then by the beginning of August. All
+in all, we have the set of runs with the characteristics shown in
+Table \ref{tab:runs}.
+%
+\begin{table*}
+\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
+\begin{center}
+\begin{tabular}{l|rr}
+\hline
+Date & \# experiments & Different IPs \\
+\hline 
+April 4th 4/4 & 57 & 191 \\
+April 24th 4/24 &  231 & 559 \\
+July 31th 7/31 & 97 & 179 \\
+\hline 
+\end{tabular}
+\end{center}
+\end{table*}
+%
+\begin{table*}
+\caption{Summary of time per run, number of IPS and number of {\tt
+    PUT}s per IP in the initial runs. \label{tab:summary}}
+\begin{center}
+\begin{tabular}{l|ccccc}
+\hline
+Date & Median \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
+\hline 
+4/4 & 5 & 2040 & 18 & 14.29\% & 5.36\% \\
+4/24 &  5 & 732 & 11 & 47.39\% & 3.91\% \\
+7/31 & 5 & 260 & 23 & 27.08\% & 1.04\%  \\
+\hline 
+\end{tabular}
+\end{center}
+\end{table*}
+%
+A summary of the characteristics of each run is also shown in Table
+\ref{tab:summary}.
 

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Response to comments
Trabajaré en explicar la figura y la implementación con web workers.
Using both figures from @mariosky and @fergunet
Helping to describe the table

@@
 hardware. This was aggravated by the increasing maintenance costs and extended downtime resulting from the continuous failures of existing clusters.
-%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
-%for buying new hardware. This was aggravated by the increasing maintance costs 
-%resulting from the continuous failures of existing clusters.
-% [JJ] Done 
 Considering this, we leveraged our experience in the design of web
@@
 
-%Algo novedoso es el uso de Javascript no solo en el cliente si no también en el Servidor.
-%Como el tema es el Software creo que esto se tiene que resaltar. En los sistemas anteriores 
-%se tenía lo de Java en el cliente, pero en el server estaba lo grueso del trabajo:
-% Administrar el server - Ahora más fácil con cloud
-% Programar el server side - Ahora se puede usar JavaScript también para eso.
-
 On the other hand, measuring performance of the resulting metacomputer
@@
 %Vamos, que todo lo que salga en la figura tiene que ir en el texto en alguna parte :)
-%
+% [Mario] Ok voy a explicar la figura. Quería saber primero su opinión tal vez se les hacía mucho detalle.
+% Creo que esta figura va mejor en la sección donde hablamos de
+% webworkers. La anterior era más adecuada como visión general. [JJ]
 \begin{figure}[!t]
 \centering
-\includegraphics[width=4in]{img/Algorithm.pdf}
+\includegraphics[width=4in]{img/system.pdf}
 \caption{Description of the proposed system. Clients execute a JS EA
@@
 %Qué diferencias hay entre los dos? (ponerlo en el SOA)
+%[Mario] Creo que es NodEO y NodIO fué mi error.
 the node.js evolutionary algorithm library. This experiment tries to 
@@
 in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:runs}.
+Table \ref{tab:runs}. In general, every experiment took several
+days. No particular care was taken about the time of the announcement
+or the particular wording. Every {\em experiment} consisted in running
+until the solution of the 40-trap problem was found. When the correct
+solution was sent to the server, the counter was updated and the pool
+of solutions reset to the void set. No special care was taken to wait
+until all clients had finished, thus it might happen that, in fact,
+the islands running in the browser {\em spill} from one experiment to
+the next. However, previous experiments have proved that the influence
+of these islands in the next experiment is, in fact negligible.
 %
-\begin{table*}
+\begin{table*}d
 \caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
@@
 \caption{Summary of time per run, number of IPS and number of {\tt
-    PUT}s per IP in the initial runs. \label{tab:summary}}
+    PUT}s per IP in the initial runs. \label{tab:summary:os}}
 \begin{center}
@@
 %
-A summary of the characteristics of each run is also shown in Table
-\ref{tab:summary}.
+The table shows that every run included more than 50 experiments. The
+number of different IPs intervening in them varied from more than one
+hundred to more than five hundred in the second experiment.
+
+A summary of the results of each run is also shown in Table
+\ref{tab:summary:os}. This table shows the median number of IPs
+intervening in each experiment,  median of the time needed
+to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
+then two columns showing the percentage of experiments that took less
+than the two baseline experiments shown in the introduction to this
+section. The first striking result is that in all cases, 50\% of the
+experiments involved 5 or less IPs. This is consistent with results
+found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
+the usual number of IPs that participated in an experiment.
+
+\subsection{{\sf NodIO-W$^2$}, a WebWorker-based architecture}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=4in]{img/Algorithm.pdf}
+\caption{Description of the W$^2$ version of {\sf NodIO}. In this
+  version, clientts use WebWorkers to run the evolutionary algorithm,
+  with two of them per browser.}  
+\label{fig:system:w2}
+\end{figure}
 

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Completing table
Examining data for the experiments

@@
 \begin{center}
-\begin{tabular}{l|ccccc}
+\begin{tabular}{l|cccccc}
 \hline
-Date & Median \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
+Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
 \hline 
-4/4 & 5 & 2040 & 18 & 14.29\% & 5.36\% \\
-4/24 &  5 & 732 & 11 & 47.39\% & 3.91\% \\
-7/31 & 5 & 260 & 23 & 27.08\% & 1.04\%  \\
+4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% \\
+4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% \\
+7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  \\
 \hline 
@@
 found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
-the usual number of IPs that participated in an experiment.
+the usual number of IPs that participated in an experiment. The
+maximum number of different IPs for each experiment is also in the
+same ballpark and of the order of 10, which is also consistent with
+prior work. The median time has a big range of variation, but 50\% of
+the time takes less than several minutes, from around 4 minutes in the
+best case to roughly 2/3 of an hour in the worst case. That figure is
+not competitive, {\em a priori}, with the baseline experiment, that is
+why it is interesting to look at the last two columns, that measure
+the percentage of experiments in which the solution was found in less
+time than the baseline. It is similar in all three cases, with around
+20\% on average for the first baseline experiment, roughly 3\% in the
+second case. The comparison is not totally fair: this experiment takes
+place in a browser and in fact most of the time is devoted to
+presenting the charts; the server is not the same since it is run in a
+PaaS instead of a high-performance desktop computer, and it is also a
+remote server, not a local server as in the baseline experiment. But
+fair is fair, and we were looking for creating a framework that
+allowed to run evolutionary computation experiments {\em faster} than
+in our very own computer. And the results is that although it will
+happen, one out of roughly 5 or roughly 20, depending on what baseline
+experiment you choose, it will not happen always or even a majority of
+the time.
+
+We will have to analyze experiment data a bit further to find out why
+this happens and also if there is some pattern in the three sets of
+experiments. An interesting question to ask, for instance, is if
+adding more computers makes the experiment take less.
 

### 2015-08-07 data/time-vs-IPs-openshift.R
Adds correlation to computation

@@ -25,5 +25,10 @@ for ( i in suffix ) {
     print(paste(i, summary(ips.time)))
     print(paste("< 69s", length(ips.time[ips.time$milliseconds < 69000,]$milliseconds)/length(ips.time$milliseconds)))
     print(paste("< 3.46s", length(ips.time[ips.time$milliseconds < 3460,]$milliseconds)/length(ips.time$milliseconds)))
+    ips.time.minus1 <- ips.time[c(1:length(ips.time$IPs)-1),]
+    ips.time.plus1 <- ips.time[c(2:length(ips.time$IPs)),]
+    print(length(ips.time.minus1$IPs))
+    print(length(ips.time.plus1$IPs))
+    print(cor(ips.time.minus1, ips.time.plus1))
 }
  

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex

@@
 
-\subsection{{\sf NodIO-W$^2$}, a WebWorker-based architecture}
+\subsection{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
+
+Javascript has commonly been used to develop client-side scripts that handle the interaction 
+with the user, communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
+Long-running scripts like those proposed in this work can interfere with the responsiveness of 
+the page. The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
+to execute scripts in parallel with user interaction scripts of the main page. 
+Worker scripts are spawned in the background allowing a thread-like operation with 
+message-passing as the coordination mechanism.
 

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adding correlation

@@
                % in two sentences. [JJ] Done.
-hardware. This was aggravated by the increasing maintenance costs and extended downtime resulting from the continuous failures of existing clusters.
+hardware. This was aggravated by the increasing maintenance costs and
+extended downtime resulting from the continuous failures of existing
+clusters.  
 Considering this, we leveraged our experience in the design of web
@@
 \begin{center}
-\begin{tabular}{l|cccccc}
+\begin{tabular}{l|ccccccc}
 \hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
+Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
+  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
 \hline 
-4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% \\
-4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% \\
-7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  \\
+4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
+4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
+7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
 \hline 
@@
 experiment you choose, it will not happen always or even a majority of
-the time.
+the time. One of the problems is shown in the last column, which
+indicates the statistical correlation between the number of IPs
+participating in one experiment and the next. In all cases,
+correlation is so low as to affirm that there is almost no
+relationship between them, that is, computing nodes are not {\em
+  staying} after the solution has been found; this is a feature of
+this implementation: the user has to actively reload the page to make
+the evolutionary algorithm start again. However, it puts a probably
+unnecesary cap on the time, or number of experiments, that an user
+will voluntarily perform.
 

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adds duration graphs

@@
 
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-4-4.png}
+\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-4-24.png}
+\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-7-31.png}
+\caption{Duration of experiments vs. number of different IPs (nodes)
+  participating in it, with averages and standard deviation shown as
+  red dots. From left to right, experiments 4/4, 4/24 and 7/31.}
+\label{fig:duration}
+\end{figure}
+%
 We will have to analyze experiment data a bit further to find out why

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adding zipf's text

@@
 
-\begin{figure}[!t]
+\begin{figure}[!htb]
 \centering
@@
   participating in it, with averages and standard deviation shown as
-  red dots. From left to right, experiments 4/4, 4/24 and 7/31.}
+  red dots. From left to right, experiments 4/4, 4/24 and 7/31. Shades
+of blue indicate how many experiments included that many unique IPs.}
 \label{fig:duration}
+\end{figure}
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-4-4.png}
+\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-4-24.png}
+\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-7-31.png}
+\caption{Duration of experiments vs rank, with $y$ axis in a
+  logarithmic scale. Dot color is related to the number of IPs
+  participating in the experiment. From left to right, experiments 4/4, 4/24 and 7/31.}
+\label{fig:zipf:os}
 \end{figure}
@@
 experiments. An interesting question to ask, for instance, is if
-adding more computers makes the experiment take less.
+adding more computers makes the experiment take less. In fact, as
+shown in Figure \ref{fig:duration}, {\em adding} more computers does
+not seem to contribute to decreasing the time needed to finish the
+experiment. However, the cause-effect relationship is not clear at
+all. It might be the opposite: since experiments take longer to finish
+and might in fact be abandoned with no one contributing for some time,
+that increases the probability of someone new joining them. In fact,
+wit experiments taking a few seconds and due to the way the
+experiments are announced, it is quite difficult that several
+volunteers join in in such a short period of time, even more so taking
+into account that volunteers are not {\em carried over} from previous
+experiments. This implies that it would be convenient to use a problem
+with a bigger size to check this hypothesis; however, at this point we
+have not found this convenient since there are several other issues
+that will be solved.
+
+It is interesting also to check the distribution of the experiment
+duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
+a Zipf's law, with similar distribution along all three runs. The 4/24
+run is the most complete and shows a S-shape, which implies an
+accumulation of experiments taking similar time and around 100
+seconds. The most interesting part is the {\em tail}, which shows how
+many experiments took a desirable amount of time, and of the order of
+10 seconds. As it can be seen, it drops sharply implying there are
+just a few of them, and with diminishing probability as time
+decreases.
 

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
More work on workers

@@
 
-Javascript has commonly been used to develop client-side scripts that handle the interaction 
-with the user, communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
+Javascript has commonly been used to develop client-side scripts that handle the user interface,
+communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
 Long-running scripts like those proposed in this work can interfere with the responsiveness of 
-the page. The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
-to execute scripts in parallel with user interaction scripts of the main page. 
-Worker scripts are spawned in the background allowing a thread-like operation with 
-message-passing as the coordination mechanism.
+the web application, because everything is handled by a single thread. 
+The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
+to execute long-running scripts in parallel with the scripts of the main page. 
+Worker scripts are spawned in the background allowing a thread-like operation using 
+message-passing as the communication mechanism between the worker's execution environment and the main thread. 
+According to the Web Workers W3C Candidate Recomendation \cite{hickson2012web} 
+workers are expected to be long-lived, they have a high start-up performance cost, 
+and a high per-instance memory cost. The Web Worker API is implemented in current
+versions of both desktop and mobile web browsers. For volunteer computing using the Web Worker API   
+brings several advantages over a single threaded implementation:
+\begin{itemize}
+\item If the browser uses a tabbed document interface, the worker script 
+keeps running in the background even it the user brings forward another tabbed window.
+\item Several instances of the NodEO script can be executed in parallel in a single web page. 
+\item Many immplemantations of the Web Worker API can use multiple CPUs for the execution
+of worker scripts, in particular the execution in multiple CPUs was tested in desktop verions of
+Chrome, FireFox and Safari.
+
+\end{itemize}
 

### 2015-08-07 data/plot-zipf-openshift.R
Trying to fit stuff

@@ -1,9 +1,25 @@
 library(ggplot2)
-puts.4.4 <- read.csv("ips-puts-openshift-4-4.dat")
-puts.4.24 <- read.csv("ips-puts-openshift-4-24.dat")
-puts.7.31 <- read.csv("ips-puts-openshift-7-31.dat")
+library(MASS)
+library(fitdistrplus)
 
-dt.puts.4.4 <- data.frame( x=1:nrow(puts.4.4),puts=sort(puts.4.4$puts,decreasing=T))
-dt.puts.4.24 <- data.frame( x=1:nrow(puts.4.24),puts=sort(puts.4.24$puts,decreasing=T))
-dt.puts.7.31 <- data.frame( x=1:nrow(puts.7.31),puts=sort(puts.7.31$puts,decreasing=T))
+process.puts <- function(suffix ) {
+    these.puts <- read.csv(paste0("ips-puts-openshift-",suffix,".dat"))
+    this.df <- data.frame(x=1:nrow(these.puts),
+                          puts=sort(these.puts$puts,decreasing=T))
+    this.hist <- hist(this.df$puts)
+    this.fit <- fitdistr(this.hist$density,"weibull")
+    this.weibull <- rweibull(length(this.hist$density),this.fit$estimate["shape"],this.fit$estimate["scale"])
+    this.w.df <- data.frame(x=1:length(this.weibull),fit=this.weibull)
+    ggplot(this.df,aes(x=x,y=puts))+geom_point(this.w.df,aes(x=x,y=fit))
+    print(this.fit)
+
+    return(this.df)
+}
+    
+
+dt.puts.4.4 <- process.puts("4-4")
+dt.puts.4.24 <- process.puts("4-24")
+dt.puts.7.31 <- process.puts("7-31")
+
+
 ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4.4,aes(x=x/length(dt.puts.4.4$x),y=puts,color='4/4'),stat='Identity')+geom_point(data=dt.puts.4.24,aes(x=x/length(dt.puts.4.24$x),y=puts,color='4/24'),stat='Identity')+geom_point(data=dt.puts.7.31,aes(x=x/length(dt.puts.7.31$x),y=puts,color='7/31'),stat='Identity')

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Homoneinizing JavaScript

@@
 public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
-and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
-gives software developers certain advantages in the cloud paradigm. JavaScript is a first-classe citizen
-in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}. Also many web services 
-have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
-developers can use the same software tools for both client and server side development. 
-Besides the advantages to deal with the same language in the full stack as previously mentioned, 
+and DropBox~\cite{mericloud}. The use of JavaScript to implement
+server side components, %FERGU: he puesto mejor el de revista en lugar
+                        %del del CEC 
+gives software developers certain advantages in the cloud
+paradigm. JavaScript is a first-classe citizen 
+in most Platforms as a Service (PaaS) as many support the Node.js
+framework \cite{wood13:nodejs:paas}. Also many web services  
+have native JavaScript libraries and use JSON for data
+interchange. Another advantage is that  
+developers can use the same software tools for both client and server
+side development.  
+Besides the advantages to deal with the same language in the full
+stack as previously mentioned,  
 this also facilitates the interoperability of the exchanged data types.
@@
 
-Javascript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
+JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
 According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
@@
 %Si no se extiende mucho se podría también hablar de la conexión con Cloud Computing,
-%se discuten las ventajas de utilizar Javascript para acceder a recursos donados por voluntarios
+%se discuten las ventajas de utilizar JavaScript para acceder a recursos donados por voluntarios
 %pero falta hablar de como se explotaría el recurso Cloud. Lo digo por el titulo de Volunteers in the Clouds. Tal vez
@@
 
-Javascript has commonly been used to develop client-side scripts that handle the user interface,
+JavaScript has commonly been used to develop client-side scripts that handle the user interface,
 communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.

### 2015-08-07 data/plot-zipf-openshift.R
Fixed fitting of #PUTs

@@ -1,18 +1,19 @@
 library(ggplot2)
-library(MASS)
-library(fitdistrplus)
+library(evd)
 
 process.puts <- function(suffix ) {
     these.puts <- read.csv(paste0("ips-puts-openshift-",suffix,".dat"))
     this.df <- data.frame(x=1:nrow(these.puts),
                           puts=sort(these.puts$puts,decreasing=T))
-    this.hist <- hist(this.df$puts)
-    this.fit <- fitdistr(this.hist$density,"weibull")
-    this.weibull <- rweibull(length(this.hist$density),this.fit$estimate["shape"],this.fit$estimate["scale"])
-    this.w.df <- data.frame(x=1:length(this.weibull),fit=this.weibull)
-    ggplot(this.df,aes(x=x,y=puts))+geom_point(this.w.df,aes(x=x,y=fit))
+    this.fit <- fgev(this.df$puts)
     print(this.fit)
-
+    this.fit.plot <- data.frame(x=1:length(this.df$puts),
+                                y=sort(rgev(length(this.df$puts),
+                                    this.fit$estimate["loc"],
+                                    this.fit$estimate["scale"],
+                                    this.fit$estimate["shape"]),decreasing=T))
+    ggplot()+geom_point(data=this.df,aes(x=x,y=puts))+geom_point(data=this.fit.plot,aes(x=x,y=y,color='red'))+scale_y_log10()
+    ggsave(paste0("puts-openshift-",suffix,".png"))
     return(this.df)
 }
     

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adds acknowledgements

@@
 
-This work has been supported in part by TIN2011-28627-C04-02 and
+This work has been supported in part by TIN2011-28627-C04-02 and. 
 TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
@@
 anonymous reviewers of previous versions of this paper who have really helped us to improve
-this paper (and our work) with their suggestions.
+this paper (and our work) with their suggestions. We are also grateful
+to Anna Saéz de Tejada for her help with the data processing scripts.
 

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adding graphs

@@
 
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.32\linewidth]{img/puts-openshift-4-4.png}
+\includegraphics[width=0.32\linewidth]{img/puts-openshift-4-24.png}
+\includegraphics[width=0.32\linewidth]{img/puts-openshift-7-31.png}
+\caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
+  Extreme Value distribution (in red). From left to right, experiments 4/4, 4/24 and 7/31.}
+\label{fig:puts:os}
+\end{figure}
+%
+
 It is interesting also to check the distribution of the experiment
@@
 just a few of them, and with diminishing probability as time
-decreases.
+decreases. This exponential distribution also appears in the
+distribution of HTTP {\tt PUT}s, equivalent to the number of
+generations divided by 100, contributed by every user, which is shown
+in Figure \ref{fig:put:os}.
 

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adds table for PUTs
Finished second subsection
Please all @pacastillo @mariosky @fergunet take a look. Analysis for the next section will follow the same pattern.
Help with making que graphs nicer would also help. All R scripts are in the `data` dir.

@@
 generations divided by 100, contributed by every user, which is shown
-in Figure \ref{fig:put:os}.
+in Figure \ref{fig:put:os}. These results show a Zipf-like behavior,
+so we have fitted it to the Generalized Extreme Value distribution,
+with the resulting parameters shown in Table \ref{tab:put:os}.
+%
+\begin{table*}
+\caption{Summary of fit to Generalized Extreme Value distribution of
+  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
+\begin{center}
+\begin{tabular}{l|ccc}
+\hline
+Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+\hline 
+4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
+0.1377 \\  
+4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\    
+7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\   
+\hline 
+\end{tabular}
+\end{center}
+\end{table*}
+%
+This distribution was originally proposed to fit extreme values
+\cite{resnick2013extreme} and contains, as a special case, the inverse
+Weibull distribution which was fitted to volunteer computing
+frameworks such as SETI@home \cite{javadi2009mining}. This
+distribution is governed by three parameters, the usual location $\mu$
+which is related to where it has its {\em center} and  scale $\sigma$,
+related to the size, but also a third shape $\xi$ parameter that is
+related to its skewness, that is, how skewed it is around the central
+location. Positive parameters indicate that the distribution {\em
+  leans} towards the origin, negative towards the other extreme
+value. In this case, Table \cite{tab:puts:os} shows $\xi$ values
+greater than one and between one and 1.4, which indicates that the
+three experiment share this origin-leaning pattern, with many users
+donating a few cycles and just a few donating extreme values. Random
+distributions with these parameters have been plotted in red in
+\ref{fig:put:os}, indicating that the fit is good enough. The
+predictive value of these fits, however, is limited, over all taking
+into account the low correlation between successive events shown
+above.
+
+At any rate, this also shows that a convenient way of increasing the
+computing power would be increasing this minimum amount per user. This
+issue led us to create the next version of the framework, which is
+presented next.
 

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
More about the algorithm
This is a brain dump just to capture the ideas. A heavy edit is
expected.

@@
 
-JavaScript has commonly been used to develop client-side scripts that handle the user interface,
-communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
-Long-running scripts like those proposed in this work can interfere with the responsiveness of 
-the web application, because everything is handled by a single thread. 
-The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
-to execute long-running scripts in parallel with the scripts of the main page. 
-Worker scripts are spawned in the background allowing a thread-like operation using 
-message-passing as the communication mechanism between the worker's execution environment and the main thread. 
-According to the Web Workers W3C Candidate Recomendation \cite{hickson2012web} 
-workers are expected to be long-lived, they have a high start-up performance cost, 
-and a high per-instance memory cost. The Web Worker API is implemented in current
-versions of both desktop and mobile web browsers. For volunteer computing using the Web Worker API   
-brings several advantages over a single threaded implementation:
+JavaScript has commonly been used to develop client-side scripts
+that handle the user interface, communicate asynchronously with servers and 
+update the content that is displayed \cite{flanagan2006javascript}.
+Long-running scripts like those proposed in this work can interfere with the
+responsiveness of the web application, because everything is handled by a 
+single thread. The Web Workers specification \cite{hickson2012web} defines an 
+API that allows Web developers to execute long-running scripts in parallel
+with the scripts of the main page. Worker scripts are spawned in the 
+background allowing a thread-like operation using message-passing as the 
+communication mechanism between the worker's execution environment and the 
+main thread. According to the Web Workers W3C Candidate Recomendation 
+\cite{hickson2012web} workers are expected to be long-lived, they have a high 
+start-up performance cost, and a high per-instance memory cost. The Web Worker
+API is implemented in current versions of both desktop and mobile web browsers.
+For volunteer computing using the Web Worker API brings several advantages 
+over a single threaded implementation:
+
 \begin{itemize}
 \item If the browser uses a tabbed document interface, the worker script 
-keeps running in the background even it the user brings forward another tabbed window.
-\item Several instances of the NodEO script can be executed in parallel in a single web page. 
-\item Many immplemantations of the Web Worker API can use multiple CPUs for the execution
-of worker scripts, in particular the execution in multiple CPUs was tested in desktop verions of
-Chrome, FireFox and Safari.
+keeps running in the background even it the user brings forward another tabbed
+window.
+\item Several instances of NodEO can be executed in parallel in a single web
+page. 
+\item Many immplemantations of the Web Worker API can use multiple CPUs for
+the execution of worker scripts, in particular the execution in multiple CPUs
+was tested in desktop verions of Chrome, FireFox and Safari.
+\item Each NodEO instance can be restarted independenly.
+\end{itemize}
 
+A sequence diagram of the W$^2$ version of {\sf NodIO} is presented 
+in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
+API to run two NodEO instances per browser. The sequence diagram shows the 
+interaction between processess in a time sequence. In this system there are
+two kinds of asynchronus messages: HTTP Requests shown as bold arrows between 
+the client and the server where PUT and GET HTTP Methods are represented by 
+red and blue arrows respectively. There are also messages between workers and 
+the main thread in locally in the client, this are shown as thin solid arrows.
+There are three processes involved:
+\begin{itemize}
+\item {\em nodejs} This is the server side process, implented in the nodejs 
+framework. Web request handlers are implemented using the Express web 
+framework. For keeping a log of the execution the Winston library is used, and 
+finally the NodeIO framework is responsible for keeping the shared pool.
+\item {\em Client main thread}  This process handles the user interaction
+using the JQuery framwork. The Chart.js library is used to display a chart of
+the best fitness in the population over time. The main script also creates 
+worker instances and handles the user interface.
+\item {\em Worker global scope} The worker execution space does not have 
+access to the Document Object Model (DOM) or JQuery objects in the main thread.
+Asynchronus HTTP requests to the server are implemented using the XHR object. 
 \end{itemize}
+
+The sequence of events are:
+\begin{enumerate}
+\item First the NodIO process is started in nodejs. As part of the 
+initialization procedure the number of experiment is initialised and a log 
+file is created. The shared pool implemented as an array is reset.
+\item A volunteer follows the link of the experiment triggering a GET HTTP 
+request and the page is loaded.
+\item The main script renders the page and creates the worker instance. In 
+order to create a dedicated worker instance another GET request is sent to the 
+server in order to recieve the worker implementation script. 
+Each worker instance executes one or more evolutionary algorithms 
+{\em islands}. Each starts with random population and it is assigned a 
+universally unique  identifier (UUID). The UUID is included in each of the HTTP
+requests sent to the server.  
+The initial paramaters for each island
+can be distinct. Once the worker is created and the parameters have been set,
+a message is sent to the worker to start the algorithm.
+\item The worker first creates the population following with the previously 
+specified parameters. Once the population has been initialised, the EA 
+generation loop is started. After n generations the state of the population 
+is evaluated. First the chromosome with the best fitness is sent to the shared
+population with a PUT HTTP request. If the best fitness is not a global best,
+the EA loop must continue, but first a chromosome is randomly selected from 
+the shared pupulation. Also the current generation number and the best 
+chromosome is sent to the client main thread in order to update the DOM to 
+display the current state of the island (described next). These messages are
+all asynchronous so even before a message is recieved by target the EA loop of
+the worker could continue. If the best fitness is a global best the best 
+chromosme is sent to the server and the loop ends. A message is sent to the 
+main thread indicating that the best individual has been found.
+\item The main thread recieves messages from workers as events handled by 
+a callback function. When an iteration message is recived ,the page is 
+updated by displaying a dinamic plot of the geneartion number and best 
+fitness, and a representation of the chromosme is also generated.   
+\item When a global best is recieved from an island the current experiment 
+ends, the experiment number is incremented and the population array is reset.  
+\item When the main thread recieves a message the a worker has ended the EA
+loop, that worker is reinitialized. The worker process is not ended, to
+avoid the cost, only the parameters and population is reset and a new UIID 
+assigned.
+\end{enumerate}
 
@@
 \caption{Description of the W$^2$ version of {\sf NodIO}. In this
-  version, clientts use WebWorkers to run the evolutionary algorithm,
+  version, clients use Web Workers to run the evolutionary algorithm,
   with two of them per browser.}  

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Revert "More about the algorithm"
This reverts commit 4e35e721982c6ae9b9bccaf38301da7457f4a2cb.

--- 2015-08-07 volunteer-clouds-ieee-ci-si.tex
@@
 brings several advantages over a single threaded implementation:
+
 \begin{itemize}
@@
 keeps running in the background even it the user brings forward another tabbed window.
-\item Several instances of the NodEO script can be executed in parallel in a single web page. 
+\item Several instances of NodEO can be executed in parallel in a single web page. 
 \item Many immplemantations of the Web Worker API can use multiple CPUs for the execution
@@
 Chrome, FireFox and Safari.
+\item Each NodEO instance can be restarted independenly.
+\end{itemize}
 
+A sequence diagram of the W$^2$ version of {\sf NodIO} is presented in Figure~\ref{fig:system:w2}. In this
+version, clients use the Web Workers API to run two NodEO instances per browser. The sequence diagram shows
+the interaction between processess in a time sequence. In this system there are two kinds of asynchronus
+messages: HTTP Requests shown as bold arrows between the client and the server where PUT and GET HTTP Methods
+are represented by red and blue arrows respectively. There are also messages between workers and the main thread in
+locally in the client, this are shown as thin solid arrows. There are three processes involved:
+
+\begin{itemize}
+\item {\em nodejs} This is the server side process, implented in the nodejs framework. Web request handlers are
+implemented using the Express web framework. For keeping a log of the execution the Winston library is used, and finally
+the NodeIO framework is responsible for keeping the shared pool.
+\item {\em Client main thread}  This process handles the user interaction using the JQuery framwork. The Chart.js library
+is used to display a chart of the best fitness in the population over time. The main script also creates 
+worker instances and handles the user interface.  
+\item {\em Worker global scope} The worker execution space does not have access to the Document Object Model (DOM) 
+or JQuery objects in the main thread. Asynchronus HTTP requests to the server are implemented using the XHR object. 
 \end{itemize}
+
+The sequence of events are:
+\begin{enumerate}
+\item First the NodIO process is started in nodejs. As part of the initialization procedure the number
+of experiment is initialised and a log file is created. The shared pool implemented as an array is reset.
+\item A volunteer follows the link of the experiment triggering a GET HTTP request and the page is 
+loaded.
+\item The main script renders the page and creates the worker 
+\end{enumerate}
 

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Description of NodIO-W2
This is a brain dump to get the ideas, heavy edit needed.
Revised intro and SoA
There are still many comments and offers to write stuff. Please do it, this should be terminated @mariosky @fergunet @pacastillo

--- 2015-08-08 volunteer-clouds-ieee-ci-si.tex
@@
   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+  \graphicspath{{img/}}
 \else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
   \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
+  \graphicspath{{img/}}
 \fi
@@
 still has a long way to go if it is going to become a mainstream technology in the
-same way the cloud technologies in general have. The lack of general purpose and open source tools
+same way the cloud technologies in general have. The lack of general
+purpose and open source tools 
 with analytics built in could be one of the reasons why this is so, as
@@
 asynchronous, distributed evolutionary algorithms running in the
-browser. The architecture, along with some initial results on the expected performance and scaling
-behavior is presented. We also use the algorithm data to  %[Pedro] I don't understand what "the algorithm analytics" stand for???
-% data - JJ
-find out what kind of patterns the user behavior follow and try to fit them to some statistical distribution.
+browser. The architecture, along with some initial results on the
+expected performance and scaling 
+behavior is presented. We also use the algorithm data to 
+find out what kind of patterns the user behavior follow and try to fit
+them to some statistical distribution. 
 \end{abstract}
@@
 Evolutionary algorithms, evolutionary computation, distributed computing, internet computing,
-social networks, volunteer computing, metacomputing. 
+social networks, volunteer computing, metacomputing, performance evaluation. 
 \end{IEEEkeywords}
@@
 
-There is, then, space for systems that do not have any of these %[Mario] these assumptions? [JJ] OK
+There is, then, space for systems that do not have any of these 
 assumptions, such as the one presented in this paper, which will be
@@
  including Flash animations and VBScript, ActiveX or Java applets were
-used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
+used; for instance, Java was pointed out in \cite{soares1998get} as a
+``language for the
 internet'', providing some advantages such as multi-architecture compatibility or 
@@
 experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to Jet, such as SuperWeb
+above. Other proposals contemporary to Jet, such as {\em SuperWeb}
 \cite{alexandrov1997superweb} are roughly a translation of classical
@@
 evolutionary computation and it has been used ever since by several
-authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
+authors, including more recent efforts
+\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}
+that even 
 used the client's GPU \cite{duda2013gpu} or create an object-oriented
@@
 and DropBox~\cite{mericloud}. The use of JavaScript to implement
-server side components, %FERGU: he puesto mejor el de revista en lugar
-                        %del del CEC 
+server side components 
 gives software developers certain advantages in the cloud
-paradigm. JavaScript is a first-classe citizen 
-in most Platforms as a Service (PaaS) as many support the Node.js
-framework \cite{wood13:nodejs:paas}. Also many web services  
-have native JavaScript libraries and use JSON for data
+paradigm, since is a first-classe citizen 
+in most Platforms as a Service (PaaS) products as most support the Node.js
+framework \cite{wood13:nodejs:paas}. Many web services  
+have native JavaScript libraries too, and use JSON (JavaScript Object Notation) for data
 interchange. Another advantage is that  
@@
 side development.
-Besides the advantages to deal with the same language in the full
-stack as previously mentioned,  
-this also facilitates the interoperability of the exchanged data types.
 %Ayuda con la mini conclusion FERGU: mira la última frase de arriba
+% Borrada por iterativa - JJ
+
+JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
+According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
+metaheuristics optimization allows an improvement in communicability, 
+reproducibility, interoperability, automated assembly, knowledge engineering 
+and efficiency.
 
 In the work presented in this paper, JavaScript has been used throughout the full 
-stack, which has the advantage of using the same lines of code for
+stack, which has the advantage of using the self-same lines of code for
 writing an evolutionary algorithm for the
@@
 %development..
+% Ya se ha dicho, ¿no? - JJ
 
-JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
-According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
-metaheuristics optimization allows an improvement in communicability, 
-reproducibility, interoperability, automated assembly, knowledge engineering 
-and efficiency.
-
-There have been relatively few efforts to analyze the
-performance that can be obtained from these volunteer computing
-experiments. 
+On the other hand, we are also interested in measuring the performance
+of volunteer computing systems, an area in which there have been
+relatively few efforts.  
 There were some initial attempts to avoid the differences in performance
@@
 %o según la demanda.
+% Ya se ha hablado. Revísalo - JJ
 

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Expanding the baseline experiment

@@
 find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$, a population size equal 512, which was found in previous
-experiments to be the correct one for finding the solution 95\% of the
-%Fergu: cita a esos experimentos?
-time. The algorithm was run until the solution (a string with all
+b=2, z=3$, a population size equal 512. The algorithm was run until the solution (a string with all
 ones) was found or five millions evaluations had been performed. It
@@
 
-These two experiments establish first a baseline result, which will
-depend on the population. The volunteer computing experiments that we
+These two experiments establish first a baseline result and show that
+it depends on the population. The volunteer computing experiments that we
 will describe next do not and can not have the same conditions, but
@@
 prove next that that target performance can be achieved by carrying
-out several experiments in different conditions.
+out several experiments in different conditions. Second, they act also
+as a basic performance test for the server, and also prove that the
+basic server can be used for many kind of different set-ups, as long
+as they follow the REST protocolo to deposit and obtain individuals
+from the pool. It is obvious that, in this case, the server is not
+really needed, since there is a single client and no interaction
+between them, but there is a cost in sending and obtaining chromosomes
+from the server so it is included for the sake of fairer comparison.
 

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Added intro to Nodio-WW section
Please @mariosky check that it is correct.

@@
 
+In order to improve the number of cycles per user, we improved the
+basic architecture in several different aspects. By using WebWorkers
+the algorithm kept running on the background even if the {\sf NodIO}
+tab was not in focus. It also allowed to run several clients per
+browser, as opposed to a single one as we did previously. Another
+improvement involved restart of the client once the solution was
+found, so that as long as the user was visiting the page it was
+running an evolutionary algorithm. Finally, a small algorithmic
+improvement was included: instead of having a fixed population,
+population size was random and different for each client.
+
 JavaScript has commonly been used to develop client-side scripts

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
JS and plugin vulnerabilities

@@
 technologies is that they are not {\em universal} in the
-sense than an extra component, namely, the Java virtual machine, has
-to be installed in the browser. JavaScript, however,
+sense than an extra component, namely, the Java virtual machine or another
+virtual environment has to be installed in the browser. 
+Also these virtual environments have become targets of malicious attacks; % [Mario] Aqui empieza 
+for instance numerous vulnerabilities have been discovered in % lo que decía de la seguridad.
+the propietary Adobe Flash runtime environment and 
+video player \cite{ford2009analyzing,watanabe2010new}. The need for
+these additional environments has been mitigated with the 
+adoption of the HTML5 standard \cite{anthes2012html5}.
+JavaScript, however,
 \cite{flanagan2006javascript} was introduced in 1997 as a
@@
 side development.
-%Ayuda con la mini conclusion FERGU: mira la última frase de arriba
+% Ayuda con la mini conclusion FERGU: mira la última frase de arriba
 % Borrada por iterativa - JJ
@@
 desktop or command line and browser clients, both based in the EC
-library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time % [Mario] Creo que si se debería
+% dedicar una oración para recordar la diferencia y relación entre NodIO y NodEO
+% En mi mente los digo igual, NodEO como Rodeo y NodIO como Radio y me costó trabajo ver la diferencia
+% Pero igual un lector que llegue a leer sin vicios, si vea claramente la relación. 
 when writing code and leverages %FERGU: qué diferencias hay entre NodIO y NodEO??
@@
 % Ya se ha dicho, ¿no? - JJ 
+% Si, eso ya quedó - Mario
 
@@
 % [JJ] Escribe algo... 
+% [Mario] Ok, aunque va unos párrafos antes.
 
@@
 %empleando las tecnologías estándar (JavaScript, JSON, Cookies, etc.) mencionadas previamente.
+%[Mario] Ya quedó esto
 
@@
 % Ya se ha hablado. Revísalo - JJ
+% Si, ya quedó -Mario
 
@@
 % Para la intro o esta sección - JJ
+% Ya se hizo - Mario
 
@@
 % webworkers. La anterior era más adecuada como visión general. [JJ]
+% [Mario] Ok
 \begin{figure}[!t]
@@
 %[Mario] Creo que es NodEO y NodIO fué mi error. 
+%[Mario] Ok, ya me quedó claro. Pero no sé si quede claro en el paper, sin leer el otro.
 the node.js evolutionary algorithm library. This experiment tries to 

### 2015-08-08 data/plot-zipf-openshift.R
Adds ww data

@@ -24,3 +24,15 @@ dt.puts.7.31 <- process.puts("7-31")
 
 
 ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4.4,aes(x=x/length(dt.puts.4.4$x),y=puts,color='4/4'),stat='Identity')+geom_point(data=dt.puts.4.24,aes(x=x/length(dt.puts.4.24$x),y=puts,color='4/24'),stat='Identity')+geom_point(data=dt.puts.7.31,aes(x=x/length(dt.puts.7.31$x),y=puts,color='7/31'),stat='Identity')
+
+all.puts <- data.frame(Experiment=rep("4-4",length(dt.puts.4.4$puts)),
+                       x = dt.puts.4.4$x,
+                       puts = dt.puts.4.4$puts )
+all.puts <- rbind(all.puts,
+                  data.frame(Experiment=rep("4-24",length(dt.puts.4.24$puts)),
+                             x = dt.puts.4.24$x,
+                             puts = dt.puts.4.24$puts ))
+all.puts <- rbind(all.puts,
+                  data.frame(Experiment=rep("7-31",length(dt.puts.7.31$puts)),
+                             x = dt.puts.7.31$x,
+                             puts = dt.puts.7.31$puts ))

### 2015-08-08 data/time-vs-IPs-openshift.R

@@ -18,17 +18,27 @@ time.vs.IPs <- function( data, preffix ) {
 }
 
 suffix <- c("4-4","4-24","7-31")
-
+all.openshift <- data.frame(Experiment=as.Date(character()),
+                            IPs=character(), 
+                            milliseconds=character()) 
 for ( i in suffix ) {
     ips.time <- read.csv(paste0("ips-duration-openshift-",i,".dat"))
+    all.openshift <- rbind(all.openshift,
+                           data.frame( Experiment=rep(i,length(ips.time$IPs)),
+                                      IPs = ips.time$IPs,
+                                      milliseconds = ips.time$milliseconds))
     time.vs.IPs(ips.time, i)
     print(paste(i, summary(ips.time)))
     print(paste("< 69s", length(ips.time[ips.time$milliseconds < 69000,]$milliseconds)/length(ips.time$milliseconds)))
     print(paste("< 3.46s", length(ips.time[ips.time$milliseconds < 3460,]$milliseconds)/length(ips.time$milliseconds)))
     ips.time.minus1 <- ips.time[c(1:length(ips.time$IPs)-1),]
     ips.time.plus1 <- ips.time[c(2:length(ips.time$IPs)),]
-    print(length(ips.time.minus1$IPs))
-    print(length(ips.time.plus1$IPs))
     print(cor(ips.time.minus1, ips.time.plus1))
 }
- 
+
+
+all.openshift.minus1 <- all.openshift[c(1:length(all.openshift$IPs)-1),]
+all.openshift.plus1 <- all.openshift[c(2:length(all.openshift$IPs)),]
+print(cor(all.openshift.minus1$IPs, all.openshift.plus1$IPs))
+print(paste("< 69s", length(all.openshift[all.openshift$milliseconds < 69000,]$milliseconds)/length(all.openshift$milliseconds)))
+print(paste("< 3.46s", length(all.openshift[all.openshift$milliseconds < 3460,]$milliseconds)/length(all.openshift$milliseconds)))

### 2015-08-08 data/time-vs-IPs.R

@@ -9,3 +9,8 @@ ips.time$i <- 1:nrow(ips.time)
 ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
 ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
 
+ print(paste("< 69s", length(ips.time[ips.time$milliseconds < 69000,]$milliseconds)/length(ips.time$milliseconds)))
+print(paste("< 3.46s", length(ips.time[ips.time$milliseconds < 3460,]$milliseconds)/length(ips.time$milliseconds)))
+ips.time.minus1 <- ips.time[c(1:length(ips.time$IPs)-1),]
+ips.time.plus1 <- ips.time[c(2:length(ips.time$IPs)),]
+print(cor(ips.time.minus1$IPs, ips.time.plus1$IPs))

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Adds ww data
Check out new results
Look, @mariosky, they are so much better than what we had before. Great job!

@@
 generations divided by 100, contributed by every user, which is shown
-in Figure \ref{fig:put:os}. These results show a Zipf-like behavior,
+in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
 so we have fitted it to the Generalized Extreme Value distribution,
-with the resulting parameters shown in Table \ref{tab:put:os}.
+with the resulting parameters shown in Table \ref{tab:puts:os}.
 %
@@
   leans} towards the origin, negative towards the other extreme
-value. In this case, Table \cite{tab:puts:os} shows $\xi$ values
+value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
 greater than one and between one and 1.4, which indicates that the
@@
 distributions with these parameters have been plotted in red in
-\ref{fig:put:os}, indicating that the fit is good enough. The
+\ref{fig:puts:os}, indicating that the fit is good enough. The
 predictive value of these fits, however, is limited, over all taking
@@
 \end{enumerate}
-
+%
 \begin{figure}[!t]
@@
 \end{figure}
+%
 
+\begin{table*}
+\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
+  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
+\begin{center}
+\begin{tabular}{l|ccccccc}
+\hline
+Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
+  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
+\hline 
+{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
+{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
+\hline 
+\end{tabular}
+\end{center}
+\end{table*}
+%
 %---------------------------------------------------------------

### 2015-08-08 data/ips-puts.R

@@ -1,0 +1,3 @@
+load(ggplot2)
+
+ips.puts.ww <- read.csv("ips-puts-ww.dat")

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Revised baseline
Closes #17

@@
 \usepackage{array}
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
 
@@
 
-% Note that keywords are not normally used for peerreview papers.
 \begin{IEEEkeywords}
@@
 
-This last point is what we will try to prove next.
+This last point is what we will try to prove in the next section.
 
@@
 \caption{Comparison of time to solution for the baseline system, using a {\tt
-    node.js} client, for different population sizes. They only compare on
-runs in which solution has been found.}
+    node.js} client, for different population sizes. Only
+runs in which solution has been found are used for this graph.}
 \label{fig:baseline}
@@
 First we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO, 
-%FERGU: es NodEO o NodIO? Quito el \sf, por cierto, que queda feo
-%Por qué usamos NodEO aquí y no NodIO, vamos a compararlos? 
-%Qué diferencias hay entre los dos? (ponerlo en el SOA)
-%[Mario] Creo que es NodEO y NodIO fué mi error. 
-%[Mario] Ok, ya me quedó claro. Pero no sé si quede claro en el paper, sin leer el otro.
-the node.js evolutionary algorithm library. This experiment tries to 
-%FERGU: quitar el "the node.js evolutionary algorithm library", si ya lo hemos dicho
-%o explicarlo mejor: "the basic js library we have used as base to built NodIO", por ejemplo
+evolutionary algorithm in a desktop client written using NodEO \cite{figshare:nodeo}, the basic JavaScript library we have used as base to built {\sf NodIO}. This experiment tries to 
 find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$, a population size equal 512. The algorithm was run until the solution (a string with all
-ones) was found or five millions evaluations had been performed. It
+b=2, z=3$ and a population size equal 512. The algorithm was run until the solution (a string with all
+ones) was found or five million evaluations had been performed. It
 took around a minute, on average, that is $68.9694$ seconds, to

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Adds description of results

@@
 well as the lack of proofs of concept. That is why in this paper we
-present NodIO, a client-server architecture for distributed
+present {\sf NodIO}, a client-server architecture for distributed
 evolutionary algorithms, written entirely in JavaScript, which needs
@@
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. NodIO is first an application that
+provisionally named {\sf NodIO}. {\sf NodIO} is first an application that
 runs in two tiers, a server and a client, the latter  can be possibly
@@
 First we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO \cite{figshare:nodeo}, the basic JavaScript library we have used as base to built {\sf NodIO}. This experiment tries to 
+evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to built {\sf NodIO}. This experiment tries to 
 find the solution to the 40-trap function with parameters $l=4, a=1,
@@
 \item {\em Client main thread}  This process handles the user interaction
-using the JQuery framwork. The Chart.js library is used to display a chart of
+using the JQuery framework. The Chart.js library is used to display a chart of
 the best fitness in the population over time. The main script also creates 
@@
 \end{itemize}
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=4in]{img/Algorithm.pdf}
+\caption{Description of the W$^2$ version of {\sf NodIO}. In this
+  version, clients use Web Workers to run the evolutionary algorithm,
+  with two of them per browser.}  
+\label{fig:system:w2}
+\end{figure}
 
-The sequence of events are:
+The sequence of events is as follows:
 \begin{enumerate}
-\item First the NodIO process is started in nodejs. As part of the 
+\item First the {\sf NodIO} process is started in nodejs. As part of the 
 initialization procedure the number of experiment is initialised and a log 
@@
 %
-\begin{figure}[!t]
-\centering
-\includegraphics[width=4in]{img/Algorithm.pdf}
-\caption{Description of the W$^2$ version of {\sf NodIO}. In this
-  version, clients use Web Workers to run the evolutionary algorithm,
-  with two of them per browser.}  
-\label{fig:system:w2}
-\end{figure}
-%
-
-\begin{table*}
+\begin{table*}[!htb]
 \caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
@@
 %
+
+The experiments were run in the same way as before, using several
+Twitter announcements throughout several days by the end of July
+2015. Eventually, more than one thousand experiments were completed in
+a matter of days. A summary of results are shown in Table
+\ref{tab:summary:ww}, comparing it with the aggregate of the results
+obtained with the initial version of {\sf NodIO}. These results are
+remarkably different, being similar only in the median and maximum
+number of IPs, although in this case it would be combination
+IP-worker, being considered every worker a different IP. The median
+number of seconds is almost 80\% better than the previous version, and
+the median number of PUTs is three times better; this causes that most
+of experiment take less time than one of the baselines, and one third
+less than the second baseline. Thus, the conclusion in this case is
+that we can obtain, through volunteer computation, a result that is,
+most of the times, better than we would using a similar, non-parallel,
+desktop setup, which makes this system suitable for massively
+distributed evolutionary algorithms.
+
+However, it is interesting to note why this is so, and the first hint
+is the inter-experiment correlation between the number of IPs in
+succesive experiments. While before it was an unremarkable 10\%, it is
+now more than 43\%, making the numbef of IPs in contiguous experiments
+highly correlated. The main reason for this is the setup in {\sf
+  NodIO-W$^2$}, which restarts an island after a solution has been
+found and also is running even if the tab is not on the
+foreground. This means that volunteers can leave the experiment
+running for as long as they want and they will be contributing to
+experiment after experiment, unlike the previous version, where they
+stopped contributing after the solution was found. This is also
+reflected in the number of PUTs, which is the number of generations
+divided by 100; more than 50\% of the volunteers run for more than
+4000 generations. All things considered, this means than there are
+many more computers running {\em simultaneously}, leading to this
+almost 5-fold increase in running time. 
 %---------------------------------------------------------------
@@
 this paper (and our work) with their suggestions. We are also grateful
-to Anna Saéz de Tejada for her help with the data processing scripts.
+to Anna S\'aez de Tejada for her help with the data processing scripts.
 

### 2015-08-08 data/ips-puts.R
Adds GEV fit

@@ -1,3 +1,5 @@
 load(ggplot2)
-
+library(evd)
 ips.puts.ww <- read.csv("ips-puts-ww.dat")
+this.fit <- fgev(ips.puts.ww$puts)
+print(this.fit)

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Graph with explanation

@@
 almost 5-fold increase in running time.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{img/ips-time-ww.png}
+\caption{Time employed in every experiment vs.
+number of IPs (in abscises).
+The red dot and line show the average and standard deviation, the shade of blue the number of experiments with the same number of IPs.
+}
+\label{fig:ipstime:w2}
+\end{figure}
+%
+It is also interesting to measure the scaling properties in this
+model, that is, the relationship between the number of IPs
+participating in an experiment and the time it takes to complete
+it. This is shown in Figure \ref{fig:ipstime:ww}, which shows every
+experiment in a time (in milliseconds) vs. number of workers participating
+in it. Although there is not a clear trend, the graph seems to say
+that the time needed for finding the solution in this evolutionary 
+algorithm does not depend on the number of workers participating in
+it. This does not mean that it is independent on the number of {\em
+  simultaneous} users, which is probably the case although it is more
+difficult to measure. There seems to be a trend towards a decreasing
+standard deviation in the time, with more volunteers adding {\em
+  robustness} to an experiment and certainty in the time it is taking;
+however, more experiments would be needed to check this hypothesis.
+
 %---------------------------------------------------------------

### 2015-08-09 data/time-vs-IPs.R

@@ -1,6 +1,6 @@
 library(ggplot2)
 
-ips.time <- read.csv("ips-time.csv")
+ips.time <- read.csv("ips-time-ww.csv")
 ips.time$howmany <- 1:nrow(ips.time)
 for ( i in 1:nrow(ips.time) ) {
     ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Adding explanation for WW
Please @mariosky @fergunet @pacastillo check the conclusions. This is almost done now, I'll write the conclusions soon.

@@
 %
+%
+\begin{table*}
+\caption{Summary of fit to Generalized Extreme Value distribution of
+  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
+\begin{center}
+\begin{tabular}{ccc}
+\hline
+Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+\hline 
+18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$ 0.04746 \\       
+\hline 
+\end{tabular}
+\end{center}
+\end{table*}
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{img/gev-fit-ww.png}
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red)
+  for the {\sf NodIO$^2$} model. }  
+\label{fig:gev:w2}
+\end{figure}
+%
 It is also interesting to measure the scaling properties in this
@@
 participating in an experiment and the time it takes to complete
-it. This is shown in Figure \ref{fig:ipstime:ww}, which shows every
+it. This is shown in Figure \ref{fig:ipstime:w2}, which shows every
 experiment in a time (in milliseconds) vs. number of workers participating
@@
 however, more experiments would be needed to check this hypothesis.
+
+As we did in the previous subsection, we have fitted also the number
+of HTTP {\tt PUT}s per worker to a GEV function, with the result shown
+in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
+in the previous experiment, \ref{tab:puts:os}, we find higher values
+for the location $\mu$ and scale $\sigma$ parameters, accounting for
+the higher number of contributions per volunteer that has been
+previously observed; however, the shape $\xi$ parameter is
+substantially similar, with a {\em bell} shape leaning towards the
+origin, indicating a pattern of a few users with many contributions
+and many with few contributions. If we plot the model and the data
+side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
+is not so tight as in the previous case, with the model overestimating
+the number of individuals with a high number contributions. This is
+probably due to the fact that in the previous case, the user needed to
+voluntarily reload the page to contribute the most generations, with a
+{\em phase change} between the people that did so and the people that
+just run the experiment until completion. In this case, however, the
+user just needs to let it run, without that phase change and thus a
+more {\em equalitarian} distribution of contributions.
 

### 2015-08-09 data/ips-puts.R

@@ -1,5 +1,13 @@
-load(ggplot2)
+library(ggplot2)
 library(evd)
 ips.puts.ww <- read.csv("ips-puts-ww.dat")
+ips.puts.ww.df <- data.frame(x=1:nrow(ips.puts.ww),
+                             puts=sort(ips.puts.ww$puts,decreasing=T))
 this.fit <- fgev(ips.puts.ww$puts)
 print(this.fit)
+this.fit.plot <- data.frame(x=1:length(ips.puts.ww$puts),
+                            y=sort(rgev(length(ips.puts.ww$puts),
+                                this.fit$estimate["loc"],
+                                this.fit$estimate["scale"],
+                                this.fit$estimate["shape"]),decreasing=T))
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=x,y=puts,color='Data'))+geom_point(data=this.fit.plot,aes(x=x,y=y,color='Fit'))+scale_y_log10()

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Fixing some typos and adding comments for @JJ and @mariosky

@@
 for instance numerous vulnerabilities have been discovered in % lo que decía de la seguridad.
-the propietary Adobe Flash runtime environment and 
+the proprietary Adobe Flash runtime environment and
 video player \cite{ford2009analyzing,watanabe2010new}. The need for
@@
 \end{figure}
-First we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to built {\sf NodIO}. This experiment tries to 
+First, we establish the baseline performance by running the
+evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to build {\sf NodIO}. This experiment tries to
 find the solution to the 40-trap function with parameters $l=4, a=1,
@@
 seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}.
+are shown in Figure \ref{fig:baseline}. %Fergu: dejar claro que es solo una isla
 
@@
 it depends on the population. The volunteer computing experiments that we
+%FERGU: explicar un poco más lo de "depends on the population"?, por ejemplo añadiendo
+%as this is a deceptive problem, and low levels or diversity may not help to
+%find the solution \cite{} o algo parecido.
 will describe next do not and can not have the same conditions, but
@@
 prove next that that target performance can be achieved by carrying
-out several experiments in different conditions. Second, they act also
+out several experiments in different conditions. Second, they also act
 as a basic performance test for the server, and also prove that the
-basic server can be used for many kind of different set-ups, as long
-as they follow the REST protocolo to deposit and obtain individuals
+basic server can be used for many kinds of different setups, as long
+as they follow the REST protocol to deposit and obtain individuals
 from the pool. It is obvious that, in this case, the server is not
@@
 Initial experiments were set up using the OpenShift
-Platform-as-a-Service, which provides a free tier within which this
+PaaS, which provides a free tier within which this %Ya se defnió antes PaaS
 experiment could be performed. That way, and as required, the hardware
@@
 the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proved that the influence
+the next. However, previous experiments have proven that the influence
 of these islands in the next experiment is, in fact negligible.
@@
 %
-The table shows that every run included more than 50 experiments. The
+The table shows that every run included more than 50 experiments.
+The %FERGU: No debería ser mejor al revés?
+Cada experimento está compuesto de varias runs.
+De todas maneras arriba lo has definido así, así que si no quieres no lo cambies.
+Pero me parece más correcto que un experimento tenga runs, en vez de al revés
 number of different IPs intervening in them varied from more than one
@@
 second case. The comparison is not totally fair: this experiment takes
-place in a browser and in fact most of the time is devoted to
+place in a browser, and in fact most of the time is devoted to
 presenting the charts; the server is not the same since it is run in a
@@
 in our very own computer. And the results is that although it will
-happen, one out of roughly 5 or roughly 20, depending on what baseline
+happen, one out of roughly 5 or roughly 20, depending on what baseline %FERGU: one->1?
 experiment you choose, it will not happen always or even a majority of
@@
 the evolutionary algorithm start again. However, it puts a probably
-unnecesary cap on the time, or number of experiments, that an user
+unnecessary cap on the time, or number of experiments, that a user %Fergu: user no tiene "an", me lo dijo paloma. es "yuser", no "iuser", por eso no lo tiene :P
 will voluntarily perform.
@@
 \includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-7-31.png}
-\caption{Duration of experiments vs rank, with $y$ axis in a
+\caption{Duration of experiments vs. rank, with $y$ axis in a
   logarithmic scale. Dot color is related to the number of IPs
@@
 that increases the probability of someone new joining them. In fact,
-wit experiments taking a few seconds and due to the way the
+with experiments taking a few seconds and due to the way the
 experiments are announced, it is quite difficult that several
@@
 have not found this convenient since there are several other issues
-that will be solved.
+that will be solved. %FERGU: and that will be explained next? Algún ejemplo?
 
@@
 issue led us to create the next version of the framework, which is
-presented next.
+presented next. %Fergu: esto mola, métele más caña en otras secciones para que se vea que es una conclusión util
 
@@
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects. By using WebWorkers
+basic architecture in several different aspects. By using WebWorkers %FERGU: WebWorker separado o junto? homogeneizar
 the algorithm kept running on the background even if the {\sf NodIO}
@@
 improvement was included: instead of having a fixed population,
-population size was random and different for each client.
+population size was random and different for each client. %FERGU: mínimo y máximo?
 
-JavaScript has commonly been used to develop client-side scripts
+JavaScript has been commonly used to develop client-side scripts
 that handle the user interface, communicate asynchronously with servers and
@@
 communication mechanism between the worker's execution environment and the
-main thread. According to the Web Workers W3C Candidate Recomendation 
+main thread. According to the Web Workers W3C Candidate Recommendation
 \cite{hickson2012web} workers are expected to be long-lived, they have a high
@@
 \item Several instances of NodEO can be executed in parallel in a single web
-page. 
-\item Many immplemantations of the Web Worker API can use multiple CPUs for
+page. %FERGU: NodEO o NodIO? (es NodEO la instancia, creo pero por si acaso :P)
+\item Many implementations of the Web Worker API can use multiple CPUs for
 the execution of worker scripts, in particular the execution in multiple CPUs
-was tested in desktop verions of Chrome, FireFox and Safari.
-\item Each NodEO instance can be restarted independenly.
+was tested in desktop versions of Chrome, FireFox and Safari.
+\item Each NodEO instance can be restarted independently.
 \end{itemize}
@@
 API to run two NodEO instances per browser. The sequence diagram shows the
-interaction between processess in a time sequence. In this system there are
-two kinds of asynchronus messages: HTTP Requests shown as bold arrows between 
+interaction between processes in a time sequence. In this system there are
+two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
 the client and the server where PUT and GET HTTP Methods are represented by
 red and blue arrows respectively. There are also messages between workers and
-the main thread in locally in the client, this are shown as thin solid arrows.
+the main thread in locally in the client, these are shown as thin solid arrows.
 There are three processes involved:
 \begin{itemize}
-\item {\em nodejs} This is the server side process, implented in the nodejs 
+\item {\em nodejs} This is the server side process, implemented in the nodejs %FERGU: creo que oficialmente es Node.js, homogeneizar en el paper
 framework. Web request handlers are implemented using the Express web
@@
 access to the Document Object Model (DOM) or JQuery objects in the main thread.
-Asynchronus HTTP requests to the server are implemented using the XHR object. 
+Asynchronous HTTP requests to the server are implemented using the XHR object. %FERGU: Qué es el objeto XHR? Alguna referencia/justificación de por qué se usa?
 \end{itemize}
@@
 \item First the {\sf NodIO} process is started in nodejs. As part of the
-initialization procedure the number of experiment is initialised and a log 
+initialization procedure the number of experiment is initialized and a log
 file is created. The shared pool implemented as an array is reset.
@@
 \item The main script renders the page and creates the worker instance. In
-order to create a dedicated worker instance another GET request is sent to the 
-server in order to recieve the worker implementation script. 
+order to create a dedicated worker instance, another GET request is sent to the
+server in order to receive the worker implementation script.
 Each worker instance executes one or more evolutionary algorithms
@@
 requests sent to the server.
-The initial paramaters for each island
+The initial parameters for each island
 can be distinct. Once the worker is created and the parameters have been set,
@@
 \item The worker first creates the population following with the previously
-specified parameters. Once the population has been initialised, the EA 
-generation loop is started. After n generations the state of the population 
+specified parameters. Once the population has been initialized, the EA
+generation loop is started. After {\em n} generations the state of the population
 is evaluated. First the chromosome with the best fitness is sent to the shared
@@
 the EA loop must continue, but first a chromosome is randomly selected from
-the shared pupulation. Also the current generation number and the best 
+the shared population. Also the current generation number and the best
 chromosome is sent to the client main thread in order to update the DOM to
 display the current state of the island (described next). These messages are
-all asynchronous so even before a message is recieved by target the EA loop of
+all asynchronous, so even before a message is received by target the EA loop of
 the worker could continue. If the best fitness is a global best the best
-chromosme is sent to the server and the loop ends. A message is sent to the 
+chromosome is sent to the server and the loop ends. A message is sent to the
 main thread indicating that the best individual has been found.
-\item The main thread recieves messages from workers as events handled by 
-a callback function. When an iteration message is recived ,the page is 
-updated by displaying a dinamic plot of the geneartion number and best 
-fitness, and a representation of the chromosme is also generated.   
-\item When a global best is recieved from an island the current experiment 
+\item The main thread receives messages from workers as events handled by
+a callback function. When an iteration message is received, the page is
+updated by displaying a dynamic plot of the generation number and best
+fitness, and a representation of the chromosome is also generated.
+\item When a global best is received from an island the current experiment
 ends, the experiment number is incremented and the population array is reset.
-\item When the main thread recieves a message the a worker has ended the EA
+\item When the main thread receives a message the a worker has ended the EA
 loop, that worker is reinitialized. The worker process is not ended, to
@@
 2015. Eventually, more than one thousand experiments were completed in
-a matter of days. A summary of results are shown in Table
+a matter of days. A summary of results is shown in Table
 \ref{tab:summary:ww}, comparing it with the aggregate of the results
@@
 the median number of PUTs is three times better; this causes that most
-of experiment take less time than one of the baselines, and one third
+of experiments take less time than one of the baselines, and one third
 less than the second baseline. Thus, the conclusion in this case is
@@
 
+%FERGU: a lo mejor algún reviewer se queja de que la comparativa no es justa
+%porque comparamos un mono-isla (el baseline) con un multi-isla. Quizás deberíamos
+%haber usado un multi-isla en el baseline (aunque sea secuencial). Para la próxima.
+
 However, it is interesting to note why this is so, and the first hint
 is the inter-experiment correlation between the number of IPs in
-succesive experiments. While before it was an unremarkable 10\%, it is
-now more than 43\%, making the numbef of IPs in contiguous experiments
+successive experiments. While before it was an unremarkable 10\%, it is
+now more than 43\%, making the number of IPs in contiguous experiments
 highly correlated. The main reason for this is the setup in {\sf
@@
 As we did in the previous subsection, we have fitted also the number
-of HTTP {\tt PUT}s per worker to a GEV function, with the result shown
+of HTTP {\tt PUT}s per worker to a GET function, with the result shown %FERGU: he cambiado GEV por GET (lo de GEV era un error, no? vayamos a que sea algo nuevo que no conozco...)
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
una letra d se había quedado suelta en la página 5, junto a la TABLE I (la he quitado)
una falta de ortografía en la línea 706 del .tex (arreglada)

@@
 %
-\begin{table*}d
+\begin{table*}
 \caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
@@
 greater than one and between one and 1.4, which indicates that the
-three experiment share this origin-leaning pattern, with many users
+three experiments share this origin-leaning pattern, with many users
 donating a few cycles and just a few donating extreme values. Random

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Web Workers (separado)

@@
 distributions with these parameters have been plotted in red in
-\ref{fig:puts:os}, indicating that the fit is good enough. The
+Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
 predictive value of these fits, however, is limited, over all taking
@@
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects. By using WebWorkers %FERGU: WebWorker separado o junto? homogeneizar
+basic architecture in several different aspects. By using Web Workers %FERGU: WebWorker separado o junto? homogeneizar
 the algorithm kept running on the background even if the {\sf NodIO}

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
iniciamos la redacción de las conclusiones con un resumen de lo que se ha presentado y lo que se ha obtenido en los experimentos

@@
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
-in the previous experiment, \ref{tab:puts:os}, we find higher values
+in the previous experiment, Table \ref{tab:puts:os}, we find higher values
 for the location $\mu$ and scale $\sigma$ parameters, accounting for
@@
 
-The conclusion goes here.
+% The conclusion goes here.
+
+In this paper a client-server architecture for distributed evolutionary algorithms, termed {\sf NodIO}, for volunteer computing has been presented.
+
+In order to establish a baseline performance, the evolutionary algorithm was run in a desktop client-program written in JavaScript using NodEO to solve the 40-trap function.
+In this first experiment, it could be observed how many users donated a few cycles and just a few donated a high amount of CPU to the experiment.
+
+As a second experiment, and trying to improve the number of cycles per user, the system architecture was modified.
+In this case, Web Workers were used so that several clients per browers could be executed.
+As a result, in this case a higher number of contributions per volunteer could be observed, although as in the previous experiments a pattern of a few users with many contributions and many with few contributions was obtained.
 

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Adds Weibull fit

--- 2015-08-09 volunteer-clouds-ieee-ci-si.tex
@@
 for instance numerous vulnerabilities have been discovered in % lo que decía de la seguridad.
-the proprietary Adobe Flash runtime environment and
+the propietary Adobe Flash runtime environment and 
 video player \cite{ford2009analyzing,watanabe2010new}. The need for
@@
 \end{figure}
-First, we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to build {\sf NodIO}. This experiment tries to
+First we establish the baseline performance by running the
+evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to built {\sf NodIO}. This experiment tries to 
 find the solution to the 40-trap function with parameters $l=4, a=1,
@@
 seconds. Results for these two experiments 
-are shown in Figure \ref{fig:baseline}. %Fergu: dejar claro que es solo una isla
+are shown in Figure \ref{fig:baseline}.
 
@@
 it depends on the population. The volunteer computing experiments that we
-%FERGU: explicar un poco más lo de "depends on the population"?, por ejemplo añadiendo
-%as this is a deceptive problem, and low levels or diversity may not help to
-%find the solution \cite{} o algo parecido.
 will describe next do not and can not have the same conditions, but
@@
 prove next that that target performance can be achieved by carrying
-out several experiments in different conditions. Second, they also act
+out several experiments in different conditions. Second, they act also
 as a basic performance test for the server, and also prove that the
-basic server can be used for many kinds of different setups, as long
-as they follow the REST protocol to deposit and obtain individuals
+basic server can be used for many kind of different set-ups, as long
+as they follow the REST protocolo to deposit and obtain individuals
 from the pool. It is obvious that, in this case, the server is not
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which this %Ya se defnió antes PaaS
+Platform-as-a-Service, which provides a free tier within which this
 experiment could be performed. That way, and as required, the hardware
@@
 the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proven that the influence
+the next. However, previous experiments have proved that the influence
 of these islands in the next experiment is, in fact negligible.
 %
-\begin{table*}
+\begin{table*}d
 \caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
@@
 %
-The table shows that every run included more than 50 experiments.
-The %FERGU: No debería ser mejor al revés?
-Cada experimento está compuesto de varias runs.
-De todas maneras arriba lo has definido así, así que si no quieres no lo cambies.
-Pero me parece más correcto que un experimento tenga runs, en vez de al revés
+The table shows that every run included more than 50 experiments. The
 number of different IPs intervening in them varied from more than one
@@
 second case. The comparison is not totally fair: this experiment takes
-place in a browser, and in fact most of the time is devoted to
+place in a browser and in fact most of the time is devoted to
 presenting the charts; the server is not the same since it is run in a
@@
 in our very own computer. And the results is that although it will
-happen, one out of roughly 5 or roughly 20, depending on what baseline %FERGU: one->1?
+happen, one out of roughly 5 or roughly 20, depending on what baseline
 experiment you choose, it will not happen always or even a majority of
@@
 the evolutionary algorithm start again. However, it puts a probably
-unnecessary cap on the time, or number of experiments, that a user %Fergu: user no tiene "an", me lo dijo paloma. es "yuser", no "iuser", por eso no lo tiene :P
+unnecesary cap on the time, or number of experiments, that an user
 will voluntarily perform.
@@
 \includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-7-31.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
+\caption{Duration of experiments vs rank, with $y$ axis in a
   logarithmic scale. Dot color is related to the number of IPs
@@
 that increases the probability of someone new joining them. In fact,
-with experiments taking a few seconds and due to the way the
+wit experiments taking a few seconds and due to the way the
 experiments are announced, it is quite difficult that several
@@
 have not found this convenient since there are several other issues
-that will be solved. %FERGU: and that will be explained next? Algún ejemplo?
+that will be solved.
 
@@
 greater than one and between one and 1.4, which indicates that the
-three experiments share this origin-leaning pattern, with many users
+three experiment share this origin-leaning pattern, with many users
 donating a few cycles and just a few donating extreme values. Random
@@
 issue led us to create the next version of the framework, which is
-presented next. %Fergu: esto mola, métele más caña en otras secciones para que se vea que es una conclusión util
+presented next.
 
@@
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects. By using WebWorkers %FERGU: WebWorker separado o junto? homogeneizar
+basic architecture in several different aspects. By using WebWorkers
 the algorithm kept running on the background even if the {\sf NodIO}
@@
 improvement was included: instead of having a fixed population,
-population size was random and different for each client. %FERGU: mínimo y máximo?
+population size was random and different for each client.
 
-JavaScript has been commonly used to develop client-side scripts
+JavaScript has commonly been used to develop client-side scripts
 that handle the user interface, communicate asynchronously with servers and 
@@
 communication mechanism between the worker's execution environment and the 
-main thread. According to the Web Workers W3C Candidate Recommendation
+main thread. According to the Web Workers W3C Candidate Recomendation 
 \cite{hickson2012web} workers are expected to be long-lived, they have a high 
@@
 \item Several instances of NodEO can be executed in parallel in a single web
-page. %FERGU: NodEO o NodIO? (es NodEO la instancia, creo pero por si acaso :P)
-\item Many implementations of the Web Worker API can use multiple CPUs for
+page. 
+\item Many immplemantations of the Web Worker API can use multiple CPUs for
 the execution of worker scripts, in particular the execution in multiple CPUs
-was tested in desktop versions of Chrome, FireFox and Safari.
-\item Each NodEO instance can be restarted independently.
+was tested in desktop verions of Chrome, FireFox and Safari.
+\item Each NodEO instance can be restarted independenly.
 \end{itemize}
@@
 API to run two NodEO instances per browser. The sequence diagram shows the 
-interaction between processes in a time sequence. In this system there are
-two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
+interaction between processess in a time sequence. In this system there are
+two kinds of asynchronus messages: HTTP Requests shown as bold arrows between 
 the client and the server where PUT and GET HTTP Methods are represented by 
 red and blue arrows respectively. There are also messages between workers and 
-the main thread in locally in the client, these are shown as thin solid arrows.
+the main thread in locally in the client, this are shown as thin solid arrows.
 There are three processes involved:
 \begin{itemize}
-\item {\em nodejs} This is the server side process, implemented in the nodejs %FERGU: creo que oficialmente es Node.js, homogeneizar en el paper
+\item {\em nodejs} This is the server side process, implented in the nodejs 
 framework. Web request handlers are implemented using the Express web 
@@
 access to the Document Object Model (DOM) or JQuery objects in the main thread.
-Asynchronous HTTP requests to the server are implemented using the XHR object. %FERGU: Qué es el objeto XHR? Alguna referencia/justificación de por qué se usa?
+Asynchronus HTTP requests to the server are implemented using the XHR object. 
 \end{itemize}
@@
 \item First the {\sf NodIO} process is started in nodejs. As part of the 
-initialization procedure the number of experiment is initialized and a log
+initialization procedure the number of experiment is initialised and a log 
 file is created. The shared pool implemented as an array is reset.
@@
 \item The main script renders the page and creates the worker instance. In 
-order to create a dedicated worker instance, another GET request is sent to the
-server in order to receive the worker implementation script.
+order to create a dedicated worker instance another GET request is sent to the 
+server in order to recieve the worker implementation script. 
 Each worker instance executes one or more evolutionary algorithms 
@@
 requests sent to the server.  
-The initial parameters for each island
+The initial paramaters for each island
 can be distinct. Once the worker is created and the parameters have been set,
@@
 \item The worker first creates the population following with the previously 
-specified parameters. Once the population has been initialized, the EA
-generation loop is started. After {\em n} generations the state of the population
+specified parameters. Once the population has been initialised, the EA 
+generation loop is started. After n generations the state of the population 
 is evaluated. First the chromosome with the best fitness is sent to the shared
@@
 the EA loop must continue, but first a chromosome is randomly selected from 
-the shared population. Also the current generation number and the best
+the shared pupulation. Also the current generation number and the best 
 chromosome is sent to the client main thread in order to update the DOM to 
 display the current state of the island (described next). These messages are
-all asynchronous, so even before a message is received by target the EA loop of
+all asynchronous so even before a message is recieved by target the EA loop of
 the worker could continue. If the best fitness is a global best the best 
-chromosome is sent to the server and the loop ends. A message is sent to the
+chromosme is sent to the server and the loop ends. A message is sent to the 
 main thread indicating that the best individual has been found.
-\item The main thread receives messages from workers as events handled by
-a callback function. When an iteration message is received, the page is
-updated by displaying a dynamic plot of the generation number and best
-fitness, and a representation of the chromosome is also generated.
-\item When a global best is received from an island the current experiment
+\item The main thread recieves messages from workers as events handled by 
+a callback function. When an iteration message is recived ,the page is 
+updated by displaying a dinamic plot of the geneartion number and best 
+fitness, and a representation of the chromosme is also generated.   
+\item When a global best is recieved from an island the current experiment 
 ends, the experiment number is incremented and the population array is reset.  
-\item When the main thread receives a message the a worker has ended the EA
+\item When the main thread recieves a message the a worker has ended the EA
 loop, that worker is reinitialized. The worker process is not ended, to
@@
 2015. Eventually, more than one thousand experiments were completed in
-a matter of days. A summary of results is shown in Table
+a matter of days. A summary of results are shown in Table
 \ref{tab:summary:ww}, comparing it with the aggregate of the results
@@
 the median number of PUTs is three times better; this causes that most
-of experiments take less time than one of the baselines, and one third
+of experiment take less time than one of the baselines, and one third
 less than the second baseline. Thus, the conclusion in this case is
@@
 
-%FERGU: a lo mejor algún reviewer se queja de que la comparativa no es justa
-%porque comparamos un mono-isla (el baseline) con un multi-isla. Quizás deberíamos
-%haber usado un multi-isla en el baseline (aunque sea secuencial). Para la próxima.
-
 However, it is interesting to note why this is so, and the first hint
 is the inter-experiment correlation between the number of IPs in
-successive experiments. While before it was an unremarkable 10\%, it is
-now more than 43\%, making the number of IPs in contiguous experiments
+succesive experiments. While before it was an unremarkable 10\%, it is
+now more than 43\%, making the numbef of IPs in contiguous experiments
 highly correlated. The main reason for this is the setup in {\sf
@@
 \centering
-\includegraphics[width=0.9\linewidth]{img/gev-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red)
-  for the {\sf NodIO$^2$} model. }
+\includegraphics[width=0.49\linewidth]{img/gev-fit-ww.png}
+\includegraphics[width=0.49\linewidth]{img/weibull-fit-ww.png}
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model. }  
 \label{fig:gev:w2}
@@
 As we did in the previous subsection, we have fitted also the number
-of HTTP {\tt PUT}s per worker to a GET function, with the result shown %FERGU: he cambiado GEV por GET (lo de GEV era un error, no? vayamos a que sea algo nuevo que no conozco...)
+of HTTP {\tt PUT}s per worker to a GEV function, with the result shown
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained

### 2015-08-09 data/ips-puts.R

@@ -1,5 +1,7 @@
 library(ggplot2)
 library(evd)
+library(fitdistrplus)
+
 ips.puts.ww <- read.csv("ips-puts-ww.dat")
 ips.puts.ww.df <- data.frame(x=1:nrow(ips.puts.ww),
                              puts=sort(ips.puts.ww$puts,decreasing=T))
@@ -11,3 +13,12 @@ this.fit.plot <- data.frame(x=1:length(ips.puts.ww$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 ggplot()+geom_point(data=ips.puts.ww.df,aes(x=x,y=puts,color='Data'))+geom_point(data=this.fit.plot,aes(x=x,y=y,color='Fit'))+scale_y_log10()
+
+print(fitdistr(ips.puts.ww$puts,"gamma"))
+this.weibull <- fitdistr(ips.puts.ww$puts,"weibull")
+
+this.fit.weib <- data.frame(x=1:length(ips.puts.ww$puts),
+y=sort(rweibull(length(ips.puts.ww$puts),
+    this.weibull$estimate["shape"],
+    this.weibull$estimate["scale"]),decreasing=T))
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=x,y=puts,color='Data'))+geom_point(data=this.fit.weib,aes(x=x,y=y,color='Fit'))+scale_y_log10()

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Volviendo a cambiar GET por GEV
Anda, que ya te vale, @fergunet. Menos mal que ha sido uno solo.

@@
 As we did in the previous subsection, we have fitted also the number
-of HTTP {\tt PUT}s per worker to a GET function, with the result shown %FERGU: he cambiado GEV por GET (lo de GEV era un error, no? vayamos a que sea algo nuevo que no conozco...)
+of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown %FERGU: he cambiado GEV por GET (lo de GEV era un error, no?
+vayamos a que sea algo nuevo que no conozco...)
+%cojollos, Fergu, si no conoces no lo cambies. 
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained

### 2015-08-09 volunteer-clouds-ieee-ci-si.tex
Added description of Weibull fit
Which is quite good, BTW.

@@
 \begin{table*}
-\caption{Summary of fit to Generalized Extreme Value distribution of
+\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
   the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
 \begin{center}
-\begin{tabular}{ccc}
+\begin{tabular}{cccc}
 \hline
-Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
 \hline
-18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$ 0.04746 \\
+GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
+0.04746 \\
+Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
 \hline
@@
 user just needs to let it run, without that phase change and thus a
-more {\em equalitarian} distribution of contributions.
+more {\em equalitarian} distribution of contributions, which rather
+corresponds to a Weibull function, as shown in Figure
+\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
+distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
+equal to 0.69, a value that is remarkably similar to the 0.5 value
+found for time devoted to games in Chambers et
+al. \cite{chambers2005measurement}.
 

### 2015-08-10 volunteer-clouds-ieee-ci-si.tex
Revisio of the intro and SoA
With deletion of completed comments.

@@
 obtained results in Section \ref{sec:experiments}.
-Finally, conclusions and future work are presented in Section \ref{sec:conclusion}.
+Finally, conclusions and future lines of work are presented in Section
+\ref{sec:conclusion}.
 
@@
 
-Fulfilling these properties could be achieved using any of the
-languages, and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first, but with the advantage of
+The fulfillment of these properties could be achieved using any of the
+languages and associated technologies, mentioned above, provided that the user had
+installed the required virtual machine plugin first, but embedding a
+ditributed computation client in the browser has the advantage of
 providing a single and universal way of accessing resources via an
@@
 side development.
-% Ayuda con la mini conclusion FERGU: mira la última frase de arriba
-% Borrada por iterativa - JJ
 
@@
 % En mi mente los digo igual, NodEO como Rodeo y NodIO como Radio y me costó trabajo ver la diferencia
-% Pero igual un lector que llegue a leer sin vicios, si vea claramente la relación.
+% Pero igual un lector que llegue a leer sin vicios, si vea claramente
+% la relación.
+% OK, lo hago más adelante - JJ
 when writing code and leverages %FERGU: qué diferencias hay entre NodIO y NodEO??
@@
 % Es el mismo código - JJ
-%The proposed software has the advantage of full stack JavaScript
-%development..
-% Ya se ha dicho, ¿no? - JJ
-% Si, eso ya quedó - Mario
 
@@
 
-%Mario: Estaba pensando que otra justificación muy fuerte para el uso de JS sobre plug-ins
-%ha sido la seguridad ya que estos presentan otra fuente de vulnerabilidad. Para complicar las cosas a veces (Flash, ActiveX)
-%son tecnologías cerradas y propietarias. Por lo menos creo que esas
-%fueron las razones que se dieron para eliminar a Flash del iPhone.
-% [JJ] Escribe algo...
-% [Mario] Ok, aunque va unos párrafos antes.
-
 As far as we know, this paper presents one of the few experiments that
@@
 well as the results obtained with the setup.
-
-%El último párrafo podría ser más preciso. Es el único experimento que utiliza recursos heterogeneos
-%empleando las tecnologías estándar (JavaScript, JSON, Cookies, etc.) mencionadas previamente.
-%[Mario] Ya quedó esto
-
-%Si no se extiende mucho se podría también hablar de la conexión con Cloud Computing,
-%se discuten las ventajas de utilizar JavaScript para acceder a recursos donados por voluntarios
-%pero falta hablar de como se explotaría el recurso Cloud. Lo digo por el titulo de Volunteers in the Clouds. Tal vez
-%se deba hablar de la posibilidad de escalar los sistemas a bajo costo
-%o según la demanda.
-% Ya se ha hablado. Revísalo - JJ
-% Si, ya quedó -Mario
-
-% Si te animas, propón un párrafo - JJ
-% Ok, lo escribo. - Mario
-% Voy escribiendo algo de HTML5 y Web Workers?, no necesariemente para
-% esta sección.
-% Para la intro o esta sección - JJ
-% Ya se hizo - Mario
 

### 2015-08-10 volunteer-clouds-ieee-ci-si.tex
Revised state of the art
Eliminated comments and done almost-final review. This closes #11.

@@
 virtual environment has to be installed in the browser.
-Also these virtual environments have become targets of malicious attacks; % [Mario] Aqui empieza
-for instance numerous vulnerabilities have been discovered in % lo que decía de la seguridad.
+Also these virtual environments have become targets of malicious attacks; 
+for instance numerous vulnerabilities have been discovered in 
 the proprietary Adobe Flash runtime environment and
@@
 paradigm, since is a first-classe citizen
-in most Platforms as a Service (PaaS) products as most support the Node.js
+in most Platforms as a Service (PaaS) products as most support the {\tt node.js}
 framework \cite{wood13:nodejs:paas}. Many web services
@@
 desktop or command line and browser clients, both based in the EC
-library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time % [Mario] Creo que si se debería
-% dedicar una oración para recordar la diferencia y relación entre NodIO y NodEO
-% En mi mente los digo igual, NodEO como Rodeo y NodIO como Radio y me costó trabajo ver la diferencia
-% Pero igual un lector que llegue a leer sin vicios, si vea claramente
-% la relación.
-% OK, lo hago más adelante - JJ
-when writing code and leverages %FERGU: qué diferencias hay entre NodIO y NodEO??
-expertise acquired in the library itself. %FERGU: tienes info de LoC
-                                %usadas para la sección experimental y
-                                %justificar esta frase?
-% Es el mismo código - JJ
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
+when writing code and leverages 
+expertise acquired in the library itself.
 
@@
 
-As far as we know, this paper presents one of the few experiments that
-uses computational
-resources that are as dissimilar as smartphones and powerful laptops
-or desktop computers in a research center. The methodology used to
-gather resources and the algorithms used will be described next, as
-well as the results obtained with the setup.
-
 On the other hand, measuring performance of the resulting metacomputer
@@
 computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
-volunteer computing, number of users and the time spent in the
-computation in browser-based volunteer computing experiments, have only been studied in a limited way in
+volunteer computing, number of users and the time spent by every one in the
+computation in browser-based volunteer computing experiments, have
+only been studied in a limited way in 
 \cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
 run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
+\cite{javadi2009mining} found out that the Weibull, log-normal and
+Gamma distribution 
 modeled quite well the availability of resources in several clusters
@@
 
+As far as we know, this paper presents one of the few experiments that
+uses computational resources that are as dissimilar as smartphones and
+powerful laptops 
+or desktop computers in a research center. The methodology used to
+gather resources and the algorithms used will be described next, as
+well as the results obtained with the setup.
+
 %---------------------------------------------------------------
@@
 
-% [Pedro] desde el abstract no se había nombrado "NodIO".
-% Creo que en este párrafo debería introducirse, principalmente el nombre.
-% Añado una frase presentándolo:
-% [Mario] Nota: se menciona también en el segundo párrafo de la
-% Introducción.
-% OK [JJ]
-In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based
-volunteer computing system, whose architecture
-has been developed using JavaScript on both the client and server sides.
+In this sense, in this paper we propose {\sf NodIO}, a cloud or metal
+based 
+volunteer evolutionary computing system based on the {\sf NodEO}
+library, whose architecture 
+has been developed using JavaScript on the client as well as the
+server sides. 
 All parts of the system are free and available with a free
@@
 Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
-\item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces
-  requests for storing and retrieving information (the CRUD cycle:
+\item A REST server, that is, a server that includes several {\em
+  routes} %remote requests suena raro, propongo algo como: remote
+          %procedures o web interfaces [jj] changed to routes, its
+          %proper name.
+  that can be called for storing and retrieving information (the CRUD cycle:
   create, request, update and delete) from the server. The information
@@
   data store \cite{jj:idc:lowcost}.
-  %The information is of two kinds: Problem Based y ¿Cual otra?
-  %¿Logging? No se entiende bien. ¿Podría ser problem domain?
-% info. Añadido [JJ]
   The server has the capability to
@@
 Figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or chart.js have
+algorithm behavior. Different web technologies, such as JQuery or {\tt
+  chart.js} have
 been used to build the elements of the system.
-%FERGU: Mario, explica en el texto de abajo qué es winston en la figura también.
-%Vamos, que todo lo que salga en la figura tiene que ir en el texto en alguna parte :)
-% [Mario] Ok voy a explicar la figura. Quería saber primero su opinión tal vez se les hacía mucho detalle.
-% Creo que esta figura va mejor en la sección donde hablamos de
-% webworkers. La anterior era más adecuada como visión general. [JJ]
-% [Mario] Ok
 \begin{figure}[!t]
@@
 First, we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO \cite{nodeo2014}, the basic JavaScript library we have used as base to build {\sf NodIO}. This experiment tries to
+evolutionary algorithm in a desktop client written using NodEO
+\cite{nodeo2014}, the basic JavaScript library we have used as base to
+build {\sf NodIO}, the framework. This experiment tries to 
 find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$ and a population size equal 512. The algorithm was run until the solution (a string with all
+b=2, z=3$ and a population size equal to 512. The algorithm in a
+single island was run until the solution (a string with all
 ones) was found or five million evaluations had been performed. It
@@
 These two experiments establish first a baseline result and show that
-it depends on the population. The volunteer computing experiments that we
+time to success depends on the population, with bigger population
+contributing more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+experiments that we 
 %FERGU: explicar un poco más lo de "depends on the population"?, por ejemplo añadiendo
@@
 %find the solution \cite{} o algo parecido.
+%hecho [JJ]
 will describe next do not and can not have the same conditions, but
@@
 %
-The table shows that every run included more than 50 experiments.
-The %FERGU: No debería ser mejor al revés?
-Cada experimento está compuesto de varias runs.
-De todas maneras arriba lo has definido así, así que si no quieres no lo cambies.
-Pero me parece más correcto que un experimento tenga runs, en vez de al revés
+The table shows that every run included more than 50 experiments. The
+%FERGU: No debería ser mejor al revés? Cada experimento está compuesto
+%de varias runs. De todas maneras arriba lo has definido así, así que
+%si no quieres no lo cambies. Pero me parece más correcto que un
+%experimento tenga runs, en vez de al revés [jj] Pues no lo sé. Habría
+%que cambiarlo todo. Yo creo que con que sea consisente...
 number of different IPs intervening in them varied from more than one
@@
 in our very own computer. And the results is that although it will
-happen, one out of roughly 5 or roughly 20, depending on what baseline %FERGU: one->1?
+happen, 1 out of roughly 5 or roughly 20, depending on what baseline
+%FERGU: one->1? OK [JJ]
 experiment you choose, it will not happen always or even a majority of
@@
 the evolutionary algorithm start again. However, it puts a probably
-unnecessary cap on the time, or number of experiments, that a user %Fergu: user no tiene "an", me lo dijo paloma. es "yuser", no "iuser", por eso no lo tiene :P
+unnecessary cap on the time, or number of experiments, that a user
+%Fergu: user no tiene "an", me lo dijo paloma. es "yuser", no "iuser",
+%por eso no lo tiene :P 
 will voluntarily perform.
@@
   logarithmic scale. Dot color is related to the number of IPs
-  participating in the experiment. From left to right, experiments 4/4, 4/24 and 7/31.}
+  participating in the experiment. From left to right, experiments
+  4/4, 4/24 and 7/31.} 
 \label{fig:zipf:os}
@@
 We will have to analyze experiment data a bit further to find out why
-this happens and also if there is some pattern in the three sets of
+this happens and also if there are some patterns in the three sets of
 experiments. An interesting question to ask, for instance, is if
@@
 have not found this convenient since there are several other issues
-that will be solved. %FERGU: and that will be explained next? Algún ejemplo?
+that have to be solved, as it will be shown next. %FERGU: and that
+                                %will be explained next? Algún
+                                %ejemplo? 
+% A ver si así - JJ
 
@@
 \caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
-  Extreme Value distribution (in red). From left to right, experiments 4/4, 4/24 and 7/31.}
+  Extreme Value distribution (in red). From left to right, experiments
+  4/4, 4/24 and 7/31.} 
 \label{fig:puts:os}
@@
 issue led us to create the next version of the framework, which is
-presented next. %Fergu: esto mola, métele más caña en otras secciones para que se vea que es una conclusión util
+presented next. %Fergu: esto mola, métele más caña en otras secciones
+                %para que se vea que es una conclusión util 
+% Está en las conclusiones - JJ
 
@@
 window.
-\item Several instances of NodEO can be executed in parallel in a single web
-page. %FERGU: NodEO o NodIO? (es NodEO la instancia, creo pero por si acaso :P)
+\item Several evolutionary algorithms can be executed in parallel in a single web
+page. %FERGU: NodEO o NodIO? (es NodEO la instancia, creo pero por si
+      %acaso :P)
+% Mejor evolutionary algorithms - JJ
 \item Many implementations of the Web Worker API can use multiple CPUs for
@@
 \begin{itemize}
-\item {\em nodejs} This is the server side process, implemented in the nodejs %FERGU: creo que oficialmente es Node.js, homogeneizar en el paper
-framework. Web request handlers are implemented using the Express web
-framework. For keeping a log of the execution the Winston library is used, and
-finally the NodeIO framework is responsible for keeping the shared pool.
+\item {\em nodejs} This is the server side process which is basically
+  the same as used previously, except for some minor log-related
+  changes. 
 \item {\em Client main thread}  This process handles the user interaction
@@
 
+In general, this second version of the {\sf NodIO} framework and the
+single experiment performed prove that it can be the foundation for a
+high-server-performance distributed evolutionary computation platform,
+providing reasonable algorithmic performance and being, in general,
+easy to use with a straightforward modification of the fitness
+function.
+
 %---------------------------------------------------------------
@@
 
-% The conclusion goes here.
+In this paper a client-server architecture for volunteer and distributed
+evolutionary algorithms that uses the browser, termed {\sf NodIO} and
+based on the {\sf NodEO} evolutionary algorithm library has been evaluated.
 
-In this paper a client-server architecture for distributed evolutionary algorithms, termed {\sf NodIO}, for volunteer computing has been presented.
+In order to establish a baseline performance, the evolutionary
+algorithm was run in a desktop client-program written in JavaScript
+using NodEO to solve the 40-trap function. The first experiments with
+{\sf NodIO} proved that, although obtaining better performance than de
+baseline was possible, it did not happen, mainly because of the
+difficulty in carrying over volunteers to other experiments when the
+one they were participating on was finished, shown in the low
+correlation between the number of IPs in successive experiments. This
+also resulted in a low number of generations allotted by users.
 
-In order to establish a baseline performance, the evolutionary algorithm was run in a desktop client-program written in JavaScript using NodEO to solve the 40-trap function.
-In this first experiment, it could be observed how many users donated a few cycles and just a few donated a high amount of CPU to the experiment.
-
-As a second experiment, and trying to improve the number of cycles per user, the system architecture was modified.
-In this case, Web Workers were used so that several clients per browers could be executed.
-As a result, in this case a higher number of contributions per volunteer could be observed, although as in the previous experiments a pattern of a few users with many contributions and many with few contributions was obtained.
+In the second implementation, {\sf NodIO$^2$}, Web Workers
+were used so that several clients per browser could be executed and
+they could run in the background, so that the the number of
+generations per user could be improved. The number of generations per
+user increased to the point that baseline performance was overcome in
+most cases. The architectural changes led to a high correlation
+between successive experiments, so that, even as the median number of
+volunteers per experiment decreased, they were probably contributing
+to the experiment at the same time, so that our objective was
+achieved. In general, one of the first conclusion for this paper is
+that the browser technologies should be used to its full extent so
+that the user time is leveraged to its full potential, and this was
+achieved with {\sf NodIO$^2$}.
 

### 2015-08-10 volunteer-clouds-ieee-ci-si.tex
A ver qué os parecen las conclusiones
Miradlas @fergunet @mariosky @pacastillo. También conviene revisar a fondo. Voy a intentar mandarlo mañana o pasado.

@@
 
+The second objective of this paper was to model the user behavior in a
+first attempt to try and predict performance. As should be expected,
+the model depends on the implementation, with contributions following
+a General Extreme Values distribution in the case of {\sf NodIO} and a
+Weibull distribution for {\sf NodIO$^2$}. These distributions are
+close enough to each other and, in the second case, reflect the fact
+that all time spent in the page is actually devoted to computing,
+which is why the time spent (represented by the number of
+contributions to the pool) follows a model quite similar to that found
+for games or other online activities. The reverse might be true: if we
+want to have returning users for the experiments, it is probable that
+we should {\em gamify} the experience so that once they've done it
+once, they might do it more times. In the spirit of Open Science, this
+gamification might involve computing in real time data such as the one
+presented in this paper and showing it in the same page.
+
+In general, linking and finding correlations between user choices and
+performance is an interesting avenue to explore in the future. Even if
+the three previous experiments were published in a similar, way, one
+obtained up to 5 times more total cycles  that the one with the least
+number of cycles. It is also essential to obtain volunteers as
+simultaneously as possible, so it is possible that the features of the
+social network in terms of real-time use will also play a big
+role. Even as it is difficult to create controlled experiments in this
+area, it is an interesting challenge to explore in the future.
+
+Although we think that, in general, the results presented in this
+paper are independent of the problem chosen, it is true that even if
+it a difficult problem for evolutionary algorithms, it takes a few
+seconds with the right settings to solve. This makes it difficult also
+for this kind of asynchronous framework, but it would be interesting
+to check the performance for more computationally demanding
+tasks. Besides, the two versions of the framework include as an
+algorithmc variant using random population size. We do not think this
+has had a big influence on the result, but it would be interesting to
+measure exactly what this influence has been. Adding a bit of more
+control to the user might contribute also to gamification. Right now
+the user has only two webworkers. It is a matter of a single click to
+open more tabs in the browser, giving more webworkers, but it would be
+interesting to put everything in a single page and under our control,
+to check how often this happens and under which conditions.
+
+Finally, the implementation needs some refinement in terms of
+programming and also ease of use. Tools such as Yeoman might be used
+to create a generator in which the user just has to create a fitness
+function, with the rest of the framework wrapped around
+automatically. All this, as the data used for this paper and the paper
+itself, will be published with a free license in GitHub.
+
 %---------------------------------------------------------------

### 2015-08-10 volunteer-clouds-ieee-ci-si.tex
Fixing a few typos

@@
 browser. As such, it can be run in any computer that loads the web
-page where the client is embedded, including many over which%[Mario]Quitar el possibly?   [JJ] Done
+page where the client is embedded, including many over which %[Mario]Quitar el possibly?   [JJ] Done
 the scientist has no control.
@@
 JavaScript, however,
-\cite{flanagan2006javascript} was introduced in 1997 as a
-browser-embedded language and has, since then, become a set of standards
+\cite{flanagan2006javascript} 
+%was introduced in 1997 as a 
+%browser-embedded language and has, since then,
+%Ya se había dicho  
+has become a set of standards
 \cite{ECMA-262} that include the language and its components. Since
@@
 red and blue arrows respectively. There are also messages between workers and
-the main thread in locally in the client, these are shown as thin solid arrows.
+the main thread locally in the client, these are shown as thin solid arrows.
 There are three processes involved:
 \begin{itemize}
-\item {\em nodejs} This is the server side process which is basically
+\item {\em node.js} This is the server side process which is basically
   the same as used previously, except for some minor log-related
@@
 \item {\em Client main thread}  This process handles the user interaction
-using the JQuery framework. The Chart.js library is used to display a chart of
+using the JQuery framework. The chart.js library is used to display a chart of
 the best fitness in the population over time. The main script also creates
@@
 paper are independent of the problem chosen, it is true that even if
-it a difficult problem for evolutionary algorithms, it takes a few
+it is a difficult problem for evolutionary algorithms, it takes a few
 seconds with the right settings to solve. This makes it difficult also

### 2015-08-11 volunteer-clouds-ieee-ci-si.tex
Spell check and some answers to questions

@@
 installed the required virtual machine plugin first, but embedding a
-ditributed computation client in the browser has the advantage of
+distributed computation client in the browser has the advantage of
 providing a single and universal way of accessing resources via an
@@
 gives software developers certain advantages in the cloud
-paradigm, since is a first-classe citizen
+paradigm, since is a first-class citizen
 in most Platforms as a Service (PaaS) products as most support the {\tt node.js}
@@
 access to the Document Object Model (DOM) or JQuery objects in the main thread.
-Asynchronous HTTP requests to the server are implemented using the XHR object. %FERGU: Qué es el objeto XHR? Alguna referencia/justificación de por qué se usa?
+Asynchronous HTTP requests to the server are implemented using the
+XMLHttpRequest (XHR) object. %FERGU: Qué es el objeto XHR? Alguna
+                             %referencia/justificación de por qué se
+                             %usa? 
 \end{itemize}
@@
 \begin{enumerate}
-\item First the {\sf NodIO} process is started in nodejs. As part of the
+\item First the {\sf NodIO} process is started in NodeJS. As part of the
 initialization procedure the number of experiment is initialized and a log
@@
 loop, that worker is reinitialized. The worker process is not ended, to
-avoid the cost, only the parameters and population is reset and a new UIID
+avoid the cost, only the parameters and population is reset and a new UUID
 assigned.
@@
 \caption{Time employed in every experiment vs.
-number of IPs (in abscises).
+number of IPs (in abscissas).
 The red dot and line show the average and standard deviation, the shade of blue the number of experiments with the same number of IPs.
@@
 user just needs to let it run, without that phase change and thus a
-more {\em equalitarian} distribution of contributions, which rather
+more {\em egalitarian} distribution of contributions, which rather
 corresponds to a Weibull function, as shown in Figure
@@
 tasks. Besides, the two versions of the framework include as an
-algorithmc variant using random population size. We do not think this
+algorithmic variant using random population size. We do not think this
 has had a big influence on the result, but it would be interesting to
@@
 control to the user might contribute also to gamification. Right now
-the user has only two webworkers. It is a matter of a single click to
-open more tabs in the browser, giving more webworkers, but it would be
+the user has only two WebWorkers. It is a matter of a single click to
+open more tabs in the browser, giving more WebWorkers, but it would be
 interesting to put everything in a single page and under our control,

### 2015-08-11 volunteer-clouds-ieee-ci-si.tex
Consistent library names

@@
 paradigm, since is a first-class citizen
-in most Platforms as a Service (PaaS) products as most support the {\tt node.js}
+in most Platforms as a Service (PaaS) products as most support the {\tt Node.js}
 framework \cite{wood13:nodejs:paas}. Many web services
@@
 algorithm behavior. Different web technologies, such as JQuery or {\tt
-  chart.js} have
+  Chart.js} have
 been used to build the elements of the system.
@@
 \item {\em Scalability} is provided via the use of a lightweight and
-  high-performance, single-threaded, server based in NodeJS and
+  high-performance, single-threaded, server based in Node.js and
   Express.js. Although this single server is a bottleneck since it
@@
 \caption{Comparison of time to solution for the baseline system, using a {\tt
-    node.js} client, for different population sizes. Only
+    Node.js} client, for different population sizes. Only
 runs in which solution has been found are used for this graph.}
@@
 \begin{itemize}
-\item {\em node.js} This is the server side process which is basically
+\item {\em Node.js} This is the server side process which is basically
   the same as used previously, except for some minor log-related
@@
 \item {\em Client main thread}  This process handles the user interaction
-using the JQuery framework. The chart.js library is used to display a chart of
+using the JQuery framework. The Chart.js library is used to display a chart of
 the best fitness in the population over time. The main script also creates
@@
 \begin{enumerate}
-\item First the {\sf NodIO} process is started in NodeJS. As part of the
+\item First the {\sf NodIO} process is started in Node.js. As part of the
 initialization procedure the number of experiment is initialized and a log

### 2015-08-11 volunteer-clouds-ieee-ci-si.tex
Minor fixes closes #21

--- 2015-08-11 volunteer-clouds-ieee-ci-si.tex
@@
 \includegraphics[width=0.49\linewidth]{img/weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model. }  
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
 \label{fig:gev:w2}
@@
 %
-It is also interesting to measure the scaling properties in this
+We are also interested in measuring the scaling properties in this
 model, that is, the relationship between the number of IPs

### 2015-08-11 volunteer-clouds-ieee-ci-si.tex
Conclusion written
And closes #19

@@
 
-In this paper a client-server architecture for volunteer and distributed
+In this paper two versions of a client-server architecture for volunteer and distributed
 evolutionary algorithms that uses the browser, termed {\sf NodIO} and
@@
 automatically. All this, as the data used for this paper and the paper
-itself, will be published with a free license in GitHub.
+itself, has been published with a free license in GitHub at
+\url{https://github.com/JJ/modeling-volunteer-computing}.
 

### 2015-08-12 volunteer-clouds-ieee-ci-si.tex
Few corrections

@@
 server in order to receive the worker implementation script.
-Each worker instance executes one or more evolutionary algorithms
-{\em islands}. Each starts with random population and it is assigned a
+Each worker instance executes over time one or more evolutionary algorithms
+{\em islands}. Each starts with a random population and it is assigned a
 universally unique  identifier (UUID). The UUID is included in each of the HTTP
@@
 a message is sent to the worker to start the algorithm.
-\item The worker first creates the population following with the previously
+\item The worker first creates the population following the previously
 specified parameters. Once the population has been initialized, the EA
@@
 display the current state of the island (described next). These messages are
-all asynchronous, so even before a message is received by target the EA loop of
-the worker could continue. If the best fitness is a global best the best
+all asynchronous, so even before a message is received by the target, the EA loop of
+the worker continues. If the best fitness is a global best the best
 chromosome is sent to the server and the loop ends. A message is sent to the
@@
 loop, that worker is reinitialized. The worker process is not ended, to
-avoid the cost, only the parameters and population is reset and a new UUID
+avoid the start-up cost, only the parameters and population are reset and a new UUID
 assigned.
@@
 foreground. This means that volunteers can leave the experiment
-running for as long as they want and they will be contributing to
+running for as long as they want and they will be contributing %borrado un to
 experiment after experiment, unlike the previous version, where they
@@
 using NodEO to solve the 40-trap function. The first experiments with
-{\sf NodIO} proved that, although obtaining better performance than de
+{\sf NodIO} proved that, although obtaining better performance than the
 baseline was possible, it did not happen, mainly because of the

### 2015-08-12 volunteer-clouds-ieee-ci-si.tex
homogeneizar cómo escribimos Web Workers

@@
   stopping when solution is found or, as we have done in this paper,
-  using WebWorkers, which run in the background even if the webpage or
+  using Web Workers, which run in the background even if the webpage or
   browser tab with the experiment is not currently on focus.
@@
 control to the user might contribute also to gamification. Right now
-the user has only two WebWorkers. It is a matter of a single click to
-open more tabs in the browser, giving more WebWorkers, but it would be
+the user has only two Web Workers. It is a matter of a single click to
+open more tabs in the browser, giving more Web Workers, but it would be
 interesting to put everything in a single page and under our control,

### 2015-08-12 volunteer-clouds-ieee-ci-si.tex
Revising final version
Closes #22. Thanks all for your work, and this is the final version unless you find some major bug.

@@
 distributed evolutionary computation software frameworks available on
-the market, usually as open source systems \cite{Parejo12Survey}. Most of these systems have %FERGU: añadida referencia
+the market, usually as open source systems \cite{Parejo12Survey}. Most of these systems have 
 two features: they are {\em desktop} applications, that is, they must
@@
 runs in two tiers, a server and a client, the latter  can be possibly
-run on a  % [Pedro] Both the server and the client in browsers?
-          % different browsers as I understand? [JJ]  No, server in
-          % server, client in browser, although there's nothing that
-          % can't be done in a browser.
-browser. As such, it can be run in any computer that loads the web
-page where the client is embedded, including many over which %[Mario]Quitar el possibly?   [JJ] Done
+run on a browser. As such, it can be run in any computer that loads the web
+page where the client is embedded, including many over which 
 the scientist has no control.
@@
 Some time ago, our group faced the problem of diminishing funds for
-buying new     % [Pedro] This sentence is so so long. I would split it
-               % in two sentences. [JJ] Done.
-hardware. This was aggravated by the increasing maintenance costs and
+buying new hardware. This was aggravated by the increasing maintenance costs and
 extended downtime resulting from the continuous failures of existing
@@
 
-This kind of systems must solve a series of issues before even starting to %FERGU: cambiado el "these systems" porque parecía que hacía referencia a "los anteriores"
+This kind of systems must solve a series of issues before even starting to 
 measure their performance; these issues are related to security as well as
@@
 state of the art (Section \ref{sec:soa}) in web-based distributed
-computation systems along with attempts to predict and model its behavior. %FERGU: El IEEE es americano
+computation systems along with attempts to predict and model its behavior. 
 Then, a detailed description of the proposed system is presented
@@
 adoption of the HTML5 standard \cite{anthes2012html5}.
-JavaScript, however,
-\cite{flanagan2006javascript} 
-%was introduced in 1997 as a 
-%browser-embedded language and has, since then,
-%Ya se había dicho  
-has become a set of standards
+JavaScript \cite{flanagan2006javascript}, however, has become a set of standards
 \cite{ECMA-262} that include the language and its components. Since
@@
 \item A REST server, that is, a server that includes several {\em
-  routes} %remote requests suena raro, propongo algo como: remote
-          %procedures o web interfaces [jj] changed to routes, its
-          %proper name.
+  routes} 
   that can be called for storing and retrieving information (the CRUD cycle:
@@
 seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. %Fergu: dejar claro que es solo una isla
+are shown in Figure \ref{fig:baseline}.
 
@@
 contributing more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
-experiments that we 
-%FERGU: explicar un poco más lo de "depends on the population"?, por ejemplo añadiendo
-%as this is a deceptive problem, and low levels or diversity may not help to
-%find the solution \cite{} o algo parecido.
-%hecho [JJ]
-will describe next do not and can not have the same conditions, but
+experiments that we will describe next do not and can not have the
+same conditions, but  
 the baseline is that if they eventually take longer than a basic
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which this %Ya se defnió antes PaaS
+PaaS, which provides a free tier within which this 
 experiment could be performed. That way, and as required, the hardware
@@
 The table shows that every run included more than 50 experiments. The
-%FERGU: No debería ser mejor al revés? Cada experimento está compuesto
-%de varias runs. De todas maneras arriba lo has definido así, así que
-%si no quieres no lo cambies. Pero me parece más correcto que un
-%experimento tenga runs, en vez de al revés [jj] Pues no lo sé. Habría
-%que cambiarlo todo. Yo creo que con que sea consisente...
 number of different IPs intervening in them varied from more than one
@@
 A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}. This table shows the median number of IPs
-intervening in each experiment,  median of the time needed
+\ref{tab:summary:os}, which shows the median number of IPs
+intervening in each experiment,  median time needed
 to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
@@
 happen, 1 out of roughly 5 or roughly 20, depending on what baseline
-%FERGU: one->1? OK [JJ]
 experiment you choose, it will not happen always or even a majority of
@@
 unnecessary cap on the time, or number of experiments, that a user
-%Fergu: user no tiene "an", me lo dijo paloma. es "yuser", no "iuser",
-%por eso no lo tiene :P 
 will voluntarily perform.
@@
 have not found this convenient since there are several other issues
-that have to be solved, as it will be shown next. %FERGU: and that
-                                %will be explained next? Algún
-                                %ejemplo? 
-% A ver si así - JJ
+that have to be solved, as it will be shown next.
 
@@
 %
-
 It is interesting also to check the distribution of the experiment
@@
 issue led us to create the next version of the framework, which is
-presented next. %Fergu: esto mola, métele más caña en otras secciones
-                %para que se vea que es una conclusión util 
-% Está en las conclusiones - JJ
-
+presented next. 
 \subsection{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
@@
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects. By using Web Workers %FERGU: WebWorker separado o junto? homogeneizar
+basic architecture in several different aspects. By using Web Workers 
 the algorithm kept running on the background even if the {\sf NodIO}
@@
 improvement was included: instead of having a fixed population,
-population size was random and different for each client. %FERGU: mínimo y máximo?
+population size was randomly distributed between 128 and 256 and
+different for each client.
 
@@
 \item Several evolutionary algorithms can be executed in parallel in a single web
-page. %FERGU: NodEO o NodIO? (es NodEO la instancia, creo pero por si
-      %acaso :P)
-% Mejor evolutionary algorithms - JJ
+page. 
 \item Many implementations of the Web Worker API can use multiple CPUs for
@@
 Asynchronous HTTP requests to the server are implemented using the
-XMLHttpRequest (XHR) object. %FERGU: Qué es el objeto XHR? Alguna
-                             %referencia/justificación de por qué se
-                             %usa? 
+XMLHttpRequest (XHR) object.  
 \end{itemize}
@@
 %porque comparamos un mono-isla (el baseline) con un multi-isla. Quizás deberíamos
-%haber usado un multi-isla en el baseline (aunque sea secuencial). Para la próxima.
+%haber usado un multi-isla en el baseline (aunque sea
+%secuencial). Para la próxima.
+% se trata de probar que podemos obtener mejores resultados que en una
+% implementación local básica. Si usamos una paralela local, ya
+% depende del ordenador que tengamos - JJ
 
@@
 foreground. This means that volunteers can leave the experiment
-running for as long as they want and they will be contributing %borrado un to
+running for as long as they want and they will be contributing 
 experiment after experiment, unlike the previous version, where they
@@
 \includegraphics[width=0.9\linewidth]{img/ips-time-ww.png}
-\caption{Time employed in every experiment vs.
-number of IPs (in abscissas).
-The red dot and line show the average and standard deviation, the shade of blue the number of experiments with the same number of IPs.
-}
+\caption{Time employed in every experiment vs. number of IPs (in
+  abscissas). The red dot and line show the average and standard
+  deviation, the shade of blue the number of experiments with the same
+  number of IPs. } 
 \label{fig:ipstime:w2}
@@
 As we did in the previous subsection, we have fitted also the number
-of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown %FERGU: he cambiado GEV por GET (lo de GEV era un error, no?
-vayamos a que sea algo nuevo que no conozco...)
-%cojollos, Fergu, si no conoces no lo cambies. 
+of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
@@
 found for time devoted to games in Chambers et
-al. \cite{chambers2005measurement}.
+al. \cite{chambers2005measurement}. It is important to note that, in
+this case, the number of PUTs made per client does not correspond to a
+uniform amount of computaton, since 100 generations with a random
+population correspond, in each case, to different number of
+oprations; in other words, the same number of PUTs will correspond to
+different number of operations in each case. However, since clients
+are also different and take a different time in each case, it is
+difficult to ascertain how this might have an influence in the
+statistical distribution.
 

### 2015-08-28 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Moved to another thir, main dir for working paper
To make compilation easier
REfs #23

--- 2015-08-12 volunteer-clouds-ieee-ci-si.tex
@@
   \usepackage[pdftex]{graphicx}
-  \graphicspath{{img/}}
+  \graphicspath{{../img/}}
 \else
   \usepackage[dvips]{graphicx}
-  \graphicspath{{img/}}
+  \graphicspath{{../img/}}
 \fi
@@
 \centering
-\includegraphics[width=4in]{img/system.pdf}
+\includegraphics[width=4in]{system.pdf}
 \caption{Description of the proposed system. Clients execute a JS EA
@@
 \centering
-\includegraphics[width=12cm]{img/baseline-times.png}
+\includegraphics[width=12cm]{baseline-times.png}
 \caption{Comparison of time to solution for the baseline system, using a {\tt
@@
 \centering
-\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{img/time-vs-ips-OS-7-31.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
 \caption{Duration of experiments vs. number of different IPs (nodes)
@@
 \centering
-\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{img/time-vs-rank-OS-7-31.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
 \caption{Duration of experiments vs. rank, with $y$ axis in a
@@
 \centering
-\includegraphics[width=0.32\linewidth]{img/puts-openshift-4-4.png}
-\includegraphics[width=0.32\linewidth]{img/puts-openshift-4-24.png}
-\includegraphics[width=0.32\linewidth]{img/puts-openshift-7-31.png}
+\includegraphics[width=0.32\linewidth]{puts-openshift-4-4.png}
+\includegraphics[width=0.32\linewidth]{puts-openshift-4-24.png}
+\includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
 \caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
@@
 \centering
-\includegraphics[width=4in]{img/Algorithm.pdf}
+\includegraphics[width=4in]{Algorithm.pdf}
 \caption{Description of the W$^2$ version of {\sf NodIO}. In this
@@
 \centering
-\includegraphics[width=0.9\linewidth]{img/ips-time-ww.png}
+\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
 \caption{Time employed in every experiment vs. number of IPs (in
@@
 \centering
-\includegraphics[width=0.49\linewidth]{img/gev-fit-ww.png}
-\includegraphics[width=0.49\linewidth]{img/weibull-fit-ww.png}
+\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
+\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
 \caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
@@
 %---------------------------------------------------------------
-% use section* for acknowledgment
 \section*{Acknowledgment}
@@
 GENIL project (CEI-BIOTIC Granada). We would also like to thank the
-anonymous reviewers of previous versions of this paper who have really helped us to improve
+anonymous reviewers of previous versions of this paper who have really
+helped us to improve 
 this paper (and our work) with their suggestions. We are also grateful

### 2015-10-23 data/log2csv.py
script para convertir log a CSV (para Incho)

@@ -1,0 +1,44 @@
+#!/usr/bin/python
+# -*- coding: utf-8 -*-
+
+import ast, sys
+
+ruta_log = sys.argv[1]
+ruta_csv = sys.argv[2]
+
+lista_lineas = list()
+
+#Abrimos el fichero en modo lectura
+fichero = open(ruta_log,"r")
+
+
+
+for linea in fichero:
+
+    lista_lineas.append(ast.literal_eval(linea))
+
+
+
+
+fichero_salida = open(ruta_csv,"w")
+
+elemento = '"chromosome","fitness","IP","date","time"\n'
+fichero_salida.write(elemento)
+
+
+for diccionario in lista_lineas:
+
+    try:
+ 
+        fh1, fh2 = diccionario["timestamp"].split("T")
+        fh = '"' + fh1 + '","' + fh2 + '"'
+
+        elemento = '"' + str(diccionario["chromosome"]) + '","' + str(diccionario["fitness"]) + '","' + str(diccionario["IP"]) + '","' + fh + '"\n'
+        fichero_salida.write(elemento)
+    except:
+        pass
+
+
+
+fichero.close()
+fichero_salida.close()

### 2015-10-23 data/log2csv.py
Controlando los argumentos de entrada

@@ -3,42 +3,48 @@
 
 import ast, sys
 
-ruta_log = sys.argv[1]
-ruta_csv = sys.argv[2]
+def procesa(input_file, output_file):
 
-lista_lineas = list()
+    ruta_log = input_file
+    ruta_csv = output_file
 
-#Abrimos el fichero en modo lectura
-fichero = open(ruta_log,"r")
+    lista_lineas = list()
 
+    #Abrimos el fichero en modo lectura
+    fichero = open(ruta_log,"r")
 
+    for linea in fichero:
 
-for linea in fichero:
-
-    lista_lineas.append(ast.literal_eval(linea))
-
-
+        lista_lineas.append(ast.literal_eval(linea))
 
 
-fichero_salida = open(ruta_csv,"w")
+    fichero_salida = open(ruta_csv,"w")
 
-elemento = '"chromosome","fitness","IP","date","time"\n'
-fichero_salida.write(elemento)
+    elemento = '"chromosome","fitness","IP","date","time"\n'
+    fichero_salida.write(elemento)
 
 
-for diccionario in lista_lineas:
+    for diccionario in lista_lineas:
 
-    try:
+        try:
  
-        fh1, fh2 = diccionario["timestamp"].split("T")
-        fh = '"' + fh1 + '","' + fh2 + '"'
+            fh1, fh2 = diccionario["timestamp"].split("T")
+            fh = '"' + fh1 + '","' + fh2 + '"'
 
-        elemento = '"' + str(diccionario["chromosome"]) + '","' + str(diccionario["fitness"]) + '","' + str(diccionario["IP"]) + '","' + fh + '"\n'
-        fichero_salida.write(elemento)
-    except:
-        pass
+            elemento = '"' + str(diccionario["chromosome"]) + '","' + str(diccionario["fitness"]) + '","' + str(diccionario["IP"]) + '","' + fh + '"\n'
+            fichero_salida.write(elemento)
+        except:
+            pass
 
 
 
-fichero.close()
-fichero_salida.close()
+    fichero.close()
+    fichero_salida.close()
+
+
+if len(sys.argv) > 2:
+
+    procesa(sys.argv[1],sys.argv[2])
+
+else:
+    print "Usage: log2.csv input_file output_file"

### 2015-10-23 data/log2csv.py
Agregados docstrings

@@ -1,28 +1,37 @@
 #!/usr/bin/python
 # -*- coding: utf-8 -*-
+"""
+Script parq convertir los ficheros *.log en *.csv.
+
+Sólo extrae las filas con campo "chromosome", y sólo parte de los campos (Los que incho necesita)
+"""
 
 import ast, sys
 
 def procesa(input_file, output_file):
+"""
+El primer argumento es El archivo *.log de entrada, el segundo el *.csv de salida
+"""
+
 
     ruta_log = input_file
     ruta_csv = output_file
 
     lista_lineas = list()
 
-    #Abrimos el fichero en modo lectura
+    #Abrimos el fichero de entrada en modo lectura
     fichero = open(ruta_log,"r")
 
     for linea in fichero:
 
         lista_lineas.append(ast.literal_eval(linea))
 
-
+    #Abrimos el fichero de salida en modo escritura
     fichero_salida = open(ruta_csv,"w")
 
+    # Cabecera del CSV
     elemento = '"chromosome","fitness","IP","date","time"\n'
     fichero_salida.write(elemento)
-
 
     for diccionario in lista_lineas:
 
@@ -36,11 +45,8 @@ def procesa(input_file, output_file):
         except:
             pass
 
-
-
     fichero.close()
     fichero_salida.close()
-
 
 if len(sys.argv) > 2:
 

### 2015-10-23 data/log2csv.py
Agregada licencia

@@ -1,10 +1,31 @@
 #!/usr/bin/python
 # -*- coding: utf-8 -*-
 """
-Script parq convertir los ficheros *.log en *.csv.
+Script para convertir los ficheros *.log en *.csv.
 
 Sólo extrae las filas con campo "chromosome", y sólo parte de los campos (Los que incho necesita)
 """
+
+
+"""
+    Copyright 2015 psicobyte@gmail.com
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+
+"""
+
+
 
 import ast, sys
 
@@ -12,7 +33,6 @@ def procesa(input_file, output_file):
 """
 El primer argumento es El archivo *.log de entrada, el segundo el *.csv de salida
 """
-
 
     ruta_log = input_file
     ruta_csv = output_file

### 2015-10-23 data/log2csv.py
Solicionado error de indentación

@@ -30,9 +30,9 @@ Sólo extrae las filas con campo "chromosome", y sólo parte de los campos (Los 
 import ast, sys
 
 def procesa(input_file, output_file):
-"""
-El primer argumento es El archivo *.log de entrada, el segundo el *.csv de salida
-"""
+    """
+    El primer argumento es El archivo *.log de entrada, el segundo el *.csv de salida
+    """
 
     ruta_log = input_file
     ruta_csv = output_file

### 2015-10-23 data/log2csv.py
Solicionado error con las comillas en la fecha

@@ -3,7 +3,7 @@
 """
 Script para convertir los ficheros *.log en *.csv.
 
-Sólo extrae las filas con campo "chromosome", y sólo parte de los campos (Los que incho necesita)
+Sólo extrae las filas con los campos "chromosome", fitness","IP" y "timestap"'; y sólo parte de los campos (Los que incho necesita)
 """
 
 
@@ -60,7 +60,7 @@ def procesa(input_file, output_file):
             fh1, fh2 = diccionario["timestamp"].split("T")
             fh = '"' + fh1 + '","' + fh2 + '"'
 
-            elemento = '"' + str(diccionario["chromosome"]) + '","' + str(diccionario["fitness"]) + '","' + str(diccionario["IP"]) + '","' + fh + '"\n'
+            elemento = '"' + str(diccionario["chromosome"]) + '","' + str(diccionario["fitness"]) + '","' + str(diccionario["IP"]) + '",' + fh + '\n'
             fichero_salida.write(elemento)
         except:
             pass

### 2015-10-24 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Change of computers commit
Some corrections

@@
   Potentially Massive Distributed Evolutionary Computation}
+% The word massive in the title is giving us problems.
+% Why not change it to something about sharing resources?.
+% Also lets not forget about volunteers, they can be a resource too
+% and they cannot be replaced by any super computer
+
+%  I propose something like:
+% A Web Based Architecture to Harness Volunteer Computational Resources 
+% for Low Cost Distributed Evolutionary Computation
+
+% Some keywords missing: Cloud, Asynchronous, Pool-Based, Javascript-Based
+% Open Source software.
+% Lets not forget this issue is about software.
 
@@
 
+% Work in progress..
+%Volunteer computing is used routinely in science and industry, but it
+%still has a long way to go if it is going to become a mainstream technology 
+%in the same way the cloud technologies in general have. In one hand, projects 
+%lack special purpose and open source tools with analytics builtin and
+%in the other there are still barriers that keep users from sharing their computing 
+%resources like the need to install software applications or plugins in order 
+%to cooperate. That is why in this paper we present {\sf NodIO}, a client-server
+%architecture for distributed
+%evolutionary algorithms, written entirely in JavaScript, which needs
+%minimal server support to work and is able to support persistent,
+%asynchronous, distributed evolutionary algorithms running in the
+%browser.  The architecture, along with some initial results on the
+%expected performance and scaling
+%behavior is presented. We also use the algorithm data to
+%find out what kind of patterns the user behavior follow and try to fit
+%them to some statistical distribution.
+
 \begin{IEEEkeywords}
@@
 
-There is, then, space for systems that do not have any of these
+There is, then, space for systems that do not make any of these
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is first an application that
-runs in two tiers, a server and a client, the latter  can be possibly
-run on a browser. As such, it can be run in any computer that loads the web
-page where the client is embedded, including many over which 
+provisionally named {\sf NodIO}. {\sf NodIO} is first a two tier client/server 
+application, in which clients can possibly run in a browser. 
+Therefore, it can be run in any web-enabled device, including many over which 
 the scientist has no control.
@@
 in Section \ref{sec:description}, followed by the experiments and 
-obtained results in Section \ref{sec:experiments}.
+results obtained in Section \ref{sec:experiments}.
 Finally, conclusions and future lines of work are presented in Section
@@
 \cite{baratloo1996charlotte}. This system, which implemented a
-distributed shared memory, was inspired by previous work
-on parallel computing systems PVM and MPI and solved most issues by
+distributed shared memory, was inspired by a previous work
+on parallel computing systems PVM and MPI and solved most of the issues by
 using the Java virtual machine embedded in the browser,
-which provides
+which provided
 security to the user, uniformity of computing environment
 independently of the operating system and access to a common data store in the
-server. Since the virtual machine is a sandbox, it is safe to the user
+server. Since the Java virtual machine is a sandbox, it is safe to the user
 and running a compiled bytecode guarantees a certain level of
@@
 Other issues such as scheduling, which are related to adaptive 
- parallelism, are solved on the server. This
+parallelism, were solved on the server. This
 paper was later extended and published in a journal in 1999
@@
 experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to Jet, such as {\em SuperWeb}
+above. Other proposals contemporary to JET, such as {\em SuperWeb}
 \cite{alexandrov1997superweb} are roughly a translation of classical
-distribution techniques, with complicated architectures that include
+distribution techniques, with overly complicated architectures that include
 brokers for resource discovery. It is focused, however, in solving the
 issues involved in browser-based computing and also in making a system
-that is able to work across different classes of computers, mentioning
+that is able to work across different types of computers, mentioning
 SPARCstations and PCs with Windows-95. Interestingly enough, this
 paper incorporates a discussion on the economic model, paving the way
-for the for-profit distributed computing systems that later on
-arose. On the other hand, the potential for volunteer computing using
+for the for-profit distributed computing systems that arose later on. 
+On the other hand, the potential for volunteer computing using
 browsers was realized
@@
 gives software developers certain advantages in the cloud
-paradigm, since is a first-class citizen
-in most Platforms as a Service (PaaS) products as most support the {\tt Node.js}
-framework \cite{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and use JSON (JavaScript Object Notation) for data
-interchange. Another advantage is that
+paradigm, since it is a first-class citizen
+in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most. \cite{wood13:nodejs:paas}. Many web services
+have native JavaScript libraries too, and also use JSON (JavaScript Object Notation) for data interchange. Another advantage is that
 developers can use the same software tools for both client and server
@@
 JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
-According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in
-metaheuristics optimization allows an improvement in communicability,
+According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
+metaheuristics optimization is that it allows an improvement in communicability,
 reproducibility, interoperability, automated assembly, knowledge engineering
@@
   additional links and information on the experiment. This code runs
-  an evolutionary algorithm {\em island} starting with random
+  an evolutionary algorithm {\em island} starting with a random
   population and sending, every 100 generations, the best individual
@@
   these constants, the algorithm can be run in many different ways:
-  stopping when solution is found or, as we have done in this paper,
+  stopping when a solution is found or, as we have done in this paper,
   using Web Workers, which run in the background even if the webpage or
-  browser tab with the experiment is not currently on focus.
+  browser tab having the experiment is not currently on focus.
 \end{enumerate}
@@
   other, communicating only through the server.
-\item Since the algorithm is run on the browser, {\em safety} is
+\item Since the algorithm runs in the browser, {\em safety} is
   achieved through its sandbox mechanisms. The user is thus assured
@@
   make sure that there is a reasonable amount of clients over which
-  the performance achieved is bigger than what you would obtain in
-  your own desktop system. If that is not done, it is a pointless
+  the performance achieved is better than what you would obtain in
+  your own desktop system. If this is not the case, it is a pointless
   academic exercise.
@@
 %
-It is interesting also to check the distribution of the experiment
+It is also interesting to check the distribution of the experiment
 duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
@@
 the algorithm kept running on the background even if the {\sf NodIO}
-tab was not in focus. It also allowed to run several clients per
+tab was not in focus. It also allowed several clients to run per
 browser, as opposed to a single one as we did previously. Another
-improvement involved restart of the client once the solution was
+improvement involved restarting the client once the solution was
 found, so that as long as the user was visiting the page it was
@@
 page. 
-\item Many implementations of the Web Worker API can use multiple CPUs for
-the execution of worker scripts, in particular the execution in multiple CPUs
+\item Many implementations of the Web Worker API can use multiple-core CPUs for
+the execution of worker scripts, in particular the execution in multiple-core CPUs
 was tested in desktop versions of Chrome, FireFox and Safari.

### 2015-10-24 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some more

@@
 \item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure the number of experiment is initialized and a log
+initialization procedure the secuential number of experiment is initialized and a log
 file is created. The shared pool implemented as an array is reset.
@@
 is evaluated. First the chromosome with the best fitness is sent to the shared
-population with a PUT HTTP request. If the best fitness is not a global best,
+population with a PUT HTTP request. If the best fitness is not the global best,
 the EA loop must continue, but first a chromosome is randomly selected from
@@
 ends, the experiment number is incremented and the population array is reset.
-\item When the main thread receives a message the a worker has ended the EA
+\item When the main thread receives a message indicating that a worker has ended the EA
 loop, that worker is reinitialized. The worker process is not ended, to
@@
 generations per user could be improved. The number of generations per
-user increased to the point that baseline performance was overcome in
+user increased to the point that baseline performance was improved in
 most cases. The architectural changes led to a high correlation

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Answer to issue #53
Closes #53

@@
 are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}. Our initial tests indicate that
+results obtained in \cite{agajaj}. Although SETI@home works
+at larger scale in both quantity of volunteers and scope of projects, 
+our initial tests indicate that
 this kind of distribution might be the most appropriate to describe

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Working on #50
Some minor clarifications

@@
 \item The system is {\em heterogeneous} since it does not need any
-  performance, operating system or even browser requirement: anyone
+  performance, operating system or even browser requirement (as long
+  as Javascript is enabled): anyone
   visiting the page, even from mobile devices, can load the algorithm.
+  In case the browser does not support HTML5 Web workers (e.g., 
+  Windows Phone or some version of Android) or their use
+  has been disabled, a basic version of NodIO can be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
   communicate with the rest of the islands.
-\item {\em Adaptiveness} is achieved simply through autonomous
-  operation of every individual island and no synchronization
+\item {\em Adaptiveness} is achieved simply through the autonomous
+  operation of every individual island with out any synchronization
   mechanism. The islands in the system are, in fact, unaware of each
-  other, communicating only through the server.
+  other, communicating only through the server. This architecture 
+  has been tested in other high demand systems with success, but 
+  in the case of NodIO, additional experiments are need to asses 
+  the scalability of the system. 
 \item Since the algorithm runs in the browser, {\em safety} is

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
About volunteer issues #50
Added a paragraph about issues of volunteer systems. I only have th
BOINC site as reference, any others?

@@
 
+As a volunteer based system, there are some issues worth 
+noting:\begin{enumerate}
+\item Volunteers are anonymous, only the IP address sent in 
+each request is known.
+\item As anonymous entities, volunteers are not accountable. 
+If a volunteer  misbehaves in some way the provider cannot 
+prosecute or discipline the volunteer.
+\item Participants must also trust the application provider, 
+regarding invation of privacy, the integrity of their devices 
+and intended application. 
+\end{enumerate}
+
 In this sense, in this paper we propose {\sf NodIO}, a cloud or metal
@@
 license from \url{https://github.com/JJ/splash-volunteer}.
+
 Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
References added #50
I hope I modified the correct file.

@@
 As a volunteer based system, there are some issues worth 
-noting:\begin{enumerate}
+noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
 \item Volunteers are anonymous, only the IP address sent in 
@@
 \end{enumerate}
+In this work security aspects related to sabotage and 
+trust management (e.g. \cite{domingues2007sabotage}) have not been addressed.
 

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Added a comment for #50  Source code is sent to clients
Closes #50
Addressing #53
Closes #53

@@
 If a volunteer  misbehaves in some way the provider cannot 
-prosecute or discipline the volunteer.
+prosecute or discipline the volunteer. In the case of JavaScript,
+this issue is agravated because even if it is obfuscated, sorce code
+and data is sent to clients. 
 \item Participants must also trust the application provider, 
@@
   that can be called for storing and retrieving information (the CRUD cycle:
-  create, request, update and delete) from the server. The information
-  is of two kinds: {\em problem} based, that is, related to the
+  create, request, update and delete) from the server. 
+  A JSON data format is used for the communication between 
+  clients and the server. There are two kinds of information:
+  {\em problem} based, that is, related to the
   evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
@@
   back to the server (via a {\tt PUT} request) and requesting a random
-  individual from the server (via a {\tt GET} request). Apart from
+  individual from the server (via a {\tt GET} request).
+
+  Apart from
   these constants, the algorithm can be run in many different ways:
@@
 \end{itemize}
+
 %

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Closing #43
Closes #43
Issue #37
Correcting; *”Another issue is that some details are described twice,
for instance, the advantages of the use of web workers.”* Closes #37

@@
   stopping when a solution is found or, as we have done in this paper,
-  using Web Workers, which run in the background even if the webpage or
-  browser tab having the experiment is not currently on focus.
+  using Web Workers. 
+  % Repeated
+  %which run in the background even if the webpage or
+  %browser tab having the experiment is not currently on focus.
 \end{enumerate}
@@
 presented next.
-\subsection{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
+
+\section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
 
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects. By using Web Workers 
-the algorithm kept running on the background even if the {\sf NodIO}
-tab was not in focus. It also allowed several clients to run per
-browser, as opposed to a single one as we did previously. Another
-improvement involved restarting the client once the solution was
+basic architecture in several different aspects.  One improvement 
+involved restarting the client once the solution was
 found, so that as long as the user was visiting the page it was
-running an evolutionary algorithm. Finally, a small algorithmic
-improvement was included: instead of having a fixed population,
+running an evolutionary algorithm. Another was a small algorithmic
+improvement: instead of having a fixed population,
 population size was randomly distributed between 128 and 256 and
-different for each client.
+different for each client.  The other improvment was the 
+use the Web Worker API as described next.
 

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changed from colloquial style
Closes #45, but we need to look for more colloquial sentences.

@@
 remote server, not a local server as in the baseline experiment. But
-fair is fair, and we were looking for creating a framework that
-allowed to run evolutionary computation experiments {\em faster} than
+even if that is the case, we were looking for creating a framework that
+allowed us to run evolutionary computation experiments {\em faster} than
 in our very own computer. And the results is that although it will

### 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Added a paragraph for issue #31

@@
 
+%PABLO: He añadido el siguiente párrafo para el issue #31
+There exist other volunteer frameworks written in compiled languages, such as C++ or Java, that can achieve more performance than JavaScript.
+However, one of the advantages of our system is that clients do not need to install anything in their machines, just to open a regular web page.
+Javascript is the only language that can be executed in this way without any installation, besides the default modern browsers, such as Chrome, Safari or Firefox, that are available in almost every desktop or mobile device.
+Other advantage is that the source code can be directly examined by the volunteer, not being compiled as other options (such as Java Applets or ActiveX).
+Also, as we discuss in this paper, it can provide some extra features such as safety or operating system independence.
+
+%PABLO: hay que enlazar con el siguiente párrafo
 This kind of systems must solve a series of issues before even starting to 
@@
 license from \url{https://github.com/JJ/splash-volunteer}.
+
+%PABLO: el siguiente párrafo justifica Javascript
 

### 2015-10-27 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Minor suggestions

--- 2015-10-26 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 
+%PABLO: He añadido el siguiente párrafo para el issue #31
+There exist other volunteer frameworks written in compiled languages, such as C++ or Java, that can achieve more performance than JavaScript.
+However, one of the advantages of our system is that clients do not need to install anything in their machines, they just need to open a regular web page
+% Somewhat redundant 
+% Javascript is the only language that can be executed in this way without any installation, 
+% also removed the default, redundant with next sentence
+in modern browsers, such as Chrome, Safari or Firefox, that are available in almost every desktop or mobile device.
+Other advantage is that the source code can be directly examined by the volunteer, not being compiled as other options (such as Java Applets or ActiveX).
+Also, as we discuss in this paper, it can provide some extra features such as enhanced security or operating system independence.
+
+%PABLO: hay que enlazar con el siguiente párrafo
 This kind of systems must solve a series of issues before even starting to 

### 2015-11-12 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Starting Section

@@
 
+\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigins Function}
+The most common use of JavaScript is in the development of
+asynchronous Web applications. Scripts are embedded in HTML pages and change
+the state of with the Document Object Model (DOM) of the page, by sending or
+retrieveing data asynchronously from a server. For these purpose JavaScript has
+been traditionally implemented as an interpreted language not designed for the
+development of high performance systems. Current JavaScript virtual machines (VMs)
+are closing the gap, for instance the Google V8 is an engine specifically designed
+for fast execution of large JavaScript applications. Also V8 compiles JavaScript
+source code directly into machine code when it is first executed. 
+There are no intermediate byte codes or interpreter.
+[\url{ https://developers.google.com/v8/design#mach_code}].
+JavaScript has gained popularity in server-side development using runtime
+environments such as Node.js and is also used in desktop and mobile applications.
+To further evaluate how a current JavaScript implemantation can be used by a
+computer scientist for the development of complex optimization problems,
+a benchmark function, which was provided by CEC2010 Special Session on
+Large-Scale Global Optimization [], is used in this section.
+
+\begin{equation}
+F_{rastrigin}(x)=\sum\limits_{i=1}^\frac{D} [x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
+\end{equation}
+
+\begin{equation}
+F_{rot_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
+\end{equation}
+
+\begin{equation}
+F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot_rastrigin}[\textbf{z}(F_{(k-1)*m+1}:P_{k*m})] 
+\end{equation}
+
 %---------------------------------------------------------------

### 2015-11-12 img/f15/plot_f15.py
Data and script for plot

@@ -1,0 +1,45 @@
+__author__ = 'mariosky'
+
+from pylab import *
+
+java_file = open("data/java.txt")
+matlab_file = open("data/matlab.txt")
+node_file = open("data/node.txt")
+chrome_file = open("data/chrome.txt")
+firefox_file = open("data/firefox.txt")
+chrome_main_file = open("data/chrome_main_thread.txt")
+firefox_main_file = open("data/firefox_main_thread.txt")
+
+java_times = [float(time) for time in  java_file ]
+matlab_times = [float(time)*1000 for time in  matlab_file ]
+node_times = [float(time) for time in  node_file ]
+chrome_times = [float(time.split(',')[0])   for time in  chrome_file ][:30]
+firefox_times = [float(time.split(',')[0])   for time in  firefox_file ][:30]
+chrome_times_main = [float(time.split(',')[0])   for time in  chrome_main_file ][:30]
+firefox_times_main = [float(time.split(',')[0])   for time in  firefox_main_file ][:30]
+
+print firefox_times_main
+data = [java_times,matlab_times,node_times, chrome_times,chrome_times_main, firefox_times,firefox_times_main]
+
+fig = figure()
+ax1 = fig.add_subplot(111)
+
+bp = plt.boxplot(data)
+
+
+
+ax1.yaxis.grid(True, linestyle='-', which='major', color='lightgrey',
+               alpha=0.5)
+
+# Hide these grid behind plot objects
+ax1.set_axisbelow(True)
+ax1.set_title('Time in miliseconds for 10k evaluations, 30 runs')
+ax1.set_xlabel('Implementation Language')
+ax1.set_ylabel('Time in seconds')
+
+xtickNames = plt.setp(ax1, xticklabels= ["Java","Matlab","node.js","chrome","chrome main","firefox","firefox main"] )
+plt.setp(xtickNames, rotation=30, fontsize=8)
+#plt.setp(xtickNames)
+plt.ylim((0,1600))
+show()
+plt.savefig('../f15_times.eps')

### 2015-11-12 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
F15 Description

@@
 a benchmark function, which was provided by CEC2010 Special Session on
-Large-Scale Global Optimization [], is used in this section.
-
+Large-Scale Global Optimization [] as Function 15 is used in this section. 
+The function is described next.
+
+The basic Rastrigins function is separable[], and is defined as follows;
 \begin{equation}
@@
 \end{equation}
-
+where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
+$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
+is a classical multimodal problem. It is difficult since the number of local
+optima grows exponentially with the increase of dimensionality. To make it 
+nonseparable, an orthogonal matrix is also used for coordinate rotation.
+The rotated Rastrgin’s function is defined as follows:
 \begin{equation}
@@
 \end{equation}
-
+Where \textbf{M}_{is} a $D×D$ orthogonal matrix.
+
+The benchmark  emulates real-world optimization problems that most likely will consist 
+of different groups of parameters with strong dependencies within
+but little interaction between the groups. This issue is reflected in the benchmark
+problems by first randomly dividing the objective variables
+into several groups, each of which contains a number of variables. The parameter $m$ 
+is used to control the number of variables in each group and hence, defining the degree
+of separability:
 \begin{equation}
@@
 \end{equation}
-
+For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, · · · , x_{D})$ is the
+candidate solution, $\textbf{o} = (o_1, o_2, · · · , o_{D})$ is the shifted global optimum,
+$\textbf{z} = \textbf{x} − \textbf{o}$ is the shifted candidate solution and
+finally $\textbf{P}$ is a random permutation of ${1, 2, · · · ,D}$.
 %---------------------------------------------------------------

### 2015-11-13 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Latex corrections

@@
 \begin{equation}
-F_{rastrigin}(x)=\sum\limits_{i=1}^\frac{D} [x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
+F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
 \end{equation}
@@
 \begin{equation}
-F_{rot_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
+F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
 \end{equation}
-Where \textbf{M}_{is} a $D×D$ orthogonal matrix.
-
-The benchmark  emulates real-world optimization problems that most likely will consist 
+Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
+emulates real-world optimization problems that most likely will consist 
 of different groups of parameters with strong dependencies within
@@
 \begin{equation}
-F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot_rastrigin}[\textbf{z}(F_{(k-1)*m+1}:P_{k*m})] 
+F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
 \end{equation}
-For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, · · · , x_{D})$ is the
-candidate solution, $\textbf{o} = (o_1, o_2, · · · , o_{D})$ is the shifted global optimum,
-$\textbf{z} = \textbf{x} − \textbf{o}$ is the shifted candidate solution and
-finally $\textbf{P}$ is a random permutation of ${1, 2, · · · ,D}$.
+For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
+candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
+$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
+finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
 %---------------------------------------------------------------

### 2015-11-13 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Why this function

@@
 The most common use of JavaScript is in the development of
-asynchronous Web applications. Scripts are embedded in HTML pages and change
-the state of with the Document Object Model (DOM) of the page, by sending or
-retrieveing data asynchronously from a server. For these purpose JavaScript has
+asynchronous Web applications. For this purpose JavaScript has
 been traditionally implemented as an interpreted language not designed for the
-development of high performance systems. Current JavaScript virtual machines (VMs)
-are closing the gap, for instance the Google V8 is an engine specifically designed
-for fast execution of large JavaScript applications. Also V8 compiles JavaScript
-source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or interpreter.
+development of high performance systems. On the other hand, current JavaScript 
+virtual machines (VMs) are closing the gap, for instance the Google V8 
+is an engine specifically designed for fast execution of 
+large JavaScript applications. In order to increase performance
+V8 compiles JavaScript source code directly into machine code when it is first executed. 
+There are no intermediate byte codes or an interpreter.
 [\url{ https://developers.google.com/v8/design#mach_code}].
@@
 environments such as Node.js and is also used in desktop and mobile applications.
-To further evaluate how a current JavaScript implemantation can be used by a
-computer scientist for the development of complex optimization problems,
-a benchmark function, which was provided by CEC2010 Special Session on
-Large-Scale Global Optimization [] as Function 15 is used in this section. 
+To further evaluate how a current JavaScript implemantation could be used by a
+computer scientist to develop complex optimization problems,
+a benchmark function, which was provided by the CEC2010 Special Session on
+Large-Scale Global Optimization [] is used in this section. 
 The function is described next.
@@
 finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
+
+This optimization problem was selected because is representative of the kind of
+algorithms, data types and structures employed in large scale optimization problems. 
+The function was implemented in both matlab and Java languages as part of  the
+Test Suite of the companion competition. To give competitors an idea of the
+computationa cost of the challenge, the runtime required for 10,000 function 
+evaluations (FEs) was published for a particular configuration, where the Java
+implementation took 7596ms and the Matlab version 1115ms. The whole experiment had
+as termination condition 3,000,000 FEs. 
 %---------------------------------------------------------------

### 2015-11-13 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Structure of section

@@
 implementation took 7596ms and the Matlab version 1115ms. The whole experiment had
-as termination condition 3,000,000 FEs. 
+3,000,000 FEs as the termination condition. Using the Java implementation as
+a guide, the function described above was implemented to compare the performence against
+both implementations, again using the test of time required for 10,000 FEs. 
+Before discussing the setup and performance results, certain details of
+the implementation in JavaScript are presented, to give readers an idea of 
+how practical and mature is the language for developing complex mathematical functions
+as well as some limitations.
+
+\begin{itemize}
+\item {\em Randomize}
+\item {\em Arrays} 
+\item {\em Timing functions}
+\item {\em Tools} 
+\end{itemize}
+
+The performance of the first test is presented in
+
+\subsubsection{Setup of experiment in NodIO-W$^2$ }
+
+\subsubsection{Results}
+
 %---------------------------------------------------------------

### 2015-11-13 img/f15/plot_f15.py
Figure of 10k FE

@@ -41,5 +41,5 @@ xtickNames = plt.setp(ax1, xticklabels= ["Java","Matlab","node.js","chrome","chr
 plt.setp(xtickNames, rotation=30, fontsize=8)
 #plt.setp(xtickNames)
 plt.ylim((0,1600))
-show()
+#show()
 plt.savefig('../f15_times.eps')

### 2015-11-13 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Figure of 10k FE
Randomizer

@@
 \begin{itemize}
-\item {\em Randomize}
+\item {\em Randomize} Both Malab and Java implementations rely on a 
+Java Randomizer library as a standardized implementation of the 
+randomization of shift vectors and matrices used for the experiments.
+The library included the generation of pseudorandom Gaussian of type
+double and uniformly distributed integers. Generators used a single
+{\tt long} seed. In order to implement the same functionality in
+JavaScript, the library random-js was used, as there are inconsistencies 
+in the implementation of {\tt Math.random()} between engines in the
+number of bits used in the generation and results are non-deterministic. 
+The library random-js, implements a Mersenne Twister algorithm producing 
+consistent results across all JavaScript implementations assuming the same seed.
 \item {\em Arrays} 
@@
 
-The performance of the first test is presented in
+The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
+The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
+using Java(TM) SE Runtime Environment (build 1.8.0_25-b17), 
+MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
+and Firefox version 41.0.2.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{f15_times.png}
+\caption{ Runtime of 10,000 function evaluations (FEs) for $F_{15}$} 
+\label{fig:f15_times}
+\end{figure}
 

### 2015-11-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Local changes

--- 2015-11-13 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
   Potentially Massive Distributed Evolutionary Computation}
-% The word massive in the title is giving us problems.
-% Why not change it to something about sharing resources?.
-% Also lets not forget about volunteers, they can be a resource too
-% and they cannot be replaced by any super computer
-
-%  I propose something like:
-% A Web Based Architecture to Harness Volunteer Computational Resources 
-% for Low Cost Distributed Evolutionary Computation
-
-% Some keywords missing: Cloud, Asynchronous, Pool-Based, Javascript-Based
-% Open Source software.
-% Lets not forget this issue is about software.
 
@@
 
-% Work in progress..
-%Volunteer computing is used routinely in science and industry, but it
-%still has a long way to go if it is going to become a mainstream technology 
-%in the same way the cloud technologies in general have. In one hand, projects 
-%lack special purpose and open source tools with analytics builtin and
-%in the other there are still barriers that keep users from sharing their computing 
-%resources like the need to install software applications or plugins in order 
-%to cooperate. That is why in this paper we present {\sf NodIO}, a client-server
-%architecture for distributed
-%evolutionary algorithms, written entirely in JavaScript, which needs
-%minimal server support to work and is able to support persistent,
-%asynchronous, distributed evolutionary algorithms running in the
-%browser.  The architecture, along with some initial results on the
-%expected performance and scaling
-%behavior is presented. We also use the algorithm data to
-%find out what kind of patterns the user behavior follow and try to fit
-%them to some statistical distribution.
-
 \begin{IEEEkeywords}
@@
 
-There is, then, space for systems that do not make any of these
+There is, then, space for systems that do not have any of these
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is first a two tier client/server 
-application, in which clients can possibly run in a browser. 
-Therefore, it can be run in any web-enabled device, including many over which 
+provisionally named {\sf NodIO}. {\sf NodIO} is first an application that
+runs in two tiers, a server and a client, the latter  can be possibly
+run on a browser. As such, it can be run in any computer that loads the web
+page where the client is embedded, including many over which 
 the scientist has no control.
@@
 
-%PABLO: He añadido el siguiente párrafo para el issue #31
-There exist other volunteer frameworks written in compiled languages, such as C++ or Java, that can achieve more performance than JavaScript.
-However, one of the advantages of our system is that clients do not need to install anything in their machines, they just need to open a regular web page
-% Somewhat redundant 
-% Javascript is the only language that can be executed in this way without any installation, 
-% also removed the default, redundant with next sentence
-in modern browsers, such as Chrome, Safari or Firefox, that are available in almost every desktop or mobile device.
-Other advantage is that the source code can be directly examined by the volunteer, not being compiled as other options (such as Java Applets or ActiveX).
-Also, as we discuss in this paper, it can provide some extra features such as enhanced security or operating system independence.
-
-%PABLO: hay que enlazar con el siguiente párrafo
 This kind of systems must solve a series of issues before even starting to 
@@
 in Section \ref{sec:description}, followed by the experiments and
-results obtained in Section \ref{sec:experiments}.
+obtained results in Section \ref{sec:experiments}.
 Finally, conclusions and future lines of work are presented in Section
@@
 \cite{baratloo1996charlotte}. This system, which implemented a
-distributed shared memory, was inspired by a previous work
-on parallel computing systems PVM and MPI and solved most of the issues by
+distributed shared memory, was inspired by previous work
+on parallel computing systems PVM and MPI and solved most issues by
 using the Java virtual machine embedded in the browser,
-which provided
+which provides
 security to the user, uniformity of computing environment
 independently of the operating system and access to a common data store in the
-server. Since the Java virtual machine is a sandbox, it is safe to the user
+server. Since the virtual machine is a sandbox, it is safe to the user
 and running a compiled bytecode guarantees a certain level of
@@
  Other issues such as scheduling, which are related to adaptive
-parallelism, were solved on the server. This
+ parallelism, are solved on the server. This
 paper was later extended and published in a journal in 1999
@@
 experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to JET, such as {\em SuperWeb}
+above. Other proposals contemporary to Jet, such as {\em SuperWeb}
 \cite{alexandrov1997superweb} are roughly a translation of classical
-distribution techniques, with overly complicated architectures that include
+distribution techniques, with complicated architectures that include
 brokers for resource discovery. It is focused, however, in solving the
 issues involved in browser-based computing and also in making a system
-that is able to work across different types of computers, mentioning
+that is able to work across different classes of computers, mentioning
 SPARCstations and PCs with Windows-95. Interestingly enough, this
 paper incorporates a discussion on the economic model, paving the way
-for the for-profit distributed computing systems that arose later on. 
-On the other hand, the potential for volunteer computing using
+for the for-profit distributed computing systems that later on
+arose. On the other hand, the potential for volunteer computing using
 browsers was realized
@@
 gives software developers certain advantages in the cloud
-paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most. \cite{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and also use JSON (JavaScript Object Notation) for data interchange. Another advantage is that
+paradigm, since is a first-class citizen
+in most Platforms as a Service (PaaS) products as most support the {\tt Node.js}
+framework \cite{wood13:nodejs:paas}. Many web services
+have native JavaScript libraries too, and use JSON (JavaScript Object Notation) for data
+interchange. Another advantage is that
 developers can use the same software tools for both client and server
@@
 JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
-According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
-metaheuristics optimization is that it allows an improvement in communicability,
+According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in
+metaheuristics optimization allows an improvement in communicability,
 reproducibility, interoperability, automated assembly, knowledge engineering
@@
 are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}. Although SETI@home works
-at larger scale in both quantity of volunteers and scope of projects, 
-our initial tests indicate that
+results obtained in \cite{agajaj}. Our initial tests indicate that
 this kind of distribution might be the most appropriate to describe
@@
 whose client is, or can be, embedded in the browser via JavaScript.
-
-As a volunteer based system, there are some issues worth 
-noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
-\item Volunteers are anonymous, only the IP address sent in 
-each request is known.
-\item As anonymous entities, volunteers are not accountable. 
-If a volunteer  misbehaves in some way the provider cannot 
-prosecute or discipline the volunteer. In the case of JavaScript,
-this issue is agravated because even if it is obfuscated, sorce code
-and data is sent to clients. 
-\item Participants must also trust the application provider, 
-regarding invation of privacy, the integrity of their devices 
-and intended application. 
-\end{enumerate}
-In this work security aspects related to sabotage and 
-trust management (e.g. \cite{domingues2007sabotage}) have not been addressed.
 
@@
 license from \url{https://github.com/JJ/splash-volunteer}.
-
-%PABLO: el siguiente párrafo justifica Javascript
-
 Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
@@
   that can be called for storing and retrieving information (the CRUD cycle:
-  create, request, update and delete) from the server. 
-  A JSON data format is used for the communication between 
-  clients and the server. There are two kinds of information:
-  {\em problem} based, that is, related to the
+  create, request, update and delete) from the server. The information
+  is of two kinds: {\em problem} based, that is, related to the
   evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
@@
   additional links and information on the experiment. This code runs
-  an evolutionary algorithm {\em island} starting with a random
+  an evolutionary algorithm {\em island} starting with random
   population and sending, every 100 generations, the best individual
   back to the server (via a {\tt PUT} request) and requesting a random
-  individual from the server (via a {\tt GET} request).
-
-  Apart from
+  individual from the server (via a {\tt GET} request). Apart from
   these constants, the algorithm can be run in many different ways:
-  stopping when a solution is found or, as we have done in this paper,
-  using Web Workers. 
-  % Repeated
-  %which run in the background even if the webpage or
-  %browser tab having the experiment is not currently on focus.
+  stopping when solution is found or, as we have done in this paper,
+  using Web Workers, which run in the background even if the webpage or
+  browser tab with the experiment is not currently on focus.
 \end{enumerate}
@@
 \item The system is {\em heterogeneous} since it does not need any
-  performance, operating system or even browser requirement (as long
-  as Javascript is enabled): anyone
+  performance, operating system or even browser requirement: anyone
   visiting the page, even from mobile devices, can load the algorithm.
-  In case the browser does not support HTML5 Web workers (e.g., 
-  Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of NodIO can be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
   communicate with the rest of the islands.
-\item {\em Adaptiveness} is achieved simply through the autonomous
-  operation of every individual island with out any synchronization
+\item {\em Adaptiveness} is achieved simply through autonomous
+  operation of every individual island and no synchronization
   mechanism. The islands in the system are, in fact, unaware of each
-  other, communicating only through the server. This architecture 
-  has been tested in other high demand systems with success, but 
-  in the case of NodIO, additional experiments are need to asses 
-  the scalability of the system. 
-\item Since the algorithm runs in the browser, {\em safety} is
+  other, communicating only through the server.
+\item Since the algorithm is run on the browser, {\em safety} is
   achieved through its sandbox mechanisms. The user is thus assured
@@
   make sure that there is a reasonable amount of clients over which
-  the performance achieved is better than what you would obtain in
-  your own desktop system. If this is not the case, it is a pointless
+  the performance achieved is bigger than what you would obtain in
+  your own desktop system. If that is not done, it is a pointless
   academic exercise.
@@
 remote server, not a local server as in the baseline experiment. But
-even if that is the case, we were looking for creating a framework that
-allowed us to run evolutionary computation experiments {\em faster} than
+fair is fair, and we were looking for creating a framework that
+allowed to run evolutionary computation experiments {\em faster} than
 in our very own computer. And the results is that although it will
@@
 %
-It is also interesting to check the distribution of the experiment
+It is interesting also to check the distribution of the experiment
 duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
@@
 presented next. 
-
-\section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
+\subsection{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
 
 In order to improve the number of cycles per user, we improved the
-basic architecture in several different aspects.  One improvement 
-involved restarting the client once the solution was
+basic architecture in several different aspects. By using Web Workers 
+the algorithm kept running on the background even if the {\sf NodIO}
+tab was not in focus. It also allowed to run several clients per
+browser, as opposed to a single one as we did previously. Another
+improvement involved restart of the client once the solution was
 found, so that as long as the user was visiting the page it was
-running an evolutionary algorithm. Another was a small algorithmic
-improvement: instead of having a fixed population,
+running an evolutionary algorithm. Finally, a small algorithmic
+improvement was included: instead of having a fixed population,
 population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvment was the 
-use the Web Worker API as described next.
+different for each client.
 
@@
 page. 
-\item Many implementations of the Web Worker API can use multiple-core CPUs for
-the execution of worker scripts, in particular the execution in multiple-core CPUs
+\item Many implementations of the Web Worker API can use multiple CPUs for
+the execution of worker scripts, in particular the execution in multiple CPUs
 was tested in desktop versions of Chrome, FireFox and Safari.
@@
 \end{itemize}
-
 %
@@
 \item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure the secuential number of experiment is initialized and a log
+initialization procedure the number of experiment is initialized and a log
 file is created. The shared pool implemented as an array is reset.
@@
 is evaluated. First the chromosome with the best fitness is sent to the shared
-population with a PUT HTTP request. If the best fitness is not the global best,
+population with a PUT HTTP request. If the best fitness is not a global best,
 the EA loop must continue, but first a chromosome is randomly selected from
@@
 ends, the experiment number is incremented and the population array is reset.
-\item When the main thread receives a message indicating that a worker has ended the EA
+\item When the main thread receives a message the a worker has ended the EA
 loop, that worker is reinitialized. The worker process is not ended, to
@@
 
-\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigins Function}
-The most common use of JavaScript is in the development of
-asynchronous Web applications. For this purpose JavaScript has
-been traditionally implemented as an interpreted language not designed for the
-development of high performance systems. On the other hand, current JavaScript 
-virtual machines (VMs) are closing the gap, for instance the Google V8 
-is an engine specifically designed for fast execution of 
-large JavaScript applications. In order to increase performance
-V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter.
-[\url{ https://developers.google.com/v8/design#mach_code}].
-JavaScript has gained popularity in server-side development using runtime
-environments such as Node.js and is also used in desktop and mobile applications.
-To further evaluate how a current JavaScript implemantation could be used by a
-computer scientist to develop complex optimization problems,
-a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization [] is used in this section. 
-The function is described next.
-
-The basic Rastrigins function is separable[], and is defined as follows;
-\begin{equation}
-F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
-\end{equation}
-where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
-$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
-is a classical multimodal problem. It is difficult since the number of local
-optima grows exponentially with the increase of dimensionality. To make it 
-nonseparable, an orthogonal matrix is also used for coordinate rotation.
-The rotated Rastrgin’s function is defined as follows:
-\begin{equation}
-F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
-\end{equation}
-Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
-emulates real-world optimization problems that most likely will consist 
-of different groups of parameters with strong dependencies within
-but little interaction between the groups. This issue is reflected in the benchmark
-problems by first randomly dividing the objective variables
-into several groups, each of which contains a number of variables. The parameter $m$ 
-is used to control the number of variables in each group and hence, defining the degree
-of separability:
-\begin{equation}
-F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
-\end{equation}
-For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
-candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
-$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
-finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
-
-This optimization problem was selected because is representative of the kind of
-algorithms, data types and structures employed in large scale optimization problems. 
-The function was implemented in both matlab and Java languages as part of  the
-Test Suite of the companion competition. To give competitors an idea of the
-computationa cost of the challenge, the runtime required for 10,000 function 
-evaluations (FEs) was published for a particular configuration, where the Java
-implementation took 7596ms and the Matlab version 1115ms. The whole experiment had
-as termination condition 3,000,000 FEs. 
 %---------------------------------------------------------------
@@
 generations per user could be improved. The number of generations per
-user increased to the point that baseline performance was improved in
+user increased to the point that baseline performance was overcome in
 most cases. The architectural changes led to a high correlation
@@
 
-This work has been supported in part by TIN2011-28627-C04-02 and.
+This work has been supported in part by TIN2011-28627-C04-02 and
 TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),

### 2015-11-15 special-issue-ci/Response to reviewers.tex
Adds boilerplate response

\documentclass[preprint]{elsarticle}
\biboptions{round, numbers}
\usepackage[latin1]{inputenc}
%\usepackage[T1]{fontenc}
%\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{color}
%\usepackage{setspace}
\usepackage{url}
\usepackage[english]{babel}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Volunteers in the Clouds: an Architecture for Low Cost and
  Potentially Massive Distributed Evolutionary Computation: Response to Reviewers\' comments}

\noindent
Dear Sirs,\\

We really appreciate the opportunity you have given us for improving
our work. Following the reviewers' requests and suggestions we have
deeply changed the contents, and  thus, the quality of the paper.

In the next paragraphs you can find these comments and our responses
to every one, along with the changes done.\\

\noindent
Yours sincerely,\\
The authors.

\end{document}

### 2015-11-15 special-issue-ci/Response to reviewers.tex
Adds response to document refs #57

@@
 
+\section{Response to Reviewer \#1}
+
+\begin{quote}
+The main contribution of this paper is the proposal of a Javascript based distributed system intended to
+run evolutionary algorithms according to the volunteer computing scheme. The idea is that user can
+contribute to the computation by using only a browser, without needing to install any plugin nor a Java
+virtual machine.
+The volunteer computing model appeared some years ago and it has been used in many fields. The
+model is effective to solve problems demanding high computational capabilities, and at this point is where
+this paper proposal introduces some concerns. The choice of Javascript code running on browsers has
+some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
+motivation of these kinds of distributed models is to solve complex problems, using Javascript doesnÕt
+appear a right choice. ItÕs true that is a very commonly used programming language, but in the context of
+Web applications, not in high performance ones. Apart from the fact of being an interpreted language
+designed for the Web, it is difficult to think on a computer scientist writing a complex math code in
+Javascrip. If the authors would have used as a case study a complex optimization problem then they
+could have provided numerical data to sustain their claims; however, the selected problem (which should
+be formulated in the paper) seems a rather academic one.
+\end{quote}
+
+\begin{quote}
+The paper organization is not clear. Some comments about this:
+- The title ÒVolunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
+Evolutionary ComputationÓ seems to suggest that the volunteers run in the cloud, but that is not the case:
+the server run􀀀 in the cloud and the volunteers do in the browsers of common PCs and laptops.
+- The paper abstract does not properly summarize the paper contents. It says that ÒThat is why in this
+paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !Ó, but there
+are two versions of the architecture, as stated later in the conclusions: ÒIn this paper two versions of a
+client-server architecture for volunteer and distributed evolutionary !Ó. That second version is neither
+commented in the introduction.
+\end{quote}
+
+\begin{quote}
+presented in Section III, followed by the experiments and obtained results in Section IV.Ó, but Section IV
+does not only contains experimental result but also the full description of the second version of the
+architecture.
+Summarizing the review, the authors should provide a convincing example showing the advantages of
+their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
+by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
+problem.
+\end{quote}
+
+\section{Response to Reviewer \#1}
+
+\begin{quote}
+Contribution:
+This paper describes a client-server architecture for distributed evolutionary algorithms named NodIO.
+The implementation of the architecture is written entirely in JavaScript, in order to support persistent,
+asynchronous, distributed evolutionary algorithms running in the browser.
+Additionally, the paper presents the results of several experiments performed to validate the proposal and
+to measure the performance improvement generated by the massive parallel computing.
+I think that the contribution is quite relevant, since it provides a computation platform that is flexible,
+extensible and potentially massive. Personally, I consider it one of the fresher and most promising
+proposals in the context of metaheuristic optimization frameworks lately. Hence, our global
+recommendation of acceptance with minor changes.
+\end{quote}
+
+\begin{quote}
+However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
+organization and description of the platform, and the presentation and writing style.
+Regarding the experimentation, I miss:
+• a section discussing the threats to validity of the experiments
+• an experiment with a more difficult problem (please use the well-known benchmark problems
+described in literature).
+• A comparison in similar conditions of the two implementations of the architecture (the algorithms
+used in the experiments are slightly different, but this could lead to dramatic changes in
+performance, and consequently, invalidate the comparison performed in the paper). This would
+enable a measurement of the actual improvement in performance obtained by the use of web
+workers (not that generated due to a different algorithm)
+\end{quote}
+
+\begin{quote}
+Regarding the organization of the paper, some important details regarding the improved framework
+design named is described in the experimentation section. I would reorganize the contents to describe the
+improved in detail in section 3, and focus on the experiments, its results and the discussion in section 4.
+Another issue is that some details are described twice, for instance, the advantages of the use of web
+workers. Fixing this issue can reduce the paper size and improve its readability.
+Regarding the presentation:
+• Figures should be readable even in gray-scale, so do not use red and blue to denote different
+elements, use a different shade or texture.
+• The axis labels and legends in most of the figures are unreadable, please increase its font size.
+• There are some typos and phrases that need rewriting.
+\end{quote}
+
+\section{Response to Reviewer \#3}
+
+\section{Response to Reviewer \#4}
+
+\begin{quote}
+This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
+architecture for low cost and “potentially” massive distributed evolutionary algorithms. The main novelty
+turns up from the fact that the architecture is developed in JavaScript.
+However, JavaScript is well-known because of its lack of capability to deal with high performance
+applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
+believe in the possibility of “potentially” massive computation as sketched by the authors.
+The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
+section must be carefully checked and extended in order to fully validate the goodness of the proposed
+architecture. At least, authors should show the behavior of the architecture in a real-world problem
+demanding high computational performance.
+Finally, English must be double-checked in order to correct minor typos such as “in the sense than[that]”
+(in page 2).
+\end{quote}
+
 \end{document}

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Adds third reviewer by @pacastillo

@@
 
-\section{Response to Reviewer \#1}
+\section{Response to Reviewer \#2}
 
@@
 
+\begin{quote}
+1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
+the system called NodIO consists of a REST server (coded in Node.js) and clients performing
+evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
+entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
+automatic restart of computations on the client's side can improve the performance of the system. The
+state of the art is comprehensive and provides good introduction into the subject.
+
+2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
+Software" special issue. Thanks to both development of Internet technology and continuous increase in
+computing power of electronic devices, including mobile devices, the web browser based distributed
+computing can be a cheap alternative for grid and high-performance computing systems that require
+high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
+architecture presented inthe paper can be potentially used for solving many different time consuming
+problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
+solved on a desktop machine, what may not be convincing for a reader looking for an effective
+optimization tool.
+
+3. Correctness: The system works correctly - it is not very complex and the authors provide its source
+code (this is a great plus!). Technical description of the system is also clear, however it lacks some
+details like parameters of the machine used to determine baselines for the experiments or the format of
+messages sent between the server and the clients. The only thing I noticed is that the authors used the
+term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
+CPUs". As far as I know, contemporary web browsers do not support multiple CPUs architecture.
+
+4. Style: Paper is generally written in a clear way. Statements like "But fair is fair," (page 6), sound a bit
+colloquially in a scientific paper and may be omitted. However, my main concern is the readability of the
+graphs. General remark - font used for description of axes should be significantly larger. In Fig. 2 the
+authors can change scale for axe Y, as the box for the population of 1024 individuals is very small.
+What do the authors mean by "times"? Time in seconds? I am not sure if this Figure is necessary at all
+as the description is understandable without it. In Fig.4 all graphs are similar and only the one in the
+middle is analyzed, so maybe others are not necessary? The same is with Fig. 5. What does "x" stand
+for in the description of axe X? (the same is with Fig. 8).
+
+5. Length: As I mentioned above the authors should think of removing some of the unnecessary graphs
+in order to improve readability. They admit themselves that it is hard to conclude from some of the
+graphs (Fig. 5 and Fig. 8). Nevertheless, overall length of the paper is fine.
+
+6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
+distributed computing system, contrary to the description in the main text of the paper. However, it
+lacks the information about the second version of the system called NodIO-W2 based on the Web
+Worker architecture.
+
+Comments to Authors:
+
+1) I like the state of the art which includes comprehensive overview of the most important papers related
+to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
+with its characteristic provided on page 4. Namely:
+
+ - Heterogeneity - it is true that almost all devices can participate in the web browser based
+ distributed computing system, providing that a user did not disable Java Script in its browser.
+ However if the authors plan to use HTML5 WebWorkers only modern web browsers can
+ participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
+ Windows Phone) are also excluded.
+
+ - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
+ implement evolutionary algorithm in that way that it wil automatically adopt to the problem type
+ and the environment in which it is computed, the authors do not mention about how real
+ adaptiveness of the system can be achieved. The same is with scalability. Only 5 different users
+ that permanently participated in the experiments are not enough to assess the scalability of the
+ system, what is a crucial parameter for distributed computing.
+
+ - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
+ code of evolutionary algorithm is sent to the client as a source code in Java Script, so some
+ experienced user may quite simple decode what has to be computed. It can may come into
+ possession of sensitive data or/and alter this code in order to disrupt the computations or make
+ them worhless. The results send to the server may also be captured and/or altered without any
+ problem since the transmission is not secure.
+
+ In my opinion you should at least highlight above issues.
+
+2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
+funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
+maximum. In the experiments you counted how many times evolutionary algorithm was able to find
+global maximum for the problem of size n=40. Initially you have found that if the population in an
+evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
+within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run (page 4, if I understood it correctly).
+Thus, what was the point of using distributed computing to solve such simple problem? Why the
+authors did not perform their experiments for larger values of n, what could justify distributed problem
+solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
+6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
+it is still behind the desktop computer in more than 60% cases (when compared to the second
+baseline). I suggest to perform additional computing experiments for some higher value of n in the
+"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
+to focus the paper on the behavior analysis of volunteers participating in a distributed computing
+system, what may be also interesting an valuable for the potential readers.
+
+3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In
+Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
+meaning probably that IPs from 16 to 29 did not performed any computations. What was the reason
+for that?
+
+For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC
+systems - the number of participants is completely incomparable and also the type of problems
+solved by the systems are of different nature and complexity.
+\end{quote}
+
 \section{Response to Reviewer \#4}

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds third reviewer by @pacastillo

@@
 The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
-using Java(TM) SE Runtime Environment (build 1.8.0_25-b17), 
+using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
 MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Response prepared closes #57

@@
 
-In the next paragraphs you can find these comments and our responses
+In the next sections you can find these comments and our responses
 to every one, along with the changes done.\\
@@
 some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
-motivation of these kinds of distributed models is to solve complex problems, using Javascript doesnÕt
-appear a right choice. ItÕs true that is a very commonly used programming language, but in the context of
+motivation of these kinds of distributed models is to solve complex problems, using Javascript doesn\'{}t
+appear a right choice. It\'{}s true that is a very commonly used programming language, but in the context of
 Web applications, not in high performance ones. Apart from the fact of being an interpreted language
@@
 The paper organization is not clear. Some comments about this:
-- The title ÒVolunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
-Evolutionary ComputationÓ seems to suggest that the volunteers run in the cloud, but that is not the case:
-the server run􀀀 in the cloud and the volunteers do in the browsers of common PCs and laptops.
-- The paper abstract does not properly summarize the paper contents. It says that ÒThat is why in this
-paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !Ó, but there
-are two versions of the architecture, as stated later in the conclusions: ÒIn this paper two versions of a
-client-server architecture for volunteer and distributed evolutionary !Ó. That second version is neither
+- The title "Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
+Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
+the server run in the cloud and the volunteers do in the browsers of common PCs and laptops.
+- The paper abstract does not properly summarize the paper contents. It says that "That is why in this
+paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !", but there
+are two versions of the architecture, as stated later in the conclusions: "In this paper two versions of a
+client-server architecture for volunteer and distributed evolutionary !". That second version is neither
 commented in the introduction.
@@
 \begin{quote}
-presented in Section III, followed by the experiments and obtained results in Section IV.Ó, but Section IV
+presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV
 does not only contains experimental result but also the full description of the second version of the
@@
 Regarding the experimentation, I miss:
-• a section discussing the threats to validity of the experiments
-• an experiment with a more difficult problem (please use the well-known benchmark problems
+-  a section discussing the threats to validity of the experiments
+-  an experiment with a more difficult problem (please use the well-known benchmark problems
 described in literature).
-• A comparison in similar conditions of the two implementations of the architecture (the algorithms
+-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
 used in the experiments are slightly different, but this could lead to dramatic changes in
@@
 Regarding the presentation:
-• Figures should be readable even in gray-scale, so do not use red and blue to denote different
+-  Figures should be readable even in gray-scale, so do not use red and blue to denote different
 elements, use a different shade or texture.
-• The axis labels and legends in most of the figures are unreadable, please increase its font size.
-• There are some typos and phrases that need rewriting.
+-  The axis labels and legends in most of the figures are unreadable, please increase its font size.
+-  There are some typos and phrases that need rewriting.
 \end{quote}
@@
 This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
-architecture for low cost and “potentially” massive distributed evolutionary algorithms. The main novelty
+architecture for low cost and "potentially" massive distributed evolutionary algorithms. The main novelty
 turns up from the fact that the architecture is developed in JavaScript.
@@
 applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
-believe in the possibility of “potentially” massive computation as sketched by the authors.
+believe in the possibility of "potentially" massive computation as sketched by the authors.
 The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
@@
 demanding high computational performance.
-Finally, English must be double-checked in order to correct minor typos such as “in the sense than[that]”
+Finally, English must be double-checked in order to correct minor typos such as "in the sense than[that]"
 (in page 2).

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Response about the use of JavaScript
In the paper and in the response to reviewers. Closes #31

@@
 \end{quote}
+
+We thank the reviewer for the comment. The only reason why JavaScript
+was chosen was because it is the only language that is native to the
+browser, allowing to carry out precisely this kind of experiments
+where the user is invited to click on an URL and start giving CPU to
+it. Although there are many other volunteer frameworks, written in
+other languages, that a priori would offer a higher performance, the
+key is that they would {\em while they are running}, that is, as soon
+as it has been installed, the user has decided to fire it up, and it
+does not affect the performance in the rest of the system, which is a
+concern in this kind of systems. We have performed speed tests on
+JavaScript and other compiled languages resulting that, in fact,
+JavaScript uses a JIT compiler that provides a good performance; we
+have published these tests as an open access technical report. This
+fact, along these explanations, have now been added to the first
+paragraph of the {\em Description of the system} section.
+
+JavaScript is probably the most popular language nowadays, and this
+has inspired the appearance of many compilers, so that the fitness
+function can, in fact, be written in any language and compiled to
+JavaScript, as is shown in this list:
+\url{https://github.com/jashkenas/coffeescript/wiki/list-of-languages-that-compile-to-js}. This
+has also been added to the text.
+
+In a nutshell, what we are saying is that, as is proven in the paper,
+a JavaScript application, running in the browser, can gather in
+seconds dozens of users. That would be, in practice, impossible with
+other kind of native clients, who would have to download and maybe
+compile the application, losing many seconds or minutes for the
+optimization problem that has to be solved.
 

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Response about the use of JavaScript
In the paper and in the response to reviewers. Closes #31

@@
 system based on the browser is simply a client-server architecture
-whose client is, or can be, embedded in the browser via JavaScript.
+whose client is, or can be, embedded in the browser via
+JavaScript. Since JavaScript is  the only language that is present
+across all browsers, the choice was quite clear. Even so, there might
+be some issues with the performance of the language itself; that is
+why we have made a comparison between JavaScript and other languages
+\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
+faster than compiled languages such as Scala and, in any case, provide
+a performance that is comparable to other languages employed usually
+in evolutionary computation and numerical optimization such as Python. Compiled languages such as
+Java or C might be, in fact, faster, but the fact that a single-user
+performance is higher is more than offset by the fact that the number
+of spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets, is several orders of magnitude
+higher, which is the objective in this kind of systems. JavaScript is
+quite a popular language nowadays, providing compilers from a number
+of languages, many of them in their family (like CoffeeScript), but
+many other outside, so researchers can, in fact, write their fitness
+function in Python, Erlang, Scala or even Java, and them compile it to
+JavaScript.
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Explains what we mean by "cloud"
Closes #32

@@
 Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
-the server run in the cloud and the volunteers do in the browsers of common PCs and laptops.
+the server run in the cloud and the volunteers do in the browsers of
+common PCs and laptops.
+\end{quote}
+
+That is correct. We just used {\em cloud} in a relatively informal
+sense because they are available as in a cloud for running the
+experiment. That is, from the point of view of the person running the
+experiment, there is a {\em cloud} out there running your experiments,
+which is actually run by volunteers in their browsers. We went to the
+original aception of ``cloud'' as a ``network'' more than the more
+precise and technical aception used today.
+
+We have added this explanation to the first paragraph in the
+conclusions; we hope this makes our intention clearer, but we can also
+eliminate that part of the title if the reviewer think it is
+necessary.
+
+\begin{quote}
+
 - The paper abstract does not properly summarize the paper contents. It says that "That is why in this
@@
 \begin{quote}
+- According the last paragraph of the introduction, ÒThen, a detailed description of the proposed system is
 presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Explains what we mean by "cloud"
Closes #32
Revising @fergunet comments
Related to issue #31, which he had not mentioned.

@@
 well as the lack of proofs of concept. That is why in this paper we
-present {\sf NodIO}, a client-server architecture for distributed
+present {\sf NodIO}, a framework used to create client-server architectures for distributed
 evolutionary algorithms, written entirely in JavaScript, which needs
@@
 %PABLO: He añadido el siguiente párrafo para el issue #31
-There exist other volunteer frameworks written in compiled languages, such as C++ or Java, that can achieve more performance than JavaScript.
-However, one of the advantages of our system is that clients do not need to install anything in their machines, they just need to open a regular web page
+There exist other volunteer frameworks written in compiled languages,
+such as C++ or Java, that can achieve more performance than
+JavaScript. However, one of the advantages of our system is that
+clients do not need to install anything in their machines, they just
+need to open a regular web page  
 % Somewhat redundant 
@@
 % also removed the default, redundant with next sentence
-in modern browsers, such as Chrome, Safari or Firefox, that are available in almost every desktop or mobile device.
-Other advantage is that the source code can be directly examined by the volunteer, not being compiled as other options (such as Java Applets or ActiveX).
-Also, as we discuss in this paper, it can provide some extra features such as enhanced security or operating system independence.
+in most browsers, such as Chrome, Safari or Firefox, that are
+available in almost every desktop or mobile device. Other advantage is
+that the source code can be directly examined by the volunteer, not
+being compiled as other options (such as Java Applets, Flash scripts or
+ActiveX). Also, as we discuss in this paper, it can provide some extra
+features such as enhanced security or operating system independence.
 
 %PABLO: hay que enlazar con el siguiente párrafo
-This kind of systems must solve a series of issues before even starting to 
-measure their performance; these issues are related to security as well as
+Those desktop clients have also the same issues as any kind of
+volunteer computing system, including our own:  these issues are
+related to security as well as
 safety, which introduces a {\em social} aspect to the design of a
@@
 In this paper two versions of a client-server architecture for volunteer and distributed
-evolutionary algorithms that uses the browser, termed {\sf NodIO} and
-based on the {\sf NodEO} evolutionary algorithm library has been evaluated.
+evolutionary algorithms that uses the browser and generated using {\sf
+  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
+evaluated. Volunteers are {\em in the cloud}, as stated in the title,
+since they are a {\em CPU as a service} for the persons running the
+experiment. In fact, in this paper we have tried to put some figures
+on the real size of that {\em cloud}.
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Explaining what is really NodIO
It's rather a framework. Closes #33

@@
 JavaScript uses a JIT compiler that provides a good performance; we
-have published these tests as an open access technical report. This
-fact, along these explanations, have now been added to the first
-paragraph of the {\em Description of the system} section.
+have published these tests as an open access technical report.
 
@@
 optimization problem that has to be solved.
+
+These facts, along the explanations, have now been added to the first
+paragraph of the {\em Description of the system} section, as well as
+to the introduction.
 
@@
 \begin{quote}
-
 - The paper abstract does not properly summarize the paper contents. It says that "That is why in this
@@
 
+In fact {\sf NodIO} is a framework that can be used to generate
+different client-server architectures. This has been clarified in the
+abstract and conclusions section.
+
 \begin{quote}
-- According the last paragraph of the introduction, ÒThen, a detailed description of the proposed system is
+- According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
 presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Including a better description of the real structure of the paper
This was oversight on our part. Closes #34

@@
 architecture.
+\end{quote}
+
+You are absolutely right. We have corrected that paragraph to include
+the actual contents of the paper.
+
+\begin{quote}
 Summarizing the review, the authors should provide a convincing example showing the advantages of

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Including a better description of the real structure of the paper
This was oversight on our part. Closes #34

@@
 Then, a detailed description of the proposed system is presented
-in Section \ref{sec:description}, followed by the experiments and 
-results obtained in Section \ref{sec:experiments}.
+in Section \ref{sec:description}, followed by the baseline experiments
+in Section \ref{sec:experiments} and experiments using a new client
+architecture that uses web workers in Section \ref{sec:w2}.
 Finally, conclusions and future lines of work are presented in Section
@@
 \section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
+\section{sec:w2}
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Addresses the last of the comments by rev1
Refs #41 but also addresses #34 and #31

@@
 \end{quote}
+
+We have added a new experiment, a more complex and taxing problem
+using a web-worker architecture, in order to prove how this can be
+used in real-life problems. In fact, our architecture {\sf NodIO} is
+not a exclusively browser-based one. Its pool-based architeture allows
+any other kind of client, including local ones, to be incorporated to
+the experiment, with the lightweight overhead of doing asynchronous
+HTTP requests, which being local are indeed negligible.
+
+Our framework, in fact, does not prevent any other kind of
+experiments. If there is a very efficient local algorithm, it only
+needs to be able to do HTTP requests to work alongside a {\em
+  volunteer cloud}. What our framework ensures is that all work done
+by volunteers can be integrated with local, cloud or other kind of
+resources. This explanation has been also added as a paragraph in the
+Introduction to the paper.
+
+We would like to thanks this reviewer for the helpful comments and
+suggestion made. We expect to have addressed them properly, but in any
+case are grateful because it has allowed us to improve the quality of
+our work.
 

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Addresses the last of the comments by rev1
Refs #41 but also addresses #34 and #31

@@
 
-In this sense, in this paper we propose {\sf NodIO}, a cloud or metal
-based 
-volunteer evolutionary computing system based on the {\sf NodEO}
-library, whose architecture 
-has been developed using JavaScript on the client as well as the
-server sides. 
-All parts of the system are free and available with a free
-license from \url{https://github.com/JJ/splash-volunteer}.
+In this sense, in this paper we propose the {\sf NodIO} framework, a
+cloud or bare metal based volunteer evolutionary computing system
+based on the {\sf NodEO} library, whose architecture has been
+developed using JavaScript on the client as well as the server sides.
+All parts of the system are free and available with a free license
+from \url{https://github.com/JJ/splash-volunteer}.
 
@@
   individual from the server (via a {\tt GET} request).
-
-  Apart from
-  these constants, the algorithm can be run in many different ways:
+\end{enumerate}
+
+  This is the invariant part of the framework, but other than that,
+  the algorithm can be run in many different ways: 
   stopping when a solution is found or, as we have done in this paper,
-  using Web Workers. 
-  % Repeated
-  %which run in the background even if the webpage or
-  %browser tab having the experiment is not currently on focus.
-\end{enumerate}
+  using Web Workers. In fact, since it is a pool-based system such as
+  the one described in \cite{LNCS86720702}, any kind of client that
+  works the application programming interface (API) can be used,
+  written in any kind of language. However, in this paper we are going
+  to focus on the dynamics and measurements of the volunteer part of
+  the system.
 
@@
 
-The researcher only has to change a function to solve a different
-problem, in this case the classical Trap function \cite{Ackley1987} has been
-used. JavaScript is a functional language and declaring a different
+The researcher only has to change a function, which can be written in
+JavaScript or in any other language with cross-compilation to
+JavaScript \cite{web:compilersjs} to solve a different
+problem. In this case the classical Trap function \cite{Ackley1987} has been
+used as well as the Rastrigin floating-point optimization
+problem. JavaScript is a functional language and declaring a different 
 function and handing at the creation of the algorithm object, called
@@
 \section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
-\section{sec:w2}
+\label{sec:w2}
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Adding threats to the system
Closes #35

@@
 
+Thanks for your appreciation.
+
 \begin{quote}
@@
 -  a section discussing the threats to validity of the experiments
+
+\end{quote}
+
+We have inserted a paragraph at the end of the section that describes
+the experimental setup describing what are the challenges that this
+kind of voluntary framework must overcome. Another paragraph analyzing
+the security from the point of view of the framework used has been
+also added. A new paragraph on the security challenges has also been
+inserted, third paragraph in this section, to clarify our integral and
+pragmatic approach to security, which so far has worked perfectly.
+
+\begin{quote}
 -  an experiment with a more difficult problem (please use the well-known benchmark problems
 described in literature).
+\end{quote}
+
+A experiment using the well known Rastrigin function has been
+performed and analyzed.
+
+\begin{quote}
 -  A comparison in similar conditions of the two implementations of the architecture (the algorithms
@@
 
+This has been included as the last subsection of the paper.
+
 \begin{quote}
@@
 design named is described in the experimentation section. I would reorganize the contents to describe the
-improved in detail in section 3, and focus on the experiments, its results and the discussion in section 4.
+improved in detail in section 3, and focus on the experiments, its
+results and the discussion in section 4.
+\end{quote}
+
+In fact, reorganizing the paper increased the chance of introducing
+new issues in this new version. We have preferred to keep it the way
+it is, although I thank you for the suggestion
+
+\begin{quote}
 Another issue is that some details are described twice, for instance, the advantages of the use of web
-workers. Fixing this issue can reduce the paper size and improve its readability.
+workers. Fixing this issue can reduce the paper size and improve its
+readability.
+\end{quote}
+
+We have eliminated duplications in this new version. Thanks for
+noticing it.
+
+\begin{quote}
 Regarding the presentation:
@@
 \end{quote}
+
+We have rewritten and revised extensively the paper, and tried to find
+a way (and failed) to make it more readable for reviewers. For the time being and
+during the revision phase, we prefer to keep it that way, but we will
+change it to grayscale if it is needed for the final version.
 

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding threats to the system
Closes #35

@@
 %---------------------------------------------------------------
-\section{Description of the system}
+\section{Description of the framework}
 \label{sec:description}
@@
 \end{enumerate}
-In this work security aspects related to sabotage and 
-trust management (e.g. \cite{domingues2007sabotage}) have not been addressed.
+
+Two kind of threats are possible in this environment; first, since the
+API is open and it is an open source system, it is relatively easy to
+sabotage the system, as indicated in \cite{domingues2007sabotage}, by
+crafting a fake request which, for instance, assigns a fake fitness to
+a particular chromosome. Another two different attacks are also
+possible: a denial of service attack as well as penetration
+testing. We have approached this threat in a pragmatic way: all
+sources, client and server, for the application, are released
+openly. We also release as open access experiments performed
+\cite{DBLP:journals/corr/GuervosG15} and all data is also, after
+cleaning it from personal information, released as open data. This
+builds trust between the user and the scientist, which indeed has an
+impact on the performance of the system, since there is no need to
+include cheating checks or other functions that would degrade the
+performance of the system. This socio-technical approach is, indeed,
+coherent with our whole approach to the study of a volunteer
+distributed system which can only be study approaching it from the
+technical as well as the social, as in social media, angle.
 
@@
   in the browser, which, every 100 generations, sends the best
-  individual and receives a random one from the server. }
+  individual and receives a random one from the server.}
 \label{fig:system}
@@
 problem. Let us see how this system addresses the different challenges
-outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
+outlined in the State of the Art (Section \ref{sec:soa}):
+\begin{itemize}
 \item {\em Scalability} is provided via the use of a lightweight and
@@
 
-This last point is what we will try to prove in the next section.
+There are many different ways to validate a framework that intends to
+address these challenges. Some of them are included by design: it is
+anonymous, since no accounts are needed to participate in the
+experiment, it is safe for the user, since we are using a the browser
+black box, and it is also heterogeneous. We have designed the
+experiments so that they show adaptability to different types of
+experiments and clients, scalability with the number of clients and
+the problem size, and also measured its performance comparing it with
+a baseline configuration, with which, as indicated above, is
+compatible and can be used in a complementary way. Fault tolerance
+will be experimented by dropping the server and check the continuity
+of the experiments.
+
+All these points will be established in the next two sections,
+starting with the baseline, single worker system next.
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Commented on the non-random population
closes #36
Answered concern about figure color
Closes #38

@@
 
-This has been included as the last subsection of the paper.
+This has been included as the last subsection of the paper. The new
+experiment has constant population, instead of using random
+population.
 
@@
 We have rewritten and revised extensively the paper, and tried to find
-a way (and failed) to make it more readable for reviewers. For the time being and
+a way (and failed) to make it more readable for reviewers. In fact,
+they use different shades, to it should mostly work in grayscale. For the time being and
 during the revision phase, we prefer to keep it that way, but we will
-change it to grayscale if it is needed for the final version.
+change it to grayscale if it is required for the final version.
 

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Addresses font size concerns.
Closes #39
Comment on the objectives of the paper.
Refs #41

@@
 change it to grayscale if it is required for the final version. 
+In fact, the labels are not so important, so increasing the font size
+would in fact distract the attention from the figure itself, which is
+the interesting information we want to convey with the graph. Once
+again, we can do it if required for the final version.
 
@@
 optimization tool.
+\end{quote}
 
+It is true that that function can be easily solved with any desktop
+machine, but we can still do that {\em and} use a volunteer
+computation framework, which will add performance to the desktop
+machine. Our intention was more to prove what kind of performance can
+be expected and how this performance is distributed, although we have
+also proved in the last section how the performance achieved by a
+volunteer computer system can, in fact, be better than a desktop, even
+a powerful one.
+
+At any rate, we have also performed a new experiment with the
+Rasrtrigin function, which in fact cannot be solved in a few minutes
+(or seconds, as in the case of Trap) by a desktop system.
+
+\begin{quote}
 3. Correctness: The system works correctly - it is not very complex and the authors provide its source

### 2015-11-16 special-issue-ci/Response to reviewers.tex
Details on baseline computer added.
Closes #42

@@
 volunteer computer system can, in fact, be better than a desktop, even
-a powerful one.
+a powerful one. This has also been clarified in the introduction and
+conclusions.
 
@@
 term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
-CPUs". As far as I know, contemporary web browsers do not support multiple CPUs architecture.
+CPUs". As far as I know, contemporary web browsers do not support
+multiple CPUs architecture.
+\end{quote}
+
+This has been clarified in the new version of the paper, third
+paragraph from the beginning of the section that explains the web
+worker-based architecture. In fact, most browsers can use multi-core
+CPUs running web workers as independent processes in each one of them.
+The base system has been now described in the first paragraph of the
+section on the experiments, it a desktop Intel i7-4770.
+
+\begin{quote}
 

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Details on baseline computer added.
Closes #42

@@
 seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}.
+are shown in Figure \ref{fig:baseline}. The baseline system has these
+characteristics: \begin{verbatim}
+Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
+2014 x86_64 x86_64 x86_64 GNU/Linux
+\end{verbatim}
+with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
 
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which this 
+PaaS, which provide a free tier within which this 
 experiment could be performed. That way, and as required, the hardware

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changed figure and caption
Closes #46
JavaScript considerations

@@
 \includegraphics[width=12cm]{baseline-times.png}
-\caption{Comparison of time to solution for the baseline system, using a {\tt
+\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
     Node.js} client, for different population sizes. Only
 runs in which solution has been found are used for this graph.}
+% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
 \label{fig:baseline}
@@
 consistent results across all JavaScript implementations assuming the same seed.
-\item {\em Arrays} 
-\item {\em Timing functions}
-\item {\em Tools} 
+\item {\em Timing functions} In order to compare the performance of 
+each implementation, an accurate and consistent way of measuring 
+the runtime is needed. In JavaScript the {\tt Date class} is often used 
+to measure execution time, but the maximum resolution is given in miliseconds.
+In order to measure the runtime intervals with higher precission two functions
+where used: for Node.js we used the native {\tt process.hrtime()} function
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array.
+For testing in the browser the {\tt Performance.now()} function was used, this
+function uses floating-point numbers with up to microsecond precision. 
+A drawback found is that the {\tt Performance.now()} function is only implemented
+in Firefox and Chrome browsers. Both functions are independent of the system clock. 
+\item {\em Tools} An important consideration when chosing certain language to 
+to write non-trivial programs is the availability of development tools. When developing
+the experiments, we used a comprehensive set of developer tools available in both
+browsers and desktop. These tools included package managers, debbugers, logs, 
+network monitors, editors and even debugging of the multi-threaded execution of Web Workers.
+A drawback found is that developer tools very from browser to browser, 
+but they all have the same functionality.
+\item {\em Data Types} JavaScript uses floating point numbers with a 
+limited precision of 64 bits that partialy implements the functionality of
+the {\tt StrictMath} library used by the Java implemantation. If more precision
+is needed developers could use {\tt math.js}, an extensive math library with support for matrices, big and complex numbers.
+
 \end{itemize}
 
-The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
+We found that the JavaScript ecosystem is mature enough for developing non-trivial 
+code, but developers have to consider the differences betwwen VMs implementations
+and use certain libraries in order to increase the precision and repeatability of
+algorithms. To test the performance of JavaScript against Java and Matlab
+
 The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
@@
 MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2.
+and Firefox version 41.0.2.  The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
 

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Discussion of F15 experiment

--- 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 virtual machines (VMs) are closing the gap, for instance the Google V8 
-is an engine specifically designed for fast execution of 
+is an engine specifically designed for the fast execution of 
 large JavaScript applications. In order to increase performance
@@
 but little interaction between the groups. This issue is reflected in the benchmark 
-problems by first randomly dividing the objective variables
+by first randomly dividing the objective variables
 into several groups, each of which contains a number of variables. The parameter $m$ 
@@
 This optimization problem was selected because is representative of the kind of
-algorithms, data types and structures employed in large scale optimization problems. 
-The function was implemented in both matlab and Java languages as part of  the
-Test Suite of the companion competition. To give competitors an idea of the
-computationa cost of the challenge, the runtime required for 10,000 function 
-evaluations (FEs) was published for a particular configuration, where the Java
-implementation took 7596ms and the Matlab version 1115ms. The whole experiment had
+algorithm, data types and structures employed in large scale optimization problems. 
+The function was implemented in both Matlab and Java languages as part of  the
+Test Suite of the companion competition of the Special Session. To give competitors an idea of the
+computational cost of the challenge, the runtime required for 10,000 function 
+evaluations (FEs) for a particular configuration was published by organizers. The Java
+implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
 3,000,000 FEs as the termination condition. Using the Java implementation as
-a guide, the function described above was implemented to compare the performence against
-both implementations, again using the test of time required for 10,000 FEs. 
+a guide, the function described above was implemented to compare the performence of JavaScript against
+both implementations, again using as a test, the time required for doing a 10,000 FEs. 
 Before discussing the setup and performance results, certain details of
@@
 \begin{itemize}
-\item {\em Randomize} Both Malab and Java implementations rely on a 
-Java Randomizer library as a standardized implementation of the 
-randomization of shift vectors and matrices used for the experiments.
-The library included the generation of pseudorandom Gaussian of type
+\item {\em Randomize} Both Malab and Java programs rely on a 
+Java Randomizer library for the implementation of the 
+randomization of shift vectors and matrices used for by the function.
+The library included the generation of pseudorandom Gaussians of type
 double and uniformly distributed integers. Generators used a single
 {\tt long} seed. In order to implement the same functionality in
-JavaScript, the library random-js was used, as there are inconsistencies 
-in the implementation of {\tt Math.random()} between engines in the
-number of bits used in the generation and results are non-deterministic. 
-The library random-js, implements a Mersenne Twister algorithm producing 
-consistent results across all JavaScript implementations assuming the same seed.
-\item {\em Arrays} 
-\item {\em Timing functions}
-\item {\em Tools} 
+JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
+in the implementation of {\tt Math.random()} between engines and their results are non-deterministic. 
+The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
+consistent results across all JavaScript VMs assuming the same seed.
+\item {\em Timing functions} In order to compare the performance of 
+each implementation, an accurate and consistent way of measuring 
+the runtime is needed. In JavaScript the {\tt Date class} is often used 
+to measure execution time, but has the limitaoun of having a maximum resolution given in miliseconds.
+In order to measure the runtime intervals with a higher precission two functions
+where used: for Node.js we used the native {\tt process.hrtime()} function
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array.
+For testing in the browser, the {\tt Performance.now()} function was used, this
+function uses floating-point numbers with up to microsecond precision. 
+A drawback found is that the {\tt Performance.now()} function is that is only implemented
+in Firefox and Chrome browsers. Both functions are independent of the system clock. 
+\item {\em Tools} An important consideration when chosing certain language to 
+to write non-trivial programs is the availability of development tools. When developing
+the experiments, we used a comprehensive set of developer tools available in both
+browsers and desktop. These tools included package managers, debbugers, logging, 
+network monitors, editors and even debugging of the multi-threaded execution of Web Workers.
+A drawback found is that developer tools very from browser to browser, 
+but they all have the same functionality.
+\item {\em Data Types} JavaScript uses floating point numbers with a 
+limited precision of 64 bits that partialy implement the functionality of
+the {\tt StrictMath} library used by the Java implemantation. If more precision
+is needed developers could use {\tt math.js}, an extensive math library with support
+for matrices, big and complex numbers.
 \end{itemize}
 
-The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
-The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
+We found that the JavaScript ecosystem is mature enough for developing non-trivial 
+code, but developers have to consider the differences betwwen VMs implementations
+and also use certain libraries in order to increase the precision and repeatability of
+algorithms. As for the performance of the JavaScript implementaion against Java and Matlab
+we conducted first the basic test described earlier in a 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
 using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
 MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2.
+and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
+In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
+991ms, JavaScipt took 35\% more time than Java with 1,270ms using Node.js and 1279ms in Chrome. 
+The results also show that there is not much difference between the Node.Js implemenation an that of a
+browser. There is also not much difference between running the code in the main thread or as workers, but
+when running the experiment in Firefox a dialog alerted of the script taking to long, showing that this
+kind of processing is not intended for the main thread.
 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
10k FE experiment

@@
 In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
-991ms, JavaScipt took 35\% more time than Java with 1,270ms using Node.js and 1279ms in Chrome. 
+991ms, JavaScipt took 32\% more time than Java with 1240ms in Chrome using a Web Worker and 1,270ms using Node.js. 
 The results also show that there is not much difference between the Node.Js implemenation an that of a
@@
 when running the experiment in Firefox a dialog alerted of the script taking to long, showing that this
-kind of processing is not intended for the main thread.
-
+kind of processing is not intended for the main thread. Also there is not much diference when running two
+Web Workers in parallel taking 1279ms each. This experiment shows that JavaScript is viable 
+language for developing complex mathematical functions for running in the browser. 
 \begin{figure}[!htb]

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Working on issue

@@
 Volunteer computing is used routinely in science and industry, but it
-still has a long way to go if it is going to become a mainstream technology in the
-same way the cloud technologies in general have. The lack of general
+still has a long way to go if it is going to become a mainstream technology in
+the same way the cloud technologies in general have. The lack of general
 purpose and open source tools
@@
 they assume that the nodes that are going to run the system are under
-your control or at least you have access to them.
+the control of developers or they at least have access to them.
 
@@
 provisionally named {\sf NodIO}. {\sf NodIO} is first a two tier client/server 
-application, in which clients can possibly run in a browser. 
-Therefore, it can be run in any web-enabled device, including many over which 
-the scientist has no control.
+application, in which clients can run in a browser. 
+Therefore, they can embedded any web-enabled device, including many over which 
+the developers have no control.
 
@@
 There exist other volunteer frameworks written in compiled languages,
-such as C++ or Java, that can achieve more performance than
+such as C++ or Java, that can achieve better performance than
 JavaScript. However, one of the advantages of our system is that
 clients do not need to install anything in their machines, they just
-need to open a regular web page  
+need to open a web page  
 % Somewhat redundant 
@@
 % also removed the default, redundant with next sentence
-in most browsers, such as Chrome, Safari or Firefox, that are
+in a compatible browser, such as Chrome, Safari or Firefox, which are
 available in almost every desktop or mobile device. Other advantage is
-that the source code can be directly examined by the volunteer, not
-being compiled as other options (such as Java Applets, Flash scripts or
+that the source code can be directly examined by volunteers, because is not
+compiled as other options (such as Java Applets, Flash scripts or
 ActiveX). Also, as we discuss in this paper, it can provide some extra
@@
 %PABLO: hay que enlazar con el siguiente párrafo
-Those desktop clients have also the same issues as any kind of
+Desktop clients have the same issues as any 
 volunteer computing system, including our own: these issues are
 related to security as well as
-safety, which introduces a {\em social} aspect to the design of a
+safety, which introduce a {\em social} aspect to the design of a
 computing system that makes it a {\em techno-social system}
@@
 security mechanisms and was proposed in \cite{chandy1996world} as the
-basis for the creation of a world-wide peer to peer system using Java
-compiled programs. The ATLAS system
+basis for the creation of a world-wide peer to peer system using compiled Java
+programs. The ATLAS system
 \cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
@@
 providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed using desktop
+URI, something that is not guaranteed by using desktop
 applications. Baratloo et al. tried in 1996 to create a system with
@@
 server. Since the Java virtual machine is a sandbox, it is safe to the user
-and running a compiled bytecode guarantees a certain level of
+and as it runs compiled bytecode it also guarantees a certain level of
 performance.
@@
 paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most. \cite{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and also use JSON (JavaScript Object Notation) for data interchange. Another advantage is that
+in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them. \cite{wood13:nodejs:paas}. Many web services
+have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
 developers can use the same software tools for both client and server
@@
 fault-tolerance of volunteer algorithms
-\cite{gonzalez2010characterizing} which can, of course, be studied in
-a more general distributed computing context
+\cite{gonzalez2010characterizing} which it can, of course, be studied in
+the more general context of distributed computing 
 \cite{nogueras2015studying} or including it in a more general study of the
@@
 
-On the other hand, measuring performance of the resulting metacomputer
+On the other hand, measuring the performance of the resulting metacomputer
 involves understanding the dynamics of this kind of systems. Initial
@@
 results obtained in \cite{agajaj}. Although SETI@home works
-at larger scale in both quantity of volunteers and scope of projects, 
+at a larger scale in both quantity of volunteers and the scope of projects, 
 our initial tests indicate that
@@
 powerful laptops 
-or desktop computers in a research center. The methodology used to
+or workstation computers in a research center. The methodology used to
 gather resources and the algorithms used will be described next, as
-well as the results obtained with the setup.
+well as the results obtained with the initial setup.
 
@@
 In general, a volunteer-based distributed evolutionary computation
-system based on the browser is simply a client-server architecture
-whose client is, or can be, embedded in the browser via
+system based on the browser is simply a client-server system
+whose client is, or it can be, embedded in the browser via
 JavaScript. Since JavaScript is  the only language that is present
@@
 function in Python, Erlang, Scala or even Java, and them compile it to
-JavaScript.
+JavaScript. For completness, an experiment comparing the performance of
+a JavaScript implementation versus Java and Matlab is included in section
+\ref{sec:w2}.
 
@@
 prosecute or discipline the volunteer. In the case of JavaScript,
-this issue is agravated because even if it is obfuscated, sorce code
+this issue is aggravated because even if it is obfuscated, source code
 and data is sent to clients. 
 \item Participants must also trust the application provider, 
-regarding invation of privacy, the integrity of their devices 
+regarding invasion of privacy, the integrity of their devices 
 and intended application. 
@@
 based on the {\sf NodEO} library, whose architecture has been
-developed using JavaScript on the client as well as the server sides.
+developed using JavaScript on the client as well as the server.
 All parts of the system are free and available with a free license
@@
   GET}ing a random chromosome from it, and {\em information} related
-  to performance and state of the experiments. It also performs logging
+  to the performance and state of the experiments. It also performs logging
   duties, but is basically a very lightweight and high performance
@@
   an evolutionary algorithm {\em island} starting with a random
-  population and sending, every 100 generations, the best individual
+  population and then sending, every 100 generations, the best individual
   back to the server (via a {\tt PUT} request) and requesting a random
@@
 In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
-991ms, JavaScipt took 32\% more time than Java with 1240ms in Chrome using a Web Worker and 1,270ms using Node.js. 
-The results also show that there is not much difference between the Node.Js implemenation an that of a
+991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a Web Worker and 1,270ms using Node.js. 
+The results also show that there is not much difference between the Node.Js implementation an that of a
 browser. There is also not much difference between running the code in the main thread or as workers, but
 when running the experiment in Firefox a dialog alerted of the script taking to long, showing that this
-kind of processing is not intended for the main thread. Also there is not much diference when running two
+kind of processing is not intended for the main thread. Also there is not much difference when running two
 Web Workers in parallel taking 1279ms each. This experiment shows that JavaScript is viable 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
More revisions, please check with diff

@@
   Chart.js} have
-been used to build the elements of the system.
+been used to build the user interface elements of the system.
 \begin{figure}[!t]
@@
 problem. In this case the classical Trap function \cite{Ackley1987} has been
-used as well as the Rastrigin floating-point optimization
+used as well as the Rastrigin's floating-point optimization
 problem. JavaScript is a functional language and declaring a different 
@@
   Express.js. Although this single server is a bottleneck since it
-  will eventually saturate, the fact that it runs as a single thread
-  allows the service of many requests. In fact, a limit with the
+  will eventually saturate, the fact that it runs as a non-blocking single thread
+  allows the service of many requests. In fact, a limit in the
   number of simultaneous requests will be reached, but so far it has
@@
   Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of NodIO can be used.
+  has been disabled, a basic version of NodIO can also be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
 compatible and can be used in a complementary way. Fault tolerance
-will be experimented by dropping the server and check the continuity
+will be experimented by dropping the server and checking the continuity
 of the experiments.
@@
 %---------------------------------------------------------------
-\section{Experiments and obtained results}
+\section{Baseline Experiment and obtained results} %Cambié el nombre
 \label{sec:experiments}
@@
 successful. The experiments were repeated for population $p=1024$,
-with success rate upgraded to 100\% and average duration $3.46$
+with a success rate upgraded to 100\% and an average duration of $3.46$
 seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. The baseline system has these
+are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
 characteristics: \begin{verbatim}
@@
 These two experiments establish first a baseline result and show that
-time to success depends on the population, with bigger population
+the time to success depends on the population size, with a bigger population
 contributing more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provide a free tier within which this 
+PaaS, which provides a free tier within which this 
 experiment could be performed. That way, and as required, the hardware
@@
 \begin{table*}
-\caption{Summary of time per run, number of IPS and number of {\tt
+\caption{Summary of time per run, number of IPs and number of {\tt
     PUT}s per IP in the initial runs. \label{tab:summary:os}}
@@
 A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of IPs
+\ref{tab:summary:os}, which shows the median number of Internet Protocol addressess (IPs)
 intervening in each experiment,  median time needed
@@
 maximum number of different IPs for each experiment is also in the
-same ballpark and of the order of 10, which is also consistent with
+same range and of the order of 10, which is also consistent with
 prior work. The median time has a big range of variation, but 50\% of
@@
 place in a browser, and in fact most of the time is devoted to
-presenting the charts; the server is not the same since it is run in a
+presenting the charts; the server is also not the same since it is run in a
 PaaS instead of a high-performance desktop computer, and it is also a
@@
 allowed us to run evolutionary computation experiments {\em faster} than
-in our very own computer. And the results is that although it will
+in our very own computer. And the results indicate that although it will
 happen, 1 out of roughly 5 or roughly 20, depending on what baseline
@@
 %
-We will have to analyze experiment data a bit further to find out why
+We will have to analyze experimental data a bit further to find out why
 this happens and also if there are some patterns in the three sets of
@@
 experiments are announced, it is quite difficult that several
-volunteers join in in such a short period of time, even more so taking
+volunteers join in in such a short period of time, even more if we take
 into account that volunteers are not {\em carried over} from previous
 experiments. This implies that it would be convenient to use a problem
-with a bigger size to check this hypothesis; however, at this point we
+of a bigger size to check this hypothesis; however, at this point we
 have not found this convenient since there are several other issues
-that have to be solved, as it will be shown next.
+that have to be solved, as it will be shown next. %Creo que este párrafo
+%saldría sobrando ahora que hicmios un experimento más grande
 
@@
 distribution is governed by three parameters, the usual location $\mu$
-which is related to where it has its {\em center} and scale $\sigma$,
+which is related to where it has its {\em center} and a scale $\sigma$,
 related to the size, but also a third shape $\xi$ parameter that is
@@
 population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvment was the 
+different for each client.  The other improvement was the 
 use the Web Worker API as described next.
@@
   changes. 
-\item {\em Client main thread}  This process handles the user interaction
-using the JQuery framework. The Chart.js library is used to display a chart of
-the best fitness in the population over time. The main script also creates
+\item {\em Client main thread} The main script creates
 worker instances and handles the user interface.
@@
 \item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure the secuential number of experiment is initialized and a log
+initialization procedure the sequential number of experiment is initialized and a log
 file is created. The shared pool implemented as an array is reset.
@@
 the shared population. Also the current generation number and the best
-chromosome is sent to the client main thread in order to update the DOM to
+chromosome is sent to the client's main thread in order to update the DOM to
 display the current state of the island (described next). These messages are
@@
 number of IPs, although in this case it would be combination
-IP-worker, being considered every worker a different IP. The median
+IP-worker, as we considered every worker as a unique IP. The median
 number of seconds is almost 80\% better than the previous version, and
@@
   NodIO-W$^2$}, which restarts an island after a solution has been
-found and also is running even if the tab is not on the
+found and also that it keeps running even if the tab is not on the
 foreground. This means that volunteers can leave the experiment
@@
 experiment after experiment, unlike the previous version, where they
-stopped contributing after the solution was found. This is also
+stopped contributing after one solution was found. This is also
 reflected in the number of PUTs, which is the number of generations
-divided by 100; more than 50\% of the volunteers run for more than
+divided by 100; more than 50\% of the volunteers run the experiment for more than
 4000 generations. All things considered, this means than there are
@@
 this case, the number of PUTs made per client does not correspond to a
-uniform amount of computaton, since 100 generations with a random
+uniform amount of computation, since 100 generations with a random
 population correspond, in each case, to different number of
-oprations; in other words, the same number of PUTs will correspond to
+operations; in other words, the same number of PUTs will correspond to
 different number of operations in each case. However, since clients
@@
 \subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigins Function}
-The most common use of JavaScript is in the development of
-asynchronous Web applications. For this purpose JavaScript has
-been traditionally implemented as an interpreted language not designed for the
-development of high performance systems. On the other hand, current JavaScript 
+JavaScript has been traditionally implemented as an interpreted language not designed for the development of high performance systems. On the other hand, current JavaScript 
 virtual machines (VMs) are closing the gap, for instance the Google V8 
@@
 environments such as Node.js and is also used in desktop and mobile applications. 
-To further evaluate how a current JavaScript implemantation could be used by a
+%Mario: Creo que el párrafo de arriba sale sobrando
+To further evaluate how a current JavaScript implementation could be used by a
 computer scientist to develop complex optimization problems,
@@
 
-The basic Rastrigins function is separable[], and is defined as follows;
+The basic Rastrigin's function is separable, and is defined as follows;
 \begin{equation}
@@
 optima grows exponentially with the increase of dimensionality. To make it 
-nonseparable, an orthogonal matrix is also used for coordinate rotation.
+non-separable, an orthogonal matrix is also used for coordinate rotation.
 The rotated Rastrgin’s function is defined as follows:
@@
 
-This optimization problem was selected because is representative of the kind of
-algorithm, data types and structures employed in large scale optimization problems. 
+This optimization problem was selected because it is representative of the kind of
+algorithms, data types and structures employed in large scale optimization problems. 
 The function was implemented in both Matlab and Java languages as part of the
-Test Suite of the companion competition of the Special Session. To give competitors an idea of the
+Test Suite of the companion competition of the Special Session and to give competitors an idea of the
 computational cost of the challenge, the runtime required for 10,000 function 
@@
 3,000,000 FEs as the termination condition. Using the Java implementation as
-a guide, the function described above was implemented to compare the performence of JavaScript against
-both implementations, again using as a test, the time required for doing a 10,000 FEs. 
+a guide, the function described above was implemented to compare the performance of JavaScript against
+both implementations, again using the time required for 10,000 FEs. 
 Before discussing the setup and performance results, certain details of
-the implementation in JavaScript are presented, to give readers an idea of 
-how practical and mature is the language for developing complex mathematical functions
-as well as some limitations.
+the implementation in JavaScript are presented to give readers an idea of 
+how practical and mature is the language for developing complex mathematical functions.
 
@@
 \item {\em Randomize} Both Malab and Java programs rely on a 
-Java Randomizer library for the implementation of the 
-randomization of shift vectors and matrices used for by the function.
-The library included the generation of pseudorandom Gaussians of type
-double and uniformly distributed integers. Generators used a single
+Java Randomizer library for the 
+randomization of the shift vectors and matrices used by the function.
+The library includes the generation of pseudorandom Gaussians of type
+double and uniformly distributed integers. Generators use a single
 {\tt long} seed. In order to implement the same functionality in
 JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
-in the implementation of {\tt Math.random()} between engines and their results are non-deterministic. 
+in the implementation of the standard {\tt Math.random()} between engines
+and also their results are non-deterministic. 
 The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
-consistent results across all JavaScript VMs assuming the same seed.
+consistent results across all JavaScript VMs.
 \item {\em Timing functions} In order to compare the performance of 
 each implementation, an accurate and consistent way of measuring 
-the runtime is needed. In JavaScript the {\tt Date class} is often used 
-to measure execution time, but has the limitaoun of having a maximum resolution given in miliseconds.
-In order to measure the runtime intervals with a higher precission two functions
+the runtime of functions is needed.
+In JavaScript the {\tt Date class} is often used 
+to measure execution time, but has the limitation of having a maximum resolution given in milliseconds.
+In order to measure the runtime intervals with a higher precision two functions
 where used: for Node.js we used the native {\tt process.hrtime()} function
-which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array.
-For testing in the browser, the {\tt Performance.now()} function was used, this
-function uses floating-point numbers with up to microsecond precision. 
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array and for measuring in the browser, the {\tt Performance.now()} function was used,
+this function uses floating-point numbers with up to microsecond precision. 
 A drawback found is that the {\tt Performance.now()} function is that is only implemented
 in Firefox and Chrome browsers. Both functions are independent of the system clock. 
-\item {\em Tools} An important consideration when chosing certain language to 
+\item {\em Tools} An important consideration when choosing certain language to 
 to write non-trivial programs is the availability of development tools. When developing
-the experiments, we used a comprehensive set of developer tools available in both
-browsers and desktop. These tools included package managers, debbugers, logging, 
+the experiments, we used a comprehensive set of developer tools available for both
+browsers and desktop. These tools included package managers, debuggers, logging libraries, 
 network monitors, editors and even debugging of the multi-threaded execution of Web Workers.
@@
 \item {\em Data Types} JavaScript uses floating point numbers with a 
-limited precision of 64 bits that partialy implement the functionality of
-the {\tt StrictMath} library used by the Java implemantation. If more precision
-is needed developers could use {\tt math.js}, an extensive math library with support
-for matrices, big and complex numbers.
+limited precision of 64 bits that partially implement the functionality of
+the {\tt StrictMath} library used by the Java implementation. If more precision
+is needed developers could use {\tt math.js}, an extensive math library with
+support for matrices and big numbers.
 \end{itemize}
@@
 We found that the JavaScript ecosystem is mature enough for developing non-trivial 
-code, but developers have to consider the differences betwwen VMs implementations
-and also use certain libraries in order to increase the precision and repeatability of
-algorithms. As for the performance of the JavaScript implementaion against Java and Matlab
+code, but developers have to consider the differences between VMs implementations
+and use certain libraries in order to increase the precision and repeatability of
+algorithms. As for the performance of JavaScript against Java and Matlab
 we conducted first the basic test described earlier in a 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
@@
 In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
-991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a Web Worker and 1,270ms using Node.js. 
-The results also show that there is not much difference between the Node.Js implementation an that of a
-browser. There is also not much difference between running the code in the main thread or as workers, but
-when running the experiment in Firefox a dialog alerted of the script taking to long, showing that this
-kind of processing is not intended for the main thread. Also there is not much difference when running two
-Web Workers in parallel taking 1279ms each. This experiment shows that JavaScript is viable 
+991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a single Web Worker and 1,270ms using Node.js. 
+The results also show that there is not much difference between the Node.js implementation an that of a
+browser.
+There is also not much difference between running the code in the main thread or in Web Workers.
+When running the experiment in Firefox a dialog alerted that the script was taking to much time, showing that this
+kind of processing is not intended for the main thread. Also there is not much overhead when running two
+Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
 language for developing complex mathematical functions for running in the browser. 
@@
 
-\subsubsection{Setup of experiment in NodIO-W$^2$ }
+\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
 

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Changed figure and caption
Closes #46
Takes out colloquialisms
Which I actually liked. Closes #45.

@@
 
-4. Style: Paper is generally written in a clear way. Statements like "But fair is fair," (page 6), sound a bit
-colloquially in a scientific paper and may be omitted. However, my main concern is the readability of the
-graphs. General remark - font used for description of axes should be significantly larger. In Fig. 2 the
-authors can change scale for axe Y, as the box for the population of 1024 individuals is very small.
-What do the authors mean by "times"? Time in seconds? I am not sure if this Figure is necessary at all
-as the description is understandable without it. In Fig.4 all graphs are similar and only the one in the
-middle is analyzed, so maybe others are not necessary? The same is with Fig. 5. What does "x" stand
-for in the description of axe X? (the same is with Fig. 8).
+4. Style: Paper is generally written in a clear way. Statements like
+"But fair is fair," (page 6), sound a bit colloquially in a scientific
+paper and may be omitted. However, my main concern is the readability
+of the graphs. General remark - font used for description of axes
+
+\end{quote}
+
+We have eliminated colloquialisms from the text.
+
+\begin{quote}
+should be significantly larger. In Fig. 2 the authors can change scale
+for axe Y, as the box for the population of 1024 individuals is very
+small.  What do the authors mean by "times"? Time in seconds? I am not
+sure if this Figure is necessary at all as the description is
+understandable without it. In Fig.4 all graphs are similar and only
+the one in the middle is analyzed, so maybe others are not necessary?
+The same is with Fig. 5. What does "x" stand for in the description of
+axe X? (the same is with Fig. 8).
+\end{quote}
+
+The labels, and indeed the label size, are not too important; the
+graph shape is. That is one reason they are grayed out. However, we
+have changed as instructed Figure 2, which is more readable now. We do
+mean times to solution, shortened for clarity. We have also clarified
+that on the caption.
+
+\begin{quote}
 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Takes out colloquialisms
Which I actually liked. Closes #45.

@@
 20\% on average for the first baseline experiment, roughly 3\% in the
-second case. The comparison is not totally fair: this experiment takes
+second case. The comparison is not totally fair: This experiment takes
 place in a browser, and in fact most of the time is devoted to
@@
 the evolutionary algorithm start again. However, it puts a probably
-unnecessary cap on the time, or number of experiments, that a user
+unnecessary upper limit on the time, or number of experiments, that a user
 will voluntarily perform.
@@
 experiments. This implies that it would be convenient to use a problem
-of a bigger size to check this hypothesis; however, at this point we
-have not found this convenient since there are several other issues
-that have to be solved, as it will be shown next. %Creo que este párrafo
+of a bigger size to check this hypothesis as we have done in the next
+experiment. 
+%  however, at this point we
+% have not found this convenient since there are several other issues
+% that have to be solved, as it will be shown next. %Creo que este párrafo
 %saldría sobrando ahora que hicmios un experimento más grande

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Explains why all graphs are necessary
Closes #47

@@
 are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
-characteristics: \begin{verbatim}
+characteristics: 
+\begin{verbatim}
 Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
@@
 many experiments took a desirable amount of time, and of the order of
-10 seconds. As it can be seen, it drops sharply implying there are
+10 seconds, and which appears in all three graphs. As it can be seen,
+it drops sharply implying there are 
 just a few of them, and with diminishing probability as time

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Explains why all graphs are necessary
Closes #47
Changing figures
Refs #48

@@
 the one in the middle is analyzed, so maybe others are not necessary?
-The same is with Fig. 5. What does "x" stand for in the description of
-axe X? (the same is with Fig. 8).
 \end{quote}
@@
 mean times to solution, shortened for clarity. We have also clarified
-that on the caption.
+that on the caption. We also think that all graphs are needed in Fig
+4, since they represent different experiments, and we have added a
+sentence to highlight that the {\em tail} appears in all three.
+
+\begin{quote}
+The same is with Fig. 5. What does "x" stand for in the description of
+axe X? (the same is with Fig. 8).
+\end{quote}
+
+$x$ has been changed to {\sf rank}, which is the usual way of
+representing Zipf's law, a measure (in this case number of {\tt PUT}s
+vs. rank. We have also changed size, color and sigils to represent
+each point to make it a bit clearer. 
+%Still fig 8 to go
 

### 2015-11-17 data/plot-zipf-openshift.R

@@ -3,17 +3,17 @@ library(evd)
 
 process.puts <- function(suffix ) {
     these.puts <- read.csv(paste0("ips-puts-openshift-",suffix,".dat"))
-    this.df <- data.frame(x=1:nrow(these.puts),
+    this.df <- data.frame(rank=1:nrow(these.puts),
                           puts=sort(these.puts$puts,decreasing=T))
     this.fit <- fgev(this.df$puts)
     print(this.fit)
-    this.fit.plot <- data.frame(x=1:length(this.df$puts),
+    this.fit.plot <- data.frame(rank=1:length(this.df$puts),
                                 y=sort(rgev(length(this.df$puts),
                                     this.fit$estimate["loc"],
                                     this.fit$estimate["scale"],
                                     this.fit$estimate["shape"]),decreasing=T))
-    ggplot()+geom_point(data=this.df,aes(x=x,y=puts))+geom_point(data=this.fit.plot,aes(x=x,y=y,color='red'))+scale_y_log10()
-    ggsave(paste0("puts-openshift-",suffix,".png"))
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))
+    ggsave(paste0("puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }
     
@@ -22,17 +22,16 @@ dt.puts.4.4 <- process.puts("4-4")
 dt.puts.4.24 <- process.puts("4-24")
 dt.puts.7.31 <- process.puts("7-31")
 
-
-ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4.4,aes(x=x/length(dt.puts.4.4$x),y=puts,color='4/4'),stat='Identity')+geom_point(data=dt.puts.4.24,aes(x=x/length(dt.puts.4.24$x),y=puts,color='4/24'),stat='Identity')+geom_point(data=dt.puts.7.31,aes(x=x/length(dt.puts.7.31$x),y=puts,color='7/31'),stat='Identity')
[…]

### 2015-11-17 special-issue-ci/Response to reviewers.tex
comentario en las respuestas a revisores
fixing a typo

@@
 JavaScript uses a JIT compiler that provides a good performance; we
-have published these tests as an open access technical report.
+have published these tests as an open access technical report.  % quizás indicar la URL aquí
 
@@
 At any rate, we have also performed a new experiment with the
-Rasrtrigin function, which in fact cannot be solved in a few minutes
+Rastrigin function, which in fact cannot be solved in a few minutes
 (or seconds, as in the case of Trap) by a desktop system.

### 2015-11-17 data/ips-puts.R
Modified plot for fig 8
And a bit of the biblio. Closes #48

@@ -3,22 +3,24 @@ library(evd)
 library(fitdistrplus)
 
 ips.puts.ww <- read.csv("ips-puts-ww.dat")
-ips.puts.ww.df <- data.frame(x=1:nrow(ips.puts.ww),
+ips.puts.ww.df <- data.frame(rank=1:nrow(ips.puts.ww),
                              puts=sort(ips.puts.ww$puts,decreasing=T))
 this.fit <- fgev(ips.puts.ww$puts)
 print(this.fit)
-this.fit.plot <- data.frame(x=1:length(ips.puts.ww$puts),
+this.fit.plot <- data.frame(rank=1:length(ips.puts.ww$puts),
                             y=sort(rgev(length(ips.puts.ww$puts),
                                 this.fit$estimate["loc"],
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=x,y=puts,color='Data'))+geom_point(data=this.fit.plot,aes(x=x,y=y,color='Fit'))+scale_y_log10()
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggsave('gev-fit-ww.png',width=5,height=5)
 
 print(fitdistr(ips.puts.ww$puts,"gamma"))
 this.weibull <- fitdistr(ips.puts.ww$puts,"weibull")
 
-this.fit.weib <- data.frame(x=1:length(ips.puts.ww$puts),
+this.fit.weib <- data.frame(rank=1:length(ips.puts.ww$puts),
 y=sort(rweibull(length(ips.puts.ww$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=x,y=puts,color='Data'))+geom_point(data=this.fit.weib,aes(x=x,y=y,color='Fit'))+scale_y_log10()
[…]

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Modified plot for fig 8
And a bit of the biblio. Closes #48
Extends abstract
To stress the main points and closes #49

@@
 each point to make it a bit clearer.
-%Still fig 8 to go
 
@@
 
+\end{quote}
+
+Finally, since the reviewer thinks that length is fine and for the
+reasons outlined above, we have decided to keep the graphs and eplain
+them a bit better.
+
+\begin{quote}
 6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
@@
 Worker architecture.
+\end{quote}
 
+we have extended the abstract stressing our main points, that were
+first to test the validity and measure the performance of our
+framework NodIO and then to use the measures as first steps to a model
+of volunteer uses.
+
+\begin{quote}
 Comments to Authors:

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changing figures
Refs #48
Modified plot for fig 8
And a bit of the biblio. Closes #48
Extends abstract
To stress the main points and closes #49

@@
 Volunteer computing is used routinely in science and industry, but it
-still has a long way to go if it is going to become a mainstream technology in
-the same way the cloud technologies in general have. The lack of general
-purpose and open source tools
-with analytics built in could be one of the reasons why this is so, as
-well as the lack of proofs of concept. That is why in this paper we
-present {\sf NodIO}, a framework used to create client-server architectures for distributed
+still has a long way to go if it is going to become a mainstream
+technology in the same way the cloud technologies in general have. The
+lack of general purpose and open source tools with analytics built in
+could be one of the reasons why this is so, as well as the lack of
+proofs of concept. That is why in this paper we present {\sf NodIO}, a
+framework used to create client-server architectures for distributed
 evolutionary algorithms, written entirely in JavaScript, which needs
@@
 asynchronous, distributed evolutionary algorithms running in the
-browser.  The architecture, along with some initial results on the
-expected performance and scaling
-behavior is presented. We also use the algorithm data to
-find out what kind of patterns the user behavior follow and try to fit
-them to some statistical distribution.
+browser, as well as any other client consuming its REST application
+programming interface.  The architecture, along with some initial
+results on the expected performance and scaling behavior when the
+number of users is increased is presented. We also use
+the algorithm data to find out what kind of patterns the user behavior
+follow by examining the resources they devote to our experiment, and
+try to fit them to some statistical distribution to create the bases
+of a model of user behavior in volunteer computing experiments of this
+kind that can be used by researchers to predict its performance. 
+In general, our experiments prove first that NodIO is a flexible and
+low-overhead platform to perform distributed computation experiments
+that include volunteers and that user behaviour follows some patterns,
+and, thus, are amenable to a model.
 \end{abstract}
@@
   4/4, 4/24 and 7/31.} 
+% Plotted with ../data/plot-zipf-openshift.R
 \label{fig:puts:os}
@@
 \label{fig:gev:w2}
+% File ../data/ips-puts.R
 \end{figure}

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English from abstract and introduction

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 \begin{abstract}
-Volunteer computing is used routinely in science and industry, but it
-still has a long way to go if it is going to become a mainstream technology in
-the same way the cloud technologies in general have. The lack of general
+Volunteer computing is used routinely 
+% (Paloma) Adverbs are always written before verbs, check this throughout the text
+in science and industry, but it
+still has a long way to go if it is going to become
+% (Paloma) I would propose "a long way to go in order to become"
+ a mainstream technology in
+the same way the cloud technologies in general have. The lack of 
+% (Paloma) Add "both" here to better complete the sentence
+general
 purpose and open source tools
-with analytics built in could be one of the reasons why this is so, as
+with analytics built in
+% (Paloma) "built-in" is an adjective and, as such, has to be written before names
+ could be one of the reasons why this is so, as
 well as the lack of proofs of concept. That is why in this paper we
 present {\sf NodIO}, a framework used to create client-server architectures for distributed
-evolutionary algorithms, written entirely in JavaScript, which needs
+evolutionary algorithms, written entirely % (Paloma) Adverb. I would also add an "and" after the comma
+ in JavaScript, which needs
 minimal server support to work and is able to support persistent,
@@
 expected performance and scaling
-behavior is presented. We also use the algorithm data to
+behavior, is presented. We also use the algorithm data to
 find out what kind of patterns the user behavior follow and try to fit
@@
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is first a two tier client/server 
-application, in which clients can run in a browser. 
-Therefore, they can embedded any web-enabled device, including many over which 
+provisionally named {\sf NodIO}. {\sf NodIO} is first
+% (Paloma) Is the "first" word really needed?
+ a two tier client/server 
+application, in which clients can run on a browser. 
+Therefore, they can
+% (Paloma) be embedded in?
+ embedded any web-enabled device, including many over which 
 the developers have no control.
@@
 clusters.
-Considering this, we leveraged our experience in the design of web
+Considering this, we leveraged 
+% (Paloma) What 'something' did you leverage?
+our experience in the design of web
 applications, including JavaScript, since their inception and other
@@
 servers and browsers as clients.
+% (Paloma) I think you should rewrite this whole sentence.
+% For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided...
 
@@
 available in almost every desktop or mobile device. Other advantage is
-that the source code can be directly examined by volunteers, because is not
+that the source code can be directly examined by volunteers, because it is not
 compiled as other options (such as Java Applets, Flash scripts or
-ActiveX). Also, as we discuss in this paper, it can provide some extra
+ActiveX)
+% (Paloma) Brackets not really needed here
+. Also, as we discuss in this paper, it can provide some extra
 features such as enhanced security or operating system independence.
@@
 Desktop clients have the same issues as any 
-volunteer computing system, including our own: these issues are
+volunteer computing system, including our % (Paloma) their?
+ own: these issues are
 related to security as well as
-safety, which introduce a {\em social} aspect to the design of a
+safety, which introduce 
+% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
+a {\em social} aspect to the design of a
 computing system that makes it a {\em techno-social system}
-\cite{vespignani2009predicting}. In this paper our intention is to
+\cite{vespignani2009predicting}. In this paper, our intention is to
 present the {\sf NodIO} framework and analyze it from two different
@@
 in Section \ref{sec:description}, followed by the baseline experiments
-in Section \ref{sec:experiments} and experiments using a new client
+in Section \ref{sec:experiments}, and experiments using a new client
 architecture that uses web workers in Section \ref{sec:w2}.

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Commenting on the value of experiments
And also the focus. Closes #51.

@@
 
+\end{quote}
+
+That is true, but it makes it anyway the most inclusive system so
+far, as we have tested it with smartphones, tablets and computers. In
+general, nowadays with rich internet applications disabling JavaScript
+is not common, but anyway we do not aim for {\em all} clients, we have
+enough with {\em most}. Windows Phone, for instance and maybe
+fortunately, is not too common.
+
+\begin{quote}
  - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
@@
  system, what is a crucial parameter for distributed computing.
+\end{quote}
+
+We have performed a new experiments in which we have had hundreds of
+simultaneous users. In fact, with the experiments already in the
+previous version of the papers 5 was a central measure, but there were
+peaks of dozens. 
+Adaptativity is by design, since it uses any browser and does not need
+any special provision for new systems. This has also been explained in
+th expanded text.
+
+\begin{quote}
 
@@
  In my opinion you should at least highlight above issues.
+\end{quote}
 
+We have also justified our {\em pragmatic} approach to security
+increasing trust by the user, but also monitoring their behavior to
+detect it and eventually ban users. So far, this has not happened.
+
+\begin{quote{
 2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
@@
 evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
-within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run (page 4, if I understood it correctly).
+within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
+(page 4, if I understood it correctly).
+\end{quote}
+I fact 3.46 is the average for all 50 runs. We have enphasized the
+word average to make that clearer.
+
+\begin{quote}
 Thus, what was the point of using distributed computing to solve such simple problem? Why the
@@
 to focus the paper on the behavior analysis of volunteers participating in a distributed computing
-system, what may be also interesting an valuable for the potential readers.
+system, what may be also interesting an valuable for the potential
+readers.
+\end{quote}
 
+In fact, that was never the point of the paper. We have chosen Trap as
+a relatively difficult problem that allows us mainly to test scaling
+and measure behavior. In a production environment, a volunteer-only
+environment would only be used if there is absolutely no other
+resource available; in general, it would be combined with the rest of
+the resources available to the experimenter: cloud or local; it would
+make no sense to not use them if they are available and in fact, NodIO
+allows to combine them seamlessly in a single system. That has been
+explained a bit better in the introduction and conclusions
+section. Besides, we have added a more difficult probmem, the shifted
+Rastrigin function, to make measures in a different scenario (one in
+which the problem is not solved in a few seconds).
+
+\begin{quote}
 3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Explains dots
Closes #52

@@
 for that?
+\end{quote}
 
+The $x$ axis runs overs number of computers participating in an
+experiment; that is, $4$ plots the time taken for all experiments in
+which 4 computing nodes participated and the red dot is the average;
+it shadows the time in the case there is a single ocassion. In the
+case you mention, there are no experiments in which there were exactly
+17 nodes, for instance. The ``red dots'' indicate that there was a
+single experiment in which that many nodes participated. We have
+clarified the caption to Figure 3 to make this clearer.
+
+\begin{quote}
 For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Commenting on the value of experiments
And also the focus. Closes #51.
Explains dots
Closes #52

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 successful. The experiments were repeated for population $p=1024$,
-with a success rate upgraded to 100\% and an average duration of $3.46$
+with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
 seconds. Results for these two experiments
@@
   participating in it, with averages and standard deviation shown as
-  red dots. From left to right, experiments 4/4, 4/24 and 7/31. Shades
-of blue indicate how many experiments included that many unique IPs.}
+  red dots; in the case there is a single red dot, there was a single
+  experiment in which so many computers participated (for instance, 16
+  computers in the experiment in the far left or 29 in the middle
+  one). 
+From left to right, experiments 4/4, 4/24 and 7/31. Shades
+of blue indicate how many experiments included that many unique IPs,
+so lighter shade for a column of dots indicates that particular number
+of computers happened less frequently, darker shadow more times.}
 \label{fig:duration}

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English from state of the art

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 internet'', providing some advantages such as multi-architecture compatibility or
-security mechanisms and was proposed in \cite{chandy1996world} as the
+security mechanisms, and was proposed in \cite{chandy1996world} as the
 basis for the creation of a world-wide peer to peer system using compiled Java
@@
 leaving us with eight desirable properties.
-
-The fulfillment of these properties could be achieved using any of the
+% (Paloma) Re-write.
+For example, "But is hierarchy, this is, the capability of preferentially using resources in the same organization, the one that could be either disregarded or included in adaptivity, which leaves us with eight desirable properties."
+
+The fulfillment of these properties could be achieved by using any of the
 languages and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first, but embedding a
+installed the required virtual machine plugin first, 
+% (Paloma) These should be separate sentences
+but embedding a
 distributed computation client in the browser has the advantage of
@@
 security to the user, uniformity of computing environment
-independently of the operating system and access to a common data store in the
+independently of the operating system, and access to a common data store in the
 server. Since the Java virtual machine is a sandbox, it is safe to the user
@@
 already in the early stages of the web the possibility of
+% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
 browser-based distributed computing was already researched.
@@
 a comprehensive statistics collection subsystem, allowing its use for science,
-since it allows the collection of algorithm analytics by the
+since it allows % (Paloma) Use synonyms, like "yield" or "permit"
+ the collection of algorithm analytics by the
 experiment designer, unlike the other systems mentioned
@@
 distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. It is focused, however, in solving the
+brokers for resource discovery. It is focused
+% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
+, however, in solving the
 issues involved in browser-based computing and also in making a system
@@
 mischief of users with code in their hands
+% (Paloma) I think this sentence is extremely informal
 \cite{sarmenta-sabotagetolerance}. However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making this effort
+Sarmenta once again used Java and not JavaScript, making this effort % (Paloma) making them
 less widespread.
@@
 scale algorithms at reasonable costs. Researchers have proposed the use of
-public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards}
+public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
 and DropBox~\cite{mericloud}. The use of JavaScript to implement
@@
 when writing code and leverages 
-expertise acquired in the library itself.
+expertise acquired
+% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
+ in the library itself.
 
@@
 computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
-volunteer computing, number of users and the time spent by every one in the
+volunteer computing, % (Paloma) Whith regard to?
+number of users and the time spent by every one in the
 computation in browser-based volunteer computing experiments, have
@@
 
-As far as we know, this paper presents one of the few experiments that
+As far as we know, this paper presents one of the few experiments that % (Paloma) Change to "which" in order to not repeating "that" all the time
 uses computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. The methodology used to
+or workstation computers in a research center. The methodology used
+% (Paloma) I would say "the used methodology"
+ to
 gather resources and the algorithms used will be described next, as

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Minor changes
Refs #31 mainly.

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 Weibull distribution which was fitted to volunteer computing
-frameworks such as SETI@home \cite{javadi2009mining}. This
+frameworks such as SETI@home \cite{javadi2009mining}; we obviously do
+not pretend to compare our system in scale or complexity with it, only
+to point out that the behavior of volunteer computing nodes follows a
+certain pattern, found in SETI@home, and which appears also in our system. This
 distribution is governed by three parameters, the usual location $\mu$

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 % (Paloma) I would propose "a long way to go in order to become"
- a mainstream technology in
-the same way the cloud technologies in general have. The lack of 
-% (Paloma) Add "both" here to better complete the sentence
-general
-purpose and open source tools
-with analytics built in
+ a mainstream
+technology in the same way the cloud technologies in general have. The
+lack of % (Paloma) Add "both" here to better complete the sentence
+general purpose and open source tools with analytics built in
 % (Paloma) "built-in" is an adjective and, as such, has to be written before names
- could be one of the reasons why this is so, as
-well as the lack of proofs of concept. That is why in this paper we
-present {\sf NodIO}, a framework used to create client-server architectures for distributed
+could be one of the reasons why this is so, as well as the lack of
+proofs of concept. That is why in this paper we present {\sf NodIO}, a
+framework used to create client-server architectures for distributed
 evolutionary algorithms, written entirely % (Paloma) Adverb. I would also add an "and" after the comma
@@
 asynchronous, distributed evolutionary algorithms running in the
-browser.  The architecture, along with some initial results on the
-expected performance and scaling
-behavior, is presented. We also use the algorithm data to
-find out what kind of patterns the user behavior follow and try to fit
-them to some statistical distribution.
+browser, as well as any other client consuming its REST application
+programming interface.  The architecture, along with some initial
+results on the expected performance and scaling behavior when the
+number of users is increased is presented. We also use
+the algorithm data to find out what kind of patterns the user behavior
+follow by examining the resources they devote to our experiment, and
+try to fit them to some statistical distribution to create the bases
+of a model of user behavior in volunteer computing experiments of this
+kind that can be used by researchers to predict its performance. 
+In general, our experiments prove, first, that NodIO is a flexible and
+low-overhead platform to perform distributed computation experiments
+that include volunteers, and then that user behavior follows some patterns,
+and, thus, are amenable to a model.
 \end{abstract}

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Minor changes
Refs #31 mainly.

@@
 
+We have added a sentence clarifying that our intention was simply to
+point out that the pattern is similar. We are obviously in a wholly
+different league.
+
 \section{Response to Reviewer \#4}
@@
 applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
-believe in the possibility of "potentially" massive computation as sketched by the authors.
+believe in the possibility of "potentially" massive computation as
+sketched by the authors.
+\end{quote}
+
+JavaScript is not, in fact, so slow and to prove its speed we have
+published a technical report comparing it with other compiled
+languages (such as Scala) and scripted languages
+\cite{2015arXiv151101088M}. In fact, node.js (used for the benchmarks)
+and the JavaScript virtual machine in Chrome (used in Android phones
+and Chrome and Chromium browsers) share the same JIT compiler, so
+their speeds are virtually the same. In general, as a conclusion of
+the report we see that the real difference between scripted and
+compiled languages will vary with the language, but also with the
+particular implementation of the algorithm and the size of the problem
+or chromosome you are using. Even if the speed for a particular
+problem is two or five times that achiever for JavaScript, this
+disadvantage is more than offset by the availability of that free
+resource in abundance; that is, you will only need that many
+volunteers cooperating to get better performance than what you would
+get with C or Java. Even so, this NodIO architecture, as highlighted
+in the new version of the paper, is pool-based so local sources using
+high-speed devices such as GPUs or even local powerful CPUs can be
+added as non-volunteer users, since the heterogeinity and simple API
+of the system allows that by design.
+
+\begin{quote}
 The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
@@
 
+\bibliographystyle{IEEEtran}
+\bibliography{geneura}
+
 \end{document}
+%%% Local Variables:
+%%% ispell-local-dictionary: "english"
+%%% End:

### 2015-11-17 data/plot-rastrigin-fitness.R
Working on Rastrigin Fitness

@@ -1,0 +1,7 @@
+library(ggplot2)
+library(scales)
+
+rastrigin.fitness <- read.csv("rastrigin-fitness.csv")
+rastrigin.fitness$date <-  as.POSIXct(rastrigin.fitness$timestamp, "%Y-%m-%d %H:%M:%S")
+qplot(date,-Fitness,data=rastrigin.fitness)+ scale_x_datetime()
+ggsave("rastrigin-fitness.png",width=4,height=3)

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

@@
 
-In general, a volunteer-based distributed evolutionary computation
+In general, a distributed volunteer-based evolutionary computation
 system based on the browser is simply a client-server system
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; that is
+be some issues with the performance of the language itself; that % (Paloma) I recommend the use of "which" here
+ is
 why we have made a comparison between JavaScript and other languages
 \cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
-faster than compiled languages such as Scala and, in any case, provide
-a performance that is comparable to other languages employed usually
-in evolutionary computation and numerical optimization such as Python. Compiled languages such as
-Java or C might be, in fact, faster, but the fact that a single-user
+faster than compiled languages such as Scala and, in any case, provides
+a performance that is comparable to other languages employed usually % (Paloma) Adverb
+in evolutionary computation and numerical optimization such as Python. 
+% (Paloma) Bit of too long sentence here. Why not changing the ; for a . ?
+Compiled languages such as
+Java or C might be, in fact, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+ that a single-user
 performance is higher is more than offset by the fact that the number
@@
 using their mobile phones or tablets, is several orders of magnitude
-higher, which is the objective in this kind of systems. JavaScript is
+higher, which is the objective in this kind of systems.
+% (Paloma) This sentence needs re-writing
+% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset.
+This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
+ JavaScript is
 quite a popular language nowadays, providing compilers from a number
@@
 many other outside, so researchers can, in fact, write their fitness
-function in Python, Erlang, Scala or even Java, and them compile it to
+function in Python, Erlang, Scala, or even Java, and them compile it to
 JavaScript. For completness, an experiment comparing the performance of
@@
 \item Participants must also trust the application provider, 
-regarding invasion of privacy, the integrity of their devices 
+regarding invasion of privacy, the integrity of their devices, 
 and intended application. 
@@
 crafting a fake request which, for instance, assigns a fake fitness to
-a particular chromosome. Another two different attacks are also
+a particular chromosome
+% (Paloma) Where is the second kind of threat?
+. Another two different attacks are also
 possible: a denial of service attack as well as penetration
@@
 sources, client and server, for the application, are released
-openly. We also release as open access experiments performed
+openly. % (Paloma) Adverb
+ We also release as open access experiments performed
 \cite{DBLP:journals/corr/GuervosG15} and all data is also, after
-cleaning it from personal information, released as open data. This
+cleaning it from personal information, released % (Paloma) Not sure if this "released" should be left and the first one in the sentence, erased...
+ as open data. This
 builds trust between the user and the scientist, which indeed has an
@@
 performance of the system. This socio-technical approach is, indeed,
-coherent with our whole approach to the study of a volunteer
-distributed system which can only be study approaching it from the
+coherent with our whole approach to the study of a distributed volunteer
+ system which can only be studied by approaching it from the
 technical as well as the social, as in social media, angle.
@@
 cloud or bare metal based volunteer evolutionary computing system
-based on the {\sf NodEO} library, whose architecture has been
+based on % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
+ the {\sf NodEO} library, whose architecture has been
 developed using JavaScript on the client as well as the server.
@@
   that can be called for storing and retrieving information (the CRUD cycle:
-  create, request, update and delete) from the server. 
+  create, request, update, and delete) from the server. 
   A JSON data format is used for the communication between 

### 2015-11-17 data/IPs-per-minute.pl
Checking out IPs per minute

@@ -1,0 +1,23 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use v5.14;
+
+use File::Slurp::Tiny qw(read_lines);
+
+my @lines = read_lines("rastrigin-IPs.csv");
+
+my %IPs_minute;
+for my $l (@lines[1..$#lines]) {
+  my ($IP, $time ) = split(/,\s+/,$l);
+  my ($minute) = ($time =~ /(\d+-\d+-\d+ \d+:\d+)/ );
+  $IPs_minute{$minute}{$IP}++;
+}
+
+for my $t ( sort { $a cmp $b } keys %IPs_minute ) {
+  say "$t, ", scalar keys %{$IPs_minute{$t}};
+}
+
+  

### 2015-11-17 data/plot-rastrigin-IPs.R

@@ -1,0 +1,6 @@
+library(ggplot2)
+
+rastrigin.IPs <- read.csv("rastrigin-IPs-per-minute.csv")
+rastrigin.IPs$minute <-  as.POSIXct(rastrigin.IPs$Minute, "%Y-%m-%d %H:%M")
+ggplot(data=rastrigin.IPs, aes(x=minute,y=IPs))+ scale_x_datetime()+geom_point()
+ggsave("rastrigin-IPs.png",width=4,height=3)

### 2015-11-17 data/plot-rastrigin-fitness.R

@@ -1,5 +1,4 @@
 library(ggplot2)
-library(scales)
 
 rastrigin.fitness <- read.csv("rastrigin-fitness.csv")
 rastrigin.fitness$date <-  as.POSIXct(rastrigin.fitness$timestamp, "%Y-%m-%d %H:%M:%S")

### 2015-11-17 data/IPs-per-hour.pl
Adds processed Rastring IPs per hour

@@ -1,0 +1,23 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use v5.14;
+
+use File::Slurp::Tiny qw(read_lines);
+
+my @lines = read_lines("rastrigin-IPs.csv");
+
+my %IPs_hour;
+for my $l (@lines[1..$#lines]) {
+  my ($IP, $time ) = split(/,\s+/,$l);
+  my ($hour) = ($time =~ /(\d+-\d+-\d+ \d+)/ );
+  $IPs_hour{$hour}{$IP}++;
+}
+
+for my $t ( sort { $a cmp $b } keys %IPs_hour ) {
+  say "$t:00:00, ", scalar keys %{$IPs_hour{$t}};
+}
+
+  

### 2015-11-17 data/plot-rastrigin-IPs.R

@@ -4,3 +4,8 @@ rastrigin.IPs <- read.csv("rastrigin-IPs-per-minute.csv")
 rastrigin.IPs$minute <-  as.POSIXct(rastrigin.IPs$Minute, "%Y-%m-%d %H:%M")
 ggplot(data=rastrigin.IPs, aes(x=minute,y=IPs))+ scale_x_datetime()+geom_point()
 ggsave("rastrigin-IPs.png",width=4,height=3)
+
+rastrigin.IPs.hour <- read.csv("rastrigin-IPs-per-hour.csv")
+rastrigin.IPs.hour$hour <-  as.POSIXct(rastrigin.IPs.hour$Hour, "%Y-%m-%d %H:%M:%S")
+ggplot(data=rastrigin.IPs.hour, aes(x=hour,y=IPs))+ scale_x_datetime()+geom_line()
+ggsave("rastrigin-IPs-hour.png",width=4,height=3)

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Rastrigin Setup

@@
 In JavaScript the {\tt Date class} is often used 
-to measure execution time, but has the limitation of having a maximum resolution given in milliseconds.
+to measure execution time, but has the limitation of having a maximum 
+resolution given in milliseconds.
 In order to measure the runtime intervals with a higher precision two functions
 where used: for Node.js we used the native {\tt process.hrtime()} function
-which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array and for measuring in the browser, the {\tt Performance.now()} function was used,
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
+array and for measuring in the browser, the {\tt Performance.now()} function was used,
 this function uses floating-point numbers with up to microsecond precision. 
@@
 algorithms. As for the performance of JavaScript against Java and Matlab
-we conducted first the basic test described earlier in a 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
+we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
+Intel Xeon E5 processor running OS X 10.10.5, 
 using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
-MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
-In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
-991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a single Web Worker and 1,270ms using Node.js. 
-The results also show that there is not much difference between the Node.js implementation an that of a
-browser.
-There is also not much difference between running the code in the main thread or in Web Workers.
-When running the experiment in Firefox a dialog alerted that the script was taking to much time, showing that this
-kind of processing is not intended for the main thread. Also there is not much overhead when running two
+MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
+Google Chrome Version 46.0.2490.86 (64-bit) 
+and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in
+Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
+performance with an average of 935ms, followed by Java with
+991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a 
+single Web Worker and 1,270ms using Node.js. 
+The results also show that there is not much difference between the Node.js
+implementation an that of a
+browser. There is also not much difference between running the code in the main thread or in Web
+Workers. When running the experiment in Firefox a dialog alerted that the
+script was taking to much time, showing that this
+kind of processing is not intended for the main thread. Also there is not much overhead
+when running two
 Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
@@
 \subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
+This experiment was deployed in a virtual private server with 512MB of memory
+runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
+experiments where the optimum was espected to be found many times, 
+this problem was expected to run constantly for many days finding only
+sub-optimal aproximations. In order to achieve this, the only change needed 
+in the system was to limit the number of chromosmes kept in the pool. 
+Previously, every PUT request which included a new chromose increased 
+the size of the pool, in a long-running experiment this has to be avoided
+in order to not exahust the available memory. In this experiment the size
+of a chromosme was 1000 double precision numbers, considereing this size
+of the pool was limited to 10,000 chromosomes.  In order to keep this size
+constant a random chromosme was pulled out of the pool before inserting a
+new one, when the pool was full.
 

### 2015-11-17 special-issue-ci/Response to reviewers.tex
Rastrigin-fitness
fixing a typo

@@
 or chromosome you are using. Even if the speed for a particular
-problem is two or five times that achiever for JavaScript, this
+problem is two or five times that achieved for JavaScript, this
 disadvantage is more than offset by the availability of that free
@@
 
+We have extended the experiments section with a new one, based on the
+hard Rastrigint problem.
+
 \bibliographystyle{IEEEtran}

### 2015-11-17 special-issue-ci/Response to reviewers.tex
fixing a typo

@@
 explained a bit better in the introduction and conclusions
-section. Besides, we have added a more difficult probmem, the shifted
+section. Besides, we have added a more difficult problem, the shifted
 Rastrigin function, to make measures in a different scenario (one in

### 2015-11-17 special-issue-ci/Response to reviewers.tex
completing a sentence

@@
 section. Besides, we have added a more difficult problem, the shifted
-Rastrigin function, to make measures in a different scenario (one in
-which the problem is not solved in a few seconds).
+Rastrigin function (a well known function optimization problem), to make 
+measures in a different scenario (one in which the problem is not solved 
+in a few seconds).
 

### 2015-11-17 special-issue-ci/Response to reviewers.tex
fixing a typo

@@
 any special provision for new systems. This has also been explained in
-th expanded text.
+the expanded text.
 
@@
 \end{quote}
-I fact 3.46 is the average for all 50 runs. We have enphasized the
-word average to make that clearer.
+
+I fact 3.46 is the average for all 50 runs. 
+We have emphasized the word average to make that clearer.
 

### 2015-11-17 special-issue-ci/Response to reviewers.tex
fixing a typo

@@
 Finally, since the reviewer thinks that length is fine and for the
-reasons outlined above, we have decided to keep the graphs and eplain
+reasons outlined above, we have decided to keep the graphs and explain
 them a bit better.
@@
 
-we have extended the abstract stressing our main points, that were
+We have extended the abstract stressing our main points, that were
 first to test the validity and measure the performance of our
@@
 enough with {\em most}. Windows Phone, for instance and maybe
-fortunately, is not too common.
+fortunately, is not too common. % ¿no existe Chrome para Windows Phone? De todas formas yo quitaría lo de "and maybe fortunately"...
 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing a typo in the paper

@@
 to the experiment at the same time, so that our objective was
-achieved. In general, one of the first conclusion for this paper is
+achieved. In general, one of the first conclusions for this paper is
 that the browser technologies should be used to its full extent so

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
a comment to the Conclusions section

@@
 performance is an interesting avenue to explore in the future. Even if
-the three previous experiments were published in a similar, way, one
+the three previous experiments were published in a similar way, one
 obtained up to 5 times more total cycles  that the one with the least
@@
 algorithmic variant using random population size. We do not think this
-has had a big influence on the result, but it would be interesting to
+has had a big influence on the results, but it would be interesting to
 measure exactly what this influence has been. Adding a bit of more
@@
 Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman might be used
+programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
 to create a generator in which the user just has to create a fitness

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Setup for 3th experiment

@@
 in order to not exahust the available memory. In this experiment the size
-of a chromosme was 1000 double precision numbers, considereing this size
-of the pool was limited to 10,000 chromosomes.  In order to keep this size
+of a chromosme was 1000 double precision numbers, considereing this
+and the memory available in the server, the the pool was limited 
+to 10,000 chromosomes.  In order to keep this size
 constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full.
+new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
+and \tt{ classic-ea-float} modules where used. The parameters for the ea algorithm 
+were tournament size = 3 and  population size = 500, returning exchanging a
+chromosme with the server every 100 generations. Again two workers in each window.
+As before, a few participation calls were published in throu posts in social networks.
 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
a comment to the results section
Results corrected

@@
 performance with an average of 935ms, followed by Java with
-991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a 
-single Web Worker and 1,270ms using Node.js. 
+991ms, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
+single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
 The results also show that there is not much difference between the Node.js

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some corrections

@@
 new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
-and \tt{ classic-ea-float} modules where used. The parameters for the ea algorithm 
+and \tt{ classic-ea-float} modules were used. The parameters for the ea algorithm 
 were tournament size = 3 and  population size = 500, returning exchanging a
 chromosme with the server every 100 generations. Again two workers in each window.
-As before, a few participation calls were published in throu posts in social networks.
+As before, a few participation calls were published in social networks.
 

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Just a correction
Starting last lines

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 In JavaScript the {\tt Date class} is often used 
-to measure execution time, but has the limitation of having a maximum 
-resolution given in milliseconds.
+to measure execution time, but has the limitation of having a maximum resolution given in milliseconds.
 In order to measure the runtime intervals with a higher precision two functions
 where used: for Node.js we used the native {\tt process.hrtime()} function
-which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
-array and for measuring in the browser, the {\tt Performance.now()} function was used,
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array and for measuring in the browser, the {\tt Performance.now()} function was used,
 this function uses floating-point numbers with up to microsecond precision. 
@@
 algorithms. As for the performance of JavaScript against Java and Matlab
-we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
-Intel Xeon E5 processor running OS X 10.10.5, 
+we conducted first the basic test described earlier in a 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
 using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
-MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
-Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in
-Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
-performance with an average of 935ms, followed by Java with
-991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a 
-single Web Worker and 1,270ms using Node.js. 
-The results also show that there is not much difference between the Node.js
-implementation an that of a
-browser. There is also not much difference between running the code in the main thread or in Web
-Workers. When running the experiment in Firefox a dialog alerted that the
-script was taking to much time, showing that this
-kind of processing is not intended for the main thread. Also there is not much overhead
-when running two
+MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
+and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
+In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
+991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a single Web Worker and 1,270ms using Node.js. 
+The results also show that there is not much difference between the Node.js implementation an that of a
+browser.
+There is also not much difference between running the code in the main thread or in Web Workers.
+When running the experiment in Firefox a dialog alerted that the script was taking to much time, showing that this
+kind of processing is not intended for the main thread. Also there is not much overhead when running two
 Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
-language for developing complex mathematical functions for running in the browser. 
-\begin{figure}[!htb]
+language for developing complex mathematical functions for running in
+the browser. 
+%
+\begin{figure}[h!tb]
 \centering
@@
 \subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
-This experiment was deployed in a virtual private server with 512MB of memory
-runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
-experiments where the optimum was espected to be found many times, 
-this problem was expected to run constantly for many days finding only
-sub-optimal aproximations. In order to achieve this, the only change needed 
-in the system was to limit the number of chromosmes kept in the pool. 
-Previously, every PUT request which included a new chromose increased 
-the size of the pool, in a long-running experiment this has to be avoided
-in order to not exahust the available memory. In this experiment the size
-of a chromosme was 1000 double precision numbers, considereing this
-and the memory available in the server, the the pool was limited 
-to 10,000 chromosomes.  In order to keep this size
-constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
-and \tt{ classic-ea-float} modules where used. The parameters for the ea algorithm 
-were tournament size = 3 and  population size = 500, returning exchanging a
-chromosme with the server every 100 generations. Again two workers in each window.
-As before, a few participation calls were published in throu posts in social networks.
 
 \subsubsection{Results}
+
+The results for this new experiment have been extremely interesting
+and quite different from the previous one.
 
@@
 to the experiment at the same time, so that our objective was
-achieved. In general, one of the first conclusions for this paper is
+achieved. In general, one of the first conclusion for this paper is
 that the browser technologies should be used to its full extent so
@@
 performance is an interesting avenue to explore in the future. Even if
-the three previous experiments were published in a similar way, one
+the three previous experiments were published in a similar, way, one
 obtained up to 5 times more total cycles  that the one with the least
@@
 algorithmic variant using random population size. We do not think this
-has had a big influence on the results, but it would be interesting to
+has had a big influence on the result, but it would be interesting to
 measure exactly what this influence has been. Adding a bit of more
@@
 Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
+programming and also ease of use. Tools such as Yeoman might be used
 to create a generator in which the user just has to create a fitness

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

--- 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 when running two
-Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
-language for developing complex mathematical functions for running in the browser. 
+Web Workers in parallel as they took 1279ms each. This experiment
+shows that JavaScript is viable as a language
+language for developing complex mathematical functions an these can run in a browser
+with acceptable performance. 
 \begin{figure}[!htb]
@@
 
+The results for this new experiment have been extremely interesting
+and quite different from the previous one.
+
 %---------------------------------------------------------------

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds graphics refs #41

@@
 \subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigins Function}
-JavaScript has been traditionally implemented as an interpreted language not designed for the development of high performance systems. On the other hand, current JavaScript 
+JavaScript has been traditionally implemented as an interpreted
+language not designed for the development of high performance
+systems. On the other hand, current JavaScript  
 virtual machines (VMs) are closing the gap, for instance the Google V8 
@@
 constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
-and \tt{ classic-ea-float} modules were used. The parameters for the ea algorithm 
+new one, when the pool was full. For this problem NodEO's {\tt chromosme-float} 
+and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 
 were tournament size = 3 and  population size = 500, returning exchanging a
@@
 
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics{rastrigin-fitness.png}
+\caption{Fitness of chromosomes sent to the server pool along time in
+  the $x$ axis.} 
+% Plotted with ../data/plot-rastrigin-fitness.R
+\label{fig:puts:rastrigin}
+\end{figure}
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
+\includegraphics[width=0.49\linewidth]{rastrigin-ips-hour.png}
+\caption{Number of unique IPs participating in the Rastrigin
+  experiment per minute (left) and per hour (right)} 
+% Plotted with ../data/plot-rastrigin-IPs.R
+\label{fig:ips:rastrigin}
+\end{figure}
+
 The results for this new experiment have been extremely interesting
-and quite different from the previous one.
+and quite different from the previous one. The most important
+characteristic that set this experiment apart from the previous one is
+the fact that it did not find the solution for the time it was
+running, instead lowering, little by little, the fitness until it was
+disconnected. The fitness of chromosomes sent by clients is shown in
+Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
+upper part, which represents the fitness of clients that have just
+started evolution from a random population, and then the fitness of
+those clients that have been already evolving for a while or obtained
+a high-fitness individual from the pool and evolved from there, and
+which managed to surpass the 9000 value by the end of the
+simulation. It incidentally shows also the number of {\em new} nodes
+coming: there is a dense {\em cloud} at the beginning, but the new
+contributions keep thinning as the evolution progresses.
 

### 2015-11-18 data/ips-puts-rastrigin.R
New plots

@@ -1,0 +1,53 @@
+library(ggplot2)
+library(evd)
+library(fitdistrplus)
+
+ips.puts.ww.rastrigin <- read.csv("IPs-Rastrigin.csv")
+ips.puts.ww.rastrigin.df <- data.frame(rank=1:nrow(ips.puts.ww.rastrigin),
+                             puts=sort(ips.puts.ww.rastrigin$puts,decreasing=T))
+this.fit <- fgev(ips.puts.ww.rastrigin$puts)
+print(this.fit)
+this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
+                            y=sort(rgev(length(ips.puts.ww.rastrigin$puts),
+                                this.fit$estimate["loc"],
+                                this.fit$estimate["scale"],
+                                this.fit$estimate["shape"]),decreasing=T))
+
+ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+
+ggsave('gev-fit-ww.rastrigin.png',width=5,height=5)
+
+print(fitdistr(ips.puts.ww.rastrigin$puts,"gamma"))
+this.weibull <- fitdistr(ips.puts.ww.rastrigin$puts,"weibull")
+
+this.fit.weib <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
+y=sort(rweibull(length(ips.puts.ww.rastrigin$puts),
+    this.weibull$estimate["shape"],
+    this.weibull$estimate["scale"]),decreasing=T))
+ggplot()+geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
[…]

### 2015-11-18 data/ips-puts.R

@@ -12,7 +12,9 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww$puts),
                                 this.fit$estimate["loc"],
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
+
 ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+
 ggsave('gev-fit-ww.png',width=5,height=5)
 
 print(fitdistr(ips.puts.ww$puts,"gamma"))

### 2015-11-18 data/ips-puts-rastrigin.R
Adds data on fit

@@ -29,8 +29,8 @@ ggsave('weibull-fit-ww.rastrigin.png',width=5,height=5)
 
 ips.puts.ww.rastrigin.workers <- read.csv("IPs-worker-Rastrigin.csv")
 ips.puts.ww.rastrigin.workers.df <- data.frame(rank=1:nrow(ips.puts.ww.rastrigin.workers),
-                             puts=sort(ips.puts.ww.rastrigin.workers$puts,decreasing=T))
-this.fit <- fgev(ips.puts.ww.rastrigin.workers$puts)
+                                               puts=sort(ips.puts.ww.rastrigin.workers$puts,decreasing=T))
+this.fit <- fgev(ips.puts.ww.rastrigin.workers$puts,std.err=FALSE)
 print(this.fit)
 this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin.workers$puts),
                             y=sort(rgev(length(ips.puts.ww.rastrigin.workers$puts),
@@ -49,5 +49,6 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww.rastrigin.workers$puts),
 y=sort(rweibull(length(ips.puts.ww.rastrigin.workers$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
+print(this.weibull)
 ggplot()+geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
 ggsave('weibull-fit-ww.rastrigin.workers.png',width=5,height=5)

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds data on fit
Citations added

@@
 V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter.
-[\url{ https://developers.google.com/v8/design#mach_code}].
+There are no intermediate byte codes or an interpreter \cite{Gray:2009:GCM:1610564.1610565}.
 JavaScript has gained popularity in server-side development using runtime
@@
 a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization [] is used in this section. 
+Large-Scale Global Optimization \cite{tang2007benchmark} is used in this section. 
 The function is described next.
@@
 \includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
-\includegraphics[width=0.49\linewidth]{rastrigin-ips-hour.png}
+\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
 \caption{Number of unique IPs participating in the Rastrigin
@@
 \end{figure}
+%
+\begin{table*}
+\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
+  the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
+\begin{center}
+\begin{tabular}{cccc}
+\hline
+Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+\hline
+GEV & 35.759 $\pm$ 0.83719  &  343.336  $\pm$  & 9.877   $\pm$
+0.04746 \\
+Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Abstract & Intro

@@
 \begin{abstract}
-Volunteer computing is used routinely
+Volunteer computing is extensively used 
 % (Paloma) Adverbs are always written before verbs, check this throughout the text
 in science and industry, but it
-still has a long way to go if it is going to become
+still has a long way to go in order to become
 % (Paloma) I would propose "a long way to go in order to become"
@@
 lack of % (Paloma) Add "both" here to better complete the sentence
-general purpose and open source tools with analytics built in
+general purpose and open source tools with built-in analytics 
 % (Paloma) "built-in" is an adjective and, as such, has to be written before names
-could be one of the reasons why this is so, as well as the lack of
+could be one of the reasons of why this is so, as well as the lack of
 proofs of concept. That is why in this paper we present {\sf NodIO}, a
 framework used to create client-server architectures for distributed
-evolutionary algorithms, written entirely % (Paloma) Adverb. I would also add an "and" after the comma
+evolutionary algorithms, entirely written  % (Paloma) Adverb. I would also add an "and" after the comma
 in JavaScript, which needs
@@
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is first
+provisionally named {\sf NodIO}. {\sf NodIO} is basically
 % (Paloma) Is the "first" word really needed? 
+% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
 a two tier client/server 
@@
 % (Paloma) be embedded in?
- embedded any web-enabled device, including many over which 
+be embedded in any web-enabled device, including many over which 
 the developers have no control.
@@
 volunteer and
-unconventional distributed evolutionary computing systems, and we
-decided to design and release a new free framework that would allow anyone to
+unconventional distributed evolutionary computing systems to 
+design and release a new free framework that would allow anyone to
 create a volunteer distributed EC experiment using cloud resources as
@@
 % For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided... 
+% (Mario) We leverage our experience in blah blah to design
 
@@
 that the source code can be directly examined by volunteers, because it is not
-compiled as other options (such as Java Applets, Flash scripts or
-ActiveX)
+compiled as other options such as Java Applets, ActionScript or
+ActiveX.
 % (Paloma) Brackets not really needed here
-. Also, as we discuss in this paper, it can provide some extra
+Also, as we discuss in this paper, it can provide some extra
 features such as enhanced security or operating system independence.
@@
 %PABLO: hay que enlazar con el siguiente párrafo
-Desktop clients have the same issues as any 
-volunteer computing system, including our % (Paloma) their?
- own: these issues are
-related to security as well as
-safety, which introduce 
+%Desktop clients have the same issues as any 
+%volunteer computing system, including our % (Paloma) their?
+% own: these issues are
+%related to security as well as
+%safety, which introduce
+
+When developing volunteer based systems, desingers must
+also consider the whole {\em social} aspect in the design, with
+issues related to security, trust and privacy among others. 
+The computing system becomes a {\em techno-social system} \cite{vespignani2009predicting}.  
 % (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
-a {\em social} aspect to the design of a
-computing system that makes it a {\em techno-social system}
-\cite{vespignani2009predicting}. In this paper, our intention is to
+%a {\em social} aspect to the design of a
+%computing system that makes it a {\em techno-social system}
+In this paper, our intention is to
 present the {\sf NodIO} framework and analyze it from two different
@@
 into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page that runs the system
+that voluntarily choose to enter the web page running the system
 determine its performance and how it can be modeled and, eventually,
@@
 state of the art (Section \ref{sec:soa}) in web-based distributed
-computation systems along with attempts to predict and model its behavior. 
+computational systems along with attempts to predict and model its behavior. 
 Then, a detailed description of the proposed system is presented

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
No more comments for Abstract and Introduction
Re-writing about hierarchy

@@
 % (Mario) We leverage our experience in blah blah to design
+% (Paloma) It may be... guess I was thinking in the noun, not in the transitive verb
 
@@
 volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, safety, anonymity, hierarchy, ease of use and
-reasonable performance. Of all these, probably {\em hierarchy}, that
-is, the capability of using preferentially resources in the same
-organization could be either disregarded or included in adaptivity,
-leaving us with eight desirable properties.
+adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
+reasonable performance.
+But hierarchy, this is, the capability of preferentially using resources in the same organization, is a feature that could be either disregarded or included within adaptivity, which leaves us with eight desirable properties.
 % (Paloma) Re-write.
 For example, "But is hierarchy, this is, the capability of preferentially using resources in the same organization, the one that could be either disregarded or included in adaptivity, which leaves us with eight desirable properties."
+% (Paloma) And I re-wrote what I proposed yesterday.
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changes in State of the art

@@
 languages and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first, 
+installed the required virtual machine plugin first. 
 % (Paloma) These should be separate sentences
-but embedding a
+However, embedding a
 distributed computation client in the browser has the advantage of
@@
 paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that
-already in the early stages of the web the possibility of
+\cite{baratloo1999charlotte}, but it is interesting to note that, even in the early stages of the web, the possibility of 
 % (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
@@
 a comprehensive statistics collection subsystem, allowing its use for science,
-since it allows % (Paloma) Use synonyms, like "yield" or "permit"
+since it yields % (Paloma) Use synonyms, like "yield" or "permit". Repetitive use of "allow" is... repetitive.
  the collection of algorithm analytics by the

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
No more comments in SotA

@@
 \cite{sarmenta-sabotagetolerance}. However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making this effort % (Paloma) making them
+Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
 less widespread.
@@
 computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
-volunteer computing, % (Paloma) Whith regard to?
-number of users and the time spent by every one in the
-computation in browser-based volunteer computing experiments, have
+volunteer computing, with regard to
+number of users and computation time spent by every one 
+in browser-based volunteer computing experiments, have
 only been studied in a limited way in 
@@
 
-As far as we know, this paper presents one of the few experiments that % (Paloma) Change to "which" in order to not repeating "that" all the time
+As far as we know, this paper presents one of the few experiments which % (Paloma) Change to "which" in order to not repeating "that" all the time
 uses computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. The methodology used
-% (Paloma) I would say "the used methodology"
- to
+or workstation computers in a research center. The methodology used to
 gather resources and the algorithms used will be described next, as

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changes in description of framework

@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; that % (Paloma) I recommend the use of "which" here
+be some issues with the performance of the language itself. This
  is
@@
 faster than compiled languages such as Scala and, in any case, provides
-a performance that is comparable to other languages employed usually % (Paloma) Adverb
+a performance that is comparable to other languages usually employed  % (Paloma) Adverb
 in evolutionary computation and numerical optimization such as Python. 
-% (Paloma) Bit of too long sentence here. Why not changing the ; for a . ?
+% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
 Compiled languages such as
-Java or C might be, in fact, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
  that a single-user
-performance is higher is more than offset by the fact that the number
+performance is higher is more than offset. This is because the number
 of spontaneous volunteers we can gather at whim, including people

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some more changes in description of framework

@@
 testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are released
-openly. % (Paloma) Adverb
- We also release as open access experiments performed
-\cite{DBLP:journals/corr/GuervosG15} and all data is also, after
-cleaning it from personal information, released % (Paloma) Not sure if this "released" should be left and the first one in the sentence, erased...
- as open data. This
+sources, client and server, for the application, are openly released. % (Paloma) Adverb
+ We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
 builds trust between the user and the scientist, which indeed has an
@@
 cloud or bare metal based volunteer evolutionary computing system
-based on % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
+derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
  the {\sf NodEO} library, whose architecture has been

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
All NodIO and NodEO words are now with the same sf format

@@
 kind that can be used by researchers to predict its performance. 
-In general, our experiments prove, first, that NodIO is a flexible and
+In general, our experiments prove, first, that {\sf NodIO} is a flexible and
 low-overhead platform to perform distributed computation experiments
@@
 desktop or command line and browser clients, both based in the EC
-library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
+library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
 when writing code and leverages 
@@
   Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of NodIO can also be used.
+  has been disabled, a basic version of {\sf NodIO} can also be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
   has been tested in other high demand systems with success, but 
-  in the case of NodIO, additional experiments are need to asses 
+  in the case of {\sf NodIO}, additional experiments are need to asses 
   the scalability of the system. 
@@
 First, we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using NodEO
+evolutionary algorithm in a desktop client written using {\sf NodEO}
 \cite{nodeo2014}, the basic JavaScript library we have used as base to
@@
 was tested in desktop versions of Chrome, FireFox and Safari.
-\item Each NodEO instance can be restarted independently.
+\item Each {\sf NodEO} instance can be restarted independently.
 \end{itemize}
@@
 in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
-API to run two NodEO instances per browser. The sequence diagram shows the
+API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
 interaction between processes in a time sequence. In this system there are
@@
 
-\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
+\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO}-W$^2$ }
 This experiment was deployed in a virtual private server with 512MB of memory
@@
 constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem NodEO's {\tt chromosme-float} 
+new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosme-float} 
 and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some more changes in description of NodIO architecture

@@
 when writing code and leverages 
-expertise acquired
-% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
- in the library itself.
+expertise acquired in the library itself.
 
@@
 faster than compiled languages such as Scala and, in any case, provides
-a performance that is comparable to other languages usually employed  % (Paloma) Adverb
+a performance that is comparable to other languages usually employed  
 in evolutionary computation and numerical optimization such as Python. 
@@
 testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are openly released. % (Paloma) Adverb
+sources, client and server, for the application, are openly released. 
  We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
@@
   to the performance and state of the experiments. It also performs logging
-  duties, but is basically a very lightweight and high performance
-  data store \cite{jj:idc:lowcost}.
+  duties, but they are basically a very lightweight and high performance
+  data storage \cite{jj:idc:lowcost}.
   The server has the capability to
@@
 \item A client that includes the evolutionary algorithm as
-  JavaScript code embedded in a web page that displays graphs and some
-  additional links and information on the experiment. This code runs
+  JavaScript code embedded in a web page that displays graphs, some
+  additional links, and information on the experiment. This code runs
   an evolutionary algorithm {\em island} starting with a random
-  population and then sending, every 100 generations, the best individual
-  back to the server (via a {\tt PUT} request) and requesting a random
-  individual from the server (via a {\tt GET} request).
+  population, then it sends, every 100 generations, the best individual
+  back to the server (via a {\tt PUT} request), and requests a random
+  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
 \end{enumerate}

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Big doubt about a used verb

@@
   the one described in \cite{LNCS86720702}, any kind of client that
-  works the application programming interface (API) can be used,
+  works the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
+But never heard of a client which "works" an API.
+Is this correct?
   written in any kind of language. However, in this paper we are going

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changes in outlined challenges list

@@
 function and handing at the creation of the algorithm object, called
-{\tt Classic} is the only change needed to work with a different
+{\tt Classic}, is the only change needed to work with a different
 problem. Let us see how this system addresses the different challenges
@@
 \item The system is {\em heterogeneous} since it does not need any
-  performance, operating system or even browser requirement (as long
-  as Javascript is enabled): anyone
+  performance, operating system, or even browser requirement as long
+  as Javascript is enabled: anyone
   visiting the page, even from mobile devices, can load the algorithm.
@@
 \item {\em Adaptiveness} is achieved simply through the autonomous
-  operation of every individual island with out any synchronization
+  operation of every individual island without any synchronization
   mechanism. The islands in the system are, in fact, unaware of each
@@
   has been tested in other high demand systems with success, but 
-  in the case of {\sf NodIO}, additional experiments are need to asses 
-  the scalability of the system. 
+  in the case of {\sf NodIO}, additional experiments are needed to asses 
+  the scalability of the system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
 \item Since the algorithm runs in the browser, {\em safety} is
@@
   that there is no unsafe access either to their local files or even
-  to more resources that the browser should be allotted.
+  to more resources than the browser should be allotted.
 \item Running the algorithm is just a matter of loading the page,
-  which means that operation is totally {\em anonymous}. For the same
+  which makes the operation totally {\em anonymous}. For the same
   reason, {\em ease of use} is optimal, being as easy as simply
-  clicking on an URL and available to anyone with access to a browser.
+  clicking on an URL, available to anyone with access to a browser.
 \item {\em Reasonable performance} is not ensured. In fact, we should

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Otro poquito

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 kind that can be used by researchers to predict its performance. 
-In general, our experiments prove, first, that {\sf NodIO} is a flexible and
+In general, our experiments prove, first, that NodIO is a flexible and
 low-overhead platform to perform distributed computation experiments
@@
 be embedded in any web-enabled device, including many over which 
-the developers have no control.
+developers have no control.
 
@@
 % (Mario) We leverage our experience in blah blah to design
-% (Paloma) It may be... guess I was thinking in the noun, not in the transitive verb
 
@@
 volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
-reasonable performance.
-But hierarchy, this is, the capability of preferentially using resources in the same organization, is a feature that could be either disregarded or included within adaptivity, which leaves us with eight desirable properties.
+adaptive parallelism, safety, anonymity, hierarchy, ease of use and
+reasonable performance. Of all these, probably {\em hierarchy}, that
+is, the capability of using preferentially resources in the same
+organization could be either disregarded or included in adaptivity,
+leaving us with eight desirable properties.
 % (Paloma) Re-write.
 For example, "But is hierarchy, this is, the capability of preferentially using resources in the same organization, the one that could be either disregarded or included in adaptivity, which leaves us with eight desirable properties."
-% (Paloma) And I re-wrote what I proposed yesterday.
 
@@
 languages and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first. 
+installed the required virtual machine plugin first, 
 % (Paloma) These should be separate sentences
-However, embedding a
+but embedding a
 distributed computation client in the browser has the advantage of
@@
 paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that, even in the early stages of the web, the possibility of 
+\cite{baratloo1999charlotte}, but it is interesting to note that,
+even in the early stages of the web, the possibility of
 % (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
@@
 the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, allowing its use for science,
-since it yields % (Paloma) Use synonyms, like "yield" or "permit". Repetitive use of "allow" is... repetitive.
+a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
+since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
 the collection of algorithm analytics by the
@@
 distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. It is focused
+brokers for resource discovery. SuperWeb is focused
 % (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
@@
 browsers was realized
-later on \cite{sarmenta-bayanihan} as well as the potential for
-mischief of users with code in their hands
+later on \cite{sarmenta-bayanihan} as well as the potential of
+user misconduct when having access to the source code.
 % (Paloma) I think this sentence is extremely informal
 \cite{sarmenta-sabotagetolerance}. However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
+Sarmenta once again used Java and not JavaScript, making this effort % (Paloma) making them
 less widespread.
@@
 desktop or command line and browser clients, both based in the EC
-library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
+library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
 when writing code and leverages 
-expertise acquired in the library itself.
+the expertise acquired
+% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
+ in the library itself.
 
@@
 al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
-volunteer computing, with regard to
-number of users and computation time spent by every one 
-in browser-based volunteer computing experiments, have
+computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
+%But the raw material of           
+%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
+But some of the essential metrics in volunteer computing like the
+number of users or the time spent by every one in the
+computation in browser-based volunteer computing experiments, have
 only been studied in a limited way in 
@@
 
-As far as we know, this paper presents one of the few experiments which % (Paloma) Change to "which" in order to not repeating "that" all the time
-uses computational resources that are as dissimilar as smartphones and
+As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
+computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. The methodology used to
-gather resources and the algorithms used will be described next, as
-well as the results obtained with the initial setup.
+or workstation computers in a research center. 
+%The methodology used
+% (Paloma) I would say "the used methodology"
+The algorithms used, as well as the methodology 
+for gathering resources will be described next, 
+together with the results obtained in this initial setup.
 
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself. This
+be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
  is
 why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
-faster than compiled languages such as Scala and, in any case, provides
-a performance that is comparable to other languages usually employed  
+\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
+faster than compiled languages such as Scala and, in any case, it provides
+a performance that is comparable to other languages usually employed % (Paloma) Adverb
 in evolutionary computation and numerical optimization such as Python. 
-% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
+% (Paloma) Bit of too long sentence here. Why not changing the ; for a . ?
 Compiled languages such as
-Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
- that a single-user
-performance is higher is more than offset. This is because the number
-of spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets, is several orders of magnitude
+Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+considering the performance of a single-user, 
+but this is more than offset by the computational power of
+the spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets. Together, the performance is several orders of magnitude
 higher, which is the objective in this kind of systems.
@@
 testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are openly released. 
- We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
+sources, client and server, for the application, are released
+openly. % (Paloma) Adverb
+ We also release as open access experiments performed
+\cite{DBLP:journals/corr/GuervosG15} and all data is also, after
+cleaning it from personal information, released % (Paloma) Not sure if this "released" should be left and the first one in the sentence, erased...
+ as open data. This
 builds trust between the user and the scientist, which indeed has an
@@
 cloud or bare metal based volunteer evolutionary computing system
-derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
+based on % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
  the {\sf NodEO} library, whose architecture has been
@@
   to the performance and state of the experiments. It also performs logging
-  duties, but they are basically a very lightweight and high performance
-  data storage \cite{jj:idc:lowcost}.
+  duties, but is basically a very lightweight and high performance
+  data store \cite{jj:idc:lowcost}.
   The server has the capability to
@@
 \item A client that includes the evolutionary algorithm as
-  JavaScript code embedded in a web page that displays graphs, some
-  additional links, and information on the experiment. This code runs
+  JavaScript code embedded in a web page that displays graphs and some
+  additional links and information on the experiment. This code runs
   an evolutionary algorithm {\em island} starting with a random
-  population, then it sends, every 100 generations, the best individual
-  back to the server (via a {\tt PUT} request), and requests a random
-  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
+  population and then sending, every 100 generations, the best individual
+  back to the server (via a {\tt PUT} request) and requesting a random
+  individual from the server (via a {\tt GET} request).
 \end{enumerate}
@@
   Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of {\sf NodIO} can also be used.
+  has been disabled, a basic version of NodIO can also be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
   has been tested in other high demand systems with success, but 
-  in the case of {\sf NodIO}, additional experiments are need to asses 
+  in the case of NodIO, additional experiments are need to asses 
   the scalability of the system. 
@@
 First, we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using {\sf NodEO}
+evolutionary algorithm in a desktop client written using NodEO
 \cite{nodeo2014}, the basic JavaScript library we have used as base to
@@
 was tested in desktop versions of Chrome, FireFox and Safari.
-\item Each {\sf NodEO} instance can be restarted independently.
+\item Each NodEO instance can be restarted independently.
 \end{itemize}
@@
 in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
-API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
+API to run two NodEO instances per browser. The sequence diagram shows the
 interaction between processes in a time sequence. In this system there are
@@
 
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO}-W$^2$ }
+\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
 This experiment was deployed in a virtual private server with 512MB of memory
@@
 constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosme-float} 
+new one, when the pool was full. For this problem NodEO's {\tt chromosme-float} 
 and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Finishing reviewing description of framework

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 anonymous, since no accounts are needed to participate in the
-experiment, it is safe for the user, since we are using a the browser
+experiment, it is safe for the user, since we are using the browser
 black box, and it is also heterogeneous. We have designed the
@@
 experiments and clients, scalability with the number of clients and
-the problem size, and also measured its performance comparing it with
+the problem size, and measured performance by comparing it with
 a baseline configuration, with which, as indicated above, is

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
adds figures

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 \begin{abstract}
-Volunteer computing is extensively used 
+Volunteer computing is used routinely
 % (Paloma) Adverbs are always written before verbs, check this throughout the text
  in science and industry, but it
-still has a long way to go in order to become
+still has a long way to go if it is going to become
 % (Paloma) I would propose "a long way to go in order to become"
@@
 lack of % (Paloma) Add "both" here to better complete the sentence
-general purpose and open source tools with built-in analytics 
+general purpose and open source tools with analytics built in
 % (Paloma) "built-in" is an adjective and, as such, has to be written before names
-could be one of the reasons of why this is so, as well as the lack of
+could be one of the reasons why this is so, as well as the lack of
 proofs of concept. That is why in this paper we present {\sf NodIO}, a
 framework used to create client-server architectures for distributed
-evolutionary algorithms, entirely written  % (Paloma) Adverb. I would also add an "and" after the comma
+evolutionary algorithms, written entirely % (Paloma) Adverb. I would also add an "and" after the comma
 in JavaScript, which needs
@@
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is basically
+provisionally named {\sf NodIO}. {\sf NodIO} is first
 % (Paloma) Is the "first" word really needed?
-% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
  a two tier client/server 
@@
 % (Paloma) be embedded in?
-be embedded in any web-enabled device, including many over which 
-developers have no control.
+ embedded any web-enabled device, including many over which 
+the developers have no control.
 
@@
 volunteer and
-unconventional distributed evolutionary computing systems to 
-design and release a new free framework that would allow anyone to
+unconventional distributed evolutionary computing systems, and we
+decided to design and release a new free framework that would allow anyone to
 create a volunteer distributed EC experiment using cloud resources as
@@
 % For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided...
-% (Mario) We leverage our experience in blah blah to design
 
@@
 that the source code can be directly examined by volunteers, because it is not
-compiled as other options such as Java Applets, ActionScript or
-ActiveX.
+compiled as other options (such as Java Applets, Flash scripts or
+ActiveX)
 % (Paloma) Brackets not really needed here
-Also, as we discuss in this paper, it can provide some extra
+. Also, as we discuss in this paper, it can provide some extra
 features such as enhanced security or operating system independence.
@@
 %PABLO: hay que enlazar con el siguiente párrafo
-%Desktop clients have the same issues as any 
-%volunteer computing system, including our % (Paloma) their?
-% own: these issues are
-%related to security as well as
-%safety, which introduce
-
-When developing volunteer based systems, desingers must
-also consider the whole {\em social} aspect in the design, with
-issues related to security, trust and privacy among others. 
-The computing system becomes a {\em techno-social system} \cite{vespignani2009predicting}.  
+Desktop clients have the same issues as any 
+volunteer computing system, including our % (Paloma) their?
+ own: these issues are
+related to security as well as
+safety, which introduce 
 % (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
-%a {\em social} aspect to the design of a
-%computing system that makes it a {\em techno-social system}
-In this paper, our intention is to
+a {\em social} aspect to the design of a
+computing system that makes it a {\em techno-social system}
+\cite{vespignani2009predicting}. In this paper, our intention is to
 present the {\sf NodIO} framework and analyze it from two different
@@
 into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page running the system
+that voluntarily choose to enter the web page that runs the system
 determine its performance and how it can be modeled and, eventually,
@@
 state of the art (Section \ref{sec:soa}) in web-based distributed
-computational systems along with attempts to predict and model its behavior. 
+computation systems along with attempts to predict and model its behavior. 
 Then, a detailed description of the proposed system is presented
@@
 paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that,
-even in the early stages of the web, the possibility of
+\cite{baratloo1999charlotte}, but it is interesting to note that
+already in the early stages of the web the possibility of
 % (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
@@
 the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
+a comprehensive statistics collection subsystem, allowing its use for science,
 since it allows % (Paloma) Use synonyms, like "yield" or "permit"
@@
 distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. SuperWeb is focused
+brokers for resource discovery. It is focused
 % (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
@@
 browsers was realized
-later on \cite{sarmenta-bayanihan} as well as the potential of
-user misconduct when having access to the source code.
+later on \cite{sarmenta-bayanihan} as well as the potential for
+mischief of users with code in their hands
 % (Paloma) I think this sentence is extremely informal
@@
 when writing code and leverages 
-the expertise acquired
+expertise acquired
 % (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
@@
 al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
-%But the raw material of           
-%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
-But some of the essential metrics in volunteer computing like the
-number of users or the time spent by every one in the
+computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
+volunteer computing, % (Paloma) Whith regard to?
+number of users and the time spent by every one in the
 computation in browser-based volunteer computing experiments, have
@@
 
-As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
-computational resources that are as dissimilar as smartphones and
+As far as we know, this paper presents one of the few experiments that % (Paloma) Change to "which" in order to not repeating "that" all the time
+uses computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. 
-%The methodology used
+or workstation computers in a research center. The methodology used
 % (Paloma) I would say "the used methodology"
-The algorithms used, as well as the methodology 
-for gathering resources will be described next, 
-together with the results obtained in this initial setup.
+ to
+gather resources and the algorithms used will be described next, as
+well as the results obtained with the initial setup.
 
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
+be some issues with the performance of the language itself; that % (Paloma) I recommend the use of "which" here
  is
 why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
-faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
+\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
+faster than compiled languages such as Scala and, in any case, provides
+a performance that is comparable to other languages employed usually % (Paloma) Adverb
 in evolutionary computation and numerical optimization such as Python. 
@@
 Compiled languages such as
-Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
-considering the performance of a single-user, 
-but this is more than offset by the computational power of
-the spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets. Together, the performance is several orders of magnitude
+Java or C might be, in fact, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+ that a single-user
+performance is higher is more than offset by the fact that the number
+of spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets, is several orders of magnitude
 higher, which is the objective in this kind of systems.
@@
 V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter \cite{Gray:2009:GCM:1610564.1610565}.
+There are no intermediate byte codes or an interpreter.
+[\url{ https://developers.google.com/v8/design#mach_code}].
 JavaScript has gained popularity in server-side development using runtime
@@
 a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization \cite{tang2007benchmark} is used in this section. 
+Large-Scale Global Optimization [] is used in this section. 
 The function is described next.
@@
 \hline
-GEV & 35.759 $\pm$ 0.83719  &  343.336  $\pm$  & 9.877   $\pm$
-0.04746 \\
+GEV & 35.759  &  343.336   & 9.877 \\
 Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
@@
 \end{table*}
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
+\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
+\caption{Fit to GEV (left) and Weibull (right) of the number of {\tt PUT}s per worker for the F15 (Rastrigin) function.} 
+% Plotted with ../data/ips-puts-rastrigin.R
+\label{fig:fit:rastrigin}
+\end{figure}
+%
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 browsers was realized
-later on \cite{sarmenta-bayanihan} as well as the potential for
-mischief of users with code in their hands
+later on \cite{sarmenta-bayanihan} as well as the potential of
+user misconduct when having access to the source code.
 % (Paloma) I think this sentence is extremely informal
@@
 when writing code and leverages 
-expertise acquired in the library itself.
+the expertise acquired
+% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
+ in the library itself.
 
@@
 
-As far as we know, this paper presents one of the few experiments which % (Paloma) Change to "which" in order to not repeating "that" all the time
-uses computational resources that are as dissimilar as smartphones and
+As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
+computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. The methodology used to
-gather resources and the algorithms used will be described next, as
-well as the results obtained with the initial setup.
-
+or workstation computers in a research center. 
+%The methodology used
+% (Paloma) I would say "the used methodology"
+The algorithms used, as well as the methodology 
+for gathering resources will be described next, 
+together with the results obtained in this initial setup.
 %---------------------------------------------------------------
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself. This
+be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
  is
 why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
-faster than compiled languages such as Scala and, in any case, provides
-a performance that is comparable to other languages usually employed  
+\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
+faster than compiled languages such as Scala and, in any case, it provides
+a performance that is comparable to other languages usually employed % (Paloma) Adverb
 in evolutionary computation and numerical optimization such as Python. 
@@
 Compiled languages such as
-Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
- that a single-user
-performance is higher is more than offset. This is because the number
-of spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets, is several orders of magnitude
+Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+considering the performance of a single-user, 
+but this is more than offset by the computational power of
+the spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets. Together, the performance is several orders of magnitude
 higher, which is the objective in this kind of systems.
@@
 anonymous, since no accounts are needed to participate in the
-experiment, it is safe for the user, since we are using the browser
+experiment, it is safe for the user, since we are using a the browser
 black box, and it is also heterogeneous. We have designed the
@@
 experiments and clients, scalability with the number of clients and
-the problem size, and measured performance by comparing it with
+the problem size, and also measured its performance comparing it with
 a baseline configuration, with which, as indicated above, is

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some changes in the experiment description

@@
 be embedded in any web-enabled device, including many over which 
-developers have no control.
+the developers have no control.
 
@@
 paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that,
-even in the early stages of the web, the possibility of
+\cite{baratloo1999charlotte}, but it is interesting to note that, even in the early stages of the web, the possibility of 
 % (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
@@
 the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
-since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
+a comprehensive statistics collection subsystem, allowing its use for science,
+since it yields % (Paloma) Use synonyms, like "yield" or "permit". Repetitive use of "allow" is... repetitive.
  the collection of algorithm analytics by the
@@
 when writing code and leverages 
-the expertise acquired
-% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
- in the library itself.
+expertise acquired in the library itself.
 
@@
 al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
-%But the raw material of           
-%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
-But some of the essential metrics in volunteer computing like the
-number of users or the time spent by every one in the
-computation in browser-based volunteer computing experiments, have
+computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
+volunteer computing, with regard to
+number of users and computation time spent by every one 
+in browser-based volunteer computing experiments, have
 only been studied in a limited way in 
@@
 
-As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
-computational resources that are as dissimilar as smartphones and
+As far as we know, this paper presents one of the few experiments which % (Paloma) Change to "which" in order to not repeating "that" all the time
+uses computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. 
-%The methodology used
-% (Paloma) I would say "the used methodology"
-The algorithms used, as well as the methodology 
-for gathering resources will be described next, 
-together with the results obtained in this initial setup.
+or workstation computers in a research center. The methodology used to
+gather resources and the algorithms used will be described next, as
+well as the results obtained with the initial setup.
+
 %---------------------------------------------------------------
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
+be some issues with the performance of the language itself. This
  is
 why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
-faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
+\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
+faster than compiled languages such as Scala and, in any case, provides
+a performance that is comparable to other languages usually employed  
 in evolutionary computation and numerical optimization such as Python. 
@@
 Compiled languages such as
-Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
-considering the performance of a single-user, 
-but this is more than offset by the computational power of
-the spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets. Together, the performance is several orders of magnitude
+Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
+ that a single-user
+performance is higher is more than offset. This is because the number
+of spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets, is several orders of magnitude
 higher, which is the objective in this kind of systems.
@@
 
-These two experiments establish first a baseline result and show that
+These two experiments establish a baseline result and show that
 the time to success depends on the population size, with a bigger population
-contributing more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
 experiments that we will describe next do not and can not have the
@@
 solution was sent to the server, the counter was updated and the pool
-of solutions reset to the void set. No special care was taken to wait
+of solutions reset to the void set. There was no special intention to wait
 until all clients had finished, thus it might happen that, in fact,
@@
 the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is, in fact negligible.
+of these islands in the next experiment is indeed negligible.
 %
@@
 \hline
-GEV & 35.759  &  343.336   & 9.877 \\
+GEV & 35.759 $\pm$ 0.83719  &  343.336  $\pm$  & 9.877   $\pm$
+0.04746 \\
 Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
@@
 \end{table*}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
-\caption{Fit to GEV (left) and Weibull (right) of the number of {\tt PUT}s per worker for the F15 (Rastrigin) function.} 
-% Plotted with ../data/ips-puts-rastrigin.R
-\label{fig:fit:rastrigin}
-\end{figure}
-%
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
The columns you talk about are not the two lasts, and also you do talk about the last column

@@
 experiments involved 5 or less IPs. This is consistent with results
-found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
+previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
 the usual number of IPs that participated in an experiment. The
@@
 not competitive, {\em a priori}, with the baseline experiment, that is
-why it is interesting to look at the last two columns, that measure
+why it is interesting to look at the columns which measure
 the percentage of experiments in which the solution was found in less

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Finishing the review of another paragraph

@@
 20\% on average for the first baseline experiment, roughly 3\% in the
-second case. The comparison is not totally fair: This experiment takes
+second case. The comparison is not totally fair: this experiment takes
 place in a browser, and in fact most of the time is devoted to
@@
 remote server, not a local server as in the baseline experiment. But
-even if that is the case, we were looking for creating a framework that
+even in this case, we were looking for creating a framework that
 allowed us to run evolutionary computation experiments {\em faster} than
 in our very own computer. And the results indicate that although it will
-happen, 1 out of roughly 5 or roughly 20, depending on what baseline
-experiment you choose, it will not happen always or even a majority of
+happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
+experiment you choose, it will not happen always or even most of
 the time. One of the problems is shown in the last column, which

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changes in Fig 3 description

@@
 relationship between them, that is, computing nodes are not {\em
-  staying} after the solution has been found; this is a feature of
+  staying} after the solution has been found. This is a feature of
 this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. However, it puts a probably
+the evolutionary algorithm start again. % (Paloma) Feature or disadvantage?
+ However, it puts a probably
 unnecessary upper limit on the time, or number of experiments, that a user
@@
   red dots; in the case there is a single red dot, there was a single
-  experiment in which so many computers participated (for instance, 16
+  experiment in which many computers participated (for instance, 16
   computers in the experiment in the far left or 29 in the middle
   one). 
-From left to right, experiments 4/4, 4/24 and 7/31. Shades
+Shades
 of blue indicate how many experiments included that many unique IPs,
-so lighter shade for a column of dots indicates that particular number
-of computers happened less frequently, darker shadow more times.}
+so lighter shade for a column of dots indicates that a particular number
+of computers happened less frequently, while darker shadow means more frequence. From left to right, experiments 4/4, 4/24 and 7/31.}
 \label{fig:duration}

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Finishing the discussion of results

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
-reasonable performance.
-But hierarchy, this is, the capability of preferentially using resources in the same organization, is a feature that could be either disregarded or included within adaptivity, which leaves us with eight desirable properties.
-% (Paloma) Re-write.
-For example, "But is hierarchy, this is, the capability of preferentially using resources in the same organization, the one that could be either disregarded or included in adaptivity, which leaves us with eight desirable properties."
+reasonable performance. But hierarchy, this is, the capability of
+preferentially using resources in the same organization, is a feature
+that could be either disregarded or included within adaptivity, which
+leaves us with eight desirable properties. 
+% (Paloma) Re-write. For example, "But is hierarchy, this is, the
+% capability of preferentially using resources in the same
+% organization, the one that could be either disregarded or included
+% in adaptivity, which leaves us with eight desirable properties." 
 % (Paloma) And I re-wrote what I proposed yesterday.
+% thanks :-) -JJ
 
@@
 paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that,
-even in the early stages of the web, the possibility of
+\cite{baratloo1999charlotte}, but it is interesting to note that, even
+in the early stages of the web, the possibility of 
 % (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
@@
 
-These two experiments establish a baseline result and show that
+These two experiments establish first a baseline result and show that
 the time to success depends on the population size, with a bigger population
-contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+contributing more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
 experiments that we will describe next do not and can not have the
@@
 solution was sent to the server, the counter was updated and the pool
-of solutions reset to the void set. There was no special intention to wait
+of solutions reset to the void set. No special care was taken to wait
 until all clients had finished, thus it might happen that, in fact,
@@
 the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible.
+of these islands in the next experiment is, in fact negligible.
 %
@@
 experiments involved 5 or less IPs. This is consistent with results
-previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
+found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
 the usual number of IPs that participated in an experiment. The
@@
 not competitive, {\em a priori}, with the baseline experiment, that is
-why it is interesting to look at the columns which measure
+why it is interesting to look at the last two columns, that measure
 the percentage of experiments in which the solution was found in less
@@
 
+This is part due to the fact that the presence of nodes is not
+constant, as shown in Figure \cite{fig:ips:rastrigin}, which plots the
+number of unique IPs participating in the experiment per minute (left)
+and hour (right). The two graphs have the same shape which matches
+also the speed in which fitness is decreased in Figure
+\ref{fig:puts:rastrigin}. At the beginning there are very few or only
+one, but later one there ar peaks of almost 30 unique machine a
+minute; these peaks appear again when new announcements are made in
+Twitter or any other social networks, to go down to a few remaining
+nodes later on. It is interesting to note that, at the most, more than
+100 unique nodes were participating in the experiment during one hour;
+also that there are sustained peaks during a day of 6-7 computers
+every minute. This is due to the fact that the computation is
+asynchronous, but also to the fact that different computers stay in
+the simulation for a different period of time. This period has been
+plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
+{\tt PUT}s per node, ranked by number, along with a fit to the two
+distributions we have used before, GEV and Weibull; the parameters of
+this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
+experiments, in which GEV fit was relatively good, in this case it is
+quite obviously not. However, the fit to the Weibull distribution,
+shown at the left of the image, is relatively good. As should be
+expected, the shape and scale are different to the one in the previous
+experiments with the Trap function, however, the shape is quite
+similar (0.40 vs 0.69), being in both cases less than one, which
+indicates a decreasing distribution of the same kind.
+
+In general, this experiment confirms what was found in the others in
+several aspects: first, there is a regularity in the distribution of
+time assigned by volunteers to it, second, there is a good and certain
+amount of volunteers that will become part of the distributed
+computer, in some cases up to one hundred; also, in some cases there
+will be users that will spend several hours, two orders of magnitude
+more generations in this case than in the previous one. There is a
+clear influence of the time to solution: a short time means that every
+experiment will only gather a few volunteers, but also that they will
+stay only for a short time. A long time will make users stay for
+longer, but if it takes too long, they will eventually lose interest.
+
+We will discuss this and the rest of the findings of the paper in the
+conclusions.
+
 %---------------------------------------------------------------

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Paragraph checked. Minor change because of grammar

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 experiments. An interesting question to ask, for instance, is if
-adding more computers makes the experiment take less. In fact, as
-shown in Figure \ref{fig:duration}, {\em adding} more computers does
+by adding more computers makes the experiment take less. In fact, as
+shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
 not seem to contribute to decreasing the time needed to finish the

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
works for calls

@@
   the one described in \cite{LNCS86720702}, any kind of client that
-  works the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
+  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
 But never heard of a client which "works" an API.
@@
   in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
+  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
+  %(Mario) Scalability of the communication system in particular?
 \item Since the algorithm runs in the browser, {\em safety} is
@@
 the evolutionary algorithm start again. % (Paloma) Feature or disadvantage?
- However, it puts a probably
+ However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
 unnecessary upper limit on the time, or number of experiments, that a user

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Expands conclusions

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 
-This is part due to the fact that the presence of nodes is not
+This is in part due to the fact that the presence of nodes is not
 constant, as shown in Figure \cite{fig:ips:rastrigin}, which plots the
@@
 experiment. In fact, in this paper we have tried to put some figures
-on the real size of that {\em cloud}.
+on the real size of that {\em cloud} and how it can be used standalone
+if there is no alternative or in conjunction with other local or
+cloud-based methods to add computing power in a seamless way through
+the pool that NodIO creates.
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
First experiments description reviewed

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
   the one described in \cite{LNCS86720702}, any kind of client that
-  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
+  works the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
 But never heard of a client which "works" an API.
@@
   in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
-  %(Mario) Scalability of the communication system in particular?
+  the scalability of the system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
 \item Since the algorithm runs in the browser, {\em safety} is
@@
 the evolutionary algorithm start again. % (Paloma) Feature or disadvantage?
- However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
+ However, it puts a probably
 unnecessary upper limit on the time, or number of experiments, that a user
@@
 seconds. The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, and of the order of
+many experiments took a desirable amount of time, of the order of
 10 seconds, and which appears in all three graphs. As it can be seen,
-it drops sharply implying there are 
+it sharply drops implying there are 
 just a few of them, and with diminishing probability as time
@@
 Weibull distribution which was fitted to volunteer computing
-frameworks such as SETI@home \cite{javadi2009mining}; we obviously do
-not pretend to compare our system in scale or complexity with it, only
+frameworks such as SETI@home \cite{javadi2009mining}. We obviously do
+not pretend to compare our system in scale or complexity with it, but
 to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which appears also in our system. This
+certain pattern, found in SETI@home, and which also appears in our system. This
 distribution is governed by three parameters, the usual location $\mu$
@@
 location. Positive parameters indicate that the distribution {\em
-  leans} towards the origin, negative towards the other extreme
+  leans} towards the origin, and negative ones towards the other extreme
 value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
@@
 issue led us to create the next version of the framework, which is
-presented next.
+presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Minor changes with regard to the use of synonyms of improvement and improve

@@
 
-In order to improve the number of cycles per user, we improved the
+In order to improve the number of cycles per user, we enhanced the
 basic architecture in several different aspects.  One improvement 
@@
 running an evolutionary algorithm. Another was a small algorithmic
-improvement: instead of having a fixed population,
+enhancement: instead of having a fixed population,
 population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvement was the 
+different for each client.  The other improvement was to 
 use the Web Worker API as described next.
@@
 JavaScript has been commonly used to develop client-side scripts
-that handle the user interface, communicate asynchronously with servers and
-update the content that is displayed \cite{flanagan2006javascript}.
+that handle the user interface, communicating asynchronously with servers and
+updating the content that is displayed \cite{flanagan2006javascript}.
 Long-running scripts like those proposed in this work can interfere with the
@@
 red and blue arrows respectively. There are also messages between workers and
-the main thread locally in the client, these are shown as thin solid arrows.
+the main thread locally in the client, which are shown as thin solid arrows.
 There are three processes involved:

### 2015-11-18 special-issue-ci/Response to reviewers.tex
REvising response to reviewers
Updating for changes in the paper. Everyone else should do it to: @pacastillo, @fergunet, @unintendedbear and @mariosky.

Refs #40

@@
 deeply changed the contents, and thus, the quality of the paper.
+
+We would like to notify the editors that, thanks to changes and
+aportations made to this new version of the paper there are two new
+coauthors.
 
@@
 JavaScript uses a JIT compiler that provides a good performance; we
-have published these tests as an open access technical report.  % quizás indicar la URL aquí
+have published these tests as an open access technical report. \cite{2015arXiv151101088M}
 
@@
 
-We have added a new experiment, a more complex and taxing problem
-using a web-worker architecture, in order to prove how this can be
-used in real-life problems. In fact, our architecture {\sf NodIO} is
-not a exclusively browser-based one. Its pool-based architeture allows
+We have added a new experiment (Section V.A), a more complex and
+taxing problem that also uses a web-worker architecture, in order to
+prove how this can be 
+used in real-life problems. In this case, we have achieved many more
+simultaneous users, up to 25 in some minutes and more than 100 in a
+single hour. All in all, more than 400 volunteers (identified by
+unique IPs) have participated in
+the experiment.
+
+In fact, our architecture {\sf NodIO} is
+not an exclusively browser-based one. Its pool-based architeture allows
 any other kind of client, including local ones, to be incorporated to
@@
 
-Our framework, in fact, does not prevent any other kind of
-experiments. If there is a very efficient local algorithm, it only
+Our framework, in fact, does not prevent the user to carry out any other kind of
+experiments. If there is a very efficient locally run algorithm, it only
 needs to be able to do HTTP requests to work alongside a {\em
@@
 Contribution:
-This paper describes a client-server architecture for distributed evolutionary algorithms named NodIO.
-The implementation of the architecture is written entirely in JavaScript, in order to support persistent,
+This paper describes a client-server architecture for distributed
+evolutionary algorithms named NodIO. 
+The implementation of the architecture is written entirely in
+JavaScript, in order to support persistent, 
 asynchronous, distributed evolutionary algorithms running in the browser.
-Additionally, the paper presents the results of several experiments performed to validate the proposal and
+Additionally, the paper presents the results of several experiments
+performed to validate the proposal and 
 to measure the performance improvement generated by the massive parallel computing.
-I think that the contribution is quite relevant, since it provides a computation platform that is flexible,
+I think that the contribution is quite relevant, since it provides a
+computation platform that is flexible, 
 extensible and potentially massive. Personally, I consider it one of the fresher and most promising
@@
 A experiment using the well known Rastrigin function has been
-performed and analyzed.
+performed and analyzed. It is now Subsection V.A.
 
@@
 
-This has been included as the last subsection of the paper. The new
+This has been included as the last subsection of the paper previous to
+the conclusions. The new
 experiment has constant population, instead of using random
-population.
+population. We have not found a big influence on the result.
 
@@
 new issues in this new version. We have preferred to keep it the way
-it is, although I thank you for the suggestion
+it is, although we thank you for the suggestion. We expect that the
+changes made to the text make it clear enough anyway.
 
 \begin{quote}
-Another issue is that some details are described twice, for instance, the advantages of the use of web
+Another issue is that some details are described twice, for instance,
+the advantages of the use of web 
 workers. Fixing this issue can reduce the paper size and improve its
@@
 a way (and failed) to make it more readable for reviewers. In fact,
-they use different shades, to it should mostly work in grayscale. For the time being and
+they use different shades, so it should mostly work in grayscale. For
+the time being and 
 during the revision phase, we prefer to keep it that way, but we will
@@
 state of the art is comprehensive and provides good introduction into the subject.
+\end{quote}
 
+We thank this reviewer for these helpful comments.
+
+\begin{quote}
 2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
@@
 The base system has been now described in the first paragraph of the
-section on the experiments, it a desktop Intel i7-4770.
+section on the experiments, it is a desktop Intel i7-4770.
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

@@
 In general, this experiment confirms what was found in the others in
-several aspects: first, there is a regularity in the distribution of
+several aspects: first, there is a regularity in the distribution of the
 time assigned by volunteers to it, second, there is a good and certain

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Sequence of events reviewed

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 
-The sequence of events is as follows:
+The sequence of events happens as follows:
 \begin{enumerate}
 \item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure the sequential number of experiment is initialized and a log
+initialization procedure, the sequential number of experiments is initialized and a log
 file is created. The shared pool implemented as an array is reset.
 \item A volunteer follows the link of the experiment triggering a GET HTTP
-request and the page is loaded.
+request, then the page is loaded.
 \item The main script renders the page and creates the worker instance. In
 order to create a dedicated worker instance, another GET request is sent to the
-server in order to receive the worker implementation script.
+server, then the client receives the second worker implementation script.
 Each worker instance executes over time one or more evolutionary algorithms
-{\em islands}. Each starts with a random population and it is assigned a
+{\em islands}. Each island starts with a random population and it is assigned a
 universally unique identifier (UUID). The UUID is included in each of the HTTP
@@
 generation loop is started. After {\em n} generations the state of the population
-is evaluated. First the chromosome with the best fitness is sent to the shared
+is evaluated. The chromosome with the best fitness is sent to the shared
 population with a PUT HTTP request. If the best fitness is not the global best,
 the EA loop must continue, but first a chromosome is randomly selected from
-the shared population. Also the current generation number and the best
-chromosome is sent to the client's main thread in order to update the DOM to
+the shared population. The current generation number and the best
+chromosome are also sent to the client's main thread in order to update the DOM to
 display the current state of the island (described next). These messages are
-all asynchronous, so even before a message is received by the target, the EA loop of
-the worker continues. If the best fitness is a global best the best
+all asynchronous, so the EA loop of
+the worker continues even before a message is received by the target. If the best fitness is a global best, the best
 chromosome is sent to the server and the loop ends. A message is sent to the
@@
 fitness, and a representation of the chromosome is also generated.
-\item When a global best is received from an island the current experiment
-ends, the experiment number is incremented and the population array is reset.
+\item When a global best is received from an island, the current experiment
+ends, the experiment number is incremented, and the population array is reset.
 \item When the main thread receives a message indicating that a worker has ended the EA

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding conclusions on Weibull
After conversation with @nuriastatgirl.

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 similar (0.40 vs 0.69), being in both cases less than one, which
-indicates a decreasing distribution of the same kind.
+indicates a decreasing distribution of the same kind. In fact, the
+difference in the scale parameter indicates that the difference
+between the time devoted by a volunteer and the next is bigger in this
+case. This is probably due to the fact that the duration of the
+experiment is bigger, allowing for bigger differences in
+behavior. While the first in the Trap function might have devoted a
+few minutes and the second a few seconds less, the difference between
+the first and second in this experiment might be in the scale of
+hours.
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Fixes error

@@
 This is in part due to the fact that the presence of nodes is not
-constant, as shown in Figure \cite{fig:ips:rastrigin}, which plots the
+constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
 number of unique IPs participating in the experiment per minute (left)

### 2015-11-18 special-issue-ci/Response to reviewers.tex
Some typos fixed for response for reviewer 1 for issue #40

--- 2015-11-17 special-issue-ci/Response to reviewers.tex
@@
 which is actually run by volunteers in their browsers. We went to the
-original aception of ``cloud'' as a ``network'' more than the more
-precise and technical aception used today.
+original acceptation of ``cloud'' as a ``network'' more than the more
+precise and technical acceptation used today. %PABLO: no existe aception, lo he cambiado a acceptation (segun WordReference)
 
@@
 conclusions; we hope this makes our intention clearer, but we can also
-eliminate that part of the title if the reviewer think it is
+eliminate that part of the title if the reviewer thinks it is
 necessary.
@@
 used in real-life problems. In fact, our architecture {\sf NodIO} is
-not a exclusively browser-based one. Its pool-based architeture allows
+not an exclusively browser-based one. Its pool-based architecture allows
 any other kind of client, including local ones, to be incorporated to
@@
 
-We would like to thanks this reviewer for the helpful comments and
+We would like to thank this reviewer for the helpful comments and
 suggestion made. We expect to have addressed them properly, but in any

### 2015-11-18 special-issue-ci/Response to reviewers.tex
Few changes

--- 2015-11-17 special-issue-ci/Response to reviewers.tex
@@
 as it has been installed, the user has decided to fire it up, and it
-does not affect the performance in the rest of the system, which is a
+does not affect the performance of the rest of his system, which is a % se confundia con el desempeño del todo el sistema
 concern in this kind of systems. We have performed speed tests on
@@
 
-That is correct. We just used {\em cloud} in a relatively informal
+That is correct. We just used the term {\em cloud} in a relatively informal
 sense because they are available as in a cloud for running the
@@
 used in real-life problems. In fact, our architecture {\sf NodIO} is
-not a exclusively browser-based one. Its pool-based architeture allows
+not an exclusively browser-based one. Its a pool-based architeture that allows
 any other kind of client, including local ones, to be incorporated to
@@
 
-A experiment using the well known Rastrigin function has been
+An experiment based on the well known Rastrigin's function has been
 performed and analyzed.
@@
 This has been included as the last subsection of the paper. The new
-experiment has constant population, instead of using random
-population.
+experiment has a constant population, instead of using  population of random size.
 
@@
 a way (and failed) to make it more readable for reviewers. In fact,
-they use different shades, to it should mostly work in grayscale. For the time being and
+they use different shades, so it should mostly work in grayscale. For the time being and
 during the revision phase, we prefer to keep it that way, but we will
@@
 also proved in the last section how the performance achieved by a
-volunteer computer system can, in fact, be better than a desktop, even
+volunteer-based computer system can, in fact, be better than a desktop, even
 a powerful one. This has also been clarified in the introduction and
@@
 
-At any rate, we have also performed a new experiment with the
-Rastrigin function, which in fact cannot be solved in a few minutes
+At any rate, we have also performed a new experiment based on the
+Rastrigin's function, which in fact cannot be solved in a few minutes
 (or seconds, as in the case of Trap) by a desktop system.
@@
  distributed computing system, providing that a user did not disable Java Script in its browser.
- However if the authors plan to use HTML5 WebWorkers only modern web browsers can
+ However if the authors plan to use HTML5 Web Workers only modern web browsers can
  participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds to conclusions refs #40 closes #2

@@
 for this kind of asynchronous framework, but it would be interesting
-to check the performance for more computationally demanding
-tasks. Besides, the two versions of the framework include as an
+to check the performance for other computationally demanding
+tasks, especially ones that cover the middle ground between a problem
+finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
+Besides, the two versions of the framework include as an
 algorithmic variant using random population size. We do not think this
-has had a big influence on the results, but it would be interesting to
-measure exactly what this influence has been. Adding a bit of more
+has had a big influence on the results and in fact this is not noticed
+for F15, which is the second version tested but it would be interesting to
+measure exactly what this influence has been. In general there are
+many issues with the evolutionary algorithm implementation itself,
+including using different, or adaptive, policies for inserting and
+sending individuals to the pool as was done in \cite{araujo2008mam},
+using different policies for population initialization, and also the
+incorporation of high-speed local resources to the pool to check what
+would be the real influence of the volunteer pool to the final
+performance.
+
+Another area is how to enhance the number and quality of
+volunteers. For instance, adding a bit of more
 control to the user might contribute also to gamification. Right now

### 2015-11-18 special-issue-ci/Response to reviewers.tex

--- 2015-11-18 special-issue-ci/Response to reviewers.tex
@@
 In fact, our architecture {\sf NodIO} is
-not an exclusively browser-based one. Its pool-based architeture allows
+not an exclusively browser-based one. Its pool-based architecture allows
 any other kind of client, including local ones, to be incorporated to

### 2015-11-18 special-issue-ci/Response to reviewers.tex
Restrigin's

@@
 
-A experiment using the well known Rastrigin function has been
+An experiment based on the the well known Rastrigin's function has been
 performed and analyzed. It is now Subsection V.A.
@@
 the conclusions. The new
-experiment has constant population, instead of using random
-population. We have not found a big influence on the result.
+experiment has a constant population instead of using  population of random size.
+We have not found a big influence on the result.
 
@@
 
-We have extended the experiments section with a new one, based on the
-hard Rastrigint problem.
+We have extended the experiments section with a new one, based on a
+hard version of the Rastrigin's problem.
 

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding meself as author, adding @mariosky explanation about the feature of first NodIO version

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
+\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$
 \thanks{Manuscript submitted for review on \today.}%
@@
 this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. % (Paloma) Feature or disadvantage?
+the evolutionary algorithm start again. Thus, the user can decide if to continue sharing resources or not.% (Paloma) Feature or disadvantage?
  However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
@@
 less than the second baseline. Thus, the conclusion in this case is
-that we can obtain, through volunteer computation, a result that is,
-most of the times, better than we would using a similar, non-parallel,
+that we can obtain, through volunteer computation, a result that most of the times is better than we would by using a similar, non-parallel,
 desktop setup, which makes this system suitable for massively
@@
 similar (0.40 vs 0.69), being in both cases less than one, which
-indicates a decreasing distribution of the same kind. In fact, the
-difference in the scale parameter indicates that the difference
-between the time devoted by a volunteer and the next is bigger in this
-case. This is probably due to the fact that the duration of the
-experiment is bigger, allowing for bigger differences in
-behavior. While the first in the Trap function might have devoted a
-few minutes and the second a few seconds less, the difference between
-the first and second in this experiment might be in the scale of
-hours.
+indicates a decreasing distribution of the same kind.
 

### 2015-11-18 special-issue-ci/Response to reviewers.tex
More fixed typos and compiling errors for issue #40

@@
 \begin{quote}
-- The paper abstract does not properly summarize the paper contents. It says that "That is why in this
-paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !", but there
+- The paper abstract does not properly summarize the paper contents. It says that ``That is why in this
+paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !'', but there
 are two versions of the architecture, as stated later in the conclusions: "In this paper two versions of a
@@
 high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
-architecture presented inthe paper can be potentially used for solving many different time consuming
-problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
+architecture presented in the paper can be potentially used for solving many different time consuming
+problems, however, the authors limited their experiments to a simple ``Trap'' function that can be easily
 solved on a desktop machine, what may not be convincing for a reader looking for an effective
@@
 messages sent between the server and the clients. The only thing I noticed is that the authors used the
-term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
+term ``multiple CPUs'' in the description of Web Worker API advantages (page 8) instead of "multi-core
 CPUs". As far as I know, contemporary web browsers do not support
@@
 4. Style: Paper is generally written in a clear way. Statements like
-"But fair is fair," (page 6), sound a bit colloquially in a scientific
+``But fair is fair,'' (page 6), sound a bit colloquially in a scientific
 paper and may be omitted. However, my main concern is the readability
@@
 for axe Y, as the box for the population of 1024 individuals is very
-small.  What do the authors mean by "times"? Time in seconds? I am not
+small.  What do the authors mean by ``times''? Time in seconds? I am not
 sure if this Figure is necessary at all as the description is
@@
 \begin{quote}
-The same is with Fig. 5. What does "x" stand for in the description of
+The same is with Fig. 5. What does ``x'' stand for in the description of
 axe X? (the same is with Fig. 8).
@@
  - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
- implement evolutionary algorithm in that way that it wil automatically adopt to the problem type
+ implement evolutionary algorithm in that way that it will automatically adopt to the problem type
  and the environment in which it is computed, the authors do not mention about how real
@@
  possession of sensitive data or/and alter this code in order to disrupt the computations or make
- them worhless. The results send to the server may also be captured and/or altered without any
+ them worthless. The results send to the server may also be captured and/or altered without any
  problem since the transmission is not secure.
@@
 
-\begin{quote{
-2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
-funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
+\begin{quote}
+2) Regarding the experiments I am very skeptical about their real value. You chose the ``Trap''
+function. in which setting all bits to zero gives local maximum and setting all bits to one gives global
 maximum. In the experiments you counted how many times evolutionary algorithm was able to find
@@
 baseline). I suggest to perform additional computing experiments for some higher value of n in the
-"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
+``Trap'' function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
 to focus the paper on the behavior analysis of volunteers participating in a distributed computing
@@
 \begin{quote}
-3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In
+3) Regarding the graphs I have already put my comments in Style section, but I have one question. In
 Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
@@
 which 4 computing nodes participated and the red dot is the average;
-it shadows the time in the case there is a single ocassion. In the
+it shadows the time in the case there is a single occasion. In the
 case you mention, there are no experiments in which there were exactly
@@
 This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
-architecture for low cost and "potentially" massive distributed evolutionary algorithms. The main novelty
+architecture for low cost and ``potentially'' massive distributed evolutionary algorithms. The main novelty
 turns up from the fact that the architecture is developed in JavaScript.
@@
 applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
-believe in the possibility of "potentially" massive computation as
+believe in the possibility of ``potentially'' massive computation as
 sketched by the authors.
@@
 their speeds are virtually the same. In general, as a conclusion of
-the report we see that the real difference between scripted and
+the report, we see that the real difference between scripted and
 compiled languages will vary with the language, but also with the
@@
 high-speed devices such as GPUs or even local powerful CPUs can be
-added as non-volunteer users, since the heterogeinity and simple API
+added as non-volunteer users, since the heterogeneity and simple API
 of the system allows that by design.
@@
 demanding high computational performance.
-Finally, English must be double-checked in order to correct minor typos such as "in the sense than[that]"
+Finally, English must be double-checked in order to correct minor typos such as ``in the sense than[that]''
 (in page 2).

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
More minor changes on the results discussion, mostly typos and use of synonyms

@@
 4000 generations. All things considered, this means than there are
-many more computers running {\em simultaneously}, leading to this
+many more computers {\em simultaneously} running, leading to this
 almost 5-fold increase in running time.
@@
 participating in an experiment and the time it takes to complete
-it. This is shown in Figure \ref{fig:ipstime:w2}, which shows every
-experiment in a time (in milliseconds) vs. number of workers participating
+it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
+experiment in terms of time (in milliseconds) vs. the number of workers participating
 in it. Although there is not a clear trend, the graph seems to say
@@
 difficult to measure. There seems to be a trend towards a decreasing
-standard deviation in the time, with more volunteers adding {\em
-  robustness} to an experiment and certainty in the time it is taking;
-however, more experiments would be needed to check this hypothesis.
-
-As we did in the previous subsection, we have fitted also the number
+standard deviation in the time, with more volunteers adding both {\em
+  robustness} to an experiment, and certainty in the time it is taking. However, more experiments would be needed to check this hypothesis.
+
+As we did in the previous subsection, we have also fitted the number
 of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
@@
 the higher number of contributions per volunteer that has been
-previously observed; however, the shape $\xi$ parameter is
+previously observed. However, the shape $\xi$ parameter is
 substantially similar, with a {\em bell} shape leaning towards the
@@
 {\em phase change} between the people that did so and the people that
-just run the experiment until completion. In this case, however, the
-user just needs to let it run, without that phase change and thus a
+just ran the experiment once until completion. In this case, however, the
+user just needs to let it run, without that phase change and thus we obtain a
 more {\em egalitarian} distribution of contributions, which rather
@@
 uniform amount of computation, since 100 generations with a random
-population correspond, in each case, to different number of
-operations; in other words, the same number of PUTs will correspond to
+population belong, in each case, to different number of
+operations. In other words, the same number of PUTs will correspond to
 different number of operations in each case. However, since clients

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Details about Javascript are reviewed

@@
 In general, this second version of the {\sf NodIO} framework and the
-single experiment performed prove that it can be the foundation for a
-high-server-performance distributed evolutionary computation platform,
+single experiment performed prove that it can be the foundation for a distributed
+high-server-performance evolutionary computation platform,
 providing reasonable algorithmic performance and being, in general,
@@
 %Mario: Creo que el párrafo de arriba sale sobrando
+% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
 To further evaluate how a current JavaScript implementation could be used by a
@@
 $D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
-is a classical multimodal problem. It is difficult since the number of local
+is a classical multimodal problem. Such problem is difficult since the number of local
 optima grows exponentially with the increase of dimensionality. To make it 
@@
 but little interaction between the groups. This issue is reflected in the benchmark 
-by first randomly dividing the objective variables
+by randomly dividing the objective variables
 into several groups, each of which contains a number of variables. The parameter $m$ 
@@
 This optimization problem was selected because it is representative of the kind of
-algorithms, data types and structures employed in large scale optimization problems. 
+algorithms, data types, and structures employed in large scale optimization problems. 
 The function was implemented in both Matlab and Java languages as part of the
@@
 implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
-3,000,000 FEs as the termination condition. Using the Java implementation as
+3,000,000 FEs as termination condition. Using the Java implementation as
 a guide, the function described above was implemented to compare the performance of JavaScript against
@@
 the implementation in JavaScript are presented to give readers an idea of 
-how practical and mature is the language for developing complex mathematical functions.
+how practical and mature this language is for developing complex mathematical functions.
 
@@
 which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
-array and for measuring in the browser, the {\tt Performance.now()} function was used,
-this function uses floating-point numbers with up to microsecond precision. 
-A drawback found is that the {\tt Performance.now()} function is that is only implemented
+array, and for measuring in the browser the {\tt Performance.now()} function was called,
+which uses floating-point numbers with up to microsecond precision. 
+A drawback found is that the {\tt Performance.now()} function is only implemented
 in Firefox and Chrome browsers. Both functions are independent of the system clock. 
-\item {\em Tools} An important consideration when choosing certain language to 
-to write non-trivial programs is the availability of development tools. When developing
+\item {\em Tools} An important consideration when choosing certain language to write non-trivial programs is the availability of development tools. While developing
 the experiments, we used a comprehensive set of developer tools available for both
 browsers and desktop. These tools included package managers, debuggers, logging libraries, 
-network monitors, editors and even debugging of the multi-threaded execution of Web Workers.
-A drawback found is that developer tools very from browser to browser, 
-but they all have the same functionality.
+network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
+Although the fact that developer tools vary from browser to browser supposed a drawback, they all have the same functionality.
 \item {\em Data Types} JavaScript uses floating point numbers with a 
@@
 the {\tt StrictMath} library used by the Java implementation. If more precision
-is needed developers could use {\tt math.js}, an extensive math library with
+is needed, developers could use {\tt math.js}, an extensive math library with
 support for matrices and big numbers.

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
More minor changes to results

@@
 code, but developers have to consider the differences between VMs implementations
-and use certain libraries in order to increase the precision and repeatability of
+and therefore use certain libraries in order to increase the precision and repeatability of
 algorithms. As for the performance of JavaScript against Java and Matlab
@@
 Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in
+and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
+ The performance of the time required for 10,000 FEs is presented in
 Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
 performance with an average of 935ms, followed by Java with
-991ms, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
+991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
 single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
@@
 Workers. When running the experiment in Firefox a dialog alerted that the
-script was taking to much time, showing that this
-kind of processing is not intended for the main thread. Also there is not much overhead
+script was taking too much time, showing that this
+kind of processing is not intended for the main thread. Additionally, there is not much overhead
 when running two

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
añadido mi nombre

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$
+\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
 \thanks{Manuscript submitted for review on \today.}%
@@
 \thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
+\thanks{$^3$Dept. of Statistics and Operational Researcha University of Granada}%
 \thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
-    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
+    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es},{\tt nrico@ugr.es}}%
 \thanks{*Corresponding author.}%

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Rearranges figures to avoid filling a page
Fixing typos
I think chromosome wasn't written correctly a single time. Closes #40

@@
 \thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
-\thanks{$^3$Dept. of Statistics and Operational Researcha University of Granada}%
+\thanks{$^3$Dept. of Statistics and Operational Research University of Granada}%
 \thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
@@
 
-When developing volunteer based systems, desingers must
+When developing volunteer based systems, designers must
 also consider the whole {\em social} aspect in the design, with
@@
 metaheuristics optimization is that it allows an improvement in communicability,
-reproducibility, interoperability, automated assembly, knowledge engineering
+reproductibility, interoperability, automated assembly, knowledge engineering
 and efficiency.
@@
 function in Python, Erlang, Scala, or even Java, and them compile it to
-JavaScript. For completness, an experiment comparing the performance of
+JavaScript. For completeness, an experiment comparing the performance of
 a JavaScript implementation versus Java and Matlab is included in section
@@
   performance, operating system, or even browser requirement as long
-  as Javascript is enabled: anyone
+  as JavaScript is enabled: anyone
   visiting the page, even from mobile devices, can load the algorithm.
@@
 A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of Internet Protocol addressess (IPs)
+\ref{tab:summary:os}, which shows the median number of Internet Protocol addresses (IPs)
 intervening in each experiment,  median time needed
@@
 so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequence. From left to right, experiments 4/4, 4/24 and 7/31.}
+of computers happened less frequently, while darker shadow means more frequency. From left to right, experiments 4/4, 4/24 and 7/31.}
 \label{fig:duration}
@@
 \end{figure}
+%
+\begin{table*}[!htb]
+\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
+  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
+\begin{center}
+\begin{tabular}{l|ccccccc}
+\hline
+Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
+  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
+\hline
+{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
+{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
 
@@
 %
-\begin{table*}[!htb]
-\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
-  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
-\begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
-{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
+\caption{Time employed in every experiment vs. number of IPs (in
+  abscissas). The red dot and line show the average and standard
+  deviation, the shade of blue the number of experiments with the same
+  number of IPs. } 
+\label{fig:ipstime:w2}
+\end{figure}
 
@@
 distributed evolutionary algorithms.
-
-%FERGU: a lo mejor algún reviewer se queja de que la comparativa no es justa
-%porque comparamos un mono-isla (el baseline) con un multi-isla. Quizás deberíamos
-%haber usado un multi-isla en el baseline (aunque sea
-%secuencial). Para la próxima.
-% se trata de probar que podemos obtener mejores resultados que en una
-% implementación local básica. Si usamos una paralela local, ya
-% depende del ordenador que tengamos - JJ
+%
+\begin{table*}
+\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
+  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
+\begin{center}
+\begin{tabular}{cccc}
+\hline
+Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+\hline
+GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
+0.04746 \\
+Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+%
 
@@
 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
-\caption{Time employed in every experiment vs. number of IPs (in
-  abscissas). The red dot and line show the average and standard
-  deviation, the shade of blue the number of experiments with the same
-  number of IPs. } 
-\label{fig:ipstime:w2}
-\end{figure}
-%
-%
-\begin{table*}
-\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
-  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
-0.04746 \\
-Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
-\label{fig:gev:w2}
-% File ../data/ips-puts.R
-\end{figure}
-%
 We are also interested in measuring the scaling properties in this
@@
 
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
+\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
+\label{fig:gev:w2}
+% File ../data/ips-puts.R
+\end{figure}
+%
 As we did in the previous subsection, we have also fitted the number
@@
 
-\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigins Function}
+\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigin's Function}
 JavaScript has been traditionally implemented as an interpreted
@@
 non-separable, an orthogonal matrix is also used for coordinate rotation.
-The rotated Rastrgin’s function is defined as follows:
+The rotated Rastrigin's function is defined as follows:
 \begin{equation}
@@
 \begin{itemize}
-\item {\em Randomize} Both Malab and Java programs rely on a 
+\item {\em Randomize} Both Matlab and Java programs rely on a 
 Java Randomizer library for the 
@@
 This experiment was deployed in a virtual private server with 512MB of memory
-runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
-experiments where the optimum was espected to be found many times, 
+running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
+experiments where the optimum was expected to be found many times, 
 this problem was expected to run constantly for many days finding only
-sub-optimal aproximations. In order to achieve this, the only change needed 
-in the system was to limit the number of chromosmes kept in the pool. 
-Previously, every PUT request which included a new chromose increased 
+sub-optimal approximations. In order to achieve this, the only change needed 
+in the system was to limit the number of chromosomes kept in the pool. 
+Previously, every PUT request which included a new chromosome increased 
 the size of the pool, in a long-running experiment this has to be avoided
-in order to not exahust the available memory. In this experiment the size
-of a chromosme was 1000 double precision numbers, considereing this
+in order to not exhaust the available memory. In this experiment the size
+of a chromosome was 1000 double precision numbers, considering this
 and the memory available in the server, the the pool was limited 
 to 10,000 chromosomes.  In order to keep this size
-constant a random chromosme was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosme-float} 
+constant a random chromosome was pulled out of the pool before inserting a
+new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
 and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 
 were tournament size = 3 and  population size = 500, returning exchanging a
-chromosme with the server every 100 generations. Again two workers in each window.
+chromosome with the server every 100 generations. Again two workers in each window.
 As before, a few participation calls were published in social networks.
@@
 \ref{fig:puts:rastrigin}. At the beginning there are very few or only
-one, but later one there ar peaks of almost 30 unique machine a
+one, but later one there are peaks of almost 30 unique machine a
 minute; these peaks appear again when new announcements are made in

### 2015-11-18 special-issue-ci/Response to reviewers.tex
Changes title

@@
 original acceptation of ``cloud'' as a ``network'' more than the more
-precise and technical acceptation used today. %PABLO: no existe aception, lo he cambiado a acceptation (segun WordReference)
+precise and technical sense used today. We have changed the title to
+avoid words like {\em potentially}, and we have no strong feelings on
+the ``volunteers in the cloud'' part. We like it, but if asked by the
+editors, we can eliminate it.
 
@@
 
-An experiment based on the the well known Rastrigin's function has been
+An experiment based on the well known Rastrigin's function has been
 performed and analyzed. It is now Subsection V.A.

### 2015-11-18 special-issue-ci/Response to reviewers.tex
More typos fixed, setting the quotes to bold, because I think it is more readable @JJ. Also, URL options to break urls properly for issue #40

--- 2015-11-18 special-issue-ci/Response to reviewers.tex
@@
 %\usepackage{setspace}
-\usepackage{url}
+\usepackage[hyphens]{url}
 \usepackage[english]{babel}
-
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}%PABLO: he añadido esto para que no se salga la URL por el borde
 \begin{document}
@@
 We would like to notify the editors that, thanks to changes and
-aportations made to this new version of the paper there are two new
+contributions made to this new version of the paper there are two new
 coauthors.
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 The main contribution of this paper is the proposal of a Javascript based distributed system intended to
@@
 be formulated in the paper) seems a rather academic one.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 The paper organization is not clear. Some comments about this:
@@
 common PCs and laptops.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 - The paper abstract does not properly summarize the paper contents. It says that ``That is why in this
@@
 commented in the introduction.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 - According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
@@
 architecture.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Summarizing the review, the authors should provide a convincing example showing the advantages of
@@
 problem.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Contribution:
@@
 recommendation of acceptance with minor changes.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
@@
 -  a section discussing the threats to validity of the experiments
-
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 -  an experiment with a more difficult problem (please use the well-known benchmark problems
 described in literature).
-\end{quote}
+}\end{quote}
 
-An experiment based on the the well known Rastrigin's function has been
+An experiment based on the well known Rastrigin's function has been
 performed and analyzed. It is now Subsection V.A.
 
-\begin{quote}
+\begin{quote}\textbf{
 -  A comparison in similar conditions of the two implementations of the architecture (the algorithms
@@
 workers (not that generated due to a different algorithm)
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Regarding the organization of the paper, some important details regarding the improved framework
@@
 results and the discussion in section 4.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Another issue is that some details are described twice, for instance,
@@
 readability.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Regarding the presentation:
@@
 -  There are some typos and phrases that need rewriting.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
@@
 state of the art is comprehensive and provides good introduction into the subject.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
@@
 optimization tool.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 3. Correctness: The system works correctly - it is not very complex and the authors provide its source
@@
 multiple CPUs architecture.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
-
+\begin{quote}\textbf{
 4. Style: Paper is generally written in a clear way. Statements like
@@
 of the graphs. General remark - font used for description of axes
-
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 should be significantly larger. In Fig. 2 the authors can change scale
@@
 the one in the middle is analyzed, so maybe others are not necessary?
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 The same is with Fig. 5. What does ``x'' stand for in the description of
 axe X? (the same is with Fig. 8).
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
-
+\begin{quote}\textbf{
 5. Length: As I mentioned above the authors should think of removing some of the unnecessary graphs
@@
 graphs (Fig. 5 and Fig. 8). Nevertheless, overall length of the paper is fine.
-
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
@@
 Worker architecture.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
-Comments to Authors:
-
+\begin{quote}\textbf{
+Comments to Authors:\\
 1) I like the state of the art which includes comprehensive overview of the most important papers related
 to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
-with its characteristic provided on page 4. Namely:
-
+with its characteristic provided on page 4. Namely:\\
  - Heterogeneity - it is true that almost all devices can participate in the web browser based
@@
  Windows Phone) are also excluded.
-
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
  - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
@@
  system, what is a crucial parameter for distributed computing.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
-
+\begin{quote}\textbf{ 
  - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
@@
  them worthless. The results send to the server may also be captured and/or altered without any
- problem since the transmission is not secure.
-
+ problem since the transmission is not secure.\\
  In my opinion you should at least highlight above issues.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 2) Regarding the experiments I am very skeptical about their real value. You chose the ``Trap''
@@
 (page 4, if I understood it correctly).
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 Thus, what was the point of using distributed computing to solve such simple problem? Why the
@@
 readers.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 3) Regarding the graphs I have already put my comments in Style section, but I have one question. In
@@
 for that?
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC
@@
 solved by the systems are of different nature and complexity.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
@@
 sketched by the authors.
-\end{quote}
+}\end{quote}
 
@@
 
-\begin{quote}
+\begin{quote}\textbf{
 The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
@@
 (in page 2).
-\end{quote}
+}\end{quote}
 

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changes title
Changes title again
Last changes before sending

@@
 
-\title{Volunteers in the Clouds: an Architecture for Low Cost and
-  Potentially Massive Distributed Evolutionary Computation}
+\title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
+for Low Cost Distributed Evolutionary Computation }
 % The word massive in the title is giving us problems.
@@
 Volunteer computing is extensively used 
-% (Paloma) Adverbs are always written before verbs, check this throughout the text
 in science and industry, but it
 still has a long way to go in order to become
-% (Paloma) I would propose "a long way to go in order to become"
  a mainstream
 technology in the same way the cloud technologies in general have. The
-lack of % (Paloma) Add "both" here to better complete the sentence
+lack of 
 general purpose and open source tools with built-in analytics 
-% (Paloma) "built-in" is an adjective and, as such, has to be written before names
 could be one of the reasons of why this is so, as well as the lack of
@@
 framework used to create client-server architectures for distributed
-evolutionary algorithms, entirely written  % (Paloma) Adverb. I would also add an "and" after the comma
+evolutionary algorithms, entirely written  
 in JavaScript, which needs

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Fixed typos for issue #67. However, tools URL needs to be added.

--- 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 the algorithm data to find out what kind of patterns the user behavior
-follow by examining the resources they devote to our experiment, and
+follows by examining the resources they devote to our experiment, and
 try to fit them to some statistical distribution to create the bases
@@
 assumptions, such as the one presented in this paper, which will be
-provisionally named {\sf NodIO}. {\sf NodIO} is basically
+ named {\sf NodIO}. {\sf NodIO} is basically
 % (Paloma) Is the "first" word really needed? 
@@
 distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. SuperWeb is focused
+brokers for resource discovery. SuperWeb is focused,
 % (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
-, however, in solving the
+however, in solving the
 issues involved in browser-based computing and also in making a system
@@
 technologies is that they are not {\em universal} in the
-sense than an extra component, namely, the Java virtual machine or another
+sense that an extra component, namely, the Java virtual machine or another
 virtual environment has to be installed in the browser.
@@
 paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them. \cite{wood13:nodejs:paas}. Many web services
+in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
 have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
@@
 crafting a fake request which, for instance, assigns a fake fitness to
-a particular chromosome
+a particular chromosome.
 % (Paloma) Where is the second kind of threat?
-. Another two different attacks are also
+Another two different attacks are also
 possible: a denial of service attack as well as penetration
@@
 purpose of this paper, we repeated the announcement several times
-through the months of April and then by the beginning of August. All
+through the month of April and then by the beginning of August. All
 in all, we have the set of runs with the characteristics shown in
@@
 this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. Thus, the user can decide if to continue sharing resources or not.% (Paloma) Feature or disadvantage?
+the evolutionary algorithm start again. Thus, the user can decide if to continue sharing resources or not. % (Paloma) Feature or disadvantage?
 However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
@@
 \item If the browser uses a tabbed document interface, the worker script
-keeps running in the background even it the user brings forward another tabbed
+keeps running in the background even in case the user brings forward another tabbed
 window.
@@
 
-A sequence diagram of the W$^2$ version of {\sf NodIO} is presented
+A sequence diagram   of the enhanced NodIO, which is called NodIO-W$^2$, is presented
 in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
@@
 new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
-and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 
+and {\tt classic-ea-float} modules were used. The parameters for the EA 
 were tournament size = 3 and  population size = 500, returning exchanging a
@@
 
-In the second implementation, {\sf NodIO$^2$}, Web Workers
+In the second implementation, {\sf NodIO-W$^2$}, Web Workers
 were used so that several clients per browser could be executed and
@@
 that the user time is leveraged to its full potential, and this was
-achieved with {\sf NodIO$^2$}.
+achieved with {\sf NodIO-W$^2$}.
 
@@
 a General Extreme Values distribution in the case of {\sf NodIO} and a
-Weibull distribution for {\sf NodIO$^2$}. These distributions are
+Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
 close enough to each other and, in the second case, reflect the fact
@@
 the three previous experiments were published in a similar way, one
-obtained up to 5 times more total cycles  that the one with the least
+obtained up to 5 times more total cycles  than the one with the least
 number of cycles. It is also essential to obtain volunteers as

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Revising extensively references
Fixing many and adding additional ones. Refs #63.

@@
 they assume that the nodes that are going to run the system are under
-the control of developers or they at least have access to them.
+the control of developers or they at least have access to them. In
+fact, these features are commonly found in most soft computing
+frameworks \cite{7094263}.
 

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding more references refs #68

@@
 application, in which clients can run on a browser. 
-Therefore, they can
-% (Paloma) be embedded in?
-be embedded in any web-enabled device, including many over which 
+Therefore, they can be embedded in any web-enabled device, including many over which 
 developers have no control.
@@
 Considering this, we leveraged 
-% (Paloma) What 'something' did you leverage?
 our experience in the design of web
@@
 % (Mario) We leverage our experience in blah blah to design
-% (Paloma) It may be... guess I was thinking in the noun, not in the transitive verb
+% (Paloma) It may be... guess I was thinking in the noun, not in the
+% transitive verb
+% Please check if there's a problem here too.
 
 %PABLO: He añadido el siguiente párrafo para el issue #31
-There exist other volunteer frameworks written in compiled languages,
-such as C++ or Java, that can achieve better performance than
-JavaScript. However, one of the advantages of our system is that
+There exist other volunteer computing frameworks written in compiled languages,
+such as C++ or Java, that, in principle, can achieve better performance than
+JavaScript \cite{2015arXiv151101088M}. However, one of the advantages of our system is that
 clients do not need to install anything in their machines, they just
@@
 ActiveX.
-% (Paloma) Brackets not really needed here
-Also, as we discuss in this paper, it can provide some extra
+Also, as we will discuss in this paper, it can provide some extra
 features such as enhanced security or operating system independence.

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds reference to node.js tools
but that paragraph was written by @mariosky, don't know what he will think of my additions.

Refs #67

@@
 the experiments, we used a comprehensive set of developer tools available for both
-browsers and desktop. These tools included package managers, debuggers, logging libraries, 
+browsers and desktop \cite{tilkov2010node}. These tools included
+package managers, in this case {\tt npm}, debuggers, logging libraries
+such as {\tt winston}, 
 network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
-Although the fact that developer tools vary from browser to browser supposed a drawback, they all have the same functionality.
+Although the fact that developer tools vary from browser to browser
+supposed a drawback, they all have the same functionality. The
+JavaScript debugger can be accessed directly from the user menu in
+most browsers. 
 \item {\em Data Types} JavaScript uses floating point numbers with a 

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
one of the acronyms

@@
 early on, it was adopted by the industry and also by scientists,
-who used them for creating an EA on
+who used them for creating an evolutionary algorithm (EA) on
 the browser as early as 1998 \cite{jj-ppsn98}.

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
 design and release a new free framework that would allow anyone to
-create a volunteer distributed EC experiment using cloud resources as
+create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
 servers and browsers as clients.

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
   routes} 
-  that can be called for storing and retrieving information (the CRUD cycle:
+  that can be called for storing and retrieving information (the ''CRUD'' cycle:
   create, request, update, and delete) from the server. 
@@
 \includegraphics[width=4in]{system.pdf}
-\caption{Description of the proposed system. Clients execute a JS EA
+\caption{Description of the proposed system. Clients execute a JavaScript EA
   in the browser, which, every 100 generations, sends the best

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
 noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
-\item Volunteers are anonymous, only the IP address sent in 
+\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
 each request is known.
@@
 A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of Internet Protocol addresses (IPs)
+\ref{tab:summary:os}, which shows the median number of IPs
 intervening in each experiment,  median time needed

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
 this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. Thus, the user can decide if to continue sharing resources or not. % (Paloma) Feature or disadvantage?
+the evolutionary algorithm start again. Thus, the user can decide if to 
+continue sharing resources or not. % (Paloma) Feature or disadvantage?
 However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
@@
   one). 
-Shades
-of blue indicate how many experiments included that many unique IPs,
+Shades of blue indicate how many experiments included that many unique IPs,
 so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequency. From left to right, experiments 4/4, 4/24 and 7/31.}
+of computers happened less frequently, while darker shadow means more frequency. 
+From left to right, experiments 4/4, 4/24 and 7/31.}
 \label{fig:duration}
@@
 two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
-the client and the server where PUT and GET HTTP Methods are represented by
+the client and the server where PUT and GET HTTP methods are represented by
 red and blue arrows respectively. There are also messages between workers and
@@
 Asynchronous HTTP requests to the server are implemented using the
-XMLHttpRequest (XHR) object.  
+XMLHttpRequest object.  
 \end{itemize}

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
 \begin{table*}
-\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
+\caption{Summary of fit to GEV and Weibull distribution of
   the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
@@
 standard deviation in the time, with more volunteers adding both {\em
-  robustness} to an experiment, and certainty in the time it is taking. However, more experiments would be needed to check this hypothesis.
+  robustness} to an experiment, and certainty in the time it is taking. 
+ However, more experiments would be needed to check this hypothesis.
 
@@
 \begin{table*}
-\caption{Summary of fit to Generalized Extreme Value (GEV) and Weibull distribution of
+\caption{Summary of fit to GEV and Weibull distribution of
   the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym

@@
 computational cost of the challenge, the runtime required for 10,000 function 
-evaluations (FEs) for a particular configuration was published by organizers. The Java
+evaluations for a particular configuration was published by organizers. The Java
 implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
-3,000,000 FEs as termination condition. Using the Java implementation as
+3,000,000 function evaluations as termination condition. Using the Java implementation as
 a guide, the function described above was implemented to compare the performance of JavaScript against
-both implementations, again using the time required for 10,000 FEs. 
+both implementations, again using the time required for 10,000 function evaluations. 
 Before discussing the setup and performance results, certain details of
@@
 and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
- The performance of the time required for 10,000 FEs is presented in
+ The performance of the time required for 10,000 function evaluations is presented in
 Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
@@
 \includegraphics[width=0.9\linewidth]{f15_times.png}
-\caption{ Runtime of 10,000 function evaluations (FEs) for $F_{15}$} 
+\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
 \label{fig:f15_times}

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing an acronym
changing some appearances of system to frameworks

@@
 distributed evolutionary computation software frameworks available on
-the market, usually as open source systems \cite{Parejo12Survey}. Most of these systems have 
+the market, usually as open source systems \cite{Parejo12Survey}. 
+Most of these frameworks have 
 two features: they are {\em desktop} applications, that is, they must
@@
 
-There is, then, space for systems that do not make any of these
+There is, then, space for frameworks that do not make any of these
 assumptions, such as the one presented in this paper, which will be
@@
 clusters.
-Considering this, we leveraged 
-our experience in the design of web
+Considering this, we leveraged our experience in the design of web
 applications, including JavaScript, since their inception and other
@@
 
-When developing volunteer based systems, designers must
+When developing volunteer based frameworks, designers must
 also consider the whole {\em social} aspect in the design, with
@@
 computational systems along with attempts to predict and model its behavior. 
-Then, a detailed description of the proposed system is presented
+Then, a detailed description of the proposed framework is presented
 in Section \ref{sec:description}, followed by the baseline experiments
@@
 providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed by using desktop
-applications. Baratloo et al. tried in 1996 to create a system with
+URI, something that is not guaranteed by using desktop applications. 
+Baratloo et al. tried in 1996 to create a framework with
 the aforementioned qualities called {\em Charlotte}
-\cite{baratloo1996charlotte}. This system, which implemented a
+\cite{baratloo1996charlotte}. This framework, which implemented a
 distributed shared memory, was inspired by a previous work
@@
 will try to come up with a more precise model. Next we describe the
-system itself.
+framework itself.
 
@@
 Two kind of threats are possible in this environment; first, since the
-API is open and it is an open source system, it is relatively easy to
+API is open and it is an open source framework, it is relatively easy to
 sabotage the system, as indicated in \cite{domingues2007sabotage}, by
@@
 sources, client and server, for the application, are openly released. 
- We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
+ We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} 
+ and all data, once anonymized, as open data. This
 builds trust between the user and the scientist, which indeed has an
@@
 developed using JavaScript on the client as well as the server.
-All parts of the system are free and available with a free license
+All parts of the framework are free and available with a free license
 from \url{https://github.com/JJ/splash-volunteer}.
@@
   Chart.js} have
-been used to build the user interface elements of the system.
+been used to build the user interface elements of the framework.
 \begin{figure}[!t]
@@
 less than the second baseline. Thus, the conclusion in this case is
-that we can obtain, through volunteer computation, a result that most of the times is better than we would by using a similar, non-parallel,
+that we can obtain, through volunteer computation, a result that most 
+of the times is better than we would by using a similar, non-parallel,
 desktop setup, which makes this system suitable for massively
@@
 \includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
+and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
 \label{fig:gev:w2}
@@
 Test Suite of the companion competition of the Special Session and to give competitors an idea of the
-computational cost of the challenge, the runtime required for 10,000 function 
-evaluations for a particular configuration was published by organizers. The Java
+computational cost of the challenge, the runtime required for 10,000 function evaluations 
+for a particular configuration was published by organizers. The Java
 implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
@@
 in Firefox and Chrome browsers. Both functions are independent of the system clock. 
-\item {\em Tools} An important consideration when choosing certain language to write non-trivial programs is the availability of development tools. While developing
+\item {\em Tools} An important consideration when choosing certain 
+language to write non-trivial programs is the availability of development tools. While developing
 the experiments, we used a comprehensive set of developer tools available for both

### 2015-12-11 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
4-core
Closes #73
NodIO-W name
Closes #80

@@
 \end{verbatim}
-with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
+with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
 
@@
 \caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
-and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
+and Weibull (in red, left) for the {\sf NodIO-W$^2$} model.}  
 \label{fig:gev:w2}
@@
 
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO}-W$^2$ }
+\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }
 This experiment was deployed in a virtual private server with 512MB of memory

### 2015-12-12 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
NodIO-W in abstract
A sentence to talk about the Web Workers version.
Fixes Latex error
Closes #68, reference 63 is now 64 and new reference added.

@@
 browser, as well as any other client consuming its REST application
-programming interface.  The architecture, along with some initial
+programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
+presented, this version leverages the HTML 5 Web Workers specification
+to execute several clients per browser in the background, so that the the number of
+generations per user could be improved. The architecture, along with some initial
 results on the expected performance and scaling behavior when the
@@
 
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }
+\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}
+
 This experiment was deployed in a virtual private server with 512MB of memory

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Threats to vilidity
Please check

@@
 
+There are several validity threats to the design of this study.
+Participants of the experiments were mainly members of our 
+own social networks, personal colleagues, friends 
+and acquaintances, when extending this work other options 
+should be considered in order to include 
+a wider audience with different motives for participation.
+Another threat is that participants were anonymous so the amount of repeated 
+participations and the effects this had in the outcome is not
+known. Another issue is that the scalability of the system was not thoroughly tested
+but for the amount of participants the basic configuration 
+was more than adequate, again more experiments are needed in order
+to assess the capability both with more clients and other server
+configurations. Although care was taken to include representative benchmarks
+of real world problems, more experiments with a broader coverage 
+should allow a better understanding of the practical limitations of
+the architecture and the tools chosen.
+
 In general, this experiment confirms what was found in the others in

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Split section V
Closes #66

@@
 
+\subsection{Experimental trials}
+\label{sec:w2Exp}
+
 The experiments were run in the same way as before, using several
@@
 
-\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigin's Function}
 JavaScript has been traditionally implemented as an interpreted

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Single subsection removed
Closes #74
Volunteer´s Safety
Closes #76

@@
 volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
+adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
 reasonable performance. But hierarchy, this is, the capability of
@@
 
-\subsection{Volunteer evolutionary computation: first experiments and results}
-
 Initial experiments were set up using the OpenShift

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Fifty runs
Closes #79
More Acknowledgements

@@
 \end{figure}
-First, we establish the baseline performance by running the
+First, we establish the baseline performance by running fifty runs of the
 evolutionary algorithm in a desktop client written using {\sf NodEO}
@@
 SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
-GENIL project (CEI-BIOTIC Granada). We would also like to thank the
+GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
+Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
+We would also like to thank the
 anonymous reviewers of previous versions of this paper who have really

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing a verb in the abstract

@@
 \begin{abstract}
-Volunteer computing is extensively used 
-in science and industry, but it
-still has a long way to go in order to become
- a mainstream
-technology in the same way the cloud technologies in general have. The
-lack of 
-general purpose and open source tools with built-in analytics 
+Volunteer computing is extensively used in science and industry, but it
+still has a long way to go in order to become a mainstream
+technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
+The lack of general purpose and open source tools with built-in analytics 
 could be one of the reasons of why this is so, as well as the lack of
@@
 framework used to create client-server architectures for distributed
-evolutionary algorithms, entirely written  
-in JavaScript, which needs
+evolutionary algorithms, entirely written in JavaScript, which needs
 minimal server support to work and is able to support persistent,

### 2015-12-14 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing some words in the abstract

@@
 evolutionary algorithms, entirely written in JavaScript, which needs
-minimal server support to work and is able to support persistent,
+minimal server support to work and that is able to support persistent,
 asynchronous, distributed evolutionary algorithms running in the
@@
 programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
-presented, this version leverages the HTML 5 Web Workers specification
+presented; this version leverages the HTML5 Web Workers specification
 to execute several clients per browser in the background, so that the the number of

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Acronyms revised
Closes #61

@@
 early on, it was adopted by the industry and also by scientists,
-who used them for creating an evolutionary algorithm (EA) on
+who used them for creating an evolutionary algorithm (EA from now on) on
 the browser as early as 1998 \cite{jj-ppsn98}.
@@
 
-Two kind of threats are possible in this environment; first, since the
-API is open and it is an open source framework, it is relatively easy to
+Several kind of threats are possible in this environment; first, since its
+programming interface is open and its code is open source, it is
+relatively easy to find vulnerabilities and 
 sabotage the system, as indicated in \cite{domingues2007sabotage}, by
@@
 a particular chromosome.
-% (Paloma) Where is the second kind of threat?
-Another two different attacks are also
+Other two different kinds of attacks are also
 possible: a denial of service attack as well as penetration
 testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are openly released. 
- We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} 
+sources, client and server, for the application, are openly released
+as free software. 
+ We also release as an open access paper the result of the 
+ experiments performed \cite{DBLP:journals/corr/GuervosG15}  
  and all data, once anonymized, as open data. This
@@
 
-In this sense, in this paper we propose the {\sf NodIO} framework, a
+In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
 cloud or bare metal based volunteer evolutionary computing system
@@
 
-Figure \ref{fig:system} describes the general system architecture and
+The figure \ref{fig:system} describes the general system architecture and
 algorithm behavior. Different web technologies, such as JQuery or {\tt
@@
 systems. On the other hand, current JavaScript  
-virtual machines (VMs) are closing the gap, for instance the Google V8 
+virtual machines (VMs from now on) are closing the gap, for instance the Google V8 
 is an engine specifically designed for the fast execution of 

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Revised
Closes #62

@@
 \begin{abstract}
-Volunteer computing is extensively used in science and industry, but it
-still has a long way to go in order to become a mainstream
-technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
-The lack of general purpose and open source tools with built-in analytics 
-could be one of the reasons of why this is so, as well as the lack of
-proofs of concept. That is why in this paper we present {\sf NodIO}, a
-framework used to create client-server architectures for distributed
-evolutionary algorithms, entirely written in JavaScript, which needs
-minimal server support to work and that is able to support persistent,
-asynchronous, distributed evolutionary algorithms running in the
-browser, as well as any other client consuming its REST application
-programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
-presented; this version leverages the HTML5 Web Workers specification
-to execute several clients per browser in the background, so that the the number of
-generations per user could be improved. The architecture, along with some initial
-results on the expected performance and scaling behavior when the
-number of users is increased is presented. We also use
-the algorithm data to find out what kind of patterns the user behavior
-follows by examining the resources they devote to our experiment, and
-try to fit them to some statistical distribution to create the bases
-of a model of user behavior in volunteer computing experiments of this
-kind that can be used by researchers to predict its performance. 
-In general, our experiments prove, first, that {\sf NodIO} is a flexible and
+Volunteer computing is extensively used in science and industry, but
+it still has a long way to go in order to become a mainstream
+technology in the same way the cloud technologies in general are. %I
+have changed "have" to "are" The lack of general purpose and open
+source tools with built-in analytics could be one of the reasons of
+why this is so, as well as the lack of proofs of concept. That is why
+in this paper we present {\sf NodIO}, a framework used to create
+client-server architectures for distributed evolutionary algorithms,
+entirely written in JavaScript, which needs minimal server support to
+work and that is able to support persistent, asynchronous, distributed
+evolutionary algorithms running in the browser, as well as any other
+client consuming its REST application programming interface. An
+enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
+also presented; this version leverages the HTML5 Web Workers
+specification to execute several clients per browser and do it in the
+background, so that the the number of generations per user could be
+improved. The architecture, along with some initial results on the
+expected performance and scaling behavior when the number of users is
+increased is presented. We also use the algorithm data to find out
+what kind of patterns the user behavior follows by examining the
+resources they devote to our experiment, and try to fit them to some
+statistical distribution to create the bases of a model of user
+behavior in volunteer computing experiments of this kind that can be
+used by researchers to predict its performance.  In general, our
+experiments prove, first, that {\sf NodIO} is a flexible and
 low-overhead platform to perform distributed computation experiments
-that include volunteers, and then that user behavior follows some patterns,
-and, thus, are amenable to a model.
+that include volunteers, and then that user behavior follows some
+patterns, and, thus, are amenable to a model.
 \end{abstract}

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changing many "system" to "framework"
Where it was more precise to do it that way. Closes #63

@@
 distributed evolutionary computation software frameworks available on
-the market, usually as open source systems \cite{Parejo12Survey}. 
+the market, usually as open source frameworks \cite{Parejo12Survey}. 
 Most of these frameworks have 
@@
 such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript \cite{2015arXiv151101088M}. However, one of the advantages of our system is that
+JavaScript \cite{2015arXiv151101088M}. However, one of the advantages of our architecture is that
 clients do not need to install anything in their machines, they just
@@
 into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page running the system
+that voluntarily choose to enter the web page running the experiment
 determine its performance and how it can be modeled and, eventually,
@@
 distributed shared memory, was inspired by a previous work
-on parallel computing systems PVM and MPI and solved most of the issues by
+on parallel computing libraries PVM and MPI and solved most of the issues by
 using the Java virtual machine embedded in the browser,
@@
 In general, a distributed volunteer-based evolutionary computation
-system based on the browser is simply a client-server system
+framework based on the browser is simply a client-server system
 whose client is, or it can be, embedded in the browser via
@@
 % (Paloma) This sentence needs re-writing
-% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset.
-This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
+% "Other compiled languages such as Java or C might be, in fact,
+% faster, but the fact that single-user performance is higher (than
+% what?) is over the offset. This is because the number of spontaneous
+% volunteers we can gather at whim, including people using their
+% mobile phones or tablets, is several orders of magnitude higher,
+% which is the objective in this kind of systems." (?) 
  JavaScript is
@@
 impact on the performance of the system, since there is no need to
-include cheating checks or other functions that would degrade the
-performance of the system. This socio-technical approach is, indeed,
+include cheating checks or other functions that would degrade it. This
+socio-technical approach is, indeed, 
 coherent with our whole approach to the study of a distributed volunteer
- system which can only be studied by approaching it from the
+ system which can only be tackled by approaching it from the
 technical as well as the social, as in social media, angle.
@@
   to focus on the dynamics and measurements of the volunteer part of
-  the system.
+  the framework.
 
@@
 \includegraphics[width=4in]{system.pdf}
-\caption{Description of the proposed system. Clients execute a JavaScript EA
+\caption{Description of the proposed framework. Clients execute a JavaScript EA
   in the browser, which, every 100 generations, sends the best
@@
   \cite{gecco07:workshop:dcor}, which had a low scalability. 
-\item The system is {\em heterogeneous} since it does not need any
+\item The framework is {\em heterogeneous} since it does not need any
   performance, operating system, or even browser requirement as long
@@
 frameworks such as SETI@home \cite{javadi2009mining}. We obviously do
-not pretend to compare our system in scale or complexity with it, but
+not pretend to compare our framework in scale or complexity with it, but
 to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which also appears in our system. This
+certain pattern, found in SETI@home, and which also appears in our framework. This
 distribution is governed by three parameters, the usual location $\mu$
@@
 API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
-interaction between processes in a time sequence. In this system there are
+interaction between processes in a time sequence. In this version there are
 two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
@@
 sub-optimal approximations. In order to achieve this, the only change needed 
-in the system was to limit the number of chromosomes kept in the pool. 
+in the framework was to limit the number of chromosomes kept in the pool. 
 Previously, every PUT request which included a new chromosome increased 

### 2015-12-15 data/ips-puts-rastrigin.R
Changes fig 8 refs #70

@@ -13,9 +13,9 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
 
-ggsave('gev-fit-ww.rastrigin.png',width=5,height=5)
+ggsave('../img/gev-fit-ww.rastrigin.png',width=5,height=5)
 
 print(fitdistr(ips.puts.ww.rastrigin$puts,"gamma"))
 this.weibull <- fitdistr(ips.puts.ww.rastrigin$puts,"weibull")
@@ -24,8 +24,8 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
 y=sort(rweibull(length(ips.puts.ww.rastrigin$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
[…]

### 2015-12-15 data/ips-puts.R
Changes fig 8 (now really) refs #70

@@ -13,9 +13,9 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
 
-ggsave('gev-fit-ww.png',width=5,height=5)
+ggsave('../img/gev-fit-ww.png',width=5,height=5)
 
 print(fitdistr(ips.puts.ww$puts,"gamma"))
 this.weibull <- fitdistr(ips.puts.ww$puts,"weibull")
@@ -24,5 +24,5 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww$puts),
 y=sort(rweibull(length(ips.puts.ww$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
-ggsave('weibull-fit-ww.png',width=5,height=5)
[…]

### 2015-12-15 data/plot-rastrigin-fitness.R

@@ -2,5 +2,5 @@ library(ggplot2)
 
 rastrigin.fitness <- read.csv("rastrigin-fitness.csv")
 rastrigin.fitness$date <-  as.POSIXct(rastrigin.fitness$timestamp, "%Y-%m-%d %H:%M:%S")
-qplot(date,-Fitness,data=rastrigin.fitness)+ scale_x_datetime()
+qplot(date,-Fitness,data=rastrigin.fitness)+ scale_x_datetime()+ scale_colour_grey()
 ggsave("rastrigin-fitness.png",width=4,height=3)

### 2015-12-15 data/plot-zipf-openshift.R
Changes fig 5 refs #70

@@ -12,8 +12,8 @@ process.puts <- function(suffix ) {
                                     this.fit$estimate["loc"],
                                     this.fit$estimate["scale"],
                                     this.fit$estimate["shape"]),decreasing=T))
-    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))
-    ggsave(paste0("puts-openshift-",suffix,".png"),width=5,height=5)
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_grey(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))
+    ggsave(paste0("../img/puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }
     
@@ -22,7 +22,7 @@ dt.puts.4.4 <- process.puts("4-4")
 dt.puts.4.24 <- process.puts("4-24")
 dt.puts.7.31 <- process.puts("7-31")
 
-ggplot()+ scale_x_continuous("Normalized # runs")+scale_y_log10()+scale_colour_hue(name = '# Runs') +geom_point(data=dt.puts.4.4,aes(x=rank/length(dt.puts.4.4$rank),y=puts,color='4/4'),stat='Identity')+geom_point(data=dt.puts.4.24,aes(x=rank/length(dt.puts.4.24$rank),y=puts,color='4/24'),stat='Identity')+geom_point(data=dt.puts.7.31,aes(x=rank/length(dt.puts.7.31$rank),y=puts,color='7/31'),stat='Identity')
[…]

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Rewrite titles of sections
Closes #65
Ditto fig 7 refs #70

@@
   one). 
-Shades of blue indicate how many experiments included that many unique IPs,
+Shades of blue (or gray in grayscale) indicate how many experiments included that many unique IPs,
 so lighter shade for a column of dots indicates that a particular number
@@
 \caption{Duration of experiments vs. rank, with $y$ axis in a
-  logarithmic scale. Dot color is related to the number of IPs
+  logarithmic scale. Dot color (or gray value) is related to the number of IPs
   participating in the experiment. From left to right, experiments
@@
 \caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
-  Extreme Value distribution (in red). From left to right, experiments
+  Extreme Value distribution (in a lighter shade of gray). From left to right, experiments
   4/4, 4/24 and 7/31.} 
@@
 \caption{Time employed in every experiment vs. number of IPs (in
-  abscissas). The red dot and line show the average and standard
-  deviation, the shade of blue the number of experiments with the same
+  abscissas). The black dot and line show the average and standard
+  deviation, the shades of gray the number of experiments with the same
   number of IPs. } 
 \label{fig:ipstime:w2}
+% ../data/time-vs-IPs.R 
 \end{figure}

### 2015-12-15 data/time-vs-IPs.R

@@ -6,7 +6,8 @@ for ( i in 1:nrow(ips.time) ) {
     ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
 }
 ips.time$i <- 1:nrow(ips.time)
-ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
+ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ scale_colour_gradient(name  ="Count", low="white",high='gray70')+ stat_summary(fun.data = "mean_cl_boot", colour = "black")
+ggsave("../img/ips-time-ww.png", width=5,height=5)
 ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
 
  print(paste("< 69s", length(ips.time[ips.time$milliseconds < 69000,]$milliseconds)/length(ips.time$milliseconds)))

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Fixing fig 12 closes #70

@@
 \includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
-and Weibull (in red, left) for the {\sf NodIO-W$^2$} model.}  
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in light color, right) 
+and Weibull (in lighter shade, left) for the {\sf NodIO-W$^2$} model.}  
 \label{fig:gev:w2}
@@
 \includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
-\caption{Fit to GEV (left) and Weibull (right) of the number of {\tt PUT}s per worker for the F15 (Rastrigin) function.} 
+\caption{Data and fit to GEV (left) and Weibull (right) of the number
+  of {\tt PUT}s per worker for the F15 (Rastrigin) function.}  
 % Plotted with ../data/ips-puts-rastrigin.R

### 2015-12-15 data/ips-puts-rastrigin.R

@@ -40,7 +40,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin.workers$puts),
 
 ggplot() +geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
 
-ggsave('../img/gev-fit-ww.rastrigin.workers.png',width=5,height=5)
+ggsave('../img/gev-fit-ww-rastrigin-workers.png',width=5,height=5)
 
 print(fitdistr(ips.puts.ww.rastrigin.workers$puts,"gamma"))
 this.weibull <- fitdistr(ips.puts.ww.rastrigin.workers$puts,"weibull")
@@ -51,4 +51,4 @@ y=sort(rweibull(length(ips.puts.ww.rastrigin.workers$puts),
     this.weibull$estimate["scale"]),decreasing=T))
 print(this.weibull)
 ggplot()+geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_grey(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
-ggsave('../img/weibull-fit-ww.rastrigin.workers.png',width=5,height=5)
+ggsave('../img/weibull-fit-ww-rastrigin-workers.png',width=5,height=5)

### 2015-12-15 data/time-vs-IPs-openshift.R
Fixing fig 3 refs #71 #70

@@ -7,10 +7,10 @@ time.vs.IPs <- function( data, preffix ) {
         data[i,]$howmany <- length(data[data$IPs==data[i,]$IPs,]$milliseconds)
     }
     data$i <- 1:nrow(data)
-    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "red")
-    ggsave(paste0("time-vs-ips-OS-",preffix,".png"))
+    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="white",high='gray70')+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=20))
+    ggsave(paste0("../img/time-vs-ips-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
-    ggsave(paste0("duration-vs-time-OS-",preffix,".png"))
+    ggsave(paste0("../img/duration-vs-time-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
     ggsave(paste0("duration-vs-time-OS-bar-",preffix,".png"))
     ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.x=element_text("Ordered experiments"))

### 2015-12-15 data/time-vs-IPs-openshift.R
Fixing fig 4 closes #71 closes #70

@@ -13,8 +13,8 @@ time.vs.IPs <- function( data, preffix ) {
     ggsave(paste0("../img/duration-vs-time-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
     ggsave(paste0("duration-vs-time-OS-bar-",preffix,".png"))
-    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.x=element_text("Ordered experiments"))
-    ggsave(paste0("time-vs-rank-OS-",preffix,".png"))
+    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='white', high='black') + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), text = element_text(size=20), panel.grid.major = element_blank(), panel.grid.minor = element_blank() )+labs( x="Ordered experiments")
+    ggsave(paste0("../img/time-vs-rank-OS-",preffix,".png"))
 }
 
 suffix <- c("4-4","4-24","7-31")

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Fixing fig 3 refs #71 #70
Fixing fig 4 closes #71 closes #70
Reorganizes sentences

@@
 %---------------------------------------------------------------
-\section{Baseline Experiment and obtained results} %Cambié el nombre
+\section{Baseline Experiments} %Cambié el nombre
 \label{sec:experiments}
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which this 
-experiment could be performed. That way, and as required, the hardware
+PaaS, which provides a free tier within which these runs could be performed. That way, and as required, the hardware
 and cloud cost for this experiment were zero. Experiments were
@@
 \end{figure}
+% ../data/time-vs-IPs-openshift.png 
 %
@@
 \label{fig:zipf:os}
+% computed via ../data/time-vs-IPs-openshift.png 
 \end{figure}
@@
 
-\section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
+\section{Enhancing NodIO with a Web Worker-based architecture}
 \label{sec:w2}

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
There was a comment inside the text
Reviewing English grammar from abstract. Couple of changes and proposal for dividing a sentence

@@
 it still has a long way to go in order to become a mainstream
-technology in the same way the cloud technologies in general are. %I
-have changed "have" to "are" The lack of general purpose and open
+technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
+The lack of general purpose and open
 source tools with built-in analytics could be one of the reasons of
-why this is so, as well as the lack of proofs of concept. That is why
+why this is so, as well as the lack of proofs of concept.
+% (Paloma) The following sentence is way too long, and the "which" after "Javascript" is confusing, so one does not know if the which refers to the framework or Javascript by reading the whole sentence.
+% I propose: "That is why in this paper we present NodIO, a framework used to create client-server architectures for distributed evolutionary algorithms, entirely written in JavaScript.
+Our framework needs minimal server support to work and is able to support persistent, asynchronous, distributed evolutionary algorithms running in the browser, as well as any other client consuming its REST application programming interface."
+That is why
 in this paper we present {\sf NodIO}, a framework used to create
@@
 expected performance and scaling behavior when the number of users is
-increased is presented. We also use the algorithm data to find out
+increased is introduced. % (Paloma) I've changed this "presented" for a synonym
+We also use the algorithm data to find out
 what kind of patterns the user behavior follows by examining the
-resources they devote to our experiment, and try to fit them to some
+resources they devote to our experiment, and try to fit them in % (Paloma) Changed. Phrasal verb was wrong.
+ some
 statistical distribution to create the bases of a model of user
 behavior in volunteer computing experiments of this kind that can be
-used by researchers to predict its performance.  In general, our
+used by researchers to predict their performance. %(Paloma) Changed to "their" because of "experiments"
+In general, our
 experiments prove, first, that {\sf NodIO} is a flexible and

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from Introduction. Couple of commas, basically, and minor change for the sake of clarity when reading

@@
 Considering this, we leveraged our experience in the design of web
-applications, including JavaScript, since their inception and other
+applications since their inception, including JavaScript, and other
 volunteer and
@@
 create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
-servers and browsers as clients.
+servers, and browsers as clients.
 % (Paloma) I think you should rewrite this whole sentence.
@@
 also consider the whole {\em social} aspect in the design, with
-issues related to security, trust and privacy among others. 
-The computing system becomes a {\em techno-social system} \cite{vespignani2009predicting}.  
+issues related to security, trust, and privacy among others. 
+The computing system becomes then a {\em techno-social system} \cite{vespignani2009predicting}.  
 % (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from State of the Art. I left a couple of comments/questions

@@
 Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers. This paper also establishes the desirable features of
+metacomputers. This paper % (Paloma) No really need for that "also", plus it complicates reading
+establishes the desirable features of
 any {\em global computing} infrastructure, that can also be applied to
@@
 independently of the operating system, and access to a common data store in the
-server. Since the Java virtual machine is a sandbox, it is safe to the user
-and as it runs compiled bytecode it also guarantees a certain level of
-performance.
+server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
+and as it runs compiled bytecode it also guaranteed a certain level of
+performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
 Other issues such as scheduling, which are related to adaptive 
@@
 later on \cite{sarmenta-bayanihan} as well as the potential of
-user misconduct when having access to the source code.
-% (Paloma) I think this sentence is extremely informal
-\cite{sarmenta-sabotagetolerance}. However, these early efforts by
+user misconduct when having access to the source code \cite{sarmenta-sabotagetolerance}. 
+However, these early efforts by
 Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
@@
 need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower.
+threshold to join them is much lower. % (Paloma) Do yo mean... higher? A low threshold means that is easy to happen something.
 
@@
 metaheuristics optimization is that it allows an improvement in communicability,
-reproductibility, interoperability, automated assembly, knowledge engineering
+reproductibility, interoperability, automated assembly, knowledge engineering,
 and efficiency.
@@
 the algorithm adaptive to the kind of resources allotted to it
-\cite{milani2004online}, which is actually not such a big problem in
+\cite{milani2004online}. This is actually not such a big problem in
 algorithms such as the EA that can easily be
-parallelized via population splitting or farming out evaluations to all
+parallelized via population splitting or by farming out evaluations to all
 the nodes available. Lately, several approaches have focused on the
@@
 the more general context of distributed computing 
-\cite{nogueras2015studying} or including it in a more general study of the
+\cite{nogueras2015studying} or be included in a more general study of the
 performance of the EA itself
@@
 run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining} found out that the Weibull, log-normal and
+\cite{javadi2009mining} found out that the Weibull, log-normal, and
 Gamma distribution 

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from description of the framework. Couple of typos.

@@
 to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
-is available in most laboratories, allowing researches to
+is available in most laboratories, allowing researchers to
 scale algorithms at reasonable costs. Researchers have proposed the use of
@@
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
+be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
  is
@@
 many other outside, so researchers can, in fact, write their fitness
-function in Python, Erlang, Scala, or even Java, and them compile it to
+function in Python, Erlang, Scala, or even Java, and then compile it to
 JavaScript. For completeness, an experiment comparing the performance of
@@
 as free software. 
- We also release as an open access paper the result of the 
+ We also released as an open access paper the result of the 
  experiments performed \cite{DBLP:journals/corr/GuervosG15}  
@@
   routes} 
-  that can be called for storing and retrieving information (the ''CRUD'' cycle:
+  that can be called for storing and retrieving information (the ``CRUD'' cycle:
   create, request, update, and delete) from the server. 

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from experiments. Minor changes

@@
 
-\section{Enhancing NodIO with a Web Worker-based architecture}
+\section{Enhancing {\sf NodIO} with a Web Worker-based architecture}
 \label{sec:w2}
@@
 Previously, every PUT request which included a new chromosome increased 
-the size of the pool, in a long-running experiment this has to be avoided
+the size of the pool. But in a long-running experiment this has to be avoided
 in order to not exhaust the available memory. In this experiment the size
-of a chromosome was 1000 double precision numbers, considering this
-and the memory available in the server, the the pool was limited 
+of a chromosome was 1000 double precision numbers, so that considering this
+and the memory available in the server, the pool was limited 
 to 10,000 chromosomes.  In order to keep this size
-constant a random chromosome was pulled out of the pool before inserting a
+constant, a random chromosome was pulled out of the pool before inserting a
 new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
@@
 were tournament size = 3 and  population size = 500, returning exchanging a
-chromosome with the server every 100 generations. Again two workers in each window.
+chromosome with the server every 100 generations. Again, there were two workers in each window.
 As before, a few participation calls were published in social networks.
@@
 
+% (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
 There are several validity threats to the design of this study.
@@
 own social networks, personal colleagues, friends 
-and acquaintances, when extending this work other options 
+and acquaintances, therefore to extend this work other options 
 should be considered in order to include 
@@
 Another threat is that participants were anonymous so the amount of repeated 
-participations and the effects this had in the outcome is not
+participations and the effects this might have had in the outcome is not
 known. Another issue is that the scalability of the system was not thoroughly tested
 but for the amount of participants the basic configuration 
-was more than adequate, again more experiments are needed in order
+was more than adequate, then again more experiments are needed in order
 to assess the capability both with more clients and other server

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewed conclusions. Some important changes about collocations or readability

@@
 In this paper two versions of a client-server architecture for volunteer and distributed
-evolutionary algorithms that uses the browser and generated using {\sf
-  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
-evaluated. Volunteers are {\em in the cloud}, as stated in the title,
+evolutionary algorithms have been evaluated. The architecture uses the browser and has been generated using {\sf
+  NodIO}, based on the {\sf NodEO} evolutionary algorithm library. Volunteers are {\em in the cloud}, as stated in the title,
 since they are a {\em CPU as a service} for the persons running the
@@
 on the real size of that {\em cloud} and how it can be used standalone
-if there is no alternative or in conjunction with other local or
+if there is no alternative, or in conjunction with other local or
 cloud-based methods to add computing power in a seamless way through
-the pool that NodIO creates.
+the pool that {\sf NodIO} creates.
 
@@
 baseline was possible, it did not happen, mainly because of the
-difficulty in carrying over volunteers to other experiments when the
-one they were participating on was finished, shown in the low
-correlation between the number of IPs in successive experiments. This
-also resulted in a low number of generations allotted by users.
+difficulty in carrying over other experiments to volunteers when the
+experiment they were participating on was finished. This has been shown in the low
+correlation between the number of IPs in successive experiments, and also resulted in a low number of generations allotted by users.
 
@@
 the model depends on the implementation, with contributions following
-a General Extreme Values distribution in the case of {\sf NodIO} and a
+a GEV distribution in the case of {\sf NodIO} and a
 Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
@@
 want to have returning users for the experiments, it is probable that
-we should {\em gamify} the experience so that once they've done it
+we should {\em gamify} the experience so that once they have done it
 once, they might do it more times. In the spirit of Open Science, this
-gamification might involve computing in real time data such as the one
-presented in this paper and showing it in the same page.
+gamification might involve computing in real-time data such as the one
+presented in this paper, while showing it in the same page.
 
@@
 role. Even as it is difficult to create controlled experiments in this
-area, it is an interesting challenge to explore in the future.
+area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation
 
@@
 has had a big influence on the results and in fact this is not noticed
-for F15, which is the second version tested but it would be interesting to
+for F15, which is the second version tested, but it would be interesting to
 measure exactly what this influence has been. In general there are

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
Adds response to reviewers refs #81

--- 2015-11-18 special-issue-ci/Response to reviewers.tex
@@
 %\usepackage{setspace}
-\usepackage{url}
+\usepackage[hyphens]{url}
 \usepackage[english]{babel}
-
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
 \begin{document}
@@
 
-\title{Volunteers in the Clouds: an Architecture for Low Cost and
-  Potentially Massive Distributed Evolutionary Computation: Response to Reviewers\' comments}
+\title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
+for Low Cost Distributed Evolutionary Computation: Response to Reviewers\' comments}
 
@@
 
-We would like to notify the editors that, thanks to changes and
-aportations made to this new version of the paper there are two new
-coauthors.
-
 In the next sections you can find these comments and our responses
-to every one, along with the changes done.\\
+to them, along with the changes done.\\
 
@@
 
+\section{Response to Editors}
+
 \section{Response to Reviewer \#1}
 
-\begin{quote}
-The main contribution of this paper is the proposal of a Javascript based distributed system intended to
-run evolutionary algorithms according to the volunteer computing scheme. The idea is that user can
-contribute to the computation by using only a browser, without needing to install any plugin nor a Java
-virtual machine.
-The volunteer computing model appeared some years ago and it has been used in many fields. The
-model is effective to solve problems demanding high computational capabilities, and at this point is where
-this paper proposal introduces some concerns. The choice of Javascript code running on browsers has
-some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
-motivation of these kinds of distributed models is to solve complex problems, using Javascript doesn\'{}t
-appear a right choice. It\'{}s true that is a very commonly used programming language, but in the context of
-Web applications, not in high performance ones. Apart from the fact of being an interpreted language
-designed for the Web, it is difficult to think on a computer scientist writing a complex math code in
-Javascrip. If the authors would have used as a case study a complex optimization problem then they
-could have provided numerical data to sustain their claims; however, the selected problem (which should
-be formulated in the paper) seems a rather academic one.
-\end{quote}
+\begin{quote}\textbf{
 
-We thank the reviewer for the comment. The only reason why JavaScript
-was chosen was because it is the only language that is native to the
-browser, allowing to carry out precisely this kind of experiments
-where the user is invited to click on an URL and start giving CPU to
-it. Although there are many other volunteer frameworks, written in
-other languages, that a priori would offer a higher performance, the
-key is that they would {\em while they are running}, that is, as soon
-as it has been installed, the user has decided to fire it up, and it
-does not affect the performance of the rest of his system, which is a % se confundia con el desempeño del todo el sistema
-concern in this kind of systems. We have performed speed tests on
-JavaScript and other compiled languages resulting that, in fact,
-JavaScript uses a JIT compiler that provides a good performance; we
-have published these tests as an open access technical report. \cite{2015arXiv151101088M}
-
-JavaScript is probably the most popular language nowadays, and this
-has inspired the appearance of many compilers, so that the fitness
-function can, in fact, be written in any language and compiled to
-JavaScript, as is shown in this list:
-\url{https://github.com/jashkenas/coffeescript/wiki/list-of-languages-that-compile-to-js}. This
-has also been added to the text.
-
-In a nutshell, what we are saying is that, as is proven in the paper,
-a JavaScript application, running in the browser, can gather in
-seconds dozens of users. That would be, in practice, impossible with
-other kind of native clients, who would have to download and maybe
-compile the application, losing many seconds or minutes for the
-optimization problem that has to be solved.
-
-These facts, along the explanations, have now been added to the first
-paragraph of the {\em Description of the system} section, as well as
-to the introduction.
-
-\begin{quote}
-The paper organization is not clear. Some comments about this:
-- The title "Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
-Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
-the server run in the cloud and the volunteers do in the browsers of
-common PCs and laptops.
-\end{quote}
-
-That is correct. We just used the term {\em cloud} in a relatively informal
-sense because they are available as in a cloud for running the
-experiment. That is, from the point of view of the person running the
-experiment, there is a {\em cloud} out there running your experiments,
-which is actually run by volunteers in their browsers. We went to the
-original acceptation of ``cloud'' as a ``network'' more than the more
-precise and technical sense used today. We have changed the title to
-avoid words like {\em potentially}, and we have no strong feelings on
-the ``volunteers in the cloud'' part. We like it, but if asked by the
-editors, we can eliminate it.
-
-We have added this explanation to the first paragraph in the
-conclusions; we hope this makes our intention clearer, but we can also
-eliminate that part of the title if the reviewer thinks it is
-necessary.
-
-\begin{quote}
-- The paper abstract does not properly summarize the paper contents. It says that ``That is why in this
-paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !'', but there
-are two versions of the architecture, as stated later in the conclusions: "In this paper two versions of a
-client-server architecture for volunteer and distributed evolutionary !". That second version is neither
-commented in the introduction.
-\end{quote}
-
-In fact {\sf NodIO} is a framework that can be used to generate
-different client-server architectures. This has been clarified in the
-abstract and conclusions section.
-
-\begin{quote}
-- According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
-presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV
-does not only contains experimental result but also the full description of the second version of the
-architecture.
-\end{quote}
-
-You are absolutely right. We have corrected that paragraph to include
-the actual contents of the paper.
-
-\begin{quote}
-Summarizing the review, the authors should provide a convincing example showing the advantages of
-their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
-by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
-problem.
-\end{quote}
-
-We have added a new experiment (Section V.A), a more complex and
-taxing problem that also uses a web-worker architecture, in order to
-prove how this can be 
-used in real-life problems. In this case, we have achieved many more
-simultaneous users, up to 25 in some minutes and more than 100 in a
-single hour. All in all, more than 400 volunteers (identified by
-unique IPs) have participated in
-the experiment.
-
-In fact, our architecture {\sf NodIO} is
-not an exclusively browser-based one. Its pool-based architecture allows
-any other kind of client, including local ones, to be incorporated to
-the experiment, with the lightweight overhead of doing asynchronous
-HTTP requests, which being local are indeed negligible.
-
-Our framework, in fact, does not prevent the user to carry out any other kind of
-experiments. If there is a very efficient locally run algorithm, it only
-needs to be able to do HTTP requests to work alongside a {\em
-  volunteer cloud}. What our framework ensures is that all work done
-by volunteers can be integrated with local, cloud or other kind of
-resources. This explanation has been also added as a paragraph in the
-Introduction to the paper.
-
-We would like to thank this reviewer for the helpful comments and
-suggestion made. We expect to have addressed them properly, but in any
-case are grateful because it has allowed us to improve the quality of
-our work.
+}\end{quote}
 
@@
 
-\begin{quote}
-Contribution:
-This paper describes a client-server architecture for distributed
-evolutionary algorithms named NodIO. 
-The implementation of the architecture is written entirely in
-JavaScript, in order to support persistent, 
-asynchronous, distributed evolutionary algorithms running in the browser.
-Additionally, the paper presents the results of several experiments
-performed to validate the proposal and 
-to measure the performance improvement generated by the massive parallel computing.
-I think that the contribution is quite relevant, since it provides a
-computation platform that is flexible, 
-extensible and potentially massive. Personally, I consider it one of the fresher and most promising
-proposals in the context of metaheuristic optimization frameworks lately. Hence, our global
-recommendation of acceptance with minor changes.
-\end{quote}
-
-Thanks for your appreciation.
-
-\begin{quote}
-However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
-organization and description of the platform, and the presentation and writing style.
-Regarding the experimentation, I miss:
--  a section discussing the threats to validity of the experiments
-
-\end{quote}
-
-We have inserted a paragraph at the end of the section that describes
-the experimental setup describing what are the challenges that this
-kind of voluntary framework must overcome. Another paragraph analyzing
-the security from the point of view of the framework used has been
-also added. A new paragraph on the security challenges has also been
-inserted, third paragraph in this section, to clarify our integral and
-pragmatic approach to security, which so far has worked perfectly.
-
-\begin{quote}
--  an experiment with a more difficult problem (please use the well-known benchmark problems
-described in literature).
-\end{quote}
-
-An experiment based on the well known Rastrigin's function has been
-performed and analyzed. It is now Subsection V.A.
-
-\begin{quote}
--  A comparison in similar conditions of the two implementations of the architecture (the algorithms
-used in the experiments are slightly different, but this could lead to dramatic changes in
-performance, and consequently, invalidate the comparison performed in the paper). This would
-enable a measurement of the actual improvement in performance obtained by the use of web
-workers (not that generated due to a different algorithm)
-\end{quote}
-
-This has been included as the last subsection of the paper previous to
-the conclusions. The new
-experiment has a constant population instead of using  population of random size.
-We have not found a big influence on the result.
-
-\begin{quote}
-Regarding the organization of the paper, some important details regarding the improved framework
-design named is described in the experimentation section. I would reorganize the contents to describe the
-improved in detail in section 3, and focus on the experiments, its
-results and the discussion in section 4.
-\end{quote}
-
-In fact, reorganizing the paper increased the chance of introducing
-new issues in this new version. We have preferred to keep it the way
-it is, although we thank you for the suggestion. We expect that the
-changes made to the text make it clear enough anyway.
-
-\begin{quote}
-Another issue is that some details are described twice, for instance,
-the advantages of the use of web 
-workers. Fixing this issue can reduce the paper size and improve its
-readability.
-\end{quote}
-
-We have eliminated duplications in this new version. Thanks for
-noticing it.
-
-\begin{quote}
-Regarding the presentation:
--  Figures should be readable even in gray-scale, so do not use red and blue to denote different
-elements, use a different shade or texture.
--  The axis labels and legends in most of the figures are unreadable, please increase its font size.
--  There are some typos and phrases that need rewriting.
-\end{quote}
-
-We have rewritten and revised extensively the paper, and tried to find
-a way (and failed) to make it more readable for reviewers. In fact,
-they use different shades, so it should mostly work in grayscale. For
-the time being and 
-during the revision phase, we prefer to keep it that way, but we will
-change it to grayscale if it is required for the final version. 
-In fact, the labels are not so important, so increasing the font size
-would in fact distract the attention from the figure itself, which is
-the interesting information we want to convey with the graph. Once
-again, we can do it if required for the final version.
+\begin{quote}\textbf{
+}\end{quote}
 
@@
 
-\begin{quote}
-1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
-the system called NodIO consists of a REST server (coded in Node.js) and clients performing
-evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
-entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
-automatic restart of computations on the client's side can improve the performance of the system. The
-state of the art is comprehensive and provides good introduction into the subject.
-\end{quote}
-
-We thank this reviewer for these helpful comments.
-
-\begin{quote}
-2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
-Software" special issue. Thanks to both development of Internet technology and continuous increase in
-computing power of electronic devices, including mobile devices, the web browser based distributed
-computing can be a cheap alternative for grid and high-performance computing systems that require
-high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
-architecture presented in the paper can be potentially used for solving many different time consuming
-problems, however, the authors limited their experiments to a simple ``Trap'' function that can be easily
-solved on a desktop machine, what may not be convincing for a reader looking for an effective
-optimization tool.
-\end{quote}
-
-It is true that that function can be easily solved with any desktop
-machine, but we can still do that {\em and} use a volunteer
-computation framework, which will add performance to the desktop
-machine. Our intention was more to prove what kind of performance can
-be expected and how this performance is distributed, although we have
-also proved in the last section how the performance achieved by a
-volunteer-based computer system can, in fact, be better than a desktop, even
-a powerful one. This has also been clarified in the introduction and
-conclusions.
-
-At any rate, we have also performed a new experiment based on the
-Rastrigin's function, which in fact cannot be solved in a few minutes
-(or seconds, as in the case of Trap) by a desktop system.
-
-\begin{quote}
-3. Correctness: The system works correctly - it is not very complex and the authors provide its source
-code (this is a great plus!). Technical description of the system is also clear, however it lacks some
-details like parameters of the machine used to determine baselines for the experiments or the format of
-messages sent between the server and the clients. The only thing I noticed is that the authors used the
-term ``multiple CPUs'' in the description of Web Worker API advantages (page 8) instead of "multi-core
-CPUs". As far as I know, contemporary web browsers do not support
-multiple CPUs architecture.
-\end{quote}
-
-This has been clarified in the new version of the paper, third
-paragraph from the beginning of the section that explains the web
-worker-based architecture. In fact, most browsers can use multi-core
-CPUs running web workers as independent processes in each one of them.
-The base system has been now described in the first paragraph of the
-section on the experiments, it is a desktop Intel i7-4770.
-
-\begin{quote}
-
-4. Style: Paper is generally written in a clear way. Statements like
-``But fair is fair,'' (page 6), sound a bit colloquially in a scientific
-paper and may be omitted. However, my main concern is the readability
-of the graphs. General remark - font used for description of axes
-
-\end{quote}
-
-We have eliminated colloquialisms from the text.
-
-\begin{quote}
-should be significantly larger. In Fig. 2 the authors can change scale
-for axe Y, as the box for the population of 1024 individuals is very
-small.  What do the authors mean by ``times''? Time in seconds? I am not
-sure if this Figure is necessary at all as the description is
-understandable without it. In Fig.4 all graphs are similar and only
-the one in the middle is analyzed, so maybe others are not necessary?
-\end{quote}
-
-The labels, and indeed the label size, are not too important; the
-graph shape is. That is one reason they are grayed out. However, we
-have changed as instructed Figure 2, which is more readable now. We do
-mean times to solution, shortened for clarity. We have also clarified
-that on the caption. We also think that all graphs are needed in Fig
-4, since they represent different experiments, and we have added a
-sentence to highlight that the {\em tail} appears in all three.
-
-\begin{quote}
-The same is with Fig. 5. What does ``x'' stand for in the description of
-axe X? (the same is with Fig. 8).
-\end{quote}
-
-$x$ has been changed to {\sf rank}, which is the usual way of
-representing Zipf's law, a measure (in this case number of {\tt PUT}s
-vs. rank. We have also changed size, color and sigils to represent
-each point to make it a bit clearer.
-
-\begin{quote}
-
-5. Length: As I mentioned above the authors should think of removing some of the unnecessary graphs
-in order to improve readability. They admit themselves that it is hard to conclude from some of the
-graphs (Fig. 5 and Fig. 8). Nevertheless, overall length of the paper is fine.
-
-\end{quote}
-
-Finally, since the reviewer thinks that length is fine and for the
-reasons outlined above, we have decided to keep the graphs and explain
-them a bit better.
-
-\begin{quote}
-6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
-distributed computing system, contrary to the description in the main text of the paper. However, it
-lacks the information about the second version of the system called NodIO-W2 based on the Web
-Worker architecture.
-\end{quote}
-
-We have extended the abstract stressing our main points, that were
-first to test the validity and measure the performance of our
-framework NodIO and then to use the measures as first steps to a model
-of volunteer uses.
-
-\begin{quote}
-Comments to Authors:
-
-1) I like the state of the art which includes comprehensive overview of the most important papers related
-to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
-with its characteristic provided on page 4. Namely:
-
- - Heterogeneity - it is true that almost all devices can participate in the web browser based
- distributed computing system, providing that a user did not disable Java Script in its browser.
- However if the authors plan to use HTML5 Web Workers only modern web browsers can
- participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
- Windows Phone) are also excluded.
-
-\end{quote}
-
-That is true, but it makes it anyway the most inclusive system so
-far, as we have tested it with smartphones, tablets and computers. In
-general, nowadays with rich internet applications disabling JavaScript
-is not common, but anyway we do not aim for {\em all} clients, we have
-enough with {\em most}. Windows Phone, for instance and maybe
-fortunately, is not too common. % ¿no existe Chrome para Windows Phone? De todas formas yo quitaría lo de "and maybe fortunately"...
-
-\begin{quote}
- - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
- implement evolutionary algorithm in that way that it will automatically adopt to the problem type
- and the environment in which it is computed, the authors do not mention about how real
- adaptiveness of the system can be achieved. The same is with scalability. Only 5 different users
- that permanently participated in the experiments are not enough to assess the scalability of the
- system, what is a crucial parameter for distributed computing.
-\end{quote}
-
-We have performed a new experiments in which we have had hundreds of
-simultaneous users. In fact, with the experiments already in the
-previous version of the papers 5 was a central measure, but there were
-peaks of dozens. 
-Adaptativity is by design, since it uses any browser and does not need
-any special provision for new systems. This has also been explained in
-the expanded text.
-
-\begin{quote}
-
- - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
- code of evolutionary algorithm is sent to the client as a source code in Java Script, so some
- experienced user may quite simple decode what has to be computed. It can may come into
- possession of sensitive data or/and alter this code in order to disrupt the computations or make
- them worthless. The results send to the server may also be captured and/or altered without any
- problem since the transmission is not secure.
-
- In my opinion you should at least highlight above issues.
-\end{quote}
-
-We have also justified our {\em pragmatic} approach to security
-increasing trust by the user, but also monitoring their behavior to
-detect it and eventually ban users. So far, this has not happened.
-
-\begin{quote}
-2) Regarding the experiments I am very skeptical about their real value. You chose the ``Trap''
-function. in which setting all bits to zero gives local maximum and setting all bits to one gives global
-maximum. In the experiments you counted how many times evolutionary algorithm was able to find
-global maximum for the problem of size n=40. Initially you have found that if the population in an
-evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
-within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
-(page 4, if I understood it correctly).
-\end{quote}
-
-I fact 3.46 is the average for all 50 runs. 
-We have emphasized the word average to make that clearer.
-
-\begin{quote}
-Thus, what was the point of using distributed computing to solve such simple problem? Why the
-authors did not perform their experiments for larger values of n, what could justify distributed problem
-solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
-6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
-it is still behind the desktop computer in more than 60% cases (when compared to the second
-baseline). I suggest to perform additional computing experiments for some higher value of n in the
-``Trap'' function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
-to focus the paper on the behavior analysis of volunteers participating in a distributed computing
-system, what may be also interesting an valuable for the potential
-readers.
-\end{quote}
-
-In fact, that was never the point of the paper. We have chosen Trap as
-a relatively difficult problem that allows us mainly to test scaling
-and measure behavior. In a production environment, a volunteer-only
-environment would only be used if there is absolutely no other
-resource available; in general, it would be combined with the rest of
-the resources available to the experimenter: cloud or local; it would
-make no sense to not use them if they are available and in fact, NodIO
-allows to combine them seamlessly in a single system. That has been
-explained a bit better in the introduction and conclusions
-section. Besides, we have added a more difficult problem, the shifted
-Rastrigin function (a well known function optimization problem), to make 
-measures in a different scenario (one in which the problem is not solved 
-in a few seconds).
-
-\begin{quote}
-3) Regarding the graphs I have already put my comments in Style section, but I have one question. In
-Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
-meaning probably that IPs from 16 to 29 did not performed any computations. What was the reason
-for that?
-\end{quote}
-
-The $x$ axis runs overs number of computers participating in an
-experiment; that is, $4$ plots the time taken for all experiments in
-which 4 computing nodes participated and the red dot is the average;
-it shadows the time in the case there is a single occasion. In the
-case you mention, there are no experiments in which there were exactly
-17 nodes, for instance. The ``red dots'' indicate that there was a
-single experiment in which that many nodes participated. We have
-clarified the caption to Figure 3 to make this clearer.
-
-\begin{quote}
-For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC
-systems - the number of participants is completely incomparable and also the type of problems
-solved by the systems are of different nature and complexity.
-\end{quote}
-
-We have added a sentence clarifying that our intention was simply to
-point out that the pattern is similar. We are obviously in a wholly
-different league.
+\begin{quote}\textbf{
+}\end{quote}
 
@@
 
-\begin{quote}
-This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
-architecture for low cost and ``potentially'' massive distributed evolutionary algorithms. The main novelty
-turns up from the fact that the architecture is developed in JavaScript.
-However, JavaScript is well-known because of its lack of capability to deal with high performance
-applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
-believe in the possibility of ``potentially'' massive computation as
-sketched by the authors.
-\end{quote}
-
-JavaScript is not, in fact, so slow and to prove its speed we have
-published a technical report comparing it with other compiled
-languages (such as Scala) and scripted languages
-\cite{2015arXiv151101088M}. In fact, node.js (used for the benchmarks)
-and the JavaScript virtual machine in Chrome (used in Android phones
-and Chrome and Chromium browsers) share the same JIT compiler, so
-their speeds are virtually the same. In general, as a conclusion of
-the report, we see that the real difference between scripted and
-compiled languages will vary with the language, but also with the
-particular implementation of the algorithm and the size of the problem
-or chromosome you are using. Even if the speed for a particular
-problem is two or five times that achieved for JavaScript, this
-disadvantage is more than offset by the availability of that free
-resource in abundance; that is, you will only need that many
-volunteers cooperating to get better performance than what you would
-get with C or Java. Even so, this NodIO architecture, as highlighted
-in the new version of the paper, is pool-based so local sources using
-high-speed devices such as GPUs or even local powerful CPUs can be
-added as non-volunteer users, since the heterogeneity and simple API
-of the system allows that by design.
-
-\begin{quote}
-The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
-section must be carefully checked and extended in order to fully validate the goodness of the proposed
-architecture. At least, authors should show the behavior of the architecture in a real-world problem
-demanding high computational performance.
-Finally, English must be double-checked in order to correct minor typos such as ``in the sense than[that]''
-(in page 2).
-\end{quote}
-
-We have extended the experiments section with a new one, based on a
-hard version of the Rastrigin's problem.
+\begin{quote}\textbf{
+}\end{quote}
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
including reviewers' questions

@@
 \begin{quote}\textbf{
+Please, correct the information about i7-4770 processor.
+Additionally, I would recommend to precise the concept of safety to "user's safety" (or "voluteer's safety") in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.
+
+I would also recommend to include at least a short information that the system utilizes HTML5 Web Workers, as from my point of view it is one of the most important aspects of the framework and to reorganize the subsections in section IV and V.
+
+I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.
+
 }\end{quote}

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
including reviewers' questions

@@
 \begin{quote}\textbf{
+1.- The section discussing the threats to validity of the experiments merely discuss the security issues of the platform but not the threats to the validity of the own experimentation performed in the paper.
+I strongly recommend to authors to follow the recommendations and guidelines described in [1] to address such comment.
+
+2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: “Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
+Additionally, the axis labels and legends in most of the figures are unreadable, please increase its font size.” I strongly recommend the authors to modify the figures to improve readability.
 }\end{quote}

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
including reviewers' questions

@@
 \begin{quote}\textbf{
+I’m not, however, fully convinced about that the new experiment of solving Rastrigin’s function be representative of a real world problem.
+The presented work, as is, has interesting merits, but the question of its usefulness remains uncertain if it would be used to solve a non-academic optimization problem.
+As the authors point out, reasonable performance is not ensured and their proposal could result in a pointless academic exercise.
 
+After reading the revised paper, an arising question is why the authors have used in Section IV a problem that is solved in around 1 second.
+An approach to simulate a hard problem could be to artificially increase the evaluation time (e.g, by adding an idle loop); this way, the problem would remain the same and the computing time could be adjusted to try to gain some speed-up with the proposed platform.
+
+Minor comments:
+- The paper structure should be enhanced: there are subsections IV.A and V.A but there are not
+IV.B nor V.B.
+- In the first paragraph of Section V is said: “! to perform the fifty runs”, but none is said before
+about those fifty runs.
+- Figures should be improved: Fig.2 has huge fonts (and the caption should include the time unit),
+while others Fig.3, Fig.4, etc include very small fonts.
+- The second version of the systems is named NodIO-W2 in Section V, but it is referred as to
+NodIO2 in the conclusions.
 }\end{quote}
@@
 I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.
-
 }\end{quote}
@@
 \begin{quote}\textbf{
-1.- The section discussing the threats to validity of the experiments merely discuss the security issues of the platform but not the threats to the validity of the own experimentation performed in the paper.
-I strongly recommend to authors to follow the recommendations and guidelines described in [1] to address such comment.
+The authors have thoroughly addressed all my previous concerns. I do not have additional comments or suggestions.
+}\end{quote}
 
-2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: “Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
-Additionally, the axis labels and legends in most of the figures are unreadable, please increase its font size.” I strongly recommend the authors to modify the figures to improve readability.
-}\end{quote}
+We would like to thank you the reviewer.
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
giving thanks to reviewer-4

@@
 Please, correct the information about i7-4770 processor.
-Additionally, I would recommend to precise the concept of safety to "user's safety" (or "voluteer's safety") in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.
+Additionally, I would recommend to precise the concept of safety to ''user's safety'' (or ''voluteer's safety'') in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.
 
@@
 
-We would like to thank you the reviewer.
+We appreciate the careful reading of the paper by Reviewer #4 and making valuable suggestions to improve the quality of this paper.
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
including Editors' questions

@@
 
+%-----------------------------------------------------------------------------
 \section{Response to Editors}
 
+\begin{quote}\textbf{
+1) Be careful with the use of acronyms.
+They must be defined the first time appearing in the text.
+Moreover, the use of too many acronyms jeopardizes readability of the manuscript.
+Please, keep only those acronyms which are really needed.
+
+2) NodIO-W2 is not presented in the abstract.
+
+3) Check the use of system vs framework.
+We recommend to use only one of these two words all along the text.
+
+4) Considering only one subsection inside a section makes harder the paper comprehensibility.
+For instance, in case of section IV, we recommend removing the caption related to subsection A and keeping everything in just one section.
+
+5) We suggest re-writing the title of section IV as "Baseline Experiments".
+
+6) We suggest re-writing the title of section V as "Enhancing NodIO with a web worker-based architecture"
+
+7) Section V should be split into two sections.
+The new section may be entitled as "Experimental trials" and it should start at "The experiments were run in the same way as before..."
+Moreover, it would be better to remove the caption of subsetion V.A.
+In case you think the section is too long then you may organize it into two subsections (A. Setup and B. Results)
+
+8) Minor typos.
+
+9) The following recent reference is closely related to your paper and it should be cited:
+J. Alcala-Fdez and Jose M. Alonso, A survey of fuzzy systems software: Taxonomy, current research trends and prospects, IEEE Transactions on Fuzzy Systems, 2015. Doi:10.1109/TFUZZ.2015.2426212
+}\end{quote}
+
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#1}
@@
 
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#2}
@@
 
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#3}
@@
 
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#4}
@@
 
+%-----------------------------------------------------------------------------
 \bibliographystyle{IEEEtran}

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
New response to reviewers was not compiling properly

--- 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
@@
 \title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
-for Low Cost Distributed Evolutionary Computation: Response to Reviewers\' comments}
+for Low Cost Distributed Evolutionary Computation: Response to Reviewers' comments}
 
@@
 
+%-----------------------------------------------------------------------------
 \section{Response to Editors}
 
+\begin{quote}
+\textbf{1) Be careful with the use of acronyms.
+They must be defined the first time appearing in the text.
+Moreover, the use of too many acronyms jeopardizes readability of the manuscript.
+Please, keep only those acronyms which are really needed.\\
+2) NodIO-W2 is not presented in the abstract.\\
+3) Check the use of system vs framework.
+We recommend to use only one of these two words all along the text.\\
+4) Considering only one subsection inside a section makes harder the paper comprehensibility.
+For instance, in case of section IV, we recommend removing the caption related to subsection A and keeping everything in just one section.\\
+5) We suggest re-writing the title of section IV as ``Baseline Experiments''.\\
+6) We suggest re-writing the title of section V as ``Enhancing NodIO with a web worker-based architecture''\\
+7) Section V should be split into two sections.
+The new section may be entitled as ``Experimental trials'' and it should start at ``The experiments were run in the same way as before...''
+Moreover, it would be better to remove the caption of subsetion V.A.
+In case you think the section is too long then you may organize it into two subsections (A. Setup and B. Results)\\
+8) Minor typos.\\
+9) The following recent reference is closely related to your paper and it should be cited:
+J. Alcala-Fdez and Jose M. Alonso, A survey of fuzzy systems software: Taxonomy, current research trends and prospects, IEEE Transactions on Fuzzy Systems, 2015. Doi:10.1109/TFUZZ.2015.2426212}
+\end{quote}
+
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#1}
 
-\begin{quote}\textbf{
-I’m not, however, fully convinced about that the new experiment of solving Rastrigin’s function be representative of a real world problem.
+\begin{quote}
+\textbf{I'm not, however, fully convinced about that the new experiment of solving Rastrigin's function be representative of a real world problem.
 The presented work, as is, has interesting merits, but the question of its usefulness remains uncertain if it would be used to solve a non-academic optimization problem.
-As the authors point out, reasonable performance is not ensured and their proposal could result in a pointless academic exercise.
-
+As the authors point out, reasonable performance is not ensured and their proposal could result in a pointless academic exercise.\\
 After reading the revised paper, an arising question is why the authors have used in Section IV a problem that is solved in around 1 second.
-An approach to simulate a hard problem could be to artificially increase the evaluation time (e.g, by adding an idle loop); this way, the problem would remain the same and the computing time could be adjusted to try to gain some speed-up with the proposed platform.
-
+An approach to simulate a hard problem could be to artificially increase the evaluation time (e.g, by adding an idle loop); this way, the problem would remain the same and the computing time could be adjusted to try to gain some speed-up with the proposed platform.\\
 Minor comments:
@@
 IV.B nor V.B.
-- In the first paragraph of Section V is said: “! to perform the fifty runs”, but none is said before
+- In the first paragraph of Section V is said: ``! to perform the fifty runs'', but none is said before
 about those fifty runs.
@@
 - The second version of the systems is named NodIO-W2 in Section V, but it is referred as to
-NodIO2 in the conclusions.
-}\end{quote}
+NodIO2 in the conclusions.}
+\end{quote}
 
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#2}
 
-\begin{quote}\textbf{
-1.- The section discussing the threats to validity of the experiments merely discuss the security issues of the platform but not the threats to the validity of the own experimentation performed in the paper.
-I strongly recommend to authors to follow the recommendations and guidelines described in [1] to address such comment.
+\begin{quote}
+\textbf{1.- The section discussing the threats to validity of the experiments merely discuss the security issues of the platform but not the threats to the validity of the own experimentation performed in the paper.
+I strongly recommend to authors to follow the recommendations and guidelines described in [1] to address such comment.\\
+2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: ``Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
+Additionally, the axis labels and legends in most of the figures are unreadable, please increase its font size.'' I strongly recommend the authors to modify the figures to improve readability.}
+\end{quote}
 
-2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: “Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
-Additionally, the axis labels and legends in most of the figures are unreadable, please increase its font size.” I strongly recommend the authors to modify the figures to improve readability.
-}\end{quote}
-
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#3}
 
-\begin{quote}\textbf{
-Please, correct the information about i7-4770 processor.
-Additionally, I would recommend to precise the concept of safety to "user's safety" (or "voluteer's safety") in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.
+\begin{quote}
+\textbf{Please, correct the information about i7-4770 processor.
+Additionally, I would recommend to precise the concept of safety to ''user's safety'' (or ''voluteer's safety'') in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.\\
+I would also recommend to include at least a short information that the system utilizes HTML5 Web Workers, as from my point of view it is one of the most important aspects of the framework and to reorganize the subsections in section IV and V.
+\\
+I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.}
+\end{quote}
 
-I would also recommend to include at least a short information that the system utilizes HTML5 Web Workers, as from my point of view it is one of the most important aspects of the framework and to reorganize the subsections in section IV and V.
-
-I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.
-}\end{quote}
-
+%-----------------------------------------------------------------------------
 \section{Response to Reviewer \#4}
@@
 
-We would like to thank you the reviewer.
+We appreciate the careful reading of the paper by Reviewer \#4 and making valuable suggestions to improve the quality of this paper.
 
+%-----------------------------------------------------------------------------
 \bibliographystyle{IEEEtran}

### 2015-12-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Modifying sections
Refs #81 and fixes suggestions by @unintendedbear

@@
   one). 
-Shades of blue (or gray in grayscale) indicate how many experiments included that many unique IPs,
+Shades of gray indicate how many experiments included that many unique IPs,
 so lighter shade for a column of dots indicates that a particular number
@@
 
+\subsection{Experimental results}
+
+After setting up baseline results for this particular problem, we will
+present now results the results obtained. Next we will present the
+details of the implementation and then the numeric results of the
+experiment.
+
 \subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}

### 2015-12-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reorganizing subsections
Also refs #81. Now it's clearer what has been done.

@@
 
-\subsection{Experimental trials}
+\subsection{First experiments}
 \label{sec:w2Exp}
@@
 function.
+
+\subsection{Tests with floating-point hard optimization problems}
 
@@
 
-\subsection{Experimental results}
-
 After setting up baseline results for this particular problem, we will
@@
 function, with the rest of the framework wrapped around
-automatically. All this, as the data used for this paper and the paper
+automatically. All this, as well as the data used for this paper and the paper
 itself, has been published with a free license in GitHub at

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
Finished response to eds refs #81

@@
 \begin{quote}
-\textbf{1) Be careful with the use of acronyms.
-They must be defined the first time appearing in the text.
-Moreover, the use of too many acronyms jeopardizes readability of the manuscript.
-Please, keep only those acronyms which are really needed.\\
+\textbf{1) Be careful with the use of acronyms. They must be defined
+  the first time appearing in the text. Moreover, the use of too many
+  acronyms jeopardizes readability of the manuscript. Please, keep
+  only those acronyms which are really needed.\\}
+\end{quote}
+
+All acronyms have been checked and defined in the first occurrence,
+some unnecessary ones (like JS for JavaScript) eliminated.
+
+\begin{quote}
+\textbf{
 2) NodIO-W2 is not presented in the abstract.\\
+}
+
+We have now made a small introduction to it.
+
+\begin{quote}
+\textbf{
 3) Check the use of system vs framework.
+
 We recommend to use only one of these two words all along the text.\\
-4) Considering only one subsection inside a section makes harder the paper comprehensibility.
-For instance, in case of section IV, we recommend removing the caption related to subsection A and keeping everything in just one section.\\
-5) We suggest re-writing the title of section IV as ``Baseline Experiments''.\\
-6) We suggest re-writing the title of section V as ``Enhancing NodIO with a web worker-based architecture''\\
-7) Section V should be split into two sections.
+}
+
+In general, we have opted for framework, since it is a client-server
+framework and it is more precise than {\em system}
+
+\begin{quote}
+\textbf{4) Considering only one subsection inside a section makes harder the paper comprehensibility.
+For instance, in case of section IV, we recommend removing the caption
+related to subsection A and keeping everything in just one section.\\}
+
+Done. Thanks for the suggestion.
+
+\begin{quote}
+\textbf{ 5) We suggest re-writing the title of section IV as ``Baseline Experiments''.\\
+6) We suggest re-writing the title of section V as ``Enhancing NodIO
+with a web worker-based architecture''\\}
+
+Done again and thanks.
+
+\begin{quote}
+\textbf{7) Section V should be split into two sections.
 The new section may be entitled as ``Experimental trials'' and it should start at ``The experiments were run in the same way as before...''
 Moreover, it would be better to remove the caption of subsetion V.A.
-In case you think the section is too long then you may organize it into two subsections (A. Setup and B. Results)\\
-8) Minor typos.\\
-9) The following recent reference is closely related to your paper and it should be cited:
+In case you think the section is too long then you may organize it
+into two subsections (A. Setup and B. Results)\\}
+
+If we understood it correctly, you thought that the section was too
+long for a single {\em segment} and you wanted two subsections (not
+two different {\em entities} of the same kind. This is what we have
+done: introduced two subsections, one focused in the experiment we
+presented on the first version, another on the experiment that
+has been introduced in this version and entitled {\bf Tests with a
+  floating-point hard optimization problem}. Thanks for the
+suggestion, it is now much clearer.
+
+\begin{quote}
+\textbf{8) Minor typos.\\}
+
+Fixed
+
+\begin{quote}
+\textbf{9) The following recent reference is closely related to your paper and it should be cited:
 J. Alcala-Fdez and Jose M. Alonso, A survey of fuzzy systems software: Taxonomy, current research trends and prospects, IEEE Transactions on Fuzzy Systems, 2015. Doi:10.1109/TFUZZ.2015.2426212}
 \end{quote}
+
+Added.
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
Continuing with rev1, refs #81

@@
 As the authors point out, reasonable performance is not ensured and their proposal could result in a pointless academic exercise.\\
-After reading the revised paper, an arising question is why the authors have used in Section IV a problem that is solved in around 1 second.
-An approach to simulate a hard problem could be to artificially increase the evaluation time (e.g, by adding an idle loop); this way, the problem would remain the same and the computing time could be adjusted to try to gain some speed-up with the proposed platform.\\
-Minor comments:
+After reading the revised paper, an arising question is why the
+authors have used in Section IV a problem that is solved in around 1
+second. An approach to simulate a hard problem could be to
+artificially increase the evaluation time (e.g, by adding an idle
+loop); this way, the problem would remain the same and the computing
+time could be adjusted to try to gain some speed-up with the proposed
+platform.\\}
+
+This is an useful suggestion. We will try to add that kind of
+experiment in the future. In fact, some of our previous papers have
+used a similar technique successfully.
+
+\begin{quote}
+\textbf{Minor comments:
 - The paper structure should be enhanced: there are subsections IV.A and V.A but there are not
-IV.B nor V.B.
-- In the first paragraph of Section V is said: ``! to perform the fifty runs'', but none is said before
-about those fifty runs.
-- Figures should be improved: Fig.2 has huge fonts (and the caption should include the time unit),
-while others Fig.3, Fig.4, etc include very small fonts.
-- The second version of the systems is named NodIO-W2 in Section V, but it is referred as to
+IV.B nor V.B.}
+
+As a response to editor's requests, this has been done.
+
+\begin{quote}
+\textbf{- In the first paragraph of Section V is said: ``! to perform the fifty runs'', but none is said before
+about those fifty runs.}
+
+It is mentioned now at the beginning of the paragraph
+
+\begin{quote}
+\textbf{- Figures should be improved: Fig.2 has huge fonts (and the caption should include the time unit),
+while others Fig.3, Fig.4, etc include very small fonts.}
+
+Fonts have been made more readable and uniform, and axis labeled.
+
+\begin{quote}
+\textbf{- The second version of the systems is named NodIO-W2 in Section V, but it is referred as to
 NodIO2 in the conclusions.}
 \end{quote}
+
+It is called the same way now. Thanks a lot for your suggestions, and
+we expect these changes satisfactory.
 

### 2015-12-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Finished response to eds refs #81
Continuing with rev1, refs #81

@@
 \centering
-\includegraphics[width=12cm]{baseline-times.png}
+\includegraphics[width=10cm]{baseline-times.png}
 \caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
@@
 took around a minute, on average, that is $68.9694$ seconds, to
-perform the fifty runs. In this experiment only 33, that is, 66\% were
+perform these runs. In this experiment only 33, that is, 66\% were
 successful. The experiments were repeated for population $p=1024$,
@@
 
-\subsection{Tests with floating-point hard optimization problems}
+\subsection{Tests with a floating-point hard optimization problem}
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
Fixes errors refs #81

@@
 \textbf{
-2) NodIO-W2 is not presented in the abstract.\\
-}
+2) NodIO-W2 is not presented in the abstract.}
+\end{quote}
 
@@
 \begin{quote}
-\textbf{
-3) Check the use of system vs framework.
-
-We recommend to use only one of these two words all along the text.\\
-}
+\textbf{3) Check the use of system vs framework.
+We recommend to use only one of these two words all along the text.}
+\end{quote}
 
@@
 For instance, in case of section IV, we recommend removing the caption
-related to subsection A and keeping everything in just one section.\\}
+related to subsection A and keeping everything in just one section.}
+\end{quote}
 
@@
 with a web worker-based architecture''\\}
+\end{quote}
 
@@
 into two subsections (A. Setup and B. Results)\\}
+\end{quote}
 
@@
 \textbf{8) Minor typos.\\}
+\end{quote}
 
@@
 time could be adjusted to try to gain some speed-up with the proposed
-platform.\\}
+platform.\\}\end{quote}
 
@@
 - The paper structure should be enhanced: there are subsections IV.A and V.A but there are not
-IV.B nor V.B.}
+IV.B nor V.B.}\end{quote}
 
@@
 \textbf{- In the first paragraph of Section V is said: ``! to perform the fifty runs'', but none is said before
-about those fifty runs.}
+about those fifty runs.}\end{quote}
 
@@
 \textbf{- Figures should be improved: Fig.2 has huge fonts (and the caption should include the time unit),
-while others Fig.3, Fig.4, etc include very small fonts.}
+while others Fig.3, Fig.4, etc include very small fonts.}\end{quote}
 
@@
 \begin{quote}
-\textbf{1.- The section discussing the threats to validity of the experiments merely discuss the security issues of the platform but not the threats to the validity of the own experimentation performed in the paper.
-I strongly recommend to authors to follow the recommendations and guidelines described in [1] to address such comment.\\
-2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: ``Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
+\textbf{1.- The section discussing the threats to validity of the
+  experiments merely discuss the security issues of the platform but
+  not the threats to the validity of the own experimentation performed
+  in the paper. I strongly recommend to authors to follow the
+  recommendations and guidelines described in [1] to address such
+  comment.\\}
+\end{quote}
+
+A comment to the threats to validity has been now included in the
+paragraph next to the conclusion.
+
+\begin{quote}
+\textbf{2.- The figures 3,4 and 5 of the paper have not been modified according to our original comments: ``Figures should be readable even in gray-scale, so do not use red and blue to denote different elements, use a different shade or texture.
 Additionally, the axis labels and legends in most of the figures are unreadable, please increase its font size.'' I strongly recommend the authors to modify the figures to improve readability.}
 \end{quote}
+
+Previously, we have checked it in grayscale and thought it was OK. We
+have done now as instructed.
+
+We would also like to thank this reviewers for his work and helpful
+comments.
 

### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
Done with rev3 refs #81

@@
 \begin{quote}
-\textbf{Please, correct the information about i7-4770 processor.
-Additionally, I would recommend to precise the concept of safety to ''user's safety'' (or ''voluteer's safety'') in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.\\
-I would also recommend to include at least a short information that the system utilizes HTML5 Web Workers, as from my point of view it is one of the most important aspects of the framework and to reorganize the subsections in section IV and V.
-\\
-I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.}
+\textbf{Please, correct the information about i7-4770
+  processor. Additionally, I would recommend to precise the concept of
+  safety to ''user's safety'' (or ''voluteer's safety'') in the
+  characteristic of the framework (page 5), as you still do not
+  address the safety of the system as a whole, at least not at this
+  point.}
+\begin{quote}
+
+This has been done, as indicated in the response to other reviewers.
+
+\begin{quote}
+\textbf{I would also recommend to include at least a short information
+  that the system utilizes HTML5 Web Workers, as from my point of view
+  it is one of the most important aspects of the framework and to
+  reorganize the subsections in section IV and V.}
 \end{quote}
+
+Done.
+
+\begin{quote}
+\textbf{I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.}
+\end{quote}
+
+Done too. Thanks a lot for your feedback.
 

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Separating in two parts
In a different branch, to keep history.

--- 2015-12-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
-\documentclass[journal,onecolumn]{IEEEtran}
+\documentclass{article}
 
@@
 
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
   \graphicspath{{../img/}}
-\else
-  \usepackage[dvips]{graphicx}
-  \graphicspath{{../img/}}
-\fi
 
@@
 
-\title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
-for Low Cost Distributed Evolutionary Computation }
-% The word massive in the title is giving us problems.
-% Why not change it to something about sharing resources?.
-% Also lets not forget about volunteers, they can be a resource too
-% and they cannot be replaced by any super computer
+\title{}
 
-%  I propose something like:
-% A Web Based Architecture to Harness Volunteer Computational Resources 
-% for Low Cost Distributed Evolutionary Computation
-
-% Some keywords missing: Cloud, Asynchronous, Pool-Based, Javascript-Based
-% Open Source software.
-% Lets not forget this issue is about software.
-
-\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
+\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro
+  A. Castillo$^1$,\\ 
+Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
 \thanks{Manuscript submitted for review on \today.}%
@@
 \begin{abstract}
-Volunteer computing is extensively used in science and industry, but
-it still has a long way to go in order to become a mainstream
-technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
-The lack of general purpose and open
-source tools with built-in analytics could be one of the reasons of
-why this is so, as well as the lack of proofs of concept.
-% (Paloma) The following sentence is way too long, and the "which" after "Javascript" is confusing, so one does not know if the which refers to the framework or Javascript by reading the whole sentence.
-% I propose: "That is why in this paper we present NodIO, a framework used to create client-server architectures for distributed evolutionary algorithms, entirely written in JavaScript.
-Our framework needs minimal server support to work and is able to support persistent, asynchronous, distributed evolutionary algorithms running in the browser, as well as any other client consuming its REST application programming interface."
-That is why
-in this paper we present {\sf NodIO}, a framework used to create
-client-server architectures for distributed evolutionary algorithms,
-entirely written in JavaScript, which needs minimal server support to
-work and that is able to support persistent, asynchronous, distributed
-evolutionary algorithms running in the browser, as well as any other
-client consuming its REST application programming interface. An
-enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
-also presented; this version leverages the HTML5 Web Workers
-specification to execute several clients per browser and do it in the
-background, so that the the number of generations per user could be
-improved. The architecture, along with some initial results on the
-expected performance and scaling behavior when the number of users is
-increased is introduced. % (Paloma) I've changed this "presented" for a synonym
-We also use the algorithm data to find out
-what kind of patterns the user behavior follows by examining the
-resources they devote to our experiment, and try to fit them in % (Paloma) Changed. Phrasal verb was wrong.
- some
-statistical distribution to create the bases of a model of user
-behavior in volunteer computing experiments of this kind that can be
-used by researchers to predict their performance. %(Paloma) Changed to "their" because of "experiments"
-In general, our
-experiments prove, first, that {\sf NodIO} is a flexible and
-low-overhead platform to perform distributed computation experiments
-that include volunteers, and then that user behavior follows some
-patterns, and, thus, are amenable to a model.
+
 \end{abstract}
@@
 
-\begin{IEEEkeywords}
+{\bf Keywords}
 Evolutionary algorithms, evolutionary computation, distributed computing, internet computing,
 social networks, volunteer computing, metacomputing, performance evaluation.
-\end{IEEEkeywords}
 
@@
 
-\IEEEPARstart{N}{owadays} there is a big and increasing number of
-distributed evolutionary computation software frameworks available on
-the market, usually as open source frameworks \cite{Parejo12Survey}. 
-Most of these frameworks have 
-two features: they are {\em desktop} applications, that is, they must
-be compiled or run directly on the operating system layer and, second,
-they assume that the nodes that are going to run the system are under
-the control of developers or they at least have access to them. In
-fact, these features are commonly found in most soft computing
-frameworks \cite{7094263}.
-
-There is, then, space for frameworks that do not make any of these
-assumptions, such as the one presented in this paper, which will be
- named {\sf NodIO}. {\sf NodIO} is basically
-% (Paloma) Is the "first" word really needed? 
-% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
-a two tier client/server 
-application, in which clients can run on a browser. 
-Therefore, they can be embedded in any web-enabled device, including many over which 
-developers have no control.
-
-Some time ago, our group faced the problem of diminishing funds for
-buying new hardware. This was aggravated by the increasing maintenance costs and
-extended downtime resulting from the continuous failures of existing
-clusters.
-Considering this, we leveraged our experience in the design of web
-applications since their inception, including JavaScript, and other
-volunteer and
-unconventional distributed evolutionary computing systems to 
-design and release a new free framework that would allow anyone to
-create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
-servers, and browsers as clients.
-% (Paloma) I think you should rewrite this whole sentence.
-% For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided... 
-% (Mario) We leverage our experience in blah blah to design
-% (Paloma) It may be... guess I was thinking in the noun, not in the
-% transitive verb
-% Please check if there's a problem here too.
-
-%PABLO: He añadido el siguiente párrafo para el issue #31
-There exist other volunteer computing frameworks written in compiled languages,
-such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript \cite{2015arXiv151101088M}. However, one of the advantages of our architecture is that
-clients do not need to install anything in their machines, they just
-need to open a web page  
-% Somewhat redundant 
-% Javascript is the only language that can be executed in this way without any installation, 
-% also removed the default, redundant with next sentence
-in a compatible browser, such as Chrome, Safari or Firefox, which are
-available in almost every desktop or mobile device. Other advantage is
-that the source code can be directly examined by volunteers, because it is not
-compiled as other options such as Java Applets, ActionScript or
-ActiveX.
-Also, as we will discuss in this paper, it can provide some extra
-features such as enhanced security or operating system independence.
-
-%PABLO: hay que enlazar con el siguiente párrafo
-%Desktop clients have the same issues as any 
-%volunteer computing system, including our % (Paloma) their?
-% own: these issues are
-%related to security as well as
-%safety, which introduce
-
-When developing volunteer based frameworks, designers must
-also consider the whole {\em social} aspect in the design, with
-issues related to security, trust, and privacy among others. 
-The computing system becomes then a {\em techno-social system} \cite{vespignani2009predicting}.  
-% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
-%a {\em social} aspect to the design of a
-%computing system that makes it a {\em techno-social system}
-In this paper, our intention is to
-present the {\sf NodIO} framework and analyze it from two different
-angles: the purely technical angle, including the decisions that went
-into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page running the experiment
-determine its performance and how it can be modeled and, eventually,
-predicted.
-
-The rest of the paper is organized as follows: Next we present the
-state of the art (Section \ref{sec:soa}) in web-based distributed
-computational systems along with attempts to predict and model its behavior. 
-Then, a detailed description of the proposed framework is presented
-in Section \ref{sec:description}, followed by the baseline experiments
-in Section \ref{sec:experiments}, and experiments using a new client
-architecture that uses web workers in Section \ref{sec:w2}.
-Finally, conclusions and future lines of work are presented in Section
-\ref{sec:conclusion}.
+The rest of the paper is organized as follows:
 
@@
 
-The web used as a resource for distributed computing has a
-long history, and almost since the beginning it was used for
-evolutionary computation even before JavaScript was
-introduced as a browser-based scripting language in 1997. Before that
-time other technologies,
- including Flash animations and VBScript, ActiveX or Java applets were
-used; for instance, Java was pointed out in \cite{soares1998get} as a
-``language for the
-internet'', providing some advantages such as multi-architecture compatibility or
-security mechanisms, and was proposed in \cite{chandy1996world} as the
-basis for the creation of a world-wide peer to peer system using compiled Java
-programs. The ATLAS system
-\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
-with another language called Cilk; all these papers indicated that
-the introduction of an architecture-independent system such as the
-Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers. This paper % (Paloma) No really need for that "also", plus it complicates reading
-establishes the desirable features of
-any {\em global computing} infrastructure, that can also be applied to
-distributed evolutionary computation experiments that rely on
-volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
-reasonable performance. But hierarchy, this is, the capability of
-preferentially using resources in the same organization, is a feature
-that could be either disregarded or included within adaptivity, which
-leaves us with eight desirable properties. 
-% (Paloma) Re-write. For example, "But is hierarchy, this is, the
-% capability of preferentially using resources in the same
-% organization, the one that could be either disregarded or included
-% in adaptivity, which leaves us with eight desirable properties." 
-% (Paloma) And I re-wrote what I proposed yesterday.
-% thanks :-) -JJ
-
-The fulfillment of these properties could be achieved by using any of the
-languages and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first. 
-% (Paloma) These should be separate sentences
-However, embedding a
-distributed computation client in the browser has the advantage of
-providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed by using desktop applications. 
-Baratloo et al. tried in 1996 to create a framework with
-the aforementioned qualities called {\em Charlotte}
-\cite{baratloo1996charlotte}. This framework, which implemented a
-distributed shared memory, was inspired by a previous work
-on parallel computing libraries PVM and MPI and solved most of the issues by
-using the Java virtual machine embedded in the browser,
-which provided
-security to the user, uniformity of computing environment
-independently of the operating system, and access to a common data store in the
-server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
-and as it runs compiled bytecode it also guaranteed a certain level of
-performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
-Other issues such as scheduling, which are related to adaptive 
-parallelism, were solved on the server. This
-paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that,
-even in the early stages of the web, the possibility of
-% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
-browser-based distributed computing was already researched.
-
-Other systems, like JET, proposed by Soares et
-al. \cite{soares1998get}, also used Java. JET supports
-the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
-since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
-the collection of algorithm analytics by the
-experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to JET, such as {\em SuperWeb}
-\cite{alexandrov1997superweb} are roughly a translation of classical
-distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. SuperWeb is focused,
-% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
-however, in solving the
-issues involved in browser-based computing and also in making a system
-that is able to work across different types of computers, mentioning
-SPARCstations and PCs with Windows-95. Interestingly enough, this
-paper incorporates a discussion on the economic model, paving the way
-for the for-profit distributed computing systems that arose later on. 
-On the other hand, the potential for volunteer computing using
-browsers was realized
-later on \cite{sarmenta-bayanihan} as well as the potential of
-user misconduct when having access to the source code \cite{sarmenta-sabotagetolerance}. 
-However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
-less widespread.
-
-The main problem with Java and all the other browser-embedded
-technologies is that they are not {\em universal} in the
-sense that an extra component, namely, the Java virtual machine or another
-virtual environment has to be installed in the browser.
-Also these virtual environments have become targets of malicious attacks; 
-for instance numerous vulnerabilities have been discovered in 
-the proprietary Adobe Flash runtime environment and
-video player \cite{ford2009analyzing,watanabe2010new}. The need for
-these additional environments has been mitigated with the
-adoption of the HTML5 standard \cite{anthes2012html5}.
-JavaScript \cite{flanagan2006javascript}, however, has become a set of standards
-\cite{ECMA-262} that include the language and its components. Since
-early on, it was adopted by the industry and also by scientists,
-who used them for creating an evolutionary algorithm (EA from now on) on
-the browser as early as 1998 \cite{jj-ppsn98}.
-
-JavaScript can be used either for unwitting
-\cite{unwitting-ec} or volunteer 
-\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
-evolutionary computation and it has been used ever since by several
-authors, including more recent efforts
-\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
-that even
-used the client's GPU \cite{duda2013gpu} or create an object-oriented
-framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
-used Java \cite{chong:1999:jDGPi} and others have gone away from the
-server-based paradigm to embrace peer to peer systems
-\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
-platforms avoid single points of failure (the server) but, since they
-need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower. % (Paloma) Do yo mean... higher? A low threshold means that is easy to happen something.
-
-Recent works propose the use of cloud computing services for the components of
-client-server distributed EAs. Cloud resources
-can reduce operational costs by outsourcing both hardware and software maintenance
-to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
-is available in most laboratories, allowing researchers to
-scale algorithms at reasonable costs. Researchers have proposed the use of
-public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
-and DropBox~\cite{mericloud}. The use of JavaScript to implement
-server side components
-gives software developers certain advantages in the cloud
-paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
-developers can use the same software tools for both client and server
-side development.
-
-JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
-According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
-metaheuristics optimization is that it allows an improvement in communicability,
-reproductibility, interoperability, automated assembly, knowledge engineering,
-and efficiency.
-
-In the work presented in this paper, JavaScript has been used throughout the full
-stack, which has the advantage of using the self-same lines of code for
-writing an evolutionary algorithm for the
-desktop or command line and browser clients, both based in the EC
-library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
-when writing code and leverages 
-the expertise acquired
-% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
- in the library itself.
-
-On the other hand, we are also interested in measuring the performance
-of volunteer computing systems, an area in which there have been
-relatively few efforts.
-There were some initial attempts to avoid the differences in performance
-that could be obtained from volunteers  by making
-the algorithm adaptive to the kind of resources allotted to it
-\cite{milani2004online}. This is actually not such a big problem in
-algorithms such as the EA that can easily be
-parallelized via population splitting or by farming out evaluations to all
-the nodes available. Lately, several approaches have focused on the
-fault-tolerance of volunteer algorithms
-\cite{gonzalez2010characterizing} which it can, of course, be studied in
-the more general context of distributed computing 
-\cite{nogueras2015studying} or be included in a more general study of the
-performance of the EA itself
-\cite{DBLP:journals/gpem/LaredoBGVAGF14}.
-
-On the other hand, measuring the performance of the resulting metacomputer
-involves understanding the dynamics of this kind of systems. Initial
-work was done for peer to peer systems by Stutzbach et
-al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
-%But the raw material of           
-%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
-But some of the essential metrics in volunteer computing like the
-number of users or the time spent by every one in the
-computation in browser-based volunteer computing experiments, have
-only been studied in a limited way in 
-\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
-run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining} found out that the Weibull, log-normal, and
-Gamma distribution 
-modeled quite well the availability of resources in several clusters
-of that framework; the shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas:
-there are many users that give a small amount of cycles, while there
-are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}. Although SETI@home works
-at a larger scale in both quantity of volunteers and the scope of projects, 
-our initial tests indicate that
-this kind of distribution might be the most appropriate to describe
-user participation in several different metrics. However, in this
-paper we will perform different experiments in different setups and
-will try to come up with a more precise model. Next we describe the
-framework itself.
-
-As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
-computational resources that are as dissimilar as smartphones and
-powerful laptops 
-or workstation computers in a research center. 
-%The methodology used
-% (Paloma) I would say "the used methodology"
-The algorithms used, as well as the methodology 
-for gathering resources will be described next, 
-together with the results obtained in this initial setup.
 %---------------------------------------------------------------
@@
   %(Mario) Scalability of the communication system in particular?
-\item Since the algorithm runs in the browser, {\em safety} is
+\item Since the algorithm runs in the browser, user's {\em safety} is
   achieved through its sandbox mechanisms. The user is thus assured
@@
 %---------------------------------------------------------------
-\section{Baseline Experiments} %Cambié el nombre
+\section{Baseline Experiments} 
 \label{sec:experiments}
@@
 the baseline is that if they eventually take longer than a basic
-desktop, their interest will be purely academic. We will try and
-prove next that that target performance can be achieved by carrying
-out several experiments in different conditions. Second, they also act
-as a basic performance test for the server, and also prove that the
-basic server can be used for many kinds of different setups, as long
-as they follow the REST protocol to deposit and obtain individuals
-from the pool. It is obvious that, in this case, the server is not
-really needed, since there is a single client and no interaction
-between them, but there is a cost in sending and obtaining chromosomes
-from the server so it is included for the sake of fairer comparison.
-
-Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which these runs could be performed. That way, and as required, the hardware
-and cloud cost for this experiment were zero. Experiments were
-announced through a post in Twitter and other social networks, and
-results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
-purpose of this paper, we repeated the announcement several times
-through the month of April and then by the beginning of August. All
-in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:runs}. In general, every experiment took several
-days. No particular care was taken about the time of the announcement
-or the particular wording. Every {\em experiment} consisted in running
-until the solution of the 40-trap problem was found. When the correct
-solution was sent to the server, the counter was updated and the pool
-of solutions reset to the void set. There was no special intention to wait
-until all clients had finished, thus it might happen that, in fact,
-the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible.
-%
-\begin{table*}
-\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
-\begin{center}
-\begin{tabular}{l|rr}
-\hline
-Date & \# experiments & Different IPs \\
-\hline
-April 4th 4/4 & 57 & 191 \\
-April 24th 4/24 &  231 & 559 \\
-July 31th 7/31 & 97 & 179 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{table*}
-\caption{Summary of time per run, number of IPs and number of {\tt
-    PUT}s per IP in the initial runs. \label{tab:summary:os}}
-\begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
-4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
-7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-The table shows that every run included more than 50 experiments. The
-number of different IPs intervening in them varied from more than one
-hundred to more than five hundred in the second experiment.
-
-A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of IPs
-intervening in each experiment,  median time needed
-to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
-then two columns showing the percentage of experiments that took less
-than the two baseline experiments shown in the introduction to this
-section. The first striking result is that in all cases, 50\% of the
-experiments involved 5 or less IPs. This is consistent with results
-previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
-the usual number of IPs that participated in an experiment. The
-maximum number of different IPs for each experiment is also in the
-same range and of the order of 10, which is also consistent with
-prior work. The median time has a big range of variation, but 50\% of
-the time takes less than several minutes, from around 4 minutes in the
-best case to roughly 2/3 of an hour in the worst case. That figure is
-not competitive, {\em a priori}, with the baseline experiment, that is
-why it is interesting to look at the columns which measure
-the percentage of experiments in which the solution was found in less
-time than the baseline. It is similar in all three cases, with around
-20\% on average for the first baseline experiment, roughly 3\% in the
-second case. The comparison is not totally fair: this experiment takes
-place in a browser, and in fact most of the time is devoted to
-presenting the charts; the server is also not the same since it is run in a
-PaaS instead of a high-performance desktop computer, and it is also a
-remote server, not a local server as in the baseline experiment. But
-even in this case, we were looking for creating a framework that
-allowed us to run evolutionary computation experiments {\em faster} than
-in our very own computer. And the results indicate that although it will
-happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
-experiment you choose, it will not happen always or even most of
-the time. One of the problems is shown in the last column, which
-indicates the statistical correlation between the number of IPs
-participating in one experiment and the next. In all cases,
-correlation is so low as to affirm that there is almost no
-relationship between them, that is, computing nodes are not {\em
-  staying} after the solution has been found. This is a feature of
-this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. Thus, the user can decide if to 
-continue sharing resources or not. % (Paloma) Feature or disadvantage?
-However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
-unnecessary upper limit on the time, or number of experiments, that a user
-will voluntarily perform.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
-\caption{Duration of experiments vs. number of different IPs (nodes)
-  participating in it, with averages and standard deviation shown as
-  red dots; in the case there is a single red dot, there was a single
-  experiment in which many computers participated (for instance, 16
-  computers in the experiment in the far left or 29 in the middle
-  one). 
-Shades of gray indicate how many experiments included that many unique IPs,
-so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequency. 
-From left to right, experiments 4/4, 4/24 and 7/31.}
-\label{fig:duration}
-\end{figure}
-% ../data/time-vs-IPs-openshift.png 
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
-  logarithmic scale. Dot color (or gray value) is related to the number of IPs
-  participating in the experiment. From left to right, experiments
-  4/4, 4/24 and 7/31.} 
-\label{fig:zipf:os}
-% computed via ../data/time-vs-IPs-openshift.png 
-\end{figure}
-%
-We will have to analyze experimental data a bit further to find out why
-this happens and also if there are some patterns in the three sets of
-experiments. An interesting question to ask, for instance, is if
-by adding more computers makes the experiment take less. In fact, as
-shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
-not seem to contribute to decreasing the time needed to finish the
-experiment. However, the cause-effect relationship is not clear at
-all. It might be the opposite: since experiments take longer to finish
-and might in fact be abandoned with no one contributing for some time,
-that increases the probability of someone new joining them. In fact,
-with experiments taking a few seconds and due to the way the
-experiments are announced, it is quite difficult that several
-volunteers join in in such a short period of time, even more if we take
-into account that volunteers are not {\em carried over} from previous
-experiments. This implies that it would be convenient to use a problem
-of a bigger size to check this hypothesis as we have done in the next
-experiment. 
-%  however, at this point we
-% have not found this convenient since there are several other issues
-% that have to be solved, as it will be shown next. %Creo que este párrafo
-%saldría sobrando ahora que hicmios un experimento más grande
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-4.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-24.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
-\caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
-  Extreme Value distribution (in a lighter shade of gray). From left to right, experiments
-  4/4, 4/24 and 7/31.} 
-% Plotted with ../data/plot-zipf-openshift.R
-\label{fig:puts:os}
-\end{figure}
-%
-It is also interesting to check the distribution of the experiment
-duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
-a Zipf's law, with similar distribution along all three runs. The 4/24
-run is the most complete and shows a S-shape, which implies an
-accumulation of experiments taking similar time and around 100
-seconds. The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, of the order of
-10 seconds, and which appears in all three graphs. As it can be seen,
-it sharply drops implying there are 
-just a few of them, and with diminishing probability as time
-decreases. This exponential distribution also appears in the
-distribution of HTTP {\tt PUT}s, equivalent to the number of
-generations divided by 100, contributed by every user, which is shown
-in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
-so we have fitted it to the Generalized Extreme Value distribution,
-with the resulting parameters shown in Table \ref{tab:puts:os}.
-%
-\begin{table*}
-\caption{Summary of fit to Generalized Extreme Value distribution of
-  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
-\begin{center}
-\begin{tabular}{l|ccc}
-\hline
-Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
-0.1377 \\
-4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\
-7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-This distribution was originally proposed to fit extreme values
-\cite{resnick2013extreme} and contains, as a special case, the inverse
-Weibull distribution which was fitted to volunteer computing
-frameworks such as SETI@home \cite{javadi2009mining}. We obviously do
-not pretend to compare our framework in scale or complexity with it, but
-to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which also appears in our framework. This
-distribution is governed by three parameters, the usual location $\mu$
-which is related to where it has its {\em center} and a scale $\sigma$,
-related to the size, but also a third shape $\xi$ parameter that is
-related to its skewness, that is, how skewed it is around the central
-location. Positive parameters indicate that the distribution {\em
-  leans} towards the origin, and negative ones towards the other extreme
-value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
-greater than one and between one and 1.4, which indicates that the
-three experiments share this origin-leaning pattern, with many users
-donating a few cycles and just a few donating extreme values. Random
-distributions with these parameters have been plotted in red in
-Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
-predictive value of these fits, however, is limited, over all taking
-into account the low correlation between successive events shown
-above.
-
-At any rate, this also shows that a convenient way of increasing the
-computing power would be increasing this minimum amount per user. This
-issue led us to create the next version of the framework, which is
-presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
+desktop, their interest will be purely academic.
 
@@
 %
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
-\caption{Time employed in every experiment vs. number of IPs (in
-  abscissas). The black dot and line show the average and standard
-  deviation, the shades of gray the number of experiments with the same
-  number of IPs. } 
-\label{fig:ipstime:w2}
-% ../data/time-vs-IPs.R 
-\end{figure}
-
-\subsection{First experiments}
-\label{sec:w2Exp}
-
-The experiments were run in the same way as before, using several
-Twitter announcements throughout several days by the end of July
-2015. Eventually, more than one thousand experiments were completed in
-a matter of days. A summary of results is shown in Table
-\ref{tab:summary:ww}, comparing it with the aggregate of the results
-obtained with the initial version of {\sf NodIO}. These results are
-remarkably different, being similar only in the median and maximum
-number of IPs, although in this case it would be combination
-IP-worker, as we considered every worker as a unique IP. The median
-number of seconds is almost 80\% better than the previous version, and
-the median number of PUTs is three times better; this causes that most
-of experiments take less time than one of the baselines, and one third
-less than the second baseline. Thus, the conclusion in this case is
-that we can obtain, through volunteer computation, a result that most 
-of the times is better than we would by using a similar, non-parallel,
-desktop setup, which makes this system suitable for massively
-distributed evolutionary algorithms.
-%
-\begin{table*}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
-0.04746 \\
-Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-
-However, it is interesting to note why this is so, and the first hint
-is the inter-experiment correlation between the number of IPs in
-successive experiments. While before it was an unremarkable 10\%, it is
-now more than 43\%, making the number of IPs in contiguous experiments
-highly correlated. The main reason for this is the setup in {\sf
-  NodIO-W$^2$}, which restarts an island after a solution has been
-found and also that it keeps running even if the tab is not on the
-foreground. This means that volunteers can leave the experiment
-running for as long as they want and they will be contributing 
-experiment after experiment, unlike the previous version, where they
-stopped contributing after one solution was found. This is also
-reflected in the number of PUTs, which is the number of generations
-divided by 100; more than 50\% of the volunteers run the experiment for more than
-4000 generations. All things considered, this means than there are
-many more computers {\em simultaneously} running, leading to this
-almost 5-fold increase in running time.
-
-We are also interested in measuring the scaling properties in this
-model, that is, the relationship between the number of IPs
-participating in an experiment and the time it takes to complete
-it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
-experiment in terms of time (in milliseconds) vs. the number of workers participating
-in it. Although there is not a clear trend, the graph seems to say
-that the time needed for finding the solution in this evolutionary
-algorithm does not depend on the number of workers participating in
-it. This does not mean that it is independent on the number of {\em
-  simultaneous} users, which is probably the case although it is more
-difficult to measure. There seems to be a trend towards a decreasing
-standard deviation in the time, with more volunteers adding both {\em
-  robustness} to an experiment, and certainty in the time it is taking. 
- However, more experiments would be needed to check this hypothesis.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in light color, right) 
-and Weibull (in lighter shade, left) for the {\sf NodIO-W$^2$} model.}  
-\label{fig:gev:w2}
-% File ../data/ips-puts.R
-\end{figure}
-%
-As we did in the previous subsection, we have also fitted the number
-of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
-in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
-in the previous experiment, Table \ref{tab:puts:os}, we find higher values
-for the location $\mu$ and scale $\sigma$ parameters, accounting for
-the higher number of contributions per volunteer that has been
-previously observed. However, the shape $\xi$ parameter is
-substantially similar, with a {\em bell} shape leaning towards the
-origin, indicating a pattern of a few users with many contributions
-and many with few contributions. If we plot the model and the data
-side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
-is not so tight as in the previous case, with the model overestimating
-the number of individuals with a high number contributions. This is
-probably due to the fact that in the previous case, the user needed to
-voluntarily reload the page to contribute the most generations, with a
-{\em phase change} between the people that did so and the people that
-just ran the experiment once until completion. In this case, however, the
-user just needs to let it run, without that phase change and thus we obtain a
-more {\em egalitarian} distribution of contributions, which rather
-corresponds to a Weibull function, as shown in Figure
-\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
-distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
-equal to 0.69, a value that is remarkably similar to the 0.5 value
-found for time devoted to games in Chambers et
-al. \cite{chambers2005measurement}. It is important to note that, in
-this case, the number of PUTs made per client does not correspond to a
-uniform amount of computation, since 100 generations with a random
-population belong, in each case, to different number of
-operations. In other words, the same number of PUTs will correspond to
-different number of operations in each case. However, since clients
-are also different and take a different time in each case, it is
-difficult to ascertain how this might have an influence in the
-statistical distribution.
-
-In general, this second version of the {\sf NodIO} framework and the
-single experiment performed prove that it can be the foundation for a distributed
-high-server-performance evolutionary computation platform,
-providing reasonable algorithmic performance and being, in general,
-easy to use with a straightforward modification of the fitness
-function.
 
@@
 
-After setting up baseline results for this particular problem, we will
-present now results the results obtained. Next we will present the
-details of the implementation and then the numeric results of the
-experiment.
-
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}
-
-This experiment was deployed in a virtual private server with 512MB of memory
-running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
-experiments where the optimum was expected to be found many times, 
-this problem was expected to run constantly for many days finding only
-sub-optimal approximations. In order to achieve this, the only change needed 
-in the framework was to limit the number of chromosomes kept in the pool. 
-Previously, every PUT request which included a new chromosome increased 
-the size of the pool. But in a long-running experiment this has to be avoided
-in order to not exhaust the available memory. In this experiment the size
-of a chromosome was 1000 double precision numbers, so that considering this
-and the memory available in the server, the pool was limited 
-to 10,000 chromosomes.  In order to keep this size
-constant, a random chromosome was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
-and {\tt classic-ea-float} modules were used. The parameters for the EA 
-were tournament size = 3 and  population size = 500, returning exchanging a
-chromosome with the server every 100 generations. Again, there were two workers in each window.
-As before, a few participation calls were published in social networks.
-
-\subsubsection{Results}
-
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics{rastrigin-fitness.png}
-\caption{Fitness of chromosomes sent to the server pool along time in
-  the $x$ axis.} 
-% Plotted with ../data/plot-rastrigin-fitness.R
-\label{fig:puts:rastrigin}
-\end{figure}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
-\caption{Number of unique IPs participating in the Rastrigin
-  experiment per minute (left) and per hour (right)} 
-% Plotted with ../data/plot-rastrigin-IPs.R
-\label{fig:ips:rastrigin}
-\end{figure}
-%
-\begin{table*}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 35.759  &  343.336   & 9.877 \\
-Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
-\caption{Data and fit to GEV (left) and Weibull (right) of the number
-  of {\tt PUT}s per worker for the F15 (Rastrigin) function.}  
-% Plotted with ../data/ips-puts-rastrigin.R
-\label{fig:fit:rastrigin}
-\end{figure}
-%
-
-The results for this new experiment have been extremely interesting
-and quite different from the previous one. The most important
-characteristic that set this experiment apart from the previous one is
-the fact that it did not find the solution for the time it was
-running, instead lowering, little by little, the fitness until it was
-disconnected. The fitness of chromosomes sent by clients is shown in
-Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
-upper part, which represents the fitness of clients that have just
-started evolution from a random population, and then the fitness of
-those clients that have been already evolving for a while or obtained
-a high-fitness individual from the pool and evolved from there, and
-which managed to surpass the 9000 value by the end of the
-simulation. It incidentally shows also the number of {\em new} nodes
-coming: there is a dense {\em cloud} at the beginning, but the new
-contributions keep thinning as the evolution progresses.
-
-This is in part due to the fact that the presence of nodes is not
-constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
-number of unique IPs participating in the experiment per minute (left)
-and hour (right). The two graphs have the same shape which matches
-also the speed in which fitness is decreased in Figure
-\ref{fig:puts:rastrigin}. At the beginning there are very few or only
-one, but later one there are peaks of almost 30 unique machine a
-minute; these peaks appear again when new announcements are made in
-Twitter or any other social networks, to go down to a few remaining
-nodes later on. It is interesting to note that, at the most, more than
-100 unique nodes were participating in the experiment during one hour;
-also that there are sustained peaks during a day of 6-7 computers
-every minute. This is due to the fact that the computation is
-asynchronous, but also to the fact that different computers stay in
-the simulation for a different period of time. This period has been
-plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
-{\tt PUT}s per node, ranked by number, along with a fit to the two
-distributions we have used before, GEV and Weibull; the parameters of
-this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
-experiments, in which GEV fit was relatively good, in this case it is
-quite obviously not. However, the fit to the Weibull distribution,
-shown at the left of the image, is relatively good. As should be
-expected, the shape and scale are different to the one in the previous
-experiments with the Trap function, however, the shape is quite
-similar (0.40 vs 0.69), being in both cases less than one, which
-indicates a decreasing distribution of the same kind. In fact, the
-difference in the scale parameter indicates that the difference
-between the time devoted by a volunteer and the next is bigger in this
-case. This is probably due to the fact that the duration of the
-experiment is bigger, allowing for bigger differences in
-behavior. While the first in the Trap function might have devoted a
-few minutes and the second a few seconds less, the difference between
-the first and second in this experiment might be in the scale of
-hours.
-
-% (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
-There are several validity threats to the design of this study.
-Participants of the experiments were mainly members of our 
-own social networks, personal colleagues, friends 
-and acquaintances, therefore to extend this work other options 
-should be considered in order to include 
-a wider audience with different motives for participation.
-Another threat is that participants were anonymous so the amount of repeated 
-participations and the effects this might have had in the outcome is not
-known. Another issue is that the scalability of the system was not thoroughly tested
-but for the amount of participants the basic configuration 
-was more than adequate, then again more experiments are needed in order
-to assess the capability both with more clients and other server
-configurations. Although care was taken to include representative benchmarks
-of real world problems, more experiments with a broader coverage 
-should allow a better understanding of the practical limitations of
-the architecture and the tools chosen.
-
-In general, this experiment confirms what was found in the others in
-several aspects: first, there is a regularity in the distribution of the
-time assigned by volunteers to it, second, there is a good and certain
-amount of volunteers that will become part of the distributed
-computer, in some cases up to one hundred; also, in some cases there
-will be users that will spend several hours, two orders of magnitude
-more generations in this case than in the previous one. There is a
-clear influence of the time to solution: a short time means that every
-experiment will only gather a few volunteers, but also that they will
-stay only for a short time. A long time will make users stay for
-longer, but if it takes too long, they will eventually lose interest.
-
-We will discuss this and the rest of the findings of the paper in the
-conclusions.
-
 %---------------------------------------------------------------
@@
 \label{sec:conclusion}
-
-In this paper two versions of a client-server architecture for volunteer and distributed
-evolutionary algorithms have been evaluated. The architecture uses the browser and has been generated using {\sf
-  NodIO}, based on the {\sf NodEO} evolutionary algorithm library. Volunteers are {\em in the cloud}, as stated in the title,
-since they are a {\em CPU as a service} for the persons running the
-experiment. In fact, in this paper we have tried to put some figures
-on the real size of that {\em cloud} and how it can be used standalone
-if there is no alternative, or in conjunction with other local or
-cloud-based methods to add computing power in a seamless way through
-the pool that {\sf NodIO} creates.
 
@@
 correlation between the number of IPs in successive experiments, and also resulted in a low number of generations allotted by users.
-
-In the second implementation, {\sf NodIO-W$^2$}, Web Workers
-were used so that several clients per browser could be executed and
-they could run in the background, so that the the number of
-generations per user could be improved. The number of generations per
-user increased to the point that baseline performance was improved in
-most cases. The architectural changes led to a high correlation
-between successive experiments, so that, even as the median number of
-volunteers per experiment decreased, they were probably contributing
-to the experiment at the same time, so that our objective was
-achieved. In general, one of the first conclusions for this paper is
-that the browser technologies should be used to its full extent so
-that the user time is leveraged to its full potential, and this was
-achieved with {\sf NodIO-W$^2$}.
-
-The second objective of this paper was to model the user behavior in a
-first attempt to try and predict performance. As should be expected,
-the model depends on the implementation, with contributions following
-a GEV distribution in the case of {\sf NodIO} and a
-Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
-close enough to each other and, in the second case, reflect the fact
-that all time spent in the page is actually devoted to computing,
-which is why the time spent (represented by the number of
-contributions to the pool) follows a model quite similar to that found
-for games or other online activities. The reverse might be true: if we
-want to have returning users for the experiments, it is probable that
-we should {\em gamify} the experience so that once they have done it
-once, they might do it more times. In the spirit of Open Science, this
-gamification might involve computing in real-time data such as the one
-presented in this paper, while showing it in the same page.
-
-In general, linking and finding correlations between user choices and
-performance is an interesting avenue to explore in the future. Even if
-the three previous experiments were published in a similar way, one
-obtained up to 5 times more total cycles  than the one with the least
-number of cycles. It is also essential to obtain volunteers as
-simultaneously as possible, so it is possible that the features of the
-social network in terms of real-time use will also play a big
-role. Even as it is difficult to create controlled experiments in this
-area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation
-
-Although we think that, in general, the results presented in this
-paper are independent of the problem chosen, it is true that even if
-it is a difficult problem for evolutionary algorithms, it takes a few
-seconds with the right settings to solve. This makes it difficult also
-for this kind of asynchronous framework, but it would be interesting
-to check the performance for other computationally demanding
-tasks, especially ones that cover the middle ground between a problem
-finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
-Besides, the two versions of the framework include as an
-algorithmic variant using random population size. We do not think this
-has had a big influence on the results and in fact this is not noticed
-for F15, which is the second version tested, but it would be interesting to
-measure exactly what this influence has been. In general there are
-many issues with the evolutionary algorithm implementation itself,
-including using different, or adaptive, policies for inserting and
-sending individuals to the pool as was done in \cite{araujo2008mam},
-using different policies for population initialization, and also the
-incorporation of high-speed local resources to the pool to check what
-would be the real influence of the volunteer pool to the final
-performance.
-
-Another area is how to enhance the number and quality of
-volunteers. For instance, adding a bit of more
-control to the user might contribute also to gamification. Right now
-the user has only two Web Workers. It is a matter of a single click to
-open more tabs in the browser, giving more Web Workers, but it would be
-interesting to put everything in a single page and under our control,
-to check how often this happens and under which conditions.
-
-Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
-to create a generator in which the user just has to create a fitness
-function, with the rest of the framework wrapped around
-automatically. All this, as well as the data used for this paper and the paper
-itself, has been published with a free license in GitHub at
-\url{https://github.com/JJ/modeling-volunteer-computing}.
 

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Proposes a title and abstract
Refs #82 #83

Please @mariosky @fergunet @unintendedbear @pacastillo have a look and make or propose changes.

@@
 
-\title{}
+\title{Revisiting evolutionary algorithms in JavaScript}
 
@@
 
+JavaScript is an interpreted language mainly known for its inclusion
+as an embedded language for programming internet browsers. This has
+inspired its use, for a long time, as a tool for evolutionary
+algorithms; several libraries have been published so far and are in
+use. However, the last years have seen a resurgence of interest in the
+language, becoming one of the most popular and thus spawning the
+improvement of its implementations, which form now the base of many
+new client-server applications. That is why in this paper we propose a series of
+measurements to establish the speed of JavaScript in evolutionary
+algorithms that can serve as a baseline for comparison with other
+distributed computing experiments. These experiments use different
+integer and floating point problems, and prove that the speed of
+JavaScript is actually competitive with other languages commonly used
+by the evolutionary algorithm practitioner.
+
 \end{abstract}
-
-% Work in progress..
-%Volunteer computing is used routinely in science and industry, but it
-%still has a long way to go if it is going to become a mainstream technology 
-%in the same way the cloud technologies in general have. In one hand, projects 
-%lack special purpose and open source tools with analytics builtin and
-%in the other there are still barriers that keep users from sharing their computing 
-%resources like the need to install software applications or plugins in order 
-%to cooperate. That is why in this paper we present {\sf NodIO}, a client-server
-%architecture for distributed
-%evolutionary algorithms, written entirely in JavaScript, which needs
-%minimal server support to work and is able to support persistent,
-%asynchronous, distributed evolutionary algorithms running in the
-%browser.  The architecture, along with some initial results on the
-%expected performance and scaling
-%behavior is presented. We also use the algorithm data to
-%find out what kind of patterns the user behavior follow and try to fit
-%them to some statistical distribution.
 
@@
 \end{figure}
-%
-\begin{table*}[!htb]
-\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
-  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
-\begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
-{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
 

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
This paper is going to need a bit of restructuring
Changing the title again. Please @mariosky @pacastillo @unintendedbear take a look. Anyone, in fact. Refs #82 #83

I have put a title in every set of paragraphs to see how we can structure it

@@
 
-\title{Revisiting evolutionary algorithms in JavaScript}
+\title{{\sf NodIO}, a framework for evolutionary algorithms in JavaScript: first results}
 
@@
 inspired its use, for a long time, as a tool for evolutionary
-algorithms; several libraries have been published so far and are in
+algorithms, mainly so in browser-based volunteer computing
+environments. Several libraries have also been published so far and are in
 use. However, the last years have seen a resurgence of interest in the
@@
 
-The rest of the paper is organized as follows:
-
-%---------------------------------------------------------------
-\section{State of the art}
-\label{sec:soa}
-
-%---------------------------------------------------------------
-\section{Description of the framework}
-\label{sec:description}
-
 In general, a distributed volunteer-based evolutionary computation
 framework based on the browser is simply a client-server system
-whose client is, or it can be, embedded in the browser via
+whose client is, or can be, embedded in the browser via
 JavaScript. Since JavaScript is  the only language that is present
 across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
+be some issues with the performance of the language itself, which 
  is
 why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
+\cite{2015arXiv151101088M}. That paper shows that JavaScript can be
 faster than compiled languages such as Scala and, in any case, it provides
@@
 \ref{sec:w2}.
+
+%%% Issues with volunteer based systems
 
@@
 
+%%% Versions of the framework
+
 In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
 cloud or bare metal based volunteer evolutionary computing system
-derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
- the {\sf NodEO} library, whose architecture has been
+derived from the {\sf NodEO} library, whose architecture has been
 developed using JavaScript on the client as well as the server.
@@
 
+% Description of the framework
 The figure \ref{fig:system} describes the general system architecture and

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Reorganizing the description of the framework.
So far we have done two things: talked about JavaScript and described the framework and how it addresses the challenges of volunteer computing frameworks. Closes #82. Refs #83

@@
 improvement of its implementations, which form now the base of many
-new client-server applications. That is why in this paper we propose a series of
+new client-server applications. We present such an application for
+running distributed or volunteer-based evolutionary algorithm
+experiments, and we make a series of
 measurements to establish the speed of JavaScript in evolutionary
@@
 
-In general, a distributed volunteer-based evolutionary computation
-framework based on the browser is simply a client-server system
-whose client is, or can be, embedded in the browser via
-JavaScript. Since JavaScript is  the only language that is present
-across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself, which 
- is
-why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}. That paper shows that JavaScript can be
-faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
-in evolutionary computation and numerical optimization such as Python. 
-% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
-Compiled languages such as
-Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
-considering the performance of a single-user, 
-but this is more than offset by the computational power of
-the spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets. Together, the performance is several orders of magnitude
-higher, which is the objective in this kind of systems.
-% (Paloma) This sentence needs re-writing
-% "Other compiled languages such as Java or C might be, in fact,
-% faster, but the fact that single-user performance is higher (than
-% what?) is over the offset. This is because the number of spontaneous
-% volunteers we can gather at whim, including people using their
-% mobile phones or tablets, is several orders of magnitude higher,
-% which is the objective in this kind of systems." (?) 
- JavaScript is
-quite a popular language nowadays, providing compilers from a number
-of languages, many of them in their family (like CoffeeScript), but
-many other outside, so researchers can, in fact, write their fitness
-function in Python, Erlang, Scala, or even Java, and then compile it to
-JavaScript. For completeness, an experiment comparing the performance of
-a JavaScript implementation versus Java and Matlab is included in section
-\ref{sec:w2}.
+The need to use non-permanent and potentially massive computing
+infrastructure for distributed scientific experiments has lead to the
+rising of {\em volunteer computing} as a set of tools that allow
+citizens to {\em donate} cycles to scientific experiments by running
+an application \cite{Anderson06thecomputational}; this application can be embedded in a web page. And
+since browsers are ubiquitous, these kind of systems are the ones that
+interest we the most and has led us to create a framework for
+evolutionary algorithm experiments that is able to work in that mode
+\cite{DBLP:journals/corr/GuervosG15}. {\sf NodIO} is a
+cloud or bare metal based volunteer evolutionary computing system
+derived from the {\sf NodEO} library, whose architecture has been
+developed using JavaScript on the client as well as the server.
+All parts of the framework are free and available with a free license
+from \url{https://github.com/JJ/splash-volunteer}.
 
@@
 
-As a volunteer based system, there are some issues worth 
-noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
+As a volunteer based system, there are some issues that have to be addressed \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
 \item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
@@
 
-%%% Versions of the framework
+% Justification for JavaScript
+In general, a distributed volunteer-based evolutionary computation
+framework based on the browser such as NodIO is simply a client-server system
+whose client is, or can be, embedded in the browser via
+JavaScript. Since JavaScript is  the only language that is present
+across all browsers, the choice was quite clear. Even so, there might
+be some issues with the performance of the language itself, which 
+ is
+why we have made a comparison between JavaScript and other languages
+\cite{2015arXiv151101088M}. In that paper we prove that JavaScript can be
+faster than compiled languages such as Scala and, in any case, it provides
+a performance that is comparable to other languages usually employed % (Paloma) Adverb
+in evolutionary computation and numerical optimization such as Python. 
+Compiled languages such as
+Java or C might be faster when 
+considering the performance of a single-user, 
+but this is more than offset by the computational power of
+the spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets. Together, the performance is several orders of magnitude
+higher, which is the objective in this kind of systems.
 
-In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
-cloud or bare metal based volunteer evolutionary computing system
-derived from the {\sf NodEO} library, whose architecture has been
-developed using JavaScript on the client as well as the server.
-All parts of the framework are free and available with a free license
-from \url{https://github.com/JJ/splash-volunteer}.
+Since JavaScript is
+quite a popular language nowadays, it also provides compilers from a number
+of languages, many of them in their family (like CoffeeScript), but
+many other outside, so researchers can, in fact, write their fitness
+function in Python, Erlang, Scala, or even Java, and then compile it to
+JavaScript. For completeness, an experiment comparing the performance of
+a JavaScript implementation versus Java and Matlab is included in section
+\ref{sec:w2}. We will first describe the {\sf NodIO} architecture in
+the next Section.
 
-%PABLO: el siguiente párrafo justifica Javascript
+\section{{\sf NodIO architecture}
 
-Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
+{\sf NodIO} architecture has two tiers. 
+The figure \ref{fig:system} describes the general system architecture and
+algorithm behavior. Different web technologies, such as JQuery or {\tt
+  Chart.js} have
+been used to build the user interface elements of the framework.
+\begin{figure}[!t]
+\centering
+\includegraphics[width=4in]{system.pdf}
+\caption{Description of the proposed framework. Clients execute a JavaScript EA
+  in the browser, which, every 100 generations, sends the best
+  individual and receives a random one from the server.}
+\label{fig:system}
+\end{figure}
+\begin{enumerate}
+
 \item A REST server, that is, a server that includes several {\em
@@
   that is reset when the solution is found.
+
 \item A client that includes the evolutionary algorithm as
@@
   the algorithm can be run in many different ways: 
-  stopping when a solution is found or, as we have done in this paper,
-  using Web Workers. In fact, since it is a pool-based system such as
+  stopping when a solution is found or using Web Workers. In fact, since it is a pool-based system such as
   the one described in \cite{LNCS86720702}, any kind of client that
-  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
-But never heard of a client which "works" an API.
-Is this correct?
+  calls the application programming interface (API) can be used,
   written in any kind of language. However, in this paper we are going
@@
 
-% Description of the framework
-The figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or {\tt
-  Chart.js} have
-been used to build the user interface elements of the framework.
-\begin{figure}[!t]
-\centering
-\includegraphics[width=4in]{system.pdf}
-\caption{Description of the proposed framework. Clients execute a JavaScript EA
-  in the browser, which, every 100 generations, sends the best
-  individual and receives a random one from the server.}
-\label{fig:system}
-\end{figure}
-
 The researcher only has to change a function, which can be written in
 JavaScript or in any other language with cross-compilation to
-JavaScript \cite{web:compilersjs} to solve a different
-problem. In this case the classical Trap function \cite{Ackley1987} has been
-used as well as the Rastrigin's floating-point optimization
-problem. JavaScript is a functional language and declaring a different 
-function and handing at the creation of the algorithm object, called
-{\tt Classic}, is the only change needed to work with a different
-problem. Let us see how this system addresses the different challenges
+JavaScript \cite{web:compilersjs} to solve different optimization
+problems.
+
+Let us see how this system addresses the different challenges
 outlined in the State of the Art (Section \ref{sec:soa}):
@@
 
-All these points will be established in the next two sections,
-starting with the baseline, single worker system next.
+However, before testing this framework in a real experiment, we have
+to establish a baseline for comparison.
 
@@
 \label{sec:experiments}
+
+We have used the classical Trap function \cite{Ackley1987}
+used as well as the Rastrigin's floating-point optimization
+problem. Since JavaScript is a functional language and declaring a different 
+function and handing at the creation of the algorithm object, called
+{\tt Classic}, is the only change needed to work with a different
+problem.
 

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Finished rearrangement of the paper.
Now let us go to the conclusions. Refs #83

@@
   in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
+  the scalability of the communication system. % (Paloma) Is it ok if
+                                % we talk about scalability when we
+                                % were talking about adaptiveness?
+                                % Scalability was also described in
+                                % the first point. 
   %(Mario) Scalability of the communication system in particular?
@@
 
-However, before testing this framework in a real experiment, we have
-to establish a baseline for comparison.
-
-%---------------------------------------------------------------
-\section{Baseline Experiments} 
-\label{sec:experiments}
-
-We have used the classical Trap function \cite{Ackley1987}
-used as well as the Rastrigin's floating-point optimization
-problem. Since JavaScript is a functional language and declaring a different 
-function and handing at the creation of the algorithm object, called
-{\tt Classic}, is the only change needed to work with a different
-problem.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=10cm]{baseline-times.png}
-\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
-    Node.js} client, for different population sizes. Only
-runs in which solution has been found are used for this graph.}
-% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
-\label{fig:baseline}
-\end{figure}
-First, we establish the baseline performance by running fifty runs of the
-evolutionary algorithm in a desktop client written using {\sf NodEO}
-\cite{nodeo2014}, the basic JavaScript library we have used as base to
-build {\sf NodIO}, the framework. This experiment tries to 
-find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$ and a population size equal to 512. The algorithm in a
-single island was run until the solution (a string with all
-ones) was found or five million evaluations had been performed. It
-took around a minute, on average, that is $68.9694$ seconds, to
-perform these runs. In this experiment only 33, that is, 66\% were
-successful. The experiments were repeated for population $p=1024$,
-with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
-seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
-characteristics: 
-\begin{verbatim}
-Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
-2014 x86_64 x86_64 x86_64 GNU/Linux
-\end{verbatim}
-with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
-
-These two experiments establish a baseline result and show that
-the time to success depends on the population size, with a bigger population
-contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
-experiments that we will describe next do not and can not have the
-same conditions, but
-the baseline is that if they eventually take longer than a basic
-desktop, their interest will be purely academic.
-
-\section{Enhancing {\sf NodIO} with a Web Worker-based architecture}
-\label{sec:w2}
-
-In order to improve the number of cycles per user, we enhanced the
-basic architecture in several different aspects.  One improvement 
+After the initial experiments \cite{DBLP:journals/corr/GuervosG15}, we
+realized that we needed to make some improvements. So, in order to
+improve the number of cycles per user, we enhanced the 
+basic architecture in several different aspects. One proposed improvement
 involved restarting the client once the solution was
@@
 population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvement was to 
-use the Web Worker API as described next.
+different for each client. We called this instance of
+the architecture NodIO-W$^2$. The other improvement was to 
+use the Web Worker API as described next
 
-JavaScript has been commonly used to develop client-side scripts
+The rationale for this change comes from the realization that JavaScript has been commonly used to develop client-side scripts
 that handle the user interface, communicating asynchronously with servers and
@@
 
-A sequence diagram of the enhanced NodIO, which is called NodIO-W$^2$, is presented
+A sequence diagram of NodIO-W$^2$, is presented
 in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
@@
 %
+
+These two versions of the framework will have to be tested {\em in the
+  wild} to measure its performance. 
+However, before testing this framework in a real experiment, we have
+to establish a baseline for comparison.
+
+%---------------------------------------------------------------
+\section{Baseline Experiments} 
+\label{sec:experiments}
+
+We have used the classical Trap function \cite{Ackley1987}
+used as well as the Rastrigin's floating-point optimization
+problem. Since JavaScript is a functional language and declaring a different 
+function and handing at the creation of the algorithm object, called
+{\tt Classic}, is the only change needed to work with a different
+problem.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=10cm]{baseline-times.png}
+\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
+    Node.js} client, for different population sizes. Only
+runs in which solution has been found are used for this graph.}
+% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
+\label{fig:baseline}
+\end{figure}
+First, we establish the baseline performance by running fifty runs of the
+evolutionary algorithm in a desktop client written using {\sf NodEO}
+\cite{nodeo2014}, the basic JavaScript library we have used as base to
+build {\sf NodIO}, the framework. This experiment tries to 
+find the solution to the 40-trap function with parameters $l=4, a=1,
+b=2, z=3$ and a population size equal to 512. The algorithm in a
+single island was run until the solution (a string with all
+ones) was found or five million evaluations had been performed. It
+took around a minute, on average, that is $68.9694$ seconds, to
+perform these runs. In this experiment only 33, that is, 66\% were
+successful. The experiments were repeated for population $p=1024$,
+with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
+seconds. Results for these two experiments
+are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
+characteristics: 
+\begin{verbatim}
+Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
+2014 x86_64 x86_64 x86_64 GNU/Linux
+\end{verbatim}
+with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
+
+These two experiments establish a baseline result and show that
+the time to success depends on the population size, with a bigger population
+contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+experiments that we will describe next do not and can not have the
+same conditions, but
+the baseline is that if they eventually take longer than a basic
+desktop, their interest will be purely academic.
 
@@
 The results also show that there is not much difference between the Node.js
-implementation an that of a
+implementation and that of a
 browser. There is also not much difference between running the code in the main thread or in Web
@@
 \label{sec:conclusion}
-
-In order to establish a baseline performance, the evolutionary
-algorithm was run in a desktop client-program written in JavaScript
-using NodEO to solve the 40-trap function. The first experiments with
-{\sf NodIO} proved that, although obtaining better performance than the
-baseline was possible, it did not happen, mainly because of the
-difficulty in carrying over other experiments to volunteers when the
-experiment they were participating on was finished. This has been shown in the low
-correlation between the number of IPs in successive experiments, and also resulted in a low number of generations allotted by users.
 

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Fixes errors
And rearranges figures a bit. Refs #83
Final draft. Refs #83. Please @mariosky @fergunet @unintendedbear revise for bad copypaste, coherence, and the rest

@@
 \usepackage{url}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works}
 
@@
 a JavaScript implementation versus Java and Matlab is included in section
-\ref{sec:w2}. We will first describe the {\sf NodIO} architecture in
+\ref{sec:experiments}. We will first describe the {\sf NodIO} architecture in
 the next Section.
 
-\section{{\sf NodIO architecture}
+\section{{\sf NodIO } architecture}
 
-{\sf NodIO} architecture has two tiers. 
+The {\sf NodIO} architecture has two tiers. 
 The figure \ref{fig:system} describes the general system architecture and
@@
 Let us see how this system addresses the different challenges
-outlined in the State of the Art (Section \ref{sec:soa}):
+outlined above:
 \begin{itemize}
@@
 how practical and mature this language is for developing complex mathematical functions.
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{f15_times.png}
+\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
+\label{fig:f15_times}
+\end{figure}
 
@@
 with acceptable performance.
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{f15_times.png}
-\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
-\label{fig:f15_times}
-\end{figure}
 
@@
 \label{sec:conclusion}
+
+This paper has been intended mainly as a description of the {\sf
+  NodIO} architecture as well as two different implementations, with
+or without web workers, as well as a collection of measurements of the
+performance of the JavaScript language in different implementations,
+and compared with other languages.
+
+The first conclusion we can draw is that, even if JavaScript is not
+the fastest language around, its performance is competitive and can be
+used profitably for medium size problems, with its speed in floating
+point problems being around 30\% less than Matlab or Java.
+
+A framework created around this language, like the {\sf NodIO} we
+systematize and introduce in this paper, can more than compensate this
+difference with the recourse of the users that can be spontaneously
+gathered in a volunteer computing experiment. The fact that it is so
+simple to participate in it, without needing to download anything,
+implies that it can be used for peak or non-permanent workloads.
+
+However, it remains to be seen how many users and the real performance
+that {\sf NodIO}, in any of its implementations, would achieve. This
+is left for future work.
 

### 2016-01-06 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Few changes

@@
 JavaScript is an interpreted language mainly known for its inclusion
-as an embedded language for programming internet browsers. This has
+as an embedded language for programming web browsers. This has
 inspired its use, for a long time, as a tool for evolutionary
@@
 language, becoming one of the most popular and thus spawning the
-improvement of its implementations, which form now the base of many
+improvement of its implementations, which are now the foundation of many
 new client-server applications. We present such an application for
-running distributed or volunteer-based evolutionary algorithm
+running distributed volunteer-based evolutionary algorithm 
 experiments, and we make a series of
@@
 infrastructure for distributed scientific experiments has lead to the
-rising of {\em volunteer computing} as a set of tools that allow
+rising of {\em volunteer computing} defined as a set of tools that allow
 citizens to {\em donate} cycles to scientific experiments by running
@@
 since browsers are ubiquitous, these kind of systems are the ones that
-interest we the most and has led us to create a framework for
+we are interested the most and this has led us to create a framework for
 evolutionary algorithm experiments that is able to work in that mode

### 2016-01-06 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Just a sentence

@@
 an application \cite{Anderson06thecomputational}; this application can be embedded in a web page. And
-since browsers are ubiquitous, these kind of systems are the ones that
-we are interested the most and this has led us to create a framework for
+since browsers are ubiquitous, we are more interested in these kind of systems, this has led us to create a framework for
 evolutionary algorithm experiments that is able to work in that mode

### 2016-01-07 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
Title proposal

@@
 \title{{\sf NodIO}, a framework for evolutionary algorithms in JavaScript: first results}
+
+%(Mario) Maybe:  {\sf NodIO}, a JavaScript framework for volunteer-based evolutionary algorithms.
 
@@
 
-As a volunteer based system, there are some issues that have to be addressed \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
+As a volunteer based system, there are some issues that have to be addressed 
+\cite{sarmenta2001volunteer,web:BOINC} :
+\begin{enumerate}
 \item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
@@
 the architecture NodIO-W$^2$. The other improvement was to 
-use the Web Worker API as described next
+use the Web Worker API as described next.
 

### 2016-01-07 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
A few changes here and there

@@
 
-\title{{\sf NodIO}, a framework for evolutionary algorithms in JavaScript: first results}
+\title{{\sf NodIO}, a JavaScript framework for volunteer-based evolutionary algorithms : first results}
 
-%(Mario) Maybe:  {\sf NodIO}, a JavaScript framework for volunteer-based evolutionary algorithms.
+%(Mario) Maybe:  {\sf NodIO}, a JavaScript framework for
+%volunteer-based evolutionary algorithms.
+% Accepted with a small change :-) - JJ
 
@@
 JavaScript is an interpreted language mainly known for its inclusion
-as an embedded language for programming web browsers. This has
+in web browsers, making them a container for rich Internet based applications. This has
 inspired its use, for a long time, as a tool for evolutionary

### 2016-01-08 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Done with rev3 refs #81
Mark eliminated sections
To be later on suppressed and the rest of the paper stitched together. Closes #82
Changing authors' names to asked format, references #84. But still I don't know how to put the corresponding author as a footnote like in other IEEE CIM papers

@@
 
-\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
-\thanks{Manuscript submitted for review on \today.}%
-\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
-\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
-\thanks{$^3$Dept. of Statistics and Operational Research University of Granada}%
-\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
-    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es},{\tt nrico@ugr.es}}%
-\thanks{*Corresponding author.}%
+\author{Juan-J.~Merelo, Dept.
+of Computer Architecture and Technology, University of Granada, Granada, Spain\\Mario Garc\'ia-Valdez, Dept.
+of Graduate Studies, Instituto Tecnológico de Tijuana, City, Mexico\\Pedro A.
+Castillo, Dept.
+of Computer Architecture and Technology, University of Granada, Granada, Spain\\Pablo Garc\'ia-S\'anchez, Dept.
+of Computer Architecture and Technology, University of Granada, Granada, Spain\\P.
+de las Cuevas, Dept.
+of Computer Architecture and Technology, University of Granada, Granada, Spain\\Nuria Rico, Dept.
+of Statistics and Operational Research, University of Granada, Granada, Spain%
+%\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
+%\thanks{Manuscript submitted for review on \today.}%
+%\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
+%\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
+%\thanks{$^3$Dept. of Statistics and Operational Research University of Granada}%
+%\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
+%    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es},{\tt nrico@ugr.es}}%
+\thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
 }
@@
 %---------------------------------------------------------------
+% Start of content for other paper ----
 \section{Description of the framework}
@@
 with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
+
+% end spin-off ------------------------------------------------------
 
@@
 
+% -- Start spin-off -------------------
 \section{Enhancing {\sf NodIO} with a Web Worker-based architecture}
@@
 %
+% End spin-off ----------------------------------------------------------------
 \begin{table*}[!htb]
@@
 \end{table*}
+
+% - Start spin-off again
+% -------------------------------------------------------
 
@@
 \end{enumerate}
-%
+% End spin-off--------------------------------------------
+
 \begin{figure}[!htb]
@@
 
+% ------------Start spin-off ----------------------------------------------------
 \subsection{Tests with a floating-point hard optimization problem}
@@
 \end{figure}
+% - end spin-off-----------------------------------------------------------------
 

### 2016-01-08 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Completing my name
Proposal for abstract reduction. References #85

@@
 of Computer Architecture and Technology, University of Granada, Granada, Spain\\Pablo Garc\'ia-S\'anchez, Dept.
-of Computer Architecture and Technology, University of Granada, Granada, Spain\\P.
-de las Cuevas, Dept.
+of Computer Architecture and Technology, University of Granada, Granada, Spain\\Paloma de las Cuevas, Dept.
 of Computer Architecture and Technology, University of Granada, Granada, Spain\\Nuria Rico, Dept.
@@
 \begin{abstract}
-Volunteer computing is extensively used in science and industry, but
-it still has a long way to go in order to become a mainstream
-technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
-The lack of general purpose and open
-source tools with built-in analytics could be one of the reasons of
-why this is so, as well as the lack of proofs of concept.
-% (Paloma) The following sentence is way too long, and the "which" after "Javascript" is confusing, so one does not know if the which refers to the framework or Javascript by reading the whole sentence.
-% I propose: "That is why in this paper we present NodIO, a framework used to create client-server architectures for distributed evolutionary algorithms, entirely written in JavaScript.
+Volunteer computing is extensively used in science and industry, yet it still has not become a mainstream technology in the same way the cloud technologies in general are.
+The lack of general purpose and open source tools with built-in analytics could be one of the reasons of why this is so, as well as the lack of proofs of concept.
+That is why we have enhenced {\sf NodIO}, a framework used to create client-server architectures for distributed evolutionary algorithms, resulting in {\sf NodIO-W$^2$}, which we present in this paper.
+{\sf NodIO-W$^2$} leverages the HTML5 Web Workers specification to execute in the background several clients per browser, so that the the number of generations per user could be improved.
+The enhancements on the NodIO architecture, along with some initial results on the expected performance and scaling behavior when the number of users is increased, are introduced.
+We also use the algorithm data to find out what kind of patterns the user behavior follows by examining the resources they devote to our experiment, and try to fit them in some statistical distribution.
+In general, our experiments prove, first, that {\sf NodIO} is a flexible and low-overhead platform to perform distributed computation experiments that include volunteers, and then that user behavior follows some patterns, and, thus, are amenable to a model.
+% Volunteer computing is extensively used in science and industry, but
+% it still has a long way to go in order to become a mainstream
+% technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
+% The lack of general purpose and open
+% source tools with built-in analytics could be one of the reasons of
+% why this is so, as well as the lack of proofs of concept.
+% % (Paloma) The following sentence is way too long, and the "which" after "Javascript" is confusing, so one does not know if the which refers to the framework or Javascript by reading the whole sentence.
+% % I propose: "That is why in this paper we present NodIO, a framework used to create client-server architectures for distributed evolutionary algorithms, entirely written in JavaScript.
 Our framework needs minimal server support to work and is able to support persistent, asynchronous, distributed evolutionary algorithms running in the browser, as well as any other client consuming its REST application programming interface."
-That is why
-in this paper we present {\sf NodIO}, a framework used to create
-client-server architectures for distributed evolutionary algorithms,
-entirely written in JavaScript, which needs minimal server support to
-work and that is able to support persistent, asynchronous, distributed
-evolutionary algorithms running in the browser, as well as any other
-client consuming its REST application programming interface. An
-enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
-also presented; this version leverages the HTML5 Web Workers
-specification to execute several clients per browser and do it in the
-background, so that the the number of generations per user could be
-improved. The architecture, along with some initial results on the
-expected performance and scaling behavior when the number of users is
-increased is introduced. % (Paloma) I've changed this "presented" for a synonym
-We also use the algorithm data to find out
-what kind of patterns the user behavior follows by examining the
-resources they devote to our experiment, and try to fit them in % (Paloma) Changed. Phrasal verb was wrong.
- some
-statistical distribution to create the bases of a model of user
-behavior in volunteer computing experiments of this kind that can be
-used by researchers to predict their performance. %(Paloma) Changed to "their" because of "experiments"
-In general, our
-experiments prove, first, that {\sf NodIO} is a flexible and
-low-overhead platform to perform distributed computation experiments
-that include volunteers, and then that user behavior follows some
-patterns, and, thus, are amenable to a model.
+% That is why
+% in this paper we present {\sf NodIO}, a framework used to create
+% client-server architectures for distributed evolutionary algorithms,
+% entirely written in JavaScript, which needs minimal server support to
+% work and that is able to support persistent, asynchronous, distributed
+% evolutionary algorithms running in the browser, as well as any other
+% client consuming its REST application programming interface. An
+% enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
+% also presented; this version leverages the HTML5 Web Workers
+% specification to execute several clients per browser and do it in the
+% background, so that the the number of generations per user could be
+% improved. The architecture, along with some initial results on the
+% expected performance and scaling behavior when the number of users is
+% increased is introduced. % (Paloma) I've changed this "presented" for a synonym
+% We also use the algorithm data to find out
+% what kind of patterns the user behavior follows by examining the
+% resources they devote to our experiment, and try to fit them in % (Paloma) Changed. Phrasal verb was wrong.
+%  some
+% statistical distribution to create the bases of a model of user
+% behavior in volunteer computing experiments of this kind that can be
+% used by researchers to predict their performance. %(Paloma) Changed to "their" because of "experiments"
+% In general, our
+% experiments prove, first, that {\sf NodIO} is a flexible and
+% low-overhead platform to perform distributed computation experiments
+% that include volunteers, and then that user behavior follows some
+% patterns, and, thus, are amenable to a model.
 \end{abstract}

### 2016-01-08 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Edited the abstract and removed long lines.

@@
 of Computer Architecture and Technology, University of Granada, Granada, Spain\\Mario Garc\'ia-Valdez, Dept.
-of Graduate Studies, Instituto Tecnológico de Tijuana, City, Mexico\\Pedro A.
+of Graduate Studies, Instituto Tecnológico de Tijuana, Tijuana, M\'exico\\Pedro A.
 Castillo, Dept.
@@
 \begin{abstract}
-Volunteer computing is extensively used in science and industry, yet it still has not become a mainstream technology in the same way the cloud technologies in general are.
-The lack of general purpose and open source tools with built-in analytics could be one of the reasons of why this is so, as well as the lack of proofs of concept.
-That is why we have enhenced {\sf NodIO}, a framework used to create client-server architectures for distributed evolutionary algorithms, resulting in {\sf NodIO-W$^2$}, which we present in this paper.
-{\sf NodIO-W$^2$} leverages the HTML5 Web Workers specification to execute in the background several clients per browser, so that the the number of generations per user could be improved.
-The enhancements on the NodIO architecture, along with some initial results on the expected performance and scaling behavior when the number of users is increased, are introduced.
-We also use the algorithm data to find out what kind of patterns the user behavior follows by examining the resources they devote to our experiment, and try to fit them in some statistical distribution.
-In general, our experiments prove, first, that {\sf NodIO} is a flexible and low-overhead platform to perform distributed computation experiments that include volunteers, and then that user behavior follows some patterns, and, thus, are amenable to a model.
+%Volunteer computing is extensively used in science and industry, yet it still has not become a mainstream technology
+%in the same way the cloud technologies in general are. The lack of general purpose and open source tools with built-in
+%analytics could be one of the reasons of why this is so, as well as the lack of proofs of concept.
+
+Using the browser as a computing resource in volunteer computing has presented several advantages, 
+but it remains a challenge to fully harness both the users’ behaviour and the browser’s capabilities.
+That is why we have enhenced {\sf NodIO}, a framework used to create client-server architectures 
+for distributed evolutionary algorithms, resulting in {\sf NodIO-W$^2$}, which we present in this paper.
+{\sf NodIO-W$^2$} leverages the HTML5 Web Workers specification to execute in the background several clients per
+browser, so that the the number of generations per user could be improved. The enhancements on the NodIO architecture,
+along with some initial results on the expected performance and scaling behavior when the number of users is increased,
+are introduced. We also use the algorithm data to find out what kind of patterns the user behavior follows by examining 
+the resources they devote to our experiment, and try to fit them in some statistical distribution. In general, 
+our experiments prove, first, that {\sf NodIO} is a flexible and low-overhead platform to perform distributed computation
+experiments that include volunteers, and then that user behavior follows some patterns, and, thus, are amenable to a model.
 % Volunteer computing is extensively used in science and industry, but

### 2016-01-09 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing @mariosky 's changes, fixing a typo I left yesterday. Abstract is now 187 words so this closes #85

@@
 but it remains a challenge to fully harness both the users’ behaviour and the browser’s capabilities.
-That is why we have enhenced {\sf NodIO}, a framework used to create client-server architectures 
+That is why we have enhanced {\sf NodIO}, a framework used to create client-server architectures 
 for distributed evolutionary algorithms, resulting in {\sf NodIO-W$^2$}, which we present in this paper.

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing the abstract
And adding the note to the response. Refs #85

@@
 That is why we have enhanced {\sf NodIO}, a framework used to create client-server architectures 
-for distributed evolutionary algorithms, resulting in {\sf NodIO-W$^2$}, which we present in this paper.
+for distributed evolutionary algorithms, resulting in {\sf
+  NodIO-W$^2$}, whose results we present in this paper.
 {\sf NodIO-W$^2$} leverages the HTML5 Web Workers specification to execute in the background several clients per
@@
 our experiments prove, first, that {\sf NodIO} is a flexible and low-overhead platform to perform distributed computation
-experiments that include volunteers, and then that user behavior follows some patterns, and, thus, are amenable to a model.
+experiments that include volunteers, and then that user behavior
+follows some patterns, and, thus, are amenable to a model and thus predictable.
 % Volunteer computing is extensively used in science and industry, but

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Revise authors
And delete index terms as the instructions in the PDF say. Closes #84.
Starting the cuts
I'll be revising on the fly but please @mariosky, @unintendedbear, @pacastillo and @fergunet check that there is no reference to what has gone to the other paper, missing links, whatever. Refs #86

@@
 % for Low Cost Distributed Evolutionary Computation
-
-% Some keywords missing: Cloud, Asynchronous, Pool-Based, Javascript-Based
-% Open Source software.
-% Lets not forget this issue is about software.
 
@@
 of Statistics and Operational Research, University of Granada, Granada, Spain%
-%\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
-%\thanks{Manuscript submitted for review on \today.}%
-%\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
-%\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
-%\thanks{$^3$Dept. of Statistics and Operational Research University of Granada}%
-%\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
-%    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es},{\tt nrico@ugr.es}}%
+
 \thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
@@
 
-\begin{IEEEkeywords}
-Evolutionary algorithms, evolutionary computation, distributed computing, internet computing,
-social networks, volunteer computing, metacomputing, performance evaluation.
-\end{IEEEkeywords}
-
 %---------------------------------------------------------------
@@
 used; for instance, Java was pointed out in \cite{soares1998get} as a
-``language for the
-internet'', providing some advantages such as multi-architecture compatibility or
+``language for the internet'', providing some advantages such as multi-architecture compatibility or
 security mechanisms, and was proposed in \cite{chandy1996world} as the
@@
 The algorithms used, as well as the methodology 
-for gathering resources will be described next, 
+for gathering resources, will be described next, %change this to
+                                %references to the other paper - JJ
 together with the results obtained in this initial setup.
-%---------------------------------------------------------------
-% Start of content for other paper ----
-\section{Description of the framework}
-\label{sec:description}
-
-In general, a distributed volunteer-based evolutionary computation
-framework based on the browser is simply a client-server system
-whose client is, or it can be, embedded in the browser via
-JavaScript. Since JavaScript is  the only language that is present
-across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
- is
-why we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
-faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
-in evolutionary computation and numerical optimization such as Python. 
-% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
-Compiled languages such as
-Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
-considering the performance of a single-user, 
-but this is more than offset by the computational power of
-the spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets. Together, the performance is several orders of magnitude
-higher, which is the objective in this kind of systems.
-% (Paloma) This sentence needs re-writing
-% "Other compiled languages such as Java or C might be, in fact,
-% faster, but the fact that single-user performance is higher (than
-% what?) is over the offset. This is because the number of spontaneous
-% volunteers we can gather at whim, including people using their
-% mobile phones or tablets, is several orders of magnitude higher,
-% which is the objective in this kind of systems." (?) 
- JavaScript is
-quite a popular language nowadays, providing compilers from a number
-of languages, many of them in their family (like CoffeeScript), but
-many other outside, so researchers can, in fact, write their fitness
-function in Python, Erlang, Scala, or even Java, and then compile it to
-JavaScript. For completeness, an experiment comparing the performance of
-a JavaScript implementation versus Java and Matlab is included in section
-\ref{sec:w2}.
-
-As a volunteer based system, there are some issues worth 
-noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
-\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
-each request is known.
-\item As anonymous entities, volunteers are not accountable. 
-If a volunteer misbehaves in some way the provider cannot 
-prosecute or discipline the volunteer. In the case of JavaScript,
-this issue is aggravated because even if it is obfuscated, source code
-and data is sent to clients. 
-\item Participants must also trust the application provider, 
-regarding invasion of privacy, the integrity of their devices, 
-and intended application. 
-\end{enumerate}
-
-Several kind of threats are possible in this environment; first, since its
-programming interface is open and its code is open source, it is
-relatively easy to find vulnerabilities and 
-sabotage the system, as indicated in \cite{domingues2007sabotage}, by
-crafting a fake request which, for instance, assigns a fake fitness to
-a particular chromosome.
-Other two different kinds of attacks are also
-possible: a denial of service attack as well as penetration
-testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are openly released
-as free software. 
- We also released as an open access paper the result of the 
- experiments performed \cite{DBLP:journals/corr/GuervosG15}  
- and all data, once anonymized, as open data. This
-builds trust between the user and the scientist, which indeed has an
-impact on the performance of the system, since there is no need to
-include cheating checks or other functions that would degrade it. This
-socio-technical approach is, indeed, 
-coherent with our whole approach to the study of a distributed volunteer
- system which can only be tackled by approaching it from the
-technical as well as the social, as in social media, angle.
-
-In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
-cloud or bare metal based volunteer evolutionary computing system
-derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
- the {\sf NodEO} library, whose architecture has been
-developed using JavaScript on the client as well as the server.
-All parts of the framework are free and available with a free license
-from \url{https://github.com/JJ/splash-volunteer}.
-
-%PABLO: el siguiente párrafo justifica Javascript
-
-Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
-\item A REST server, that is, a server that includes several {\em
-  routes} 
-  that can be called for storing and retrieving information (the ``CRUD'' cycle:
-  create, request, update, and delete) from the server. 
-  A JSON data format is used for the communication between 
-  clients and the server. There are two kinds of information:
-  {\em problem} based, that is, related to the
-  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
-  GET}ing a random chromosome from it, and {\em information} related
-  to the performance and state of the experiments. It also performs logging
-  duties, but they are basically a very lightweight and high performance
-  data storage \cite{jj:idc:lowcost}.
-  The server has the capability to
-  run a single experiment, storing the chromosomes in a data structure
-  that is reset when the solution is found.
-\item A client that includes the evolutionary algorithm as
-  JavaScript code embedded in a web page that displays graphs, some
-  additional links, and information on the experiment. This code runs
-  an evolutionary algorithm {\em island} starting with a random
-  population, then it sends, every 100 generations, the best individual
-  back to the server (via a {\tt PUT} request), and requests a random
-  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
-\end{enumerate}
-
-  This is the invariant part of the framework, but other than that,
-  the algorithm can be run in many different ways: 
-  stopping when a solution is found or, as we have done in this paper,
-  using Web Workers. In fact, since it is a pool-based system such as
-  the one described in \cite{LNCS86720702}, any kind of client that
-  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
-But never heard of a client which "works" an API.
-Is this correct?
-  written in any kind of language. However, in this paper we are going
-  to focus on the dynamics and measurements of the volunteer part of
-  the framework.
-
-The figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or {\tt
-  Chart.js} have
-been used to build the user interface elements of the framework.
-\begin{figure}[!t]
-\centering
-\includegraphics[width=4in]{system.pdf}
-\caption{Description of the proposed framework. Clients execute a JavaScript EA
-  in the browser, which, every 100 generations, sends the best
-  individual and receives a random one from the server.}
-\label{fig:system}
-\end{figure}
-
-The researcher only has to change a function, which can be written in
-JavaScript or in any other language with cross-compilation to
-JavaScript \cite{web:compilersjs} to solve a different
-problem. In this case the classical Trap function \cite{Ackley1987} has been
-used as well as the Rastrigin's floating-point optimization
-problem. JavaScript is a functional language and declaring a different 
-function and handing at the creation of the algorithm object, called
-{\tt Classic}, is the only change needed to work with a different
-problem. Let us see how this system addresses the different challenges
-outlined in the State of the Art (Section \ref{sec:soa}):
-\begin{itemize}
-\item {\em Scalability} is provided via the use of a lightweight and
-  high-performance, single-threaded, server based in Node.js and
-  Express.js. Although this single server is a bottleneck since it
-  will eventually saturate, the fact that it runs as a non-blocking single thread
-  allows the service of many requests. In fact, a limit in the
-  number of simultaneous requests will be reached, but so far it has
-  not been found, unlike what we found in our previous systems, DCoR,
-  \cite{gecco07:workshop:dcor}, which had a low scalability. 
-\item The framework is {\em heterogeneous} since it does not need any
-  performance, operating system, or even browser requirement as long
-  as JavaScript is enabled: anyone
-  visiting the page, even from mobile devices, can load the algorithm.
-  In case the browser does not support HTML5 Web workers (e.g., 
-  Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of {\sf NodIO} can also be used.
-\item {\em Fault tolerance} is always an issue, and in this case, the
-  single point of failure would be the server: the system, as such,
-  would break down if the server fails. However, the individual
-  islands in every browser would continue running, and having access
-  to just one of them would allow the local algorithm to proceed. In
-  fact, the island does not need the server to run: it runs locally if
-  needed, with the only exception that it is obviously unable to
-  communicate with the rest of the islands.
-\item {\em Adaptiveness} is achieved simply through the autonomous
-  operation of every individual island without any synchronization
-  mechanism. The islands in the system are, in fact, unaware of each
-  other, communicating only through the server. This architecture 
-  has been tested in other high demand systems with success, but 
-  in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
-  %(Mario) Scalability of the communication system in particular?
-\item Since the algorithm runs in the browser, user's {\em safety} is
-  achieved through its sandbox mechanisms. The user is thus assured
-  that there is no unsafe access either to their local files or even
-  to more resources than the browser should be allotted.
-\item Running the algorithm is just a matter of loading the page,
-  which makes the operation totally {\em anonymous}. For the same
-  reason, {\em ease of use} is optimal, being as easy as simply
-  clicking on an URL, available to anyone with access to a browser.
-\item {\em Reasonable performance} is not ensured. In fact, we should
-  make sure that there is a reasonable amount of clients over which
-  the performance achieved is better than what you would obtain in
-  your own desktop system. If this is not the case, it is a pointless
-  academic exercise.
-\end{itemize}
-
-There are many different ways to validate a framework that intends to
-address these challenges. Some of them are included by design: it is
-anonymous, since no accounts are needed to participate in the
-experiment, it is safe for the user, since we are using the browser
-black box, and it is also heterogeneous. We have designed the
-experiments so that they show adaptability to different types of
-experiments and clients, scalability with the number of clients and
-the problem size, and measured performance by comparing it with
-a baseline configuration, with which, as indicated above, is
-compatible and can be used in a complementary way. Fault tolerance
-will be experimented by dropping the server and checking the continuity
-of the experiments.
-
-All these points will be established in the next two sections,
-starting with the baseline, single worker system next.
-
-%---------------------------------------------------------------
-\section{Baseline Experiments} %Cambié el nombre
-\label{sec:experiments}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=10cm]{baseline-times.png}
-\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
-    Node.js} client, for different population sizes. Only
-runs in which solution has been found are used for this graph.}
-% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
-\label{fig:baseline}
-\end{figure}
-First, we establish the baseline performance by running fifty runs of the
-evolutionary algorithm in a desktop client written using {\sf NodEO}
-\cite{nodeo2014}, the basic JavaScript library we have used as base to
-build {\sf NodIO}, the framework. This experiment tries to 
-find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$ and a population size equal to 512. The algorithm in a
-single island was run until the solution (a string with all
-ones) was found or five million evaluations had been performed. It
-took around a minute, on average, that is $68.9694$ seconds, to
-perform these runs. In this experiment only 33, that is, 66\% were
-successful. The experiments were repeated for population $p=1024$,
-with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
-seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
-characteristics: 
-\begin{verbatim}
-Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
-2014 x86_64 x86_64 x86_64 GNU/Linux
-\end{verbatim}
-with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
-
-% end spin-off ------------------------------------------------------
+
+\section{Experimental results}
+\label{sec:exp1}
 

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adapting general intro...
... and intro to section, previous section. Refs #86.

@@
 state of the art (Section \ref{sec:soa}) in web-based distributed
-computational systems along with attempts to predict and model its behavior. 
-Then, a detailed description of the proposed framework is presented
-in Section \ref{sec:description}, followed by the baseline experiments
-in Section \ref{sec:experiments}, and experiments using a new client
+computational systems along with attempts to predict and model its
+behavior. The first results in a volunteer setting are described
+in Section \ref{sec:exp1}, and experiments using a new client
 architecture that uses web workers in Section \ref{sec:w2}.
@@
 The algorithms used, as well as the methodology 
-for gathering resources, will be described next, %change this to
-                                %references to the other paper - JJ
-together with the results obtained in this initial setup.
+for gathering resources and baseline results that we will use for
+comparison have been described in
+\cite{2016arXiv160101607M}. We will focus next on the experimental
+results using the first version of the framework.
 
@@
 
-These two experiments establish a baseline result and show that
-the time to success depends on the population size, with a bigger population
-contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+Initial results using the same code base and a single computer have
+been described in \cite{2016arXiv160101607M}. They prove that a bigger population
+contributes to diversity and thus speeds up the solution
+\cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing 
 experiments that we will describe next do not and can not have the
@@
 the baseline is that if they eventually take longer than a basic
-desktop, their interest will be purely academic. We will try and
+desktop, their interest will be purely academic, although they can
+complement them. We will try and
 prove next that that target performance can be achieved by carrying

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reducing a bit names using acronyms refs #84

@@
 
-\author{Juan-J.~Merelo, Dept.
-of Computer Architecture and Technology, University of Granada, Granada, Spain\\Mario Garc\'ia-Valdez, Dept.
-of Graduate Studies, Instituto Tecnológico de Tijuana, Tijuana, M\'exico\\Pedro A.
-Castillo, Dept.
-of Computer Architecture and Technology, University of Granada, Granada, Spain\\Pablo Garc\'ia-S\'anchez, Dept.
-of Computer Architecture and Technology, University of Granada, Granada, Spain\\Paloma de las Cuevas, Dept.
-of Computer Architecture and Technology, University of Granada, Granada, Spain\\Nuria Rico, Dept.
-of Statistics and Operational Research, University of Granada, Granada, Spain%
+\author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
+  (CAT), University of Granada, Granada, Spain (UGR)\\
+Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
+Tecnológico de Tijuana, Tijuana, M\'exico\\
+Pedro A. Castillo, Dept. of CAT, UGR\\
+Pablo Garc\'ia-S\'anchez, Dept. of CAT, UGR\\
+Paloma de las Cuevas, Dept. of CAT, UGR\\Nuria Rico, Dept. of Statistics and Operational Research, UGR%
 

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Removing second chunk
And adding a paragrah with a very short resume.

@@
 
-% -- Start spin-off -------------------
-\section{Enhancing {\sf NodIO} with a Web Worker-based architecture}
-\label{sec:w2}
-
-In order to improve the number of cycles per user, we enhanced the
-basic architecture in several different aspects.  One improvement 
-involved restarting the client once the solution was
-found, so that as long as the user was visiting the page it was
-running an evolutionary algorithm. Another was a small algorithmic
-enhancement: instead of having a fixed population,
-population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvement was to 
-use the Web Worker API as described next.
-
-JavaScript has been commonly used to develop client-side scripts
-that handle the user interface, communicating asynchronously with servers and
-updating the content that is displayed \cite{flanagan2006javascript}.
-Long-running scripts like those proposed in this work can interfere with the
-responsiveness of the web application, because everything is handled by a
-single thread. The Web Workers specification \cite{hickson2012web} defines an
-API that allows Web developers to execute long-running scripts in parallel
-with the scripts of the main page. Worker scripts are spawned in the
-background allowing a thread-like operation using message-passing as the
-communication mechanism between the worker's execution environment and the
-main thread. According to the Web Workers W3C Candidate Recommendation
-\cite{hickson2012web} workers are expected to be long-lived, they have a high
-start-up performance cost, and a high per-instance memory cost. The Web Worker
-API is implemented in current versions of both desktop and mobile web browsers.
-For volunteer computing using the Web Worker API brings several advantages
-over a single threaded implementation:
-
-\begin{itemize}
-\item If the browser uses a tabbed document interface, the worker script
-keeps running in the background even in case the user brings forward another tabbed
-window.
-\item Several evolutionary algorithms can be executed in parallel in a single web
-page. 
-\item Many implementations of the Web Worker API can use multiple-core CPUs for
-the execution of worker scripts, in particular the execution in multiple-core CPUs
-was tested in desktop versions of Chrome, FireFox and Safari.
-\item Each {\sf NodEO} instance can be restarted independently.
-\end{itemize}
-
-A sequence diagram of the enhanced NodIO, which is called NodIO-W$^2$, is presented
-in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
-API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
-interaction between processes in a time sequence. In this version there are
-two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
-the client and the server where PUT and GET HTTP methods are represented by
-red and blue arrows respectively. There are also messages between workers and
-the main thread locally in the client, which are shown as thin solid arrows.
-There are three processes involved:
-\begin{itemize}
-\item {\em Node.js} This is the server side process which is basically
-  the same as used previously, except for some minor log-related
-  changes. 
-\item {\em Client main thread} The main script creates
-worker instances and handles the user interface.
-\item {\em Worker global scope} The worker execution space does not have
-access to the Document Object Model (DOM) or JQuery objects in the main thread.
-Asynchronous HTTP requests to the server are implemented using the
-XMLHttpRequest object.  
-\end{itemize}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=4in]{Algorithm.pdf}
-\caption{Description of the W$^2$ version of {\sf NodIO}. In this
-  version, clients use Web Workers to run the evolutionary algorithm,
-  with two of them per browser.}
-\label{fig:system:w2}
-\end{figure}
-%
-% End spin-off ----------------------------------------------------------------
 \begin{table*}[!htb]
@@
 \end{table*}
-
-% - Start spin-off again
-% -------------------------------------------------------
-
-The sequence of events happens as follows:
-\begin{enumerate}
-\item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure, the sequential number of experiments is initialized and a log
-file is created. The shared pool implemented as an array is reset.
-\item A volunteer follows the link of the experiment triggering a GET HTTP
-request, then the page is loaded.
-\item The main script renders the page and creates the worker instance. In
-order to create a dedicated worker instance, another GET request is sent to the
-server, then the client receives the second worker implementation script.
-Each worker instance executes over time one or more evolutionary algorithms
-{\em islands}. Each island starts with a random population and it is assigned a
-universally unique identifier (UUID). The UUID is included in each of the HTTP
-requests sent to the server.
-The initial parameters for each island
-can be distinct. Once the worker is created and the parameters have been set,
-a message is sent to the worker to start the algorithm.
-\item The worker first creates the population following the previously
-specified parameters. Once the population has been initialized, the EA
-generation loop is started. After {\em n} generations the state of the population
-is evaluated. The chromosome with the best fitness is sent to the shared
-population with a PUT HTTP request. If the best fitness is not the global best,
-the EA loop must continue, but first a chromosome is randomly selected from
-the shared population. The current generation number and the best
-chromosome are also sent to the client's main thread in order to update the DOM to
-display the current state of the island (described next). These messages are
-all asynchronous, so the EA loop of
-the worker continues even before a message is received by the target. If the best fitness is a global best, the best
-chromosome is sent to the server and the loop ends. A message is sent to the
-main thread indicating that the best individual has been found.
-\item The main thread receives messages from workers as events handled by
-a callback function. When an iteration message is received, the page is
-updated by displaying a dynamic plot of the generation number and best
-fitness, and a representation of the chromosome is also generated.
-\item When a global best is received from an island, the current experiment
-ends, the experiment number is incremented, and the population array is reset.
-\item When the main thread receives a message indicating that a worker has ended the EA
-loop, that worker is reinitialized. The worker process is not ended, to
-avoid the start-up cost, only the parameters and population are reset and a new UUID
-assigned.
-\end{enumerate}
-% End spin-off--------------------------------------------
-
+%
 \begin{figure}[!htb]
@@
 
-\subsection{First experiments}
+\subsection{Experiments using web workers}
 \label{sec:w2Exp}
 
+The results reported above indicated that we should look for a way of
+increasing the time that every user devotes. This led us to create an
+architecture that uses Web Workers, which is described in
+\cite{2016arXiv160101607M}; web workers work in the background even if
+the web page is not in focus. A few improvements were also tested,
+basically using a random population and restarting the web worker when
+the solution was found.
+
 The experiments were run in the same way as before, using several
-Twitter announcements throughout several days by the end of July
+ announcements in Twitter throughout several days by the end of July
 2015. Eventually, more than one thousand experiments were completed in

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Last segment eliminated
Refers to #86 and adds some glue paragraphs.

@@
 in Section \ref{sec:exp1}, and experiments using a new client
-architecture that uses web workers in Section \ref{sec:w2}.
+architecture that uses web workers in Section \ref{sec:w2} and then
+applied to a different problem in Section \ref{sec:rastrigin}. 
 Finally, conclusions and future lines of work are presented in Section
@@
 
-% ------------Start spin-off ----------------------------------------------------
-\subsection{Tests with a floating-point hard optimization problem}
-
-JavaScript has been traditionally implemented as an interpreted
-language not designed for the development of high performance
-systems. On the other hand, current JavaScript  
-virtual machines (VMs from now on) are closing the gap, for instance the Google V8 
-is an engine specifically designed for the fast execution of 
-large JavaScript applications. In order to increase performance
-V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter \cite{Gray:2009:GCM:1610564.1610565}.
-JavaScript has gained popularity in server-side development using runtime
-environments such as Node.js and is also used in desktop and mobile applications. 
-%Mario: Creo que el párrafo de arriba sale sobrando
-% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
-To further evaluate how a current JavaScript implementation could be used by a
-computer scientist to develop complex optimization problems,
-a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization \cite{tang2007benchmark} is used in this section. 
-The function is described next.
-
-The basic Rastrigin's function is separable, and is defined as follows;
-\begin{equation}
-F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
-\end{equation}
-where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
-$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
-is a classical multimodal problem. Such problem is difficult since the number of local
-optima grows exponentially with the increase of dimensionality. To make it 
-non-separable, an orthogonal matrix is also used for coordinate rotation.
-The rotated Rastrigin's function is defined as follows:
-\begin{equation}
-F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
-\end{equation}
-Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
-emulates real-world optimization problems that most likely will consist 
-of different groups of parameters with strong dependencies within
-but little interaction between the groups. This issue is reflected in the benchmark 
-by randomly dividing the objective variables
-into several groups, each of which contains a number of variables. The parameter $m$ 
-is used to control the number of variables in each group and hence, defining the degree
-of separability:
-\begin{equation}
-F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
-\end{equation}
-For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
-candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
-$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
-finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
-
-This optimization problem was selected because it is representative of the kind of
-algorithms, data types, and structures employed in large scale optimization problems. 
-The function was implemented in both Matlab and Java languages as part of the
-Test Suite of the companion competition of the Special Session and to give competitors an idea of the
-computational cost of the challenge, the runtime required for 10,000 function evaluations 
-for a particular configuration was published by organizers. The Java
-implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
-3,000,000 function evaluations as termination condition. Using the Java implementation as
-a guide, the function described above was implemented to compare the performance of JavaScript against
-both implementations, again using the time required for 10,000 function evaluations. 
-Before discussing the setup and performance results, certain details of
-the implementation in JavaScript are presented to give readers an idea of 
-how practical and mature this language is for developing complex mathematical functions.
-
-\begin{itemize}
-\item {\em Randomize} Both Matlab and Java programs rely on a 
-Java Randomizer library for the 
-randomization of the shift vectors and matrices used by the function.
-The library includes the generation of pseudorandom Gaussians of type
-double and uniformly distributed integers. Generators use a single
-{\tt long} seed. In order to implement the same functionality in
-JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
-in the implementation of the standard {\tt Math.random()} between engines
-and also their results are non-deterministic. 
-The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
-consistent results across all JavaScript VMs.
-\item {\em Timing functions} In order to compare the performance of 
-each implementation, an accurate and consistent way of measuring 
-the runtime of functions is needed.
-In JavaScript the {\tt Date class} is often used 
-to measure execution time, but has the limitation of having a maximum 
-resolution given in milliseconds.
-In order to measure the runtime intervals with a higher precision two functions
-where used: for Node.js we used the native {\tt process.hrtime()} function
-which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
-array, and for measuring in the browser the {\tt Performance.now()} function was called,
-which uses floating-point numbers with up to microsecond precision. 
-A drawback found is that the {\tt Performance.now()} function is only implemented
-in Firefox and Chrome browsers. Both functions are independent of the system clock. 
-\item {\em Tools} An important consideration when choosing certain 
-language to write non-trivial programs is the availability of development tools. While developing
-the experiments, we used a comprehensive set of developer tools available for both
-browsers and desktop \cite{tilkov2010node}. These tools included
-package managers, in this case {\tt npm}, debuggers, logging libraries
-such as {\tt winston}, 
-network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
-Although the fact that developer tools vary from browser to browser
-supposed a drawback, they all have the same functionality. The
-JavaScript debugger can be accessed directly from the user menu in
-most browsers. 
-\item {\em Data Types} JavaScript uses floating point numbers with a 
-limited precision of 64 bits that partially implement the functionality of
-the {\tt StrictMath} library used by the Java implementation. If more precision
-is needed, developers could use {\tt math.js}, an extensive math library with
-support for matrices and big numbers.
-\end{itemize}
-
-We found that the JavaScript ecosystem is mature enough for developing non-trivial 
-code, but developers have to consider the differences between VMs implementations
-and therefore use certain libraries in order to increase the precision and repeatability of
-algorithms. As for the performance of JavaScript against Java and Matlab
-we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
-Intel Xeon E5 processor running OS X 10.10.5, 
-using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
-MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
-Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
- The performance of the time required for 10,000 function evaluations is presented in
-Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
-performance with an average of 935ms, followed by Java with
-991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
-single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
-The results also show that there is not much difference between the Node.js
-implementation an that of a
-browser. There is also not much difference between running the code in the main thread or in Web
-Workers. When running the experiment in Firefox a dialog alerted that the
-script was taking too much time, showing that this
-kind of processing is not intended for the main thread. Additionally, there is not much overhead
-when running two
-Web Workers in parallel as they took 1279ms each. This experiment
-shows that JavaScript is viable as a language
-language for developing complex mathematical functions an these can run in a browser
-with acceptable performance. 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{f15_times.png}
-\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
-\label{fig:f15_times}
-\end{figure}
-% - end spin-off-----------------------------------------------------------------
-
-After setting up baseline results for this particular problem, we will
-present now results the results obtained. Next we will present the
-details of the implementation and then the numeric results of the
-experiment.
-
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}
-
-This experiment was deployed in a virtual private server with 512MB of memory
+\section{Working on a hard optimization problem}
+\label{sec:rastrigin}
+
+The fact that the solution of this particular problem was found in a
+few seconds and restarts were necessary yielded interesting results,
+but we needed to work on a harder problem, that is why created a new
+client version that optimized Rastrigin function, a multimodal
+floating-point problem used as a benchmark in evolutionary
+algorithms. The set up for this function, and a comparison of the
+baseline speed for JavaScript in an standalone experiment and other
+languages is shown in \cite{2016arXiv160101607M}, where the details of
+the implementation have also been extensively described.
+
+Specifically, this experiment was deployed in a virtual private server with 512MB of memory
 running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
@@
 chromosome with the server every 100 generations. Again, there were two workers in each window.
-As before, a few participation calls were published in social networks.
-
-\subsubsection{Results}
+As in previous experiments, a few calls to participation were published in social networks.
 

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Edited Intro: Is NodIO, presented in this paper?

@@
 
+%Mario: Is NodIO presented here after the cut? 
+%TO DO: Add reference
 There is, then, space for frameworks that do not make any of these
-assumptions, such as the one presented in this paper, which will be
- named {\sf NodIO}. {\sf NodIO} is basically
+assumptions, such as {\sf NodIO} presented in \cite{}. {\sf NodIO} is basically 
+%assumptions, such as the one presented in this paper, which will be
+% named {\sf NodIO}. {\sf NodIO} is basically
 % (Paloma) Is the "first" word really needed? 
@@
 In this paper, our intention is to
-present the {\sf NodIO} framework and analyze it from two different
+analyze the {\sf NodIO} framework from two different
+%Mario: Edit this?
+%present the {\sf NodIO} framework and analyze it from two different
 angles: the purely technical angle, including the decisions that went

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Getting down to 12 pages.
Eliminated 1 reference

--- 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
 the aforementioned qualities called {\em Charlotte}
-\cite{baratloo1996charlotte}. This framework, which implemented a
+\cite{baratloo1999charlotte}. This framework, which implemented a
 distributed shared memory, was inspired by a previous work
@@
 Other issues such as scheduling, which are related to adaptive 
-parallelism, were solved on the server. This
-paper was later extended and published in a journal in 1999
-\cite{baratloo1999charlotte}, but it is interesting to note that,
+parallelism, were solved on the server. It is interesting to note that,
 even in the early stages of the web, the possibility of
@@
 \subsection{Experiments using web workers}
-\label{sec:w2Exp}
+\label{sec:w2}
 
@@
 %
-\begin{figure}[!htb]
-\centering
-\includegraphics{rastrigin-fitness.png}
-\caption{Fitness of chromosomes sent to the server pool along time in
-  the $x$ axis.} 
-% Plotted with ../data/plot-rastrigin-fitness.R
-\label{fig:puts:rastrigin}
-\end{figure}
+% \begin{figure}[!htb]
+% \centering
+% \includegraphics{rastrigin-fitness.png}
+% \caption{Fitness of chromosomes sent to the server pool along time in
+%   the $x$ axis.} 
+% % Plotted with ../data/plot-rastrigin-fitness.R
+% \label{fig:puts:rastrigin}
+% \end{figure}
 %
@@
 running, instead lowering, little by little, the fitness until it was
-disconnected. The fitness of chromosomes sent by clients is shown in
-Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
-upper part, which represents the fitness of clients that have just
-started evolution from a random population, and then the fitness of
-those clients that have been already evolving for a while or obtained
-a high-fitness individual from the pool and evolved from there, and
-which managed to surpass the 9000 value by the end of the
-simulation. It incidentally shows also the number of {\em new} nodes
-coming: there is a dense {\em cloud} at the beginning, but the new
-contributions keep thinning as the evolution progresses.
+disconnected. The fitness of chromosomes sent by clients managed to
+surpass the 9000 value by the end of the; and although initially there
+was a high number of new nodes, this number decreases with time.
 
@@
 number of unique IPs participating in the experiment per minute (left)
-and hour (right). The two graphs have the same shape which matches
-also the speed in which fitness is decreased in Figure
-\ref{fig:puts:rastrigin}. At the beginning there are very few or only
+and hour (right). The two graphs have the same shape: At the beginning there are very few or only
 one, but later one there are peaks of almost 30 unique machine a

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Permission to cut this sections in SOA
Fully removed all material in the other paper
And additionally reduced content to 12 pages to comply with requirements.

@@
 The fulfillment of these properties could be achieved by using any of the
-languages and associated technologies, mentioned above, provided that the user had
+languages and associated technologies mentioned above, provided that the user had
 installed the required virtual machine plugin first. 
@@
 in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:runs}. In general, every experiment took several
+Table \ref{tab:summary:os}. In general, every experiment took several
 days. No particular care was taken about the time of the announcement
@@
 \begin{table*}
-\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
-\begin{center}
-\begin{tabular}{l|rr}
-\hline
-Date & \# experiments & Different IPs \\
-\hline
-April 4th 4/4 & 57 & 191 \\
-April 24th 4/24 &  231 & 559 \\
-July 31th 7/31 & 97 & 179 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{table*}
 \caption{Summary of time per run, number of IPs and number of {\tt
@@
 \begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
-4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
-7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
+\begin{tabular}{l|ccccccccc}
+\hline
+Date & Experiments &\#IPs &\#IPs &  \#IPs & Time (s) &  \#{\tt PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment\\
+     & \# & Different & Median & Max & Median &  Median & & & correlation \\ 
+\hline
+4/4 & 57 & 191 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
+4/24 & 231 & 559 & 5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
+7/31 & 97 & 179 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
 \hline

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

@@
 Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers. This paper % (Paloma) No really need for that "also", plus it complicates reading
+metacomputers. 
+%% We need to only reference this section?:
+%%------------------- BEGIN
+This paper % (Paloma) No really need for that "also", plus it complicates reading
 establishes the desirable features of
@@
 % thanks :-) -JJ
-
+%%------------------- END
+%% EDIT:
+%%------------------- BEGIN
 The fulfillment of these properties could be achieved by using any of the
@@
 installed the required virtual machine plugin first. 
+%%------------------- END
 % (Paloma) These should be separate sentences

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
SoA edited
Deleted section of desirable features described in the other paper.

@@
 %%------------------- BEGIN
-This paper % (Paloma) No really need for that "also", plus it complicates reading
-establishes the desirable features of
-any {\em global computing} infrastructure, that can also be applied to
-distributed evolutionary computation experiments that rely on
-volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
-reasonable performance. But hierarchy, this is, the capability of
-preferentially using resources in the same organization, is a feature
-that could be either disregarded or included within adaptivity, which
-leaves us with eight desirable properties. 
-% (Paloma) Re-write. For example, "But is hierarchy, this is, the
-% capability of preferentially using resources in the same
-% organization, the one that could be either disregarded or included
-% in adaptivity, which leaves us with eight desirable properties." 
-% (Paloma) And I re-wrote what I proposed yesterday.
-% thanks :-) -JJ
-%%------------------- END
-%% EDIT:
-%%------------------- BEGIN
-The fulfillment of these properties could be achieved by using any of the
-languages and associated technologies mentioned above, provided that the user had
+% This paper % (Paloma) No really need for that "also", plus it complicates reading
+% establishes the desirable features of
+% any {\em global computing} infrastructure, that can also be applied to
+% distributed evolutionary computation experiments that rely on
+% volunteer computing: scalability, heterogeneity, fault tolerance,
+% adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
+% reasonable performance. But hierarchy, this is, the capability of
+% preferentially using resources in the same organization, is a feature
+% that could be either disregarded or included within adaptivity, which
+% leaves us with eight desirable properties. 
+% % (Paloma) Re-write. For example, "But is hierarchy, this is, the
+% % capability of preferentially using resources in the same
+% % organization, the one that could be either disregarded or included
+% % in adaptivity, which leaves us with eight desirable properties." 
+% % (Paloma) And I re-wrote what I proposed yesterday.
+% % thanks :-) -JJ
+% %%------------------- END
+% %% EDIT:
+% %%------------------- BEGIN
+% The fulfillment of these properties could be achieved by using any of the
+% languages and associated technologies mentioned above, provided that the user had
+% installed the required virtual machine plugin first. 
+% %%------------------- END
+The fulfillment of the desirable features of any distributed evolutionary 
+computation experiments that rely on volunteer computing \cite{2016arXiv160101607M} could be achieved by using any of the
+languages and associated technologies, mentioned above, provided that the user had
 installed the required virtual machine plugin first.
-%%------------------- END
+
 % (Paloma) These should be separate sentences

### 2016-01-10 data/time-vs-IPs-openshift.R
Better fig 1 refs #89

@@ -7,7 +7,7 @@ time.vs.IPs <- function( data, preffix ) {
         data[i,]$howmany <- length(data[data$IPs==data[i,]$IPs,]$milliseconds)
     }
     data$i <- 1:nrow(data)
-    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="white",high='gray70')+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=20))
+    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="#11AA99",high='#EE3333')+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=30), axis.text.y = element_text(angle=90, vjust=0.5)) 
     ggsave(paste0("../img/time-vs-ips-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
     ggsave(paste0("../img/duration-vs-time-OS-",preffix,".png"))

### 2016-01-10 data/time-vs-IPs-openshift.R
Better fig 2 refs #89

@@ -13,7 +13,7 @@ time.vs.IPs <- function( data, preffix ) {
     ggsave(paste0("../img/duration-vs-time-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
     ggsave(paste0("duration-vs-time-OS-bar-",preffix,".png"))
-    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='white', high='black') + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), text = element_text(size=20), panel.grid.major = element_blank(), panel.grid.minor = element_blank() )+labs( x="Ordered experiments")
+    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='#11AA99', high='#EE3333') + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.y = element_text(angle=90, vjust=0.5) )+labs( x="Ordered experiments")
     ggsave(paste0("../img/time-vs-rank-OS-",preffix,".png"))
 }
 

### 2016-01-10 data/plot-zipf-openshift.R
Better fig 3 refs #89

@@ -12,7 +12,7 @@ process.puts <- function(suffix ) {
                                     this.fit$estimate["loc"],
                                     this.fit$estimate["scale"],
                                     this.fit$estimate["shape"]),decreasing=T))
-    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_grey(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=30), axis.text.y = element_text(angle=90, vjust=0.5))
     ggsave(paste0("../img/puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }

### 2016-01-10 data/time-vs-IPs.R
Better fig 4 refs #89

@@ -6,7 +6,7 @@ for ( i in 1:nrow(ips.time) ) {
     ips.time[i,]$howmany <- length(ips.time[ips.time$IPs==ips.time[i,]$IPs,]$milliseconds)
 }
 ips.time$i <- 1:nrow(ips.time)
-ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ scale_colour_gradient(name  ="Count", low="white",high='gray70')+ stat_summary(fun.data = "mean_cl_boot", colour = "black")
+ggplot(data=ips.time, aes(x=IPs,y=milliseconds,color=howmany))+geom_point()+scale_y_log10()+ scale_colour_gradient(name  ="Count", low="#11AA99",high='#EE3333')+ stat_summary(fun.data = "mean_cl_boot", colour = "black")
 ggsave("../img/ips-time-ww.png", width=5,height=5)
 ggplot(data=ips.time, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
 

### 2016-01-10 data/ips-puts.R
Better fig 5 refs #89

@@ -13,7 +13,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
 
 ggsave('../img/gev-fit-ww.png',width=5,height=5)
 
@@ -24,5 +24,5 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww$puts),
 y=sort(rweibull(length(ips.puts.ww$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_grey(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
[…]

### 2016-01-10 data/ips-puts-rastrigin.R
Better fig 5 refs #89

@@ -13,7 +13,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))+ theme(text = element_text(size=20), axis.text.y = element_text(angle=90, vjust=0.5))
 
 ggsave('../img/gev-fit-ww.rastrigin.png',width=5,height=5)
 

### 2016-01-10 data/ips-puts.R

@@ -13,7 +13,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))+ theme(text = element_text(size=20), axis.text.y = element_text(angle=90, vjust=0.5))
 
 ggsave('../img/gev-fit-ww.png',width=5,height=5)
 
@@ -24,5 +24,5 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww$puts),
 y=sort(rweibull(length(ips.puts.ww$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
-ggplot()+geom_point(data=ips.puts.ww.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_manual(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
[…]

### 2016-01-10 data/ips-puts-rastrigin.R
All graphics changed to color and bigger labels.
The palette comes from here https://l.facebook.com/l.php?u=https%3A%2F%2Fpersonal.sron.nl%2F~pault%2F&h=IAQFbQxkq recommended by @psicobyte. This refs #89 and, after revision, will close it. Please @unintendedbear or @fergunet check it out.

@@ -13,7 +13,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_grey(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit")) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))+ theme(text = element_text(size=20), axis.text.y = element_text(angle=90, vjust=0.5))
+ggplot() +geom_point(data=ips.puts.ww.rastrigin.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))+ theme(text = element_text(size=20), axis.text.y = element_text(angle=90, vjust=0.5))
 
 ggsave('../img/gev-fit-ww.rastrigin.png',width=5,height=5)
 
@@ -24,7 +24,7 @@ this.fit.weib <- data.frame(rank=1:length(ips.puts.ww.rastrigin$puts),
 y=sort(rweibull(length(ips.puts.ww.rastrigin$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
[…]

### 2016-01-10 data/ips-puts-rastrigin.R
Adds comments to editor's note

@@ -38,7 +38,7 @@ this.fit.plot <- data.frame(rank=1:length(ips.puts.ww.rastrigin.workers$puts),
                                 this.fit$estimate["scale"],
                                 this.fit$estimate["shape"]),decreasing=T))
 
-ggplot() +geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
+ggplot() +geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data')) +geom_point(data=this.fit.plot,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10() + scale_colour_manual(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))+ theme(text = element_text(size=20), axis.text.y = element_text(angle=90, vjust=0.5))
 
 ggsave('../img/gev-fit-ww-rastrigin-workers.png',width=5,height=5)
 
@@ -50,5 +50,5 @@ y=sort(rweibull(length(ips.puts.ww.rastrigin.workers$puts),
     this.weibull$estimate["shape"],
     this.weibull$estimate["scale"]),decreasing=T))
 print(this.weibull)
-ggplot()+geom_point(data=ips.puts.ww.rastrigin.workers.df,aes(x=rank,y=puts,color='Data',shape='Data'))+geom_point(data=this.fit.weib,aes(x=rank,y=y,color='Fit',shape='Fit'))+scale_y_log10()+ scale_colour_manual(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name ="Puts", breaks=c("Data", "Fit"), labels=c("Data", "Fit"))
[…]

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Better fig 1 refs #89
Better fig 2 refs #89
Final version, about to send
This has no major changes except for the paper split.
Last edits

@@
 Using the browser as a computing resource in volunteer computing has presented several advantages, 
-but it remains a challenge to fully harness both the users’ behaviour and the browser’s capabilities.
+but it remains a challenge to fully harness both the users' behaviour and the browser's capabilities.
 That is why we have enhanced {\sf NodIO}, a framework used to create client-server architectures 
@@
 along with some initial results on the expected performance and scaling behavior when the number of users is increased,
-are introduced. We also use the algorithm data to find out what kind of patterns the user behavior follows by examining 
+are introduced. We also use the experimental data to find out what kind of patterns the user behavior follows by examining 
 the resources they devote to our experiment, and try to fit them in some statistical distribution. In general, 
@@
 There is, then, space for frameworks that do not make any of these
-assumptions, such as {\sf NodIO} presented in \cite{}. {\sf NodIO} is basically 
+assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607M}. {\sf NodIO} is basically 
 %assumptions, such as the one presented in this paper, which will be
@@
 parallelized via population splitting or by farming out evaluations to all
-the nodes available. Lately, several approaches have focused on the
+available nodes. Lately, several approaches have focused on the
 fault-tolerance of volunteer algorithms
@@
 \end{figure}
-% ../data/time-vs-IPs-openshift.png 
+% ../data/time-vs-IPs-openshift.R
 %
@@
 \label{fig:zipf:os}
-% computed via ../data/time-vs-IPs-openshift.png 
+% computed via ../data/time-vs-IPs-openshift.R
 \end{figure}

### 2016-01-11 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Post-final changes
Under the request of the editor.

@@
   participating in it, with averages and standard deviation shown as
-  red dots; in the case there is a single red dot, there was a single
+  red dots; in the case there is a single dot, there was a single
   experiment in which many computers participated (for instance, 16
@@
 so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequency. 
+of computers happened less frequently, while darker shade or closer to
+redmeans more frequency. 
 From left to right, experiments 4/4, 4/24 and 7/31.}
@@
 \caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
-  Extreme Value distribution (in a lighter shade of gray). From left to right, experiments
+  Extreme Value distribution (in a lighter shade or blue). From left to right, experiments
   4/4, 4/24 and 7/31.} 
@@
   abscissas). The black dot and line show the average and standard
-  deviation, the shades of gray the number of experiments with the same
+  deviation, the blue to red shades the number of experiments with the same
   number of IPs. } 
@@
 \includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in light color, right) 
-and Weibull (in lighter shade, left) for the {\sf NodIO-W$^2$} model.}  
+\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in light
+  or blue color, left) 
+and Weibull (in lighter or blue shade, right) for the {\sf NodIO-W$^2$} model.}  
 \label{fig:gev:w2}
@@
 function, with the rest of the framework wrapped around
-automatically. All this, as well as the data used for this paper and the paper
-itself, has been published with a free license in GitHub at
-\url{https://github.com/JJ/modeling-volunteer-computing}.
+automatically.
+%  All this, as well as the data used for this paper and the paper
+% itself, has been published with a free license in GitHub at
+% \url{https://github.com/JJ/modeling-volunteer-computing}.
 

### 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Putting full affiliations
Closes #90

@@
 \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
-  (CAT), University of Granada, Granada, Spain (UGR)\\
+  and CITIC, University of Granada, Granada, Spain \\
 Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
 Tecnológico de Tijuana, Tijuana, M\'exico\\
-Pedro A. Castillo, Dept. of CAT, UGR\\
-Pablo Garc\'ia-S\'anchez, Dept. of CAT, UGR\\
-Paloma de las Cuevas, Dept. of CAT, UGR\\Nuria Rico, Dept. of Statistics and Operational Research, UGR%
+Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
+Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
+Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
+University of Granada, Granada, Spain\\
+Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%
 

### 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Use normal font for PUT
Closes #92

@@
 \begin{table*}
-\caption{Summary of time per run, number of IPs and number of {\tt
-    PUT}s per IP in the initial runs. \label{tab:summary:os}}
+\caption{Summary of time per run, number of IPs and number of PUTs per IP in the initial runs. \label{tab:summary:os}}
 \begin{center}
@@
 \hline
-Date & Experiments &\#IPs &\#IPs &  \#IPs & Time (s) &  \#{\tt PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment\\
+Date & Experiments &\#IPs &\#IPs &  \#IPs & Time (s) &  \#PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment\\
      & \# & Different & Median & Max & Median &  Median & & & correlation \\ 
@@
 intervening in each experiment,  median time needed
-to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
+to finish the experiment, median number of HTTP PUTs per IP, and
 then two columns showing the percentage of experiments that took less
@@
 \includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
-\caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
+\caption{Number of PUTs per unique IP and fit to a Generalized
   Extreme Value distribution (in a lighter shade or blue). From left to right, experiments
@@
 decreases. This exponential distribution also appears in the
-distribution of HTTP {\tt PUT}s, equivalent to the number of
+distribution of HTTP PUTs, equivalent to the number of
 generations divided by 100, contributed by every user, which is shown
@@
 \caption{Summary of fit to Generalized Extreme Value distribution of
-  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
+  the number of PUTs per unique IP. \label{tab:puts:os}}
 \begin{center}
@@
 \hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
+Date & Median \#IPs & Max \#IPs & Median time (s) & Median \# PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
 \hline
@@
 \caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
+  the number of PUTs per worker. \label{tab:puts:ww}}
 \begin{center}
@@
 \includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in light
+\caption{Ranked number of PUTs per worker and GEV fit (in light
   or blue color, left) 
@@
 As we did in the previous subsection, we have also fitted the number
-of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
+of HTTP PUTs per worker to a  Generalized Extreme Value (GEV) function, with the result shown
 in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
@@
 \caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
+  the number of PUTs per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
 \begin{center}
@@
 \caption{Data and fit to GEV (left) and Weibull (right) of the number
-  of {\tt PUT}s per worker for the F15 (Rastrigin) function.}  
+  of PUTs per worker for the F15 (Rastrigin) function.}  
 % Plotted with ../data/ips-puts-rastrigin.R
@@
 plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
-{\tt PUT}s per node, ranked by number, along with a fit to the two
+PUTs per node, ranked by number, along with a fit to the two
 distributions we have used before, GEV and Weibull; the parameters of

### 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Bigger dots and bigger legends
Improving graphics refs #91

@@
   participating in it, with averages and standard deviation shown as
-  red dots; in the case there is a single dot, there was a single
-  experiment in which many computers participated (for instance, 16
+  black dots and lines; in the case there is a single dot, there was a single
+  experiment in which so many computers participated (for instance, 16
   computers in the experiment in the far left or 29 in the middle
   one). 
-Shades of gray indicate how many experiments included that many unique IPs,
-so lighter shade for a column of dots indicates that a particular number
+Shades of gray or color indicate how many experiments included that many unique IPs,
+so lighter (bluer) shade for a column of dots indicates that a particular number
 of computers happened less frequently, while darker shade or closer to
-redmeans more frequency. 
+red means more frequency. 
 From left to right, experiments 4/4, 4/24 and 7/31.}

### 2016-01-18 data/time-vs-IPs-openshift.R

@@ -1,5 +1,5 @@
 library(ggplot2)
-
+library(grid)
                         
 time.vs.IPs <- function( data, preffix ) {
     data$howmany <- 1:nrow(data)
@@ -7,7 +7,7 @@ time.vs.IPs <- function( data, preffix ) {
         data[i,]$howmany <- length(data[data$IPs==data[i,]$IPs,]$milliseconds)
     }
     data$i <- 1:nrow(data)
-    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="#11AA99",high='#EE3333')+geom_point()+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=30), axis.text.y = element_text(angle=90, vjust=0.5)) 
+    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="#11AA99",high='#EE3333')+geom_point(size=4)+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=25), legend.key.height=unit(2, "line"), axis.text.y = element_text(angle=90, vjust=0.5)) 
     ggsave(paste0("../img/time-vs-ips-OS-",preffix,".png"))
     ggplot(data=data, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
     ggsave(paste0("../img/duration-vs-time-OS-",preffix,".png"))

### 2016-01-18 data/plot-zipf-openshift.R
Changes letter size fig 3
Refs #91

@@ -12,7 +12,7 @@ process.puts <- function(suffix ) {
                                     this.fit$estimate["loc"],
                                     this.fit$estimate["scale"],
                                     this.fit$estimate["shape"]),decreasing=T))
-    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=30), axis.text.y = element_text(angle=90, vjust=0.5))
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=25), axis.text.y = element_text(angle=90, vjust=0.5))
     ggsave(paste0("../img/puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }

### 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adjusts caption closes #91

@@
   one). 
-Shades of gray or color indicate how many experiments included that many unique IPs,
+Shade of gray or color indicate how many experiments included that many unique IPs,
 so lighter (bluer) shade for a column of dots indicates that a particular number

### 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
References checked closes #93

@@
 
-This work has been supported in part by TIN2011-28627-C04-02 and
-TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
+This work has been supported in part by
+TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
 SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
First too big version for GECCO.
Will have to be cut down.

--- 2016-01-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
@@
-\documentclass[journal,onecolumn]{IEEEtran}
-
-\usepackage{cite}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
+% This is "sig-alternate.tex" V2.1 April 2013
+% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
+%
+% This example file demonstrates the use of the 'sig-alternate.cls'
+% V2.5 LaTeX2e document class file. It is for those submitting
+% articles to ACM Conference Proceedings WHO DO NOT WISH TO
+% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
+% The 'sig-alternate.cls' file will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+%
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V2.5) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) NO page numbers
+%
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'sig-alternate.cls' you have control, however, from within
+% the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the ACM Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% e.g.
+% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
+% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% ================= IF YOU HAVE QUESTIONS =======================
+% Questions regarding the SIGS styles, SIGS policies and
+% procedures, Conferences etc. should be sent to
+% Adrienne Griscti (griscti@acm.org)
+%
+% Technical questions _only_ to
+% Gerald Murray (murray@hq.acm.org)
+% ===============================================================
+%
+% For tracking purposes - this is V2.0 - May 2012
+
+\documentclass{sig-alternate}
+
 \graphicspath{{../img/}}
-\else
-  \usepackage[dvips]{graphicx}
-  \graphicspath{{../img/}}
-\fi
 
@@
 
-% correct bad hyphenation here
-\hyphenation{op-tical net-works}
-
 \begin{document}
 
-\title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
+% Copyright
+\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\doi{10.475/123_4}
+
+% ISBN
+\isbn{123-4567-24-567/08/06}
+
+%Conference
+\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}
+
+\acmPrice{\$15.00}
+
+%
+% --- Author Metadata here ---
+\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources 
 for Low Cost Distributed Evolutionary Computation }
-% The word massive in the title is giving us problems.
-% Why not change it to something about sharing resources?.
-% Also lets not forget about volunteers, they can be a resource too
-% and they cannot be replaced by any super computer
-
-%  I propose something like:
-% A Web Based Architecture to Harness Volunteer Computational Resources 
-% for Low Cost Distributed Evolutionary Computation
-
-\author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
-  and CITIC, University of Granada, Granada, Spain \\
-Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
-Tecnológico de Tijuana, Tijuana, M\'exico\\
-Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
-University of Granada, Granada, Spain\\
-Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%
-
-\thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{5} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+% \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
+%   and CITIC, University of Granada, Granada, Spain \\
+% Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
+% Tecnológico de Tijuana, Tijuana, M\'exico\\
+% Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
+% Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
+% Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
+% University of Granada, Granada, Spain\\
+% Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%
+
+% \thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
+% }
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
+       \affaddr{Institute for Clarity in Documentation}\\
+       \affaddr{1932 Wallamaloo Lane}\\
+       \affaddr{Wallamaloo, New Zealand}\\
+       \email{trovato@corporation.com}
+% 2nd. author
+\alignauthor
+G.K.M. Tobin\titlenote{The secretary disavows
+any knowledge of this author's actions.}\\
+       \affaddr{Institute for Clarity in Documentation}\\
+       \affaddr{P.O. Box 1212}\\
+       \affaddr{Dublin, Ohio 43017-6221}\\
+       \email{webmaster@marysville-ohio.com}
+% 3rd. author
+\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
+one who did all the really hard work.}\\
+       \affaddr{The Th{\o}rv{\"a}ld Group}\\
+       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
+       \affaddr{Hekla, Iceland}\\
+       \email{larst@affiliation.org}
+\and  % use '\and' if you need 'another row' of author names
+% 4th. author
+\alignauthor Lawrence P. Leipuner\\
+       \affaddr{Brookhaven Laboratories}\\
+       \affaddr{Brookhaven National Lab}\\
+       \affaddr{P.O. Box 5000}\\
+       \email{lleipuner@researchlabs.org}
+% 5th. author
+\alignauthor Sean Fogarty\\
+       \affaddr{NASA Ames Research Center}\\
+       \affaddr{Moffett Field}\\
+       \affaddr{California 94035}\\
+       \email{fogartys@amesres.org}
 }
@@
 
+\maketitle
+
 \begin{abstract}
-%Volunteer computing is extensively used in science and industry, yet it still has not become a mainstream technology
-%in the same way the cloud technologies in general are. The lack of general purpose and open source tools with built-in
-%analytics could be one of the reasons of why this is so, as well as the lack of proofs of concept.
 
@@
 follows some patterns, and, thus, are amenable to a model and thus predictable.
-% Volunteer computing is extensively used in science and industry, but
-% it still has a long way to go in order to become a mainstream
-% technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
-% The lack of general purpose and open
-% source tools with built-in analytics could be one of the reasons of
-% why this is so, as well as the lack of proofs of concept.
-% % (Paloma) The following sentence is way too long, and the "which" after "Javascript" is confusing, so one does not know if the which refers to the framework or Javascript by reading the whole sentence.
-% % I propose: "That is why in this paper we present NodIO, a framework used to create client-server architectures for distributed evolutionary algorithms, entirely written in JavaScript.
-Our framework needs minimal server support to work and is able to support persistent, asynchronous, distributed evolutionary algorithms running in the browser, as well as any other client consuming its REST application programming interface."
-% That is why
-% in this paper we present {\sf NodIO}, a framework used to create
-% client-server architectures for distributed evolutionary algorithms,
-% entirely written in JavaScript, which needs minimal server support to
-% work and that is able to support persistent, asynchronous, distributed
-% evolutionary algorithms running in the browser, as well as any other
-% client consuming its REST application programming interface. An
-% enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
-% also presented; this version leverages the HTML5 Web Workers
-% specification to execute several clients per browser and do it in the
-% background, so that the the number of generations per user could be
-% improved. The architecture, along with some initial results on the
-% expected performance and scaling behavior when the number of users is
-% increased is introduced. % (Paloma) I've changed this "presented" for a synonym
-% We also use the algorithm data to find out
-% what kind of patterns the user behavior follows by examining the
-% resources they devote to our experiment, and try to fit them in % (Paloma) Changed. Phrasal verb was wrong.
-%  some
-% statistical distribution to create the bases of a model of user
-% behavior in volunteer computing experiments of this kind that can be
-% used by researchers to predict their performance. %(Paloma) Changed to "their" because of "experiments"
-% In general, our
-% experiments prove, first, that {\sf NodIO} is a flexible and
-% low-overhead platform to perform distributed computation experiments
-% that include volunteers, and then that user behavior follows some
-% patterns, and, thus, are amenable to a model.
+
 \end{abstract}
 
-% Work in progress..
-%Volunteer computing is used routinely in science and industry, but it
-%still has a long way to go if it is going to become a mainstream technology 
-%in the same way the cloud technologies in general have. In one hand, projects 
-%lack special purpose and open source tools with analytics builtin and
-%in the other there are still barriers that keep users from sharing their computing 
-%resources like the need to install software applications or plugins in order 
-%to cooperate. That is why in this paper we present {\sf NodIO}, a client-server
-%architecture for distributed
-%evolutionary algorithms, written entirely in JavaScript, which needs
-%minimal server support to work and is able to support persistent,
-%asynchronous, distributed evolutionary algorithms running in the
-%browser.  The architecture, along with some initial results on the
-%expected performance and scaling
-%behavior is presented. We also use the algorithm data to
-%find out what kind of patterns the user behavior follow and try to fit
-%them to some statistical distribution.
+\keywords{Volunteer computing, distributed computing, cloud computing}
 
@@
 
-\IEEEPARstart{N}{owadays} there is a big and increasing number of
+Nowadays there is a big and increasing number of
 distributed evolutionary computation software frameworks available on
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
 \bibliography{geneura,volunteer,javascript,ror-js,GA-general}

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Eliminating unneeded references
Acknowledgements commented for double-blind. References #95

@@
 
-\maketitle
-
 \begin{abstract}
@@
 they assume that the nodes that are going to run the system are under
-the control of developers or they at least have access to them. In
-fact, these features are commonly found in most soft computing
-frameworks \cite{7094263}.
+the control of developers or they at least have access to them.
 
@@
 %---------------------------------------------------------------
-\section*{Acknowledgment}
-
-This work has been supported in part by
-TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
-SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
-GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
-Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
-We would also like to thank the
-anonymous reviewers of previous versions of this paper who have really
-helped us to improve 
-this paper (and our work) with their suggestions. We are also grateful
-to Anna S\'aez de Tejada for her help with the data processing scripts.
+% \section*{Acknowledgment}
+% 
+% This work has been supported in part by
+% TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
+% SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
+% GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
+% Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
+% We would also like to thank the
+% anonymous reviewers of previous versions of this paper who have really
+% helped us to improve 
+% this paper (and our work) with their suggestions. We are also grateful
+% to Anna S\'aez de Tejada for her help with the data processing scripts.
 

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Anonymizing a couple of references, the only ones I found that can discover us. References #95

@@
 There is, then, space for frameworks that do not make any of these
-assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607M}. {\sf NodIO} is basically 
+assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607Manom}. {\sf NodIO} is basically 
 %assumptions, such as the one presented in this paper, which will be
@@
 The fulfillment of the desirable features of any distributed evolutionary 
-computation experiments that rely on volunteer computing \cite{2016arXiv160101607M} could be achieved by using any of the
+computation experiments that rely on volunteer computing \cite{2016arXiv160101607Manom} could be achieved by using any of the
 languages and associated technologies, mentioned above, provided that the user had
@@
 comparison have been described in
-\cite{2016arXiv160101607M}. We will focus next on the experimental
+\cite{2016arXiv160101607Manom}. We will focus next on the experimental
 results using the first version of the framework.
@@
 Initial results using the same code base and a single computer have
-been described in \cite{2016arXiv160101607M}. They prove that a bigger population
+been described in \cite{2016arXiv160101607Manom}. They prove that a bigger population
 contributes to diversity and thus speeds up the solution
@@
 announced through a post in Twitter and other social networks, and
-results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
+results were published here \cite{DBLP:conf/gecco/GuervosG15anom}. For the
 purpose of this paper, we repeated the announcement several times
@@
 experiments involved 5 or less IPs. This is consistent with results
-previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
+previously obtained \cite{DBLP:conf/gecco/GuervosG15anom} which found 6 to be
 the usual number of IPs that participated in an experiment. The
@@
 architecture that uses Web Workers, which is described in
-\cite{2016arXiv160101607M}; web workers work in the background even if
+\cite{2016arXiv160101607Manom}; web workers work in the background even if
 the web page is not in focus. A few improvements were also tested,
@@
 baseline speed for JavaScript in an standalone experiment and other
-languages is shown in \cite{2016arXiv160101607M}, where the details of
+languages is shown in \cite{2016arXiv160101607Manom}, where the details of
 the implementation have also been extensively described.

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Deleting boilerplate
New abstract written
The deadline for submitting abstracts is tomorrow. It can be change from now to final submission, but it is better to have this right now. We can modify the abstract later on. Refs #96. Please @mariosky @unintendedbear @fergunet check it out.

@@
 
-% DOI
-\doi{10.475/123_4}
-
-% ISBN
-\isbn{123-4567-24-567/08/06}
-
 %Conference
-\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}
-
-\acmPrice{\$15.00}
-
-%
-% --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+\conferenceinfo{GECCO '16}{Denver, CO, USA}
 % --- End of Author Metadata ---
@@
 
-Using the browser as a computing resource in volunteer computing has presented several advantages, 
-but it remains a challenge to fully harness both the users' behaviour and the browser's capabilities.
-That is why we have enhanced {\sf NodIO}, a framework used to create client-server architectures 
-for distributed evolutionary algorithms, resulting in {\sf
-  NodIO-W$^2$}, whose results we present in this paper.
-{\sf NodIO-W$^2$} leverages the HTML5 Web Workers specification to execute in the background several clients per
-browser, so that the the number of generations per user could be improved. The enhancements on the NodIO architecture,
-along with some initial results on the expected performance and scaling behavior when the number of users is increased,
-are introduced. We also use the experimental data to find out what kind of patterns the user behavior follows by examining 
-the resources they devote to our experiment, and try to fit them in some statistical distribution. In general, 
-our experiments prove, first, that {\sf NodIO} is a flexible and low-overhead platform to perform distributed computation
-experiments that include volunteers, and then that user behavior
-follows some patterns, and, thus, are amenable to a model and thus predictable.
+Using the browser as a computing resource in volunteer computing presents several advantages, 
+but it remains a challenge to harness the browser's capabilities and
+to model the user's behavior so that those capabilities can be
+leveraged optimally. In this paper we present the results of several
+experiments with a volunteer computing framework called {\sf NodIO}
+that uses JavaScript on the client and on the server and is easily
+deployable in any kind of cloud infrastructure, including free
+ones. We use different client configurations to find out which one is
+able to get the user to lend more computing cycles and test different
+problems to check the influence it has on said performance.
+From these experiments we can already draw some conclusions, besides
+the fact that volunteer computing can be a valuable computing
+resource: the user has to be kept engaged in some way to be able to
+use as many computing cycles as possible, the client has to be built
+to use the computer capabilities fully, and there is some common model
+to all sets of users, following a common statistical distribution. As
+a pragmatic result, establishing trust with the user by open sourcing
+not only the software, but also the data, is essential to have an user
+base as wide as possible.
 

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Few changes
Plese chech

@@
 experiments with a volunteer computing framework called {\sf NodIO}
-that uses JavaScript on the client and on the server and is easily
-deployable in any kind of cloud infrastructure, including free
+that uses JavaScript both on the client and the server and is easily
+deployed in any kind of cloud infrastructure, including free
 ones. We use different client configurations to find out which one is

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Final touches abstract

@@
 experiments with a volunteer computing framework called {\sf NodIO}
-that uses JavaScript both on the client and the server and is easily
-deployed in any kind of cloud infrastructure, including free
-ones. We use different client configurations to find out which one is
+that is easily deployable in using free cloud resources. We use
+different browser configurations to find out which one is 
 able to get the user to lend more computing cycles and test different

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex

@@
 %Mario: Is NodIO presented here after the cut? 
+% Not sure. Maybe not. - JJ
 %TO DO: Add reference
@@
 
-The main problem with Java and all the other browser-embedded
+The main problem with using Java and all the other browser-embedded
 technologies is that they are not {\em universal} in the
@@
 adoption of the HTML5 standard \cite{anthes2012html5}.
-JavaScript \cite{flanagan2006javascript}, however, has become a set of standards
-\cite{ECMA-262} that include the language and its components. Since
-early on, it was adopted by the industry and also by scientists,
-who used them for creating an evolutionary algorithm (EA from now on) on
-the browser as early as 1998 \cite{jj-ppsn98}.
-
-JavaScript can be used either for unwitting
+JavaScript however, has become a set of standards
+\cite{ECMA-262} that include the language and its components.
+
+JavaScript can be used either for unwitting, that is, stealth,
 \cite{unwitting-ec} or volunteer 
@@
 authors, including more recent efforts
-\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
 that even
-used the client's GPU \cite{duda2013gpu} or create an object-oriented
-framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
+used the client's GPU \cite{duda2013gpu}. Even if, as said above, Java
+is not the best option, many other researchers have
 used Java \cite{chong:1999:jDGPi} and others have gone away from the
 server-based paradigm to embrace peer to peer systems
-\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
+\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
+computing 
 platforms avoid single points of failure (the server) but, since they
 need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower. % (Paloma) Do yo mean... higher? A low threshold means that is easy to happen something.
-
-Recent works propose the use of cloud computing services for the components of
-client-server distributed EAs. Cloud resources
-can reduce operational costs by outsourcing both hardware and software maintenance
-to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
-is available in most laboratories, allowing researchers to
-scale algorithms at reasonable costs. Researchers have proposed the use of
-public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
-and DropBox~\cite{mericloud}. The use of JavaScript to implement
-server side components
-gives software developers certain advantages in the cloud
-paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
-developers can use the same software tools for both client and server
-side development.
+threshold to join them is much higher. % (Paloma) Do yo
+                                % mean... higher? A low threshold
+                                % means that is easy to happen
+                                % something.
+%Correct. But maybe in this new version we should eliminate this
+%paragraph altogether. 
+% The next paragraph has been suppresed.
 
@@
 reproductibility, interoperability, automated assembly, knowledge engineering,
-and efficiency.
-
-In the work presented in this paper, JavaScript has been used throughout the full
+and efficiency. In the work presented in this paper, JavaScript has been used throughout the full
 stack, which has the advantage of using the self-same lines of code for
@@
 the expertise acquired
-% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
+% (Paloma) Not sure if "acquired" should be written before
+% "expertise"... sounds better to me, anyways
+% Don't know what you mean. Write it down and we'll see - JJ
  in the library itself.
@@
 of volunteer computing systems, an area in which there have been
-relatively few efforts.
-There were some initial attempts to avoid the differences in performance
+relatively few efforts. There were some initial attempts to avoid the differences in performance
 that could be obtained from volunteers  by making
@@
 On the other hand, measuring the performance of the resulting metacomputer
-involves understanding the dynamics of this kind of systems. Initial
+involves understanding its dynamics. Initial
 work was done for peer to peer systems by Stutzbach et
@@
 
-As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
+As far as we know, this paper presents one of the few experimental
+results using a volunteer computing system that uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
 computational resources that are as dissimilar as smartphones and
 powerful laptops 
-or workstation computers in a research center. 
-%The methodology used
-% (Paloma) I would say "the used methodology"
-The algorithms used, as well as the methodology 
-for gathering resources and baseline results that we will use for
-comparison have been described in
-\cite{2016arXiv160101607Manom}. We will focus next on the experimental
-results using the first version of the framework.
+or workstation computers in a research center. This will be described next.
 
@@
 \label{sec:exp1}
+
+% A bit of description of the framework, with reference to published
+% reports - JJ
 

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex
Minor comments and changes for @JJ although I guess it is already sent :P

@@
 to model the user's behavior so that those capabilities can be
-leveraged optimally. In this paper we present the results of several
+leveraged optimally. In this paper, we present the results of several
 experiments with a volunteer computing framework called {\sf NodIO}
 that is easily deployable in using free cloud resources. We use 
+%Pablo: "deployable in using free cloud resources" is correct?
 different browser configurations to find out which one is 
@@
 problems to check the influence it has on said performance.
+% Pablo: so the objective is "to determine which browser conf. is able
+% to get more computing power from the users", it isn't? 
+% (to keep this in mind the rest of the paper). What is a "browser configuration"?
+% (explain it with a phrase in the abstract)
 From these experiments we can already draw some conclusions, besides
@@
 a pragmatic result, establishing trust with the user by open sourcing
-not only the software, but also the data, is essential to have an user
+not only the software, but also the data, is essential to have a user %Pablo: remember, is "yiuser" :P
 base as wide as possible.

### 2016-02-01 GECCO/volunteer-clouds-gecco16.tex
Reduced it to 8 pages
Please @unintendedbear @mariosky @fergunet check it for consistency, explained things (or not) and anything you can see. GECCO is a very difficult venue, we really have to nail it.

@@
 %Pablo: "deployable in using free cloud resources" is correct?
-different browser configurations to find out which one is 
+% why not? - JJ
+different implementations to find out which one is 
 able to get the user to lend more computing cycles and test different
@@
 % (explain it with a phrase in the abstract)
+% No, we don't. Please read the paper!!!! - JJ
 From these experiments we can already draw some conclusions, besides
@@
 use as many computing cycles as possible, the client has to be built
-to use the computer capabilities fully, and there is some common model
-to all sets of users, following a common statistical distribution. As
-a pragmatic result, establishing trust with the user by open sourcing
-not only the software, but also the data, is essential to have a user %Pablo: remember, is "yiuser" :P
-base as wide as possible.
+to use the computer capabilities fully, and, finally, the user
+contributions follow a common statistical distribution. It is also
+essential to be as open as possible with software and data.
 
@@
 such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript \cite{2015arXiv151101088M}. However, one of the advantages of our architecture is that
+JavaScript. However, one of the advantages of our architecture is that
 clients do not need to install anything in their machines, they just
@@
 adoption of the HTML5 standard \cite{anthes2012html5}.
-JavaScript however, has become a set of standards
-\cite{ECMA-262} that include the language and its components.
+JavaScript however, has become a set of standards that include the language and its components.
 
@@
 that even
-used the client's GPU \cite{duda2013gpu}. Even if, as said above, Java
-is not the best option, many other researchers have
-used Java \cite{chong:1999:jDGPi} and others have gone away from the
-server-based paradigm to embrace peer to peer systems
-\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
-computing 
-platforms avoid single points of failure (the server) but, since they
-need a certain amount of infrastructure installed to start, the
-threshold to join them is much higher. % (Paloma) Do yo
+used the client's GPU \cite{duda2013gpu}. 
+%Also a bit out of scope. Have to reduce the paper. 
+% Even if, as said above, Java
+% is not the best option, many other researchers have
+% used Java \cite{chong:1999:jDGPi} and others have gone away from the
+% server-based paradigm to embrace peer to peer systems
+% \cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
+% computing 
+% platforms avoid single points of failure (the server) but, since they
+% need a certain amount of infrastructure installed to start, the
+% threshold to join them is much higher. % (Paloma) Do yo
                                 % mean... higher? A low threshold
@@
 
-JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
-According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
-metaheuristics optimization is that it allows an improvement in communicability,
-reproductibility, interoperability, automated assembly, knowledge engineering,
-and efficiency. In the work presented in this paper, JavaScript has been used throughout the full
-stack, which has the advantage of using the self-same lines of code for
-writing an evolutionary algorithm for the
-desktop or command line and browser clients, both based in the EC
-library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
-when writing code and leverages 
-the expertise acquired
-% (Paloma) Not sure if "acquired" should be written before
-% "expertise"... sounds better to me, anyways
-% Don't know what you mean. Write it down and we'll see - JJ
- in the library itself.
+%Also suppressing this one which is probably not needed in this paper
+%- JJ
+% JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
+% According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
+% metaheuristics optimization is that it allows an improvement in communicability,
+% reproductibility, interoperability, automated assembly, knowledge engineering,
+% and efficiency. In the work presented in this paper, JavaScript has been used throughout the full
+% stack, which has the advantage of using the self-same lines of code for
+% writing an evolutionary algorithm for the
+% desktop or command line and browser clients, both based in the EC
+% library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
+% when writing code and leverages 
+% the expertise acquired
+% % (Paloma) Not sure if "acquired" should be written before
+% % "expertise"... sounds better to me, anyways
+% % Don't know what you mean. Write it down and we'll see - JJ
+%  in the library itself.
 
@@
 As far as we know, this paper presents one of the few experimental
-results using a volunteer computing system that uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
-computational resources that are as dissimilar as smartphones and
-powerful laptops 
-or workstation computers in a research center. This will be described next.
+results using a volunteer computing system which uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
+very diverse computational resources; these will be described next.
 
@@
 
-Initial results using the same code base and a single computer have
-been described in \cite{2016arXiv160101607Manom}. They prove that a bigger population
-contributes to diversity and thus speeds up the solution
-\cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing 
-experiments that we will describe next do not and can not have the
-same conditions, but
-the baseline is that if they eventually take longer than a basic
-desktop, their interest will be purely academic, although they can
-complement them. We will try and
-prove next that that target performance can be achieved by carrying
-out several experiments in different conditions. Second, they also act
-as a basic performance test for the server, and also prove that the
-basic server can be used for many kinds of different setups, as long
-as they follow the REST protocol to deposit and obtain individuals
-from the pool. It is obvious that, in this case, the server is not
-really needed, since there is a single client and no interaction
-between them, but there is a cost in sending and obtaining chromosomes
-from the server so it is included for the sake of fairer comparison.
-
-Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which these runs could be performed. That way, and as required, the hardware
-and cloud cost for this experiment were zero. Experiments were
-announced through a post in Twitter and other social networks, and
-results were published here \cite{DBLP:conf/gecco/GuervosG15anom}. For the
-purpose of this paper, we repeated the announcement several times
-through the month of April and then by the beginning of August. All
-in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:summary:os}. In general, every experiment took several
-days. No particular care was taken about the time of the announcement
-or the particular wording. Every {\em experiment} consisted in running
-until the solution of the 40-trap problem was found. When the correct
-solution was sent to the server, the counter was updated and the pool
-of solutions reset to the void set. There was no special intention to wait
-until all clients had finished, thus it might happen that, in fact,
-the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible.
+We set up different experiments by deploying the server in a cloud
+service, which changed in every one. This server runs a web service
+which stores chromosomes sent from clients and sends them back. It is
+an open source system that has been described extensively in
+\cite{2016arXiv160101607Manom}. The client runs an evolutionary
+algorithm written in JavaScript, that after a number of generations
+sends back the best individual to the server, requesting at the same
+time a random individual of those stored in the server. The experiment
+starts by setting up a website that is sent by the server and announcing it on several social
+networks so that people contribute to them by going to the page and
+staying there for a while.
+
+\subsection{Initial tests}
+
+Initial tests with an evolutionary algorithm solving the 40-Trap
+problem were set up using the free PaaS (Platform as a
+Service) OpenShift; they are described in
+Table \ref{tab:summary:os}. In general, every tests took several
+days and during it the problem was solved more than 50 times with
+several hundreds different IPs participating in it; the
+evolutionary algorithm stopped when the solution was found. 
 %
@@
 %
-The table shows that every run included more than 50 experiments. The
-number of different IPs intervening in them varied from more than one
-hundred to more than five hundred in the second experiment.
-
-A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of IPs
-intervening in each experiment,  median time needed
-to finish the experiment, median number of HTTP PUTs per IP, and
-then two columns showing the percentage of experiments that took less
-than the two baseline experiments shown in the introduction to this
-section. The first striking result is that in all cases, 50\% of the
-experiments involved 5 or less IPs. This is consistent with results
-previously obtained \cite{DBLP:conf/gecco/GuervosG15anom} which found 6 to be
-the usual number of IPs that participated in an experiment. The
-maximum number of different IPs for each experiment is also in the
-same range and of the order of 10, which is also consistent with
-prior work. The median time has a big range of variation, but 50\% of
-the time takes less than several minutes, from around 4 minutes in the
-best case to roughly 2/3 of an hour in the worst case. That figure is
-not competitive, {\em a priori}, with the baseline experiment, that is
-why it is interesting to look at the columns which measure
-the percentage of experiments in which the solution was found in less
-time than the baseline. It is similar in all three cases, with around
-20\% on average for the first baseline experiment, roughly 3\% in the
-second case. The comparison is not totally fair: this experiment takes
-place in a browser, and in fact most of the time is devoted to
-presenting the charts; the server is also not the same since it is run in a
-PaaS instead of a high-performance desktop computer, and it is also a
-remote server, not a local server as in the baseline experiment. But
-even in this case, we were looking for creating a framework that
-allowed us to run evolutionary computation experiments {\em faster} than
-in our very own computer. And the results indicate that although it will
-happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
-experiment you choose, it will not happen always or even most of
-the time. One of the problems is shown in the last column, which
-indicates the statistical correlation between the number of IPs
-participating in one experiment and the next. In all cases,
-correlation is so low as to affirm that there is almost no
-relationship between them, that is, computing nodes are not {\em
-  staying} after the solution has been found. This is a feature of
-this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. Thus, the user can decide if to 
-continue sharing resources or not. % (Paloma) Feature or disadvantage?
-However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
-unnecessary upper limit on the time, or number of experiments, that a user
-will voluntarily perform.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
-\caption{Duration of experiments vs. number of different IPs (nodes)
-  participating in it, with averages and standard deviation shown as
-  black dots and lines; in the case there is a single dot, there was a single
-  experiment in which so many computers participated (for instance, 16
-  computers in the experiment in the far left or 29 in the middle
-  one). 
-Shade of gray or color indicate how many experiments included that many unique IPs,
-so lighter (bluer) shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shade or closer to
-red means more frequency. 
-From left to right, experiments 4/4, 4/24 and 7/31.}
-\label{fig:duration}
-\end{figure}
-% ../data/time-vs-IPs-openshift.R
+
+The last column shows the percentage of experiments in which running the
+experiment in the browser takes {\em less} time than in the
+desktop. Although volunteer computing systems do not exclude desktop
+computing and, in fact, the server can handle any client, not only
+those running JavaScript on the browser, we include it here for the
+sake of comparison and to show that, being a free resource, it can
+anyway make a valid and valuable contribution to any distributed or
+desktop computing experiment. An interesting result is also that the
+correlation between users of an experiment and the next is
+non-existent, that is, computing nodes are not {\em
+  staying} after the solution has been found. This is an issue that we
+will address in this paper.
+
+% \begin{figure}[!htb]
+% \centering
+% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
+% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
+% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
+% \caption{Duration of experiments vs. number of different IPs (nodes)
+%   participating in it, with averages and standard deviation shown as
+%   black dots and lines, single dot if it was 1). Shade of gray or
+%   color indicate how many experiments included that many unique IPs.
+% From left to right, experiments 4/4, 4/24 and 7/31.}
+% \label{fig:duration}
+% \end{figure}
+% % ../data/time-vs-IPs-openshift.R
 %
@@
 %
-We will have to analyze experimental data a bit further to find out why
-this happens and also if there are some patterns in the three sets of
-experiments. An interesting question to ask, for instance, is if
-by adding more computers makes the experiment take less. In fact, as
-shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
-not seem to contribute to decreasing the time needed to finish the
-experiment. However, the cause-effect relationship is not clear at
-all. It might be the opposite: since experiments take longer to finish
-and might in fact be abandoned with no one contributing for some time,
-that increases the probability of someone new joining them. In fact,
-with experiments taking a few seconds and due to the way the
-experiments are announced, it is quite difficult that several
-volunteers join in in such a short period of time, even more if we take
-into account that volunteers are not {\em carried over} from previous
-experiments. This implies that it would be convenient to use a problem
-of a bigger size to check this hypothesis as we have done in the next
-experiment. 
-%  however, at this point we
+% We will have to analyze experimental data a bit further to find out why
+% this happens and also if there are some patterns in the three sets of
+% experiments. An interesting question to ask, for instance, is if
+% by adding more computers makes the experiment take less. In fact, as
+% shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
+% not seem to contribute to decreasing the time needed to finish the
+% experiment. However, the cause-effect relationship is not clear at
+% all. It might be the opposite: since experiments take longer to finish
+% and might in fact be abandoned with no one contributing for some time,
+% that increases the probability of someone new joining them. In fact,
+% with experiments taking a few seconds and due to the way the
+% experiments are announced, it is quite difficult that several
+% volunteers join in in such a short period of time, even more if we take
+% into account that volunteers are not {\em carried over} from previous
+% experiments. This implies that it would be convenient to use a problem
+% of a bigger size to check this hypothesis as we have done in the next
+% experiment. 
+%  However, at this point we
 % have not found this convenient since there are several other issues
@@
 for F15, which is the second version tested, but it would be interesting to
-measure exactly what this influence has been. In general there are
-many issues with the evolutionary algorithm implementation itself,
-including using different, or adaptive, policies for inserting and
-sending individuals to the pool as was done in \cite{araujo2008mam},
-using different policies for population initialization, and also the
-incorporation of high-speed local resources to the pool to check what
-would be the real influence of the volunteer pool to the final
-performance.
+measure exactly what this influence has been. 
+% In general there are
+% many issues with the evolutionary algorithm implementation itself,
+% including using different, or adaptive, policies for inserting and
+% sending individuals to the pool as was done in \cite{araujo2008mam},
+% using different policies for population initialization, and also the
+% incorporation of high-speed local resources to the pool to check what
+% would be the real influence of the volunteer pool to the final
+% performance.
 

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Revised the intro and rewritten in part.
Remember that now we will only have the experiments, so we must tell as much about the server as possible in the intro. Hope it is enough. Please @mariosky @unintendedbear have a look.

@@
 Using the browser as a computing resource in volunteer computing presents several advantages, 
-but it remains a challenge to harness the browser's capabilities and
+but it remains a challenge to fully harness the browser's capabilities and
 to model the user's behavior so that those capabilities can be
 leveraged optimally. In this paper, we present the results of several
-experiments with a volunteer computing framework called {\sf NodIO}
-that is easily deployable in using free cloud resources. We use 
+evolutionary computation experiments with different implementations of
+a volunteer computing framework called {\sf NodIO}, designed to be
+easily deployable in using free cloud resources. We use  
 %Pablo: "deployable in using free cloud resources" is correct?
@@
 
-Nowadays there is a big and increasing number of
-distributed evolutionary computation software frameworks available on
-the market, usually as open source frameworks \cite{Parejo12Survey}. 
-Most of these frameworks have 
+Nowadays there is a big and increasing number of distributed
+evolutionary computation (EC) software frameworks available, usually
+as open source frameworks \cite{Parejo12Survey}.  Most of them have
 two features: they are {\em desktop} applications, that is, they must
@@
 
-%Mario: Is NodIO presented here after the cut? 
-% Not sure. Maybe not. - JJ
-%TO DO: Add reference
-There is, then, space for frameworks that do not make any of these
-assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607Manom}. {\sf NodIO} is basically 
-%assumptions, such as the one presented in this paper, which will be
-% named {\sf NodIO}. {\sf NodIO} is basically
-% (Paloma) Is the "first" word really needed? 
-% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
-a two tier client/server 
-application, in which clients can run on a browser. 
-Therefore, they can be embedded in any web-enabled device, including many over which 
-developers have no control.
+There is, then, space for other EC frameworks that do not make any of
+these assumptions, such as {\sf NodIO} presented in
+\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of
+client/server application and has some support to program an
+evolutionary algorithm embedded in pages served by them which can then
+communicate via that server. The client part can run in any
+web-enabled device, including many over which developers have no
+control, that is, volunteers.
 
@@
 need to open a web page  
-% Somewhat redundant 
-% Javascript is the only language that can be executed in this way without any installation, 
-% also removed the default, redundant with next sentence
-in a compatible browser, such as Chrome, Safari or Firefox, which are
+in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
 available in almost every desktop or mobile device. Other advantage is
@@
 compiled as other options such as Java Applets, ActionScript or
-ActiveX.
-Also, as we will discuss in this paper, it can provide some extra
-features such as enhanced security or operating system independence.
+ActiveX are. Also, as we will discuss in this paper, it can provide some extra
+features such as enhanced security or operating system
+independence. This system, however, is not intended to substitute
+frameworks running in dedicated hardware or the cloud, but to
+complement it. In general, several volunteers will be needed to
+overcome the overhead involved in using a client-server architecture
+and this number of volunteers is obviously not guaranteed.
 
@@
 
-When developing volunteer based frameworks, designers must
-also consider the whole {\em social} aspect in the design, with
+One of the aspects that must be taken into account in volunteer based
+frameworks is its {\em social} aspect in the design, with 
 issues related to security, trust, and privacy among others. 
-The computing system becomes then a {\em techno-social system} \cite{vespignani2009predicting}.  
-% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
-%a {\em social} aspect to the design of a
-%computing system that makes it a {\em techno-social system}
-In this paper, our intention is to
-analyze the {\sf NodIO} framework from two different
-%Mario: Edit this?
-%present the {\sf NodIO} framework and analyze it from two different
-angles: the purely technical angle, including the decisions that went
-into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page running the experiment
-determine its performance and how it can be modeled and, eventually,
-predicted.
+The computing system becomes then a {\em techno-social system}
+\cite{vespignani2009predicting}. In that sense, its social aspects
+cannot be separated from its technical specifications, with one
+feeding into the other. One of these aspects is safety for the user
+and the scientist running it: the user must be reasonably sure that the
+experiment is not misusing its resources and the scientist must have a
+certain safety in the sense that its server is not going to be
+attacked via cheating or denial or service attacks. We have decided to
+approach this in a pragmatic way. Including many security features
+would decrease performance; on the other hand, we publish all data
+resulting from the experiment once private information, such as IPs,
+has been processed to a untraceable hash. The experiments themselves
+are a proof that there are enough users trusting the experiment to
+give it its CPU time and that there has been no attempt to cheat by
+sending specially crafted chromosomes or to attack the server to make
+it stop working because there was no incentive in doing so. The
+system, besides, is running in a dedicated Platform as a Service
+container or Virtual Private Server so accessing the server would only
+show a few data obtained from the experiment, nothing more. Making it
+free software ensured this was known by anyone in the experiment, via
+links in the web page that hosted it.
 
 The rest of the paper is organized as follows: Next we present the
-state of the art (Section \ref{sec:soa}) in web-based distributed
+state of the art in web-based distributed
 computational systems along with attempts to predict and model its
-behavior. The first results in a volunteer setting are described
+behavior  in Section \ref{sec:soa}. The first results in a volunteer
+setting are described 
 in Section \ref{sec:exp1}, and experiments using a new client

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Adds a paragraph at the end of the intro
To make clear what we are doing in the paper in a few sentences.
Refs #97

@@
 
-The rest of the paper is organized as follows: Next we present the
+In this paper we will carry out three experiments, after establishing
+the baseline of the performance that can be expected in a JavaScript
+based system and its comparison with other languages in
+\cite{2016arXiv160101607Manom}. First, several experiments using a
+Platform as a Service, and a simple Ajax client that does not
+restart. Second, we change a bit the client so that it uses web
+workers, that is, it runs in the background and it restarts. Third, we use the same
+client but with a harder problem that cannot be solved by a single
+client and thus needs the contribution of many volunteers to
+proceed. With these three experiments we want mainly to try and model
+the behavior of users and how different problems and implementations
+have an influence on it.
+
+So, the rest of the paper is organized as follows: Next we present the
 state of the art in web-based distributed

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Rewriting conclusions
Mainly to get back to 8 pages. Abstract has been quite revised for now, so this closes #96

@@
 client but with a harder problem that cannot be solved by a single
-client and thus needs the contribution of many volunteers to
+volunteer and thus needs the contribution of many volunteers to
 proceed. With these three experiments we want mainly to try and model
@@
 issues involved in browser-based computing and also in making a system
-that is able to work across different types of computers, mentioning
-SPARCstations and PCs with Windows-95. Interestingly enough, this
+that is able to work across different types of computers. Interestingly enough, this
 paper incorporates a discussion on the economic model, paving the way
@@
 virtual environment has to be installed in the browser.
-Also these virtual environments have become targets of malicious attacks; 
-for instance numerous vulnerabilities have been discovered in 
-the proprietary Adobe Flash runtime environment and
-video player \cite{ford2009analyzing,watanabe2010new}. The need for
+% Also these virtual environments have become targets of malicious attacks; 
+% for instance numerous vulnerabilities have been discovered in 
+% the proprietary Adobe Flash runtime environment and
+% video player \cite{ford2009analyzing,watanabe2010new}. 
+The need for
 these additional environments has been mitigated with the
-adoption of the HTML5 standard \cite{anthes2012html5}.
-JavaScript however, has become a set of standards that include the language and its components.
+adoption of the HTML5 standard \cite{anthes2012html5}, which is
+tightly integrated with JavaScript.
 
@@
 %
-\begin{table*}
+\begin{table}
 \caption{Summary of fit to Generalized Extreme Value distribution of
@@
 \end{center}
-\end{table*}
+\end{table}
 %
@@
 %
-\begin{table*}
+\begin{table}
 \caption{Summary of fit to GEV and Weibull distribution of
@@
 \hline
-GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
-0.04746 \\
-Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
+GEV & 18.275 $\pm$ 0.837  &  25.96  $\pm$ 1.22 & 1.242   $\pm$ 0.047 \\
+Weibull & ND & 75.11 $\pm$ 2.97  & 0.697 $\pm$ 0.014 \\
 \hline
@@
 %---------------------------------------------------------------
-\section{Conclusion}
+\section{Conclusions and future work}
 \label{sec:conclusion}
@@
 In general, linking and finding correlations between user choices and
-performance is an interesting avenue to explore in the future. Even if
-the three previous experiments were published in a similar way, one
-obtained up to 5 times more total cycles  than the one with the least
-number of cycles. It is also essential to obtain volunteers as
+performance is an interesting avenue to explore in the future.
+%  Even if
+% the three previous experiments were published in a similar way, one
+% obtained up to 5 times more total cycles  than the one with the least
+% number of cycles. 
+It is also essential to obtain volunteers as
 simultaneously as possible, so it is possible that the features of the
@@
 
-Although we think that, in general, the results presented in this
-paper are independent of the problem chosen, it is true that even if
-it is a difficult problem for evolutionary algorithms, it takes a few
-seconds with the right settings to solve. This makes it difficult also
-for this kind of asynchronous framework, but it would be interesting
-to check the performance for other computationally demanding
+This paper also shows that the time needed to find a solution or if it
+is eventually found have an influence on the time people devote to
+it. It would be so interesting to check the performance for other computationally demanding
 tasks, especially ones that cover the middle ground between a problem
@@
 has had a big influence on the results and in fact this is not noticed
-for F15, which is the second version tested, but it would be interesting to
+for F15, which is the last version tested, but it would be interesting to
 measure exactly what this influence has been. 
@@
 
-Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
-to create a generator in which the user just has to create a fitness
-function, with the rest of the framework wrapped around
-automatically.
+% Finally, the implementation needs some refinement in terms of
+% programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
+% to create a generator in which the user just has to create a fitness
+% function, with the rest of the framework wrapped around
+% automatically.
 %  All this, as well as the data used for this paper and the paper

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Revises again conclusions
Moving a part from results to there. Refs #98

@@
 
-% (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
+In general, this experiment confirms what was found in the others in
+several aspects: first, there is a regularity in the distribution of the
+time assigned by volunteers to it, second, there is a good and certain
+amount of volunteers that will become part of the distributed
+computer, in some cases up to one hundred; also, in some cases there
+will be users that will spend several hours, two orders of magnitude
+more generations in this case than in the previous one. There is a
+clear influence of the time to solution: a short time means that every
+experiment will only gather a few volunteers, but also that they will
+stay only for a short time. A long time will make users stay for
+longer, but if it takes too long, they will eventually lose interest.
+
+We will discuss this and the rest of the findings of the paper in the
+conclusions.
+
+%---------------------------------------------------------------
+\section{Conclusions and future work}
+\label{sec:conclusion}
+
+In this paper several experiments with a client-server architecture for volunteer and distributed
+evolutionary algorithms have been performed. The architecture uses the
+browser and has been generated using the {\sf
+  NodIO} framework, which includes  {\sf NodEO} as an evolutionary
+algorithm library. Volunteers are {\em in the cloud}, as stated in the
+title, 
+since they are a {\em CPU as a service} for the persons running the
+experiment. In fact, in this paper we have tried to put some figures
+on the real size of that {\em cloud} in case it is used by itself or
+in conjunction with other local or 
+cloud-based methods to add computing power in a seamless way through
+the pool that {\sf NodIO} creates.
+
+In order to establish a baseline performance, which was published in
+\cite{2016arXiv160101607Manom}, the evolutionary algorithm was run in a desktop client application written in JavaScript
+using NodEO to solve the 40-trap function. The first experiments in a
+volunteer computing scenario proved that, although obtaining better performance than the
+baseline  was possible, it did not happen, mainly because of the
+difficulty in carrying over other experiments to volunteers when the
+experiment they were participating on was finished. This has been shown in the low
+correlation between the number of IPs in successive experiments, and
+also resulted in a low number of generations allotted by users.
+
+In the second implementation, Web Workers
+were used so that several clients per browser page could be executed and
+run in the background; our intention was to increase the time spent by
+the users and, in fact, the number of generations per
+user increased to the point that we obtained a performance better than
+baseline in most cases. The architectural changes led to a high correlation
+between successive experiments, so that, even as the median number of
+volunteers per experiment decreased, they were probably contributing
+to the experiment simultaneously, so that our objective was
+achieved. In general, one of the first conclusions for this paper is
+that the browser technologies should be used to its full extent so
+that the user time is leveraged to its full potential, and this was
+achieved using web workers.
+
+The second objective of this paper was to model the user behavior in a
+first attempt to try and predict performance. As should be expected,
+the model and its parameters depends on the implementation, with contributions following
+a GEV distribution with single-thread and a
+Weibull distribution when web workers are used. These distributions are
+close enough to each other and, in the second case, reflect the fact
+that most time spent in the page is actually devoted to computing, as
+opposed to presenting the chart or not doing anything,
+which is why the time spent (represented by the number of
+contributions to the pool) follows a model quite similar to that found
+for games or other online activities. The reverse implication, from
+games to volunteer computing,  might also be true: if we
+want to have returning users for the experiments, it is probable that
+we should {\em gamify} the experience so that once they have participated
+once, they would be encouraged to do for a longer time or to share
+it. In the spirit of Open Science, this 
+gamification might involve computing in real-time data such as the one
+presented in this paper, while showing it in the same page.
+
+In general, linking and finding correlations between user choices and
+performance is an interesting avenue to explore in the future. % (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
+% Moved to conclusions. 
 There are several validity threats to the design of this study.
@@
 the architecture and the tools chosen.
-
-In general, this experiment confirms what was found in the others in
-several aspects: first, there is a regularity in the distribution of the
-time assigned by volunteers to it, second, there is a good and certain
-amount of volunteers that will become part of the distributed
-computer, in some cases up to one hundred; also, in some cases there
-will be users that will spend several hours, two orders of magnitude
-more generations in this case than in the previous one. There is a
-clear influence of the time to solution: a short time means that every
-experiment will only gather a few volunteers, but also that they will
-stay only for a short time. A long time will make users stay for
-longer, but if it takes too long, they will eventually lose interest.
-
-We will discuss this and the rest of the findings of the paper in the
-conclusions.
-
-%---------------------------------------------------------------
-\section{Conclusions and future work}
-\label{sec:conclusion}
-
-In this paper two versions of a client-server architecture for volunteer and distributed
-evolutionary algorithms have been evaluated. The architecture uses the browser and has been generated using {\sf
-  NodIO}, based on the {\sf NodEO} evolutionary algorithm library. Volunteers are {\em in the cloud}, as stated in the title,
-since they are a {\em CPU as a service} for the persons running the
-experiment. In fact, in this paper we have tried to put some figures
-on the real size of that {\em cloud} and how it can be used standalone
-if there is no alternative, or in conjunction with other local or
-cloud-based methods to add computing power in a seamless way through
-the pool that {\sf NodIO} creates.
-
-In order to establish a baseline performance, the evolutionary
-algorithm was run in a desktop client-program written in JavaScript
-using NodEO to solve the 40-trap function. The first experiments with
-{\sf NodIO} proved that, although obtaining better performance than the
-baseline was possible, it did not happen, mainly because of the
-difficulty in carrying over other experiments to volunteers when the
-experiment they were participating on was finished. This has been shown in the low
-correlation between the number of IPs in successive experiments, and also resulted in a low number of generations allotted by users.
-
-In the second implementation, {\sf NodIO-W$^2$}, Web Workers
-were used so that several clients per browser could be executed and
-they could run in the background, so that the the number of
-generations per user could be improved. The number of generations per
-user increased to the point that baseline performance was improved in
-most cases. The architectural changes led to a high correlation
-between successive experiments, so that, even as the median number of
-volunteers per experiment decreased, they were probably contributing
-to the experiment at the same time, so that our objective was
-achieved. In general, one of the first conclusions for this paper is
-that the browser technologies should be used to its full extent so
-that the user time is leveraged to its full potential, and this was
-achieved with {\sf NodIO-W$^2$}.
-
-The second objective of this paper was to model the user behavior in a
-first attempt to try and predict performance. As should be expected,
-the model depends on the implementation, with contributions following
-a GEV distribution in the case of {\sf NodIO} and a
-Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
-close enough to each other and, in the second case, reflect the fact
-that all time spent in the page is actually devoted to computing,
-which is why the time spent (represented by the number of
-contributions to the pool) follows a model quite similar to that found
-for games or other online activities. The reverse might be true: if we
-want to have returning users for the experiments, it is probable that
-we should {\em gamify} the experience so that once they have done it
-once, they might do it more times. In the spirit of Open Science, this
-gamification might involve computing in real-time data such as the one
-presented in this paper, while showing it in the same page.
-
-In general, linking and finding correlations between user choices and
-performance is an interesting avenue to explore in the future.
 %  Even if

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Back to 8 pages
Revised structure of the main section
To make it a single section of experiments with subsections. Also revises intro and thus closes #97 if nobody else has anything to say.

@@
 \title{Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources 
-for Low Cost Distributed Evolutionary Computation }
+for Low Cost Distributed Evolutionary Computation}
 %
@@
 computational systems along with attempts to predict and model its
-behavior  in Section \ref{sec:soa}. The first results in a volunteer
-setting are described 
-in Section \ref{sec:exp1}, and experiments using a new client
-architecture that uses web workers in Section \ref{sec:w2} and then
-applied to a different problem in Section \ref{sec:rastrigin}. 
+behavior  in Section \ref{sec:soa}. A brief presentation of the
+methodology used and the results of the experiments are presented in
+Section \ref{sec:exp1}, including new results using web workers on the
+same problem and on a new problem, Rastrigin's or F15 function. 
 Finally, conclusions and future lines of work are presented in Section
@@
 \cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}
-that even
-used the client's GPU \cite{duda2013gpu}. 
+
+% that even
+% used the client's GPU \cite{duda2013gpu}. 
 %Also a bit out of scope. Have to reduce the paper. 
@@
 
-\section{Experimental results}
+\section{Methodology and experimental results}
 \label{sec:exp1}
@@
 
-We set up different experiments by deploying the server in a cloud
-service, which changed in every one. This server runs a web service
-which stores chromosomes sent from clients and sends them back. It is
+We set up different experiments by deploying the server in free cloud
+services, changing it from one experiment to the next if needed. This server runs a web service
+which stores chromosomes sent from clients and serves requests to
+them; it is 
 an open source system that has been described extensively in
@@
 The results reported above indicated that we should look for a way of
-increasing the time that every user devotes. This led us to create an
+increasing the time users spend on the page. This led us to create an
 architecture that uses Web Workers, which is described in
@@
 
-\section{Working on a hard optimization problem}
+\subsection{A hard optimization problem: Rastrigin's function}
 \label{sec:rastrigin}

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Please @JJ explain better the objective and the measures we are going to use in the abstract
Minor English corrections in the intro

@@
 % No, we don't. Please read the paper!!!! - JJ
+% FERGU: Sorry, JJ, but the objective is not clear to me yet here, 
+% even if the implementation comparison I wrote is in the conclusions as the first 
+% objective of the paper. Also, the metrics used should be mentioned in the abstract.
+% For example: "The first objective of this paper is to compare different implementations 
+% of NodIO (with and without Web Workers), measuring the time spent by the users. 
+% The second objective is to model the user behavior by comparing whatever".
+% Also, the two compared implementations (w and w/ Web Worker) should be mentioned.
 From these experiments we can already draw some conclusions, besides
@@
 JavaScript. However, one of the advantages of our architecture is that
-clients do not need to install anything in their machines, they just
+clients do not need to install anything on their machines, they just
+%Pablo: According English Stack Exchange, you install "ON" a machine (but "IN" a directory)
 need to open a web page  
 in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
-available in almost every desktop or mobile device. Other advantage is
+available in almost every desktop or mobile device. Another advantage is
 that the source code can be directly examined by volunteers, because it is not
@@
 independence. This system, however, is not intended to substitute
-frameworks running in dedicated hardware or the cloud, but to
+frameworks running on dedicated hardware or the cloud, but to
+%Pablo: same here, I learnt this with my paper of EAs "on clusters"
 complement it. In general, several volunteers will be needed to
@@
 
-In this paper we will carry out three experiments, after establishing
+In this paper, we will carry out three experiments, after establishing
 the baseline of the performance that can be expected in a JavaScript
@@
 workers, that is, it runs in the background and it restarts. Third, we use the same
-client but with a harder problem that cannot be solved by a single
+client, but with a harder problem that cannot be solved by a single
 volunteer and thus needs the contribution of many volunteers to

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Comments at the end of the introduction

@@
 workers, that is, it runs in the background and it restarts. Third, we use the same
+% Pablo: it restarts when the solution is found (it has not been mentioned yet)
 client, but with a harder problem that cannot be solved by a single
@@
 have an influence on it. 
-
-So, the rest of the paper is organized as follows: Next we present the
+%Pablo: To do this we compare the time spent by each user because...
+
+So, the rest of the paper is organized as follows: next we present the
 state of the art in web-based distributed
@@
 methodology used and the results of the experiments are presented in
+%Pablo: mention the problems used in the first experiments
 Section \ref{sec:exp1}, including new results using web workers on the
-same problem and on a new problem, Rastrigin's or F15 function. 
+same problem and on a new problem, Rastrigin or F15 function. 
+%Pablo: Rastrigin is not the owner of the function
 Finally, conclusions and future lines of work are presented in Section

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Minor corrections and comments for conclusions

@@
 to the experiment simultaneously, so that our objective was
-achieved. In general, one of the first conclusions for this paper is
-that the browser technologies should be used to its full extent so
+achieved. In general, one of the first conclusions of this paper is
+that the browser technologies should be used to its full extent, so
 that the user time is leveraged to its full potential, and this was
@@
 first attempt to try and predict performance. As should be expected,
-the model and its parameters depends on the implementation, with contributions following
+the model and its parameters depend on the implementation, with contributions following
 a GEV distribution with single-thread and a
@@
 own social networks, personal colleagues, friends 
-and acquaintances, therefore to extend this work other options 
+and acquaintances, therefore, to extend this work other options 
 should be considered in order to include 
 a wider audience with different motives for participation.
-Another threat is that participants were anonymous so the amount of repeated 
+Another threat is that participants were anonymous, so the amount of repeated 
 participations and the effects this might have had in the outcome is not
@@
 % \url{https://github.com/JJ/modeling-volunteer-computing}.  
+% Pablo: I should uncomment this last paragraph, but using the url: https://github.com/ANONYMOUS
 

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Intro
Please check diff :)

@@
 a volunteer computing framework called {\sf NodIO}, designed to be
-easily deployable in using free cloud resources. We use  
+easily deployable on freely available cloud resources. We use  
 %Pablo: "deployable in using free cloud resources" is correct?
 % why not? - JJ
+% is easy deployable? yes! where? in free cloud resources! 
+% is easy deployable? yes! when?  (when) using free cloud resources
+% is one or the other, I like "on"      
 different implementations to find out which one is 
@@
 these assumptions, such as {\sf NodIO} presented in
-\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of
+\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of a
 client/server application and has some support to program an
 evolutionary algorithm embedded in pages served by them which can then
-communicate via that server. The client part can run in any
+communicate each other via that server. The client part can run in any
 web-enabled device, including many over which developers have no
@@
 %Pablo: same here, I learnt this with my paper of EAs "on clusters"
-complement it. In general, several volunteers will be needed to
+complement them. In general, several volunteers will be needed to
 overcome the overhead involved in using a client-server architecture
@@
 
-One of the aspects that must be taken into account in volunteer based
-frameworks is its {\em social} aspect in the design, with 
+%Mario: Rewrote: 
+%One of the aspects that must be taken into account in volunteer based
+%frameworks is its {\em social} aspect in the design, with 
+The {\em social} aspect must be taken into account when designing 
+volunteer based frameworks, with 
 issues related to security, trust, and privacy among others. 
@@
 certain safety in the sense that its server is not going to be
-attacked via cheating or denial or service attacks. We have decided to
+attacked via cheating or denial of service attacks. We have decided to
 approach this in a pragmatic way. Including many security features
@@
 resulting from the experiment once private information, such as IPs,
-has been processed to a untraceable hash. The experiments themselves
+has been processed to an untraceable hash. The experiments themselves
 are a proof that there are enough users trusting the experiment to
@@
 it stop working because there was no incentive in doing so. The
-system, besides, is running in a dedicated Platform as a Service
+system, besides, is running in a dedicated Platform as a Service (PaaS)
 container or Virtual Private Server so accessing the server would only
-show a few data obtained from the experiment, nothing more. Making it
+show the few data obtained from the experiment, nothing more. Making it
 free software ensured this was known by anyone in the experiment, via
@@
 based system and its comparison with other languages in
-\cite{2016arXiv160101607Manom}. First, several experiments using a
-Platform as a Service, and a simple Ajax client that does not
-restart. Second, we change a bit the client so that it uses web
-workers, that is, it runs in the background and it restarts. Third, we use the same
+\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
+restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and it restarts. Third, we use the same
 % Pablo: it restarts when the solution is found (it has not been mentioned yet)
-client, but with a harder problem that cannot be solved by a single
+client, but with a harder problem that cannot be easily solved by a single
 volunteer and thus needs the contribution of many volunteers to
-proceed. With these three experiments we want mainly to try and model
-the behavior of users and how different problems and implementations
-have an influence on it. 
+proceed. 
+%Mario: refrased: 
+%With these three experiments we want mainly to try and model
+With these three experiments our main objective is to try and model
+the behavior of users and learn how different problems and implementation choices have an influence on it.
+
 %Pablo: To do this we compare the time spent by each user because...
@@
 state of the art in web-based distributed
-computational systems along with attempts to predict and model its
+computational systems along with attempts to predict and model their
 behavior in Section \ref{sec:soa}. A brief presentation of the
@@
 Section \ref{sec:exp1}, including new results using web workers on the
-same problem and on a new problem, Rastrigin or F15 function. 
+same problem and on a new problem, a modified Rastrigin (F15) function. 
 %Pablo: Rastrigin is not the owner of the function
+%Mario: In many places is called  Rastrigin's function because it was proposed by him.
 Finally, conclusions and future lines of work are presented in Section

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
More sections

@@
 the more general context of distributed computing 
-\cite{nogueras2015studying} or be included in a more general study of the
+\cite{nogueras2015studying} or be included in a broader study of the
 performance of the EA itself
@@
 \cite{javadi2009mining} found out that the Weibull, log-normal, and
-Gamma distribution 
+Gamma distributions 
 modeled quite well the availability of resources in several clusters
@@
 paper we will perform different experiments in different setups and
-will try to come up with a more precise model. Next we describe the
-framework itself.
-
+will try to come up with a more precise model. 
+%Mario: erase: Next we describe the framework itself.
 As far as we know, this paper presents one of the few experimental
@@
 algorithm written in JavaScript, that after a number of generations
-sends back the best individual to the server, requesting at the same
+sends the best individual back to the server, requesting at the same
 time a random individual of those stored in the server. The experiment
@@
 
-\subsection{Initial tests}
+\subsection{Initial tests with a simple Ajax client}
+%% Above we mention three experiments:  Simple Ajax client, Web Workers, Harder Problem
+%% Im renaming subsection to  reference the first experiment
 
 Initial tests with an evolutionary algorithm solving the 40-Trap
-problem were set up using the free PaaS (Platform as a
-Service) OpenShift; they are described in
+problem were set up using the free PaaS OpenShift; they are described in
 Table \ref{tab:summary:os}. In general, every tests took several
 days and during it the problem was solved more than 50 times with
-several hundreds different IPs participating in it; the
+several hundreds different IPs participating in it; each
 evolutionary algorithm stopped when the solution was found. 
@@
 non-existent, that is, computing nodes are not {\em
-  staying} after the solution has been found. This is an issue that we
-will address in this paper.
+  staying} after the solution has been found. This is the issue that we
+address in the next experiments. 
+%Mario: changed: will address in this paper.
 
@@
 At any rate, this also shows that a convenient way of increasing the
-computing power would be increasing this minimum amount per user. This
+%Mario: As I understend is: 
+computing power would be increasing this minimum amount of experiments per user. This
 issue led us to create the next version of the framework, which is
@@
 4000 generations. All things considered, this means than there are
+%Mario: many more more threads {\em simultaneously} running?
 many more computers {\em simultaneously} running, leading to this

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Last changes

@@
 Section \ref{sec:exp1}, including new results using web workers on the
-same problem and on a new problem, a modified Rastrigin (F15) function. 
+same problem and on a new problem, a modified Rastrigin's function. 
 %Pablo: Rastrigin is not the owner of the function
@@
 
-\subsection{A hard optimization problem: Rastrigin's function}
+\subsection{A hard optimization problem: A Shifted Rotated Rastrigin's function}
 \label{sec:rastrigin}
@@
 but we needed to work on a harder problem, that is why created a new
-client version that optimized Rastrigin function, a multimodal
-floating-point problem used as a benchmark in evolutionary
+client version that optimized a harder variation of the Rastrigin's function, 
+used as a benchmark in evolutionary
 algorithms. The set up for this function, and a comparison of the
@@
 The results for this new experiment have been extremely interesting
-and quite different from the previous one. The most important
-characteristic that set this experiment apart from the previous one is
+and quite different from the others. The most important
+characteristic that set this experiment apart is
 the fact that it did not find the solution for the time it was
@@
 disconnected. The fitness of chromosomes sent by clients managed to
-surpass the 9000 value by the end of the; and although initially there
-was a high number of new nodes, this number decreases with time.
+surpass the 9000 value by the end of the execution; and although initially there
+was a high number of new nodes, this number decreased with time.
 
@@
 and hour (right). The two graphs have the same shape: At the beginning there are very few or only
-one, but later one there are peaks of almost 30 unique machine a
+one, but later one there are peaks of almost 30 unique machines a
 minute; these peaks appear again when new announcements are made in
@@
 will be users that will spend several hours, two orders of magnitude
-more generations in this case than in the previous one. There is a
-clear influence of the time to solution: a short time means that every
+more generations in this case than in the previous one.
+%Mario: than in the previous one. Are we refereing to the Wˆ2 version?
+There is a clear influence of the time to solution: a short time means that every
 experiment will only gather a few volunteers, but also that they will

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Making objectives clear and measurable
Thanks, @fergunet, for calling my bluff, but the objectives were already there, although not with measures. Made them clearer, added measures. Now, @unintendedbear, I do need you to take a look at everything. I have a meeting today.

@@
 to model the user's behavior so that those capabilities can be
-leveraged optimally. In this paper, we present the results of several
+leveraged optimally. These are the objectives of this paper, where we present the results of several
 evolutionary computation experiments with different implementations of
@@
 % is one or the other, I like "on"   
+% on it is. - JJ   
 different implementations to find out which one is 
 able to get the user to lend more computing cycles and test different
-problems to check the influence it has on said performance.
+problems to check the influence it has on said performance, as
+measured by the time needed to find a solution, but also by the number
+of users that it is able to gather and keep. 
 % Pablo: so the objective is "to determine which browser conf. is able
@@
 % The second objective is to model the user behavior by comparing whatever".
-% Also, the two compared implementations (w and w/ Web Worker) should be mentioned.
+% Also, the two compared implementations (w and w/ Web Worker) should
+% be mentioned.
+% The objectives are in the first sentence. I have changed the second
+% one to make it clearer - JJ
 From these experiments we can already draw some conclusions, besides

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Try to use 'secondly'... though this was an American paper OK. Anyway, try to avoid 'second' if you do not have a 'first'

@@
 two features: they are {\em desktop} applications, that is, they must
-be compiled or run directly on the operating system layer and, second,
+be compiled or run directly on the operating system layer and, in addition,
 they assume that the nodes that are going to run the system are under

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Comment for @JJ and please, next time we 'import' material from a bigger paper, erase previous comments or it will be impossible to review and to distinguish which are the new ones.
Couple of changes and connecting paragraphs.

@@
 these assumptions, such as {\sf NodIO} presented in
-\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of a
+\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of a % (Paloma) Only the server?
 client/server application and has some support to program an
 evolutionary algorithm embedded in pages served by them which can then
-communicate each other via that server. The client part can run in any
+communicate between each other via that server. The client part can run in any
 web-enabled device, including many over which developers have no
@@
 servers, and browsers as clients.
-% (Paloma) I think you should rewrite this whole sentence.
-% For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided... 
-% (Mario) We leverage our experience in blah blah to design
-% (Paloma) It may be... guess I was thinking in the noun, not in the
-% transitive verb
-% Please check if there's a problem here too.
+% (Paloma) It seems that you've started the introduction twice.
+The first two paragraphs "introduce the introduction", and so does the third.
+It is specially noticeable when you find "evolutionary computation (EC)" twice.
 
@@
 independence. This system, however, is not intended to substitute
-frameworks running on dedicated hardware or the cloud, but to
+already existing frameworks running on dedicated hardware or the cloud, but to
 %Pablo: same here, I learnt this with my paper of EAs "on clusters"
-complement them. In general, several volunteers will be needed to
-overcome the overhead involved in using a client-server architecture
-and this number of volunteers is obviously not guaranteed.
+complement them.
 
@@
 %frameworks is its {\em social} aspect in the design, with 
-The {\em social} aspect must be taken into account when designing 
+In general, several volunteers will be needed to
+overcome the overhead involved in using a client-server architecture
+and this number of volunteers is obviously not guaranteed. Therefore, the {\em social} aspect must be taken into account when designing 
 volunteer based frameworks, with 

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
More changes and more comments

@@
 these assumptions, such as {\sf NodIO} presented in
-\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of a % (Paloma) Only the server?
-client/server application and has some support to program an
-evolutionary algorithm embedded in pages served by them which can then
-communicate between each other via that server. The client part can run in any
+\cite{2016arXiv160101607Manom}. {\sf NodIO} implements a client-server application; the server offers the possibility to program an
+evolutionary algorithm that is embedded in the served pages, and those pages can communicate between each other via the server. On the other hand, the client part can run in any
 web-enabled device, including many over which developers have no
@@
 such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript. However, one of the advantages of our architecture is that
+JavaScript. However, one of the advantages of {\sf NodIO} architecture is that
 clients do not need to install anything on their machines, they just
@@
 that the source code can be directly examined by volunteers, because it is not
-compiled as other options such as Java Applets, ActionScript or
+compiled as other options such as Java Applets, ActionScript, or
 ActiveX are. Also, as we will discuss in this paper, it can provide some extra
@@
 The computing system becomes then a {\em techno-social system}
-\cite{vespignani2009predicting}. In that sense, its social aspects
+\cite{vespignani2009predicting}, in the sense that its social aspects
 cannot be separated from its technical specifications, with one
-feeding into the other. One of these aspects is safety for the user
-and the scientist running it: the user must be reasonably sure that the
-experiment is not misusing its resources and the scientist must have a
-certain safety in the sense that its server is not going to be
+feeding into the other. One of these aspects is safety for the user running it, as well as for the scientist: the user must be reasonably sure that the
+experiment is not misusing their resources and the scientist must have a
+certain safety in the sense that their server is not going to be
 attacked via cheating or denial of service attacks. We have decided to
-approach this in a pragmatic way. Including many security features
+approach this in a pragmatic way. Including many security features % (Paloma) I'm kind of lost here... Sentences seem to be unconnected and unrelated...
 would decrease performance; on the other hand, we publish all data

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Introduction has been reviewed

@@
 are a proof that there are enough users trusting the experiment to
-give it its CPU time and that there has been no attempt to cheat by
+give it their CPU time and that there has been no attempt to cheat by
 sending specially crafted chromosomes or to attack the server to make
-it stop working because there was no incentive in doing so. The
+it stop working because there was no incentive in doing so.
+% (Paloma) Is this really true?
+Can we really state that there were no attacks because there was no incentive?
+Attackers don't always need an incentive to perform an attack, or the incentive could not be perceived by us.
+And why do the experiments prove this?
+Are there specific measures?
+The
 system, besides, is running in a dedicated Platform as a Service (PaaS)
 container or Virtual Private Server so accessing the server would only
-show the few data obtained from the experiment, nothing more. Making it
+show the few data obtained from the experiment, adding security against attackers. In addition, making it
 free software ensured this was known by anyone in the experiment, via
-links in the web page that hosted it.
+links in the web page that hosted it. % (Paloma) What was known? The fact that it was running in a PaaS? Why was that known by links?
 
@@
 \cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
-restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and it restarts. Third, we use the same
+restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and then it restarts. Third, we use the same
 % Pablo: it restarts when the solution is found (it has not been mentioned yet)

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
State of the art has been reviewed

@@
 time other technologies,
- including Flash animations and VBScript, ActiveX or Java applets were
-used; for instance, Java was pointed out in \cite{soares1998get} as a
+ including Flash animations and VBScript, ActiveX, or Java applets were
+used with the same purpose; for instance, Java was pointed out in \cite{soares1998get} as a
 ``language for the internet'', providing some advantages such as multi-architecture compatibility or
@@
 computation experiments that rely on volunteer computing \cite{2016arXiv160101607Manom} could be achieved by using any of the
-languages and associated technologies, mentioned above, provided that the user had
+languages and associated technologies mentioned above, provided that the user had
 installed the required virtual machine plugin first.
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}.
 
@@
 
-On the other hand, we are also interested in measuring the performance
+Additionally, we are also interested in measuring the performance
 of volunteer computing systems, an area in which there have been
@@
 \cite{milani2004online}. This is actually not such a big problem in
-algorithms such as the EA that can easily be
+algorithms such as the Evolutionary Algorithm (EA) that can easily be
 parallelized via population splitting or by farming out evaluations to all
@@
 fault-tolerance of volunteer algorithms
-\cite{gonzalez2010characterizing} which it can, of course, be studied in
+\cite{gonzalez2010characterizing} which can, of course, be studied in
 the more general context of distributed computing 
@@
 
-On the other hand, measuring the performance of the resulting metacomputer
+Moreover, measuring the performance of the resulting metacomputer
 involves understanding its dynamics. Initial
@@
 are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}. Although SETI@home works
+results obtained in \cite{agajaj}. % (Paloma) I have a doubt here, should this reference be anonymised or not?
+Although SETI@home works
 at a larger scale in both quantity of volunteers and the scope of projects, 

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Little changes in the beginning of the methodology

@@
 them; it is 
-an open source system that has been described extensively in
+an open source system that has been extensively described in
 \cite{2016arXiv160101607Manom}. The client runs an evolutionary
-algorithm written in JavaScript, that after a number of generations
+algorithm written in JavaScript that after a number of generations
 sends the best individual back to the server, requesting at the same
 time a random individual of those stored in the server. The experiment
-starts by setting up a website that is sent by the server and announcing it on several social
-networks so that people contribute to them by going to the page and
+starts by setting up a website that is sent by the server and then announcing it on several social
+networks so that people contribute to it by visiting the page and
 staying there for a while.

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Some more changes

@@
 Initial tests with an evolutionary algorithm solving the 40-Trap
-problem were set up using the free PaaS OpenShift; they are described in
-Table \ref{tab:summary:os}. In general, every tests took several
-days and during it the problem was solved more than 50 times with
+problem were set up using the free PaaS OpenShift; the results are described in
+Table \ref{tab:summary:os}. In general, every test took several
+days and during its performance the problem was solved more than 50 times with
 several hundreds different IPs participating in it; each
@@
 
-The last column shows the percentage of experiments in which running the
+The second-to-last column in Table \ref{tab:summary:os} shows the percentage of experiments in which running the
 experiment in the browser takes {\em less} time than in the
-desktop. Although volunteer computing systems do not exclude desktop
+desktop. Volunteer computing systems do not exclude desktop
 computing and, in fact, the server can handle any client, not only
-those running JavaScript on the browser, we include it here for the
+those running JavaScript on the browser. Although, we include it here for the
 sake of comparison and to show that, being a free resource, it can
 anyway make a valid and valuable contribution to any distributed or
-desktop computing experiment. An interesting result is also that the
+desktop computing experiment. Another interesting result is that the
 correlation between users of an experiment and the next is
@@
 a Zipf's law, with similar distribution along all three runs. The 4/24
-run is the most complete and shows a S-shape, which implies an
+run is the most complete and shows an S-shape, which implies an
 accumulation of experiments taking similar time and around 100
@@
 just a few of them, and with diminishing probability as time
-decreases. This exponential distribution also appears in the
+decreases. This exponential distribution also appears in Figure \ref{fig:puts:os}, which shows the
 distribution of HTTP PUTs, equivalent to the number of
-generations divided by 100, contributed by every user, which is shown
-in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
+generations divided by 100, contributed by every user. These results show a Zipf-like behavior,
 so we have fitted it to the Generalized Extreme Value distribution,

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Adding a reference to the name of NodIO-W2 and erasing the explanation of number of PUTs, because it was already explained

@@
 architecture that uses Web Workers, which is described in
-\cite{2016arXiv160101607Manom}; web workers work in the background even if
+\cite{2016arXiv160101607Manom} and is called {\sf
+  NodIO-W$^2$}; web workers work in the background even if
 the web page is not in focus. A few improvements were also tested,
@@
 stopped contributing after one solution was found. This is also
-reflected in the number of PUTs, which is the number of generations
-divided by 100; more than 50\% of the volunteers run the experiment for more than
+reflected in the number of PUTs, more than 50\% of the volunteers run the experiment for more than
 4000 generations. All things considered, this means than there are

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Section 3.2 also reviewed

@@
 
-We are also interested in measuring the scaling properties in this
+We are also interested in measuring the scaling properties of this
 model, that is, the relationship between the number of IPs
@@
 it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
-experiment in terms of time (in milliseconds) vs. the number of workers participating
+experiment in terms of time (in milliseconds) vs. % (Paloma) Isn't "vs." informal?
+the number of workers participating
 in it. Although there is not a clear trend, the graph seems to say
@@
 standard deviation in the time, with more volunteers adding both {\em
-  robustness} to an experiment, and certainty in the time it is taking. 
+  robustness} to an experiment, and certainty in the time it is taking. % (Paloma) Should "certainty" go in italics? Otherwise it's confusing. 
  However, more experiments would be needed to check this hypothesis.
@@
 and many with few contributions. If we plot the model and the data
-side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
+side by side, as shown in Figure \ref{fig:gev:w2}, we see that the fit
 is not so tight as in the previous case, with the model overestimating
 the number of individuals with a high number contributions. This is
-probably due to the fact that in the previous case, the user needed to
+probably due to the fact that in the previous case the user needed to
 voluntarily reload the page to contribute the most generations, with a

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Little changes and comments

@@
 
+% (Paloma) I would propose as section title "A hard optimization problem: The Shifted Rotated Rastrigin's function"
+
 The fact that the solution of this particular problem was found in a
-few seconds and restarts were necessary yielded interesting results,
+few seconds and restarts were necessary yielded interesting results, % (Paloma) Where?
 but we needed to work on a harder problem, that is why created a new
@@
 experiments where the optimum was expected to be found many times, 
-this problem was expected to run constantly for many days finding only
+this problem was expected to constantly run for many days finding only
 sub-optimal approximations. In order to achieve this, the only change needed 
@@
 the size of the pool. But in a long-running experiment this has to be avoided
-in order to not exhaust the available memory. In this experiment the size
+in order to not exhausting the available memory. In this experiment the size
 of a chromosome was 1000 double precision numbers, so that considering this

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Addressing some issues
But at this stage when we have to submit the paper I would have preferred more direct changes than suggestions.

@@
 With these three experiments our main objective is to try and model
-the behavior of users and learn how different problems and implementation choices have an influence on it.
+the behavior of users and learn how different problems and
+implementation choices have an influence on it. That is why in this
+paper we will collect information on the number of users participating
+in each experiment and also the time they spent doing so. In this
+paper we will show how this participation follows certain patterns,
+and that they depend on the implementation itself as well as the
+problem and its difficulty; however, certain implementations allow to
+extract more CPU cycles from volunteers.
 
 %Pablo: To do this we compare the time spent by each user because...
+% Pablo, please, don't JJ me. Do the changes, because I'm not learning
+% now. I have added a bit anyway.
 
@@
 Other issues such as scheduling, which are related to adaptive 
-parallelism, were solved on the server. It is interesting to note that,
+parallelism, were solved on the server. But it is interesting to note that,
 even in the early stages of the web, the possibility of
@@
 %Mario: As I understend is: 
-computing power would be increasing this minimum amount of experiments per user. This
-issue led us to create the next version of the framework, which is
-presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
+computing power would be increasing this minimum amount of experiments
+per user. For this reason, we made some changes to the implementation,
+whose results are presented in the next subsection. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
 
 \begin{table*}[!htb]
-\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
+\caption{Summary of {\sf NodIO-W$^2$} (an acronym for the version that
+  uses web workers) and comparison with the previous
   experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
@@
 increasing the time users spend on the page. This led us to create an
-architecture that uses Web Workers, which is described in
-\cite{2016arXiv160101607Manom} and is called {\sf
+architecture that uses web workers, which is described in
+\cite{2016arXiv160101607Manom} and we abbreviate to {\sf
   NodIO-W$^2$}; web workers work in the background even if
@@
 now more than 43\%, making the number of IPs in contiguous experiments
-highly correlated. The main reason for this is the setup in {\sf
-  NodIO-W$^2$}, which restarts an island after a solution has been
+highly correlated. The main reason for this is the setup in the
+version using web workers, which restarts an island after a solution has been
 found and also that it keeps running even if the tab is not on the
@@
 %Mario: many more more threads {\em simultaneously} running?
-many more computers {\em simultaneously} running, leading to this
-almost 5-fold increase in running time.
+more threads and many more computers {\em simultaneously} running,
+leading to this almost 5-fold increase in running time.
 

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Minor fixes in experimental section

--- 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
@@
 seconds. The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, of the order of
+many experiments took a desirable amount of time, on the order of
 10 seconds, and which appears in all three graphs. As it can be seen,
@@
 predictive value of these fits, however, is limited, over all taking
+%Pablo: question for Paloma: is over all or overall?
 into account the low correlation between successive events shown
@@
 are also different and take a different time in each case, it is
-difficult to ascertain how this might have an influence in the
+difficult to ascertain how this might have an influence on the
 statistical distribution.
@@
 algorithms. The set up for this function, and a comparison of the
-baseline speed for JavaScript in an standalone experiment and other
+baseline speed for JavaScript in a standalone experiment and other
 languages is shown in \cite{2016arXiv160101607Manom}, where the details of
@@
 the size of the pool. But in a long-running experiment this has to be avoided
-in order to not exhausting the available memory. In this experiment the size
+in order to not exhaust the available memory. In this experiment the size
 of a chromosome was 1000 double precision numbers, so that considering this
@@
 chromosome with the server every 100 generations. Again, there were two workers in each window.
-As in previous experiments, a few calls to participation were published in social networks.
+As in previous experiments, a few calls to participation were published on social networks.
 
@@
 and quite different from the others. The most important
-characteristic that set this experiment apart is
+characteristic that sets this experiment apart is
 the fact that it did not find the solution for the time it was

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Managed to put it back to 8 pages
Closes #95
Closes #98

--- 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
@@
 % University of Granada, Granada, Spain\\
-% Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%
 
@@
 
-Some time ago, our group faced the problem of diminishing funds for
+We were interested in this kind of systems because, during the big
+global crisis, our group faced the problem of diminishing funds for
 buying new hardware. This was aggravated by the increasing maintenance costs and
@@
 servers, and browsers as clients.
-% (Paloma) It seems that you've started the introduction twice.
-The first two paragraphs "introduce the introduction", and so does the third.
-It is specially noticeable when you find "evolutionary computation (EC)" twice.
+% (Paloma) It seems that you've started the introduction twice. The
+% first two paragraphs "introduce the introduction", and so does the
+% third. It is specially noticeable when you find "evolutionary
+% computation (EC)" twice.
+% Yep, I want to say it all, but don't know how to put it together.
 
@@
 cannot be separated from its technical specifications, with one
-feeding into the other. One of these aspects is safety for the user running it, as well as for the scientist: the user must be reasonably sure that the
+feeding into the other. One of these aspects is safety for the user
+running it, as well as for the scientist: the user must be reasonably
+sure that the 
 experiment is not misusing their resources and the scientist must have a
 certain safety in the sense that their server is not going to be
-attacked via cheating or denial of service attacks. We have decided to
-approach this in a pragmatic way. Including many security features % (Paloma) I'm kind of lost here... Sentences seem to be unconnected and unrelated...
+attacked via cheating or denial of service attacks.
+
+We have decided to approach security in a pragmatic way. Including
+many security features % (Paloma) I'm kind of lost here... Sentences
+                       % seem to be unconnected and unrelated...  
+% Separated paragraph and substituting wildcard words - JJ
 would decrease performance; on the other hand, we publish all data
@@
 it stop working because there was no incentive in doing so.
-% (Paloma) Is this really true?
-Can we really state that there were no attacks because there was no incentive?
-Attackers don't always need an incentive to perform an attack, or the incentive could not be perceived by us.
-And why do the experiments prove this?
-Are there specific measures?
+% (Paloma) Is this really true? Can we really state that there were no
+% attacks because there was no incentive? Attackers don't always need
+% an incentive to perform an attack, or the incentive could not be
+% perceived by us. And why do the experiments prove this? Are there
+% specific measures? 
+% There were no attacks, that's a fact. There could be in the future,
+% but not for the time being - JJ
 The
@@
 show the few data obtained from the experiment, adding security against attackers. In addition, making it
-free software ensured this was known by anyone in the experiment, via
-links in the web page that hosted it. % (Paloma) What was known? The fact that it was running in a PaaS? Why was that known by links?
+free software ensured the meager bounty obtained by attackers were
+known by anyone participating in the experiment, via 
+links in the web page that hosted it. % (Paloma) What was known? The
+                                % fact that it was running in a PaaS?
+                                % Why was that known by links?
+%That there wasn't anything to hunt. Cleared, I expect.
 
@@
 based system and its comparison with other languages in
-\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
-restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and then it restarts. Third, we use the same
-% Pablo: it restarts when the solution is found (it has not been mentioned yet)
+\cite{2016arXiv160101607Manom}. First, several experiments are
+performed using a simple Ajax client that does not 
+restart when a solution is found. Second, we change the client a bit
+so that it uses web workers, that is, it runs in the background
+process and then it restarts when the solution is found. Third, we use the same
+% Pablo: it restarts when the solution is found (it has not been
+% mentioned yet)
+% Just do it! - JJ
 client, but with a harder problem that cannot be easily solved by a single
@@
 GEV & 35.759  &  343.336   & 9.877 \\
-Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
+Weibull & ND & 29.00 $\pm$ 2.15  & 0.40 $\pm$ 0.01 \\
 \hline

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
This will try to be the last change. Uploading
Reducing it to 8 pages again
We are on the edge, @fergunet. Every time you make a change you should check that it does not push it over the page limit.

@@
 
-Another area is how to enhance the number and quality of
-volunteers. For instance, adding a bit of more
+Other improvements could include adding a bit of more
 control to the user might contribute also to gamification. Right now

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
First version submitted
Still 16 hours to go, that is, until late tomorrow morning. I prefer to leave it this way, but if there's any major error, please let me know. Also check from the email you have received that it's done correctly.

thanks @fergunet @unintendedbear @mariosky @pacastillo for your help.

@@
 
-Using the browser as a computing resource in volunteer computing presents several advantages, 
-but it remains a challenge to fully harness the browser's capabilities and
-to model the user's behavior so that those capabilities can be
-leveraged optimally. These are the objectives of this paper, where we present the results of several
-evolutionary computation experiments with different implementations of
-a volunteer computing framework called {\sf NodIO}, designed to be
-easily deployable on freely available cloud resources. We use  
-%Pablo: "deployable in using free cloud resources" is correct?
-% why not? - JJ
-% is easy deployable? yes! where? in free cloud resources! 
-% is easy deployable? yes! when?  (when) using free cloud resources
-% is one or the other, I like "on"   
-% on it is. - JJ   
-different implementations to find out which one is 
-able to get the user to lend more computing cycles and test different
-problems to check the influence it has on said performance, as
-measured by the time needed to find a solution, but also by the number
-of users that it is able to gather and keep. 
-% Pablo: so the objective is "to determine which browser conf. is able
-% to get more computing power from the users", it isn't? 
-% (to keep this in mind the rest of the paper). What is a "browser configuration"?
-% (explain it with a phrase in the abstract)
-% No, we don't. Please read the paper!!!! - JJ
-% FERGU: Sorry, JJ, but the objective is not clear to me yet here, 
-% even if the implementation comparison I wrote is in the conclusions as the first 
-% objective of the paper. Also, the metrics used should be mentioned in the abstract.
-% For example: "The first objective of this paper is to compare different implementations 
-% of NodIO (with and without Web Workers), measuring the time spent by the users. 
-% The second objective is to model the user behavior by comparing whatever".
-% Also, the two compared implementations (w and w/ Web Worker) should
-% be mentioned.
-% The objectives are in the first sentence. I have changed the second
-% one to make it clearer - JJ
-From these experiments we can already draw some conclusions, besides
-the fact that volunteer computing can be a valuable computing
-resource: the user has to be kept engaged in some way to be able to
-use as many computing cycles as possible, the client has to be built
-to use the computer capabilities fully, and, finally, the user
-contributions follow a common statistical distribution. It is also
-essential to be as open as possible with software and data.
+Using volunteer's browsers as a computing resource presents several
+advantages, but it remains a challenge to fully harness the browser's
+capabilities and to model the user's behavior so that those
+capabilities can be leveraged optimally. These are the objectives of
+this paper, where we present the results of several evolutionary
+computation experiments with different implementations of a volunteer
+computing framework called {\sf NodIO}, designed to be easily
+deployable on freely available cloud resources. We use different
+implementations to find out which one is able to get the user to lend
+more computing cycles and test different problems to check the
+influence it has on said performance, as measured by the time needed
+to find a solution, but also by the number of users engaged.  From
+these experiments we can already draw some conclusions, besides the
+fact that volunteer computing can be a valuable computing resource and
+that it is essential to be as open as possible with software and data:
+the user has to be kept engaged to obtain as many computing cycles as
+possible, the client has to be built to use the computer capabilities
+fully, and, finally, that the user contributions follow a common
+statistical distribution.
 

### 2016-02-28 Alife16/paper.tex
Starting new paper

\documentclass[letterpaper]{article}
\usepackage{natbib,alifeconf}
\usepackage{hyperref}
\usepackage{url}

%\usepackage[pdftex]{graphicx}
\graphicspath{{../img/}}
\DeclareGraphicsExtensions{.pdf}

\title{ Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
for Low Cost Distributed Evolutionary Computation}
\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$\\
\mbox{}\\
$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada \\
$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana \\
{\tt jmerelo@ugr.es}, {\tt mario@tectijuana.edu.mx}}

\begin{document}
\maketitle

\begin{abstract}
Volunteer computing is extensively used 
in science and industry, but it
still has a long way to go in order to become
 a mainstream
technology in the same way the cloud technologies in general have. The
lack of 
general purpose and open source tools with built-in analytics 
could be one of the reasons of why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, entirely written  
in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
presented, this version leverages the HTML 5 Web Workers specification
to execute several clients per browser in the background, so that the the number of
generations per user could be improved. The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follows by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that {\sf NodIO} is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

\section{Introduction}
\label{introduction}
Nodays there is a big and increasing number of
distributed evolutionary computation software frameworks available on
the market, usually as open source systems \citep{Parejo12Survey}. 
Most of these frameworks have 
two features: they are {\em desktop} applications, that is, they must
be compiled or run directly on the operating system layer and, second,
they assume that the nodes that are going to run the system are under
the control of developers or they at least have access to them.

There is, then, space for frameworks that do not make any of these
assumptions, such as the one presented in this paper, which will be
 named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can be embedded in any web-enabled device, including many over which 
developers have no control.

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and
unconventional distributed evolutionary computing systems to 
design and release a new free framework that would allow anyone to
create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
servers and browsers as clients.

There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript \citep{2015arXiv151101088M}. However, one of the advantages of our system is that
clients do not need to install anything in their machines, they just
need to open a web page  
in a compatible browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX.
Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system independence.

When developing volunteer based frameworks, designers must
also consider the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. 
The computing system becomes a {\em techno-social system} \citep{vespignani2009predicting}.

In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the system
determine its performance and how it can be modeled and, eventually,
predicted.

The rest of the paper is organized as follows: Next we present the
state of the art (Section \ref{sec:soa}) in web-based distributed
computational systems along with attempts to predict and model its behavior. 
Then, a detailed description of the proposed framework is presented
in Section \ref{sec:description}, followed by the baseline experiments
in Section \ref{sec:experiments}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2}.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

%---------------------------------------------------------------
\section{State of the art}
\label{sec:soa}

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \citep{soares1998get} as a
``language for the
internet'', providing some advantages such as multi-architecture compatibility or
security mechanisms, and was proposed in \citep{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using compiled Java
programs. The ATLAS system
\citep{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper also establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely on
volunteer computing: scalability, heterogeneity, fault tolerance,
adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
reasonable performance. But hierarchy, this is, the capability of
preferentially using resources in the same organization, is a feature
that could be either disregarded or included within adaptivity, which
leaves us with eight desirable properties.

The fulfillment of these properties could be achieved by using any of the
languages and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first.

However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\citep{baratloo1996charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing systems PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Since the Java virtual machine is a sandbox, it is safe to the user
and as it runs compiled bytecode it also guarantees a certain level of
performance.
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. This
paper was later extended and published in a journal in 1999
\citep{baratloo1999charlotte}, but it is interesting to note that,
even in the early stages of the web, the possibility of
browser-based distributed computing was already researched.

Other systems, like JET, proposed by Soares et
al. \citep{soares1998get}, also used Java. JET supports
the execution of parallel applications over the Internet and features
a comprehensive statistics collection subsystem, enabling its use for science, 
since it allows 
the collection of algorithm analytics by the
experiment designer, unlike the other systems mentioned
above. Other proposals contemporary to JET, such as {\em SuperWeb}
\citep{alexandrov1997superweb} are roughly a translation of classical
distribution techniques, with overly complicated architectures that include
brokers for resource discovery. SuperWeb is focused,
however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different types of computers, mentioning
SPARCstations and PCs with Windows-95. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that arose later on. 
On the other hand, the potential for volunteer computing using
browsers was realized
later on \citep{sarmenta-bayanihan} as well as the potential of
user misconduct when having access to the source code.
\citep{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
less widespread.

The main problem with Java and all the other browser-embedded
technologies is that they are not {\em universal} in the
sense that an extra component, namely, the Java virtual machine or another
virtual environment has to be installed in the browser.
Also these virtual environments have become targets of malicious attacks; 
for instance numerous vulnerabilities have been discovered in 
the proprietary Adobe Flash runtime environment and
video player \citep{ford2009analyzing,watanabe2010new}. The need for
these additional environments has been mitigated with the
adoption of the HTML5 standard \citep{anthes2012html5}.
JavaScript \citep{flanagan2006javascript}, however, has become a set of standards
\citep{ECMA-262} that include the language and its components. Since
early on, it was adopted by the industry and also by scientists,
who used them for creating an evolutionary algorithm (EA) on
the browser as early as 1998 \citep{jj-ppsn98}.

JavaScript can be used either for unwitting
\citep{unwitting-ec} or volunteer 
\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \citep{duda2013gpu} or create an object-oriented
framework for evolution in the browser \citep{EvoStar2014:jsEO}. Many other researchers have
used Java \citep{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

Recent works propose the use of cloud computing services for the components of
client-server distributed EAs. Cloud resources
can reduce operational costs by outsourcing both hardware and software maintenance
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
is available in most laboratories, allowing researches to
scale algorithms at reasonable costs. Researchers have proposed the use of
public clouds such as Amazon EC2 \citep{CloudScale}, Google App Engine\citep{di2013towards},
and DropBox~\citep{mericloud}. The use of JavaScript to implement
server side components
gives software developers certain advantages in the cloud
paradigm, since it is a first-class citizen
in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \citep{wood13:nodejs:paas}. Many web services
have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
developers can use the same software tools for both client and server
side development.

JavaScript is also compatible with the functional programming paradigm \citep{Cousineau1998,MacLennan1990,Thompson1996}.
According to \citep{swanresearch2015}, the benefit of using this kind of paradigm in
metaheuristics optimization is that it allows an improvement in communicability,
reproductibility, interoperability, automated assembly, knowledge engineering
and efficiency.

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \citep{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
the expertise acquired in the library itself.

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}.

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}.

But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\citep{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\citep{javadi2009mining} found out that the Weibull, log-normal and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \citep{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
framework itself.

As far as we know, this paper presents one of the few experiments using 
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. 
%The methodology used
% (Paloma) I would say "the used methodology"
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

\section{Description of the framework}
\label{sec:description}

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\citep{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

As a volunteer based system, there are some issues worth 
noting \citep{sarmenta2001volunteer} :\begin{enumerate}
\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
each request is known.
\item As anonymous entities, volunteers are not accountable. 
If a volunteer misbehaves in some way the provider cannot 
prosecute or discipline the volunteer. In the case of JavaScript,
this issue is aggravated because even if it is obfuscated, source code
and data is sent to clients. 
\item Participants must also trust the application provider, 
regarding invasion of privacy, the integrity of their devices, 
and intended application. 
\end{enumerate}

Two kind of threats are possible in this environment; first, since the
API is open and it is an open source framework, it is relatively easy to
sabotage the system, as indicated in \citep{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome.
% (Paloma) Where is the second kind of threat?
Another two different attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released. 
 We also release as open access the performed experiments \citep{DBLP:journals/corr/GuervosG15} 
 and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade the
performance of the system. This socio-technical approach is, indeed,
coherent with our whole approach to the study of a distributed volunteer
 system which can only be studied by approaching it from the
technical as well as the social, as in social media, angle.

In this sense, in this paper we propose the {\sf NodIO} framework, a
cloud or bare metal based volunteer evolutionary computing system
derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
 the {\sf NodEO} library, whose architecture has been
developed using JavaScript on the client as well as the server.
All parts of the framework are free and available with a free license
from \url{https://github.com/JJ/splash-volunteer}.

%PABLO: el siguiente párrafo justifica Javascript

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the ''CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \citep{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population, then it sends, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request), and requests a random
  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
\end{enumerate}

  This is the invariant part of the framework, but other than that,
  the algorithm can be run in many different ways: 
  stopping when a solution is found or, as we have done in this paper,
  using Web Workers. In fact, since it is a pool-based system such as
  the one described in \citep{LNCS86720702}, any kind of client that
  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API. But never heard of a client which "works" an API. Is this correct?
  written in any kind of language. However, in this paper we are going
  to focus on the dynamics and measurements of the volunteer part of
  the system.

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the user interface elements of the framework.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{system.pdf}
\caption{Description of the proposed system. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server.}
\label{fig:system}
\end{figure}

The researcher only has to change a function, which can be written in
JavaScript or in any other language with cross-compilation to
JavaScript \citep{web:compilersjs} to solve a different
problem. In this case the classical Trap function \citep{Ackley1987} has been
used as well as the Rastrigin's floating-point optimization
problem. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic}, is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a non-blocking single thread
  allows the service of many requests. In fact, a limit in the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \citep{gecco07:workshop:dcor}, which had a low scalability. 
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system, or even browser requirement as long
  as JavaScript is enabled: anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of {\sf NodIO} can also be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island without any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. This architecture 
  has been tested in other high demand systems with success, but 
  in the case of {\sf NodIO}, additional experiments are needed to asses 
  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
  %(Mario) Scalability of the communication system in particular?
\item Since the algorithm runs in the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources than the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which makes the operation totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL, available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and measured performance by comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and checking the continuity
of the experiments.

All these points will be established in the next two sections,
starting with the baseline, single worker system next.

%---------------------------------------------------------------
\section{Baseline Experiment and obtained results} %Cambié el nombre
\label{sec:experiments}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO}
\citep{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

These two experiments establish a baseline result and show that
the time to success depends on the population size, with a bigger population
contributing to have more diversity and thus speeding up the solution \citep{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
experiments that we will describe next do not and can not have the
same conditions, but
the baseline is that if they eventually take longer than a basic
desktop, their interest will be purely academic. We will try and
prove next that that target performance can be achieved by carrying
out several experiments in different conditions. Second, they also act
as a basic performance test for the server, and also prove that the
basic server can be used for many kinds of different setups, as long
as they follow the REST protocol to deposit and obtain individuals
from the pool. It is obvious that, in this case, the server is not
really needed, since there is a single client and no interaction
between them, but there is a cost in sending and obtaining chromosomes
from the server so it is included for the sake of fairer comparison.

\subsection{Volunteer evolutionary computation: first experiments and results}

Initial experiments were set up using the OpenShift
PaaS, which provides a free tier within which this 
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible.
%
\begin{table*}
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPs and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-exper. correlation\\
\hline
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}, which shows the median number of IPs
intervening in each experiment,  median time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
previously obtained \citep{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment. The
maximum number of different IPs for each experiment is also in the
same range and of the order of 10, which is also consistent with
prior work. The median time has a big range of variation, but 50\% of
the time takes less than several minutes, from around 4 minutes in the
best case to roughly 2/3 of an hour in the worst case. That figure is
not competitive, {\em a priori}, with the baseline experiment, that is
why it is interesting to look at the columns which measure
the percentage of experiments in which the solution was found in less
time than the baseline. It is similar in all three cases, with around
20\% on average for the first baseline experiment, roughly 3\% in the
second case. The comparison is not totally fair: this experiment takes
place in a browser, and in fact most of the time is devoted to
presenting the charts; the server is also not the same since it is run in a
PaaS instead of a high-performance desktop computer, and it is also a
remote server, not a local server as in the baseline experiment. But
even in this case, we were looking for creating a framework that
allowed us to run evolutionary computation experiments {\em faster} than
in our very own computer. And the results indicate that although it will
happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
experiment you choose, it will not happen always or even most of
the time. One of the problems is shown in the last column, which
indicates the statistical correlation between the number of IPs
participating in one experiment and the next. In all cases,
correlation is so low as to affirm that there is almost no
relationship between them, that is, computing nodes are not {\em
  staying} after the solution has been found. This is a feature of
this implementation: the user has to actively reload the page to make
the evolutionary algorithm start again. Thus, the user can decide if to 
continue sharing resources or not. % (Paloma) Feature or disadvantage?
However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
unnecessary upper limit on the time, or number of experiments, that a user
will voluntarily perform.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
\caption{Duration of experiments vs. number of different IPs (nodes)
  participating in it, with averages and standard deviation shown as
  red dots; in the case there is a single red dot, there was a single
  experiment in which many computers participated (for instance, 16
  computers in the experiment in the far left or 29 in the middle
  one). 
Shades of blue indicate how many experiments included that many unique IPs,
so lighter shade for a column of dots indicates that a particular number
of computers happened less frequently, while darker shadow means more frequency. 
From left to right, experiments 4/4, 4/24 and 7/31.}
\label{fig:duration}
\end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
\caption{Duration of experiments vs. rank, with $y$ axis in a
  logarithmic scale. Dot color is related to the number of IPs
  participating in the experiment. From left to right, experiments
  4/4, 4/24 and 7/31.} 
\label{fig:zipf:os}
\end{figure}
%
We will have to analyze experimental data a bit further to find out why
this happens and also if there are some patterns in the three sets of
experiments. An interesting question to ask, for instance, is if
by adding more computers makes the experiment take less. In fact, as
shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
not seem to contribute to decreasing the time needed to finish the
experiment. However, the cause-effect relationship is not clear at
all. It might be the opposite: since experiments take longer to finish
and might in fact be abandoned with no one contributing for some time,
that increases the probability of someone new joining them. In fact,
with experiments taking a few seconds and due to the way the
experiments are announced, it is quite difficult that several
volunteers join in in such a short period of time, even more if we take
into account that volunteers are not {\em carried over} from previous
experiments. This implies that it would be convenient to use a problem
of a bigger size to check this hypothesis as we have done in the next
experiment. 
%  however, at this point we
% have not found this convenient since there are several other issues
% that have to be solved, as it will be shown next. %Creo que este párrafo
%saldría sobrando ahora que hicmios un experimento más grande

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{puts-openshift-4-4.png}
\includegraphics[width=0.32\linewidth]{puts-openshift-4-24.png}
\includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
\caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
  Extreme Value distribution (in red). From left to right, experiments
  4/4, 4/24 and 7/31.} 
% Plotted with ../data/plot-zipf-openshift.R
\label{fig:puts:os}
\end{figure}
%
It is also interesting to check the distribution of the experiment
duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
a Zipf's law, with similar distribution along all three runs. The 4/24
run is the most complete and shows a S-shape, which implies an
accumulation of experiments taking similar time and around 100
seconds. The most interesting part is the {\em tail}, which shows how
many experiments took a desirable amount of time, of the order of
10 seconds, and which appears in all three graphs. As it can be seen,
it sharply drops implying there are 
just a few of them, and with diminishing probability as time
decreases. This exponential distribution also appears in the
distribution of HTTP {\tt PUT}s, equivalent to the number of
generations divided by 100, contributed by every user, which is shown
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
so we have fitted it to the Generalized Extreme Value distribution,
with the resulting parameters shown in Table \ref{tab:puts:os}.
%
\begin{table*}
\caption{Summary of fit to Generalized Extreme Value distribution of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|ccc}
\hline
Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline
4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
0.1377 \\
4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\
7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
This distribution was originally proposed to fit extreme values
\citep{resnick2013extreme} and contains, as a special case, the inverse
Weibull distribution which was fitted to volunteer computing
frameworks such as SETI@home \citep{javadi2009mining}. We obviously do
not pretend to compare our system in scale or complexity with it, but
to point out that the behavior of volunteer computing nodes follows a
certain pattern, found in SETI@home, and which also appears in our system. This
distribution is governed by three parameters, the usual location $\mu$
which is related to where it has its {\em center} and a scale $\sigma$,
related to the size, but also a third shape $\xi$ parameter that is
related to its skewness, that is, how skewed it is around the central
location. Positive parameters indicate that the distribution {\em
  leans} towards the origin, and negative ones towards the other extreme
value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
greater than one and between one and 1.4, which indicates that the
three experiments share this origin-leaning pattern, with many users
donating a few cycles and just a few donating extreme values. Random
distributions with these parameters have been plotted in red in
Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
predictive value of these fits, however, is limited, over all taking
into account the low correlation between successive events shown
above.

At any rate, this also shows that a convenient way of increasing the
computing power would be increasing this minimum amount per user. This
issue led us to create the next version of the framework, which is
presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?

\section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
\label{sec:w2}

In order to improve the number of cycles per user, we enhanced the
basic architecture in several different aspects.  One improvement 
involved restarting the client once the solution was
found, so that as long as the user was visiting the page it was
running an evolutionary algorithm. Another was a small algorithmic
enhancement: instead of having a fixed population,
population size was randomly distributed between 128 and 256 and
different for each client.  The other improvement was to 
use the Web Worker API as described next.

JavaScript has been commonly used to develop client-side scripts
that handle the user interface, communicating asynchronously with servers and
updating the content that is displayed \citep{flanagan2006javascript}.
Long-running scripts like those proposed in this work can interfere with the
responsiveness of the web application, because everything is handled by a
single thread. The Web Workers specification \citep{hickson2012web} defines an
API that allows Web developers to execute long-running scripts in parallel
with the scripts of the main page. Worker scripts are spawned in the
background allowing a thread-like operation using message-passing as the
communication mechanism between the worker's execution environment and the
main thread. According to the Web Workers W3C Candidate Recommendation
\citep{hickson2012web} workers are expected to be long-lived, they have a high
start-up performance cost, and a high per-instance memory cost. The Web Worker
API is implemented in current versions of both desktop and mobile web browsers.
For volunteer computing using the Web Worker API brings several advantages
over a single threaded implementation:

\begin{itemize}
\item If the browser uses a tabbed document interface, the worker script
keeps running in the background even in case the user brings forward another tabbed
window.
\item Several evolutionary algorithms can be executed in parallel in a single web
page. 
\item Many implementations of the Web Worker API can use multiple-core CPUs for
the execution of worker scripts, in particular the execution in multiple-core CPUs
was tested in desktop versions of Chrome, FireFox and Safari.
\item Each {\sf NodEO} instance can be restarted independently.
\end{itemize}

A sequence diagram of the enhanced NodIO, which is called NodIO-W$^2$, is presented
in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
interaction between processes in a time sequence. In this system there are
two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
the client and the server where PUT and GET HTTP methods are represented by
red and blue arrows respectively. There are also messages between workers and
the main thread locally in the client, which are shown as thin solid arrows.
There are three processes involved:
\begin{itemize}
\item {\em Node.js} This is the server side process which is basically
  the same as used previously, except for some minor log-related
  changes. 
\item {\em Client main thread} The main script creates
worker instances and handles the user interface.
\item {\em Worker global scope} The worker execution space does not have
access to the Document Object Model (DOM) or JQuery objects in the main thread.
Asynchronous HTTP requests to the server are implemented using the
XMLHttpRequest object.  
\end{itemize}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=3in]{Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clients use Web Workers to run the evolutionary algorithm,
  with two of them per browser.}
\label{fig:system:w2}
\end{figure}
%
\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}

The sequence of events happens as follows:
\begin{enumerate}
\item First the {\sf NodIO} process is started in Node.js. As part of the
initialization procedure, the sequential number of experiments is initialized and a log
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP
request, then the page is loaded.
\item The main script renders the page and creates the worker instance. In
order to create a dedicated worker instance, another GET request is sent to the
server, then the client receives the second worker implementation script.
Each worker instance executes over time one or more evolutionary algorithms
{\em islands}. Each island starts with a random population and it is assigned a
universally unique identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.
The initial parameters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following the previously
specified parameters. Once the population has been initialized, the EA
generation loop is started. After {\em n} generations the state of the population
is evaluated. The chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not the global best,
the EA loop must continue, but first a chromosome is randomly selected from
the shared population. The current generation number and the best
chromosome are also sent to the client's main thread in order to update the DOM to
display the current state of the island (described next). These messages are
all asynchronous, so the EA loop of
the worker continues even before a message is received by the target. If the best fitness is a global best, the best
chromosome is sent to the server and the loop ends. A message is sent to the
main thread indicating that the best individual has been found.
\item The main thread receives messages from workers as events handled by
a callback function. When an iteration message is received, the page is
updated by displaying a dynamic plot of the generation number and best
fitness, and a representation of the chromosome is also generated.
\item When a global best is received from an island, the current experiment
ends, the experiment number is incremented, and the population array is reset.
\item When the main thread receives a message indicating that a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the start-up cost, only the parameters and population are reset and a new UUID
assigned.
\end{enumerate}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
\caption{Time employed in every experiment vs. number of IPs (in
  abscissas). The red dot and line show the average and standard
  deviation, the shade of blue the number of experiments with the same
  number of IPs. } 
\label{fig:ipstime:w2}
\end{figure}

The experiments were run in the same way as before, using several
Twitter announcements throughout several days by the end of July
2015. Eventually, more than one thousand experiments were completed in
a matter of days. A summary of results is shown in Table
\ref{tab:summary:ww}, comparing it with the aggregate of the results
obtained with the initial version of {\sf NodIO}. These results are
remarkably different, being similar only in the median and maximum
number of IPs, although in this case it would be combination
IP-worker, as we considered every worker as a unique IP. The median
number of seconds is almost 80\% better than the previous version, and
the median number of PUTs is three times better; this causes that most
of experiments take less time than one of the baselines, and one third
less than the second baseline. Thus, the conclusion in this case is
that we can obtain, through volunteer computation, a result that most 
of the times is better than we would by using a similar, non-parallel,
desktop setup, which makes this system suitable for massively
distributed evolutionary algorithms.
%
\begin{table*}
\caption{Summary of fit to GEV and Weibull distribution of
  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
\begin{center}
\begin{tabular}{cccc}
\hline
Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline
GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
0.04746 \\
Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%

However, it is interesting to note why this is so, and the first hint
is the inter-experiment correlation between the number of IPs in
successive experiments. While before it was an unremarkable 10\%, it is
now more than 43\%, making the number of IPs in contiguous experiments
highly correlated. The main reason for this is the setup in {\sf
  NodIO-W$^2$}, which restarts an island after a solution has been
found and also that it keeps running even if the tab is not on the
foreground. This means that volunteers can leave the experiment
running for as long as they want and they will be contributing 
experiment after experiment, unlike the previous version, where they
stopped contributing after one solution was found. This is also
reflected in the number of PUTs, which is the number of generations
divided by 100; more than 50\% of the volunteers run the experiment for more than
4000 generations. All things considered, this means than there are
many more computers {\em simultaneously} running, leading to this
almost 5-fold increase in running time.

We are also interested in measuring the scaling properties in this
model, that is, the relationship between the number of IPs
participating in an experiment and the time it takes to complete
it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
experiment in terms of time (in milliseconds) vs. the number of workers participating
in it. Although there is not a clear trend, the graph seems to say
that the time needed for finding the solution in this evolutionary
algorithm does not depend on the number of workers participating in
it. This does not mean that it is independent on the number of {\em
  simultaneous} users, which is probably the case although it is more
difficult to measure. There seems to be a trend towards a decreasing
standard deviation in the time, with more volunteers adding both {\em
  robustness} to an experiment, and certainty in the time it is taking. 
 However, more experiments would be needed to check this hypothesis.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
and Weibull (in red, left) for the {\sf NodIO-W$^2$} model.}  
\label{fig:gev:w2}
% File ../data/ips-puts.R
\end{figure}
%
As we did in the previous subsection, we have also fitted the number
of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
in the previous experiment, Table \ref{tab:puts:os}, we find higher values
for the location $\mu$ and scale $\sigma$ parameters, accounting for
the higher number of contributions per volunteer that has been
previously observed. However, the shape $\xi$ parameter is
substantially similar, with a {\em bell} shape leaning towards the
origin, indicating a pattern of a few users with many contributions
and many with few contributions. If we plot the model and the data
side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
is not so tight as in the previous case, with the model overestimating
the number of individuals with a high number contributions. This is
probably due to the fact that in the previous case, the user needed to
voluntarily reload the page to contribute the most generations, with a
{\em phase change} between the people that did so and the people that
just ran the experiment once until completion. In this case, however, the
user just needs to let it run, without that phase change and thus we obtain a
more {\em egalitarian} distribution of contributions, which rather
corresponds to a Weibull function, as shown in Figure
\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
equal to 0.69, a value that is remarkably similar to the 0.5 value
found for time devoted to games in Chambers et
al. \citep{chambers2005measurement}. It is important to note that, in
this case, the number of PUTs made per client does not correspond to a
uniform amount of computation, since 100 generations with a random
population belong, in each case, to different number of
operations. In other words, the same number of PUTs will correspond to
different number of operations in each case. However, since clients
are also different and take a different time in each case, it is
difficult to ascertain how this might have an influence in the
statistical distribution.

In general, this second version of the {\sf NodIO} framework and the
single experiment performed prove that it can be the foundation for a distributed
high-server-performance evolutionary computation platform,
providing reasonable algorithmic performance and being, in general,
easy to use with a straightforward modification of the fitness
function.

\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigin's Function}
JavaScript has been traditionally implemented as an interpreted
language not designed for the development of high performance
systems. On the other hand, current JavaScript  
virtual machines (VMs) are closing the gap, for instance the Google V8 
is an engine specifically designed for the fast execution of 
large JavaScript applications. In order to increase performance
V8 compiles JavaScript source code directly into machine code when it is first executed. 
There are no intermediate byte codes or an interpreter \citep{Gray:2009:GCM:1610564.1610565}.
JavaScript has gained popularity in server-side development using runtime
environments such as Node.js and is also used in desktop and mobile applications. 
%Mario: Creo que el párrafo de arriba sale sobrando
% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
To further evaluate how a current JavaScript implementation could be used by a
computer scientist to develop complex optimization problems,
a benchmark function, which was provided by the CEC2010 Special Session on
Large-Scale Global Optimization \citep{tang2007benchmark} is used in this section. 
The function is described next.

The basic Rastrigin's function is separable, and is defined as follows;
\begin{equation}
F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
\end{equation}
where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
is a classical multimodal problem. Such problem is difficult since the number of local
optima grows exponentially with the increase of dimensionality. To make it 
non-separable, an orthogonal matrix is also used for coordinate rotation.
The rotated Rastrigin's function is defined as follows:
\begin{equation}
F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
\end{equation}
Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
emulates real-world optimization problems that most likely will consist 
of different groups of parameters with strong dependencies within
but little interaction between the groups. This issue is reflected in the benchmark 
by randomly dividing the objective variables
into several groups, each of which contains a number of variables. The parameter $m$ 
is used to control the number of variables in each group and hence, defining the degree
of separability:
\begin{equation}
F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
\end{equation}
For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.

This optimization problem was selected because it is representative of the kind of
algorithms, data types, and structures employed in large scale optimization problems. 
The function was implemented in both Matlab and Java languages as part of the
Test Suite of the companion competition of the Special Session and to give competitors an idea of the
computational cost of the challenge, the runtime required for 10,000 function evaluations 
for a particular configuration was published by organizers. The Java
implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
3,000,000 function evaluations as termination condition. Using the Java implementation as
a guide, the function described above was implemented to compare the performance of JavaScript against
both implementations, again using the time required for 10,000 function evaluations. 
Before discussing the setup and performance results, certain details of
the implementation in JavaScript are presented to give readers an idea of 
how practical and mature this language is for developing complex mathematical functions.

\begin{itemize}
\item {\em Randomize} Both Matlab and Java programs rely on a 
Java Randomizer library for the 
randomization of the shift vectors and matrices used by the function.
The library includes the generation of pseudorandom Gaussians of type
double and uniformly distributed integers. Generators use a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
in the implementation of the standard {\tt Math.random()} between engines
and also their results are non-deterministic. 
The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript VMs.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime of functions is needed.
In JavaScript the {\tt Date class} is often used 
to measure execution time, but has the limitation of having a maximum 
resolution given in milliseconds.
In order to measure the runtime intervals with a higher precision two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
array, and for measuring in the browser the {\tt Performance.now()} function was called,
which uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome browsers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when choosing certain 
language to write non-trivial programs is the availability of development tools. While developing
the experiments, we used a comprehensive set of developer tools available for both
browsers and desktop \citep{tilkov2010node}. These tools included
package managers, in this case {\tt npm}, debuggers, logging libraries
such as {\tt winston}, 
network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
Although the fact that developer tools vary from browser to browser
supposed a drawback, they all have the same functionality. The
JavaScript debugger can be accessed directly from the user menu in
most browsers. 
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partially implement the functionality of
the {\tt StrictMath} library used by the Java implementation. If more precision
is needed, developers could use {\tt math.js}, an extensive math library with
support for matrices and big numbers.
\end{itemize}

We found that the JavaScript ecosystem is mature enough for developing non-trivial 
code, but developers have to consider the differences between VMs implementations
and therefore use certain libraries in order to increase the precision and repeatability of
algorithms. As for the performance of JavaScript against Java and Matlab
we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
 The performance of the time required for 10,000 function evaluations is presented in
Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
performance with an average of 935ms, followed by Java with
991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
The results also show that there is not much difference between the Node.js
implementation an that of a
browser. There is also not much difference between running the code in the main thread or in Web
Workers. When running the experiment in Firefox a dialog alerted that the
script was taking too much time, showing that this
kind of processing is not intended for the main thread. Additionally, there is not much overhead
when running two
Web Workers in parallel as they took 1279ms each. This experiment
shows that JavaScript is viable as a language
language for developing complex mathematical functions an these can run in a browser
with acceptable performance. 
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{f15_times.png}
\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
\label{fig:f15_times}
\end{figure}

\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}
This experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, considering this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.

\subsubsection{Results}

%
\begin{figure}[!htb]
\centering
\includegraphics[width=3in]{rastrigin-fitness.png}
\caption{Fitness of chromosomes sent to the server pool along time in
  the $x$ axis.} 
% Plotted with ../data/plot-rastrigin-fitness.R
\label{fig:puts:rastrigin}
\end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
\caption{Number of unique IPs participating in the Rastrigin
  experiment per minute (left) and per hour (right)} 
% Plotted with ../data/plot-rastrigin-IPs.R
\label{fig:ips:rastrigin}
\end{figure}
%
\begin{table*}
\caption{Summary of fit to GEV and Weibull distribution of
  the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
\begin{center}
\begin{tabular}{cccc}
\hline
Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline
GEV & 35.759  &  343.336   & 9.877 \\
Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
\caption{Fit to GEV (left) and Weibull (right) of the number of {\tt PUT}s per worker for the F15 (Rastrigin) function.} 
% Plotted with ../data/ips-puts-rastrigin.R
\label{fig:fit:rastrigin}
\end{figure}
%

The results for this new experiment have been extremely interesting
and quite different from the previous one. The most important
characteristic that set this experiment apart from the previous one is
the fact that it did not find the solution for the time it was
running, instead lowering, little by little, the fitness until it was
disconnected. The fitness of chromosomes sent by clients is shown in
Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
upper part, which represents the fitness of clients that have just
started evolution from a random population, and then the fitness of
those clients that have been already evolving for a while or obtained
a high-fitness individual from the pool and evolved from there, and
which managed to surpass the 9000 value by the end of the
simulation. It incidentally shows also the number of {\em new} nodes
coming: there is a dense {\em cloud} at the beginning, but the new
contributions keep thinning as the evolution progresses.

This is in part due to the fact that the presence of nodes is not
constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
number of unique IPs participating in the experiment per minute (left)
and hour (right). The two graphs have the same shape which matches
also the speed in which fitness is decreased in Figure
\ref{fig:puts:rastrigin}. At the beginning there are very few or only
one, but later one there are peaks of almost 30 unique machine a
minute; these peaks appear again when new announcements are made in
Twitter or any other social networks, to go down to a few remaining
nodes later on. It is interesting to note that, at the most, more than
100 unique nodes were participating in the experiment during one hour;
also that there are sustained peaks during a day of 6-7 computers
every minute. This is due to the fact that the computation is
asynchronous, but also to the fact that different computers stay in
the simulation for a different period of time. This period has been
plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
{\tt PUT}s per node, ranked by number, along with a fit to the two
distributions we have used before, GEV and Weibull; the parameters of
this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
experiments, in which GEV fit was relatively good, in this case it is
quite obviously not. However, the fit to the Weibull distribution,
shown at the left of the image, is relatively good. As should be
expected, the shape and scale are different to the one in the previous
experiments with the Trap function, however, the shape is quite
similar (0.40 vs 0.69), being in both cases less than one, which
indicates a decreasing distribution of the same kind. In fact, the
difference in the scale parameter indicates that the difference
between the time devoted by a volunteer and the next is bigger in this
case. This is probably due to the fact that the duration of the
experiment is bigger, allowing for bigger differences in
behavior. While the first in the Trap function might have devoted a
few minutes and the second a few seconds less, the difference between
the first and second in this experiment might be in the scale of
hours.

In general, this experiment confirms what was found in the others in
several aspects: first, there is a regularity in the distribution of the
time assigned by volunteers to it, second, there is a good and certain
amount of volunteers that will become part of the distributed
computer, in some cases up to one hundred; also, in some cases there
will be users that will spend several hours, two orders of magnitude
more generations in this case than in the previous one. There is a
clear influence of the time to solution: a short time means that every
experiment will only gather a few volunteers, but also that they will
stay only for a short time. A long time will make users stay for
longer, but if it takes too long, they will eventually lose interest.

We will discuss this and the rest of the findings of the paper in the
conclusions.

%---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

In this paper two versions of a client-server architecture for volunteer and distributed
evolutionary algorithms that uses the browser and generated using {\sf
  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
evaluated. Volunteers are {\em in the cloud}, as stated in the title,
since they are a {\em CPU as a service} for the persons running the
experiment. In fact, in this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative or in conjunction with other local or
cloud-based methods to add computing power in a seamless way through
the pool that NodIO creates.

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than the
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over volunteers to other experiments when the
one they were participating on was finished, shown in the low
correlation between the number of IPs in successive experiments. This
also resulted in a low number of generations allotted by users.

In the second implementation, {\sf NodIO-W$^2$}, Web Workers
were used so that several clients per browser could be executed and
they could run in the background, so that the the number of
generations per user could be improved. The number of generations per
user increased to the point that baseline performance was improved in
most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment at the same time, so that our objective was
achieved. In general, one of the first conclusions for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved with {\sf NodIO-W$^2$}.

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model depends on the implementation, with contributions following
a General Extreme Values distribution in the case of {\sf NodIO} and a
Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
close enough to each other and, in the second case, reflect the fact
that all time spent in the page is actually devoted to computing,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse might be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they've done it
once, they might do it more times. In the spirit of Open Science, this
gamification might involve computing in real time data such as the one
presented in this paper and showing it in the same page.

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar way, one
obtained up to 5 times more total cycles  than the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to explore in the future.

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for other computationally demanding
tasks, especially ones that cover the middle ground between a problem
finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool as was done in \citep{araujo2008mam},
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

Another area is how to enhance the number and quality of
volunteers. For instance, adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.

Finally, the implementation needs some refinement in terms of
programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
to create a generator in which the user just has to create a fitness
function, with the rest of the framework wrapped around
automatically. All this, as the data used for this paper and the paper
itself, has been published with a free license in GitHub at
\url{https://github.com/JJ/modeling-volunteer-computing}.

%---------------------------------------------------------------
\section*{Acknowledgment}

This work has been supported in part by TIN2011-28627-C04-02 and
TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
GENIL project (CEI-BIOTIC Granada). We would also like to thank the
anonymous reviewers of previous versions of this paper who have really
helped us to improve 
this paper (and our work) with their suggestions. We are also grateful
to Anna S\'aez de Tejada for her help with the data processing scripts.

\bibliographystyle{apalike}
\bibliography{volunteer,GA-general,geneura,javascript,ror-js}

\end{document}

### 2016-02-28 Alife16/paper.tex
Reduced to 11 pages a a bit
But still a way to go. Almost everything must go refs #99

@@
 for Low Cost Distributed Evolutionary Computation}
-\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$\\
+\author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
+  García-Sánchez $^1$, Mario Garc\'ia-Valdez$^2$\\
 \mbox{}\\
@@
 browser, as well as any other client consuming its REST application
-programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
-presented, this version leverages the HTML 5 Web Workers specification
-to execute several clients per browser in the background, so that the the number of
-generations per user could be improved. The architecture, along with some initial
+programming interface. The architecture, along with some initial
 results on the expected performance and scaling behavior when the
@@
 
-\section{{\sf NodIO-W$^2$}, a Web Worker-based architecture}
-\label{sec:w2}
-
-In order to improve the number of cycles per user, we enhanced the
-basic architecture in several different aspects.  One improvement 
-involved restarting the client once the solution was
-found, so that as long as the user was visiting the page it was
-running an evolutionary algorithm. Another was a small algorithmic
-enhancement: instead of having a fixed population,
-population size was randomly distributed between 128 and 256 and
-different for each client.  The other improvement was to 
-use the Web Worker API as described next.
-
-JavaScript has been commonly used to develop client-side scripts
-that handle the user interface, communicating asynchronously with servers and
-updating the content that is displayed \citep{flanagan2006javascript}.
-Long-running scripts like those proposed in this work can interfere with the
-responsiveness of the web application, because everything is handled by a
-single thread. The Web Workers specification \citep{hickson2012web} defines an
-API that allows Web developers to execute long-running scripts in parallel
-with the scripts of the main page. Worker scripts are spawned in the
-background allowing a thread-like operation using message-passing as the
-communication mechanism between the worker's execution environment and the
-main thread. According to the Web Workers W3C Candidate Recommendation
-\citep{hickson2012web} workers are expected to be long-lived, they have a high
-start-up performance cost, and a high per-instance memory cost. The Web Worker
-API is implemented in current versions of both desktop and mobile web browsers.
-For volunteer computing using the Web Worker API brings several advantages
-over a single threaded implementation:
-
-\begin{itemize}
-\item If the browser uses a tabbed document interface, the worker script
-keeps running in the background even in case the user brings forward another tabbed
-window.
-\item Several evolutionary algorithms can be executed in parallel in a single web
-page. 
-\item Many implementations of the Web Worker API can use multiple-core CPUs for
-the execution of worker scripts, in particular the execution in multiple-core CPUs
-was tested in desktop versions of Chrome, FireFox and Safari.
-\item Each {\sf NodEO} instance can be restarted independently.
-\end{itemize}
-
-A sequence diagram of the enhanced NodIO, which is called NodIO-W$^2$, is presented
-in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
-API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
-interaction between processes in a time sequence. In this system there are
-two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
-the client and the server where PUT and GET HTTP methods are represented by
-red and blue arrows respectively. There are also messages between workers and
-the main thread locally in the client, which are shown as thin solid arrows.
-There are three processes involved:
-\begin{itemize}
-\item {\em Node.js} This is the server side process which is basically
-  the same as used previously, except for some minor log-related
-  changes. 
-\item {\em Client main thread} The main script creates
-worker instances and handles the user interface.
-\item {\em Worker global scope} The worker execution space does not have
-access to the Document Object Model (DOM) or JQuery objects in the main thread.
-Asynchronous HTTP requests to the server are implemented using the
-XMLHttpRequest object.  
-\end{itemize}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=3in]{Algorithm.pdf}
-\caption{Description of the W$^2$ version of {\sf NodIO}. In this
-  version, clients use Web Workers to run the evolutionary algorithm,
-  with two of them per browser.}
-\label{fig:system:w2}
-\end{figure}
-%
-\begin{table*}[!htb]
-\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
-  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
-\begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
-{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-The sequence of events happens as follows:
-\begin{enumerate}
-\item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure, the sequential number of experiments is initialized and a log
-file is created. The shared pool implemented as an array is reset.
-\item A volunteer follows the link of the experiment triggering a GET HTTP
-request, then the page is loaded.
-\item The main script renders the page and creates the worker instance. In
-order to create a dedicated worker instance, another GET request is sent to the
-server, then the client receives the second worker implementation script.
-Each worker instance executes over time one or more evolutionary algorithms
-{\em islands}. Each island starts with a random population and it is assigned a
-universally unique identifier (UUID). The UUID is included in each of the HTTP
-requests sent to the server.
-The initial parameters for each island
-can be distinct. Once the worker is created and the parameters have been set,
-a message is sent to the worker to start the algorithm.
-\item The worker first creates the population following the previously
-specified parameters. Once the population has been initialized, the EA
-generation loop is started. After {\em n} generations the state of the population
-is evaluated. The chromosome with the best fitness is sent to the shared
-population with a PUT HTTP request. If the best fitness is not the global best,
-the EA loop must continue, but first a chromosome is randomly selected from
-the shared population. The current generation number and the best
-chromosome are also sent to the client's main thread in order to update the DOM to
-display the current state of the island (described next). These messages are
-all asynchronous, so the EA loop of
-the worker continues even before a message is received by the target. If the best fitness is a global best, the best
-chromosome is sent to the server and the loop ends. A message is sent to the
-main thread indicating that the best individual has been found.
-\item The main thread receives messages from workers as events handled by
-a callback function. When an iteration message is received, the page is
-updated by displaying a dynamic plot of the generation number and best
-fitness, and a representation of the chromosome is also generated.
-\item When a global best is received from an island, the current experiment
-ends, the experiment number is incremented, and the population array is reset.
-\item When the main thread receives a message indicating that a worker has ended the EA
-loop, that worker is reinitialized. The worker process is not ended, to
-avoid the start-up cost, only the parameters and population are reset and a new UUID
-assigned.
-\end{enumerate}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
-\caption{Time employed in every experiment vs. number of IPs (in
-  abscissas). The red dot and line show the average and standard
-  deviation, the shade of blue the number of experiments with the same
-  number of IPs. } 
-\label{fig:ipstime:w2}
-\end{figure}
-
-The experiments were run in the same way as before, using several
-Twitter announcements throughout several days by the end of July
-2015. Eventually, more than one thousand experiments were completed in
-a matter of days. A summary of results is shown in Table
-\ref{tab:summary:ww}, comparing it with the aggregate of the results
-obtained with the initial version of {\sf NodIO}. These results are
-remarkably different, being similar only in the median and maximum
-number of IPs, although in this case it would be combination
-IP-worker, as we considered every worker as a unique IP. The median
-number of seconds is almost 80\% better than the previous version, and
-the median number of PUTs is three times better; this causes that most
-of experiments take less time than one of the baselines, and one third
-less than the second baseline. Thus, the conclusion in this case is
-that we can obtain, through volunteer computation, a result that most 
-of the times is better than we would by using a similar, non-parallel,
-desktop setup, which makes this system suitable for massively
-distributed evolutionary algorithms.
-%
-\begin{table*}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker. \label{tab:puts:ww}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 18.275 $\pm$ 0.83719  &  25.962  $\pm$ 1.22015 & 1.242   $\pm$
-0.04746 \\
-Weibull & ND & 75.11457519 $\pm$ 2.97470251  & 0.69714410 $\pm$ 0.01383976 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-
-However, it is interesting to note why this is so, and the first hint
-is the inter-experiment correlation between the number of IPs in
-successive experiments. While before it was an unremarkable 10\%, it is
-now more than 43\%, making the number of IPs in contiguous experiments
-highly correlated. The main reason for this is the setup in {\sf
-  NodIO-W$^2$}, which restarts an island after a solution has been
-found and also that it keeps running even if the tab is not on the
-foreground. This means that volunteers can leave the experiment
-running for as long as they want and they will be contributing 
-experiment after experiment, unlike the previous version, where they
-stopped contributing after one solution was found. This is also
-reflected in the number of PUTs, which is the number of generations
-divided by 100; more than 50\% of the volunteers run the experiment for more than
-4000 generations. All things considered, this means than there are
-many more computers {\em simultaneously} running, leading to this
-almost 5-fold increase in running time.
-
-We are also interested in measuring the scaling properties in this
-model, that is, the relationship between the number of IPs
-participating in an experiment and the time it takes to complete
-it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
-experiment in terms of time (in milliseconds) vs. the number of workers participating
-in it. Although there is not a clear trend, the graph seems to say
-that the time needed for finding the solution in this evolutionary
-algorithm does not depend on the number of workers participating in
-it. This does not mean that it is independent on the number of {\em
-  simultaneous} users, which is probably the case although it is more
-difficult to measure. There seems to be a trend towards a decreasing
-standard deviation in the time, with more volunteers adding both {\em
-  robustness} to an experiment, and certainty in the time it is taking. 
- However, more experiments would be needed to check this hypothesis.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
-and Weibull (in red, left) for the {\sf NodIO-W$^2$} model.}  
-\label{fig:gev:w2}
-% File ../data/ips-puts.R
-\end{figure}
-%
-As we did in the previous subsection, we have also fitted the number
-of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
-in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
-in the previous experiment, Table \ref{tab:puts:os}, we find higher values
-for the location $\mu$ and scale $\sigma$ parameters, accounting for
-the higher number of contributions per volunteer that has been
-previously observed. However, the shape $\xi$ parameter is
-substantially similar, with a {\em bell} shape leaning towards the
-origin, indicating a pattern of a few users with many contributions
-and many with few contributions. If we plot the model and the data
-side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
-is not so tight as in the previous case, with the model overestimating
-the number of individuals with a high number contributions. This is
-probably due to the fact that in the previous case, the user needed to
-voluntarily reload the page to contribute the most generations, with a
-{\em phase change} between the people that did so and the people that
-just ran the experiment once until completion. In this case, however, the
-user just needs to let it run, without that phase change and thus we obtain a
-more {\em egalitarian} distribution of contributions, which rather
-corresponds to a Weibull function, as shown in Figure
-\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
-distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
-equal to 0.69, a value that is remarkably similar to the 0.5 value
-found for time devoted to games in Chambers et
-al. \citep{chambers2005measurement}. It is important to note that, in
-this case, the number of PUTs made per client does not correspond to a
-uniform amount of computation, since 100 generations with a random
-population belong, in each case, to different number of
-operations. In other words, the same number of PUTs will correspond to
-different number of operations in each case. However, since clients
-are also different and take a different time in each case, it is
-difficult to ascertain how this might have an influence in the
-statistical distribution.
-
-In general, this second version of the {\sf NodIO} framework and the
-single experiment performed prove that it can be the foundation for a distributed
-high-server-performance evolutionary computation platform,
-providing reasonable algorithmic performance and being, in general,
-easy to use with a straightforward modification of the fitness
-function.
-
-\subsection{$F_{15}:\frac{D}{m}$-group Shifted and $m$-rotated Rastrigin's Function}
-JavaScript has been traditionally implemented as an interpreted
-language not designed for the development of high performance
-systems. On the other hand, current JavaScript  
-virtual machines (VMs) are closing the gap, for instance the Google V8 
-is an engine specifically designed for the fast execution of 
-large JavaScript applications. In order to increase performance
-V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter \citep{Gray:2009:GCM:1610564.1610565}.
-JavaScript has gained popularity in server-side development using runtime
-environments such as Node.js and is also used in desktop and mobile applications. 
-%Mario: Creo que el párrafo de arriba sale sobrando
-% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
-To further evaluate how a current JavaScript implementation could be used by a
-computer scientist to develop complex optimization problems,
-a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization \citep{tang2007benchmark} is used in this section. 
-The function is described next.
-
-The basic Rastrigin's function is separable, and is defined as follows;
-\begin{equation}
-F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
-\end{equation}
-where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
-$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin’s function 
-is a classical multimodal problem. Such problem is difficult since the number of local
-optima grows exponentially with the increase of dimensionality. To make it 
-non-separable, an orthogonal matrix is also used for coordinate rotation.
-The rotated Rastrigin's function is defined as follows:
-\begin{equation}
-F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
-\end{equation}
-Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
-emulates real-world optimization problems that most likely will consist 
-of different groups of parameters with strong dependencies within
-but little interaction between the groups. This issue is reflected in the benchmark 
-by randomly dividing the objective variables
-into several groups, each of which contains a number of variables. The parameter $m$ 
-is used to control the number of variables in each group and hence, defining the degree
-of separability:
-\begin{equation}
-F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
-\end{equation}
-For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
-candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
-$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
-finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
-
-This optimization problem was selected because it is representative of the kind of
-algorithms, data types, and structures employed in large scale optimization problems. 
-The function was implemented in both Matlab and Java languages as part of the
-Test Suite of the companion competition of the Special Session and to give competitors an idea of the
-computational cost of the challenge, the runtime required for 10,000 function evaluations 
-for a particular configuration was published by organizers. The Java
-implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
-3,000,000 function evaluations as termination condition. Using the Java implementation as
-a guide, the function described above was implemented to compare the performance of JavaScript against
-both implementations, again using the time required for 10,000 function evaluations. 
-Before discussing the setup and performance results, certain details of
-the implementation in JavaScript are presented to give readers an idea of 
-how practical and mature this language is for developing complex mathematical functions.
-
-\begin{itemize}
-\item {\em Randomize} Both Matlab and Java programs rely on a 
-Java Randomizer library for the 
-randomization of the shift vectors and matrices used by the function.
-The library includes the generation of pseudorandom Gaussians of type
-double and uniformly distributed integers. Generators use a single
-{\tt long} seed. In order to implement the same functionality in
-JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
-in the implementation of the standard {\tt Math.random()} between engines
-and also their results are non-deterministic. 
-The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
-consistent results across all JavaScript VMs.
-\item {\em Timing functions} In order to compare the performance of 
-each implementation, an accurate and consistent way of measuring 
-the runtime of functions is needed.
-In JavaScript the {\tt Date class} is often used 
-to measure execution time, but has the limitation of having a maximum 
-resolution given in milliseconds.
-In order to measure the runtime intervals with a higher precision two functions
-where used: for Node.js we used the native {\tt process.hrtime()} function
-which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
-array, and for measuring in the browser the {\tt Performance.now()} function was called,
-which uses floating-point numbers with up to microsecond precision. 
-A drawback found is that the {\tt Performance.now()} function is only implemented
-in Firefox and Chrome browsers. Both functions are independent of the system clock. 
-\item {\em Tools} An important consideration when choosing certain 
-language to write non-trivial programs is the availability of development tools. While developing
-the experiments, we used a comprehensive set of developer tools available for both
-browsers and desktop \citep{tilkov2010node}. These tools included
-package managers, in this case {\tt npm}, debuggers, logging libraries
-such as {\tt winston}, 
-network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
-Although the fact that developer tools vary from browser to browser
-supposed a drawback, they all have the same functionality. The
-JavaScript debugger can be accessed directly from the user menu in
-most browsers. 
-\item {\em Data Types} JavaScript uses floating point numbers with a 
-limited precision of 64 bits that partially implement the functionality of
-the {\tt StrictMath} library used by the Java implementation. If more precision
-is needed, developers could use {\tt math.js}, an extensive math library with
-support for matrices and big numbers.
-\end{itemize}
-
-We found that the JavaScript ecosystem is mature enough for developing non-trivial 
-code, but developers have to consider the differences between VMs implementations
-and therefore use certain libraries in order to increase the precision and repeatability of
-algorithms. As for the performance of JavaScript against Java and Matlab
-we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
-Intel Xeon E5 processor running OS X 10.10.5, 
-using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
-MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
-Google Chrome Version 46.0.2490.86 (64-bit) 
-and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
- The performance of the time required for 10,000 function evaluations is presented in
-Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
-performance with an average of 935ms, followed by Java with
-991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
-single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
-The results also show that there is not much difference between the Node.js
-implementation an that of a
-browser. There is also not much difference between running the code in the main thread or in Web
-Workers. When running the experiment in Firefox a dialog alerted that the
-script was taking too much time, showing that this
-kind of processing is not intended for the main thread. Additionally, there is not much overhead
-when running two
-Web Workers in parallel as they took 1279ms each. This experiment
-shows that JavaScript is viable as a language
-language for developing complex mathematical functions an these can run in a browser
-with acceptable performance. 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{f15_times.png}
-\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
-\label{fig:f15_times}
-\end{figure}
-
-\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}
-This experiment was deployed in a virtual private server with 512MB of memory
-running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
-experiments where the optimum was expected to be found many times, 
-this problem was expected to run constantly for many days finding only
-sub-optimal approximations. In order to achieve this, the only change needed 
-in the system was to limit the number of chromosomes kept in the pool. 
-Previously, every PUT request which included a new chromosome increased 
-the size of the pool, in a long-running experiment this has to be avoided
-in order to not exhaust the available memory. In this experiment the size
-of a chromosome was 1000 double precision numbers, considering this
-and the memory available in the server, the the pool was limited 
-to 10,000 chromosomes.  In order to keep this size
-constant a random chromosome was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
-and {\tt classic-ea-float} modules were used. The parameters for the EA 
-were tournament size = 3 and  population size = 500, returning exchanging a
-chromosome with the server every 100 generations. Again two workers in each window.
-As before, a few participation calls were published in social networks.
-
-\subsubsection{Results}
-
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=3in]{rastrigin-fitness.png}
-\caption{Fitness of chromosomes sent to the server pool along time in
-  the $x$ axis.} 
-% Plotted with ../data/plot-rastrigin-fitness.R
-\label{fig:puts:rastrigin}
-\end{figure}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
-\caption{Number of unique IPs participating in the Rastrigin
-  experiment per minute (left) and per hour (right)} 
-% Plotted with ../data/plot-rastrigin-IPs.R
-\label{fig:ips:rastrigin}
-\end{figure}
-%
-\begin{table*}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of {\tt PUT}s per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 35.759  &  343.336   & 9.877 \\
-Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
-\caption{Fit to GEV (left) and Weibull (right) of the number of {\tt PUT}s per worker for the F15 (Rastrigin) function.} 
-% Plotted with ../data/ips-puts-rastrigin.R
-\label{fig:fit:rastrigin}
-\end{figure}
-%
-
-The results for this new experiment have been extremely interesting
-and quite different from the previous one. The most important
-characteristic that set this experiment apart from the previous one is
-the fact that it did not find the solution for the time it was
-running, instead lowering, little by little, the fitness until it was
-disconnected. The fitness of chromosomes sent by clients is shown in
-Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
-upper part, which represents the fitness of clients that have just
-started evolution from a random population, and then the fitness of
-those clients that have been already evolving for a while or obtained
-a high-fitness individual from the pool and evolved from there, and
-which managed to surpass the 9000 value by the end of the
-simulation. It incidentally shows also the number of {\em new} nodes
-coming: there is a dense {\em cloud} at the beginning, but the new
-contributions keep thinning as the evolution progresses.
-
-This is in part due to the fact that the presence of nodes is not
-constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
-number of unique IPs participating in the experiment per minute (left)
-and hour (right). The two graphs have the same shape which matches
-also the speed in which fitness is decreased in Figure
-\ref{fig:puts:rastrigin}. At the beginning there are very few or only
-one, but later one there are peaks of almost 30 unique machine a
-minute; these peaks appear again when new announcements are made in
-Twitter or any other social networks, to go down to a few remaining
-nodes later on. It is interesting to note that, at the most, more than
-100 unique nodes were participating in the experiment during one hour;
-also that there are sustained peaks during a day of 6-7 computers
-every minute. This is due to the fact that the computation is
-asynchronous, but also to the fact that different computers stay in
-the simulation for a different period of time. This period has been
-plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
-{\tt PUT}s per node, ranked by number, along with a fit to the two
-distributions we have used before, GEV and Weibull; the parameters of
-this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
-experiments, in which GEV fit was relatively good, in this case it is
-quite obviously not. However, the fit to the Weibull distribution,
-shown at the left of the image, is relatively good. As should be
-expected, the shape and scale are different to the one in the previous
-experiments with the Trap function, however, the shape is quite
-similar (0.40 vs 0.69), being in both cases less than one, which
-indicates a decreasing distribution of the same kind. In fact, the
-difference in the scale parameter indicates that the difference
-between the time devoted by a volunteer and the next is bigger in this
-case. This is probably due to the fact that the duration of the
-experiment is bigger, allowing for bigger differences in
-behavior. While the first in the Trap function might have devoted a
-few minutes and the second a few seconds less, the difference between
-the first and second in this experiment might be in the scale of
-hours.
-
-In general, this experiment confirms what was found in the others in
-several aspects: first, there is a regularity in the distribution of the
-time assigned by volunteers to it, second, there is a good and certain
-amount of volunteers that will become part of the distributed
-computer, in some cases up to one hundred; also, in some cases there
-will be users that will spend several hours, two orders of magnitude
-more generations in this case than in the previous one. There is a
-clear influence of the time to solution: a short time means that every
-experiment will only gather a few volunteers, but also that they will
-stay only for a short time. A long time will make users stay for
-longer, but if it takes too long, they will eventually lose interest.
-
-We will discuss this and the rest of the findings of the paper in the
-conclusions.
-
 %---------------------------------------------------------------
@@
 also resulted in a low number of generations allotted by users.
-
-In the second implementation, {\sf NodIO-W$^2$}, Web Workers
-were used so that several clients per browser could be executed and
-they could run in the background, so that the the number of
-generations per user could be improved. The number of generations per
-user increased to the point that baseline performance was improved in
-most cases. The architectural changes led to a high correlation
-between successive experiments, so that, even as the median number of
-volunteers per experiment decreased, they were probably contributing
-to the experiment at the same time, so that our objective was
-achieved. In general, one of the first conclusions for this paper is
-that the browser technologies should be used to its full extent so
-that the user time is leveraged to its full potential, and this was
-achieved with {\sf NodIO-W$^2$}.
 
@@
 this paper (and our work) with their suggestions. We are also grateful
-to Anna S\'aez de Tejada for her help with the data processing scripts.
+to Anna S\'aez de Tejada for her help with the data processing
+scripts. We are also grateful to {\tt @otisdriftwood} for his help
+gathering users for the new experiments.
 

### 2016-02-28 Alife16/paper.tex
Rewritten abstract and title
Please @mariosky @unintendedbear @fergunet have a good look.
Refs #99 #100

@@
 
-\title{ Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
-for Low Cost Distributed Evolutionary Computation}
+\title{The human in the loop: volunteer computing as a socio-technical
+system}
 \author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
@@
 \begin{abstract}
-Volunteer computing is extensively used 
-in science and industry, but it
-still has a long way to go in order to become
- a mainstream
-technology in the same way the cloud technologies in general have. The
-lack of 
-general purpose and open source tools with built-in analytics 
-could be one of the reasons of why this is so, as well as the lack of
-proofs of concept. That is why in this paper we present {\sf NodIO}, a
-framework used to create client-server architectures for distributed
-evolutionary algorithms, entirely written  
-in JavaScript, which needs
-minimal server support to work and is able to support persistent,
-asynchronous, distributed evolutionary algorithms running in the
-browser, as well as any other client consuming its REST application
-programming interface. The architecture, along with some initial
-results on the expected performance and scaling behavior when the
-number of users is increased is presented. We also use
-the algorithm data to find out what kind of patterns the user behavior
-follows by examining the resources they devote to our experiment, and
-try to fit them to some statistical distribution to create the bases
-of a model of user behavior in volunteer computing experiments of this
-kind that can be used by researchers to predict its performance. 
-In general, our experiments prove, first, that {\sf NodIO} is a flexible and
-low-overhead platform to perform distributed computation experiments
-that include volunteers, and then that user behavior follows some patterns,
-and, thus, are amenable to a model.
+Volunteer computing is a form of distributed computing where users
+decide on their participation and the amount of time they will
+``lend''. This makes them an essential part of the algorithm and of
+the performance of the whole system. As a sociotechnical system, this
+participation follows some patterns and in this paper we examine the
+result of several volunteer distributed evolutionary computation
+experiments and try to find out which are those patterns and what
+makes an experiment successful or not, including the feedback loop
+that is created between the users and the algorithm itself.
 \end{abstract}
@@
 \end{document}
+
+%%% Local Variables:
+%%% ispell-local-dictionary: "english"
+%%% End:

### 2016-02-28 Alife16/paper.tex
Reduced state of the art to the relevant parts.
Refs #101 but I still have to update it.

@@
 
-The web used as a resource for distributed computing has a
-long history, and almost since the beginning it was used for
-evolutionary computation even before JavaScript was
-introduced as a browser-based scripting language in 1997. Before that
-time other technologies,
- including Flash animations and VBScript, ActiveX or Java applets were
-used; for instance, Java was pointed out in \citep{soares1998get} as a
-``language for the
-internet'', providing some advantages such as multi-architecture compatibility or
-security mechanisms, and was proposed in \citep{chandy1996world} as the
-basis for the creation of a world-wide peer to peer system using compiled Java
-programs. The ATLAS system
-\citep{Baldeschwieler:1996:TIG:504450.504482} also used Java together
-with another language called Cilk; all these papers indicated that
-the introduction of an architecture-independent system such as the
-Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers. This paper also establishes the desirable features of
-any {\em global computing} infrastructure, that can also be applied to
-distributed evolutionary computation experiments that rely on
-volunteer computing: scalability, heterogeneity, fault tolerance,
-adaptive parallelism, safety, anonymity, hierarchy, ease of use, and
-reasonable performance. But hierarchy, this is, the capability of
-preferentially using resources in the same organization, is a feature
-that could be either disregarded or included within adaptivity, which
-leaves us with eight desirable properties.
-
-The fulfillment of these properties could be achieved by using any of the
-languages and associated technologies, mentioned above, provided that the user had
-installed the required virtual machine plugin first.
-
-However, embedding a
-distributed computation client in the browser has the advantage of
-providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed by using desktop applications. 
-Baratloo et al. tried in 1996 to create a framework with
-the aforementioned qualities called {\em Charlotte}
-\citep{baratloo1996charlotte}. This framework, which implemented a
-distributed shared memory, was inspired by a previous work
-on parallel computing systems PVM and MPI and solved most of the issues by
-using the Java virtual machine embedded in the browser,
-which provided
-security to the user, uniformity of computing environment
-independently of the operating system, and access to a common data store in the
-server. Since the Java virtual machine is a sandbox, it is safe to the user
-and as it runs compiled bytecode it also guarantees a certain level of
-performance.
-Other issues such as scheduling, which are related to adaptive 
-parallelism, were solved on the server. This
-paper was later extended and published in a journal in 1999
-\citep{baratloo1999charlotte}, but it is interesting to note that,
-even in the early stages of the web, the possibility of
-browser-based distributed computing was already researched.
-
-Other systems, like JET, proposed by Soares et
-al. \citep{soares1998get}, also used Java. JET supports
-the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, enabling its use for science, 
-since it allows 
-the collection of algorithm analytics by the
-experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to JET, such as {\em SuperWeb}
-\citep{alexandrov1997superweb} are roughly a translation of classical
-distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. SuperWeb is focused,
-however, in solving the
-issues involved in browser-based computing and also in making a system
-that is able to work across different types of computers, mentioning
-SPARCstations and PCs with Windows-95. Interestingly enough, this
-paper incorporates a discussion on the economic model, paving the way
-for the for-profit distributed computing systems that arose later on. 
-On the other hand, the potential for volunteer computing using
-browsers was realized
-later on \citep{sarmenta-bayanihan} as well as the potential of
-user misconduct when having access to the source code.
-\citep{sarmenta-sabotagetolerance}. However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
-less widespread.
-
-The main problem with Java and all the other browser-embedded
-technologies is that they are not {\em universal} in the
-sense that an extra component, namely, the Java virtual machine or another
-virtual environment has to be installed in the browser.
-Also these virtual environments have become targets of malicious attacks; 
-for instance numerous vulnerabilities have been discovered in 
-the proprietary Adobe Flash runtime environment and
-video player \citep{ford2009analyzing,watanabe2010new}. The need for
-these additional environments has been mitigated with the
-adoption of the HTML5 standard \citep{anthes2012html5}.
-JavaScript \citep{flanagan2006javascript}, however, has become a set of standards
-\citep{ECMA-262} that include the language and its components. Since
-early on, it was adopted by the industry and also by scientists,
-who used them for creating an evolutionary algorithm (EA) on
-the browser as early as 1998 \citep{jj-ppsn98}.
+Volunteer computing involves an user running a program voluntarily
+and, as such, has been deployed in many different ways from the
+beginning of the Internet, starting with the SETI@home framework for
+processing extraterrestrial signals \cite{david-seti:home}. However
+the dual introduction of JavaScript as an universal language for the
+browser and the browser as an ubiquitous web and Internet client has
+made this combination the most popular for volunteer computing
+frameworks such as the one we are using here and whose first version
+was described in \cite{DBLP:conf/gecco/GuervosG15}.
 
@@
 
-Recent works propose the use of cloud computing services for the components of
-client-server distributed EAs. Cloud resources
-can reduce operational costs by outsourcing both hardware and software maintenance
-to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
-is available in most laboratories, allowing researches to
-scale algorithms at reasonable costs. Researchers have proposed the use of
-public clouds such as Amazon EC2 \citep{CloudScale}, Google App Engine\citep{di2013towards},
-and DropBox~\citep{mericloud}. The use of JavaScript to implement
-server side components
-gives software developers certain advantages in the cloud
-paradigm, since it is a first-class citizen
-in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \citep{wood13:nodejs:paas}. Many web services
-have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
-developers can use the same software tools for both client and server
-side development.
-
-JavaScript is also compatible with the functional programming paradigm \citep{Cousineau1998,MacLennan1990,Thompson1996}.
-According to \citep{swanresearch2015}, the benefit of using this kind of paradigm in
-metaheuristics optimization is that it allows an improvement in communicability,
-reproductibility, interoperability, automated assembly, knowledge engineering
-and efficiency.
-
-In the work presented in this paper, JavaScript has been used throughout the full
-stack, which has the advantage of using the self-same lines of code for
-writing an evolutionary algorithm for the
-desktop or command line and browser clients, both based in the EC
-library {\sf NodEO} \citep{DBLP:conf/gecco/GuervosVGES14}. This saves time 
-when writing code and leverages 
-the expertise acquired in the library itself.
-
 On the other hand, we are also interested in measuring the performance
@@
 performance of the EA itself
-\citep{DBLP:journals/gpem/LaredoBGVAGF14}.
-
-On the other hand, measuring the performance of the resulting metacomputer
-involves understanding the dynamics of this kind of systems. Initial
+\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
+measured without understanding the dynamics of this kind of systems. Initial
 work was done for peer to peer systems by Stutzbach et
@@
 
-But some of the essential metrics in volunteer computing like the
+Some of the essential metrics in volunteer computing like the
 number of users or the time spent by every one in the
@@
 run. Studies using volunteer computing platforms such as SETI@home
-\citep{javadi2009mining} found out that the Weibull, log-normal and
+\citep{javadi2009mining,agajaj} found out that the Weibull, log-normal and
 Gamma distribution 
@@
 there are many users that give a small amount of cycles, while there
-are just a few that give many cycles. This is in concordance with the
-results obtained in \citep{agajaj}. Although SETI@home works
-at a larger scale in both quantity of volunteers and the scope of projects, 
-our initial tests indicate that
-this kind of distribution might be the most appropriate to describe
-user participation in several different metrics. However, in this
-paper we will perform different experiments in different setups and
-will try to come up with a more precise model. Next we describe the
-framework itself.
+are just a few that give many cycles.
 
-As far as we know, this paper presents one of the few experiments using 
-computational resources that are as dissimilar as smartphones and
-powerful laptops 
-or workstation computers in a research center. 
-%The methodology used
-% (Paloma) I would say "the used methodology"
+As far as we know, this paper presents one of the few experiments
+that measure the performance of a sociotechnical metacomputer. 
 The algorithms used, as well as the methodology 
@@
 function in Python, Erlang, Scala, or even Java, and them compile it to
-JavaScript. For completeness, an experiment comparing the performance of
-a JavaScript implementation versus Java and Matlab is included in section
-\ref{sec:w2}.
+JavaScript.
 

### 2016-02-28 Alife16/paper.tex
Updates state of the art
Reducing it at the same time, but adding some heretofore unmentioned works. Closes #101

@@
 
-%\usepackage[pdftex]{graphicx}
 \graphicspath{{../img/}}
@@
 \author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
-  García-Sánchez $^1$, Mario Garc\'ia-Valdez$^2$\\
+  Garc\'ia-S\'anchez $^1$, Mario Garc\'ia-Valdez$^2$\\
 \mbox{}\\
@@
 determine its performance and how it can be modeled and, eventually,
-predicted.
-
-The rest of the paper is organized as follows: Next we present the
-state of the art (Section \ref{sec:soa}) in web-based distributed
-computational systems along with attempts to predict and model its behavior. 
-Then, a detailed description of the proposed framework is presented
-in Section \ref{sec:description}, followed by the baseline experiments
-in Section \ref{sec:experiments}, and experiments using a new client
-architecture that uses web workers in Section \ref{sec:w2}.
-Finally, conclusions and future lines of work are presented in Section
-\ref{sec:conclusion}.
+predicted. Next we present the state of the art in web-based
+volunteer-based systems along with attempts to predict and model its
+behavior.
 
@@
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \cite{david-seti:home}. However
+processing extraterrestrial signals \citep{david-seti:home}. However
 the dual introduction of JavaScript as an universal language for the
@@
 frameworks such as the one we are using here and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15}.
+was described in \citep{DBLP:conf/gecco/GuervosG15}.
 
 JavaScript can be used either for unwitting
-\citep{unwitting-ec} or volunteer 
+\citep{unwitting-ec,boldrin2007distributed} or volunteer 
 \citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
@@
 authors, including more recent efforts
-\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
-that even
-used the client's GPU \citep{duda2013gpu} or create an object-oriented
-framework for evolution in the browser \citep{EvoStar2014:jsEO}. Many other researchers have
+\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
 used Java \citep{chong:1999:jDGPi} and others have gone away from the
@@
 al. \citep{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \citep{churn08,laredo2008rcp}.
+computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
+study was performed by Martínez et al. on the Capataz system
+\citep{martinez2015capataz}; however in this case the number of
+computers used was known in advance and the main focus was on
+measuring speed up and how job bundling helped to reduce overhead and
+enhance performance.
 
@@
 As far as we know, this paper presents one of the few experiments
-that measure the performance of a sociotechnical metacomputer. 
+that measure the performance of a sociotechnical
+metacomputer. Apolónia et al. \citep{apolonia2012enhancing} used the
+Facebook protocol to distribute tasks among the {\em walls} of
+friends, explicitly using the social network for computing. However,
+it stopped short of relating performance to the macro measures of the
+users' social networks. 
 The algorithms used, as well as the methodology 
@@
 it is a difficult problem for evolutionary algorithms, it takes a few
-seconds with the right settings to solve. This makes it difficult also
-for this kind of asynchronous framework, but it would be interesting
-to check the performance for other computationally demanding
-tasks, especially ones that cover the middle ground between a problem
-finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
+seconds with the right settings to solve. 
 Besides, the two versions of the framework include as an
@@
 including using different, or adaptive, policies for inserting and
-sending individuals to the pool as was done in \citep{araujo2008mam},
+sending individuals to the pool,
 using different policies for population initialization, and also the

### 2016-02-28 Alife16/paper.tex
Revises intro
Please @fergunet @unintendedbear @mariosky revise and, if needed close. Refs #102

@@
 \label{introduction}
-Nodays there is a big and increasing number of
-distributed evolutionary computation software frameworks available on
-the market, usually as open source systems \citep{Parejo12Survey}. 
-Most of these frameworks have 
-two features: they are {\em desktop} applications, that is, they must
-be compiled or run directly on the operating system layer and, second,
-they assume that the nodes that are going to run the system are under
-the control of developers or they at least have access to them.
-
-There is, then, space for frameworks that do not make any of these
-assumptions, such as the one presented in this paper, which will be
- named {\sf NodIO}. {\sf NodIO} is basically
-% (Paloma) Is the "first" word really needed? 
-% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
-a two tier client/server 
-application, in which clients can run on a browser. 
-Therefore, they can be embedded in any web-enabled device, including many over which 
-developers have no control.
 
 Some time ago, our group faced the problem of diminishing funds for
-buying new hardware. This was aggravated by the increasing maintenance costs and
-extended downtime resulting from the continuous failures of existing
-clusters.
-Considering this, we leveraged our experience in the design of web
-applications, including JavaScript, since their inception and other
-volunteer and
-unconventional distributed evolutionary computing systems to 
-design and release a new free framework that would allow anyone to
-create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
-servers and browsers as clients.
+buying new hardware. This was aggravated by the increasing maintenance
+costs and extended downtime resulting from the continuous failures of
+existing clusters.  Considering this, we leveraged our experience in
+the design of web applications with JavaScript and other volunteer and
+unconventional distributed evolutionary computing systems to design
+and release a new free framework that would allow anyone to create a
+volunteer distributed evolutionary computation (EC) experiment using
+cloud resources as servers and browsers as clients. This framework was
+called NodIO \citep{2016arXiv160101607M}. NodIO provides server
+infraestructure for volunteer-based distributed evolutionary computing
+experiments by providing a chromosome {\em pool}. This pool is used by
+clients in browsers and any other using the application programming
+interface (API) to put chromosomes and retrieve them, working then as
+a loose, asynchronous and {\em ad hoc} connection among all clients
+using it.
 
-There exist other volunteer computing frameworks written in compiled languages,
-such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript \citep{2015arXiv151101088M}. However, one of the advantages of our system is that
-clients do not need to install anything in their machines, they just
-need to open a web page  
-in a compatible browser, such as Chrome, Safari or Firefox, which are
-available in almost every desktop or mobile device. Other advantage is
-that the source code can be directly examined by volunteers, because it is not
-compiled as other options such as Java Applets, ActionScript or
-ActiveX.
-Also, as we will discuss in this paper, it can provide some extra
-features such as enhanced security or operating system independence.
+This loose connection provides a low-overhead way to connect desktop
+experiments with volunteer-based ones, with all of them contributing
+to the pool, but every one of them working as separate islands that
+carry out their own evolutionary algorithms. This is why NodIO is
+proposed mainly as a {\em complement} to existent resources such as
+desktop systems or laptops. As long as it provides a non-null
+computational capability that can help existing resources find the
+solution faster it will have found its purpose. The main use case is
+someone setting up NodIO in the cloud, writing a fitness evaluation
+function and running a client from his or her own computer, but
+requesting help in social networks for additional resources. That is
+why, in this system,  we have considered the whole {\em social} aspect
+in the design, with issues related to security, trust and privacy
+among others. The computing system becomes a {\em techno-social
+  system} \citep{vespignani2009predicting}. In this paper we are going
+to focus on measuring the response of users to experiments, but at the
+same time we will also focus on the technical aspects of the server and how
+those might change the behavior of users, improving the capability of
+the system.
 
-When developing volunteer based frameworks, designers must
-also consider the whole {\em social} aspect in the design, with
-issues related to security, trust and privacy among others. 
-The computing system becomes a {\em techno-social system} \citep{vespignani2009predicting}.
-
-In this paper, our intention is to
-present the {\sf NodIO} framework and analyze it from two different
-angles: the purely technical angle, including the decisions that went
-into its features, and the social angle, to account on how the persons
-that voluntarily choose to enter the web page running the system
-determine its performance and how it can be modeled and, eventually,
-predicted. Next we present the state of the art in web-based
+Our research group is committed to open science, and we think this is
+a very important part of the techno-social system. By being
+transparent, incentives to cheat are reduced and, in fact, we have
+detected no issue for the time being. Next we present the state of the art in web-based
 volunteer-based systems along with attempts to predict and model its

### 2016-02-28 Alife16/paper.tex
Intro

@@
 Volunteer computing is a form of distributed computing where users
-decide on their participation and the amount of time they will
+decide on their participation and the amount of time and other resources they will
 ``lend''. This makes them an essential part of the algorithm and of
@@
 experiments with volunteer-based ones, with all of them contributing
-to the pool, but every one of them working as separate islands that
-carry out their own evolutionary algorithms. This is why NodIO is
+to the pool, but every one of them working as separate island 
+carrying out their own evolutionary algorithms. This is why NodIO is
 proposed mainly as a {\em complement} to existent resources such as
@@
 same time we will also focus on the technical aspects of the server and how
-those might change the behavior of users, improving the capability of
+those might change the behavior of users, improving the capability of % Mario: How these might change
 the system.

### 2016-02-28 Alife16/paper.tex
SoA
Few edits,

@@
 
-JavaScript can be used either for unwitting
+JavaScript can be used for either unwitting
 \citep{unwitting-ec,boldrin2007distributed} or volunteer 
@@
 \citep{milani2004online}, which is actually not such a big problem in
-algorithms such as the EA that can easily be
-parallelized via population splitting or farming out evaluations to all
+algorithms such as the EA that can be easily 
+parallelized via population splitting or by farming out the evaluation to all
 the nodes available. Lately, several approaches have focused on the
-fault-tolerance of volunteer algorithms
+fault-tolerance of volunteer algorithms %volunteer systems?
 \citep{gonzalez2010characterizing} which it can, of course, be studied in
@@
 \citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
-measured without understanding the dynamics of this kind of systems. Initial
+measured without first understanding the dynamics of this kind of systems. Initial
 work was done for peer to peer systems by Stutzbach et
@@
 computers used was known in advance and the main focus was on
-measuring speed up and how job bundling helped to reduce overhead and
+measuring the speed up and how job bundling helped to reduce overhead and
 enhance performance.

### 2016-02-28 Alife16/paper.tex
Description
@JJ : I think a paragraph is not needed it references Web Workers and
repeats things we are saying elsewhere.

@@
   additional links, and information on the experiment. This code runs
-  an evolutionary algorithm {\em island} starting with a random
-  population, then it sends, every 100 generations, the best individual
-  back to the server (via a {\tt PUT} request), and requests a random
-  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
+  an evolutionary algorithm {\em island} that starts with a random
+  population, then after every 100 generations, it sends the best individual
+  back to the server (via a {\tt PUT} request), and then requests a random
+  individual back from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
 \end{enumerate}
+
+%%%%% DELETE THIS PARAGRAPH? References Web Workers  %%%%%%
 
@@
   the algorithm can be run in many different ways: 
-  stopping when a solution is found or, as we have done in this paper,
+  stopping when a solution is found or , as we have done in this paper,
   using Web Workers. In fact, since it is a pool-based system such as
@@
   the system. 
+%%%%%%%
 

### 2016-02-28 Alife16/paper.tex
Description
Again @JJ please delete or change a paragraph that repeats info, one or
the other.

@@
   in the browser, which, every 100 generations, sends the best
-  individual and receives a random one from the server.}
+  individual and receives a random one back from the server.}
 \label{fig:system}
@@
 
+%%%% DELETE? Repeats lines 186 - 191
 The researcher only has to change a function, which can be written in
@@
 JavaScript \citep{web:compilersjs} to solve a different
-problem. In this case the classical Trap function \citep{Ackley1987} has been
-used as well as the Rastrigin's floating-point optimization
-problem. JavaScript is a functional language and declaring a different 
+problem. 
+%%%%  
+In this case the classical Trap function \citep{Ackley1987} has been
+used. JavaScript is a functional language and declaring a different 
 function and handing at the creation of the algorithm object, called
@@
   visiting the page, even from mobile devices, can load the algorithm.
-  In case the browser does not support HTML5 Web workers (e.g., 
-  Windows Phone or some version of Android) or their use
-  has been disabled, a basic version of {\sf NodIO} can also be used.
 \item {\em Fault tolerance} is always an issue, and in this case, the
@@
 All these points will be established in the next two sections,
-starting with the baseline, single worker system next.
+starting with the baseline, single worker system next. %%%  Single worker?
 

### 2016-02-28 Alife16/paper.tex

@@
 %---------------------------------------------------------------
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 
@@
 TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
+PROY-PP2015-06 (Plan Propio 2015 UGR),  % please, remember to include all the current projects   ;)   Thanks!
 SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17

### 2016-02-28 Alife16/paper.tex
Deleted some paragraphs

@@
 
-%%%%% DELETE THIS PARAGRAPH? References Web Workers  %%%%%%
-
-  This is the invariant part of the framework, but other than that,
-  the algorithm can be run in many different ways: 
-  stopping when a solution is found or , as we have done in this paper,
-  using Web Workers. In fact, since it is a pool-based system such as
-  the one described in \citep{LNCS86720702}, any kind of client that
-  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API.
-But never heard of a client which "works" an API.
-Is this correct?
-  written in any kind of language. However, in this paper we are going
-  to focus on the dynamics and measurements of the volunteer part of
-  the system. 
-%%%%%%%
-
 Figure \ref{fig:system} describes the general system architecture and
@@
 
-%%%% DELETE? Repeats lines 186 - 191
-The researcher only has to change a function, which can be written in
-JavaScript or in any other language with cross-compilation to
-JavaScript \citep{web:compilersjs} to solve a different
-problem. 
-%%%%  
 In this case the classical Trap function \citep{Ackley1987} has been

### 2016-02-28 Alife16/paper.tex
Fixed error with refs clashing with hyperref

--- 2016-02-28 Alife16/paper.tex
@@
 
-Figure \ref{fig:system} describes the general system architecture and
+Figure \nameref{fig:system} describes the general system architecture and
 algorithm behavior. Different web technologies, such as JQuery or {\tt
@@
 problem. Let us see how this system addresses the different challenges
-outlined in the State of the Art (Section \ref{sec:soa}):
+outlined in the State of the Art (Section \nameref{sec:soa}):
 \begin{itemize}
@@
 seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
+are shown in Figure \nameref{fig:baseline}. The baseline hardware system has these
 characteristics: 
@@
 in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:runs}. In general, every experiment took several
+Table \nameref{tab:runs}. In general, every experiment took several
 days. No particular care was taken about the time of the announcement
@@
 A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of IPs
+\nameref{tab:summary:os}, which shows the median number of IPs
 intervening in each experiment,  median time needed
@@
 by adding more computers makes the experiment take less. In fact, as
-shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
+shown in Figure \nameref{fig:duration}, the {\em addition} of more computers does
 not seem to contribute to decreasing the time needed to finish the
@@
 experiment.
-%  however, at this point we
-% have not found this convenient since there are several other issues
-% that have to be solved, as it will be shown next. %Creo que este párrafo
-%saldría sobrando ahora que hicmios un experimento más grande
 
@@
 It is also interesting to check the distribution of the experiment
-duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
+duration, shown in Figure \nameref{fig:zipf:os} and which roughly follows
 a Zipf's law, with similar distribution along all three runs. The 4/24
@@
 generations divided by 100, contributed by every user, which is shown
-in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
+in Figure \nameref{fig:puts:os}. These results show a Zipf-like behavior,
 so we have fitted it to the Generalized Extreme Value distribution,
-with the resulting parameters shown in Table \ref{tab:puts:os}.
+with the resulting parameters shown in Table \nameref{tab:puts:os}.
 %
@@
   leans} towards the origin, and negative ones towards the other extreme
-value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
+value. In this case, Table \nameref{tab:puts:os} shows $\xi$ values
 greater than one and between one and 1.4, which indicates that the
@@
 distributions with these parameters have been plotted in red in
-Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
+Figure \nameref{fig:puts:os}, indicating that the fit is good enough. The
 predictive value of these fits, however, is limited, over all taking

### 2016-02-28 Alife16/paper.tex
Changing stuff
Minor corrections and two questions related with the metrics to detect the patterns for issue #102

--- 2016-02-28 Alife16/paper.tex
@@
 that is created between the users and the algorithm itself.
+%Pablo: The patterns we are looking for are... 
 \end{abstract}
@@
 called NodIO \citep{2016arXiv160101607M}. NodIO provides server
-infraestructure for volunteer-based distributed evolutionary computing
+infrastructure for volunteer-based distributed evolutionary computing
 experiments by providing a chromosome {\em pool}. This pool is used by
@@
 carrying out their own evolutionary algorithms. This is why NodIO is
-proposed mainly as a {\em complement} to existent resources such as
+proposed mainly as a {\em complement} to existing resources such as
 desktop systems or laptops. As long as it provides a non-null
@@
 among others. The computing system becomes a {\em techno-social
-  system} \citep{vespignani2009predicting}. In this paper we are going
+%Pablo: In abstract is named 'sociotechnical'. Which one is the correct form for the rest of the paper?
+  system} \citep{vespignani2009predicting}. In this paper, we are going
 to focus on measuring the response of users to experiments, but at the
+%Pablo: which metric are we going to measure this response?
 same time we will also focus on the technical aspects of the server and how

### 2016-02-28 Alife16/paper.tex

@@
 \usepackage{natbib,alifeconf}
-\usepackage{hyperref}
+\usepackage[breaklinks=true]{hyperref}
 \usepackage{url}

### 2016-02-28 data/plot-zipf-openshift.R
Let's put stuff in order now
I'm re-writing the old R scripts to compute the Weibull distribution, which I wasn't for some reason. The GEV does not work on these new runs, so we probably cannot use it here. Let's stick to Weibull.

@@ -13,7 +13,18 @@ process.puts <- function(suffix ) {
                                     this.fit$estimate["scale"],
                                     this.fit$estimate["shape"]),decreasing=T))
     ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.plot,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=25), axis.text.y = element_text(angle=90, vjust=0.5))
-    ggsave(paste0("../img/puts-openshift-",suffix,".png"),width=5,height=5)
+    ggsave(paste0("../img/gev-puts-openshift-",suffix,".png"),width=5,height=5)
+
+    ## Weibull
+    this.weibull <- fitdistr(this.df$puts,"weibull")
+    print(this.weibull)
+
+    this.fit.weibull <- data.frame(rank=1:length(this.df$puts),
+                                y=sort(rweibull(length(this.df$puts),
+                                                this.weibull$estimate["shape"],
+                                                this.weibull$estimate["scale"]),decreasing=T))
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.weibull,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=25), axis.text.y = element_text(angle=90, vjust=0.5))
+    ggsave(paste0("../img/weibull-puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }
     

### 2016-02-28 Alife16/paper.tex
Taking out verbose explanations
Managed to reduce it to 8 pages
Closes #99 but we'll have to keep it that way.

@@
 JavaScript. Since JavaScript is  the only language that is present
-across all browsers, the choice was quite clear. Even so, there might
-be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
- is
-why we have made a comparison between JavaScript and other languages
-\citep{2015arXiv151101088M}. The results of this work show that JavaScript can be
-faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
-in evolutionary computation and numerical optimization such as Python. 
-% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
-Compiled languages such as
-Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
-considering the performance of a single-user, 
-but this is more than offset by the computational power of
-the spontaneous volunteers we can gather at whim, including people
-using their mobile phones or tablets. Together, the performance is several orders of magnitude
-higher, which is the objective in this kind of systems.
-% (Paloma) This sentence needs re-writing
-% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset.
-This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
- JavaScript is
-quite a popular language nowadays, providing compilers from a number
-of languages, many of them in their family (like CoffeeScript), but
-many other outside, so researchers can, in fact, write their fitness
-function in Python, Erlang, Scala, or even Java, and them compile it to
-JavaScript.
-
-As a volunteer based system, there are some issues worth 
-noting \citep{sarmenta2001volunteer} :\begin{enumerate}
-\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
-each request is known.
-\item As anonymous entities, volunteers are not accountable. 
-If a volunteer misbehaves in some way the provider cannot 
-prosecute or discipline the volunteer. In the case of JavaScript,
-this issue is aggravated because even if it is obfuscated, source code
-and data is sent to clients. 
-\item Participants must also trust the application provider, 
-regarding invasion of privacy, the integrity of their devices, 
-and intended application. 
-\end{enumerate}
-
-Two kind of threats are possible in this environment; first, since the
-API is open and it is an open source framework, it is relatively easy to
-sabotage the system, as indicated in \citep{domingues2007sabotage}, by
-crafting a fake request which, for instance, assigns a fake fitness to
-a particular chromosome.
-% (Paloma) Where is the second kind of threat?
-Another two different attacks are also
-possible: a denial of service attack as well as penetration
-testing. We have approached this threat in a pragmatic way: all
-sources, client and server, for the application, are openly released. 
- We also release as open access the performed experiments \citep{DBLP:journals/corr/GuervosG15} 
- and all data, once anonymized, as open data. This
-builds trust between the user and the scientist, which indeed has an
-impact on the performance of the system, since there is no need to
-include cheating checks or other functions that would degrade the
-performance of the system. This socio-technical approach is, indeed,
-coherent with our whole approach to the study of a distributed volunteer
- system which can only be studied by approaching it from the
-technical as well as the social, as in social media, angle.
+across all browsers, the choice was quite clear. We should emphasize
+that NodIO is more intended as an auxiliary computing engine, more
+than the main one, so performance of JavaScript as a language is not
+so important; even so, we have made a comparison between JavaScript and other languages
+\citep{2015arXiv151101088M} that shows that performance of JavaScript
+is comparable to other interpreted languages; compiled languages would
+be faster, but, of course, it is impossible to gather volunteers
+spontaneously and without any installation with them.
 
@@
 {\tt Classic}, is the only change needed to work with a different
-problem. Let us see how this system addresses the different challenges
-outlined in the State of the Art (Section \ref{sec:soa}):
-\begin{itemize}
-\item {\em Scalability} is provided via the use of a lightweight and
-  high-performance, single-threaded, server based in Node.js and
-  Express.js. Although this single server is a bottleneck since it
-  will eventually saturate, the fact that it runs as a non-blocking single thread
-  allows the service of many requests. In fact, a limit in the
-  number of simultaneous requests will be reached, but so far it has
-  not been found, unlike what we found in our previous systems, DCoR,
-  \citep{gecco07:workshop:dcor}, which had a low scalability. 
-\item The system is {\em heterogeneous} since it does not need any
-  performance, operating system, or even browser requirement as long
-  as JavaScript is enabled: anyone
-  visiting the page, even from mobile devices, can load the algorithm.
-\item {\em Fault tolerance} is always an issue, and in this case, the
-  single point of failure would be the server: the system, as such,
-  would break down if the server fails. However, the individual
-  islands in every browser would continue running, and having access
-  to just one of them would allow the local algorithm to proceed. In
-  fact, the island does not need the server to run: it runs locally if
-  needed, with the only exception that it is obviously unable to
-  communicate with the rest of the islands.
-\item {\em Adaptiveness} is achieved simply through the autonomous
-  operation of every individual island without any synchronization
-  mechanism. The islands in the system are, in fact, unaware of each
-  other, communicating only through the server. This architecture 
-  has been tested in other high demand systems with success, but 
-  in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
-  %(Mario) Scalability of the communication system in particular?
-\item Since the algorithm runs in the browser, {\em safety} is
-  achieved through its sandbox mechanisms. The user is thus assured
-  that there is no unsafe access either to their local files or even
-  to more resources than the browser should be allotted.
-\item Running the algorithm is just a matter of loading the page,
-  which makes the operation totally {\em anonymous}. For the same
-  reason, {\em ease of use} is optimal, being as easy as simply
-  clicking on an URL, available to anyone with access to a browser.
-\item {\em Reasonable performance} is not ensured. In fact, we should
-  make sure that there is a reasonable amount of clients over which
-  the performance achieved is better than what you would obtain in
-  your own desktop system. If this is not the case, it is a pointless
-  academic exercise.
-\end{itemize}
-
-There are many different ways to validate a framework that intends to
-address these challenges. Some of them are included by design: it is
-anonymous, since no accounts are needed to participate in the
-experiment, it is safe for the user, since we are using the browser
-black box, and it is also heterogeneous. We have designed the
-experiments so that they show adaptability to different types of
-experiments and clients, scalability with the number of clients and
-the problem size, and measured performance by comparing it with
-a baseline configuration, with which, as indicated above, is
-compatible and can be used in a complementary way. Fault tolerance
-will be experimented by dropping the server and checking the continuity
-of the experiments.
-
-All these points will be established in the next two sections,
-starting with the baseline, single worker system next. %%%  Single worker?
+problem. The next Section will describe experiments performed to
+establish baseline performance and gather initial performance
+results.
 

### 2016-02-28 Alife16/paper.tex
Leaving it for today
Up to the results section, pretty much finished. But please @mariosky @fergunet @unintendedbear have a look and revise fof badly written things, things missing, and things that shouldn't be there.

Main thing now is to insert new results, which include different cache sizes and the new problem, with 50 traps, in the loop. There is space for one page, but conclusions will be reduced, so I think there's enough. Will do it tomorrow, but please all pitch in, because tomorrow is the last day!

@@
   run a single experiment, storing the chromosomes in a data structure
-  that is reset when the solution is found.
+  that is reset when the solution is found. This data structure can
+  hold every chromosome in a particular experiment, or have a finite
+  size that erases the oldest chromosomes once it has filled to
+  capacity. In this paper we will test both implementations.  
 \item A client that includes the evolutionary algorithm as
@@
   back to the server (via a {\tt PUT} request), and then requests a random
-  individual back from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
+  individual back from the server (via a {\tt GET} request). We have
+  kept the number of generations between migrations fixed since it is
+  a way of finding out how many real work every client has done. 
 \end{enumerate}
@@
 %---------------------------------------------------------------
-\section{Baseline Experiment and obtained results} %Cambié el nombre
+\section{Modeling the performance of a volunteer-based distributed computer} %Cambié el nombre
 \label{sec:experiments}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3in]{baseline-times.png}
-\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
-    Node.js} client, for different population sizes. Only
-runs in which solution has been found are used for this graph.}
-% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
-\label{fig:baseline}
-\end{figure}
-First, we establish the baseline performance by running the
-evolutionary algorithm in a desktop client written using {\sf NodEO}
-\citep{nodeo2014}, the basic JavaScript library we have used as base to
-build {\sf NodIO}, the framework. This experiment tries to 
-find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$ and a population size equal to 512. The algorithm in a
-single island was run until the solution (a string with all
-ones) was found or five million evaluations had been performed. It
-took around a minute, on average, that is $68.9694$ seconds, to
-perform the fifty runs. In this experiment only 33, that is, 66\% were
-successful. The experiments were repeated for population $p=1024$,
-with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
-seconds. Results for these two experiments
-are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
-characteristics: 
-\begin{verbatim}
-Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
-2014 x86_64 x86_64 x86_64 GNU/Linux
-\end{verbatim}
-with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
-
-These two experiments establish a baseline result and show that
-the time to success depends on the population size, with a bigger population
-contributing to have more diversity and thus speeding up the solution \citep{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
-experiments that we will describe next do not and can not have the
-same conditions, but
-the baseline is that if they eventually take longer than a basic
-desktop, their interest will be purely academic. We will try and
-prove next that that target performance can be achieved by carrying
-out several experiments in different conditions. Second, they also act
-as a basic performance test for the server, and also prove that the
-basic server can be used for many kinds of different setups, as long
-as they follow the REST protocol to deposit and obtain individuals
-from the pool. It is obvious that, in this case, the server is not
-really needed, since there is a single client and no interaction
-between them, but there is a cost in sending and obtaining chromosomes
-from the server so it is included for the sake of fairer comparison.
-
-\subsection{Volunteer evolutionary computation: first experiments and results}
 

### 2016-02-29 Alife16/paper.tex
Processing new results

@@
 %---------------------------------------------------------------
-\section{Modeling the performance of a volunteer-based distributed computer} %Cambié el nombre
+\section{Modeling the performance of a volunteer-based distributed computer} 
 \label{sec:experiments}
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier within which this 
-experiment could be performed. That way, and as required, the hardware
-and cloud cost for this experiment were zero. Experiments were
-announced through a post in Twitter and other social networks, and
+PaaS, which provides a free tier, making the whole experiment cost
+equal to 0. Experiments were
+announced through a posts in Twitter and, the latest case, Telegram, and
 results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
 purpose of this paper, we repeated the announcement several times
-through the month of April and then by the beginning of August. All
+through the month of April and then by the end of February. All
 in all, we have the set of runs with the characteristics shown in
@@
 %
-\begin{table*}
+\begin{table}
 \caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}

### 2016-02-29 data/IPs-per-hour.pl

@@ -7,7 +7,9 @@ use v5.14;
 
 use File::Slurp::Tiny qw(read_lines);
 
-my @lines = read_lines("rastrigin-IPs.csv");
+my $file_name = shift || "rastrigin-IPs.csv";
+
+my @lines = read_lines($file_name);
 
 my %IPs_hour;
 for my $l (@lines[1..$#lines]) {

### 2016-02-29 data/IPs-per-minute.pl

@@ -7,15 +7,22 @@ use v5.14;
 
 use File::Slurp::Tiny qw(read_lines);
 
-my @lines = read_lines("rastrigin-IPs.csv");
+my $file_name = shift || "rastrigin-IPs.csv";
+my @lines = read_lines($file_name);
 
 my %IPs_minute;
+my %total_IPs;
 for my $l (@lines[1..$#lines]) {
-  my ($IP, $time ) = split(/,\s+/,$l);
-  my ($minute) = ($time =~ /(\d+-\d+-\d+ \d+:\d+)/ );
+  my ($IP, $time ) = split(/;\s+/,$l);
+  my ($minute) = ($time =~ /(\d+-\d+-\d+.\d+:\d+)/ );
+  if (!$minute ) {
+    die "Wrong format for $time ";
+  }
+  $total_IPs{$IP}++;
   $IPs_minute{$minute}{$IP}++;
 }
 
+say "time,IPs-Total:".scalar keys %total_IPs;
 for my $t ( sort { $a cmp $b } keys %IPs_minute ) {
   say "$t, ", scalar keys %{$IPs_minute{$t}};
 }

### 2016-02-29 data/jq-fitness-time.sh
Files processed and data

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+cat Rastrigin.log| /usr/local/bin/jq 'select(.chromosome != null) | [ .IP, .timestamp ] | map(tostring) | join( ", " ) ' > rastrigin-IPs.csv

### 2016-02-29 data/IPs-per-minute.pl
Changing files

@@ -14,15 +14,15 @@ my %IPs_minute;
 my %total_IPs;
 for my $l (@lines[1..$#lines]) {
   my ($IP, $time ) = split(/;\s+/,$l);
-  my ($minute) = ($time =~ /(\d+-\d+-\d+.\d+:\d+)/ );
+  my ($day,$minute) = ($time =~ /(\d+-\d+-\d+).(\d+:\d+)/ );
   if (!$minute ) {
     die "Wrong format for $time ";
   }
   $total_IPs{$IP}++;
-  $IPs_minute{$minute}{$IP}++;
+  $IPs_minute{"$day $minute"}{$IP}++;
 }
 
-say "time,IPs-Total:".scalar keys %total_IPs;
+say "time,IPs";
 for my $t ( sort { $a cmp $b } keys %IPs_minute ) {
   say "$t, ", scalar keys %{$IPs_minute{$t}};
 }

### 2016-02-29 data/plot-IPs-per-minute.R
Adds more images and files

@@ -1,0 +1,16 @@
+library(ggplot2)
+
+
+ips.per.minute <- function( suffix ) {
+    file.name <- paste0("ips-per-minute-cache=",suffix)
+    IPs <- read.csv(paste0("2016-alife/",file.name,".csv"))
+    IPs$minute <-  as.POSIXct(IPs$time, "%Y-%m-%dT%H:%M")
+    ggplot(data=IPs, aes(x=minute,y=IPs))+ scale_x_datetime()+geom_line()+stat_smooth()
+    ggsave(paste0("../img/",file.name,".png"),width=4,height=3)
+    
+}
+
+suffix <- c("128","64","32")
+for ( i in suffix ) {
+    ips.per.minute( i )
+}

### 2016-02-29 Alife16/paper.tex
Files processed and data
Updates table

@@
 establish baseline performance and gather initial performance
-results.
+results. In the experiments performed last year we used the 40-trap
+problem, while the current experiments changed to the more difficult 50-trap.
 
@@
 \begin{table}
-\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
+\caption{Characteristics of the experiments. \label{tab:runs}}
 \begin{center}
-\begin{tabular}{l|rr}
+\begin{tabular}{l|rrl}
 \hline
-Date & \# experiments & Different IPs \\
+Experiment & \# experiments & Different IPs & Traps \\
 \hline
-April 4th 4/4 & 57 & 191 \\
-April 24th 4/24 &  231 & 559 \\
-July 31th 7/31 & 97 & 179 \\
+April 4th 4/4 & 57 & 191 & 40 \\
+April 24th 4/24 &  231 & 559& 40  \\
+July 31th 7/31 & 97 & 179 & 40 \\
+February, cache=128 & 61 & 75 & 50  \\
+February, cache=64 & 61 & 220 & 50  \\
+February, cache=32 & 39 & 86 & 50  \\
+% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
 \hline
@@
 %
-\begin{table*}
+\begin{table}
 \caption{Summary of time per run, number of IPs and number of {\tt
@@
 \begin{center}
-\begin{tabular}{l|ccccccc}
+\begin{tabular}{l|cccc}
 \hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
-  PUT}s & $<$ 69s & $<$ 3.46s & Inter-exper. correlation\\
+     & \multicolumn{2}{c}{IPs} & \multicolumn{2}{c}{Median} \\
+Experiment & Median & Max & time (s) &  \#{\tt  PUT}s \\
 \hline
-4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
-4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
-7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
+4/4 & 5 & 16 & 2040 & 18   \\
+4/24 &  5 & 29 & 732 & 11  \\
+7/31 & 5 & 14 & 260 & 23   \\
 \hline

### 2016-02-29 Alife16/paper.tex
Updates table and data
Explaining new results

@@
 
+\begin{table}[htb]
+\caption{Experiment table, with summary of results. \label{tab:runs}}
+\begin{center}
+\begin{tabular}{l|rrl}
+\hline
+Experiment & \#Runs & Different IPs & Traps \\
+\hline
+April 4th 4/4 & 57 & 191 & 40 \\
+April 24th 4/24 &  231 & 559& 40  \\
+July 31th 7/31 & 97 & 179 & 40 \\
+February, cache=128 & 61 & 75 & 50  \\
+February, cache=64 & 61 & 220 & 50  \\
+February, cache=32 & 39 & 86 & 50  \\
+% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+%
 Initial experiments were set up using the OpenShift
@@
 of these islands in the next experiment is indeed negligible.
+
+The table shows that every run included more than 30 experiments. The
+number of different IPs intervening in them varied from more than one
+hundred to more than five hundred in the second experiment, with a
+number around 50 in the second, and most recent, batch of experiments. 
 %
-\begin{table}
-\caption{Characteristics of the experiments. \label{tab:runs}}
-\begin{center}
-\begin{tabular}{l|rrl}
-\hline
-Experiment & \# experiments & Different IPs & Traps \\
-\hline
-April 4th 4/4 & 57 & 191 & 40 \\
-April 24th 4/24 &  231 & 559& 40  \\
-July 31th 7/31 & 97 & 179 & 40 \\
-February, cache=128 & 61 & 75 & 50  \\
-February, cache=64 & 61 & 220 & 50  \\
-February, cache=32 & 39 & 86 & 50  \\
-% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{table}
+\begin{table}[htb]
 \caption{Summary of time per run, number of IPs and number of {\tt
@@
 7/31 & 5 & 14 & 260 & 23   \\
+Cache=128 & 5 & 17 & 222.2 & 124 \\ 
+Cache=64 & 8 & 38 & 51.3 & 100 \\
+Cache=32 & 6 & 19 & 58.9 & 45 \\
 \hline
+% Data in .RData file in Splash-volunteer
 \end{tabular}
@@
 %
-The table shows that every run included more than 50 experiments. The
-number of different IPs intervening in them varied from more than one
-hundred to more than five hundred in the second experiment.
 
@@
 intervening in each experiment,  median time needed
-to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
-then two columns showing the percentage of experiments that took less
-than the two baseline experiments shown in the introduction to this
-section. The first striking result is that in all cases, 50\% of the
-experiments involved 5 or less IPs. This is consistent with results
-previously obtained \citep{DBLP:conf/gecco/GuervosG15} which found 6 to be
-the usual number of IPs that participated in an experiment. The
+to finish the experiment, median number of HTTP {\tt PUT}s per IP. The
+first striking result is that in all cases, 50\% of the 
+experiments involved 5 or less IPs. This is consistent with previous results
+\citep{DBLP:conf/gecco/GuervosG15} which found 6 to be
+the expected number of volunteer IPs. The
 maximum number of different IPs for each experiment is also in the
 same range and of the order of 10, which is also consistent with
-prior work. The median time has a big range of variation, but 50\% of
+prior work and does not vary across the two different batches.
+
+We will have to analyze differently the median time, since the two
+batches are solving different problems. In both cases it possesses a
+big range of variation, but 50\% of 
 the time takes less than several minutes, from around 4 minutes in the
-best case to roughly 2/3 of an hour in the worst case. That figure is
-not competitive, {\em a priori}, with the baseline experiment, that is
-why it is interesting to look at the columns which measure
-the percentage of experiments in which the solution was found in less
-time than the baseline. It is similar in all three cases, with around
-20\% on average for the first baseline experiment, roughly 3\% in the
-second case. The comparison is not totally fair: this experiment takes
-place in a browser, and in fact most of the time is devoted to
-presenting the charts; the server is also not the same since it is run in a
-PaaS instead of a high-performance desktop computer, and it is also a
-remote server, not a local server as in the baseline experiment. But
-even in this case, we were looking for creating a framework that
-allowed us to run evolutionary computation experiments {\em faster} than
-in our very own computer. And the results indicate that although it will
-happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
-experiment you choose, it will not happen always or even most of
-the time. One of the problems is shown in the last column, which
-indicates the statistical correlation between the number of IPs
-participating in one experiment and the next. In all cases,
-correlation is so low as to affirm that there is almost no
-relationship between them, that is, computing nodes are not {\em
-  staying} after the solution has been found. This is a feature of
-this implementation: the user has to actively reload the page to make
-the evolutionary algorithm start again. Thus, the user can decide if to 
-continue sharing resources or not. % (Paloma) Feature or disadvantage?
-However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
-unnecessary upper limit on the time, or number of experiments, that a user
-will voluntarily perform.
+best case to roughly 2/3 of an hour in the worst case. Remarkably
+enough, the time is more consistent in the second batch and always
+around one minute, in two cases even less, and that happens when the
+median number of IPs is higher. It should be noted that while the
+first batch of experiments took several days in each case, the second
+only lasted for a few hours, with a more continued effort of
+publicizing it in social networks; this is specially true in the
+case of cache equal to 64, which is noticed by the high number of
+volunteers participating in the experiment. The conclusion is that,
+in general, the key factor in the time needed to find the solution
+is, as expected, the number of volunteers it is able to gather on a
+short notice.
 

### 2016-02-29 Alife16/paper.tex
Adds new graphics

@@
 July 31th 7/31 & 97 & 179 & 40 \\
+\hline
 February, cache=128 & 61 & 75 & 50  \\
@@
 7/31 & 5 & 14 & 260 & 23   \\
+\hline
 Cache=128 & 5 & 17 & 222.2 & 124 \\ 
@@
 
+The number of {\tt PUTs}, every one corresponding to 100 generations,
+is the algorithmic result. It is relatively unchanged for the first
+batch and around 20, that is, 2000 generations or 2000*128 = 256000
+evaluations. In this case, an ``unlimited'' cache was used, with all
+individuals sent from clients stored until the end of the
+experiment. However, we were interested in measuring also the
+performance of the algorithm itself by changing the cache size, after
+making it limited. As it can be seen in the table, there is a clear
+change in the number of evaluations needed, with smaller cache sizes
+producing solutions in less evaluations, until it is for the cache
+size = 32 roughly twice as much as with the previous problem, with 40
+traps. This is a good result and is also algorithmically consistent
+with other results obtained using the same type of problems. Since in
+this paper we were interested in leveraging the user's CPU cycles by
+improving the algorithm, a good conclusion of this paper is that
+having a small pool size helps clients to obtain ``good'' individuals
+from the pool, as opposed to any individual that could be obtained
+before. Besides, cache policy deletes the oldest individuals, which
+makes those in the pool be {\em current}, helping then newcomers and
+any participant obtain the best individuals found in the last part of
+the experiment. Besides, a limited cache helps also in cases with a
+bigger search space or longer running times when the server simply
+crashed due to lack of RAM.
+
 \begin{figure}[!htb]
@@
 \label{fig:zipf:os}
+% First ones made with time-vs-IPs-openshift.R:  
 \end{figure}

### 2016-02-29 data/time-vs-IPs-alife.R

@@ -1,0 +1,41 @@
+library(ggplot2)
+library(grid)
+                        
+time.vs.IPs <- function( data, preffix ) {
+    data$howmany <- 1:nrow(data)
+    for ( i in 1:nrow(data) ) {
+        data[i,]$howmany <- length(data[data$IPs==data[i,]$IPs,]$milliseconds)
+    }
+    data$i <- 1:nrow(data)
+    ggplot(data=data, aes(x=IPs,y=milliseconds,color=howmany))+ scale_colour_gradient(name  ="Count", low="#11AA99",high='#EE3333')+geom_point(size=4)+scale_y_log10()+ stat_summary(fun.data = "mean_cl_boot", colour = "black")+ theme(text = element_text(size=25), legend.key.height=unit(2, "line"), axis.text.y = element_text(angle=90, vjust=0.5)) 
+    ggsave(paste0("../img/time-vs-ips-alife-",preffix,".png"))
+    ggplot(data=data, aes(x=i,y=milliseconds,fill=IPs))+geom_bar(stat='identity')+scale_y_log10()
+    ggsave(paste0("../img/duration-vs-time-alife-",preffix,".png"))
+    ggplot(data=data, aes(x=i,y=milliseconds,color=IPs))+geom_point()+scale_y_log10() + scale_color_gradient(low='blue', high='yellow')
+    ggsave(paste0("duration-vs-time-alife-bar-",preffix,".png"))
+    ggplot(data=data, aes(x=reorder(i,-milliseconds),y=milliseconds,color=IPs)) +geom_point()+scale_y_log10() + scale_color_gradient(low='#11AA99', high='#EE3333') + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.y = element_text(angle=90, vjust=0.5) )+labs( x="Ordered experiments")
+    ggsave(paste0("../img/time-vs-rank-alife-",preffix,".png"))
+}
+
+suffix <- c("128","64","32")
+all.openshift <- data.frame(Experiment=as.Date(character()),
+                            IPs=character(), 
+                            milliseconds=character()) 
+for ( i in suffix ) {
+    ips.time <- read.csv(paste0("2016-alife/ips-time-openshift-2016-02-cache=",i,".csv"))
+    all.openshift <- rbind(all.openshift,
[…]

### 2016-02-29 Alife16/paper.tex
First comments of new results
Please @mariosky @fergunet and @unintendedbear revise it to see if this checks out.

@@
 \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-128.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-64.png}
+\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-32.png}
 \caption{Duration of experiments vs. number of different IPs (nodes)
@@
 %
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
-  logarithmic scale. Dot color is related to the number of IPs
-  participating in the experiment. From left to right, experiments
-  4/4, 4/24 and 7/31.} 
-\label{fig:zipf:os}
-% First ones made with time-vs-IPs-openshift.R:  
-\end{figure}
-%
 We will have to analyze experimental data a bit further to find out why
@@
 and might in fact be abandoned with no one contributing for some time,
-that increases the probability of someone new joining them. In fact,
+the probability of someone new joining them is higher. In fact,
 with experiments taking a few seconds and due to the way the
@@
 into account that volunteers are not {\em carried over} from previous
-experiments. This implies that it would be convenient to use a problem
-of a bigger size to check this hypothesis as we have done in the next
-experiment.
+experiments.
+
+That is why we used a more difficult problem in the second batch of
+experiments, which is shown in bottom row of Figure
+\ref{fig:duration}. The pattern is remarkably similar, showing a
+positive correlation between the time for solving the problem and the
+number of computers, at least for cache sizes 128 and 64. However, it
+is interesting to observe that, for cache=32, the time needed to find
+the solution decreases from one to approximately 4-5 nodes, to then
+increase for a higher number of participating computers, distinguished
+by IP. The green dot at the bottom is probably and outlier that we
+will try to explain later on. This leads us to conclude that a better
+amount of computers might contribute to speed up the solution, if the
+time the experiment ideally takes is sufficient, that is, of the order of a
+minute, and enough volunteers concur simultaneously. This is also
+observed, not so clearly, in the case of cache=64, with an interval of
+around 10 IPs obtaining less time than experiments with less or more
+IPs, and of the same order, between 10 and 100 seconds.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
+\caption{Duration of experiments vs. rank, with $y$ axis in a
+  logarithmic scale. Dot color is related to the number of IPs
+  participating in the experiment. From left to right, top to bottom, experiments
+  4/4, 4/24 and 7/31 and caches=128,64, 32} 
+\label{fig:zipf:os}
+% First ones made with time-vs-IPs-openshift.R:  
+\end{figure}
+%
 

### 2016-02-29 Alife16/paper.tex
Adds weibull fits

@@
 of computers happened less frequently, while darker shadow means more frequency. 
-From left to right, experiments 4/4, 4/24 and 7/31.}
+From left to right and top to bottom experiments 4/4, 4/24 and 7/31,
+followed by experiments with 50 traps, cache=128, 64, 32.}
 \label{fig:duration}
@@
 \includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-128.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-64.png}
+\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-32.png}
 \caption{Duration of experiments vs. rank, with $y$ axis in a
   logarithmic scale. Dot color is related to the number of IPs
-  participating in the experiment. From left to right, top to bottom, experiments
+  participating in the experiment. From left to right and top to bottom, experiments
   4/4, 4/24 and 7/31 and caches=128,64, 32} 
@@
 %
+It is also interesting to check the distribution of the experiment
+duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
+a Zipf's law, with similar distribution along all three runs. The 4/24
+run is the most complete and shows a S-shape, which implies an
+accumulation of experiments taking similar time and around 100
+seconds; this S-shape appears too in the experiments with cache=128
+(bottom row, left). The most interesting part is the {\em tail}, which shows how
+many experiments took a desirable amount of time, of the order of
+10 seconds, and which appears in all three graphs. As it can be seen,
+it sharply drops implying there are 
+just a few of them, and with diminishing probability as time
+decreases. However, since they have a greenish color, implying a low
+number of IPs, they might be due to clients {\em carrying over} from
+the previous one. This is a characteristic of this implementation
+which will be examined later on, but at any rate if we discard those
+experiments that take too much or too little, there is a decreasing
+exponential distribution that corresponds to the Zipf's law.
 
@@
 \centering
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-4.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-24.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
+\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
+\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
+\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
+\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
+\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
+\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
 \caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
@@
 %
-It is also interesting to check the distribution of the experiment
-duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
-a Zipf's law, with similar distribution along all three runs. The 4/24
-run is the most complete and shows a S-shape, which implies an
-accumulation of experiments taking similar time and around 100
-seconds. The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, of the order of
-10 seconds, and which appears in all three graphs. As it can be seen,
-it sharply drops implying there are 
-just a few of them, and with diminishing probability as time
-decreases. This exponential distribution also appears in the
+ This exponential distribution also appears in the
 distribution of HTTP {\tt PUT}s, equivalent to the number of

### 2016-02-29 Alife16/paper.tex
Adding all new graphs.
For comparison. Changed GEV to Weibull. Please @unintendedbear @fergunet @mariosky Do something! Check for typos, inconsistencies, and so on..

Still some stuff to go, comparison with the new cache and general conclusions.

@@
 \includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
-\caption{Number of {\tt PUT}s per unique IP and fit to a Generalized
-  Extreme Value distribution (in red). From left to right, experiments
-  4/4, 4/24 and 7/31.} 
+\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
+  distribution (in red). From left to right and top to bottom, experiments
+  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
 % Plotted with ../data/plot-zipf-openshift.R
@@
 
-This work has been supported in part by TIN2011-28627-C04-02 and
-TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
-PROY-PP2015-06 (Plan Propio 2015 UGR),  % please, remember to include all the current projects   ;)   Thanks!
-SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
-GENIL project (CEI-BIOTIC Granada). We would also like to thank the
+This work has been supported in part by TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
+PROY-PP2015-06 (Plan Propio 2015 UGR)  % please, remember to include all the current projects   ;)   Thanks!
+. We would also like to thank the
 anonymous reviewers of previous versions of this paper who have really

### 2016-02-29 Alife16/paper.tex
Minor corrections and some questions for @JJ in SOA section for issue #101 (although it was closed)

@@
 
-Volunteer computing involves an user running a program voluntarily
+Volunteer computing involves a user running a program voluntarily
 and, as such, has been deployed in many different ways from the
@@
 processing extraterrestrial signals \citep{david-seti:home}. However
-the dual introduction of JavaScript as an universal language for the
+the dual introduction of JavaScript as a universal language for the
 browser and the browser as an ubiquitous web and Internet client has
 made this combination the most popular for volunteer computing
-frameworks such as the one we are using here and whose first version
+frameworks such as the one we are using here, and whose first version
 was described in \citep{DBLP:conf/gecco/GuervosG15}.
@@
 \citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
+%Pablo: Desell in 2008 is not so recent
 used Java \citep{chong:1999:jDGPi} and others have gone away from the
@@
 need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower.
+threshold to join them is much lower. %Pablo: this is useful for us because...
 
@@
 study was performed by Martínez et al. on the Capataz system
-\citep{martinez2015capataz}; however in this case the number of
+\citep{martinez2015capataz}; however, in this case the number of
 computers used was known in advance and the main focus was on
 measuring the speed up and how job bundling helped to reduce overhead and
-enhance performance.
+enhance performance. %Pablo: On the contrary, in this paper, we...
 
@@
 that measure the performance of a sociotechnical
-metacomputer. Apolónia et al. \citep{apolonia2012enhancing} used the
+%Pablo: I miss some definition for the concept "sociotechnical metacomputer" in the introduction
+metacomputer. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the
 Facebook protocol to distribute tasks among the {\em walls} of
@@
 it stopped short of relating performance to the macro measures of the
-users' social networks. 
+users' social networks. %Pablo: Add a phrase here to better link with the previous description of the work of Apolonia (it is a sudden huge jump from one topic to another)
 The algorithms used, as well as the methodology 

### 2016-02-29 Alife16/paper.tex
Addressing @fergunet comments
Refs #27b1590

@@
 makes an experiment successful or not, including the feedback loop
-that is created between the users and the algorithm itself.
+that is created between the users and the algorithm itself. In this
+paper we will examine what kind of patterns we can expect and what
+influence those patterns have on the implementation of the
+evolutionary algorithm. 
 %Pablo: The patterns we are looking for are... 
@@
 need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower. %Pablo: this is useful for us because...
+threshold to join them is much lower; this makes gathering new users
+immediate and spontaneous, so that they can be used for short
+algorithms on call.
 
@@
 the nodes available. Lately, several approaches have focused on the
-fault-tolerance of volunteer algorithms %volunteer systems?
+fault-tolerance of volunteer systems
 \citep{gonzalez2010characterizing} which it can, of course, be studied in
@@
 measuring the speed up and how job bundling helped to reduce overhead and
-enhance performance. %Pablo: On the contrary, in this paper, we...
+enhance performance. On the contrary, in this paper, we will use {\em
+  actual} volunteers, anonymous persons that are requested to join an
+experiment through social networks.
 
@@
 that measure the performance of a sociotechnical
-%Pablo: I miss some definition for the concept "sociotechnical metacomputer" in the introduction
-metacomputer. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the
+metacomputer, that is, a spontaneously created parallel computer that
+uses social networks for operations such as gathering new users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the
 Facebook protocol to distribute tasks among the {\em walls} of
@@
 it stopped short of relating performance to the macro measures of the
-users' social networks. %Pablo: Add a phrase here to better link with the previous description of the work of Apolonia (it is a sudden huge jump from one topic to another)
+users' social networks. As in the previous example, a social network
+was used to get new network nodes; in the previous case a web page was
+used, while Facebook's wall was used here.
 The algorithms used, as well as the methodology 

### 2016-02-29 Alife16/paper.tex
processing the Weibull distribution

@@
 metacomputer, that is, a spontaneously created parallel computer that
-uses social networks for operations such as gathering new users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the
+uses social networks for operations such as gathering new
+users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
 Facebook protocol to distribute tasks among the {\em walls} of
@@
 cloud or bare metal based volunteer evolutionary computing system
-derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
- the {\sf NodEO} library, whose architecture has been
+derived from the {\sf NodEO} library, whose architecture has been
 developed using JavaScript on the client as well as the server.
@@
 from \url{https://github.com/JJ/splash-volunteer}.
-
-%PABLO: el siguiente párrafo justifica Javascript
 
@@
 %
- This exponential distribution also appears in the
-distribution of HTTP {\tt PUT}s, equivalent to the number of
-generations divided by 100, contributed by every user, which is shown
+A similar exponential distribution also appears if we rank HTTP {\tt
+  PUT}s, equivalents to the number of 
+generations divided by 100, or to evaluations divided by 12800,
+contributed by every user, which is shown 
 in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
-so we have fitted it to the Generalized Extreme Value distribution,
-with the resulting parameters shown in Table \ref{tab:puts:os}.
+that is, a power law with a small {\em bump} in the lowest
+values. After testing the Generalized Extreme Value distribution and
+failing for the new batch of experiments, we have fitted it to a
+Weibull distribution \cite{thoman1969inferences} with the resulting
+parameters shown in Table \ref{tab:puts:os}. 
 %
 \begin{table}
-\caption{Summary of fit to Generalized Extreme Value distribution of
+\caption{Weibull distribution  parameters of the fit of
   the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
 \begin{center}
-\begin{tabular}{l|ccc}
+\begin{tabular}{l|cc}
 \hline
-Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
+Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
 \hline
-4/4 &  8.541 $\pm$ 1.0926  & 12.442 $\pm$ 1.7302 &  1.388 $\pm$
-0.1377 \\
-4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\
-7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\
+4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
+4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
+7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
 \hline
+Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
+Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
+Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
+\hline
+% from plot-zipf-openshift.R and ips-puts-alife.R
 \end{tabular}

### 2016-02-29 data/plot-zipf-openshift.R

@@ -1,5 +1,6 @@
 library(ggplot2)
 library(evd)
+library(fitdistrplus)
 
 process.puts <- function(suffix ) {
     these.puts <- read.csv(paste0("ips-puts-openshift-",suffix,".dat"))
@@ -17,13 +18,14 @@ process.puts <- function(suffix ) {
 
     ## Weibull
     this.weibull <- fitdistr(this.df$puts,"weibull")
+    print(paste0("Fit ",preffix, " Weibull " ))
     print(this.weibull)
 
     this.fit.weibull <- data.frame(rank=1:length(this.df$puts),
                                 y=sort(rweibull(length(this.df$puts),
                                                 this.weibull$estimate["shape"],
                                                 this.weibull$estimate["scale"]),decreasing=T))
-    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.weibull,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=25), axis.text.y = element_text(angle=90, vjust=0.5))
+    ggplot()+geom_point(data=this.df,aes(x=rank,y=puts,color='data',shape='data'))+geom_point(data=this.fit.weibull,aes(x=rank,y=y,color='fit',shape='fit'))+scale_y_log10()+ scale_colour_manual(name  ="Puts", breaks=c("data", "fit"), labels=c("Data", "Fit"), values=c("#11AA99",'#EE3333')) + scale_shape_discrete(name  ="PUTs", breaks=c("data", "fit"), labels=c("Data", "Fit"))+theme(text = element_text(size=25), axis.text.y = element_text(angle=90, vjust=0.5))
     ggsave(paste0("../img/weibull-puts-openshift-",suffix,".png"),width=5,height=5)
     return(this.df)
 }

### 2016-02-29 Alife16/paper.tex
Finishing the explanation of Weibull

@@
 %
-This distribution was originally proposed to fit extreme values
-\citep{resnick2013extreme} and contains, as a special case, the inverse
-Weibull distribution which was fitted to volunteer computing
-frameworks such as SETI@home \citep{javadi2009mining}. We obviously do
-not pretend to compare our system in scale or complexity with it, but
-to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which also appears in our system. This
-distribution is governed by three parameters, the usual location $\mu$
-which is related to where it has its {\em center} and a scale $\sigma$,
-related to the size, but also a third shape $\xi$ parameter that is
-related to its skewness, that is, how skewed it is around the central
-location. Positive parameters indicate that the distribution {\em
-  leans} towards the origin, and negative ones towards the other extreme
-value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
-greater than one and between one and 1.4, which indicates that the
-three experiments share this origin-leaning pattern, with many users
-donating a few cycles and just a few donating extreme values. Random
-distributions with these parameters have been plotted in red in
-Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
-predictive value of these fits, however, is limited, over all taking
-into account the low correlation between successive events shown
-above.
+The inverse
+Weibull distribution is a special case of the GEV distribution we have
+used in those papers, and appears usually in natural sciences and
+artificial life, usually related to decay. I has been frequently
+fitted to volunteer computing frameworks such as SETI@home
+\citep{javadi2009mining}. The model that user behavior follows can be
+explained straighforwardly: when users visit the page, it draws their
+attention for a limited amount of time. They give it a chance for a
+few seconds. If something there amuses them or they can engage in a
+conversation about it, they stay for a while longer, otherwise, they
+leave. The {\em scale} parameter, which is around 20-40 in the first
+batch for the 40 Trap problem and betwen 160 and 200 in the 60 traps
+problem, depends mainly on the maximum number of generations people
+leave it running. Since it finishes or stalls after a number of
+generations shown in Figure \cite{tab:summary:os}, volunteer just
+leave after that. Curiously enough, the scale parameter is roughly
+twice the median number of {\tt PUT}s per experiment, showing that, on
+average, the most loyal users reload the page twice after finishing or
+after seeing the evolution does not progress. This rule of thumb
+breaks down with the last experiment, however, which is interesting by
+itself, too.
 
-At any rate, this also shows that a convenient way of increasing the
-computing power would be increasing this minimum amount per user. This
-issue led us to create the next version of the framework, which is
-presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
+The slope or shape parameter, on the other hand, indicates the overall
+shape of the curve. A value less than 1 indicates a concave (in an
+non-algorithmic scale) curve,
+with figures closer to one indicating a smaller slope. In all cases
+values are between 0.54 and 0.77, independently of the experiment. It
+might be the case that this number depends more on the total number of
+experiments carried out, with sets with more experiments, both in the
+middle, having values between 0.60 and 0.70. The distribution is
+remarkably similar which gives us a model of user behavior that is, to
+a certain extent, independent of the experiment.
 

### 2016-02-29 Alife16/paper.tex
Adds graph with IPs per minute

@@
 experiments. 
+%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.95\linewidth]{ips-per-minute-cache=64.png}
+\caption{Simultaneous IPs every minute of the experiment with cache =
+  64. \label{fig:otisdriftwood}}
+\end{figure}
 
@@
 around 10 IPs obtaining less time than experiments with less or more
-IPs, and of the same order, between 10 and 100 seconds.
+IPs, and of the same order, between 10 and 100 seconds. If we look at
+the graph that shows the number of IPs or volunteers per minute for
+this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
+there are peaks of more than 25 volunteers, and a period of several
+hours with a minimum of 6 computers and peaks of more than 10. The
+long period after midnight where there is a single volunteer left
+masks the success achieved during this set of experiments, from which
+we draw two lessons: first, you need a social network influencer to
+announce your experiments and second, no matter what, do not do any
+experiment after midnight.
 
@@
 
+These experiments show that, as it was proved for other volunteer
+computer frameworks and also in the case of games, user engagement
+follows a Weibull distribution. This makes engagement the key for
+leveraging the performance of the sociotechnical metacomputer and a
+way to improve results in the future.
+
 %---------------------------------------------------------------

### 2016-02-29 Alife16/paper.tex
explaining a joke makes it less funny, but...
First part of experiments
@JJ please check some inconsistencies

@@
 
-In this case the classical Trap function \citep{Ackley1987} has been
-used. JavaScript is a functional language and declaring a different 
-function and handing at the creation of the algorithm object, called
-{\tt Classic}, is the only change needed to work with a different
-problem. The next Section will describe experiments performed to
-establish baseline performance and gather initial performance
+JavaScript is a functional language, so in order to work with a
+problem, a fitness function must be supplied at the creation of the algorithm object, 
+called {\tt Classic}. In this case the classical Trap 
+function \citep{Ackley1987} has been used. Depending on the
+problem additional handing functions and configurations 
+can be supplied.
+
+The next Section will describe experiments performed to
+establish a baseline performance and gather initial performance
 results. In the experiments performed last year we used the 40-trap
-problem, while the current experiments changed to the more difficult 50-trap.
+problem, while the current experiments changed to the more difficult 50-trap
+in order to compare the performance with a more computationally intensive problem.
 
@@
 PaaS, which provides a free tier, making the whole experiment cost
-equal to 0. Experiments were
-announced through a posts in Twitter and, the latest case, Telegram, and
+equal to \$0.00. Experiments were
+announced through a series of posts in Twitter and, in the latest case, Telegram, and
 results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
 purpose of this paper, we repeated the announcement several times
-through the month of April and then by the end of February. All
+through the month of February. All  %% April --> January ??
 in all, we have the set of runs with the characteristics shown in
@@
 or the particular wording. Every {\em experiment} consisted in running
-until the solution of the 40-trap problem was found. When the correct
+until the solution of the 50-trap problem was found. When the correct
 solution was sent to the server, the counter was updated and the pool
@@
 announce your experiments and second, no matter what, do not do any
-experiment after midnight.
+experiment after midnight. This statement, which might seem tongue in
+cheek, in fact it is a conclusion drawn from the experimental data and
+to what extent the social network is an essential part of the
+description and performance of the NodIO volunteer computing system.
 

### 2016-02-29 Alife16/paper.tex
Starting with conclusion

--- 2016-02-29 Alife16/paper.tex
@@
 
-In this paper two versions of a client-server architecture for volunteer and distributed
-evolutionary algorithms that uses the browser and generated using {\sf
-  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
-evaluated. Volunteers are {\em in the cloud}, as stated in the title,
-since they are a {\em CPU as a service} for the persons running the
-experiment. In fact, in this paper we have tried to put some figures
+Our intention in this paper was to assess the capabilities of a
+sociotechnical system formed by a client-server web-based framework
+running a distributed evolutionary algorithms and the volunteers that
+participate in the experiment. These volunteers are {\em in the
+  cloud}, that is, available as {\em CPU as a service} for the persons
+running the experiment. In this paper we have tried to put some figures
 on the real size of that {\em cloud} and how it can be used standalone
-if there is no alternative or in conjunction with other local or
-cloud-based methods to add computing power in a seamless way through
-the pool that NodIO creates.
+if there is no alternative, or, if other computing resources are
+available, in conjunction with other local or cloud-based methods to
+add computing power in a seamless way through the pool that NodIO creates.
 
-In order to establish a baseline performance, the evolutionary
-algorithm was run in a desktop client-program written in JavaScript
-using NodEO to solve the 40-trap function. The first experiments with
-{\sf NodIO} proved that, although obtaining better performance than the
-baseline was possible, it did not happen, mainly because of the
-difficulty in carrying over volunteers to other experiments when the
-one they were participating on was finished, shown in the low
-correlation between the number of IPs in successive experiments. This
-also resulted in a low number of generations allotted by users.
+After running the experiment on the 40-trap problem whose running time
+could be as low as a few seconds, we switched to another batch of
+experiments where we used the 50-trap problem and also to a pool of
+limited, and dwindling through the three experiments, size. Since our
+initial results indicated that what happened on the screen, a flat
+graph with no improvements or the experiment finished, influenced the
+amount of time that the users devoted to the experiment, a longer one
+could yield different results and, at the same time, result in a big
+pool that might either crash the server or return useless individuals
+to the volunteers. These new experiments have proved that using the
+limited pool is beneficial to finding the solution, since less
+evaluations are needed, but also that since the problem is more
+difficult, the users stay for longer in the web page, making less
+ephemeral the sociotechnical computing system created by the
+simulation.
 
@@
 the model depends on the implementation, with contributions following
-a General Extreme Values distribution in the case of {\sf NodIO} and a
-Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
-close enough to each other and, in the second case, reflect the fact
-that all time spent in the page is actually devoted to computing,
-which is why the time spent (represented by the number of
-contributions to the pool) follows a model quite similar to that found
+a Weibull distribution, which reflects the fact
+that volunteer computing follows a model quite similar to that found
 for games or other online activities. The reverse might be true: if we
@@
 gamification might involve computing in real time data such as the one
-presented in this paper and showing it in the same page.
+presented in this paper and showing it in the same page or presenting
+user results alongside others.
 

### 2016-02-29 Alife16/paper.tex
Accepting changes by @mariosky
But don't erase graphs, although it's easy to generate them, we might need them in its current state later on.

@@
 establish a baseline performance and gather initial performance
-results. In the experiments performed last year we used the 40-trap
+results. In the first set of experiments, performed last year, we used the 40-trap
 problem, while the current experiments changed to the more difficult 50-trap

### 2016-02-29 Alife16/paper.tex
Includes screenshot

@@
 algorithm behavior. Different web technologies, such as JQuery or {\tt
-  Chart.js} have
-been used to build the user interface elements of the framework.
+  Chart.js} have been used to build the user interface elements of the
+framework, a part of which is shown in Figure \ref{fig:screenshot} and
+should be running in \url{http://nodio-jmerelo.rhcloud.com}.
 
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.95\linewidth]{nodio-screenshot.png}
+\caption{Screenshot showing the fitness and number of users as seen by
+volunteers. The one on the left shows fitness, on the right the
+accumulated number of users. \label{fig:screenshot}}
+\end{figure}
+%
 \begin{figure}[!t]
@@
 leave it running. Since it finishes or stalls after a number of
-generations shown in Figure \cite{tab:summary:os}, volunteer just
+generations shown in Figure \ref{tab:summary:os}, volunteer just
 leave after that. Curiously enough, the scale parameter is roughly

### 2016-02-29 Alife16/paper.tex
Hope this is better refs 939f886
First full draft

@@
 used, while Facebook's wall was used here.
-The algorithms used, as well as the methodology 
+
+In our case, social networks are an integral part of the system and
+used to spontaneously obtain users. The algorithms used, as well as the methodology 
 for gathering resources will be described next, 
@@
 performance is an interesting avenue to explore in the future. Even if
-the three previous experiments were published in a similar way, one
+these  experiments were published in a similar way, one
 obtained up to 5 times more total cycles than the one with the least
-number of cycles. It is also essential to obtain volunteers as
+number of cycles. It is also essential to obtain volunteers as fast and
 simultaneously as possible, so it is possible that the features of the
 social network in terms of real-time use will also play a big
-role. Even as it is difficult to create controlled experiments in this
+role; synchronous webs such as Snapchat, which mystifies the writers
+of this paper, and Twitter, thanks to its real time nature, might be
+better suited than Facebook, LinkedIn or Google plus. Even as it is
+difficult to create controlled experiments in this 
 area, it is an interesting challenge to explore in the future.
 
-Although we think that, in general, the results presented in this
-paper are independent of the problem chosen, it is true that even if
-it is a difficult problem for evolutionary algorithms, it takes a few
-seconds with the right settings to solve. 
-Besides, the two versions of the framework include as an
-algorithmic variant using random population size. We do not think this
-has had a big influence on the results and in fact this is not noticed
-for F15, which is the second version tested but it would be interesting to
-measure exactly what this influence has been. In general there are
+The other area to explore is the algorithmic area itself. Are there
+ways to change the evolutionary algorithm, or its visualization, so
+that the user has a bigger impact on the result? One of the users in
+Twitter even suggested to embed videos so that people spent time
+looking at them, but other possible way was to make the user engage
+the algorithm by giving him or her buttons to change the mutation rate
+when the algorithm is stalled, for instance. If this is combined with
+a score board where local performance is compared to other users,
+engagement might be increased and thus the performance of the
+system. In general there are 
 many issues with the evolutionary algorithm implementation itself,
@@
 
-Another area is how to enhance the number and quality of
-volunteers. For instance, adding a bit of more
-control to the user might contribute also to gamification. Right now
-the user has only two Web Workers. It is a matter of a single click to
-open more tabs in the browser, giving more Web Workers, but it would be
-interesting to put everything in a single page and under our control,
-to check how often this happens and under which conditions.
+Finally, the implementation needs some refinement in terms of
+programming and also ease of use. Tools such as Yeoman for generating
+easily new experiments might be used, so that the user would have  to
+create only a fitness
+function, with the rest of the framework wrapped around
+automatically.
 
-Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
-to create a generator in which the user just has to create a fitness
-function, with the rest of the framework wrapped around
-automatically. All this, as the data used for this paper and the paper
-itself, has been published with a free license in GitHub at
+All these avenues of experimentation will be done openly following the
+Open Science policy of our group, which, in fact, contributes to
+establish trust and security between us and volunteers and is an
+essential feature of the system. That is why this paper, as well as
+the data and processing scripts, are published with a free license in GitHub at
 \url{https://github.com/JJ/modeling-volunteer-computing}.

### 2016-02-29 Alife16/paper.tex
Final draft

@@
 
-\title{The human in the loop: volunteer computing as a socio-technical
+\title{The human in the loop: volunteer-based metacomputers as a socio-technical
   system}
@@
 these  experiments were published in a similar way, one
-obtained up to 5 times more total cycles than the one with the least
+obtained up to five times more total cycles than the one with the least
 number of cycles. It is also essential to obtain volunteers as fast and

### 2016-02-29 Alife16/paper.tex
More typos and suggestions in results section, although it has been already submitted :P

--- 2016-02-29 Alife16/paper.tex
@@
 algorithm behavior. Different web technologies, such as JQuery or {\tt
-  Chart.js} have been used to build the user interface elements of the
-framework, a part of which is shown in Figure \ref{fig:screenshot} and
-should be running in \url{http://nodio-jmerelo.rhcloud.com}.
+  Chart.js} have
+been used to build the user interface elements of the framework.
 
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.95\linewidth]{nodio-screenshot.png}
-\caption{Screenshot showing the fitness and number of users as seen by
-volunteers. The one on the left shows fitness, on the right the
-accumulated number of users. \label{fig:screenshot}}
-\end{figure}
-%
 \begin{figure}[!t]
@@
 equal to \$0.00. Experiments were
-announced through a series of posts in Twitter and, in the latest case, Telegram, and
+announced through a series of posts on Twitter and, in the latest case, Telegram, and
 results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
@@
 days. No particular care was taken about the time of the announcement
+%Pablo: but it would great at least describe how many times was announced and how long
 or the particular wording. Every {\em experiment} consisted in running
@@
 solution was sent to the server, the counter was updated and the pool
-of solutions reset to the void set. There was no special intention to wait
-until all clients had finished, thus it might happen that, in fact,
+of solutions resets to the void set. There was no special intention to wait
+until all clients had finished, thus it might happen that. In fact,
 the islands running in the browser {\em spill} from one experiment to
 the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible.
+of these islands in the next experiment is indeed negligible. %Pablo: cite
 
 The table shows that every run included more than 30 experiments. The
+%Pablo: reference to table? (maybe the final print will misplace it)
 number of different IPs intervening in them varied from more than one
@@
 We will have to analyze differently the median time, since the two
+%Pablo: different verbal times ("will have", and before, "is shown"). Homogenize.
 batches are solving different problems. In both cases it possesses a
@@
 only lasted for a few hours, with a more continued effort of
-publicizing it in social networks; this is specially true in the
+publicizing it in social networks. This is specially true in the
+%Pablo: removed semicolon before "This". @unintendedbear does not like them :P
 case of cache equal to 64, which is noticed by the high number of
@@
 making it limited. As it can be seen in the table, there is a clear
+%Pablo: which table?
 change in the number of evaluations needed, with smaller cache sizes
@@
 before. Besides, cache policy deletes the oldest individuals, which
-makes those in the pool be {\em current}, helping then newcomers and
+makes those in the pool be {\em current}, helping then newcomers and %Pablo: "then helping newcomers", instead?
 any participant obtain the best individuals found in the last part of
@@
 That is why we used a more difficult problem in the second batch of
-experiments, which is shown in bottom row of Figure
+experiments, which is shown in the bottom row of Figure
 \ref{fig:duration}. The pattern is remarkably similar, showing a
@@
 increase for a higher number of participating computers, distinguished
-by IP. The green dot at the bottom is probably and outlier that we
+by IP. The green dot at the bottom is probably an outlier that we
 will try to explain later on. This leads us to conclude that a better
@@
 we draw two lessons: first, you need a social network influencer to
+%Pablo: people know the meaning of "influencer"? Probably yes, but I mention it just in case.
 announce your experiments and second, no matter what, do not do any
 experiment after midnight. This statement, which might seem tongue in
-cheek, in fact it is a conclusion drawn from the experimental data and
+cheek, in fact, it is a conclusion drawn from the experimental data and
 to what extent the social network is an essential part of the
@@
 (bottom row, left). The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, of the order of
+many experiments took a desirable amount of time, on the order of
 10 seconds, and which appears in all three graphs. As it can be seen,
@@
 the previous one. This is a characteristic of this implementation
-which will be examined later on, but at any rate if we discard those
+which will be examined later on, but at any rate, if we discard those
 experiments that take too much or too little, there is a decreasing
@@
 used in those papers, and appears usually in natural sciences and
-artificial life, usually related to decay. I has been frequently
+artificial life, usually related to decay. It has been frequently
 fitted to volunteer computing frameworks such as SETI@home
@@
 leave. The {\em scale} parameter, which is around 20-40 in the first
-batch for the 40 Trap problem and betwen 160 and 200 in the 60 traps
+batch for the 40 Trap problem and between 160 and 200 in the 60 traps
 problem, depends mainly on the maximum number of generations people
 leave it running. Since it finishes or stalls after a number of
-generations shown in Figure \ref{tab:summary:os}, volunteer just
-leave after that. Curiously enough, the scale parameter is roughly
+generations shown in Figure \cite{tab:summary:os}, volunteer just
+leaves after that. Curiously enough, the scale parameter is roughly
 twice the median number of {\tt PUT}s per experiment, showing that, on
@@
 The slope or shape parameter, on the other hand, indicates the overall
-shape of the curve. A value less than 1 indicates a concave (in an
+shape of the curve. A value less than 1 indicates a concave (in a
 non-algorithmic scale) curve,

### 2016-02-29 Alife16/paper.tex

@@
 leave it running. Since it finishes or stalls after a number of
-generations shown in Figure \cite{tab:summary:os}, volunteer just
-leaves after that. Curiously enough, the scale parameter is roughly
+generations shown in Figure \cite{tab:summary:os}, volunteers just
+leave after that. Curiously enough, the scale parameter is roughly
 twice the median number of {\tt PUT}s per experiment, showing that, on

### 2016-03-01 Alife16/paper.tex
Final changes after submission
Closes #100
Closes #102

@@
 Volunteer computing is a form of distributed computing where users
-decide on their participation and the amount of time and other resources they will
-``lend''. This makes them an essential part of the algorithm and of
-the performance of the whole system. As a sociotechnical system, this
-participation follows some patterns and in this paper we examine the
-result of several volunteer distributed evolutionary computation
-experiments and try to find out which are those patterns and what
-makes an experiment successful or not, including the feedback loop
-that is created between the users and the algorithm itself. In this
-paper we will examine what kind of patterns we can expect and what
-influence those patterns have on the implementation of the
-evolutionary algorithm. 
-%Pablo: The patterns we are looking for are... 
+decide on their participation and the amount of time and other
+resources they will ``lend''. This makes them an essential part of the
+algorithm and of the performance of the whole system. As a
+sociotechnical system, this participation follows some patterns and in
+this paper we examine the result of several volunteer distributed
+evolutionary computation experiments and try to find out which are
+those patterns and what makes an experiment successful or not,
+including the feedback loop that is created between the users and the
+algorithm itself. In this paper we will examine what kind of patterns
+we can expect and what influence those patterns have on the
+implementation of the evolutionary algorithm.
 \end{abstract}
@@
 experiments with volunteer-based ones, with all of them contributing
-to the pool, but every one of them working as separate island 
-carrying out their own evolutionary algorithms. This is why NodIO is
-proposed mainly as a {\em complement} to existing resources such as
-desktop systems or laptops. As long as it provides a non-null
-computational capability that can help existing resources find the
-solution faster it will have found its purpose. The main use case is
-someone setting up NodIO in the cloud, writing a fitness evaluation
-function and running a client from his or her own computer, but
-requesting help in social networks for additional resources. That is
-why, in this system,  we have considered the whole {\em social} aspect
-in the design, with issues related to security, trust and privacy
-among others. The computing system becomes a {\em techno-social
-%Pablo: In abstract is named 'sociotechnical'. Which one is the correct form for the rest of the paper?
-  system} \citep{vespignani2009predicting}. In this paper, we are going
-to focus on measuring the response of users to experiments, but at the
-%Pablo: which metric are we going to measure this response?
-same time we will also focus on the technical aspects of the server and how
-those might change the behavior of users, improving the capability of % Mario: How these might change
-the system.
+to the pool, but every one of them working as separate island carrying
+out their own evolutionary algorithms. This is why NodIO is proposed
+mainly as a {\em complement} to existing resources such as desktop
+systems or laptops. As long as it provides a non-null computational
+capability that can help existing resources find the solution faster
+it will have found its purpose. The main use case is someone setting
+up NodIO in the cloud, writing a fitness evaluation function and
+running a client from his or her own computer, but requesting help in
+social networks for additional resources. That is why, in this system,
+we have considered the whole {\em social} aspect in the design, with
+issues related to security, trust and privacy among others. The
+computing system becomes a {\em sociotechnical} system}
+\citep{vespignani2009predicting}. In this paper, we are going to focus
+on measuring the response of users to experiments, that is, the time
+they spend running it, but at the same time we will also focus on the
+technical aspects of the server and how these might change the
+behavior of users, improving the capability of the system.
 
@@
 transparent, incentives to cheat are reduced and, in fact, we have
-detected no issue for the time being. Next we present the state of the art in web-based
-volunteer-based systems along with attempts to predict and model its
-behavior.
+detected no issue for the time being. Next we present the state of the
+art in web-based volunteer computing systems along with attempts to
+predict and model its behavior.
 

### 2016-03-01 PPSN2016/paper.tex
Changed template for PPSN, closes #104

--- 2016-03-01 Alife16/paper.tex
@@
-\documentclass[letterpaper]{article}
-\usepackage{natbib,alifeconf}
-\usepackage[breaklinks=true]{hyperref}
+\documentclass[runningheads,a4paper]{llncs}
+
+\usepackage[latin1]{inputenc}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\setcounter{tocdepth}{3}
+\usepackage{graphicx}
+\usepackage{multirow}
+\usepackage{rotating}
+\usepackage{subfigure}
 \usepackage{url}
+\usepackage{caption}
+
+\newcommand{\keywords}[1]{\par\addvspace\baselineskip
+\noindent\keywordname\enspace\ignorespaces#1}
+
+\providecommand{\tabularnewline}{\\}
+
+%\documentclass[letterpaper]{article}
+%\usepackage{natbib,alifeconf}
+%\usepackage[breaklinks=true]{hyperref}
+%\usepackage{url}
 
@@
 
+\begin{document}
 \title{The human in the loop: volunteer computing as a socio-technical
 system}
-\author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
-  Garc\'ia-S\'anchez $^1$, Mario Garc\'ia-Valdez$^2$\\
-\mbox{}\\
-$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada \\
-$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana \\
-{\tt jmerelo@ugr.es}, {\tt mario@tectijuana.edu.mx}}
 
-\begin{document}
+\author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
+  Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+
+\institute{Dept. of Computer Architecture and Technology, University
+of Granada, Spain \and
+Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+
 \maketitle
+\thispagestyle{empty}
+\pagestyle{empty}
 
@@
 cloud resources as servers and browsers as clients. This framework was
-called NodIO \citep{2016arXiv160101607M}. NodIO provides server
+called NodIO \cite{2016arXiv160101607M}. NodIO provides server
 infrastructure for volunteer-based distributed evolutionary computing
@@
 issues related to security, trust and privacy among others. The
-computing system becomes a {\em sociotechnical} system}
-\citep{vespignani2009predicting}. In this paper, we are going to focus
+computing system becomes a {\em sociotechnical} system
+\cite{vespignani2009predicting}. In this paper, we are going to focus
 on measuring the response of users to experiments, that is, the time
@@
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \citep{david-seti:home}. However
+processing extraterrestrial signals \cite{david-seti:home}. However
 the dual introduction of JavaScript as a universal language for the
@@
 frameworks such as the one we are using here, and whose first version
-was described in \citep{DBLP:conf/gecco/GuervosG15}.
+was described in \cite{DBLP:conf/gecco/GuervosG15}.
 
 JavaScript can be used for either unwitting
-\citep{unwitting-ec,boldrin2007distributed} or volunteer 
-\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
+\cite{unwitting-ec,boldrin2007distributed} or volunteer 
+\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
 evolutionary computation and it has been used ever since by several
 authors, including more recent efforts
-\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
+\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
 %Pablo: Desell in 2008 is not so recent
-used Java \citep{chong:1999:jDGPi} and others have gone away from the
+used Java \cite{chong:1999:jDGPi} and others have gone away from the
 server-based paradigm to embrace peer to peer systems
-\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
+\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
 platforms avoid single points of failure (the server) but, since they
@@
 the algorithm adaptive to the kind of resources allotted to it
-\citep{milani2004online}, which is actually not such a big problem in
+\cite{milani2004online}, which is actually not such a big problem in
 algorithms such as the EA that can be easily 
@@
 fault-tolerance of volunteer systems
-\citep{gonzalez2010characterizing} which it can, of course, be studied in
+\cite{gonzalez2010characterizing} which it can, of course, be studied in
 the more general context of distributed computing 
-\citep{nogueras2015studying} or including it in a more general study of the
+\cite{nogueras2015studying} or including it in a more general study of the
 performance of the EA itself
-\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
+\cite{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
 measured without first understanding the dynamics of this kind of systems. Initial
 work was done for peer to peer systems by Stutzbach et
-al. \citep{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
+al. \cite{stutzbach2006understanding} and extended to volunteer
+computing by Laredo et al. \cite{churn08,laredo2008rcp}. A similar
 study was performed by Martínez et al. on the Capataz system
-\citep{martinez2015capataz}; however, in this case the number of
+\cite{martinez2015capataz}; however, in this case the number of
 computers used was known in advance and the main focus was on
@@
 only been studied in a limited way in 
-\citep{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
+\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
 run. Studies using volunteer computing platforms such as SETI@home
-\citep{javadi2009mining,agajaj} found out that the Weibull, log-normal and
+\cite{javadi2009mining,agajaj} found out that the Weibull, log-normal and
 Gamma distribution 
@@
 uses social networks for operations such as gathering new
-users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
+users. Apol{\'o}nia et al. \cite{apolonia2012enhancing} used the 
 Facebook protocol to distribute tasks among the {\em walls} of
@@
 so important; even so, we have made a comparison between JavaScript and other languages
-\citep{2015arXiv151101088M} that shows that performance of JavaScript
+\cite{2015arXiv151101088M} that shows that performance of JavaScript
 is comparable to other interpreted languages; compiled languages would
@@
   duties, but they are basically a very lightweight and high performance
-  data storage \citep{jj:idc:lowcost}.
+  data storage \cite{jj:idc:lowcost}.
   The server has the capability to
@@
 called {\tt Classic}. In this case the classical Trap 
-function \citep{Ackley1987} has been used. Depending on the
+function \cite{Ackley1987} has been used. Depending on the
 problem additional handing functions and configurations 
@@
 announced through a series of posts on Twitter and, in the latest case, Telegram, and
-results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
+results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
 purpose of this paper, we repeated the announcement several times
@@
 experiments involved 5 or less IPs. This is consistent with previous results
-\citep{DBLP:conf/gecco/GuervosG15} which found 6 to be
+\cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
 the expected number of volunteer IPs. The
@@
 fitted to volunteer computing frameworks such as SETI@home
-\citep{javadi2009mining}. The model that user behavior follows can be
+\cite{javadi2009mining}. The model that user behavior follows can be
 explained straighforwardly: when users visit the page, it draws their
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{splncs}
 \bibliography{volunteer,GA-general,geneura,javascript,ror-js}

### 2016-03-09 PPSN2016/paper.Rnw
R is the most irritant thing ever

--- 2016-03-01 PPSN2016/paper.tex
@@
 
-%\documentclass[letterpaper]{article}
-%\usepackage{natbib,alifeconf}
-%\usepackage[breaklinks=true]{hyperref}
-%\usepackage{url}
-
 \graphicspath{{../img/}}
@@
 \begin{document}
+
+<<setup, cache=FALSE,echo=FALSE>>=
+library("ggplot2")
+experiment.data <- <- data.frame(Experiment=character(),
+                                 newgraph = character(),
+                                 nooverlap = character(),
+                                 reboot = character(),
+                                 cachereboot = character(),
+                                 milliseconds=character(),                           
+                                 IPs=character(), 
+                                 actualIPs=character(),
+                                 PUTs=character(),
+                                 actualPUTs=character())
+these.experiments <- c("nooverlap-1","nooverlap-newgraph", "nooverlap-newgraph-2","nooverlap-newgraph-3","nooverlap-newgraph-reboot")
+
+for ( i in these.experiments ) {
+    print(i)
+    this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
+    newgraph = TRUE
+    nooverlap = TRUE
+    reboot = TRUE
+    cachereboot = TRUE
+    if ( ! grep("newgraph",i)) {
+        newgraph = FALSE
+    }
+    if ( ! grep("noover",i)) {
+        nooverlap = FALSE
+    }
+    if ( ! grep("reboot",i)) {
+        reboot = FALSE
+    }
+    if ( ! grep("cache",i)) {
+        cachereboot = FALSE
+    }
+
+    these.experiments <- rbind( these.experiments,
+                               data.frame( Experiment=rep(i, length(this.data$IPs)),
+                                          newgraph = rep(newgraph, length(this.data$IPs)),
+                                          nooverlap = rep(nooverlap, length(this.data$IPs)),
+                                          reboot = rep(reboot, length(this.data$IPs)),
+                                          cachereboot = rep(cachereboot, length(this.data$IPs)),
+                                          milliseconds=this.data$milliseconds,
+                                          IPs=this.data$IPs,
+                                          actualIPs=this.data$actualIPs,
+                                          PUTs=this.data$PUTs,
+                                          actualPUTs=this.data$actualPUTs))
+}
+
+@
 \title{The human in the loop: volunteer computing as a socio-technical

### 2016-03-09 PPSN2016/paper.Rnw
Solved data frame issue and plotted some things

--- 2016-03-01 PPSN2016/paper.tex
@@
 \begin{document}
+
+<<setup, cache=FALSE,echo=FALSE>>=
+library("ggplot2")
+experiment.data <- data.frame(Experiment=character(),
+                                 newgraph = character(),
+                                 nooverlap = character(),
+                                 reboot = character(),
+                                 cachereboot = character(),
+                                 milliseconds=character(),                           
+                                 IPs=character(), 
+                                 actualIPs=character(),
+                                 PUTs=character(),
+                                 actualPUTs=character())
+
+these.experiments <- c("nooverlap-1","nooverlap-newgraph", "nooverlap-newgraph-2","nooverlap-newgraph-3","nooverlap-newgraph-reboot")
+
+for ( i in these.experiments ) {
+    print(i)
+    this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
+    newgraph = TRUE
+    nooverlap = TRUE
+    reboot = TRUE
+    cachereboot = TRUE
+
+    experiment.data <- rbind( experiment.data,
+                             data.frame( Experiment=rep(i, length(this.data$IPs)),
+                                        newgraph = rep(newgraph, length(this.data$IPs)),
+                                        nooverlap = rep(nooverlap, length(this.data$IPs)),
+                                        reboot = rep(reboot, length(this.data$IPs)),
+                                        cachereboot = rep(cachereboot, length(this.data$IPs)),
+                                        milliseconds=as.integer(this.data$milliseconds),
+                                        IPs=as.integer(this.data$IPs),
+                                        actualIPs=as.integer(this.data$actualIPs),
+                                        PUTs=as.integer(this.data$PUTs),
+                                        actualPUTs=as.integer(this.data$actualPUTs)))
+}
+
+@
 \title{The human in the loop: volunteer computing as a socio-technical

### 2016-03-09 PPSN2016/paper.Rnw
Inserted data in paper

--- 2016-03-01 PPSN2016/paper.tex
@@
 \begin{document}
+
+<<setup, cache=FALSE,echo=FALSE>>=
+library("ggplot2")
+experiment.data <- data.frame(Experiment=character(),
+                              milliseconds=character(),                           
+                              IPs=character(), 
+                              actualIPs=character(),
+                              PUTs=character(),
+                              actualPUTs=character())
+
+these.experiments <- c("nooverlap-1","nooverlap-newgraph", "nooverlap-newgraph-2","nooverlap-newgraph-3","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
+
+for ( i in these.experiments ) {
+    this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
+
+    experiment.data <- rbind( experiment.data,
+                             data.frame( Experiment=rep(i, length(this.data$IPs)),                   
+                                        milliseconds=as.integer(this.data$milliseconds),
+                                        IPs=as.integer(this.data$IPs),
+                                        actualIPs=as.integer(this.data$actualIPs),
+                                        PUTs=as.integer(this.data$PUTs),
+                                        actualPUTs=as.integer(this.data$actualPUTs)))
+
+}
+
+experiment.data$newgraph <- grepl("newgraph",experiment.data$Experiment)
+experiment.data$nooverlap <- grepl("nooverlap",experiment.data$Experiment)
+experiment.data$reboot <- grepl("reboot",experiment.data$Experiment)
+experiment.data$cachecrash <- grepl("cache",experiment.data$Experiment)
+
+time.data <- data.frame(Experiment=character(),
+                        minute=as.Date(character()),
+                        IPs=character())
+
+these.experiments <- c("nooverlap-0","nooverlap","nooverlap-newgraph","nooverlap-newgraph-reboot-cachecrash")
+
+for ( i in these.experiments ) {
+    this.data <- read.csv(paste0("../data/2016-PPSN/ips-per-minute-cache=32-",i,".csv"))
+    time.data <- rbind(time.data,
+                       data.frame(Experiment=rep(i,length(this.data$IPs)),
+                                  minute=strptime(this.data$time,"%Y-%m-%d %H:%M"),
+                                  IPs=as.integer(this.data$IPs))
+                       )
+}
+@
+
 \title{The human in the loop: volunteer computing as a socio-technical

### 2016-03-11 PPSN2016/paper.Rnw
Rewrite abstract
I have done several changes, mainly on the client, as well as extensive experiments. Main change now is that overlap is almost entirely eliminated, that is, it is quite improbable that the clients started in a previous experiment contribute to a new one. However, this also makes changes in behavior of the users: they see changes if they leave it long enough. So let's see what goes on.

So far only thing changed is title and abstract. Please @mariosky @unintendedbear @fergunet take a look

@@
 
+experiment.aggregates <- data.frame(Experiment=character(),
+                                     milliseconds=character(),
+                                     IPs=character(),
+                                     actualIPs = character(),
+                                     PUTss = character(),
+                                     actualPUTs = character());
+
 for ( i in these.experiments ) {
     this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
+
+    experiment.aggregates <- rbind( experiment.aggregates,
+                                   data.frame( Experiment=i,                   
+                                              milliseconds=median(as.integer(this.data$milliseconds)),
+                                              IPs=median(as.integer(this.data$IPs)),
+                                              actualIPs=median(as.integer(this.data$actualIPs)),
+                                              PUTs=median(as.integer(this.data$PUTs)),
+                                              actualPUTs=median(as.integer(this.data$actualPUTs)))
+                                   )
 
@@
 these.experiments <- c("nooverlap-0","nooverlap","nooverlap-newgraph","nooverlap-newgraph-reboot-cachecrash")
+ip.correlations <- data.frame(Experiment=character(),
+                              correlation=character())
 
@@
     this.data <- read.csv(paste0("../data/2016-PPSN/ips-per-minute-cache=32-",i,".csv"))
+    this.data.plus1 <- this.data[2:length(this.data$IPs),]
+    ip.correlations <- rbind(ip.correlations, 
+                             data.frame(Experiment=i,
+                                        correlation = cor(this.data[c(1:length(this.data$IPs)-1),]$IPs
+                                                         ,this.data.plus1$IPs)))
     time.data <- rbind(time.data,
@@
 
-\title{The human in the loop: volunteer computing as a socio-technical
-system}
+\title{Evolutionary algorithms in a volunteer computing system:
+  performance assessment}
 
@@
 \begin{abstract}
-Volunteer computing is a form of distributed computing where users
-decide on their participation and the amount of time and other
-resources they will ``lend''. This makes them an essential part of the
-algorithm and of the performance of the whole system. As a
-sociotechnical system, this participation follows some patterns and in
-this paper we examine the result of several volunteer distributed
-evolutionary computation experiments and try to find out which are
-those patterns and what makes an experiment successful or not,
-including the feedback loop that is created between the users and the
-algorithm itself. In this paper we will examine what kind of patterns
-we can expect and what influence those patterns have on the
-implementation of the evolutionary algorithm.
+In a volunteer distributed computing system, users run a program in
+their own machine to contribute to a common effort. If they are
+embedded in a web page, collaboration is easy, but also ephemeral. In
+this paper we analyze a volunteer evolutionary computing system called
+NodIO and try to find out a few rules that encourage volunteer
+participation and how to make the most of their contributions to the
+evolutionary algorithm. We will show different measures of
+participation and contribution to the algorithm and how different
+volunteer usage patterns and tweaks in the algorithm contribute to
+improvements and leveraging of these contributions. 
 \end{abstract}

### 2016-03-11 PPSN2016/paper.Rnw
Filtering now by number of PUTs
This is a rather lame heuristic and I'm sure it can be beat, but it shows that if you take that "dummies" out experiments take much longer and need many more inputs.

@@
 
+experiment.aggregates.filtered <- data.frame(Experiment=character(),
+                                             milliseconds=character(),
+                                             IPs=character(),
+                                             actualIPs = character(),
+                                             PUTss = character(),
+                                             actualPUTs = character());
+
 for ( i in these.experiments ) {
@@
                                         actualPUTs=as.integer(this.data$actualPUTs)))
+
+    this.data <- this.data[this.data$actualPUTs>=10,]
+    experiment.aggregates.filtered <- rbind( experiment.aggregates.filtered,
+                                            data.frame( Experiment=i,                   
+                                                       milliseconds=median(as.integer(this.data$milliseconds)),
+                                                       IPs=median(as.integer(this.data$IPs)),
+                                                       actualIPs=median(as.integer(this.data$actualIPs)),
+                                                       PUTs=median(as.integer(this.data$PUTs)),
+                                                       actualPUTs=median(as.integer(this.data$actualPUTs))))
 

### 2016-03-11 PPSN2016/paper.Rnw
Completely rewritten introduction
Refs #109

Please @mariosky @unintendedbear @fergunet take a look; it describes the approach and what we are going to do in the paper and introduces new references. It probably should be shorter, because there is a 10 page limit, divided thus

* 1 page: intro
* 1 page: SoA
* 2 pages: Description framework
* 3 pages: results
* 1 page: conclusion
* 2 pages: Acknowledgements and references.

the 10 page limit is hard to do...

@@
 
-Some time ago, our group faced the problem of diminishing funds for
-buying new hardware. This was aggravated by the increasing maintenance
-costs and extended downtime resulting from the continuous failures of
-existing clusters.  Considering this, we leveraged our experience in
-the design of web applications with JavaScript and other volunteer and
-unconventional distributed evolutionary computing systems to design
-and release a new free framework that would allow anyone to create a
-volunteer distributed evolutionary computation (EC) experiment using
-cloud resources as servers and browsers as clients. This framework was
-called NodIO \cite{2016arXiv160101607M}. NodIO provides server
-infrastructure for volunteer-based distributed evolutionary computing
-experiments by providing a chromosome {\em pool}. This pool is used by
-clients in browsers and any other using the application programming
-interface (API) to put chromosomes and retrieve them, working then as
-a loose, asynchronous and {\em ad hoc} connection among all clients
-using it.
+The world wide web provides not only a platform for content
+distribution, but an increasingly reliable and high-performance
+platform for running distributed applications. This is based in two
+key facts: the Javascript virtual machine every browser runs and a
+simplified standard interface for interacting with servers. Thus
+creating a distributed computing experiment is just a matter of making
+a JavaScript application interchange information with a server, called
+rest. From the point of view of the programmer, this involves
+relatively common skills and no special libraries, since that
+interface is built in the browser, and a simple application that
+responds to those requests on the server side; both involve just a few
+dozens lines of code additionally to whatever business logic your
+application has. But, more importantly and from the point of view of
+the user, that application can be run by simply visiting a web page.
 
-This loose connection provides a low-overhead way to connect desktop
-experiments with volunteer-based ones, with all of them contributing
-to the pool, but every one of them working as separate island carrying
-out their own evolutionary algorithms. This is why NodIO is proposed
-mainly as a {\em complement} to existing resources such as desktop
-systems or laptops. As long as it provides a non-null computational
-capability that can help existing resources find the solution faster
-it will have found its purpose. The main use case is someone setting
-up NodIO in the cloud, writing a fitness evaluation function and
-running a client from his or her own computer, but requesting help in
-social networks for additional resources. That is why, in this system,
-we have considered the whole {\em social} aspect in the design, with
-issues related to security, trust and privacy among others. The
-computing system becomes a {\em sociotechnical} system
-\cite{vespignani2009predicting}. In this paper, we are going to focus
-on measuring the response of users to experiments, that is, the time
-they spend running it, but at the same time we will also focus on the
-technical aspects of the server and how these might change the
-behavior of users, improving the capability of the system.
+Using this system for creating distributed experiments is called {\em
+  volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing
+\cite{sarmenta2001volunteer} and it dates back, in different shapes
+and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
+here, however, is to use it as a resource for evolutionary
+computation. This has been a line of research in our group for a long
+time \cite{jj-ppsn98}, but has lately been revisited thanks to the
+ubiquity of social networks and handheld computing devices.
 
-Our research group is committed to open science, and we think this is
-a very important part of the techno-social system. By being
-transparent, incentives to cheat are reduced and, in fact, we have
-detected no issue for the time being. Next we present the state of the
-art in web-based volunteer computing systems along with attempts to
-predict and model its behavior.
+In this line of research that uses volunteer computing for
+evolutionary algorithms, there are several pending issues. The first
+and maybe most important is approaching volunteer computing as a
+techno-social system \cite{vespignani2009predicting} which integrates
+user decisions and behavioral patterns in the system model; this
+includes trying to optimize the number of users in a particular
+experiment. The second line of research, although related to the
+first, is more focused on the evolutionary algorithm itself and how
+different design decisions will affect its performance.
+
+In this paper our focus will be in the second aspect: we will try to
+design a decentralized system that, at the same time, is able to use
+all resources available. This design will be done incrementally by
+changing the system on the client and the server and measuring its
+impact on the overall performance: time and evaluations needed to find
+the solution. Eventually, we want to find a system that, whatever the
+number of users available to perform the experiment, is able to
+maximize their contribution to the evolutionary algorithm, at the same
+time that the evolutionary algorithm itself makes the most of those
+contributions, independently of how many they are.
+
+The rest of the paper is organized as follows: Next we will briefly
+present the state of the art in opportunistic distributed evolutionary
+computation (EC). Section \ref{sec:description} will describe the
+framework and problem used in the experiments, which are publicly
+available under a free license. We will present the results of the
+different steps in the incremental design in Section
+\ref{sec:experiments}, to finally wrap up with the conclusions.
 

### 2016-03-11 PPSN2016/paper.Rnw
Including aggregate count

@@
                                     PUTss = character(),
-                                     actualPUTs = character());
+                                    actualPUTs = character(),
+                                    count = character())
 
-experiment.aggregates.filtered <- data.frame(Experiment=character(),
-                                             milliseconds=character(),
-                                             IPs=character(),
-                                             actualIPs = character(),
-                                             PUTss = character(),
-                                             actualPUTs = character());
+experiment.aggregates.filtered = experiment.aggregates
 
@@
                                               PUTs=median(as.integer(this.data$PUTs)),
-                                              actualPUTs=median(as.integer(this.data$actualPUTs)))
+                                              actualPUTs=median(as.integer(this.data$actualPUTs)),
+                                              count= length(this.data$IPs))
                                    )
@@
                                                        PUTs=median(as.integer(this.data$PUTs)),
-                                                       actualPUTs=median(as.integer(this.data$actualPUTs))))
+                                                       actualPUTs=median(as.integer(this.data$actualPUTs)),
+                                                       count= length(this.data$IPs)))
 

### 2016-03-12 PPSN2016/paper.Rnw
Rewriting state of the art
Introducing new factors in human computing. Refs #108.

Please @unintendedbear @mariosky @fergunet take a look at the new approach to the state of the art.

@@
 the dual introduction of JavaScript as a universal language for the
-browser and the browser as an ubiquitous web and Internet client has
+browser and the browser itself as an ubiquitous web and Internet client has
 made this combination the most popular for volunteer computing
@@
 authors, including more recent efforts
-\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
-%Pablo: Desell in 2008 is not so recent
-used Java \cite{chong:1999:jDGPi} and others have gone away from the
-server-based paradigm to embrace peer to peer systems
-\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
-platforms avoid single points of failure (the server) but, since they
-need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower; this makes gathering new users
-immediate and spontaneous, so that they can be used for short
-algorithms on call.
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.
 
-On the other hand, we are also interested in measuring the performance
-of volunteer computing systems, an area in which there have been
-relatively few efforts.
-There were some initial attempts to avoid the differences in performance
-that could be obtained from volunteers  by making
-the algorithm adaptive to the kind of resources allotted to it
-\cite{milani2004online}, which is actually not such a big problem in
-algorithms such as the EA that can be easily 
-parallelized via population splitting or by farming out the evaluation to all
-the nodes available. Lately, several approaches have focused on the
-fault-tolerance of volunteer systems
-\cite{gonzalez2010characterizing} which it can, of course, be studied in
-the more general context of distributed computing 
-\cite{nogueras2015studying} or including it in a more general study of the
-performance of the EA itself
-\cite{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
-measured without first understanding the dynamics of this kind of systems. Initial
-work was done for peer to peer systems by Stutzbach et
-al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. A similar
-study was performed by Martínez et al. on the Capataz system
-\cite{martinez2015capataz}; however, in this case the number of
-computers used was known in advance and the main focus was on
-measuring the speed up and how job bundling helped to reduce overhead and
-enhance performance. On the contrary, in this paper, we will use {\em
-  actual} volunteers, anonymous persons that are requested to join an
-experiment through social networks.
+In general, those papers go no further than a proof of concept,
+attempting to gauge how many users join the effort and also how long
+would it take a concrete amount of users to reach the solution. In
+general, they do not try to assess the performance of the algorithm
+itself in the precise conditions they find in a volunteer computing
+system, although papers such as the one by Laredo et
+al. \cite{churn08:ijhpsa}, using models, try to find out the resulting
+performance when the users are changing continously according to a
+Weibull distribution. In the case of Klein and Spector \cite{unwitting-ec}, the algorithm
+is actually run on the server, using the browser mainly for fitness
+evaluation so no actual contribution to the evaluation of the
+evolutionary algorithm itself is made.
 
-Some of the essential metrics in volunteer computing like the
-number of users or the time spent by every one in the
-computation in browser-based volunteer computing experiments, have
-only been studied in a limited way in 
-\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
-run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining,agajaj} found out that the Weibull, log-normal and
-Gamma distribution 
-modeled quite well the availability of resources in several clusters
-of that framework; the shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas:
-there are many users that give a small amount of cycles, while there
-are just a few that give many cycles.
+There is another factor that should be taken into account: since the
+user has control of the browser, there is a limited amount of
+interaction with it, namely, the fact that by reloading the webpage he
+can apply a kind of hypermutation killing the current population and
+generating new individuals some of which will make their way to the
+common pool via migration. In that sense, volunteer computing is also
+a way of {\em human computation} \cite{quinn2011human}, a concept that
+has also been applied to evolutionary algorithms \cite{972056}, in
+this case extensively and with all operators. It is quite difficult to
+find out how many times this happens and what is their effect on the
+overall algorithm, but this only reaffirms the fact that there is a
+long list of issues with volunteer evolutionary computation, and that
+the volunteer him or herself is at the center of many of them.
 
-As far as we know, this paper presents one of the few experiments
-that measure the performance of a sociotechnical
-metacomputer, that is, a spontaneously created parallel computer that
-uses social networks for operations such as gathering new
-users. Apol{\'o}nia et al. \cite{apolonia2012enhancing} used the 
-Facebook protocol to distribute tasks among the {\em walls} of
-friends, explicitly using the social network for computing. However,
-it stopped short of relating performance to the macro measures of the
-users' social networks. As in the previous example, a social network
-was used to get new network nodes; in the previous case a web page was
-used, while Facebook's wall was used here.
-
-In our case, social networks are an integral part of the system and
-used to spontaneously obtain users. The algorithms used, as well as the methodology 
-for gathering resources will be described next, 
-together with the results obtained in this initial setup.
+In this paper we will make contributions towards the design of
+efficient volunteer evolutionary algorithms by studying the effect of
+several design decisions on performance. In the next Section we will
+present the general framework and the initial setup.
 

### 2016-03-12 PPSN2016/paper.Rnw
Adds table

@@
 
-\title{Evolutionary algorithms in a volunteer computing system:
-  performance assessment}
+\title{A performance assessment of evolutionary algorithms in a volunteer computing system}
 
@@
 system based on the browser is simply a client-server system
-whose client is, or it can be, embedded in the browser via
-JavaScript. Since JavaScript is  the only language that is present
-across all browsers, the choice was quite clear. We should emphasize
-that NodIO is more intended as an auxiliary computing engine, more
-than the main one, so performance of JavaScript as a language is not
-so important; even so, we have made a comparison between JavaScript and other languages
-\cite{2015arXiv151101088M} that shows that performance of JavaScript
-is comparable to other interpreted languages; compiled languages would
-be faster, but, of course, it is impossible to gather volunteers
-spontaneously and without any installation with them.
+with a client embedded in the browser via JavaScript. We call this
+system {\sf NodIO}. All parts of the framework are free and available
+with a free license from
+\url{https://github.com/JJ/splash-volunteer}.
 
-In this sense, in this paper we propose the {\sf NodIO} framework, a
-cloud or bare metal based volunteer evolutionary computing system
-derived from the {\sf NodEO} library, whose architecture has been
-developed using JavaScript on the client as well as the server.
-All parts of the framework are free and available with a free license
-from \url{https://github.com/JJ/splash-volunteer}.
-
-Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
-\item A REST server, that is, a server that includes several {\em
-  routes} 
-  that can be called for storing and retrieving information (the ''CRUD'' cycle:
-  create, request, update, and delete) from the server. 
-  A JSON data format is used for the communication between 
-  clients and the server. There are two kinds of information:
-  {\em problem} based, that is, related to the
-  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
-  GET}ing a random chromosome from it, and {\em information} related
-  to the performance and state of the experiments. It also performs logging
-  duties, but they are basically a very lightweight and high performance
-  data storage \cite{jj:idc:lowcost}.
-  The server has the capability to
-  run a single experiment, storing the chromosomes in a data structure
-  that is reset when the solution is found. This data structure can
-  hold every chromosome in a particular experiment, or have a finite
+The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
+\item A REST (representational state transfer) server that responds to
+  HTTP requests made by the browsers. These HTTP requests use JSON for
+  encoding information, in this case chromosomes and associates
+  information send by the clients and return requests made by the
+  server.  The server has the capability to
+  run a single experiment, storing the incoming chromosomes in a data structure
+  that is reset when the solution is found. This data has a finite
   size that erases the oldest chromosomes once it has filled to
-  capacity. In this paper we will test both implementations.  
+  capacity.  
 \item A client that includes the evolutionary algorithm as
-  JavaScript code embedded in a web page that displays graphs, some
+  JavaScript code embedded in a web page that displays the fitness and
+  other graphs, some
   additional links, and information on the experiment. This code runs
@@
   kept the number of generations between migrations fixed since it is
-  a way of finding out how many real work every client has done. 
+  a way of finding out how much real work every client is doing. 
 \end{enumerate}
@@
   Chart.js} have been used to build the user interface elements of the
-framework, a part of which is shown in Figure \ref{fig:screenshot} and
-should be running in \url{http://nodio-jmerelo.rhcloud.com}.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.95\linewidth]{nodio-screenshot.png}
-\caption{Screenshot showing the fitness and number of users as seen by
-volunteers. The one on the left shows fitness, on the right the
-accumulated number of users. \label{fig:screenshot}}
-\end{figure}
+framework, which should be running in
+\url{http://nodio-jmerelo.rhcloud.com}, a free resource hosted in the
+OpenShift Platform as a Service.\footnote{It is not guaranteed to be running,
+or running the same version, when you read this, however; you can always get
+the sources from GitHub and set it up yourself.}
 %
@@
 
-JavaScript is a functional language, so in order to work with a
-problem, a fitness function must be supplied at the creation of the algorithm object, 
-called {\tt Classic}. In this case the classical Trap 
-function \cite{Ackley1987} has been used. Depending on the
-problem additional handing functions and configurations 
-can be supplied.
+NodIO needs a fitness function to work with; in this case the classical Trap 
+function \cite{Ackley1987} has been used; this function runs in the
+browser, but the server needs to know when the solution has been
+found, so a termination condition is also included as a configuration
+option in the server. In all cases we are using 50 Traps, a problem
+that is difficult enough to need the intervention of several users for
+an extended amount of minutes.
 
-The next Section will describe experiments performed to
-establish a baseline performance and gather initial performance
-results. In the first set of experiments, performed last year, we used the 40-trap
-problem, while the current experiments changed to the more difficult 50-trap
-in order to compare the performance with a more computationally intensive problem.
+All experiments are announced in the same way: we use social networks,
+mainly Twitter, to announce the start of a new experiment. LinkedIn,
+Facebook, and also private groups in Telegram and WhatsApp have been
+used too for announcements. We also encourage friends and followers to
+put an original post for announcing it, instead of just sharing, in
+the way allowed by the platform, the original post. During the running
+of the experiments, users are engaged, questions answered, and all
+kind of explanations given if requested. In general, announcements
+have been made in Spanish, but in some cases English was used
+too. This makes for a certainly unrealiable experimentation framework,
+but it is a realistic one. The steps in the design of the evolutionary
+algorithm will be explained next.
 
 %---------------------------------------------------------------
-\section{Modeling the performance of a volunteer-based distributed computer} 
+\section{Experiments and results} 
 \label{sec:experiments}
@@
 \begin{center}
-\begin{tabular}{l|rrl}
-\hline
-Experiment & \#Runs & Different IPs & Traps \\
-\hline
-April 4th 4/4 & 57 & 191 & 40 \\
-April 24th 4/24 &  231 & 559& 40  \\
-July 31th 7/31 & 97 & 179 & 40 \\
-\hline
-February, cache=128 & 61 & 75 & 50  \\
-February, cache=64 & 61 & 220 & 50  \\
-February, cache=32 & 39 & 86 & 50  \\
-% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
-\hline
-\end{tabular}
+
+<<experiments, cache=FALSE,echo=FALSE>>=
+experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
+kable(experiment.aggregates)
+@ %def
+
 \end{center}

### 2016-03-12 PPSN2016/paper.Rnw
Reinserting the table

@@
 \begin{center}
-
-<<experiments, cache=FALSE,echo=FALSE>>=
-experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
-kable(experiment.aggregates)
-@ %def
+\begin{tabular}{|l|c|c|c|c|c|c|}
+\hline
+Experiment & Time (ms) & \multicolumn{2}{|c|}{IPs} & \multicolumn{2}{|c|}{PUTs} & \#Runs\\
+           &           & Total & Contributing & Total & Contributing & \\
+\hline
+base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81\\
+\hline
+graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
+graph2 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
+graph3 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
+\hline
+reboot & 513354.5 & 9 & 6 & 545.5 & 62.5 & 40\\
+\hline
+cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55\\
+\hline
+\end{tabular}
+%% Initially generated with this code, 
+%% <<experiments, cache=FALSE,echo=FALSE>>=
+%% experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
+%% kable(experiment.aggregates)
+%% @ %def
 
@@
 of these islands in the next experiment is indeed negligible. %Pablo: cite
-
-The table shows that every run included more than 30 experiments. The
-%Pablo: reference to table? (maybe the final print will misplace it)
-number of different IPs intervening in them varied from more than one
-hundred to more than five hundred in the second experiment, with a
-number around 50 in the second, and most recent, batch of experiments. 
-%
-\begin{table}[htb]
-\caption{Summary of time per run, number of IPs and number of {\tt
-    PUT}s per IP in the initial runs. \label{tab:summary:os}}
-\begin{center}
-\begin{tabular}{l|cccc}
-\hline
-     & \multicolumn{2}{c}{IPs} & \multicolumn{2}{c}{Median} \\
-Experiment & Median & Max & time (s) &  \#{\tt  PUT}s \\
-\hline
-4/4 & 5 & 16 & 2040 & 18   \\
-4/24 &  5 & 29 & 732 & 11  \\
-7/31 & 5 & 14 & 260 & 23   \\
-\hline
-Cache=128 & 5 & 17 & 222.2 & 124 \\ 
-Cache=64 & 8 & 38 & 51.3 & 100 \\
-Cache=32 & 6 & 19 & 58.9 & 45 \\
-\hline
-% Data in .RData file in Splash-volunteer
-\end{tabular}
-\end{center}
-\end{table}
-%
-
-A summary of the results of each run is also shown in Table
-\ref{tab:summary:os}, which shows the median number of IPs
-intervening in each experiment,  median time needed
-to finish the experiment, median number of HTTP {\tt PUT}s per IP. The
-first striking result is that in all cases, 50\% of the 
-experiments involved 5 or less IPs. This is consistent with previous results
-\cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
-the expected number of volunteer IPs. The
-maximum number of different IPs for each experiment is also in the
-same range and of the order of 10, which is also consistent with
-prior work and does not vary across the two different batches.
-
-We will have to analyze differently the median time, since the two
-%Pablo: different verbal times ("will have", and before, "is shown"). Homogenize.
-batches are solving different problems. In both cases it possesses a
-big range of variation, but 50\% of 
-the time takes less than several minutes, from around 4 minutes in the
-best case to roughly 2/3 of an hour in the worst case. Remarkably
-enough, the time is more consistent in the second batch and always
-around one minute, in two cases even less, and that happens when the
-median number of IPs is higher. It should be noted that while the
-first batch of experiments took several days in each case, the second
-only lasted for a few hours, with a more continued effort of
-publicizing it in social networks. This is specially true in the
-%Pablo: removed semicolon before "This". @unintendedbear does not like them :P
-case of cache equal to 64, which is noticed by the high number of
-volunteers participating in the experiment. The conclusion is that,
-in general, the key factor in the time needed to find the solution
-is, as expected, the number of volunteers it is able to gather on a
-short notice.
-
-The number of {\tt PUTs}, every one corresponding to 100 generations,
-is the algorithmic result. It is relatively unchanged for the first
-batch and around 20, that is, 2000 generations or 2000*128 = 256000
-evaluations. In this case, an ``unlimited'' cache was used, with all
-individuals sent from clients stored until the end of the
-experiment. However, we were interested in measuring also the
-performance of the algorithm itself by changing the cache size, after
-making it limited. As it can be seen in the table, there is a clear
-%Pablo: which table?
-change in the number of evaluations needed, with smaller cache sizes
-producing solutions in less evaluations, until it is for the cache
-size = 32 roughly twice as much as with the previous problem, with 40
-traps. This is a good result and is also algorithmically consistent
-with other results obtained using the same type of problems. Since in
-this paper we were interested in leveraging the user's CPU cycles by
-improving the algorithm, a good conclusion of this paper is that
-having a small pool size helps clients to obtain ``good'' individuals
-from the pool, as opposed to any individual that could be obtained
-before. Besides, cache policy deletes the oldest individuals, which
-makes those in the pool be {\em current}, helping then newcomers and %Pablo: "then helping newcomers", instead?
-any participant obtain the best individuals found in the last part of
-the experiment. Besides, a limited cache helps also in cases with a
-bigger search space or longer running times when the server simply
-crashed due to lack of RAM.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-128.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-64.png}
-\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-32.png}
-\caption{Duration of experiments vs. number of different IPs (nodes)
-  participating in it, with averages and standard deviation shown as
-  red dots; in the case there is a single red dot, there was a single
-  experiment in which many computers participated (for instance, 16
-  computers in the experiment in the far left or 29 in the middle
-  one). 
-Shades of blue indicate how many experiments included that many unique IPs,
-so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequency. 
-From left to right and top to bottom experiments 4/4, 4/24 and 7/31,
-followed by experiments with 50 traps, cache=128, 64, 32.}
-\label{fig:duration}
-\end{figure}
-%
-We will have to analyze experimental data a bit further to find out why
-this happens and also if there are some patterns in the three sets of
-experiments. An interesting question to ask, for instance, is if
-by adding more computers makes the experiment take less. In fact, as
-shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
-not seem to contribute to decreasing the time needed to finish the
-experiment. However, the cause-effect relationship is not clear at
-all. It might be the opposite: since experiments take longer to finish
-and might in fact be abandoned with no one contributing for some time,
-the probability of someone new joining them is higher. In fact,
-with experiments taking a few seconds and due to the way the
-experiments are announced, it is quite difficult that several
-volunteers join in in such a short period of time, even more if we take
-into account that volunteers are not {\em carried over} from previous
-experiments. 
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.95\linewidth]{ips-per-minute-cache=64.png}
-\caption{Simultaneous IPs every minute of the experiment with cache =
-  64. \label{fig:otisdriftwood}}
-\end{figure}
-
-That is why we used a more difficult problem in the second batch of
-experiments, which is shown in the bottom row of Figure
-\ref{fig:duration}. The pattern is remarkably similar, showing a
-positive correlation between the time for solving the problem and the
-number of computers, at least for cache sizes 128 and 64. However, it
-is interesting to observe that, for cache=32, the time needed to find
-the solution decreases from one to approximately 4-5 nodes, to then
-increase for a higher number of participating computers, distinguished
-by IP. The green dot at the bottom is probably an outlier that we
-will try to explain later on. This leads us to conclude that a better
-amount of computers might contribute to speed up the solution, if the
-time the experiment ideally takes is sufficient, that is, of the order of a
-minute, and enough volunteers concur simultaneously. This is also
-observed, not so clearly, in the case of cache=64, with an interval of
-around 10 IPs obtaining less time than experiments with less or more
-IPs, and of the same order, between 10 and 100 seconds. If we look at
-the graph that shows the number of IPs or volunteers per minute for
-this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
-there are peaks of more than 25 volunteers, and a period of several
-hours with a minimum of 6 computers and peaks of more than 10. The
-long period after midnight where there is a single volunteer left
-masks the success achieved during this set of experiments, from which
-we draw two lessons: first, you need a social network influencer to
-%Pablo: people know the meaning of "influencer"? Probably yes, but I mention it just in case.
-announce your experiments and second, no matter what, do not do any
-experiment after midnight. This statement, which might seem tongue in
-cheek, in fact, it is a conclusion drawn from the experimental data and
-to what extent the social network is an essential part of the
-description and performance of the NodIO volunteer computing system.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-128.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-64.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-alife-32.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
-  logarithmic scale. Dot color is related to the number of IPs
-  participating in the experiment. From left to right and top to bottom, experiments
-  4/4, 4/24 and 7/31 and caches=128,64, 32} 
-\label{fig:zipf:os}
-% First ones made with time-vs-IPs-openshift.R:  
-\end{figure}
-%
-It is also interesting to check the distribution of the experiment
-duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
-a Zipf's law, with similar distribution along all three runs. The 4/24
-run is the most complete and shows an S-shape, which implies an
-accumulation of experiments taking similar time and around 100
-seconds; this S-shape appears too in the experiments with cache=128
-(bottom row, left). The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, on the order of
-10 seconds, and which appears in all three graphs. As it can be seen,
-it sharply drops implying there are 
-just a few of them, and with diminishing probability as time
-decreases. However, since they have a greenish color, implying a low
-number of IPs, they might be due to clients {\em carrying over} from
-the previous one. This is a characteristic of this implementation
-which will be examined later on, but at any rate, if we discard those
-experiments that take too much or too little, there is a decreasing
-exponential distribution that corresponds to the Zipf's law.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
-\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
-\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
-\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
-\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
-\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
-\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
-  distribution (in red). From left to right and top to bottom, experiments
-  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
-% Plotted with ../data/plot-zipf-openshift.R
-\label{fig:puts:os}
-\end{figure}
-%
-A similar exponential distribution also appears if we rank HTTP {\tt
-  PUT}s, equivalents to the number of 
-generations divided by 100, or to evaluations divided by 12800,
-contributed by every user, which is shown 
-in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
-that is, a power law with a small {\em bump} in the lowest
-values. After testing the Generalized Extreme Value distribution and
-failing for the new batch of experiments, we have fitted it to a
-Weibull distribution \cite{thoman1969inferences} with the resulting
-parameters shown in Table \ref{tab:puts:os}. 
-%
-\begin{table}
-\caption{Weibull distribution  parameters of the fit of
-  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
-\begin{center}
-\begin{tabular}{l|cc}
-\hline
-Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
-\hline
-4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
-4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
-7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
-\hline
-Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
-Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
-Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
-\hline
-% from plot-zipf-openshift.R and ips-puts-alife.R
-\end{tabular}
-\end{center}
-\end{table}
-%
-The inverse
-Weibull distribution is a special case of the GEV distribution we have
-used in those papers, and appears usually in natural sciences and
-artificial life, usually related to decay. It has been frequently
-fitted to volunteer computing frameworks such as SETI@home
-\cite{javadi2009mining}. The model that user behavior follows can be
-explained straighforwardly: when users visit the page, it draws their
-attention for a limited amount of time. They give it a chance for a
-few seconds. If something there amuses them or they can engage in a
-conversation about it, they stay for a while longer, otherwise, they
-leave. The {\em scale} parameter, which is around 20-40 in the first
-batch for the 40 Trap problem and between 160 and 200 in the 60 traps
-problem, depends mainly on the maximum number of generations people
-leave it running. Since it finishes or stalls after a number of
-generations shown in Figure \cite{tab:summary:os}, volunteers just
-leave after that. Curiously enough, the scale parameter is roughly
-twice the median number of {\tt PUT}s per experiment, showing that, on
-average, the most loyal users reload the page twice after finishing or
-after seeing the evolution does not progress. This rule of thumb
-breaks down with the last experiment, however, which is interesting by
-itself, too.
-
-The slope or shape parameter, on the other hand, indicates the overall
-shape of the curve. A value less than 1 indicates a concave (in a
-non-algorithmic scale) curve,
-with figures closer to one indicating a smaller slope. In all cases
-values are between 0.54 and 0.77, independently of the experiment. It
-might be the case that this number depends more on the total number of
-experiments carried out, with sets with more experiments, both in the
-middle, having values between 0.60 and 0.70. The distribution is
-remarkably similar which gives us a model of user behavior that is, to
-a certain extent, independent of the experiment.
-
-These experiments show that, as it was proved for other volunteer
-computer frameworks and also in the case of games, user engagement
-follows a Weibull distribution. This makes engagement the key for
-leveraging the performance of the sociotechnical metacomputer and a
-way to improve results in the future.
 

### 2016-03-12 PPSN2016/paper.Rnw
Solving data problems
One of the files was duplicated, it's been rearranged now. Table also fixed.

@@
 
-these.experiments <- c("nooverlap-1","nooverlap-newgraph", "nooverlap-newgraph-2","nooverlap-newgraph-3","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
+these.experiments <- c("nooverlap-1","nooverlap-newgraph-1", "nooverlap-newgraph-2","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
 
@@
 graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
-graph2 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
-graph3 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
+graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
 \hline

### 2016-03-12 PPSN2016/paper.Rnw
Updating the description o the experiments
Which now together with Intro and Abstract take exactly 4 pages as intended. Please @unintendedbear @mariosky @fergunet take a good look at it to see if it's clear enough.

@@
   kept the number of generations between migrations fixed since it is
-  a way of finding out how much real work every client is doing. 
+  a way of finding out how much real work every client is doing. In
+  the version of the server used in this paper, we implemented a
+  policy by which the combination IP-fitness value could only be
+  inserted once into the cache. That means that if a particular IP
+  sends the fitness 99 into the server, it will be accepted only
+  once. The main intention of this policy was to avoid {\em
+    overlapping} of contributions, with clients still running an {\em
+    old} run of the experiment contributing after a new one had
+  started. This did not completely avoid these contributions: if the
+  clients of the {\em old} run increased their new fitness, it would
+  still be allowed. 
 \end{enumerate}
@@
 
+For every new individual sent by the clients to the cache, several
+pieces of meta data were stored: a time stamp, the client IP, the chromosome and
+value, the cache size in that particular moment, and also if that
+individual actually updated the cache or not.
+
 All experiments are announced in the same way: we use social networks,
@@
 \hline
-Experiment & Time (ms) & \multicolumn{2}{|c|}{IPs} & \multicolumn{2}{|c|}{PUTs} & \#Runs\\
+Experiment & Median Time (ms) & \multicolumn{2}{|c|}{Median IPs} & \multicolumn{2}{|c|}{Median PUTs} & \#Runs\\
            &           & Total & Contributing & Total & Contributing & \\
@@
 %
-Initial experiments were set up using the OpenShift
+Since the point of using volunteer computing system is to cut costs,
+all experiments were set up using the OpenShift
 PaaS, which provides a free tier, making the whole experiment cost
-equal to \$0.00. Experiments were
-announced through a series of posts on Twitter and, in the latest case, Telegram, and
-results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
-purpose of this paper, we repeated the announcement several times
-through the month of February. All  %% April --> January ??
-in all, we have the set of runs with the characteristics shown in
-Table \ref{tab:runs}. In general, every experiment took several
-days. No particular care was taken about the time of the announcement
-%Pablo: but it would great at least describe how many times was announced and how long
-or the particular wording. Every {\em experiment} consisted in running
-until the solution of the 50-trap problem was found. When the correct
-solution was sent to the server, the counter was updated and the pool
-of solutions resets to the void set. There was no special intention to wait
-until all clients had finished, thus it might happen that. In fact,
-the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible. %Pablo: cite
+equal to \$0.00. In fact, the NodIO system can be deployed to any
+system as long as it can run node.js and access to the filesystem, for
+log files, is available. All these logs are available under
+request. Processed files are available in the same repository that
+hosts this paper, as well as the scripts needed to process them.
+
+Every run had slightly different conditions, although in some cases it
+was just a server reboot and a new round of announcements. Each run
+batch included at least 30 competions, that is, running until the
+solution was found. The number of runs was controlled by polling an
+URL that indicated the number of runs made in the present batch so
+far. A summary of the experiments and  its results is shown in Table
+\ref{tab:runs}. This table shows the median time needed to find the
+solution as well as the median number of IPs and PUTs.
 

### 2016-03-12 PPSN2016/paper.Rnw
Explaining the table

@@
 \ref{tab:runs}. This table shows the median time needed to find the
-solution as well as the median number of IPs and PUTs.
+solution as well as the median number of IPs and PUTs. In this case we
+distinguish between the total number of clients participating in that
+particular run, and the {\em actual} number whose contributions were
+accepted into the cache, which was set up to accept those only if {\em
+  new} chromosomes were sent to the server in order to avoid overlaps
+between one run and the next. That is why, in both cases, the
+``Contributing'' column includes a value that is less than the total
+number. The final column includes the total number of runs in that
+particular batch, finished after a minimum number of 30 was reached.
 

### 2016-03-12 PPSN2016/paper.Rnw
Intro looks good

@@
 In a volunteer distributed computing system, users run a program in
-their own machine to contribute to a common effort. If they are
+their own machine to contribute to a common effort. If the program is
 embedded in a web page, collaboration is easy, but also ephemeral. In
@@
 
-The world wide web provides not only a platform for content
-distribution, but an increasingly reliable and high-performance
+The world wide web not only provides a platform for content
+distribution, but also an increasingly reliable and high-performance
 platform for running distributed applications. This is based in two
@@
 a JavaScript application interchange information with a server, called
-rest. From the point of view of the programmer, this involves
+REST. From the point of view of the programmer, this involves
 relatively common skills and no special libraries, since that

### 2016-03-13 PPSN2016/paper.Rnw
Comment
cache

@@
   server.  The server has the capability to
-  run a single experiment, storing the incoming chromosomes in a data structure
+  run a single experiment, storing the incoming chromosomes in a cache data structure
   that is reset when the solution is found. This data has a finite
@@
   still be allowed. 
+  % As I remember we kept the run_number so IPs with old runs
+  % did not interfere with newer ones. But you are saying: what if they are super slow 
+  % and always PUT when the run is already over?, we want to receive 
+  % those individuals? 
+  % "That means that if a particular IP
+  % sends the fitness 99 into the server, it will be accepted only
+  % once."  
+  % once per run?
+  %  but different individuals could have the same fitness, are we limiting the search? 
+  %  maybe we could have multiple versions of the pool a la MVCC  :)
+  % I think if we are just trying to limit overlapping of the experiments
+  % then PUTs should send the run's seq_number and completly remove it.
+
 \end{enumerate}
@@
 Figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or {\tt
+algorithm behavior. Different JavaScript libraries, such as JQuery or {\tt
   Chart.js} have been used to build the user interface elements of the
@@
 For every new individual sent by the clients to the cache, several
+% We have not presented the "cache",  only the server and
+% ..storing the incoming chromosomes in a data structure
+% Im changing over there to cache
 pieces of meta data were stored: a time stamp, the client IP, the chromosome and

### 2016-03-13 PPSN2016/paper.Rnw
Reprocessed files
Added a new column with "reboots" and nothing seems to be broken, so...
Answering @mariosky suggestions
Adding data per IP to the paper script
It looks like there is a certain difference, with nnumber of reboots going up depending on the presentation. I will have to do a Wilcoxon difference or something

@@
                                     actualIPs = character(),
-                                    PUTss = character(),
+                                    PUTs = character(),
                                     actualPUTs = character(),
@@
                                   IPs=as.integer(this.data$IPs))
+                       )
+}
+
+ips.data <- data.frame(Experiment=character(),                      
+                       PUTs=character(),
+                       reboots=character())
+
+ips.aggregates <- data.frame(Experiment=character(),
+                             PUTs = character(),
+                             reboots = character(),
+                             avgPUTs = character(),
+                             avgreboots = character() )
+
+for ( i in these.experiments ) {
+    this.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
+    ips.data <- rbind(ips.data,
+                      data.frame(Experiment=rep(i,length(this.data$PUTs)),
+                                 PUTs=as.integer(this.data$PUTs),
+                                 reboots=as.integer(this.data$reboots))
+                       )
+
+    ips.aggregates <- rbind( ips.aggregates,
+                            data.frame(Experiment=i,
+                                       PUTs = median(this.data$PUTs),
+                                       reboots = median(this.data$reboots),
+                                       avgPUTs = mean(this.data$PUTs),
+                                       avgreboots = mean(this.data$reboots) )
                             )
@@
   % once per run?
+% JJ - once per run or experiment. But we didn't keep that run_number
+% to avoid interference, just to keep track
   %  but different individuals could have the same fitness, are we limiting the search? 
+% We are actually checking for the same individual, not only the fitness
   %  maybe we could have multiple versions of the pool a la MVCC  :)
+% Let's leave that for the future...- JJ
   % I think if we are just trying to limit overlapping of the experiments
   % then PUTs should send the run's seq_number and completly remove it.  
-
+% The client does not keep track of the seq number. Maybe we should do
+% just that, but sending back and forth the seq number implies
+% refactoring. - JJ
 \end{enumerate}
@@
 particular batch, finished after a minimum number of 30 was reached.
+
+The baseline, called {\em base} in the
 

### 2016-03-13 PPSN2016/paper.Rnw
Reshaping data for graphs

@@
                        PUTs=character(),
-                       reboots=character())
+                       reboots=character(),
+                       idu=character(),
+                       idx=character())
 
@@
                              avgPUTs = character(),
-                             avgreboots = character() )
+                             avgreboots = character(),
+                             totalIPs = character())
 
@@
     this.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
+    this.len <- length(this.data$PUTs)
+    seq <- 1:this.len
     ips.data <- rbind(ips.data,
                       data.frame(Experiment=rep(i,length(this.data$PUTs)),
+                                 idu=seq,
+                                 idx=seq/this.len,
                                  PUTs=as.integer(this.data$PUTs),
@@
                                        avgPUTs = mean(this.data$PUTs),
-                                       avgreboots = mean(this.data$reboots) )
+                                       avgreboots = mean(this.data$reboots),
+                                       totalIPs =this.len )
                             )

### 2016-03-13 PPSN2016/paper.Rnw
Explaining the different experiments

@@
                                     actualPUTs = character(),
-                                    count = character())
+                                    count = character(),
+                                    totalIPs = character())
 
@@
 
+ips.data <- data.frame(Experiment=character(),                      
+                       PUTs=character(),
+                       reboots=character(),
+                       idu=character(),
+                       idx=character())
+
+ips.aggregates <- data.frame(Experiment=character(),
+                             PUTs = character(),
+                             reboots = character(),
+                             avgPUTs = character(),
+                             avgreboots = character(),
+                             totalIPs = character())
+
 for ( i in these.experiments ) {
     this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
+
+    this.ip.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
+    this.len <- length(this.ip.data$PUTs)
+    seq <- 1:this.len
+    ips.data <- rbind(ips.data,
+                      data.frame(Experiment=rep(i,length(this.ip.data$PUTs)),
+                                 idu=seq,
+                                 idx=seq/this.len,
+                                 PUTs=as.integer(this.ip.data$PUTs),
+                                 reboots=as.integer(this.ip.data$reboots))
+                       )
+
+    ips.aggregates <- rbind( ips.aggregates,
+                            data.frame(Experiment=i,
+                                       PUTs = median(this.ip.data$PUTs),
+                                       reboots = median(this.ip.data$reboots),
+                                       avgPUTs = mean(this.ip.data$PUTs),
+                                       avgreboots = mean(this.ip.data$reboots),
+                                       totalIPs =this.len )
+                            )
 
@@
                                               actualPUTs=median(as.integer(this.data$actualPUTs)),
-                                              count= length(this.data$IPs))
+                                              count= length(this.data$IPs),
+                                              totalIPs=this.len)
                                    )
@@
 
-ips.data <- data.frame(Experiment=character(),                      
-                       PUTs=character(),
-                       reboots=character(),
-                       idu=character(),
-                       idx=character())
-
-ips.aggregates <- data.frame(Experiment=character(),
-                             PUTs = character(),
-                             reboots = character(),
-                             avgPUTs = character(),
-                             avgreboots = character(),
-                             totalIPs = character())
-
-for ( i in these.experiments ) {
-    this.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
-    this.len <- length(this.data$PUTs)
-    seq <- 1:this.len
-    ips.data <- rbind(ips.data,
-                      data.frame(Experiment=rep(i,length(this.data$PUTs)),
-                                 idu=seq,
-                                 idx=seq/this.len,
-                                 PUTs=as.integer(this.data$PUTs),
-                                 reboots=as.integer(this.data$reboots))
-                       )
-
-    ips.aggregates <- rbind( ips.aggregates,
-                            data.frame(Experiment=i,
-                                       PUTs = median(this.data$PUTs),
-                                       reboots = median(this.data$reboots),
-                                       avgPUTs = mean(this.data$PUTs),
-                                       avgreboots = mean(this.data$reboots),
-                                       totalIPs =this.len )
-                            )
-}
 @
@@
 \begin{center}
-\begin{tabular}{|l|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-Experiment & Median Time (ms) & \multicolumn{2}{|c|}{Median IPs} & \multicolumn{2}{|c|}{Median PUTs} & \#Runs\\
-           &           & Total & Contributing & Total & Contributing & \\
+Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &
+\multicolumn{2}{|c|}{Median PUTs} & \#Runs & Different IPs\\
+           &   (ms)        & Total & Used & Total & Used &
+           &\\
 \hline
-base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81\\
+base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81 & 262\\
 \hline
-graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
-graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
+graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36 & 96\\
+graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36 & 73\\
 \hline
-reboot & 513354.5 & 9 & 6 & 545.5 & 62.5 & 40\\
+reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 45 & 168\\
 \hline
-cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55\\
+cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55 & 225\\
 \hline
@@
 %% <<experiments, cache=FALSE,echo=FALSE>>=
-%% experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
+%% experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
 %% kable(experiment.aggregates)
@@
 %
+\begin{figure}[htbp]
+\centering
+
+<<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
+ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))
+@ %def graph for IPs and puts
+
+\caption{Ranked and normalized number of contributions (PUTs) per IP,
+  plotted in a log $y$ scale.}
+\label{fig:zipf}
+\end{figure}
+
 Since the point of using volunteer computing system is to cut costs,
@@
 log files, is available. All these logs are available under
-request. Processed files are available in the same repository that
+rrequest. Processed files are available in the same repository that
 hosts this paper, as well as the scripts needed to process them.
@@
 
-The baseline, called {\em base} in the
+The baseline, called {\em base} in the Table, uses simply the
+mechanism of dropping contributions from clients if they are repeated
+to avoid overlap. Clients still get a random chromosome from the
+server, so they can in fact proceed with the algorithm and even finish
+it. The fact that these users were not contributing to the pool was
+conveyed by inserting a graph in the client that represented the size
+of the cache and labeled ``How am I contributing''. The user could
+then realize if his contribution was being 0 and thus do something
+about it; these are labeled {\tt graph1} and 2. However, a stuck
+client could still add to the pool if evolution got unstuck, so we
+added a mechanism for rebooting the client, that is, reloading the
+page, if the cache size was found to be less than 1, which indicated a
+run recently initiated. This batch of runs is labelled {\tt
+  reboot}. However, we found that this size of the cache was actually
+skipped by some slow clients or too fast evolutions, so that in the
+last case, labeled {\tt cachecrash}, clients rebooted if the cache
+collapsed by more than a certain length, indicating a cache that had
+been voided and was filling again.
 

### 2016-03-13 PPSN2016/paper.Rnw
Adding objectives
Changes data processing

@@
                               PUTs=character(),
-                              actualPUTs=character())
+                              actualPUTs=character(),
+                              PUTsIP =character())
 
@@
                                     count = character(),
-                                    totalIPs = character())
+                                    totalIPs = character(),
+                                    PUTsIP = character())
 
@@
                                               count= length(this.data$IPs),
-                                              totalIPs=this.len)
+                                              totalIPs=this.len,
+                                              PUTsIP=median(as.integer(this.data$PUTs)/as.integer(this.data$IPs)))
                                    )
@@
                                         PUTs=as.integer(this.data$PUTs),
-                                        actualPUTs=as.integer(this.data$actualPUTs)))
+                                        actualPUTs=as.integer(this.data$actualPUTs),
+                                        PUTsIP=as.integer(this.data$PUTs)/as.integer(this.data$IPs)))
 
@@
                                                        actualPUTs=median(as.integer(this.data$actualPUTs)),
-                                                       count= length(this.data$IPs)))
+                                                       count= length(this.data$IPs),
+                                                       PUTsIP=median(as.integer(this.data$PUTs)/as.integer(this.data$IPs))
+                                                       )
+                                            )
 
@@
 collapsed by more than a certain length, indicating a cache that had
-been voided and was filling again.
+been voided and was filling again. Our objective was twofold: to
+encourage engagement by the users and also to increase the involvement
+of every client in the common experiment, eliminating at the same time
+overlaps between runs.
 

### 2016-03-13 PPSN2016/paper.Rnw
Adds PUTratio

@@
                               actualPUTs=character(),
+                              PUTratio = character(),
                               PUTsIP =character())
@@
                                               actualPUTs=median(as.integer(this.data$actualPUTs)),
+                                              PUTratio=median(as.integer(this.data$actualPUTs)/as.integer(this.data$PUTs)),
                                               count= length(this.data$IPs),
@@
 
-    this.data <- this.data[this.data$actualPUTs>=10,]
+    this.data <- this.data[this.data$actualPUTs>=16,]
     experiment.aggregates.filtered <- rbind( experiment.aggregates.filtered,

### 2016-03-13 PPSN2016/paper.Rnw
Remaking table
Fixes graph and table

@@
 \begin{center}
-\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|}
 \hline
 Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &
-\multicolumn{2}{|c|}{Median PUTs} & \#Runs & Different IPs\\
-           &   (ms)        & Total & Used & Total & Used &
-           &\\
+\multicolumn{3}{|c|}{Median PUTs} & \#Runs & Unique IPs & Puts/IP \\
+           &   (ms)        & Total & Used & Total & Used & Ratio &  & & \\
 \hline
-base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81 & 262\\
+base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 0.1290323 & 81 & 262 & 21.33333\\
 \hline
-graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36 & 96\\
-graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36 & 73\\
+graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 0.1282238 & 36 & 96 & 24.30000\\
+
+graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 0.0520509 & 36 & 73 & 129.04545\\
 \hline
-reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 45 & 168\\
+reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 0.0955121 & 45 & 168 & 75.88889\\
 \hline
-cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55 & 225\\
+cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 0.1575092 & 55 & 225 & 43.28571\\
 \hline
 \end{tabular}
+%% \end{tabular}
 %% Initially generated with this code, 
-%% <<experiments, cache=FALSE,echo=FALSE>>=
-%% experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
-%% kable(experiment.aggregates)
-%% @ %def
+% <<experiments, cache=FALSE,echo=FALSE>>=
+%experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
+%kable(experiment.aggregates)
+%@ %def
 
@@
 <<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
-ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))
+ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
 @ %def graph for IPs and puts

### 2016-03-13 PPSN2016/paper.Rnw
Inserts new table

@@
 %@ %def 
+\end{center}
+\end{table}
+%
+%
+\begin{table}
+  \caption{Experiments with overlaps eliminated:  summary of results. \label{tab:filtered}}
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &\multicolumn{2}{|c|}{Median PUTs} & \#Runs &  Puts/IP \\
+&   (ms)        & Total & Used & Total & Used &  & \\
+\hline
+base & 293685 & 13.0 & 8.0 & 483.0 & 57.0 & 47 & 35.50000\\
+\hline
+graph1 & 768053 & 12.0 & 5.0 & 601.0 & 79.0 & 21 & 45.84615\\
+graph2 & 1477969 & 6.0 & 4.0 & 901.0 & 52.0 & 25 & 149.42857\\
+\hline
+reboot & 799471 & 9.5 & 6.5 & 656.5 & 67.0 & 44 & 76.48016\\
+\hline
+cachereboot & 463347 & 12.0 & 12.0 & 638.5 & 80.5 & 52 & 44.25911\\
+\hline
+\end{tabular}
+%% Initially generated with this code, 
+%%<<experiments.filtered, cache=FALSE,echo=FALSE>>=
+%%experiment.aggregates.filtered$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
+%%kable(experiment.aggregates.filtered)
+%%@ %def
 
@@
 \end{table}
+%
 %
@@
 overlaps between runs.
+
+So we will have to check whether we have effectively eliminated these
+overlaps. In order to do that, we will have to establish a threshold
+under which reasonably the solution cannot be found. In every
+evolutionary algorithm it might happen, by chance, that the solution
+is found in just a few evaluations, but that is usually not the
+case. By looking at the logs in the last case we have established that
+minimun as 16 PUTs; in less than 1600 generations no solution to that
+difficult problem will be found. Let us filter then the runs,
+eliminating those that use less than 1600 generations to find the
+solution which can reasonably be said that they did so because some
+browser ``carried over'' a chromosome from the previous run. The
+results are shown in Table \ref{tab:filtered}
 

### 2016-03-13 PPSN2016/paper.Rnw
Including and commenting results.
Please @mariosky @fergunet and @unintendedbear take a look at what has been written and comment and so on.

@@
 \caption{Ranked and normalized number of contributions (PUTs) per IP,
-  plotted in a log $y$ scale.}
+  plotted in a log $y$ scale.\label{fig:zipf}}
 \label{fig:zipf}
@@
 was just a server reboot and a new round of announcements. Each run
-batch included at least 30 competions, that is, running until the
+batch included at least 30 completions, that is, evolution until the
 solution was found. The number of runs was controlled by polling an
 URL that indicated the number of runs made in the present batch so
-far. A summary of the experiments and  its results is shown in Table
+far. Every volunteer contributed a certain number of individuals
+(represented by the PUT HTTP order) to the server. The ranked number
+of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
+log-log scale. Besides following the usual Zipf's law found in our
+previous work and having a very similar appearance for all 5 sets, it
+is interesting to see how some sets have a number of contributions per
+IP that is slightly superior to the rest, specially so in {\tt
+  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
+show a regime in the middle of the graph that shows that some users
+are contributing more than in other similar sets. In general, these
+differences in behavior are due more to the specific users that show
+up in these experiments than to general laws, but it should also be
+noted that those behaviors can be encouraged by design.
+
+A summary of the experiments and  its results is shown in Table
 \ref{tab:runs}. This table shows the median time needed to find the
@@
 browser ``carried over'' a chromosome from the previous run. The
-results are shown in Table \ref{tab:filtered}
+results are shown in Table \ref{tab:filtered}. The first change is in
+the column \#Runs, that shows that most of the runs in the initial
+configurations found the solution thanks to these carried over
+clients. However, in the last two rows, the ones that rebooted clients
+when they detected changes in cache, the difference is minimal. This
+means that one objective has been reached: if clients reboot when they
+detect changes in the cache that imply a new run, overlaps are almost
+eliminated.
+
+However, the rest of the measures also changes and so we will refer to
+this table from now on.
 

### 2016-03-13 PPSN2016/paper.Rnw
Commenting on experiments

@@
 %%@ %def 
-
+%
 \end{center}
@@
 \centering
-
 <<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
@@
 @ %def graph for IPs and puts
-
+%
 \caption{Ranked and normalized number of contributions (PUTs) per IP,
@@
 detect changes in the cache that imply a new run, overlaps are almost
-eliminated.
+eliminated. Going back to Table \ref{tab:runs}, we can see than the
+ratio of effective contributions, those that feed the cache, increases
+from around 13\% in the first case to almost 16\% in the last,
+reaching lows of less than 6\% in some cases ({\tt graph2}). This
+indicates that rebooting the clients has the positive  effect of
+increasing the contribution of clients to the common pool. However, we
+should remark that still the clients will stop contributing if they
+get stuck in a particular solution, so the ratio keeps being low
+resulting in a slow change in the cache and a lower diversity of
+clients that tap that cache. At any rate, an increased number of
+effective PUTs will have a positive impact in the algorithm.
 
-However, the rest of the measures also changes and so we will refer to
-this table from now on.
+We can check this impact by looking at the column that refers to
+median time in Table \ref{tab:filtered} and its relation to the median
+number of PUTs, which reflects the number of evaluations performed by
+the clients. The two last rows have a low number of evaluations,
+around 650, which is higher than the first two, probably due to the
+lower volunteer supply, but better than the middle one; besides, the
+amount is remarkably similar. But in the last case, it takes much less
+time to find the solution {\em using the same number of
+  evaluations}. All variables are difficult to control in this
+environment, but it might be due to the fact that parallelism has
+increased, there are more volunteers contributing at the same time. In
+fact, as can be seen in the last column, the median number of
+contributions per user is {\em much lower} in the last row, which
+indicates that {\tt cachereboot} needs less contributions by users to
+find the same solution. Besides, the last row shows that {\em all}
+users actively contribute to the cache, while in the {\tt reboot}
+strategy 30\% of users do not. This might be due to the fact that {\tt
+  reboot} only does so when cache size is less than one. With very
+fast clients there might be some clients that do not detect that
+change and might, in fact continue with their former run, thus not
+contributing to the common cache.
 
@@
 
-Our intention in this paper was to assess the capabilities of a
-sociotechnical system formed by a client-server web-based framework
-running a distributed evolutionary algorithms and the volunteers that
-participate in the experiment. These volunteers are {\em in the
-  cloud}, that is, available as {\em CPU as a service} for the persons
-running the experiment. In this paper we have tried to put some figures
-on the real size of that {\em cloud} and how it can be used standalone
-if there is no alternative, or, if other computing resources are
-available, in conjunction with other local or cloud-based methods to
-add computing power in a seamless way through the pool that NodIO creates.
-
-After running the experiment on the 40-trap problem whose running time
-could be as low as a few seconds, we switched to another batch of
-experiments where we used the 50-trap problem and also to a pool of
-limited, and dwindling through the three experiments, size. Since our
-initial results indicated that what happened on the screen, a flat
-graph with no improvements or the experiment finished, influenced the
-amount of time that the users devoted to the experiment, a longer one
-could yield different results and, at the same time, result in a big
-pool that might either crash the server or return useless individuals
-to the volunteers. These new experiments have proved that using the
-limited pool is beneficial to finding the solution, since less
-evaluations are needed, but also that since the problem is more
-difficult, the users stay for longer in the web page, making less
-ephemeral the sociotechnical computing system created by the
-simulation.
-
-The second objective of this paper was to model the user behavior in a
-first attempt to try and predict performance. As should be expected,
-the model depends on the implementation, with contributions following
-a Weibull distribution, which reflects the fact
-that volunteer computing follows a model quite similar to that found
-for games or other online activities. The reverse might be true: if we
-want to have returning users for the experiments, it is probable that
-we should {\em gamify} the experience so that once they've done it
-once, they might do it more times. In the spirit of Open Science, this
-gamification might involve computing in real time data such as the one
-presented in this paper and showing it in the same page or presenting
-user results alongside others.
-
-In general, linking and finding correlations between user choices and
-performance is an interesting avenue to explore in the future. Even if
-these  experiments were published in a similar way, one
-obtained up to 5 times more total cycles than the one with the least
-number of cycles. It is also essential to obtain volunteers as fast and
-simultaneously as possible, so it is possible that the features of the
-social network in terms of real-time use will also play a big
-role; synchronous webs such as Snapchat, which mystifies the writers
-of this paper, and Twitter, thanks to its real time nature, might be
-better suited than Facebook, LinkedIn or Google plus. Even as it is
-difficult to create controlled experiments in this 
-area, it is an interesting challenge to explore in the future.
-
-The other area to explore is the algorithmic area itself. Are there
-ways to change the evolutionary algorithm, or its visualization, so
-that the user has a bigger impact on the result? One of the users in
-Twitter even suggested to embed videos so that people spent time
-looking at them, but other possible way was to make the user engage
-the algorithm by giving him or her buttons to change the mutation rate
-when the algorithm is stalled, for instance. If this is combined with
-a score board where local performance is compared to other users,
-engagement might be increased and thus the performance of the
-system. In general there are 
-many issues with the evolutionary algorithm implementation itself,
-including using different, or adaptive, policies for inserting and
-sending individuals to the pool,
-using different policies for population initialization, and also the
-incorporation of high-speed local resources to the pool to check what
-would be the real influence of the volunteer pool to the final
-performance.
-
-Finally, the implementation needs some refinement in terms of
-programming and also ease of use. Tools such as Yeoman for generating
-easily new experiments might be used, so that the user would have  to
-create only a fitness
-function, with the rest of the framework wrapped around
-automatically.
+% TODO
 

### 2016-03-13 PPSN2016/paper.Rnw
Computing correlation

@@
 
-these.experiments <- c("nooverlap-0","nooverlap","nooverlap-newgraph","nooverlap-newgraph-reboot-cachecrash")
+these.experiments <- c("nooverlap-0","nooverlap-newgraph","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
 ip.correlations <- data.frame(Experiment=character(),
@@
 
+\begin{table}[htbp]
+  \caption{Minute-wise correlation for number of participating volunteers. \label{tab:correlation}}
+\begin{center}
+<<correlations, cache=FALSE,echo=FALSE>>=
+#experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
+kable(ip.correlations)
+@ %def correlations
+\end{center}
+\end{table}
+%
 We can check this impact by looking at the column that refers to

### 2016-03-13 PPSN2016/paper.Rnw
Explaining correlation but maybe @mariosky or @fergunet have a better idea

@@
 <<correlations, cache=FALSE,echo=FALSE>>=
-#experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
+ip.correlations$Experiment <- c("base","graph1","reboot","cachereboot")
 kable(ip.correlations)
@@
 contributing to the common cache.
+
+In fact, as can be seen in Table \ref{tab:correlations}, which shows
+the correlation between the number of IPs from one minute to the next,
+the behavior of all experiments but the last one are similar, with a
+very high correlation indicating that the number of volunteers does
+not change from one minute to the next. However, there is a sharp drop
+for the {\tt cachereboot} experiment, which indicates a bigger
+dynamism, with individuals joining and leaving the experiment all the
+time. This might be circumstancial, but it also reflects the fact
+that the solution is found in less time and that users need less
+contributions to find it personally.
 

### 2016-03-13 PPSN2016/paper.Rnw
First full version
Refs #109 #108 #107

Please all revise the paper. I would like to submit it next week to go on to another paper for PPSN. @fergunet @unintendedbear @mariosky revise English, conclusions, everything...

@@
 time that the evolutionary algorithm itself makes the most of those
-contributions, independently of how many they are.
+contributions, independently of how many they are, and is able to find
+the solution to the problem in less time and with less contributions.
 
@@
 %
-\begin{table}
-  \caption{Experiments with overlaps eliminated:  summary of results. \label{tab:filtered}}
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|c|c|c|}
-\hline
-Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &\multicolumn{2}{|c|}{Median PUTs} & \#Runs &  Puts/IP \\
-&   (ms)        & Total & Used & Total & Used &  & \\
-\hline
-base & 293685 & 13.0 & 8.0 & 483.0 & 57.0 & 47 & 35.50000\\
-\hline
-graph1 & 768053 & 12.0 & 5.0 & 601.0 & 79.0 & 21 & 45.84615\\
-graph2 & 1477969 & 6.0 & 4.0 & 901.0 & 52.0 & 25 & 149.42857\\
-\hline
-reboot & 799471 & 9.5 & 6.5 & 656.5 & 67.0 & 44 & 76.48016\\
-\hline
-cachereboot & 463347 & 12.0 & 12.0 & 638.5 & 80.5 & 52 & 44.25911\\
-\hline
-\end{tabular}
-%% Initially generated with this code, 
-%%<<experiments.filtered, cache=FALSE,echo=FALSE>>=
-%%experiment.aggregates.filtered$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
-%%kable(experiment.aggregates.filtered)
-%%@ %def 
 %
-\end{center}
-\end{table}
+Since the point of using volunteer computing system is to cut costs,
+all experiments were set up using the OpenShift
+PaaS, which provides a free tier, making the whole experiment cost
+equal to \$0.00. In fact, the NodIO system can be deployed to any
+system as long as it can run node.js and access to the filesystem, for
+log files, is available. All these logs are available under
+rrequest. Processed files are available in the same repository that
+hosts this paper, as well as the scripts needed to process them.
+
 %
@@
 \end{figure}
-
-Since the point of using volunteer computing system is to cut costs,
-all experiments were set up using the OpenShift
-PaaS, which provides a free tier, making the whole experiment cost
-equal to \$0.00. In fact, the NodIO system can be deployed to any
-system as long as it can run node.js and access to the filesystem, for
-log files, is available. All these logs are available under
-rrequest. Processed files are available in the same repository that
-hosts this paper, as well as the scripts needed to process them.
-
+%
 Every run had slightly different conditions, although in some cases it
@@
 
+%
+\begin{table}
+  \caption{Experiments with overlaps eliminated:  summary of results. \label{tab:filtered}}
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &\multicolumn{2}{|c|}{Median PUTs} & \#Runs &  Puts/IP \\
+&   (ms)        & Total & Used & Total & Used &  & \\
+\hline
+base & 293685 & 13.0 & 8.0 & 483.0 & 57.0 & 47 & 35.50000\\
+\hline
+graph1 & 768053 & 12.0 & 5.0 & 601.0 & 79.0 & 21 & 45.84615\\
+graph2 & 1477969 & 6.0 & 4.0 & 901.0 & 52.0 & 25 & 149.42857\\
+\hline
+reboot & 799471 & 9.5 & 6.5 & 656.5 & 67.0 & 44 & 76.48016\\
+\hline
+cachereboot & 463347 & 12.0 & 12.0 & 638.5 & 80.5 & 52 & 44.25911\\
+\hline
+\end{tabular}
+%% Initially generated with this code, 
+%%<<experiments.filtered, cache=FALSE,echo=FALSE>>=
+%%experiment.aggregates.filtered$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
+%%kable(experiment.aggregates.filtered)
+%%@ %def 
+%
+\end{center}
+\end{table}
+%
 So we will have to check whether we have effectively eliminated these
@@
 
-% TODO
+Our main intention in this paper was to focus on the algorithm itself
+so that it became as effective as possible, and able to use
+computation by volunteers in the best possible way, or at least better
+than the initial configuration. This initial configuration included
+{\em overlaps} or clients running islands that had actually started in
+a previous experiment in the set. If we are interested in measuring
+how many users or their behavior in bulk this is not so important, but
+in this case our focus was on the algorithm so we had to try that all
+clients participating in it started more or less at the same time and
+from a random population. In fact, we managed to do so, except in some
+byzantine cases, by making the clients restart when the cache size
+collapses to 1 or to a size much smaller than the current; we called
+this client configuration {\tt reboot} and {\tt cachecrash}.  This
+configuration achiever a positive secondary effect: all clients
+contribute to fill the cache, having thus an actual effect on the
+chromosome pool and effectively making all clients present contribute
+to the simulation.
+
+The last configuration managed to find, with a relatively small number
+of evaluations, solutions in half the time that in previous sets of
+experiments. Although in the {\tt base} and {\tt graph} the median
+time and number of contributions was smaller, we cannot disregard the
+possibility that many of those solutions were found thanks to
+contributions from {\em carried over} islands with an algorithm that
+has run a big amount of evaluations already. At any rate, the {\tt
+  cachereboot} finds on average solutions four times as fast as {\tt
+  graph2} with twice as many average volunteers (or three times as
+many if we consider {\em effective} volunteers), needing 50\% less
+evaluations, thus meaning a big improvement in the performance of the
+evolutionary algorithm, which was the target of this paper. This
+improvement is less with respect to {\tt reboot}, but in any case the
+method to detect the beginning of a new run is much more effective and
+will be kept in the future.
+
+This future holds many challenges which once again will go in two
+different directions: gathering more users and improving the
+efficiency of the algorithm run in an asynchronous way by the
+users. In the first case, it will be interesting to make it more {\em
+  social} by comparing local performance to other's contributions, and
+showing it in a graph or being able to tweet advances or the fact that
+the solution has been found. Anything that keeps the users running the
+system and attracts new users will contribute to the speed of finding
+the solution. And the algorithm will have to be improved, mainly by
+making the population more diverse. Right now there is a rigid policy
+of contribution to the pool: send the best individual, but this could
+be relaxed. Some other policies could also be tested, as well as
+running heterogenous algorithms in the same way that was done in
+\cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other
+evolutionary algorith, seems to be the key, but there are many ways of
+approaching it.
 
@@
 
-This work has been supported in part by TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
-PROY-PP2015-06 (Plan Propio 2015 UGR)  % please, remember to include all the current projects   ;)   Thanks!
-. We would also like to thank the
-anonymous reviewers of previous versions of this paper who have really
-helped us to improve 
+This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
+2015 UGR) . We would also like to thank the anonymous reviewers of
+previous versions of this paper who have really helped us to improve
 this paper (and our work) with their suggestions. We are also grateful

### 2016-03-14 PPSN2016/paper.Rnw
REST & Experiments sction

@@
 \item A REST (representational state transfer) server that responds to
-  HTTP requests made by the browsers. These HTTP requests use JSON for
-  encoding information, in this case chromosomes and associates
-  information send by the clients and return requests made by the
+  HTTP requests made by the browsers. These HTTP requests use JSON for information 
+  encoding, in this case chromosomes and related
+  information sent by the clients and responses returned by the
   server.  The server has the capability to
   run a single experiment, storing the incoming chromosomes in a cache data structure
-  that is reset when the solution is found. This data has a finite
+  that is reset when the solution is found. This cache has a finite
   size that erases the oldest chromosomes once it has filled to
@@
   the version of the server used in this paper, we implemented a
-  policy by which the combination IP-fitness value could only be
+  policy by which the combination IP-fitness chromosome could only be
   inserted once into the cache. That means that if a particular IP
@@
 browser, but the server needs to know when the solution has been
-found, so a termination condition is also included as a configuration
-option in the server. In all cases we are using 50 Traps, a problem
+found, so a termination condition is also included in the server, as a configuration
+option. In all cases we are using 50 Traps, a problem
 that is difficult enough to need the intervention of several users for
@@
 pieces of meta data were stored: a time stamp, the client IP, the chromosome and
-value, the cache size in that particular moment, and also if that
+fitness value, the cache size in that particular moment, and also if that
 individual actually updated the cache or not.
@@
 equal to \$0.00. In fact, the NodIO system can be deployed to any
-system as long as it can run node.js and access to the filesystem, for
-log files, is available. All these logs are available under
-rrequest. Processed files are available in the same repository that
+system as long as it can run node.js and has access to the filesystem, for
+log files, if available. All these logs are available upon
+request. Processed files are available in the same repository that
 hosts this paper, as well as the scripts needed to process them.
@@
 far. Every volunteer contributed a certain number of individuals
-(represented by the PUT HTTP order) to the server. The ranked number
+(represented by the PUT HTTP request) to the server. The ranked number
 of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
@@
 is interesting to see how some sets have a number of contributions per
-IP that is slightly superior to the rest, specially so in {\tt
+IP that are slightly superior to the rest, specially so in {\tt
   graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
@@
 
-The baseline, called {\em base} in the Table, uses simply the
+The baseline, called {\em base} in the Table, uses the simple
 mechanism of dropping contributions from clients if they are repeated

### 2016-03-14 PPSN2016/paper.Rnw
Experiments

@@
 
-The baseline, called {\em base} in the Table, uses the simple
+The baseline, called {\em base} in the Table, uses the
 mechanism of dropping contributions from clients if they are repeated
@@
 overlaps. In order to do that, we will have to establish a threshold
-under which reasonably the solution cannot be found. In every
+under which we can reasonably expect that the solution cannot be found. In every
 evolutionary algorithm it might happen, by chance, that the solution
@@
 difficult problem will be found. Let us filter then the runs,
-eliminating those that use less than 1600 generations to find the
+eliminating those that used less than 1600 generations to find the
 solution which can reasonably be said that they did so because some
@@
 increasing the contribution of clients to the common pool. However, we
-should remark that still the clients will stop contributing if they
+should remark that the clients will still stop contributing if they
 get stuck in a particular solution, so the ratio keeps being low
@@
 clients that tap that cache. At any rate, an increased number of
-effective PUTs will have a positive impact in the algorithm.
+effective PUTs will have a positive impact on the algorithm.
 
@@
 increased, there are more volunteers contributing at the same time. In
-fact, as can be seen in the last column, the median number of
+fact, as it can be seen in the last column, the median number of
 contributions per user is {\em much lower} in the last row, which
@@
 change and might, in fact continue with their former run, thus not
-contributing to the common cache.
+contributing to the current common cache.
 

### 2016-03-14 PPSN2016/paper.Rnw
Conclusions

@@
 Our main intention in this paper was to focus on the algorithm itself
-so that it became as effective as possible, and able to use
-computation by volunteers in the best possible way, or at least better
+so that it could became as effective as possible, and would be able to use
+computations by volunteers in the best possible way, or at least better
 than the initial configuration. This initial configuration included
@@
 
-This future holds many challenges which once again will go in two
+This future work holds many challenges which once again will go in two
 different directions: gathering more users and improving the

### 2016-03-14 PPSN2016/paper.Rnw
Little change in the abstract. References #107

@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 this paper we analyze a volunteer evolutionary computing system called
-NodIO and try to find out a few rules that encourage volunteer
+NodIO, trying to find out a few rules that encourage volunteer
 participation and how to make the most of their contributions to the

### 2016-03-14 PPSN2016/paper.Rnw
Reviewing Introduction. References #109

@@
 key facts: the Javascript virtual machine every browser runs and a
-simplified standard interface for interacting with servers. Thus
+simplified standard interface for interacting with servers.
+% These are "things", not "facts". As facts would be... "every browser runs a Javascript virtual machine and that there is a simplified standard interface for interacting with servers (?)
+Thus
 creating a distributed computing experiment is just a matter of making
@@
 different design decisions will affect its performance.
-
+% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
 In this paper our focus will be in the second aspect: we will try to
 design a decentralized system that, at the same time, is able to use
-all resources available. This design will be done incrementally by
+all available resources. This design will be done incrementally by
 changing the system on the client and the server and measuring its

### 2016-03-14 PPSN2016/paper.Rnw
Reviewing State of the Art.

@@
 attempting to gauge how many users join the effort and also how long
-would it take a concrete amount of users to reach the solution. In
+would it take for a concrete amount of users to reach the solution. In
 general, they do not try to assess the performance of the algorithm
@@
 user has control of the browser, there is a limited amount of
-interaction with it, namely, the fact that by reloading the webpage he
-can apply a kind of hypermutation killing the current population and
+interaction with it, namely, the fact that by reloading the webpage they
+can apply a kind of hypermutation, killing the current population and
 generating new individuals some of which will make their way to the

### 2016-03-14 PPSN2016/paper.Rnw
Reviewing Description of the framework.

@@
 system based on the browser is simply a client-server system
-with a client embedded in the browser via JavaScript. We call this
+which client is embedded in the browser via JavaScript. We call this
 system {\sf NodIO}. All parts of the framework are free and available
@@
   that is reset when the solution is found. This cache has a finite
-  size that erases the oldest chromosomes once it has filled to
+  size that erases the oldest chromosomes once it has reached its % Also "it has been filled to its"
   capacity.  
@@
 NodIO needs a fitness function to work with; in this case the classical Trap 
-function \cite{Ackley1987} has been used; this function runs in the
+function \cite{Ackley1987} has been used; % I don't think this use of the ; is correct
+this function runs in the
 browser, but the server needs to know when the solution has been
@@
 Facebook, and also private groups in Telegram and WhatsApp have been
-used too for announcements. We also encourage friends and followers to
+used for announcements too. We also encourage friends and followers to
 put an original post for announcing it, instead of just sharing, in

### 2016-03-16 EvoSoft16/nodio-evosoft16.tex
Starting the new paper for EvoSoft

--- 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
@@
 
-\title{Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources 
-for Low Cost Distributed Evolutionary Computation}
+\title{NodIO: a framework for pool-based evolutionary computation}
 %
@@
 %
-% \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
-%   and CITIC, University of Granada, Granada, Spain \\
-% Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
-% Tecnológico de Tijuana, Tijuana, M\'exico\\
-% Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-% Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-% Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
-% University of Granada, Granada, Spain\\
-
-% \thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
-% }
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
+Juan-J.~Merelo, Pedro A. Castillo,Pablo Garc\'ia-S\'anchez,Paloma de las Cuevas\\
+\affaddr{Dept. of Computer Architecture and Technology
+  and CITIC, University of Granada, Granada, Spain} \\
+\email{(jmerelo|pacv|pgarcia|palomacd)@ugr.es}
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
+Mario Garc\'ia-Valdez,\affaddr{ Dept. of Graduate Studies, Instituto
+Tecnológico de Tijuana, Tijuana, M\'exico}\\
 }
@@
 \begin{abstract}
-
-Using volunteer's browsers as a computing resource presents several
-advantages, but it remains a challenge to fully harness the browser's
-capabilities and to model the user's behavior so that those
-capabilities can be leveraged optimally. These are the objectives of
-this paper, where we present the results of several evolutionary
-computation experiments with different implementations of a volunteer
-computing framework called {\sf NodIO}, designed to be easily
-deployable on freely available cloud resources. We use different
-implementations to find out which one is able to get the user to lend
-more computing cycles and test different problems to check the
-influence it has on said performance, as measured by the time needed
-to find a solution, but also by the number of users engaged.  From
-these experiments we can already draw some conclusions, besides the
-fact that volunteer computing can be a valuable computing resource and
-that it is essential to be as open as possible with software and data:
-the user has to be kept engaged to obtain as many computing cycles as
-possible, the client has to be built to use the computer capabilities
-fully, and, finally, that the user contributions follow a common
-statistical distribution.
 
@@
 \section{Introduction}
-
-Nowadays there is a big and increasing number of distributed
-evolutionary computation (EC) software frameworks available, usually
-as open source frameworks \cite{Parejo12Survey}.  Most of them have
-two features: they are {\em desktop} applications, that is, they must
-be compiled or run directly on the operating system layer and, in addition,
-they assume that the nodes that are going to run the system are under
-the control of developers or they at least have access to them.
-
-There is, then, space for other EC frameworks that do not make any of
-these assumptions, such as {\sf NodIO} presented in
-\cite{2016arXiv160101607Manom}. {\sf NodIO} implements a client-server application; the server offers the possibility to program an
-evolutionary algorithm that is embedded in the served pages, and those pages can communicate between each other via the server. On the other hand, the client part can run in any
-web-enabled device, including many over which developers have no
-control, that is, volunteers.
-
-We were interested in this kind of systems because, during the big
-global crisis, our group faced the problem of diminishing funds for
-buying new hardware. This was aggravated by the increasing maintenance costs and
-extended downtime resulting from the continuous failures of existing
-clusters.
-Considering this, we leveraged our experience in the design of web
-applications since their inception, including JavaScript, and other
-volunteer and
-unconventional distributed evolutionary computing systems to 
-design and release a new free framework that would allow anyone to
-create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
-servers, and browsers as clients.
-% (Paloma) It seems that you've started the introduction twice. The
-% first two paragraphs "introduce the introduction", and so does the
-% third. It is specially noticeable when you find "evolutionary
-% computation (EC)" twice.
-% Yep, I want to say it all, but don't know how to put it together.
-
-%PABLO: He añadido el siguiente párrafo para el issue #31
-There exist other volunteer computing frameworks written in compiled languages,
-such as C++ or Java, that, in principle, can achieve better performance than
-JavaScript. However, one of the advantages of {\sf NodIO} architecture is that
-clients do not need to install anything on their machines, they just
-%Pablo: According English Stack Exchange, you install "ON" a machine (but "IN" a directory)
-need to open a web page  
-in a compatible, that is, almost any, browser, such as Chrome, Safari or  Firefox, which are
-available in almost every desktop or mobile device. Another advantage is
-that the source code can be directly examined by volunteers, because it is not
-compiled as other options such as Java Applets, ActionScript, or
-ActiveX are. Also, as we will discuss in this paper, it can provide some extra
-features such as enhanced security or operating system
-independence. This system, however, is not intended to substitute
-already existing frameworks running on dedicated hardware or the cloud, but to
-%Pablo: same here, I learnt this with my paper of EAs "on clusters"
-complement them.
-
-%PABLO: hay que enlazar con el siguiente párrafo
-%Desktop clients have the same issues as any 
-%volunteer computing system, including our % (Paloma) their?
-% own: these issues are
-%related to security as well as
-%safety, which introduce
-
-%Mario: Rewrote: 
-%One of the aspects that must be taken into account in volunteer based
-%frameworks is its {\em social} aspect in the design, with 
-In general, several volunteers will be needed to
-overcome the overhead involved in using a client-server architecture
-and this number of volunteers is obviously not guaranteed. Therefore, the {\em social} aspect must be taken into account when designing 
-volunteer based frameworks, with 
-issues related to security, trust, and privacy among others. 
-The computing system becomes then a {\em techno-social system}
-\cite{vespignani2009predicting}, in the sense that its social aspects
-cannot be separated from its technical specifications, with one
-feeding into the other. One of these aspects is safety for the user
-running it, as well as for the scientist: the user must be reasonably
-sure that the 
-experiment is not misusing their resources and the scientist must have a
-certain safety in the sense that their server is not going to be
-attacked via cheating or denial of service attacks.
-
-We have decided to approach security in a pragmatic way. Including
-many security features % (Paloma) I'm kind of lost here... Sentences
-                       % seem to be unconnected and unrelated...  
-% Separated paragraph and substituting wildcard words - JJ
-would decrease performance; on the other hand, we publish all data
-resulting from the experiment once private information, such as IPs,
-has been processed to an untraceable hash. The experiments themselves
-are a proof that there are enough users trusting the experiment to
-give it their CPU time and that there has been no attempt to cheat by
-sending specially crafted chromosomes or to attack the server to make
-it stop working because there was no incentive in doing so.
-% (Paloma) Is this really true? Can we really state that there were no
-% attacks because there was no incentive? Attackers don't always need
-% an incentive to perform an attack, or the incentive could not be
-% perceived by us. And why do the experiments prove this? Are there
-% specific measures? 
-% There were no attacks, that's a fact. There could be in the future,
-% but not for the time being - JJ
-The
-system, besides, is running in a dedicated Platform as a Service (PaaS)
-container or Virtual Private Server so accessing the server would only
-show the few data obtained from the experiment, adding security against attackers. In addition, making it
-free software ensured the meager bounty obtained by attackers were
-known by anyone participating in the experiment, via 
-links in the web page that hosted it. % (Paloma) What was known? The
-                                % fact that it was running in a PaaS?
-                                % Why was that known by links?
-%That there wasn't anything to hunt. Cleared, I expect.
-
-In this paper, we will carry out three experiments, after establishing
-the baseline of the performance that can be expected in a JavaScript
-based system and its comparison with other languages in
-\cite{2016arXiv160101607Manom}. First, several experiments are
-performed using a simple Ajax client that does not 
-restart when a solution is found. Second, we change the client a bit
-so that it uses web workers, that is, it runs in the background
-process and then it restarts when the solution is found. Third, we use the same
-% Pablo: it restarts when the solution is found (it has not been
-% mentioned yet)
-% Just do it! - JJ
-client, but with a harder problem that cannot be easily solved by a single
-volunteer and thus needs the contribution of many volunteers to
-proceed. 
-%Mario: refrased: 
-%With these three experiments we want mainly to try and model
-With these three experiments our main objective is to try and model
-the behavior of users and learn how different problems and
-implementation choices have an influence on it. That is why in this
-paper we will collect information on the number of users participating
-in each experiment and also the time they spent doing so. In this
-paper we will show how this participation follows certain patterns,
-and that they depend on the implementation itself as well as the
-problem and its difficulty; however, certain implementations allow to
-extract more CPU cycles from volunteers.
-
-%Pablo: To do this we compare the time spent by each user because...
-% Pablo, please, don't JJ me. Do the changes, because I'm not learning
-% now. I have added a bit anyway.
 
@@
 
-The web used as a resource for distributed computing has a
-long history, and almost since the beginning it was used for
-evolutionary computation even before JavaScript was
-introduced as a browser-based scripting language in 1997. Before that
-time other technologies,
- including Flash animations and VBScript, ActiveX, or Java applets were
-used with the same purpose; for instance, Java was pointed out in \cite{soares1998get} as a
-``language for the internet'', providing some advantages such as multi-architecture compatibility or
-security mechanisms, and was proposed in \cite{chandy1996world} as the
-basis for the creation of a world-wide peer to peer system using compiled Java
-programs. The ATLAS system
-\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
-with another language called Cilk; all these papers indicated that
-the introduction of an architecture-independent system such as the
-Java virtual machine opened the doors for ultrascale, Internet-wide,
-metacomputers. 
-%% We need to only reference this section?:
-%%------------------- BEGIN
-% This paper % (Paloma) No really need for that "also", plus it complicates reading
-% establishes the desirable features of
-% any {\em global computing} infrastructure, that can also be applied to
-% distributed evolutionary computation experiments that rely on
-% volunteer computing: scalability, heterogeneity, fault tolerance,
-% adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
-% reasonable performance. But hierarchy, this is, the capability of
-% preferentially using resources in the same organization, is a feature
-% that could be either disregarded or included within adaptivity, which
-% leaves us with eight desirable properties. 
-% % (Paloma) Re-write. For example, "But is hierarchy, this is, the
-% % capability of preferentially using resources in the same
-% % organization, the one that could be either disregarded or included
-% % in adaptivity, which leaves us with eight desirable properties." 
-% % (Paloma) And I re-wrote what I proposed yesterday.
-% % thanks :-) -JJ
-% %%------------------- END
-% %% EDIT:
-% %%------------------- BEGIN
-% The fulfillment of these properties could be achieved by using any of the
-% languages and associated technologies mentioned above, provided that the user had
-% installed the required virtual machine plugin first. 
-% %%------------------- END
-The fulfillment of the desirable features of any distributed evolutionary 
-computation experiments that rely on volunteer computing \cite{2016arXiv160101607Manom} could be achieved by using any of the
-languages and associated technologies mentioned above, provided that the user had
-installed the required virtual machine plugin first.
-
-% (Paloma) These should be separate sentences
-However, embedding a
-distributed computation client in the browser has the advantage of
-providing a single and universal way of accessing resources via an
-URI, something that is not guaranteed by using desktop applications. 
-Baratloo et al. tried in 1996 to create a framework with
-the aforementioned qualities called {\em Charlotte}
-\cite{baratloo1999charlotte}. This framework, which implemented a
-distributed shared memory, was inspired by a previous work
-on parallel computing libraries PVM and MPI and solved most of the issues by
-using the Java virtual machine embedded in the browser,
-which provided
-security to the user, uniformity of computing environment
-independently of the operating system, and access to a common data store in the
-server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
-and as it runs compiled bytecode it also guaranteed a certain level of
-performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
-Other issues such as scheduling, which are related to adaptive 
-parallelism, were solved on the server. But it is interesting to note that,
-even in the early stages of the web, the possibility of
-% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
-browser-based distributed computing was already researched.
-
-Other systems, like JET, proposed by Soares et
-al. \cite{soares1998get}, also used Java. JET supports
-the execution of parallel applications over the Internet and features
-a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
-since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
-the collection of algorithm analytics by the
-experiment designer, unlike the other systems mentioned
-above. Other proposals contemporary to JET, such as {\em SuperWeb}
-\cite{alexandrov1997superweb} are roughly a translation of classical
-distribution techniques, with overly complicated architectures that include
-brokers for resource discovery. SuperWeb is focused,
-% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
-however, in solving the
-issues involved in browser-based computing and also in making a system
-that is able to work across different types of computers. Interestingly enough, this
-paper incorporates a discussion on the economic model, paving the way
-for the for-profit distributed computing systems that arose later on. 
-On the other hand, the potential for volunteer computing using
-browsers was realized
-later on \cite{sarmenta-bayanihan} as well as the potential of
-user misconduct when having access to the source code \cite{sarmenta-sabotagetolerance}. 
-However, these early efforts by
-Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
-less widespread.
-
-The main problem with using Java and all the other browser-embedded
-technologies is that they are not {\em universal} in the
-sense that an extra component, namely, the Java virtual machine or another
-virtual environment has to be installed in the browser.
-% Also these virtual environments have become targets of malicious attacks; 
-% for instance numerous vulnerabilities have been discovered in 
-% the proprietary Adobe Flash runtime environment and
-% video player \cite{ford2009analyzing,watanabe2010new}. 
-The need for
-these additional environments has been mitigated with the
-adoption of the HTML5 standard \cite{anthes2012html5}, which is
-tightly integrated with JavaScript.
-
-JavaScript can be used either for unwitting, that is, stealth,
-\cite{unwitting-ec} or volunteer 
-\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
-evolutionary computation and it has been used ever since by several
-authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}.
-
-% that even
-% used the client's GPU \cite{duda2013gpu}. 
-%Also a bit out of scope. Have to reduce the paper. 
-% Even if, as said above, Java
-% is not the best option, many other researchers have
-% used Java \cite{chong:1999:jDGPi} and others have gone away from the
-% server-based paradigm to embrace peer to peer systems
-% \cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
-% computing 
-% platforms avoid single points of failure (the server) but, since they
-% need a certain amount of infrastructure installed to start, the
-% threshold to join them is much higher. % (Paloma) Do yo
-                                % mean... higher? A low threshold
-                                % means that is easy to happen
-                                % something.
-%Correct. But maybe in this new version we should eliminate this
-%paragraph altogether. 
-% The next paragraph has been suppresed.
-
-%Also suppressing this one which is probably not needed in this paper
-%- JJ
-% JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
-% According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
-% metaheuristics optimization is that it allows an improvement in communicability,
-% reproductibility, interoperability, automated assembly, knowledge engineering,
-% and efficiency. In the work presented in this paper, JavaScript has been used throughout the full
-% stack, which has the advantage of using the self-same lines of code for
-% writing an evolutionary algorithm for the
-% desktop or command line and browser clients, both based in the EC
-% library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
-% when writing code and leverages 
-% the expertise acquired
-% % (Paloma) Not sure if "acquired" should be written before
-% % "expertise"... sounds better to me, anyways
-% % Don't know what you mean. Write it down and we'll see - JJ
-%  in the library itself.
-
-Additionally, we are also interested in measuring the performance
-of volunteer computing systems, an area in which there have been
-relatively few efforts. There were some initial attempts to avoid the differences in performance
-that could be obtained from volunteers  by making
-the algorithm adaptive to the kind of resources allotted to it
-\cite{milani2004online}. This is actually not such a big problem in
-algorithms such as the Evolutionary Algorithm (EA) that can easily be
-parallelized via population splitting or by farming out evaluations to all
-available nodes. Lately, several approaches have focused on the
-fault-tolerance of volunteer algorithms
-\cite{gonzalez2010characterizing} which can, of course, be studied in
-the more general context of distributed computing 
-\cite{nogueras2015studying} or be included in a broader study of the
-performance of the EA itself
-\cite{DBLP:journals/gpem/LaredoBGVAGF14}.
-
-Moreover, measuring the performance of the resulting metacomputer
-involves understanding its dynamics. Initial
-work was done for peer to peer systems by Stutzbach et
-al. \cite{stutzbach2006understanding} and extended to volunteer
-computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
-%But the raw material of           
-%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
-But some of the essential metrics in volunteer computing like the
-number of users or the time spent by every one in the
-computation in browser-based volunteer computing experiments, have
-only been studied in a limited way in 
-\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
-run. Studies using volunteer computing platforms such as SETI@home
-\cite{javadi2009mining} found out that the Weibull, log-normal, and
-Gamma distributions 
-modeled quite well the availability of resources in several clusters
-of that framework; the shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas:
-there are many users that give a small amount of cycles, while there
-are just a few that give many cycles. This is in concordance with the
-results obtained in \cite{agajaj}. % (Paloma) I have a doubt here, should this reference be anonymised or not?
-Although SETI@home works
-at a larger scale in both quantity of volunteers and the scope of projects, 
-our initial tests indicate that
-this kind of distribution might be the most appropriate to describe
-user participation in several different metrics. However, in this
-paper we will perform different experiments in different setups and
-will try to come up with a more precise model. 
-%Mario: erase: Next we describe the framework itself.
-As far as we know, this paper presents one of the few experimental
-results using a volunteer computing system which uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
-very diverse computational resources; these will be described next.
-
-\section{Methodology and experimental results}
+\section{The NodIO framework and API}
 \label{sec:exp1}
-
-% A bit of description of the framework, with reference to published
-% reports - JJ
-
-We set up different experiments by deploying the server in free cloud
-services, changing it from one experiment to the next if needed. This server runs a web service
-which stores chromosomes sent from clients and serves requests to
-them; it is 
-an open source system that has been extensively described in
-\cite{2016arXiv160101607Manom}. The client runs an evolutionary
-algorithm written in JavaScript that after a number of generations
-sends the best individual back to the server, requesting at the same
-time a random individual of those stored in the server. The experiment
-starts by setting up a website that is sent by the server and then announcing it on several social
-networks so that people contribute to it by visiting the page and
-staying there for a while.
-
-\subsection{Initial tests with a simple Ajax client}
-%% Above we mention three experiments:  Simple Ajax client, Web Workers, Harder Problem
-%% Im renaming subsection to  reference the first experiment
-
-Initial tests with an evolutionary algorithm solving the 40-Trap
-problem were set up using the free PaaS OpenShift; the results are described in
-Table \ref{tab:summary:os}. In general, every test took several
-days and during its performance the problem was solved more than 50 times with
-several hundreds different IPs participating in it; each
-evolutionary algorithm stopped when the solution was found. 
-%
-\begin{table*}
-\caption{Summary of time per run, number of IPs and number of PUTs per IP in the initial runs. \label{tab:summary:os}}
-\begin{center}
-\begin{tabular}{l|ccccccccc}
-\hline
-Date & Experiments &\#IPs &\#IPs &  \#IPs & Time (s) &  \#PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment\\
-     & \# & Different & Median & Max & Median &  Median & & & correlation \\ 
-\hline
-4/4 & 57 & 191 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
-4/24 & 231 & 559 & 5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
-7/31 & 97 & 179 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-
-The second-to-last column in Table \ref{tab:summary:os} shows the percentage of experiments in which running the
-experiment in the browser takes {\em less} time than in the
-desktop. Volunteer computing systems do not exclude desktop
-computing and, in fact, the server can handle any client, not only
-those running JavaScript on the browser. Although, we include it here for the
-sake of comparison and to show that, being a free resource, it can
-anyway make a valid and valuable contribution to any distributed or
-desktop computing experiment. Another interesting result is that the
-correlation between users of an experiment and the next is
-non-existent, that is, computing nodes are not {\em
-  staying} after the solution has been found. This is the issue that we
-address in the next experiments. 
-%Mario: changed: will address in this paper.
-
-% \begin{figure}[!htb]
-% \centering
-% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
-% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
-% \includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
-% \caption{Duration of experiments vs. number of different IPs (nodes)
-%   participating in it, with averages and standard deviation shown as
-%   black dots and lines, single dot if it was 1). Shade of gray or
-%   color indicate how many experiments included that many unique IPs.
-% From left to right, experiments 4/4, 4/24 and 7/31.}
-% \label{fig:duration}
-% \end{figure}
-% % ../data/time-vs-IPs-openshift.R
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
-\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
-  logarithmic scale. Dot color (or gray value) is related to the number of IPs
-  participating in the experiment. From left to right, experiments
-  4/4, 4/24 and 7/31.} 
-\label{fig:zipf:os}
-% computed via ../data/time-vs-IPs-openshift.R
-\end{figure}
-%
-% We will have to analyze experimental data a bit further to find out why
-% this happens and also if there are some patterns in the three sets of
-% experiments. An interesting question to ask, for instance, is if
-% by adding more computers makes the experiment take less. In fact, as
-% shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
-% not seem to contribute to decreasing the time needed to finish the
-% experiment. However, the cause-effect relationship is not clear at
-% all. It might be the opposite: since experiments take longer to finish
-% and might in fact be abandoned with no one contributing for some time,
-% that increases the probability of someone new joining them. In fact,
-% with experiments taking a few seconds and due to the way the
-% experiments are announced, it is quite difficult that several
-% volunteers join in in such a short period of time, even more if we take
-% into account that volunteers are not {\em carried over} from previous
-% experiments. This implies that it would be convenient to use a problem
-% of a bigger size to check this hypothesis as we have done in the next
-% experiment. 
-%  However, at this point we
-% have not found this convenient since there are several other issues
-% that have to be solved, as it will be shown next. %Creo que este párrafo
-%saldría sobrando ahora que hicmios un experimento más grande
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-4.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-4-24.png}
-\includegraphics[width=0.32\linewidth]{puts-openshift-7-31.png}
-\caption{Number of PUTs per unique IP and fit to a Generalized
-  Extreme Value distribution (in a lighter shade or blue). From left to right, experiments
-  4/4, 4/24 and 7/31.} 
-% Plotted with ../data/plot-zipf-openshift.R
-\label{fig:puts:os}
-\end{figure}
-%
-It is also interesting to check the distribution of the experiment
-duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
-a Zipf's law, with similar distribution along all three runs. The 4/24
-run is the most complete and shows an S-shape, which implies an
-accumulation of experiments taking similar time and around 100
-seconds. The most interesting part is the {\em tail}, which shows how
-many experiments took a desirable amount of time, on the order of
-10 seconds, and which appears in all three graphs. As it can be seen,
-it sharply drops implying there are 
-just a few of them, and with diminishing probability as time
-decreases. This exponential distribution also appears in Figure \ref{fig:puts:os}, which shows the
-distribution of HTTP PUTs, equivalent to the number of
-generations divided by 100, contributed by every user. These results show a Zipf-like behavior,
-so we have fitted it to the Generalized Extreme Value distribution,
-with the resulting parameters shown in Table \ref{tab:puts:os}.
-%
-\begin{table}
-\caption{Summary of fit to Generalized Extreme Value distribution of
-  the number of PUTs per unique IP. \label{tab:puts:os}}
-\begin{center}
-\begin{tabular}{l|ccc}
-\hline
-Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
-0.1377 \\
-4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\
-7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-This distribution was originally proposed to fit extreme values
-\cite{resnick2013extreme} and contains, as a special case, the inverse
-Weibull distribution which was fitted to volunteer computing
-frameworks such as SETI@home \cite{javadi2009mining}. We obviously do
-not pretend to compare our framework in scale or complexity with it, but
-to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which also appears in our framework. This
-distribution is governed by three parameters, the usual location $\mu$
-which is related to where it has its {\em center} and a scale $\sigma$,
-related to the size, but also a third shape $\xi$ parameter that is
-related to its skewness, that is, how skewed it is around the central
-location. Positive parameters indicate that the distribution {\em
-  leans} towards the origin, and negative ones towards the other extreme
-value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
-greater than one and between one and 1.4, which indicates that the
-three experiments share this origin-leaning pattern, with many users
-donating a few cycles and just a few donating extreme values. Random
-distributions with these parameters have been plotted in red in
-Figure \ref{fig:puts:os}, indicating that the fit is good enough. The
-predictive value of these fits, however, is limited, over all taking
-%Pablo: question for Paloma: is over all or overall?
-into account the low correlation between successive events shown
-above.
-
-At any rate, this also shows that a convenient way of increasing the
-%Mario: As I understend is: 
-computing power would be increasing this minimum amount of experiments
-per user. For this reason, we made some changes to the implementation,
-whose results are presented in the next subsection. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
-
-\begin{table*}[!htb]
-\caption{Summary of {\sf NodIO-W$^2$} (an acronym for the version that
-  uses web workers) and comparison with the previous
-  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
-\begin{center}
-\begin{tabular}{l|ccccccc}
-\hline
-Date & Median \#IPs & Max \#IPs & Median time (s) & Median \# PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
-\hline
-{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
-{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
-\caption{Time employed in every experiment vs. number of IPs (in
-  abscissas). The black dot and line show the average and standard
-  deviation, the blue to red shades the number of experiments with the same
-  number of IPs. } 
-\label{fig:ipstime:w2}
-% ../data/time-vs-IPs.R 
-\end{figure}
-
-\subsection{Experiments using web workers}
-\label{sec:w2}
-
-The results reported above indicated that we should look for a way of
-increasing the time users spend on the page. This led us to create an
-architecture that uses web workers, which is described in
-\cite{2016arXiv160101607Manom} and we abbreviate to {\sf
-  NodIO-W$^2$}; web workers work in the background even if
-the web page is not in focus. A few improvements were also tested,
-basically using a random population and restarting the web worker when
-the solution was found.
-
-The experiments were run in the same way as before, using several
- announcements in Twitter throughout several days by the end of July
-2015. Eventually, more than one thousand experiments were completed in
-a matter of days. A summary of results is shown in Table
-\ref{tab:summary:ww}, comparing it with the aggregate of the results
-obtained with the initial version of {\sf NodIO}. These results are
-remarkably different, being similar only in the median and maximum
-number of IPs, although in this case it would be combination
-IP-worker, as we considered every worker as a unique IP. The median
-number of seconds is almost 80\% better than the previous version, and
-the median number of PUTs is three times better; this causes that most
-of experiments take less time than one of the baselines, and one third
-less than the second baseline. Thus, the conclusion in this case is
-that we can obtain, through volunteer computation, a result that most 
-of the times is better than we would by using a similar, non-parallel,
-desktop setup, which makes this system suitable for massively
-distributed evolutionary algorithms.
-%
-\begin{table}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of PUTs per worker. \label{tab:puts:ww}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 18.275 $\pm$ 0.837  &  25.96  $\pm$ 1.22 & 1.242   $\pm$ 0.047 \\
-Weibull & ND & 75.11 $\pm$ 2.97  & 0.697 $\pm$ 0.014 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-
-However, it is interesting to note why this is so, and the first hint
-is the inter-experiment correlation between the number of IPs in
-successive experiments. While before it was an unremarkable 10\%, it is
-now more than 43\%, making the number of IPs in contiguous experiments
-highly correlated. The main reason for this is the setup in the
-version using web workers, which restarts an island after a solution has been
-found and also that it keeps running even if the tab is not on the
-foreground. This means that volunteers can leave the experiment
-running for as long as they want and they will be contributing 
-experiment after experiment, unlike the previous version, where they
-stopped contributing after one solution was found. This is also
-reflected in the number of PUTs, more than 50\% of the volunteers run the experiment for more than
-4000 generations. All things considered, this means than there are
-%Mario: many more more threads {\em simultaneously} running?
-more threads and many more computers {\em simultaneously} running,
-leading to this almost 5-fold increase in running time.
-
-We are also interested in measuring the scaling properties of this
-model, that is, the relationship between the number of IPs
-participating in an experiment and the time it takes to complete
-it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
-experiment in terms of time (in milliseconds) vs. % (Paloma) Isn't "vs." informal?
-the number of workers participating
-in it. Although there is not a clear trend, the graph seems to say
-that the time needed for finding the solution in this evolutionary
-algorithm does not depend on the number of workers participating in
-it. This does not mean that it is independent on the number of {\em
-  simultaneous} users, which is probably the case although it is more
-difficult to measure. There seems to be a trend towards a decreasing
-standard deviation in the time, with more volunteers adding both {\em
-  robustness} to an experiment, and certainty in the time it is taking. % (Paloma) Should "certainty" go in italics? Otherwise it's confusing. 
- However, more experiments would be needed to check this hypothesis.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
-\caption{Ranked number of PUTs per worker and GEV fit (in light
-  or blue color, left) 
-and Weibull (in lighter or blue shade, right) for the {\sf NodIO-W$^2$} model.}  
-\label{fig:gev:w2}
-% File ../data/ips-puts.R
-\end{figure}
-%
-As we did in the previous subsection, we have also fitted the number
-of HTTP PUTs per worker to a  Generalized Extreme Value (GEV) function, with the result shown
-in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
-in the previous experiment, Table \ref{tab:puts:os}, we find higher values
-for the location $\mu$ and scale $\sigma$ parameters, accounting for
-the higher number of contributions per volunteer that has been
-previously observed. However, the shape $\xi$ parameter is
-substantially similar, with a {\em bell} shape leaning towards the
-origin, indicating a pattern of a few users with many contributions
-and many with few contributions. If we plot the model and the data
-side by side, as shown in Figure \ref{fig:gev:w2}, we see that the fit
-is not so tight as in the previous case, with the model overestimating
-the number of individuals with a high number contributions. This is
-probably due to the fact that in the previous case the user needed to
-voluntarily reload the page to contribute the most generations, with a
-{\em phase change} between the people that did so and the people that
-just ran the experiment once until completion. In this case, however, the
-user just needs to let it run, without that phase change and thus we obtain a
-more {\em egalitarian} distribution of contributions, which rather
-corresponds to a Weibull function, as shown in Figure
-\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
-distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
-equal to 0.69, a value that is remarkably similar to the 0.5 value
-found for time devoted to games in Chambers et
-al. \cite{chambers2005measurement}. It is important to note that, in
-this case, the number of PUTs made per client does not correspond to a
-uniform amount of computation, since 100 generations with a random
-population belong, in each case, to different number of
-operations. In other words, the same number of PUTs will correspond to
-different number of operations in each case. However, since clients
-are also different and take a different time in each case, it is
-difficult to ascertain how this might have an influence on the
-statistical distribution.
-
-In general, this second version of the {\sf NodIO} framework and the
-single experiment performed prove that it can be the foundation for a distributed
-high-server-performance evolutionary computation platform,
-providing reasonable algorithmic performance and being, in general,
-easy to use with a straightforward modification of the fitness
-function.
-
-\subsection{A hard optimization problem: A Shifted Rotated Rastrigin's function}
-\label{sec:rastrigin}
-
-% (Paloma) I would propose as section title "A hard optimization problem: The Shifted Rotated Rastrigin's function"
-
-The fact that the solution of this particular problem was found in a
-few seconds and restarts were necessary yielded interesting results, % (Paloma) Where?
-but we needed to work on a harder problem, that is why created a new
-client version that optimized a harder variation of the Rastrigin's function, 
-used as a benchmark in evolutionary
-algorithms. The set up for this function, and a comparison of the
-baseline speed for JavaScript in a standalone experiment and other
-languages is shown in \cite{2016arXiv160101607Manom}, where the details of
-the implementation have also been extensively described.
-
-Specifically, this experiment was deployed in a virtual private server with 512MB of memory
-running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
-experiments where the optimum was expected to be found many times, 
-this problem was expected to constantly run for many days finding only
-sub-optimal approximations. In order to achieve this, the only change needed 
-in the framework was to limit the number of chromosomes kept in the pool. 
-Previously, every PUT request which included a new chromosome increased 
-the size of the pool. But in a long-running experiment this has to be avoided
-in order to not exhausting the available memory. In this experiment the size
-of a chromosome was 1000 double precision numbers, so that considering this
-and the memory available in the server, the pool was limited 
-to 10,000 chromosomes.  In order to keep this size
-constant, a random chromosome was pulled out of the pool before inserting a
-new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
-and {\tt classic-ea-float} modules were used. The parameters for the EA 
-were tournament size = 3 and  population size = 500, returning exchanging a
-chromosome with the server every 100 generations. Again, there were two workers in each window.
-As in previous experiments, a few calls to participation were published on social networks.
-
-%
-% \begin{figure}[!htb]
-% \centering
-% \includegraphics{rastrigin-fitness.png}
-% \caption{Fitness of chromosomes sent to the server pool along time in
-%   the $x$ axis.} 
-% % Plotted with ../data/plot-rastrigin-fitness.R
-% \label{fig:puts:rastrigin}
-% \end{figure}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
-\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
-\caption{Number of unique IPs participating in the Rastrigin
-  experiment per minute (left) and per hour (right)} 
-% Plotted with ../data/plot-rastrigin-IPs.R
-\label{fig:ips:rastrigin}
-\end{figure}
-%
-\begin{table}
-\caption{Summary of fit to GEV and Weibull distribution of
-  the number of PUTs per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
-\begin{center}
-\begin{tabular}{cccc}
-\hline
-Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
-\hline
-GEV & 35.759  &  343.336   & 9.877 \\
-Weibull & ND & 29.00 $\pm$ 2.15  & 0.40 $\pm$ 0.01 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
-\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
-\caption{Data and fit to GEV (left) and Weibull (right) of the number
-  of PUTs per worker for the F15 (Rastrigin) function.}  
-% Plotted with ../data/ips-puts-rastrigin.R
-\label{fig:fit:rastrigin}
-\end{figure}
-%
-
-The results for this new experiment have been extremely interesting
-and quite different from the others. The most important
-characteristic that sets this experiment apart is
-the fact that it did not find the solution for the time it was
-running, instead lowering, little by little, the fitness until it was
-disconnected. The fitness of chromosomes sent by clients managed to
-surpass the 9000 value by the end of the execution; and although initially there
-was a high number of new nodes, this number decreased with time.
-
-This is in part due to the fact that the presence of nodes is not
-constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
-number of unique IPs participating in the experiment per minute (left)
-and hour (right). The two graphs have the same shape: At the beginning there are very few or only
-one, but later one there are peaks of almost 30 unique machines a
-minute; these peaks appear again when new announcements are made in
-Twitter or any other social networks, to go down to a few remaining
-nodes later on. It is interesting to note that, at the most, more than
-100 unique nodes were participating in the experiment during one hour;
-also that there are sustained peaks during a day of 6-7 computers
-every minute. This is due to the fact that the computation is
-asynchronous, but also to the fact that different computers stay in
-the simulation for a different period of time. This period has been
-plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
-PUTs per node, ranked by number, along with a fit to the two
-distributions we have used before, GEV and Weibull; the parameters of
-this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
-experiments, in which GEV fit was relatively good, in this case it is
-quite obviously not. However, the fit to the Weibull distribution,
-shown at the left of the image, is relatively good. As should be
-expected, the shape and scale are different to the one in the previous
-experiments with the Trap function, however, the shape is quite
-similar (0.40 vs 0.69), being in both cases less than one, which
-indicates a decreasing distribution of the same kind. In fact, the
-difference in the scale parameter indicates that the difference
-between the time devoted by a volunteer and the next is bigger in this
-case. This is probably due to the fact that the duration of the
-experiment is bigger, allowing for bigger differences in
-behavior. While the first in the Trap function might have devoted a
-few minutes and the second a few seconds less, the difference between
-the first and second in this experiment might be in the scale of
-hours.
-
-In general, this experiment confirms what was found in the others in
-several aspects: first, there is a regularity in the distribution of the
-time assigned by volunteers to it, second, there is a good and certain
-amount of volunteers that will become part of the distributed
-computer, in some cases up to one hundred; also, in some cases there
-will be users that will spend several hours, two orders of magnitude
-more generations in this case than in the previous one.
-%Mario: than in the previous one. Are we refereing to the Wˆ2 version?
-There is a clear influence of the time to solution: a short time means that every
-experiment will only gather a few volunteers, but also that they will
-stay only for a short time. A long time will make users stay for
-longer, but if it takes too long, they will eventually lose interest.
-
-We will discuss this and the rest of the findings of the paper in the
-conclusions.
 
@@
 
-In this paper several experiments with a client-server architecture for volunteer and distributed
-evolutionary algorithms have been performed. The architecture uses the
-browser and has been generated using the {\sf
-  NodIO} framework, which includes  {\sf NodEO} as an evolutionary
-algorithm library. Volunteers are {\em in the cloud}, as stated in the
-title, 
-since they are a {\em CPU as a service} for the persons running the
-experiment. In fact, in this paper we have tried to put some figures
-on the real size of that {\em cloud} in case it is used by itself or
-in conjunction with other local or 
-cloud-based methods to add computing power in a seamless way through
-the pool that {\sf NodIO} creates.
+%---------------------------------------------------------------
+\section*{Acknowledgment}
 
-In order to establish a baseline performance, which was published in
-\cite{2016arXiv160101607Manom}, the evolutionary algorithm was run in a desktop client application written in JavaScript
-using NodEO to solve the 40-trap function. The first experiments in a
-volunteer computing scenario proved that, although obtaining better performance than the
-baseline  was possible, it did not happen, mainly because of the
-difficulty in carrying over other experiments to volunteers when the
-experiment they were participating on was finished. This has been shown in the low
-correlation between the number of IPs in successive experiments, and
-also resulted in a low number of generations allotted by users.
-
-In the second implementation, Web Workers
-were used so that several clients per browser page could be executed and
-run in the background; our intention was to increase the time spent by
-the users and, in fact, the number of generations per
-user increased to the point that we obtained a performance better than
-baseline in most cases. The architectural changes led to a high correlation
-between successive experiments, so that, even as the median number of
-volunteers per experiment decreased, they were probably contributing
-to the experiment simultaneously, so that our objective was
-achieved. In general, one of the first conclusions of this paper is
-that the browser technologies should be used to its full extent, so
-that the user time is leveraged to its full potential, and this was
-achieved using web workers.
-
-The second objective of this paper was to model the user behavior in a
-first attempt to try and predict performance. As should be expected,
-the model and its parameters depend on the implementation, with contributions following
-a GEV distribution with single-thread and a
-Weibull distribution when web workers are used. These distributions are
-close enough to each other and, in the second case, reflect the fact
-that most time spent in the page is actually devoted to computing, as
-opposed to presenting the chart or not doing anything,
-which is why the time spent (represented by the number of
-contributions to the pool) follows a model quite similar to that found
-for games or other online activities. The reverse implication, from
-games to volunteer computing,  might also be true: if we
-want to have returning users for the experiments, it is probable that
-we should {\em gamify} the experience so that once they have participated
-once, they would be encouraged to do for a longer time or to share
-it. In the spirit of Open Science, this 
-gamification might involve computing in real-time data such as the one
-presented in this paper, while showing it in the same page.
-
-In general, linking and finding correlations between user choices and
-performance is an interesting avenue to explore in the future. % (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
-% Moved to conclusions. 
-There are several validity threats to the design of this study.
-Participants of the experiments were mainly members of our 
-own social networks, personal colleagues, friends 
-and acquaintances, therefore, to extend this work other options 
-should be considered in order to include 
-a wider audience with different motives for participation.
-Another threat is that participants were anonymous, so the amount of repeated 
-participations and the effects this might have had in the outcome is not
-known. Another issue is that the scalability of the system was not thoroughly tested
-but for the amount of participants the basic configuration 
-was more than adequate, then again more experiments are needed in order
-to assess the capability both with more clients and other server
-configurations. Although care was taken to include representative benchmarks
-of real world problems, more experiments with a broader coverage 
-should allow a better understanding of the practical limitations of
-the architecture and the tools chosen.
-%  Even if
-% the three previous experiments were published in a similar way, one
-% obtained up to 5 times more total cycles  than the one with the least
-% number of cycles. 
-It is also essential to obtain volunteers as
-simultaneously as possible, so it is possible that the features of the
-social network in terms of real-time use will also play a big
-role. Even as it is difficult to create controlled experiments in this
-area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation
-
-This paper also shows that the time needed to find a solution or if it
-is eventually found have an influence on the time people devote to
-it. It would be so interesting to check the performance for other computationally demanding
-tasks, especially ones that cover the middle ground between a problem
-finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
-Besides, the two versions of the framework include as an
-algorithmic variant using random population size. We do not think this
-has had a big influence on the results and in fact this is not noticed
-for F15, which is the last version tested, but it would be interesting to
-measure exactly what this influence has been. 
-% In general there are
-% many issues with the evolutionary algorithm implementation itself,
-% including using different, or adaptive, policies for inserting and
-% sending individuals to the pool as was done in \cite{araujo2008mam},
-% using different policies for population initialization, and also the
-% incorporation of high-speed local resources to the pool to check what
-% would be the real influence of the volunteer pool to the final
-% performance.
-
-Other improvements could include adding a bit of more
-control to the user might contribute also to gamification. Right now
-the user has only two Web Workers. It is a matter of a single click to
-open more tabs in the browser, giving more Web Workers, but it would be
-interesting to put everything in a single page and under our control,
-to check how often this happens and under which conditions.
-
-% Finally, the implementation needs some refinement in terms of
-% programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
-% to create a generator in which the user just has to create a fitness
-% function, with the rest of the framework wrapped around
-% automatically.
-%  All this, as well as the data used for this paper and the paper
-% itself, has been published with a free license in GitHub at
-% \url{https://github.com/JJ/modeling-volunteer-computing}.  
-% Pablo: I should uncomment this last paragraph, but using the url: https://github.com/ANONYMOUS
-
-%---------------------------------------------------------------
-% \section*{Acknowledgment}
-% 
-% This work has been supported in part by
-% TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
-% SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
-% GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
-% Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
-% We would also like to thank the
-% anonymous reviewers of previous versions of this paper who have really
-% helped us to improve 
-% this paper (and our work) with their suggestions. We are also grateful
-% to Anna S\'aez de Tejada for her help with the data processing scripts.
+This work has been supported in part by
+TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
+SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
+GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
+Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
+We would also like to thank the
+anonymous reviewers of previous versions of this paper who have really
+helped us to improve 
+this paper (and our work) with their suggestions. We are also grateful
+to Anna S\'aez de Tejada for her help with the data processing scripts.
 
@@
 \end{document}
+
 %%% Local Variables:

### 2016-03-18 PPSN2016/paper.Rnw

@@
 \begin{abstract}
-In a volunteer distributed computing system, users run a program in
+In a volunteer distributed computing system, users run a program on
 their own machine to contribute to a common effort. If the program is
 embedded in a web page, collaboration is easy, but also ephemeral. In
-this paper we analyze a volunteer evolutionary computing system called
+this paper, we analyse a volunteer evolutionary computing system called
 NodIO, trying to find out a few rules that encourage volunteer
@@
 volunteer usage patterns and tweaks in the algorithm contribute to
+%Pablo: añadiría un ", such as...," con un ejemplo entre "algorithm" y "contribute"
 improvements and leveraging of these contributions. 

### 2016-03-19 PPSN2016/paper.Rnw
Back to 10 pages
Please @mariosky @fergunet @unintendedbear check that out when making corrections

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 \url{http://nodio-jmerelo.rhcloud.com}, a free resource hosted in the
-OpenShift Platform as a Service.\footnote{It is not guaranteed to be running,
-or running the same version, when you read this, however; you can always get
-the sources from GitHub and set it up yourself.}
+OpenShift Platform as a Service.
+%% \footnote{It is not guaranteed to be running,
+%% or running the same version, when you read this, however; you can always get
+%% the sources from GitHub and set it up yourself.}
 %
@@
 fast clients there might be some clients that do not detect that
-change and might, in fact continue with their former run, thus not
+change and might, in fact, continue with their former run, thus not
 contributing to the current common cache.
 
-In fact, as can be seen in Table \ref{tab:correlations}, which shows
-the correlation between the number of IPs from one minute to the next,
-the behavior of all experiments but the last one are similar, with a
-very high correlation indicating that the number of volunteers does
-not change from one minute to the next. However, there is a sharp drop
-for the {\tt cachereboot} experiment, which indicates a bigger
+In fact, Table \ref{tab:correlation}, which shows
+the correlation between the number of IPs from one minute to the
+next, presents a similar  behavior for all experiments but the last one: a
+very high correlation indicating that the number of volunteers changes
+slowly. The {\tt cachereboot} experiment correlation indicates a bigger
 dynamism, with individuals joining and leaving the experiment all the
-time. This might be circumstancial, but it also reflects the fact
+time. This might be circumstancial, but it also informs the fact
 that the solution is found in less time and that users need less
-contributions to find it personally.
+contributions to find it.
 

### 2016-03-19 PPSN2016/paper.Rnw
Very little changes in section 4

@@
 
-A summary of the experiments and  its results is shown in Table
+A summary of the experiments and  their results is shown in Table
 \ref{tab:runs}. This table shows the median time needed to find the
 solution as well as the median number of IPs and PUTs. In this case we
-distinguish between the total number of clients participating in that
+distinguish between the total number of clients participating in every
 particular run, and the {\em actual} number whose contributions were
@@
 conveyed by inserting a graph in the client that represented the size
-of the cache and labeled ``How am I contributing''. The user could
+of the cache, labeled ``How am I contributing''. The user could
 then realize if his contribution was being 0 and thus do something

### 2016-03-20 PPSN2016/paper.Rnw
Fixes latex
Fixes URL and typos

@@
 embedded in a web page, collaboration is easy, but also ephemeral. In
-this paper, we analyse a volunteer evolutionary computing system called
+this paper, we analyze a volunteer evolutionary computing system called
 NodIO, trying to find out a few rules that encourage volunteer
@@
 participation and contribution to the algorithm and how different
-volunteer usage patterns and tweaks in the algorithm contribute to
-%Pablo: añadiría un ", such as...," con un ejemplo entre "algorithm" y "contribute"
+volunteer usage patterns and tweaks in the algorithm, such as restarting clients when solution has been found, contribute to
 improvements and leveraging of these contributions. 
@@
 
-The world wide web not only provides a platform for content
+The world wide web  provides not only a platform for content
 distribution, but also an increasingly reliable and high-performance
-platform for running distributed applications. This is based in two
-key facts: the Javascript virtual machine every browser runs and a
+operating system for running distributed applications. This is based in two
+key things: the Javascript virtual machine every browser runs and a
 simplified standard interface for interacting with servers.
 % These are "things", not "facts". As facts would be... "every browser runs a Javascript virtual machine and that there is a simplified standard interface for interacting with servers (?)
+% changed to things - JJ
 Thus
@@
 processing extraterrestrial signals \cite{david-seti:home}. However
-the dual introduction of JavaScript as a universal language for the
+the dual facts of the introduction of JavaScript as a universal language for the
 browser and the browser itself as an ubiquitous web and Internet client has
@@
 al. \cite{churn08:ijhpsa}, using models, try to find out the resulting
-performance when the users are changing continously according to a
+performance when the users are changing continuously according to a
 Weibull distribution. In the case of Klein and Spector \cite{unwitting-ec}, the algorithm
@@
 have been made in Spanish, but in some cases English was used
-too. This makes for a certainly unrealiable experimentation framework,
+too. This makes for a certainly unreliable experimentation framework,
 but it is a realistic one. The steps in the design of the evolutionary
@@
   plotted in a log $y$ scale.\label{fig:zipf}}
-\label{fig:zipf}
 \end{figure}
@@
 case. By looking at the logs in the last case we have established that
-minimun as 16 PUTs; in less than 1600 generations no solution to that
+minimum as 16 PUTs; in less than 1600 generations no solution to that
 difficult problem will be found. Let us filter then the runs,
@@
 dynamism, with individuals joining and leaving the experiment all the
-time. This might be circumstancial, but it also informs the fact
+time. This might be circumstantial, but it also informs the fact
 that the solution is found in less time and that users need less
@@
 be relaxed. Some other policies could also be tested, as well as
-running heterogenous algorithms in the same way that was done in
+running heterogeneous algorithms in the same way that was done in
 \cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other
-evolutionary algorith, seems to be the key, but there are many ways of
+evolutionary algorithm, seems to be the key, but there are many ways of
 approaching it.
@@
 the data and processing scripts, are published with a free license in GitHub at
-\url{https://github.com/JJ/modeling-volunteer-computing}.
+\url{https://git.io/gecco-es-15}.
 
@@
 Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-2015 UGR) . We would also like to thank the anonymous reviewers of
-previous versions of this paper who have really helped us to improve
-this paper (and our work) with their suggestions. We are also grateful
-to Anna S\'aez de Tejada for her help with the data processing
-scripts. We are also grateful to {\tt @otisdriftwood} for his help
+2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
 gathering users for the new experiments.

### 2016-03-20 PPSN2016/paper.Rnw
Back to 10 pages

@@
 here, however, is to use it as a resource for evolutionary
-computation. This has been a line of research in our group for a long
-time \cite{jj-ppsn98}, but has lately been revisited thanks to the
-ubiquity of social networks and handheld computing devices.
+computation, as our group has done for a long time \cite{jj-ppsn98}.
 
@@
 % Is it ok that we say that the first line of research is the most important... but then we focus on the second?
-In this paper our focus will be in the second aspect: we will try to
+We have approached the first issue in our previous work,
+% I have justified to focus, in this case, on the second
+but in this paper our focus will be in the second aspect: we will try to
 design a decentralized system that, at the same time, is able to use
-all available resources. This design will be done incrementally by
-changing the system on the client and the server and measuring its
+all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
+changing client and the server and measuring its
 impact on the overall performance: time and evaluations needed to find
@@
 time that the evolutionary algorithm itself makes the most of those
-contributions, independently of how many they are, and is able to find
-the solution to the problem in less time and with less contributions.
+contributions and is able to find
+the solution to the problem in a minimum time and with the least
+number of contributions.
 
@@
   server.  The server has the capability to
-  run a single experiment, storing the incoming chromosomes in a cache data structure
-  that is reset when the solution is found. This cache has a finite
-  size that erases the oldest chromosomes once it has reached its % Also "it has been filled to its"
+  run a single experiment, storing the incoming chromosomes in a
+  key-value cache that uses as key the chromosome string itself and is
+  reset when the solution is found. This cache has a finite 
+  size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
   capacity.  
@@
 \url{http://nodio-jmerelo.rhcloud.com}, a free resource hosted in the
-OpenShift Platform as a Service.
-%% \footnote{It is not guaranteed to be running,
-%% or running the same version, when you read this, however; you can always get
-%% the sources from GitHub and set it up yourself.}
+OpenShift Platform as a Service\footnote{It is not guaranteed to be running,
+or running the same version, when you read this, however; you can always get
+the sources from GitHub and set it up yourself.}.
 %
@@
 NodIO needs a fitness function to work with; in this case the classical Trap 
-function \cite{Ackley1987} has been used; % I don't think this use of the ; is correct
-this function runs in the
+function \cite{Ackley1987} has been used. % I don't think this use of the ; is correct
+% Changed to .
+The evolutionary algorithm that evaluates that function runs in the
 browser, but the server needs to know when the solution has been
-found, so a termination condition is also included in the server, as a configuration
+found, so a function that checks termination condition is also included in the server, as a configuration
 option. In all cases we are using 50 Traps, a problem
@@
 an extended amount of minutes.
-
 For every new individual sent by the clients to the cache, several
@@
 % Im changing over there to cache
-pieces of meta data were stored: a time stamp, the client IP, the chromosome and
+% That's been changed - JJ
+pieces of metadata were stored: a time stamp, the client IP, the chromosome and
 fitness value, the cache size in that particular moment, and also if that
@@
 number. The final column includes the total number of runs in that
-particular batch, finished after a minimum number of 30 was reached.
+particular batch, finished roughly after a minimum number of 30 was reached.
 
@@
 \begin{table}
-  \caption{Experiments with overlaps eliminated:  summary of results. \label{tab:filtered}}
+  \caption{Experiments with {\em overlapped} runs filtered out: summary of results. \label{tab:filtered}}
 \begin{center}

### 2016-03-20 PPSN2016/paper.Rnw
Back to 11 pages after fixes

@@
 system as long as it can run node.js and has access to the filesystem, for
-log files, if available. All these logs are available upon
-request. Processed files are available in the same repository that
-hosts this paper, as well as the scripts needed to process them.
+log files, if available%% . All these logs are available upon
+%% request. Processed files are available in the same repository that
+%% hosts this paper, as well as the scripts needed to process them.
 
@@
 log-log scale. Besides following the usual Zipf's law found in our
-previous work and having a very similar appearance for all 5 sets, it
+previous work and thus having a very similar appearance for all 5 sets, it
 is interesting to see how some sets have a number of contributions per
@@
 ``Contributing'' column includes a value that is less than the total
-number. The final column includes the total number of runs in that
+number. The next column includes the total number of runs in that
 particular batch, finished roughly after a minimum number of 30 was reached.
@@
 server, so they can in fact proceed with the algorithm and even finish
-it. The fact that these users were not contributing to the pool was
+it. However, this was not known to the volunteer, so the fact that they were not contributing to the pool was
 conveyed by inserting a graph in the client that represented the size
 of the cache, labeled ``How am I contributing''. The user could
-then realize if his contribution was being 0 and thus do something
-about it; these are labeled {\tt graph1} and 2. However, a stuck
-client could still add to the pool if evolution got unstuck, so we
+then realize there was no contribution and do something
+about it. These sets of runs are labeled {\tt graph1} and 2. However, a stuck
+client could still add to the pool if it found a new optimum, so we
 added a mechanism for rebooting the client, that is, reloading the
 page, if the cache size was found to be less than 1, which indicated a
-run recently initiated. This batch of runs is labelled {\tt
-  reboot}. However, we found that this size of the cache was actually
-skipped by some slow clients or too fast evolutions, so that in the
+run recently started. This batch of runs is labelled {\tt
+  reboot}. This size of the cache was not really noticed by some slow clients or too fast runs, so that in the
 last case, labeled {\tt cachecrash}, clients rebooted if the cache
@@
 %
-So we will have to check whether we have effectively eliminated these
+The first thin we will have to check is whether we have effectively eliminated these
 overlaps. In order to do that, we will have to establish a threshold

### 2016-03-20 PPSN2016/paper.Rnw
Back to 10 pages

@@
 \begin{table}[htb]
-\caption{Experiment table, with summary of results. \label{tab:runs}}
+\caption{Summary of results for the 5 sets of experiments.\label{tab:runs}}
 \begin{center}
@@
 %
-\caption{Ranked and normalized number of contributions (PUTs) per IP,
-  plotted in a log $y$ scale.\label{fig:zipf}}
+\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
 \end{figure}
@@
 fact, as it can be seen in the last column, the median number of
-contributions per user is {\em much lower} in the last row, which
-indicates that {\tt cachereboot} needs less contributions by users to
+contributions per user is the lowest for {\tt cachereboot}, which so needs less contributions by users to
 find the same solution. Besides, the last row shows that {\em all}
@@
 strategy 30\% of users do not. This might be due to the fact that {\tt
-  reboot} only does so when cache size is less than one. With very
-fast clients there might be some clients that do not detect that
-change and might, in fact, continue with their former run, thus not
-contributing to the current common cache.
+  reboot} only does so when cache size is less than one, indicating
+that this strategy is only partly successful in avoiding overlaps.
 

### 2016-03-20 PPSN2016/paper.Rnw
Back to 10 pages

@@
 the correlation between the number of IPs from one minute to the
-next, presents a similar  behavior for all experiments but the last one: a
-very high correlation indicating that the number of volunteers changes
-slowly. The {\tt cachereboot} experiment correlation indicates a bigger
-dynamism, with individuals joining and leaving the experiment all the
-time. This might be circumstantial, but it also informs the fact
-that the solution is found in less time and that users need less
-contributions to find it.
+next, presents a similar  behavior for all experiments but the last one. A high correlation indicating that the number of volunteers changes
+slowly from one minute to the next; however, the {\tt cachereboot} experiment correlation indicates a bigger
+dynamism, with volunteers joining and leaving the experiment all the
+time, maybe because experiments take less time and need less resources.
 

### 2016-03-20 PPSN2016/paper.Rnw
Back to 10 pages by reducing conclusions

@@
 number of PUTs, which reflects the number of evaluations performed by
-the clients. The two last rows have a low number of evaluations,
+the clients. The last two rows have a low number of evaluations,
 around 650, which is higher than the first two, probably due to the
@@
 time to find the solution {\em using the same number of
-  evaluations}. All variables are difficult to control in this
+  evaluations}; the same situation applies if we compare {\tt graph1}
+to {\tt cachereboot}. All variables are difficult to control in this
 environment, but it might be due to the fact that parallelism has
@@
 
-Our main intention in this paper was to focus on the algorithm itself
-so that it could became as effective as possible, and would be able to use
-computations by volunteers in the best possible way, or at least better
-than the initial configuration. This initial configuration included
+Our main intention in this paper was to configure an evolutionary
+algorithm in a volunteer computing environment to improve its
+efficiency. This initial configuration included
 {\em overlaps} or clients running islands that had actually started in

### 2016-04-01 data/process-time.pl
Erasing old processing scripts
Closes #3

@@ -1,32 +1,0 @@
-#!/usr/bin/env perl
-
-use File::Slurp::Tiny qw(read_file);
-use JSON;
-use v5.14;
-
-my $file_name = shift || "log";
-
-my $log_json = read_file( $file_name ) || die "Can't read file $file_name\n";
-
-my $log = decode_json $log_json;
-
-my %IPs;
-
-my @times;
-for my $l (@$log ) {
-  if ( $l->{'put'} ) {
-    push @times, $l->{'put'}[0];
-  } else {
-    push @times, $l->{'get'}[0];
-  }
-  if ( $l->{'put'} ){
-    $IPs{$l->{'IP'}}++
-  };
-}
-
-my $total_time = $times[$#times]-$times[0];
-my $total_ips = keys(%IPs);
-
-
-say "$total_time;$total_ips";
-

### 2016-04-03 EvoSoft16/nodio-evosoft16.tex
Making small changes
Well, time for this one
I wasn't going to do it, but this is published as a TechRep and I think we should give it a try for EvoSoft.

Please @unintendedbear @fergunet @pacv @mariosky give it a shot more or less now. I'll submit a first version anyways and send another version when it is ready. I have to update it with the latest changes to the framework, for instance.

@@
-% This is "sig-alternate.tex" V2.1 April 2013
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
 \documentclass{sig-alternate}
@@
 
-\title{NodIO: a framework for pool-based evolutionary computation}
+\title{NodIO: a framework and architecture for pool-based evolutionary
+  computation} 
 %
@@
 \alignauthor
-Juan-J.~Merelo, Pedro A. Castillo,Pablo Garc\'ia-S\'anchez,Paloma de las Cuevas\\
+Juan-J.~Merelo, Pedro A. Castillo, Pablo Garc\'ia-S\'anchez, Paloma de las Cuevas\\
 \affaddr{Dept. of Computer Architecture and Technology
@@
 \begin{abstract}
-
+JavaScript is an interpreted language mainly known for its inclusion
+in web browsers, making them a container for rich Internet based
+applications. This has inspired its use, for a long time, as a tool
+for evolutionary algorithms, mainly so in browser-based volunteer
+computing environments. Several libraries have also been published so
+far and are in use. However, the last years have seen a resurgence of
+interest in the language, becoming one of the most popular and thus
+spawning the improvement of its implementations, which are now the
+foundation of many new client-server applications. We present such an
+application for running distributed volunteer-based evolutionary
+algorithm experiments, and we make a series of measurements to
+establish the speed of JavaScript in evolutionary algorithms that can
+serve as a baseline for comparison with other distributed computing
+experiments. These experiments use different integer and floating
+point problems, and prove that the speed of JavaScript is actually
+competitive with other languages commonly used by the evolutionary
+algorithm practitioner.
 \end{abstract}
@@
 
-So, the rest of the paper is organized as follows: next we present the
-state of the art in web-based distributed
-computational systems along with attempts to predict and model their
-behavior in Section \ref{sec:soa}. A brief presentation of the
+The need to use non-permanent and potentially massive computing
+infrastructure for distributed scientific experiments has lead to the
+rising of {\em volunteer computing} defined as a set of tools that allow
+citizens to {\em donate} cycles to scientific experiments by running
+an application \cite{Anderson06thecomputational}; this application can be embedded in a web page. And
+since browsers are ubiquitous, we are more interested in these kind of systems, this has led us to create a framework for
+evolutionary algorithm experiments that is able to work in that mode
+\cite{DBLP:journals/corr/GuervosG15}. {\sf NodIO} is a
+cloud or bare metal based volunteer evolutionary computing system
+derived from the {\sf NodEO} library, whose architecture has been
+developed using JavaScript on the client as well as the server.
+All parts of the framework are free and available with a free license
+from \url{https://github.com/JJ/splash-volunteer}.
+
+%%% Issues with volunteer based systems
+
+As a volunteer based system, there are some issues that have to be addressed 
+\cite{sarmenta2001volunteer,web:BOINC} :
+\begin{enumerate}
+\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
+each request is known.
+\item As anonymous entities, volunteers are not accountable. 
+If a volunteer misbehaves in some way the provider cannot 
+prosecute or discipline the volunteer. In the case of JavaScript,
+this issue is aggravated because even if it is obfuscated, source code
+and data is sent to clients. 
+\item Participants must also trust the application provider, 
+regarding invasion of privacy, the integrity of their devices, 
+and intended application. 
+\end{enumerate}
+
+Several kind of threats are possible in this environment; first, since its
+programming interface is open and its code is open source, it is
+relatively easy to find vulnerabilities and 
+sabotage the system, as indicated in \cite{domingues2007sabotage}, by
+crafting a fake request which, for instance, assigns a fake fitness to
+a particular chromosome.
+Other two different kinds of attacks are also
+possible: a denial of service attack as well as penetration
+testing. We have approached this threat in a pragmatic way: all
+sources, client and server, for the application, are openly released
+as free software. 
+ We also released as an open access paper the result of the 
+ experiments performed \cite{DBLP:journals/corr/GuervosG15}  
+ and all data, once anonymized, as open data. This
+builds trust between the user and the scientist, which indeed has an
+impact on the performance of the system, since there is no need to
+include cheating checks or other functions that would degrade it. This
+socio-technical approach is, indeed, 
+coherent with our whole approach to the study of a distributed volunteer
+ system which can only be tackled by approaching it from the
+technical as well as the social, as in social media, angle.
+
+% Justification for JavaScript
+In general, a distributed volunteer-based evolutionary computation
+framework based on the browser such as NodIO is simply a client-server system
+whose client is, or can be, embedded in the browser via
+JavaScript. Since JavaScript is  the only language that is present
+across all browsers, the choice was quite clear. Even so, there might
+be some issues with the performance of the language itself, which 
+ is
+why we have made a comparison between JavaScript and other languages
+\cite{2015arXiv151101088M}. In that paper we prove that JavaScript can be
+faster than compiled languages such as Scala and, in any case, it provides
+a performance that is comparable to other languages usually employed % (Paloma) Adverb
+in evolutionary computation and numerical optimization such as Python. 
+Compiled languages such as
+Java or C might be faster when 
+considering the performance of a single-user, 
+but this is more than offset by the computational power of
+the spontaneous volunteers we can gather at whim, including people
+using their mobile phones or tablets. Together, the performance is several orders of magnitude
+higher, which is the objective in this kind of systems.
+
+Since JavaScript is
+quite a popular language nowadays, it also provides compilers from a number
+of languages, many of them in their family (like CoffeeScript), but
+many other outside, so researchers can, in fact, write their fitness
+function in Python, Erlang, Scala, or even Java, and then compile it to
+JavaScript. For completeness, an experiment comparing the performance of
+a JavaScript implementation versus Java and Matlab is included in section
+\ref{sec:experiments}. We will first describe the {\sf NodIO} architecture in
+the next Section.
+
+So, the rest of the paper is organized as follows: A brief presentation of the
 methodology used and the results of the experiments are presented in
@@
 %---------------------------------------------------------------
-\section{State of the art}
-\label{sec:soa}
+%\section{State of the art}
+%\label{sec:soa}
 
@@
 
+The {\sf NodIO} architecture has two tiers. 
+The figure \ref{fig:system} describes the general system architecture and
+algorithm behavior. Different web technologies, such as JQuery or {\tt
+  Chart.js} have
+been used to build the user interface elements of the framework.
+%
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=4in]{system.pdf}
+\caption{Description of the proposed framework. Clients execute a JavaScript EA
+  in the browser, which, every 100 generations, sends the best
+  individual and receives a random one from the server.}
+\label{fig:system}
+\end{figure*}
+
+\begin{enumerate}
+\item A REST server, that is, a server that includes several {\em
+  routes} 
+  that can be called for storing and retrieving information (the ``CRUD'' cycle:
+  create, request, update, and delete) from the server. 
+  A JSON data format is used for the communication between 
+  clients and the server. There are two kinds of information:
+  {\em problem} based, that is, related to the
+  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
+  GET}ing a random chromosome from it, and {\em information} related
+  to the performance and state of the experiments. It also performs logging
+  duties, but they are basically a very lightweight and high performance
+  data storage \cite{jj:idc:lowcost}.
+  The server has the capability to
+  run a single experiment, storing the chromosomes in a data structure
+  that is reset when the solution is found.
+
+\item A client that includes the evolutionary algorithm as
+  JavaScript code embedded in a web page that displays graphs, some
+  additional links, and information on the experiment. This code runs
+  an evolutionary algorithm {\em island} starting with a random
+  population, then it sends, every 100 generations, the best individual
+  back to the server (via a {\tt PUT} request), and requests a random
+  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
+\end{enumerate}
+
+  This is the invariant part of the framework, but other than that,
+  the algorithm can be run in many different ways: 
+  stopping when a solution is found or using Web Workers. In fact, since it is a pool-based system such as
+  the one described in \cite{LNCS86720702}, any kind of client that
+  calls the application programming interface (API) can be used,
+  written in any kind of language. However, in this paper we are going
+  to focus on the dynamics and measurements of the volunteer part of
+  the framework.
+
+The researcher only has to change a function, which can be written in
+JavaScript or in any other language with cross-compilation to
+JavaScript \cite{web:compilersjs} to solve different optimization
+problems.
+
+Let us see how this system addresses the different challenges
+outlined above:
+\begin{itemize}
+\item {\em Scalability} is provided via the use of a lightweight and
+  high-performance, single-threaded, server based in Node.js and
+  Express.js. Although this single server is a bottleneck since it
+  will eventually saturate, the fact that it runs as a non-blocking single thread
+  allows the service of many requests. In fact, a limit in the
+  number of simultaneous requests will be reached, but so far it has
+  not been found, unlike what we found in our previous systems, DCoR,
+  \cite{gecco07:workshop:dcor}, which had a low scalability. 
+\item The framework is {\em heterogeneous} since it does not need any
+  performance, operating system, or even browser requirement as long
+  as JavaScript is enabled: anyone
+  visiting the page, even from mobile devices, can load the algorithm.
+  In case the browser does not support HTML5 Web workers (e.g., 
+  Windows Phone or some version of Android) or their use
+  has been disabled, a basic version of {\sf NodIO} can also be used.
+\item {\em Fault tolerance} is always an issue, and in this case, the
+  single point of failure would be the server: the system, as such,
+  would break down if the server fails. However, the individual
+  islands in every browser would continue running, and having access
+  to just one of them would allow the local algorithm to proceed. In
+  fact, the island does not need the server to run: it runs locally if
+  needed, with the only exception that it is obviously unable to
+  communicate with the rest of the islands.
+\item {\em Adaptiveness} is achieved simply through the autonomous
+  operation of every individual island without any synchronization
+  mechanism. The islands in the system are, in fact, unaware of each
+  other, communicating only through the server. This architecture 
+  has been tested in other high demand systems with success, but 
+  in the case of {\sf NodIO}, additional experiments are needed to asses 
+  the scalability of the communication system. % (Paloma) Is it ok if
+                                % we talk about scalability when we
+                                % were talking about adaptiveness?
+                                % Scalability was also described in
+                                % the first point. 
+  %(Mario) Scalability of the communication system in particular?
+\item Since the algorithm runs in the browser, user's {\em safety} is
+  achieved through its sandbox mechanisms. The user is thus assured
+  that there is no unsafe access either to their local files or even
+  to more resources than the browser should be allotted.
+\item Running the algorithm is just a matter of loading the page,
+  which makes the operation totally {\em anonymous}. For the same
+  reason, {\em ease of use} is optimal, being as easy as simply
+  clicking on an URL, available to anyone with access to a browser.
+\item {\em Reasonable performance} is not ensured. In fact, we should
+  make sure that there is a reasonable amount of clients over which
+  the performance achieved is better than what you would obtain in
+  your own desktop system. If this is not the case, it is a pointless
+  academic exercise.
+\end{itemize}
+
+There are many different ways to validate a framework that intends to
+address these challenges. Some of them are included by design: it is
+anonymous, since no accounts are needed to participate in the
+experiment, it is safe for the user, since we are using the browser
+black box, and it is also heterogeneous. We have designed the
+experiments so that they show adaptability to different types of
+experiments and clients, scalability with the number of clients and
+the problem size, and measured performance by comparing it with
+a baseline configuration, with which, as indicated above, is
+compatible and can be used in a complementary way. Fault tolerance
+will be experimented by dropping the server and checking the continuity
+of the experiments.
+
+After the initial experiments \cite{DBLP:journals/corr/GuervosG15}, we
+realized that we needed to make some improvements. So, in order to
+improve the number of cycles per user, we enhanced the 
+basic architecture in several different aspects. One proposed improvement
+involved restarting the client once the solution was
+found, so that as long as the user was visiting the page it was
+running an evolutionary algorithm. Another was a small algorithmic
+enhancement: instead of having a fixed population,
+population size was randomly distributed between 128 and 256 and
+different for each client. We called this instance of
+the architecture NodIO-W$^2$. The other improvement was to 
+use the Web Worker API as described next.
+
+The rationale for this change comes from the realization that JavaScript has been commonly used to develop client-side scripts
+that handle the user interface, communicating asynchronously with servers and
+updating the content that is displayed \cite{flanagan2006javascript}.
+Long-running scripts like those proposed in this work can interfere with the
+responsiveness of the web application, because everything is handled by a
+single thread. The Web Workers specification \cite{hickson2012web} defines an
+API that allows Web developers to execute long-running scripts in parallel
+with the scripts of the main page. Worker scripts are spawned in the
+background allowing a thread-like operation using message-passing as the
+communication mechanism between the worker's execution environment and the
+main thread. According to the Web Workers W3C Candidate Recommendation
+\cite{hickson2012web} workers are expected to be long-lived, they have a high
+start-up performance cost, and a high per-instance memory cost. The Web Worker
+API is implemented in current versions of both desktop and mobile web browsers.
+For volunteer computing using the Web Worker API brings several advantages
+over a single threaded implementation:
+
+\begin{itemize}
+\item If the browser uses a tabbed document interface, the worker script
+keeps running in the background even in case the user brings forward another tabbed
+window.
+\item Several evolutionary algorithms can be executed in parallel in a single web
+page. 
+\item Many implementations of the Web Worker API can use multiple-core CPUs for
+the execution of worker scripts, in particular the execution in multiple-core CPUs
+was tested in desktop versions of Chrome, FireFox and Safari.
+\item Each {\sf NodEO} instance can be restarted independently.
+\end{itemize}
+
+A sequence diagram of NodIO-W$^2$, is presented
+in Figure~\ref{fig:system:w2}. In this version, clients use the Web Workers
+API to run two {\sf NodEO} instances per browser. The sequence diagram shows the
+interaction between processes in a time sequence. In this version there are
+two kinds of asynchronous messages: HTTP Requests shown as bold arrows between
+the client and the server where PUT and GET HTTP methods are represented by
+red and blue arrows respectively. There are also messages between workers and
+the main thread locally in the client, which are shown as thin solid arrows.
+There are three processes involved:
+\begin{itemize}
+\item {\em Node.js} This is the server side process which is basically
+  the same as used previously, except for some minor log-related
+  changes. 
+\item {\em Client main thread} The main script creates
+worker instances and handles the user interface.
+\item {\em Worker global scope} The worker execution space does not have
+access to the Document Object Model (DOM) or JQuery objects in the main thread.
+Asynchronous HTTP requests to the server are implemented using the
+XMLHttpRequest object.  
+\end{itemize}
+%
+\begin{figure*}[!htb]
+\centering
+\includegraphics[width=4in]{Algorithm.pdf}
+\caption{Description of the W$^2$ version of {\sf NodIO}. In this
+  version, clients use Web Workers to run the evolutionary algorithm,
+  with two of them per browser.}
+\label{fig:system:w2}
+\end{figure*}
+
+The sequence of events happens as follows:
+\begin{enumerate}
+\item First the {\sf NodIO} process is started in Node.js. As part of the
+initialization procedure, the sequential number of experiments is initialized and a log
+file is created. The shared pool implemented as an array is reset.
+\item A volunteer follows the link of the experiment triggering a GET HTTP
+request, then the page is loaded.
+\item The main script renders the page and creates the worker instance. In
+order to create a dedicated worker instance, another GET request is sent to the
+server, then the client receives the second worker implementation script.
+Each worker instance executes over time one or more evolutionary algorithms
+{\em islands}. Each island starts with a random population and it is assigned a
+universally unique identifier (UUID). The UUID is included in each of the HTTP
+requests sent to the server.
+The initial parameters for each island
+can be distinct. Once the worker is created and the parameters have been set,
+a message is sent to the worker to start the algorithm.
+\item The worker first creates the population following the previously
+specified parameters. Once the population has been initialized, the EA
+generation loop is started. After {\em n} generations the state of the population
+is evaluated. The chromosome with the best fitness is sent to the shared
+population with a PUT HTTP request. If the best fitness is not the global best,
+the EA loop must continue, but first a chromosome is randomly selected from
+the shared population. The current generation number and the best
+chromosome are also sent to the client's main thread in order to update the DOM to
+display the current state of the island (described next). These messages are
+all asynchronous, so the EA loop of
+the worker continues even before a message is received by the target. If the best fitness is a global best, the best
+chromosome is sent to the server and the loop ends. A message is sent to the
+main thread indicating that the best individual has been found.
+\item The main thread receives messages from workers as events handled by
+a callback function. When an iteration message is received, the page is
+updated by displaying a dynamic plot of the generation number and best
+fitness, and a representation of the chromosome is also generated.
+\item When a global best is received from an island, the current experiment
+ends, the experiment number is incremented, and the population array is reset.
+\item When the main thread receives a message indicating that a worker has ended the EA
+loop, that worker is reinitialized. The worker process is not ended, to
+avoid the start-up cost, only the parameters and population are reset and a new UUID
+assigned.
+\end{enumerate}
+%
+
+These two versions of the framework will have to be tested {\em in the
+  wild} to measure its performance. 
+However, before testing this framework in a real experiment, we have
+to establish a baseline for comparison.
+
+\section{Baseline Experiments} 
+\label{sec:experiments}
+
+We have used the classical Trap function \cite{Ackley1987}
+used as well as the Rastrigin's floating-point optimization
+problem. Since JavaScript is a functional language and declaring a different 
+function and handing at the creation of the algorithm object, called
+{\tt Classic}, is the only change needed to work with a different
+problem.
+
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=10cm]{baseline-times.png}
+\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
+    Node.js} client, for different population sizes. Only
+runs in which solution has been found are used for this graph.}
+% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
+\label{fig:baseline}
+\end{figure*}
+First, we establish the baseline performance by running fifty runs of the
+evolutionary algorithm in a desktop client written using {\sf NodEO}
+\cite{nodeo2014}, the basic JavaScript library we have used as base to
+build {\sf NodIO}, the framework. This experiment tries to 
+find the solution to the 40-trap function with parameters $l=4, a=1,
+b=2, z=3$ and a population size equal to 512. The algorithm in a
+single island was run until the solution (a string with all
+ones) was found or five million evaluations had been performed. It
+took around a minute, on average, that is $68.9694$ seconds, to
+perform these runs. In this experiment only 33, that is, 66\% were
+successful. The experiments were repeated for population $p=1024$,
+with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
+seconds. Results for these two experiments
+are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
+characteristics: 
+\begin{verbatim}
+Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
+2014 x86_64 x86_64 x86_64 GNU/Linux
+\end{verbatim}
+with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.
+
+These two experiments establish a baseline result and show that
+the time to success depends on the population size, with a bigger population
+contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+experiments that we will describe next do not and can not have the
+same conditions, but
+the baseline is that if they eventually take longer than a basic
+desktop, their interest will be purely academic.
+
+\subsection{Tests with a floating-point hard optimization problem}
+
+JavaScript has been traditionally implemented as an interpreted
+language not designed for the development of high performance
+systems. On the other hand, current JavaScript  
+virtual machines (VMs from now on) are closing the gap, for instance the Google V8 
+is an engine specifically designed for the fast execution of 
+large JavaScript applications. In order to increase performance
+V8 compiles JavaScript source code directly into machine code when it is first executed. 
+There are no intermediate byte codes or an interpreter \cite{Gray:2009:GCM:1610564.1610565}.
+JavaScript has gained popularity in server-side development using runtime
+environments such as Node.js and is also used in desktop and mobile applications. 
+%Mario: Creo que el párrafo de arriba sale sobrando
+% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
+To further evaluate how a current JavaScript implementation could be used by a
+computer scientist to develop complex optimization problems,
+a benchmark function, which was provided by the CEC2010 Special Session on
+Large-Scale Global Optimization \cite{tang2007benchmark} is used in this section. 
+The function is described next.
+
+The basic Rastrigin's function is separable, and is defined as follows;
+\begin{equation}
+F_{rastrigin}(x)=\sum\limits_{i=1}^D [ x_{i}^{2}-10\cos(2\pi x_i)+10  ] 
+\end{equation}
+where $D$ is the dimension and $x = (x1, x2, · · · , x_{D})$ is a
+$D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin's function 
+is a classical multimodal problem. Such problem is difficult since the number of local
+optima grows exponentially with the increase of dimensionality. To make it 
+non-separable, an orthogonal matrix is also used for coordinate rotation.
+The rotated Rastrigin's function is defined as follows:
+\begin{equation}
+F_{rot\_rastrigin}(x)=F_{rastrigin}(\textbf{z}), \textbf{z}= \textbf{x} * \textbf{M}
+\end{equation}
+Where $\textbf{M}$ is  a $D \times{D}$ orthogonal matrix. The benchmark
+emulates real-world optimization problems that most likely will consist 
+of different groups of parameters with strong dependencies within
+but little interaction between the groups. This issue is reflected in the benchmark 
+by randomly dividing the objective variables
+into several groups, each of which contains a number of variables. The parameter $m$ 
+is used to control the number of variables in each group and hence, defining the degree
+of separability:
+\begin{equation}
+F_{15}(x)=\sum\limits_{k=1}^\frac{D}{m}F_{rot\_rastrigin}[\textbf{z}(P_{(k-1)*m+1}:P_{k*m})] 
+\end{equation}
+For the benchmark $D = 1000$, Group size $m = 50$, $\textbf{x} = (x_1, x_2, \ldots , x_{D})$ is the
+candidate solution, $\textbf{o} = (o_1, o_2, \ldots  , o_{D})$ is the shifted global optimum,
+$\textbf{z} = \textbf{x} - \textbf{o}$ is the shifted candidate solution and
+finally $\textbf{P}$ is a random permutation of $\lbrace1, 2, \ldots  ,D\rbrace$.
+
+This optimization problem was selected because it is representative of the kind of
+algorithms, data types, and structures employed in large scale optimization problems. 
+The function was implemented in both Matlab and Java languages as part of the
+Test Suite of the companion competition of the Special Session and to give competitors an idea of the
+computational cost of the challenge, the runtime required for 10,000 function evaluations 
+for a particular configuration was published by organizers. The Java
+implementation took 7596ms and the Matlab version 1115ms. The whole experiment had the value of
+3,000,000 function evaluations as termination condition. Using the Java implementation as
+a guide, the function described above was implemented to compare the performance of JavaScript against
+both implementations, again using the time required for 10,000 function evaluations. 
+Before discussing the setup and performance results, certain details of
+the implementation in JavaScript are presented to give readers an idea of 
+how practical and mature this language is for developing complex mathematical functions.
+%
+\begin{figure*}[!htb]
+\centering
+\includegraphics[width=0.9\linewidth]{f15_times.png}
+\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
+\label{fig:f15_times}
+\end{figure*}
+
+\begin{itemize}
+\item {\em Randomize} Both Matlab and Java programs rely on a 
+Java Randomizer library for the 
+randomization of the shift vectors and matrices used by the function.
+The library includes the generation of pseudorandom Gaussians of type
+double and uniformly distributed integers. Generators use a single
+{\tt long} seed. In order to implement the same functionality in
+JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
+in the implementation of the standard {\tt Math.random()} between engines
+and also their results are non-deterministic. 
+The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
+consistent results across all JavaScript VMs.
+\item {\em Timing functions} In order to compare the performance of 
+each implementation, an accurate and consistent way of measuring 
+the runtime of functions is needed.
+In JavaScript the {\tt Date class} is often used 
+to measure execution time, but has the limitation of having a maximum 
+resolution given in milliseconds.
+In order to measure the runtime intervals with a higher precision two functions
+where used: for Node.js we used the native {\tt process.hrtime()} function
+which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
+array, and for measuring in the browser the {\tt Performance.now()} function was called,
+which uses floating-point numbers with up to microsecond precision. 
+A drawback found is that the {\tt Performance.now()} function is only implemented
+in Firefox and Chrome browsers. Both functions are independent of the system clock. 
+\item {\em Tools} An important consideration when choosing certain 
+language to write non-trivial programs is the availability of development tools. While developing
+the experiments, we used a comprehensive set of developer tools available for both
+browsers and desktop \cite{tilkov2010node}. These tools included
+package managers, in this case {\tt npm}, debuggers, logging libraries
+such as {\tt winston}, 
+network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
+Although the fact that developer tools vary from browser to browser
+supposed a drawback, they all have the same functionality. The
+JavaScript debugger can be accessed directly from the user menu in
+most browsers. 
+\item {\em Data Types} JavaScript uses floating point numbers with a 
+limited precision of 64 bits that partially implement the functionality of
+the {\tt StrictMath} library used by the Java implementation. If more precision
+is needed, developers could use {\tt math.js}, an extensive math library with
+support for matrices and big numbers.
+\end{itemize}
+
+We found that the JavaScript ecosystem is mature enough for developing non-trivial 
+code, but developers have to consider the differences between VMs implementations
+and therefore use certain libraries in order to increase the precision and repeatability of
+algorithms. As for the performance of JavaScript against Java and Matlab
+we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
+Intel Xeon E5 processor running OS X 10.10.5, 
+using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
+MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
+Google Chrome Version 46.0.2490.86 (64-bit) 
+and Firefox version 41.0.2. % (Paloma) Table here? Or at least the characteristics could be represented the same way as for the baseline experiment
+ The performance of the time required for 10,000 function evaluations is presented in
+Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
+performance with an average of 935ms, followed by Java with
+991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
+single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
+The results also show that there is not much difference between the Node.js
+implementation and that of a
+browser. There is also not much difference between running the code in the main thread or in Web
+Workers. When running the experiment in Firefox a dialog alerted that the
+script was taking too much time, showing that this
+kind of processing is not intended for the main thread. Additionally, there is not much overhead
+when running two
+Web Workers in parallel as they took 1279ms each. This experiment
+shows that JavaScript is viable as a language
+language for developing complex mathematical functions an these can run in a browser
+with acceptable performance.
+
 %---------------------------------------------------------------
@@
 \bibliographystyle{abbrv}
-\bibliography{geneura,volunteer,javascript,ror-js,GA-general}
+\bibliography{geneura,volunteer,javascript,GA-general}
 

### 2016-04-03 EvoSoft16/nodio-evosoft16.tex
Adding an expression of intentions

@@
 \item First the {\sf NodIO} process is started in Node.js. As part of the
-initialization procedure, the sequential number of experiments is initialized and a log
-file is created. The shared pool implemented as an array is reset.
-\item A volunteer follows the link of the experiment triggering a GET HTTP
-request, then the page is loaded.
-\item The main script renders the page and creates the worker instance. In
+initialization procedure, the sequential number of experiments held in
+a variable is initialized and a log
+file is created. The shared pool implemented as hash that uses as keys
+the chromosome strings is reset.
+\item A volunteer follows the link to the experiment triggering a GET HTTP
+request for static content and then the page is loaded.
+\item The main script renders the page and creates the worker
+  instance. In
 order to create a dedicated worker instance, another GET request is sent to the
@@
 universally unique identifier (UUID). The UUID is included in each of the HTTP
-requests sent to the server.
+requests sent to the server so that that particular worker can be
+identified and tracked in the logs. 
 The initial parameters for each island
@@
 specified parameters. Once the population has been initialized, the EA
-generation loop is started. After {\em n} generations the state of the population
-is evaluated. The chromosome with the best fitness is sent to the shared
+generation loop is started. After {\em n} generations  the chromosome
+with the best fitness is sent to the shared 
 population with a PUT HTTP request. If the best fitness is not the global best,
 the EA loop must continue, but first a chromosome is randomly selected from
-the shared population. The current generation number and the best
+the shared pool. The current generation number and the best
 chromosome are also sent to the client's main thread in order to update the DOM to
@@
 all asynchronous, so the EA loop of
-the worker continues even before a message is received by the target. If the best fitness is a global best, the best
+the worker continues even before a message is received by the
+target. If the best fitness is a global best, the best 
 chromosome is sent to the server and the loop ends. A message is sent to the
@@
 a callback function. When an iteration message is received, the page is
-updated by displaying a dynamic plot of the generation number and best
+updated by plotting a dynamic graph of the generation number and best
 fitness, and a representation of the chromosome is also generated.
@@
 
-These two versions of the framework will have to be tested {\em in the
-  wild} to measure its performance. 
-However, before testing this framework in a real experiment, we have
-to establish a baseline for comparison.
+These two versions of the framework have been tested {\em in the
+  wild} to measure its performance; results are published elsewhere,
+but we also have to establish a baseline for comparison. This is only
+for comparison purposes and not intended to prove that a volunteer
+framework, if used exclusively, is faster than other
+implementations. The main intention of this line of research is to
+check how evolutionary algorithms work in this context, and also how
+volunteer computing frameworks can be used to {\em supplement}
+whatever resources, desktop computers, laptop or cloud, are availab.e
 
@@
 We have used the classical Trap function \cite{Ackley1987}
-used as well as the Rastrigin's floating-point optimization
+as well as the Rastrigin's floating-point optimization
 problem. Since JavaScript is a functional language and declaring a different 

### 2016-04-03 EvoSoft16/nodio-evosoft16.tex
Final retouches

@@
 citizens to {\em donate} cycles to scientific experiments by running
-an application \cite{Anderson06thecomputational}; this application can be embedded in a web page. And
-since browsers are ubiquitous, we are more interested in these kind of systems, this has led us to create a framework for
+an application \cite{Anderson06thecomputational}; this application can
+be embedded in a web page. And 
+since browsers are ubiquitous, we are more interested in these kind of
+systems, this has led us to create a framework for 
 evolutionary algorithm experiments that is able to work in that mode
@@
 As a volunteer based system, there are some issues that have to be addressed 
-\cite{sarmenta2001volunteer,web:BOINC} :
+\cite{sarmenta2001volunteer,web:BOINC}:
 \begin{enumerate}
-\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
-each request is known.
+\item Volunteers are anonymous, only the Internet Protocol addresses
+  (IP) sent in each request is known.
 \item As anonymous entities, volunteers are not accountable. 
@@
 crafting a fake request which, for instance, assigns a fake fitness to
-a particular chromosome.
+a particular chromosome or sends a hand made solution to the server. 
 Other two different kinds of attacks are also
@@
  experiments performed \cite{DBLP:journals/corr/GuervosG15}  
- and all data, once anonymized, as open data. This
+ and all data, once IPs have been anonymized, as open data. This
 builds trust between the user and the scientist, which indeed has an
@@
 faster than compiled languages such as Scala and, in any case, it provides
-a performance that is comparable to other languages usually employed % (Paloma) Adverb
+a performance that is comparable to other languages usually employed 
 in evolutionary computation and numerical optimization such as Python. 
@@
 Since JavaScript is
-quite a popular language nowadays, it also provides compilers from a number
-of languages, many of them in their family (like CoffeeScript), but
-many other outside, so researchers can, in fact, write their fitness
+quite a popular language nowadays, it can also be generated by
+compiling from a number 
+of languages, many of them in the same family (like CoffeeScript), but
+many other outside. Tesearchers can, in fact, write their fitness
 function in Python, Erlang, Scala, or even Java, and then compile it to
@@
 
-%---------------------------------------------------------------
-%\section{State of the art}
-%\label{sec:soa}
-
 \section{The NodIO framework and API}
@@
 The figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or {\tt
+algorithm behavior. Different client-side web technologies, such as JQuery or {\tt
   Chart.js} have
@@
   back to the server (via a {\tt PUT} request), and requests a random
-  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
+  individual from the server (via a {\tt GET} request). 
 \end{enumerate}
@@
   mechanism. The islands in the system are, in fact, unaware of each
-  other, communicating only through the server. This architecture 
-  has been tested in other high demand systems with success, but 
-  in the case of {\sf NodIO}, additional experiments are needed to asses 
-  the scalability of the communication system. % (Paloma) Is it ok if
-                                % we talk about scalability when we
-                                % were talking about adaptiveness?
-                                % Scalability was also described in
-                                % the first point. 
-  %(Mario) Scalability of the communication system in particular?
+  other, communicating only through the server. The webpage itself can
+  be displayed, using responsive design, in either desktop or mobile
+  systems. 
 \item Since the algorithm runs in the browser, user's {\em safety} is
@@
 
-The rationale for this change comes from the realization that JavaScript has been commonly used to develop client-side scripts
+The rationale for this change comes from the realization that
+JavaScript has been commonly used to develop client-side scripts 
 that handle the user interface, communicating asynchronously with servers and
@@
 language not designed for the development of high performance
-systems. On the other hand, current JavaScript  
-virtual machines (VMs from now on) are closing the gap, for instance the Google V8 
-is an engine specifically designed for the fast execution of 
-large JavaScript applications. In order to increase performance
-V8 compiles JavaScript source code directly into machine code when it is first executed. 
-There are no intermediate byte codes or an interpreter \cite{Gray:2009:GCM:1610564.1610565}.
-JavaScript has gained popularity in server-side development using runtime
-environments such as Node.js and is also used in desktop and mobile applications. 
-%Mario: Creo que el párrafo de arriba sale sobrando
-% (Paloma) Agree, I mean, it adds information but in case you have space problems... is ok if you start the section without it
-To further evaluate how a current JavaScript implementation could be used by a
-computer scientist to develop complex optimization problems,
-a benchmark function, which was provided by the CEC2010 Special Session on
-Large-Scale Global Optimization \cite{tang2007benchmark} is used in this section. 
-The function is described next.
+systems. On the other hand, current JavaScript virtual machines (VMs
+from now on) are closing the gap, for instance the Google V8 is an
+engine specifically designed for the fast execution of large
+JavaScript applications. In order to increase performance V8 compiles
+JavaScript source code directly into machine code when it is first
+executed.  There are no intermediate byte codes or an interpreter
+\cite{Gray:2009:GCM:1610564.1610565}.  To further evaluate how a
+current JavaScript implementation could be used by a computer
+scientist to develop complex optimization problems, a benchmark
+function, which was provided by the CEC2010 Special Session on
+Large-Scale Global Optimization \cite{tang2007benchmark} is used in
+this section.  The function is described next.
 
@@
 $D$-dimensional row vector (i.e., a $1 × D$ matrix). Rastrigin's function 
-is a classical multimodal problem. Such problem is difficult since the number of local
+is a classical multimodal problem. Such problem is difficult since the
+number of local 
 optima grows exponentially with the increase of dimensionality. To make it 
@@
 991ms. On the other hand, JavaScript took 32\% more time than Java with 1238ms in Chrome using a 
-single Web Worker and 1,234ms using Node.js. % Are these results correct? JS=>1240ms and Node.js=>1,270ms ???
+single Web Worker and 1,234ms using Node.js. % Are these results
+                                % correct? JS=>1240ms and
+                                % Node.js=>1,270ms ???
+%They should be... - JJ
 The results also show that there is not much difference between the Node.js

### 2016-04-18 ASOC/paper.Rnw
Starting version for Applied and Soft Computing

--- 2016-03-20 PPSN2016/paper.Rnw
@@
-\documentclass[runningheads,a4paper]{llncs}
+\documentclass[preprint]{elsarticle}
+\biboptions{longnamesfirst,angle,semicolon}
+
+\journal{Applied and Soft Computing}
 
 \usepackage[latin1]{inputenc}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\setcounter{tocdepth}{3}
 \usepackage{graphicx}
@@
 \usepackage{caption}
-
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
 
@@
 
-\title{A performance assessment of evolutionary algorithms in a volunteer computing system}
+\title{A performance assessment of evolutionary algorithms in a
+  volunteer computing system}
 
-\author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-  Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+\author[jjmg]{Juan-J.~Merelo\corref{cor1}\fnref{fn1}}
+\ead{jmerelo@ugr.es}
 
-\institute{Dept. of Computer Architecture and Technology, University
-of Granada, Spain \and
-Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+\author[pcd]{Paloma de las Cuevas \fnref{fn1}}
+\ead{palomacd@ugr.es}
 
-\maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+\author{Pablo Garc\'ia-S\'anchez\fnref{fn1}}
+\ead{fergunet@gmail.com}
+
+\author{Mario Garc\'ia-Valdez \fnref{fn2}}
+\ead{mario@tectijuana.edu.mx}
+
+\cortext[cor1]{Corresponding author}
+\fntext[fn1]{Dept. of Computer Architecture and Technology and CITIC, University
+of Granada, Spain}
+\fntext[fn2]{Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
 
@@
 
+\begin{keyword}
+
+  Volunteer computing, pool-based evolutionary algorithms
+\end{keyword}
+
+\maketitle
+
 \section{Introduction}
@@
 distribution, but also an increasingly reliable and high-performance
-operating system for running distributed applications. This is based in two
-key things: the Javascript virtual machine every browser runs and a
-simplified standard interface for interacting with servers.
-% These are "things", not "facts". As facts would be... "every browser runs a Javascript virtual machine and that there is a simplified standard interface for interacting with servers (?)
-% changed to things - JJ
-Thus
-creating a distributed computing experiment is just a matter of making
-a JavaScript application interchange information with a server, called
-REST. From the point of view of the programmer, this involves
+operating system for running distributed applications. This is based
+in two key things: the Javascript virtual machine every browser runs
+and a simplified standard interface for interacting with servers.
+Thus creating a distributed computing experiment is just a matter of
+making a JavaScript application interchange information with a server,
+called REST. From the point of view of the programmer, this involves
 relatively common skills and no special libraries, since that
@@
 
-\bibliographystyle{splncs}
+\section*{References}
+
+\bibliographystyle{elsarticle-num}
 \bibliography{volunteer,GA-general,geneura,javascript,ror-js}

### 2016-04-23 Alife16/paper.tex
minor issues
References #116

@@
 this paper we examine the result of several volunteer distributed
-evolutionary computation experiments and try to find out which are
-those patterns and what makes an experiment successful or not,
+evolutionary computation experiments and try to find out what
+those patterns are and what makes an experiment successful or not,
 including the feedback loop that is created between the users and the
-algorithm itself. In this paper we will examine what kind of patterns
-we can expect and what influence those patterns have on the
-implementation of the evolutionary algorithm.
+algorithm itself.
 \end{abstract}
@@
 the more general context of distributed computing 
-\citep{nogueras2015studying} or including it in a more general study of the
-performance of the EA itself
+\citep{nogueras2015studying}  or included in a more general 
+study of the performance of the EA itself
 \citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
-measured without first understanding the dynamics of this kind of systems. Initial
+measured without first understanding the dynamics of this kinds of systems. Initial
 work was done for peer to peer systems by Stutzbach et
@@
 so important; even so, we have made a comparison between JavaScript and other languages
-\citep{2015arXiv151101088M} that shows that performance of JavaScript
+\citep{2015arXiv151101088M} that shows that the performance of JavaScript
 is comparable to other interpreted languages; compiled languages would
@@
   routes} 
-  that can be called for storing and retrieving information (the ''CRUD'' cycle:
+  that can be called for storing and retrieving information (the `CRUD'' cycle:
   create, request, update, and delete) from the server. 
@@
   kept the number of generations between migrations fixed since it is
-  a way of finding out how many real work every client has done. 
+  a way of finding out how much real work every client has done. 
 \end{enumerate}
@@
 Figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different web technologies, such as JQuery or {\tt
+algorithm behavior. Different web technologies, such as JQuery and {\tt
   Chart.js} have been used to build the user interface elements of the
@@
 Initial experiments were set up using the OpenShift
-PaaS, which provides a free tier, making the whole experiment cost
+PaaS \url{www.openshift.com}, which provides a free tier, making the whole experiment cost
 equal to \$0.00. Experiments were
 announced through a series of posts on Twitter and, in the latest case, Telegram, and
-results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
+results were published in\citep{DBLP:conf/gecco/GuervosG15}. For the
 purpose of this paper, we repeated the announcement several times
@@
 from the pool, as opposed to any individual that could be obtained
-before. Besides, cache policy deletes the oldest individuals, which
+before. Besides, the cache policy deletes the oldest individuals, which
 makes those in the pool be {\em current}, helping then newcomers and %Pablo: "then helping newcomers", instead?
@@
   red dots; in the case there is a single red dot, there was a single
-  experiment in which many computers participated (for instance, 16
+  experiment in which that many computers participated (for instance, 16
   computers in the experiment in the far left or 29 in the middle
@@
 experiments. An interesting question to ask, for instance, is if 
-by adding more computers makes the experiment take less. In fact, as
+by adding more computers it makes the experiment take less time. In fact, as
 shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
@@
 by IP. The green dot at the bottom is probably an outlier that we
-will try to explain later on. This leads us to conclude that a better
+will try to explain later on. This leads us to conclude that a larger
 amount of computers might contribute to speed up the solution, if the
@@
 The inverse
-Weibull distribution is a special case of the GEV distribution we have
-used in those papers, and appears usually in natural sciences and
+Weibull distribution is a special case of the GEV distribution
+in those papers, and appears usually in natural sciences and
 artificial life, usually related to decay. It has been frequently

### 2016-04-23 Alife16/paper.tex
Fixing Issue #115
Issue #115 @JJ please read it to see if it is more clear

@@
 \citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
-platforms avoid single points of failure (the server) but, since they
-need a certain amount of infrastructure installed to start, the
-threshold to join them is much lower; this makes gathering new users
+platforms to avoid single points of failure (the server) but, since they
+need a certain amount of infrastructure installed in the client, the
+threshold to join them is much lower; but once the infrastructure of P2P systems is in place
+the gathering of new users becomes 
 immediate and spontaneous, so that they can be used for short

### 2016-04-24 Alife16/paper.tex
Making the description more clear #113

@@
 the islands running in the browser {\em spill} from one experiment to
-the next. However, previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible. %Pablo: cite
+the next. 
+%However, previous experiments have proven that the influence
+%of these islands in the next experiment is indeed negligible. %Pablo: cite
+If and individual that is close or even at optimum value is stored in the cache 
+the next experiment will be affected, because it will require less time 
+to finish, this problem has been addressed in later versions of the system. 
+But time to solutions is not the important measure here, is the computational
+power of system established by the number of volunteers, in this measure 
+previous experiments have proven that the influence
+of these islands in the next experiment is indeed negligible.
 

### 2016-04-24 Alife16/paper.tex
Frequently

@@
 so lighter shade for a column of dots indicates that a particular number
-of computers happened less frequently, while darker shadow means more frequency. 
+of computers happened less frequently, while darker shadow means more. 
 From left to right and top to bottom experiments 4/4, 4/24 and 7/31,

### 2016-04-24 Alife16/paper.tex
Minor error
ref

@@
 issues related to security, trust and privacy among others. The
-computing system becomes a {\em sociotechnical} system}
+computing system becomes a {\em sociotechnical system}
 \citep{vespignani2009predicting}. In this paper, we are going to focus
@@
 leave it running. Since it finishes or stalls after a number of
-generations shown in Figure \cite{tab:summary:os}, volunteers just
+generations shown in Figure \ref{tab:summary:os}, volunteers just
 leave after that. Curiously enough, the scale parameter is roughly

### 2016-04-24 Alife16/paper.tex
Cut a bit on the paragraph refs f731441

@@
 authors, including more recent efforts
-\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
-%Pablo: Desell in 2008 is not so recent
-used Java \citep{chong:1999:jDGPi} and others have gone away from the
-server-based paradigm to embrace peer to peer systems
-\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
-platforms to avoid single points of failure (the server) but, since they
-need a certain amount of infrastructure installed in the client, the
-threshold to join them is much lower; but once the infrastructure of P2P systems is in place
-the gathering of new users becomes 
-immediate and spontaneous, so that they can be used for short
-algorithms on call.
+\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many
+other researchers have used Java \citep{chong:1999:jDGPi}; others have
+embraced peer to peer systems
+\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
+computing platforms avoid single points of failure (the server) and
+once installed need no effort to gather new users for an experiment,
+but the cost of acquiring new users is high since they need to be set
+up.
 

### 2016-04-24 Alife16/paper.tex
Shortening the state of the art to cut the paper down to 8 pages

@@
 
-On the other hand, we are also interested in measuring the performance
-of volunteer computing systems, an area in which there have been
-relatively few efforts.
-There were some initial attempts to avoid the differences in performance
-that could be obtained from volunteers  by making
-the algorithm adaptive to the kind of resources allotted to it
-\citep{milani2004online}, which is actually not such a big problem in
-algorithms such as the EA that can be easily 
-parallelized via population splitting or by farming out the evaluation to all
-the nodes available. Lately, several approaches have focused on the
-fault-tolerance of volunteer systems
-\citep{gonzalez2010characterizing} which it can, of course, be studied in
-the more general context of distributed computing 
-\citep{nogueras2015studying}  or included in a more general 
-study of the performance of the EA itself
-\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
-measured without first understanding the dynamics of this kinds of systems. Initial
+The number of users is key in the performance of these systems, but it
+also essential to adapt
+the algorithm itself to the available resources, as shown by
+\citep{milani2004online}, although EAs can be readily distributed via
+population splitting or by farming out the evaluation to all 
+the nodes available. However, user churn affects experiment performance 
+\citep{gonzalez2010characterizing,nogueras2015studying} and also the
+performance of the algorithm itself
+\citep{DBLP:journals/gpem/LaredoBGVAGF14}. All these issues imply that
+a the performance of a volunteer system cannot be
+measured without first understanding its dynamics. Initial
 work was done for peer to peer systems by Stutzbach et
@@
 enhance performance. On the contrary, in this paper, we will use {\em
-  actual} volunteers, anonymous persons that are requested to join an
-experiment through social networks.
+  actual} volunteers.
 

### 2016-04-24 Alife16/paper.tex
Revising the SoS
Closes #115

@@
 \citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
-computing platforms avoid single points of failure (the server) and
+computing platforms avoid single points of failure, the server, and
 once installed need no effort to gather new users for an experiment,

### 2016-04-24 Alife16/paper.tex
Reviewing revision by @mariosky
And changing it a bit to avoid confusion. Closes #113.

@@
 the next. 
-%However, previous experiments have proven that the influence
-%of these islands in the next experiment is indeed negligible. %Pablo: cite
-If and individual that is close or even at optimum value is stored in the cache 
-the next experiment will be affected, because it will require less time 
-to finish, this problem has been addressed in later versions of the system. 
-But time to solutions is not the important measure here, is the computational
-power of system established by the number of volunteers, in this measure 
-previous experiments have proven that the influence
-of these islands in the next experiment is indeed negligible.
+If an individual that is close or even at optimum value is stored in the cache 
+the next experiment will be affected, and it will require less time 
+to finish. This problem has been addressed in later versions of the
+system, however, time to solution is not the important measure here,
+where we are concerned with the computational 
+power of the system established by the number of volunteers.
 
-The table shows that every run included more than 30 experiments. The
+The table shows that every experiment included more than 30 runs. The
 %Pablo: reference to table? (maybe the final print will misplace it)

### 2016-04-24 Alife16/paper.tex
Explaining a bit figures.
But there's not much space to do more. Closes #112

@@
 \caption{Duration of experiments vs. number of different IPs (nodes)
-  participating in it, with averages and standard deviation shown as
+  participating in it, with averages per number of IPs and standard deviation shown as
   red dots; in the case there is a single red dot, there was a single
@@
 \includegraphics[width=0.32\linewidth]{time-vs-rank-alife-32.png}
-\caption{Duration of experiments vs. rank, with $y$ axis in a
+\caption{Duration of experiments vs. rank of experiments sorted by
+  descending duration, with $y$ axis in a
   logarithmic scale. Dot color is related to the number of IPs
   participating in the experiment. From left to right and top to bottom, experiments
-  4/4, 4/24 and 7/31 and caches=128,64, 32} 
+  4/4, 4/24 and 7/31 and caches=128, 64, 32.} 
 \label{fig:zipf:os}
@@
 \caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
-  distribution (in red). From left to right and top to bottom, experiments
+  distribution (in red); axis $x$ shows IPs sorted by descending
+  number of PUTs. From left to right and top to bottom, experiments
   4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 

### 2016-04-24 Alife16/paper.tex
Including a definition of the cache
Which was there, only not too clear. Closes #111

@@
   The server has the capability to
-  run a single experiment, storing the chromosomes in a data structure
-  that is reset when the solution is found. This data structure can
+  run a single experiment, storing the chromosomes in a key-value store
+  that is reset when the solution is found. This store can
   hold every chromosome in a particular experiment, or have a finite
   size that erases the oldest chromosomes once it has filled to
-  capacity. In this paper we will test both implementations.  
+  capacity, acting as a cache. In this paper we will test both implementations.  
 \item A client that includes the evolutionary algorithm as
@@
 purpose of this paper, we repeated the announcement several times
-through the month of February. All  %% April --> January ??
+through the month of February of the next year. All  
 in all, we have the set of runs with the characteristics shown in
@@
 days. No particular care was taken about the time of the announcement
-%Pablo: but it would great at least describe how many times was announced and how long
+%Pablo: but it would great at least describe how many times was
+%announced and how long
+% We don't really tracked this... - JJ
 or the particular wording. Every {\em experiment} consisted in running
@@
 
-The table shows that every experiment included more than 30 runs. The
-%Pablo: reference to table? (maybe the final print will misplace it)
+The table \ref{tab:runs}shows that every experiment included more than 30 runs. The
 number of different IPs intervening in them varied from more than one

### 2016-04-24 Alife16/paper.tex
Changed all to socio-technical to be coherent with the title closes #119

@@
 algorithm and of the performance of the whole system. As a
-sociotechnical system, this participation follows some patterns and in
+socio-technical system, this participation follows some patterns and in
 this paper we examine the result of several volunteer distributed
@@
 issues related to security, trust and privacy among others. The
-computing system becomes a {\em sociotechnical system}
+computing system becomes a {\em socio-technical system}
 \citep{vespignani2009predicting}. In this paper, we are going to focus
@@
 As far as we know, this paper presents one of the few experiments
-that measure the performance of a sociotechnical
+that measure the performance of a socio-technical
 metacomputer, that is, a spontaneously created parallel computer that
@@
 follows a Weibull distribution. This makes engagement the key for
-leveraging the performance of the sociotechnical metacomputer and a
+leveraging the performance of the socio-technical metacomputer and a
 way to improve results in the future.
@@
 Our intention in this paper was to assess the capabilities of a
-sociotechnical system formed by a client-server web-based framework
+socio-technical system formed by a client-server web-based framework
 running a distributed evolutionary algorithms and the volunteers that
@@
 difficult, the users stay for longer in the web page, making less
-ephemeral the sociotechnical computing system created by the
+ephemeral the socio-technical computing system created by the
 simulation.

### 2016-04-26 GECCO/volunteer-clouds-gecco16.tex
First final version submitted to Sheridan
@mariosky, I could do it, no problem finally :-)

@@
 
-\documentclass{sig-alternate}
+\documentclass{sig-alternate-05-2015}
+\CopyrightYear{2016}
+\setcopyright{rightsretained}
+\conferenceinfo{GECCO'16 Companion}{July 20-24, 2016, Denver, CO, USA}
+\isbn{978-1-4503-4323-7/16/07}
+\doi{http://dx.doi.org/10.1145/2908961.2926999}
+
+\clubpenalty=10000
+\widowpenalty = 10000
 
@@
 
-\numberofauthors{5} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-% \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
-%   and CITIC, University of Granada, Granada, Spain \\
-% Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
-% Tecnológico de Tijuana, Tijuana, M\'exico\\
-% Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-% Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
-% Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
-% University of Granada, Granada, Spain\\
-
-% \thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
-% }
+\numberofauthors{6}
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
+Juan-J.~Merelo, Pedro A. Castillo, Pablo Garc\'ia-S\'anchez, Paloma de las Cuevas\\
+\affaddr{Dept. of Computer Architecture and Technology
+  and CITIC, University of Granada, Granada, Spain} \\
+\email{(jmerelo|pacv|pgarcia|palomacd)@ugr.es}
+\alignauthor Nuria Rico\\
+\affaddr{Dept. of Statistics, University of Granada, Granada, Spain}\\
+\email{nrico@ugr.es}
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
+Mario Garc\'ia-Valdez,\affaddr{ Dept. of Graduate Studies, Instituto
+Tecnol\'ogico de Tijuana, Tijuana, M\'exico}\\
 }
@@
 %---------------------------------------------------------------
-% \section*{Acknowledgment}
-% 
-% This work has been supported in part by
-% TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
-% SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
-% GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
-% Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
-% We would also like to thank the
-% anonymous reviewers of previous versions of this paper who have really
-% helped us to improve 
-% this paper (and our work) with their suggestions. We are also grateful
-% to Anna S\'aez de Tejada for her help with the data processing scripts.
+\section*{Acknowledgment}
+
+This work has been supported in part by
+TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR), 
+SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
+GENIL project (CEI-BIOTIC Granada). Additional support was recieved by
+Projects 5622.15-P (ITM) and  PROINNOVA 2015: 220590 (CONACYT).
 

### 2016-04-27 GECCO/volunteer-clouds-gecco16.tex
Corrected version submitted again

@@
 \CopyrightYear{2016}
-\setcopyright{rightsretained}
-\conferenceinfo{GECCO'16 Companion}{July 20-24, 2016, Denver, CO, USA}
-\isbn{978-1-4503-4323-7/16/07}
-\doi{http://dx.doi.org/10.1145/2908961.2926999}
+\CopyrightYear{2016}
+\setcopyright{acmlicensed}
+\conferenceinfo{GECCO '16,}{July 20 - 24, 2016, Denver, CO, USA}
+\isbn{978-1-4503-4206-3/16/07}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/2908812.2908849}
 
@@
 \alignauthor
-Juan-J.~Merelo, Pedro A. Castillo, Pablo Garc\'ia-S\'anchez, Paloma de las Cuevas\\
+Juan-J.~Merelo\\ 
+Pedro A. Castillo\\ 
+Pablo Garc\'ia-S\'anchez\\
+Paloma de las Cuevas\\
 \affaddr{Dept. of Computer Architecture and Technology
@@
 available in almost every desktop or mobile device. Another advantage is
-that the source code can be directly examined by volunteers, because it is not
+that the source code can be directly examined by volunteers as long as
+it is not obscured, because it is not
 compiled as other options such as Java Applets, ActionScript, or
@@
 %With these three experiments we want mainly to try and model
+
 With these three experiments our main objective is to try and model
@@
 \end{table}
-%
+
 This distribution was originally proposed to fit extreme values
@@
 to point out that the behavior of volunteer computing nodes follows a
-certain pattern, found in SETI@home, and which also appears in our framework. This
+certain pattern, found in SETI@home, and which also appears in our
+framework.
+
+This
 distribution is governed by three parameters, the usual location $\mu$
@@
 predictive value of these fits, however, is limited, over all taking
-%Pablo: question for Paloma: is over all or overall?
-into account the low correlation between successive events shown
-above.
+into account the low correlation between successive events which is
+shown in the figures afigure bove.
 
 At any rate, this also shows that a convenient way of increasing the
-%Mario: As I understend is: 
-computing power would be increasing this minimum amount of experiments
-per user. For this reason, we made some changes to the implementation,
-whose results are presented in the next subsection. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?
+computing power would be to try and improve this minimum amount of experiments
+per user. This is checked next.
 

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Fixing some typos, closes #121

@@
 universally unique identifier (UUID). The UUID is included in each of the HTTP
-requests sent to the server so that that particular worker can be
+requests sent to the server so that particular worker can be
 identified and tracked in the logs. 
@@
 volunteer computing frameworks can be used to {\em supplement}
-whatever resources, desktop computers, laptop or cloud, are availab.e
+whatever resources, desktop computers, laptop or cloud, are available
 
@@
 the time to success depends on the population size, with a bigger population
-contributing to have more diversity and thus speeding up the solution \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
+contributing to have more diversity and thus speeding up the solution searching \cite{DBLP:conf/lion/LaredoDFGB13}. The volunteer computing
 experiments that we will describe next do not and can not have the

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Modified the script to generate f15 boxplots as requested by the reviewer, and fixed figure 3 width for issue #123

--- 2016-04-03 EvoSoft16/nodio-evosoft16.tex
@@
 
-\begin{figure*}[!t]
+\begin{figure}[!t]
 \centering
-\includegraphics[width=10cm]{baseline-times.png}
+\includegraphics[width=8cm]{baseline-times.png}
 \caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
@@
 \label{fig:baseline}
-\end{figure*}
+\end{figure}
 First, we establish the baseline performance by running fifty runs of the

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Solving hyphenation problems. References #125

--- 2016-05-03 EvoSoft16/nodio-evosoft16.tex
@@
 \usepackage{url}
+
+% *** EXTRA HYPHENATION RULES ***
+%
+\usepackage[english]{babel}
+\hyphenation{
+  au-tonomous
+  non-de-ter-min-is-tic}
 
@@
 \item {\em Scalability} is provided via the use of a lightweight and
-  high-performance, single-threaded, server based in Node.js and
-  Express.js. Although this single server is a bottleneck since it
+  high-performance, single-threaded, server based in Express.js and
+  Node.js. Although this single server is a bottleneck since it
   will eventually saturate, the fact that it runs as a non-blocking single thread
@@
   communicate with the rest of the islands.
-\item {\em Adaptiveness} is achieved simply through the autonomous
+\item {\em Adaptiveness} is achieved through the autonomous
   operation of every individual island without any synchronization
@@
 \begin{verbatim}
-Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
-2014 x86_64 x86_64 x86_64 GNU/Linux
+Linux penny 3.13.0-34-generic #60-Ubuntu SMP
+Wed Aug 13 15:45:27 UTC 2014
+x86_64 x86_64 x86_64 GNU/Linux
 \end{verbatim}
@@
 in the implementation of the standard {\tt Math.random()} between engines
-and also their results are non-deterministic. 
-The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
+and also their results are non-de\-ter\-min\-is\-tic. 
+The library {\tt random-js} implements an algorithm called Mersenne Twister, producing 
 consistent results across all JavaScript VMs.
@@
 which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
-array, and for measuring in the browser the {\tt Performance.now()} function was called,
+array, and for measuring in the browser the {\tt Per\-for\-mance.now()} function was called,
 which uses floating-point numbers with up to microsecond precision. 
 A drawback found is that the {\tt Performance.now()} function is only implemented
-in Firefox and Chrome browsers. Both functions are independent of the system clock. 
+in Firefox and Chrome brows\-ers. Both functions are independent of the system clock. 
 \item {\em Tools} An important consideration when choosing certain 

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Fixing more typos, closes #125

@@
 of languages, many of them in the same family (like CoffeeScript), but
-many other outside. Tesearchers can, in fact, write their fitness
+many other outside. Researchers can, in fact, write their fitness
 function in Python, Erlang, Scala, or even Java, and then compile it to

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Completing fig.4 label
Adds copyright and corrects @fergunet's email
Checking figure revisions
Closes #123
Adding a reference to justify that
That was not the main point of the paper, anyway, it was only done for the sake of comparison

Closes #124

@@
 
-% Copyright
-\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-%Conference
-\conferenceinfo{GECCO '16}{Denver, CO, USA}
-% --- End of Author Metadata ---
+\CopyrightYear{2016}
+\setcopyright{acmlicensed}
+\conferenceinfo{GECCO'16 Companion,}{July 20 - 24, 2016, Denver, CO, USA}
+\isbn{978-1-4503-4323-7/16/07}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/2908961.2931723}
 
@@
   and CITIC, University of Granada, Granada, Spain} \\
-\email{(jmerelo|pacv|pgarcia|palomacd)@ugr.es}
+\email{(jmerelo|pacv|pablogarcia|palomacd)@ugr.es}
 \alignauthor
@@
 find the solution to the 40-trap function with parameters $l=4, a=1,
-b=2, z=3$ and a population size equal to 512. The algorithm in a
+b=2, z=3$ and a population size equal to 512.
+
+The algorithm in a
 single island was run until the solution (a string with all
@@
 took around a minute, on average, that is $68.9694$ seconds, to
-perform these runs. In this experiment only 33, that is, 66\% were
-successful. The experiments were repeated for population $p=1024$,
+perform these runs, but in this experiment only 33 out of 50, that is, 66\% were
+successful.
+
+In order to check running time for different initial conditions, the
+experiments were repeated for population $p=1024$, 
 with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
-seconds. Results for these two experiments
+seconds. In general, deceptive problems such as this one are very
+sensitive to initial diversity. A bigger population size implies that
+there will be more schemata in it, leading to a greater degree of
+success in exploitation. In general, that seems to be the case in many
+deceptive problems so that explains this order-of-magnitude drop in
+running time \cite{goldberg1991genetic}.
+
+Results for these two experiments
 are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
@@
 \includegraphics[width=0.9\linewidth]{f15_times.png}
-\caption{ Runtime of 10,000 function evaluations for $F_{15}$} 
+\caption{Runtime of 10,000 function evaluations for $F_{15}$ in
+  several implementation languages. } 
 \label{fig:f15_times}

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
ACM Classification added
Please compile and check, I don’t have latex here.

--- 2016-05-03 EvoSoft16/nodio-evosoft16.tex
@@
 \end{abstract}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10010147.10010178.10010219</concept_id>
+<concept_desc>Computing methodologies~Distributed artificial intelligence</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010178.10010219.10010220</concept_id>
+<concept_desc>Computing methodologies~Multi-agent systems</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Computing methodologies~Distributed artificial intelligence}
+\ccsdesc[300]{Computing methodologies~Multi-agent systems}
 

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Some final touches

--- 2016-05-03 EvoSoft16/nodio-evosoft16.tex
@@
 \alignauthor
-Juan-J.~Merelo, Pedro A. Castillo, Pablo Garc\'ia-S\'anchez, Paloma de las Cuevas\\
+Juan-J.~Merelo\\
+Pedro A. Castillo\\
+Pablo Garc\'ia-S\'anchez\\
+Paloma de las Cuevas\\
 \affaddr{Dept. of Computer Architecture and Technology
@@
 \alignauthor
-Mario Garc\'ia-Valdez,\affaddr{ Dept. of Graduate Studies, Instituto
-Tecnológico de Tijuana, Tijuana, M\'exico}\\
+Mario Garc\'ia-Valdez,
+\affaddr{ Dept. of Graduate Studies, Instituto
+Tecnológico de Tijuana, Tijuana, M\'exico}
+\email{mario@tectijuana.edu.mx}\\
 }

### 2016-05-04 EvoSoft16/nodio-evosoft16.tex
Keywords added when submitting
Thanks anyway @mariosky Closes #122
New new version for EvoSoft

--- 2016-05-03 EvoSoft16/nodio-evosoft16.tex
@@
 
+\setlength{\paperheight}{11in}
+\setlength{\paperwidth}{8.5in}
+\usepackage[
+  pass,% keep layout unchanged 
+  % showframe,% show the layout
+]{geometry}
+
 \begin{document}
@@
 
-\title{NodIO: a framework and architecture for pool-based evolutionary
-  computation} 
+\title{NodIO: a Framework and Architecture for Pool-based Evolutionary Computation} 
 %

### 2016-10-28 EvoPar2017/paper.tex
Lo siento, borré la historia...

--- 2016-03-20 PPSN2016/paper.Rnw
@@
-\documentclass[runningheads,a4paper]{llncs}
+\documentclass[runningheads,a4paper]{llncs}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \DeclareGraphicsExtensions{.pdf}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<setup, cache=FALSE,echo=FALSE>>=
-library("ggplot2")
-experiment.data <- data.frame(Experiment=character(),
-                              milliseconds=character(),                           
-                              IPs=character(), 
-                              actualIPs=character(),
-                              PUTs=character(),
-                              actualPUTs=character(),
-                              PUTratio = character(),
-                              PUTsIP =character())
-
-these.experiments <- c("nooverlap-1","nooverlap-newgraph-1", "nooverlap-newgraph-2","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
-
-experiment.aggregates <- data.frame(Experiment=character(),
-                                    milliseconds=character(),
-                                    IPs=character(),
-                                    actualIPs = character(),
-                                    PUTs = character(),
-                                    actualPUTs = character(),
-                                    count = character(),
-                                    totalIPs = character(),
-                                    PUTsIP = character())
-
-experiment.aggregates.filtered = experiment.aggregates
-
-ips.data <- data.frame(Experiment=character(),                      
-                       PUTs=character(),
-                       reboots=character(),
-                       idu=character(),
-                       idx=character())
-
-ips.aggregates <- data.frame(Experiment=character(),
-                             PUTs = character(),
-                             reboots = character(),
-                             avgPUTs = character(),
-                             avgreboots = character(),
-                             totalIPs = character())
-
-for ( i in these.experiments ) {
-    this.data <- read.csv(paste0("../data/2016-PPSN/ips-time-cache-2016-03-cache=32-",i,".csv"))
-
-    this.ip.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
-    this.len <- length(this.ip.data$PUTs)
-    seq <- 1:this.len
-    ips.data <- rbind(ips.data,
-                      data.frame(Experiment=rep(i,length(this.ip.data$PUTs)),
-                                 idu=seq,
-                                 idx=seq/this.len,
-                                 PUTs=as.integer(this.ip.data$PUTs),
-                                 reboots=as.integer(this.ip.data$reboots))
-                       )
-
-    ips.aggregates <- rbind( ips.aggregates,
-                            data.frame(Experiment=i,
-                                       PUTs = median(this.ip.data$PUTs),
-                                       reboots = median(this.ip.data$reboots),
-                                       avgPUTs = mean(this.ip.data$PUTs),
-                                       avgreboots = mean(this.ip.data$reboots),
-                                       totalIPs =this.len )
-                            )
-
-    experiment.aggregates <- rbind( experiment.aggregates,
-                                   data.frame( Experiment=i,                   
-                                              milliseconds=median(as.integer(this.data$milliseconds)),
-                                              IPs=median(as.integer(this.data$IPs)),
-                                              actualIPs=median(as.integer(this.data$actualIPs)),
-                                              PUTs=median(as.integer(this.data$PUTs)),
-                                              actualPUTs=median(as.integer(this.data$actualPUTs)),
-                                              PUTratio=median(as.integer(this.data$actualPUTs)/as.integer(this.data$PUTs)),
-                                              count= length(this.data$IPs),
-                                              totalIPs=this.len,
-                                              PUTsIP=median(as.integer(this.data$PUTs)/as.integer(this.data$IPs)))
-                                   )
-
-    experiment.data <- rbind( experiment.data,
-                             data.frame( Experiment=rep(i, length(this.data$IPs)),                   
-                                        milliseconds=as.integer(this.data$milliseconds),
-                                        IPs=as.integer(this.data$IPs),
-                                        actualIPs=as.integer(this.data$actualIPs),
-                                        PUTs=as.integer(this.data$PUTs),
-                                        actualPUTs=as.integer(this.data$actualPUTs),
-                                        PUTsIP=as.integer(this.data$PUTs)/as.integer(this.data$IPs)))
-
-    this.data <- this.data[this.data$actualPUTs>=16,]
-    experiment.aggregates.filtered <- rbind( experiment.aggregates.filtered,
-                                            data.frame( Experiment=i,                   
-                                                       milliseconds=median(as.integer(this.data$milliseconds)),
-                                                       IPs=median(as.integer(this.data$IPs)),
-                                                       actualIPs=median(as.integer(this.data$actualIPs)),
-                                                       PUTs=median(as.integer(this.data$PUTs)),
-                                                       actualPUTs=median(as.integer(this.data$actualPUTs)),
-                                                       count= length(this.data$IPs),
-                                                       PUTsIP=median(as.integer(this.data$PUTs)/as.integer(this.data$IPs))
-                                                       )
-                                            )
-
-}
-
-experiment.data$newgraph <- grepl("newgraph",experiment.data$Experiment)
-experiment.data$nooverlap <- grepl("nooverlap",experiment.data$Experiment)
-experiment.data$reboot <- grepl("reboot",experiment.data$Experiment)
-experiment.data$cachecrash <- grepl("cache",experiment.data$Experiment)
-
-time.data <- data.frame(Experiment=character(),
-                        minute=as.Date(character()),
-                        IPs=character())
-
-these.experiments <- c("nooverlap-0","nooverlap-newgraph","nooverlap-newgraph-reboot","nooverlap-newgraph-reboot-cachecrash")
-ip.correlations <- data.frame(Experiment=character(),
-                              correlation=character())
-
-for ( i in these.experiments ) {
-    this.data <- read.csv(paste0("../data/2016-PPSN/ips-per-minute-cache=32-",i,".csv"))
-    this.data.plus1 <- this.data[2:length(this.data$IPs),]
-    ip.correlations <- rbind(ip.correlations, 
-                             data.frame(Experiment=i,
-                                        correlation = cor(this.data[c(1:length(this.data$IPs)-1),]$IPs
-                                                         ,this.data.plus1$IPs)))
-    time.data <- rbind(time.data,
-                       data.frame(Experiment=rep(i,length(this.data$IPs)),
-                                  minute=strptime(this.data$time,"%Y-%m-%d %H:%M"),
-                                  IPs=as.integer(this.data$IPs))
-                       )
-}
-
-@
-
-\title{A performance assessment of evolutionary algorithms in a volunteer computing system}
+\title{A performance assessment of evolutionary algorithms in
+  volunteer computing environments}
 
@@
 \maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+
+%\thispagestyle{empty}
+%\pagestyle{empty}
 
@@
 \centering
-<<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
-ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
-@ %def graph for IPs and puts
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/zipf-1}
+
+\end{knitrout}
 %
@@
 \begin{center}
-<<correlations, cache=FALSE,echo=FALSE>>=
-ip.correlations$Experiment <- c("base","graph1","reboot","cachereboot")
-kable(ip.correlations)
-@ %def correlations
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\begin{tabular}{l|r}
+\hline
+Experiment & correlation\\
+\hline
+base & 0.8408602\\
+\hline
+graph1 & 0.7346538\\
+\hline
+reboot & 0.7280738\\
+\hline
+cachereboot & 0.4657099\\
+\hline
+\end{tabular}
+
+\end{knitrout}
 \end{center}

### 2016-10-30 EvoPar2017/paper.Rnw
Lo siento, borré la historia...
Anonymizes authors closes #128

--- 2016-03-20 PPSN2016/paper.Rnw
@@
 
-\title{A performance assessment of evolutionary algorithms in a volunteer computing system}
+\title{A performance assessment of evolutionary algorithms in
+  volunteer computing environments}
 
-\author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-  Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
+%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
 
-\institute{Dept. of Computer Architecture and Technology, University
-of Granada, Spain \and
-Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+% \institute{Dept. of Computer Architecture and Technology, University
+% of Granada, Spain \and
+% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+
+\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
+  \and Jet A. Nother \inst{2}}
+
+\institute{Invisible University, Frantzia \and
+Institute of Higher Studies, Zootropa}
 
 \maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+
+%\thispagestyle{empty}
+%\pagestyle{empty}
 
@@
 
-This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
-Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
-gathering users for the new experiments.
+% This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+% Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
+% 2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
+% gathering users for the new experiments.
+
+Several projects\\
+Taking\\
+This\\
+Much\\
+Space\\
 

### 2016-10-30 EvoPar2017/paper.tex
Anonymizes authors closes #128

@@
 
-\author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-  Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
+%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
 
-\institute{Dept. of Computer Architecture and Technology, University
-of Granada, Spain \and
-Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+% \institute{Dept. of Computer Architecture and Technology, University
+% of Granada, Spain \and
+% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+
+\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
+  \and Jet A. Nother \inst{2}}
+
+\institute{Invisible University, Frantzia \and
+Institute of Higher Studies, Zootropa}
 

### 2016-10-30 EvoPar2017/paper.Rnw
Anonymising URLs. Closes #129

--- 2016-03-20 PPSN2016/paper.Rnw
@@
 
-\title{A performance assessment of evolutionary algorithms in a volunteer computing system}
+\title{A performance assessment of evolutionary algorithms in
+  volunteer computing environments}
 
@@
 system {\sf NodIO}. All parts of the framework are free and available
-with a free license from
-\url{https://github.com/JJ/splash-volunteer}.
+with a free license from \url{https://ANONYMOUS-URL}.
+%\url{https://github.com/JJ/splash-volunteer}.
 
@@
   Chart.js} have been used to build the user interface elements of the
-framework, which should be running in
-\url{http://nodio-jmerelo.rhcloud.com}, a free resource hosted in the
+framework, which should be running in \url{http://ANONYMOUS-URL}, 
+%\url{http://nodio-jmerelo.rhcloud.com}, 
+a free resource hosted in the
 OpenShift Platform as a Service\footnote{It is not guaranteed to be running,
@@
 the data and processing scripts, are published with a free license in GitHub at
-\url{https://git.io/gecco-es-15}.
+\url{ANONYMOUS-URL}.
+%\url{https://git.io/gecco-es-15}.
 
@@
 
-This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
-Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
-gathering users for the new experiments.
+%This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+%Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
+%2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
+%gathering users for the new experiments.
 

### 2016-10-31 EvoPar2017/paper.tex
Anonymising References. Closes #130

--- 2016-10-28 EvoPar2017/paper.tex
@@
 here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98}.
+computation, as it has been done for a long time \cite{jj-ppsn98anonymous}.
 
@@
 frameworks such as the one we are using here, and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15}.
+was described in \cite{DBLP:conf/gecco/GuervosG15anonymous}.
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/zipf-1}
+%\includegraphics[width=\maxwidth]{figure/zipf-1}
 
@@
 running heterogeneous algorithms in the same way that was done in
-\cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other
+\cite{DBLP:journals/grid/ValdezTGVO15anonymous}. Diversity, as in any other
 evolutionary algorithm, seems to be the key, but there are many ways of
@@
 \bibliographystyle{splncs}
-\bibliography{volunteer,GA-general,geneura,javascript,ror-js}
+\bibliography{volunteer,GA-general,geneura,javascript,ror-js,anonymous}
 

### 2016-11-07 EvoPar2017/paper.Rnw
Anonymizes this

@@
 
-\author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-  Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+%% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
+%%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+\author{First Author \and Second Author}
 
-\institute{Dept. of Computer Architecture and Technology, University
-of Granada, Spain \and
-Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+%% \institute{Dept. of Computer Architecture and Technology, University
+%% of Granada, Spain \and
+%% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+\institute{First Institute in Somecountry \and Second University in Someothercountry}
 
@@
 %gathering users for the new experiments.
+Acknowledgements\\
+taking\\
+this\\
+space
 

### 2016-11-07 EvoPar2017/paper.tex
ooooops! sorry, had to comment that line due to compiling errors. Fixed.
Anonymizes this

--- 2016-10-28 EvoPar2017/paper.tex
@@
 
-This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
-Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
-gathering users for the new experiments.
+%This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+%Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
+%2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
+%gathering users for the new experiments.
+Acknowledgements\\
+taking\\
+this\\
+space
 

### 2016-11-13 EvoPar2017/paper.Rnw
Adds new graph refs #131

@@
 library("ggplot2")
+library("RColorBrewer")
+
 experiment.data <- data.frame(Experiment=character(),
@@
 }
+
+# New data for this paper
+
+IPs <- read.csv("../data/nodio-2016-3-17-0_per_minute.csv")
+IPs$minute <-  as.POSIXct(IPs$time, format="%Y-%m-%dT%H:%M")
+
+minutes.data <- data.frame(Experiment="17-0",                      
+                           minute=IPs$minute,
+                           IPs=IPs$IPs,
+                           PUTs=IPs$PUTs,
+                           updates=IPs$updates,
+                           ratio=IPs$updates/IPs$PUTs)
 
@@
 For every new individual sent by the clients to the cache, several
-% We have not presented the "cache",  only the server and
-% ..storing the incoming chromosomes in a data structure
-% Im changing over there to cache
-% That's been changed - JJ
 pieces of metadata were stored: a time stamp, the client IP, the chromosome and
@@
 \centering
-<<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
+<<zipf, cache=FALSE,echo=FALSE,fig.height=4>>=
 ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@@
 
+\section{Finding the keys to a good performance}
+%
+%
+\begin{figure}[ht!b]
+\centering
+<<minutedata, cache=FALSE,echo=FALSE,fig.height=6>>=
+ggplot(data=minutes.data, aes(x=minute,y=PUTs,color=ratio))+ scale_x_datetime()+geom_point()+stat_smooth()
+@ %def graph on PUTs per minute
+%
+\caption{Plotting the contributions per minute in a new and massive
+  experiment, as well as the real number as color. Lighter color
+  indicates that all PUTs contributed new individuals to the
+  experiment, darker color few to none. \label{fig:minute}}
+\end{figure}
+
 %---------------------------------------------------------------

### 2016-11-13 EvoPar2017/paper.tex
Cleaning up a bit
Adds new graph refs #131

@@
 
-%% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-%%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
-\author{First Author \and Second Author}
+% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
+%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
 
-%% \institute{Dept. of Computer Architecture and Technology, University
-%% of Granada, Spain \and
-%% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
-\institute{First Institute in Somecountry \and Second University in Someothercountry}
+% \institute{Dept. of Computer Architecture and Technology, University
+% of Granada, Spain \and
+% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+
+\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
+  \and Jet A. Nother \inst{2}}
+
+\institute{Invisible University, Frantzia \and
+Institute of Higher Studies, Zootropa}
 
@@
 
+\section{Finding the keys to a good performance}
+%
+%
+\begin{figure}[ht!b]
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/minutedata-1}
+
+\end{knitrout}
+%
+\caption{Plotting the contributions per minute in a new and massive
+  experiment, as well as the real number as color. Lighter color
+  indicates that all PUTs contributed new individuals to the
+  experiment, darker color few to none. \label{fig:minute}}
+\end{figure}
+
 %---------------------------------------------------------------
@@
 
-%This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
-%Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-%2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
-%gathering users for the new experiments.
-Acknowledgements\\
-taking\\
-this\\
-space
+% This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+% Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
+% 2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
+% gathering users for the new experiments.
+
+Several projects\\
+Taking\\
+This\\
+Much\\
+Space\\
 

### 2016-11-13 EvoPar2017/paper.Rnw
Checking internet connection

@@
 NodIO needs a fitness function to work with; in this case the classical Trap 
-function \cite{Ackley1987} has been used. % I don't think this use of the ; is correct
-% Changed to .
-The evolutionary algorithm that evaluates that function runs in the
+function \cite{Ackley1987} has been used. 
+The evolutionary algorithm that uses Trap as fitness function is
+completely run in the
 browser, but the server needs to know when the solution has been
-found, so a function that checks termination condition is also included in the server, as a configuration
+found, so a subroutine that checks if the solution been found  is also included in the server, as a configuration
 option. In all cases we are using 50 Traps, a problem

### 2016-11-13 EvoPar2017/paper.Rnw
Reviewing text and eliminating comments

@@
   still be allowed. 
-  % As I remember we kept the run_number so IPs with old runs
-  % did not interfere with newer ones. But you are saying: what if they are super slow 
-  % and always PUT when the run is already over?, we want to receive 
-  % those individuals? 
-  % "That means that if a particular IP
-  % sends the fitness 99 into the server, it will be accepted only
-  % once."  
-  % once per run?
-% JJ - once per run or experiment. But we didn't keep that run_number
-% to avoid interference, just to keep track
-  %  but different individuals could have the same fitness, are we limiting the search? 
-% We are actually checking for the same individual, not only the fitness
-  %  maybe we could have multiple versions of the pool a la MVCC  :)
-% Let's leave that for the future...- JJ
-  % I think if we are just trying to limit overlapping of the experiments
-  % then PUTs should send the run's seq_number and completly remove it.  
-% The client does not keep track of the seq number. Maybe we should do
-% just that, but sending back and forth the seq number implies
-% refactoring. - JJ
 \end{enumerate}
@@
 PaaS, which provides a free tier, making the whole experiment cost
-equal to \$0.00. In fact, the NodIO system can be deployed to any
-system as long as it can run node.js and has access to the filesystem, for
-log files, if available%% . All these logs are available upon
-%% request. Processed files are available in the same repository that
-%% hosts this paper, as well as the scripts needed to process them.
+equal to nil. In fact, the NodIO system can be deployed to any
+Platform or Infraestructure as a Service 
+system as long as it can run node.js and can store logs on the same
+place or in some other Log as a Service. Local storage can be used,
+but it is not a requisite and, in fact, Heroku does not allow
+persistent storage of files in its service. 
+All these logs are available upon
+request. Processed files are available in the same repository that
+hosts this paper, as well as the scripts needed to process them.
 

### 2016-11-13 EvoPar2017/paper.Rnw
Adds references to intro refs #132

@@
 \title{A performance assessment of evolutionary algorithms in
-  volunteer computing environments}
+  volunteer computing environments: the importance of entropy}
 
@@
 their own machine to contribute to a common effort. If the program is
-embedded in a web page, collaboration is easy, but also ephemeral. In
+embedded in a web page, collaboration is straightforward, but also ephemeral. In
 this paper, we analyze a volunteer evolutionary computing system called
-NodIO, trying to find out a few rules that encourage volunteer
-participation and how to make the most of their contributions to the
-evolutionary algorithm. We will show different measures of
+NodIO, by running several experiments, some of them massive. Our
+objective is to discover rules that encourage volunteer
+participation and also the interplay of these contributions with the
+dynamics of the algorithm itself, making it more or less efficient. We will show different measures of
 participation and contribution to the algorithm and how different 
-volunteer usage patterns and tweaks in the algorithm, such as restarting clients when solution has been found, contribute to
-improvements and leveraging of these contributions. 
+volunteer usage patterns and tweaks in the algorithm, such as
+restarting clients when solution has been found, contribute to 
+improvements and leveraging of these contributions. We will also try
+to find out what is the key factor in the early termination of the
+experiments, measuring entropy in the contributions and other large
+scale indicators. 
 \end{abstract}
@@
 
-The world wide web provides not only a platform for content
-distribution, but also an increasingly reliable and high-performance
-operating system for running distributed applications. This is based
-in two key things: the Javascript virtual machine every browser runs
-and a simplified standard interface for interacting with servers.
+The World Wide Web provides not only a platform for content
+distribution, but also, thanks to the maturity and reliability of the
+HTTP protocol, an increasingly reliable and high-performance
+operating system for running distributed applications. Besides the
+protocol itself, there are two factors that contribute to this fact:
+the Javascript virtual machine every browser runs \cite{paulson2005building}
+and the simplified standard interface for interacting with servers
+exemplified by the REST application interface convention \cite{masse2011rest}.
 Thus creating a distributed computing experiment is just a matter of

### 2016-11-13 EvoPar2017/paper.Rnw
Reviewed state of the art
Someone like @mariosky should try to review this, maybe adding a bit more. EvoStar allows up to 16 pages.

@@
 
-Volunteer computing involves a user running a program voluntarily
+Volunteer computing involves users deciding to run a program that acts
+as a client or as a peer in an experiment
 and, as such, has been deployed in many different ways from the
@@
 frameworks such as the one we are using here, and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15}.
+was described in [citation hidden]. 
+%\cite{DBLP:conf/gecco/GuervosG15}.
 
-JavaScript can be used for either unwitting
+Several authors have already described system using JavaScript  either
+for unwitting
 \cite{unwitting-ec,boldrin2007distributed} or volunteer 
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}. Most
+of these papers focused on testing the system by itself as proofs of
+concepts, more than on its performance or their actual influence on
+the result of an evolutionary algorithm. In fact, some of them do not
+even work in as a distributed computing system
+\cite{EvoStar2014:jsEO}, describing the expressive capabilities of
+JavaScript for evolutionary computation.
 
-In general, those papers go no further than a proof of concept,
-attempting to gauge how many users join the effort and also how long
-would it take for a concrete amount of users to reach the solution. In
-general, they do not try to assess the performance of the algorithm
+The proof of concepts described above do not go further than trying to
+find out  how many users join the effort and how many the system is
+able to support. In fact, systems described in
+\cite{gecco07:workshop:dcor} had serious scaling issues; some of them
+also tried to find out how much time was needed to find the solution
+or, alternatively, how many users would be needed to be competitive
+with single-user single-computer implementations of the same
+algorithm. However, the latest systems try to enphasize the seamless
+integration of peer and volunteer computing systems \cite{leclerc2016seamless}.
+
+This latest system offloads evaluation to browsers so the actual
+algorithm runs in some other place; in general, most researchers do
+not try to assess the performance of the algorithm 
 itself in the precise conditions they find in a volunteer computing

### 2016-11-14 EvoPar2017/paper.Rnw
Adds the entropy graph

@@
                            ratio=IPs$updates/IPs$PUTs)
+
+# Entropy data
+new.data <- read.csv("../data/nodio-2016-3-17-0_by_exp.csv")
+new.data <- new.data[new.data$actualPUTs>=16,] # Filters short experiments (uses old data)
+new.experiment.data <- data.frame( Experiment=rep("17-0", length(new.data$IPs)),                   
+                                  milliseconds=as.integer(new.data$milliseconds),
+                                  IPs=as.integer(new.data$IPs),
+                                  actualIPs=as.integer(new.data$actualIPs),
+                                  PUTs=as.integer(new.data$PUTs),
+                                  actualPUTs=as.integer(new.data$actualPUTs),
+                                  PUTsIP=as.integer(new.data$PUTs)/as.integer(new.data$IPs),
+                                  entropy=as.numeric(new.data$entropy),
+                                  cacheEntropy=as.numeric(new.data$cacheEntropy)
+                                  )
 
@@
 
+In order to have enough data to perform more analysis on it, a new
+experiment was performed along several days with, eventually, several
+hundreds of persons participating on it. It was announced by several
+popular Twitter figures and found its way into Reddit. The number of
+PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
+are due to new announcements. Please check also the dot color, which
+corresponds to the {\em ratio} or, as indicated above, the actual
+number of contributing PUTs, those that have not been discarded due to
+repeated sending by clients or (briefly) belonging to the previous
+instance of the experiment. It reveals that, in some cases where the
+amount of PUTs really increases, they are actually not contributing so
+much, while light dots, corresponding to a ration close to 1, usually
+stay in the lowest part of the graphic.
+
+This is interesting in the sense that in indicates that there might be
+something more to performance that just the number of volunteers
+contributing.
+%
+\begin{figure}[ht!b]
+\centering
+<<entropy, cache=FALSE,echo=FALSE>>=
+max <- max(new.experiment.data$IPs)
+exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
+ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
+@ %def entropy data
+%
+\caption{Entropy of the user contribution vs. time to solution in the
+  volunteer computing experiment. \label{fig:entropy}}
+\end{figure}
+
 %---------------------------------------------------------------

### 2016-11-14 EvoPar2017/paper.tex
Adds the entropy graph

@@
 
+In order to have enough data to perform more analysis on it, a new
+experiment was performed along several days with, eventually, several
+hundreds of persons participating on it. It was announced by several
+popular Twitter figures and found its way into Reddit. The number of
+PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
+are due to new announcements. Please check also the dot color, which
+corresponds to the {\em ratio} or, as indicated above, the actual
+number of contributing PUTs, those that have not been discarded due to
+repeated sending by clients or (briefly) belonging to the previous
+instance of the experiment. It reveals that, in some cases where the
+amount of PUTs really increases, they are actually not contributing so
+much, while light dots, corresponding to a ration close to 1, usually
+stay in the lowest part of the graphic.
+
+This is interesting in the sense that in indicates that there might be
+something more to performance that just the number of volunteers
+contributing.
+%
+\begin{figure}[ht!b]
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/entropy-1}
+
+\end{knitrout}
+%
+\caption{Entropy of the user contribution vs. time to solution in the
+  volunteer computing experiment. \label{fig:entropy}}
+\end{figure}
+
 %---------------------------------------------------------------

### 2016-11-14 EvoPar2017/paper.Rnw
Adds entropy and references

@@
 something more to performance that just the number of volunteers
-contributing.
+contributing. In this experiment we include an additional measure:
+user-contribution entropy, which is measure by computing the
+compression entropy of the user contributions every for the whole
+experiment.  This is computed converting the sequence of the number of
+PUTs into a string and then compressing it using {\tt Zlib}, a
+standard compression library. Since the compression rate is related to
+the randomness of the sequence, a higher compression rate will
+correspond to an experiment with a low entropy, while a lower
+compression rate will indicate that there has been a certain amount of
+unexpected changes during the experiment. Although less popular than
+Jenssen-Shannon entropy, this way of measuring it is widely used
+\cite{ornstein1993entropy} in case where we want to compute the
+entropy of sequences of events, such as arrhythmias
+\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
+range of the compression entropy will vary between 0 (lower, or worse
+compression rate) and 1 (no compression). 
 %

### 2016-11-14 EvoPar2017/paper.Rnw
Adds entropy linear model

@@
 \centering
-<<zipf, cache=FALSE,echo=FALSE,fig.height=4>>=
+<<zipf, cache=FALSE,echo=FALSE,fig.height=6>>=
 ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@@
 PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
-are due to new announcements. Please check also the dot color, which
+are due to new announcements in social networks, usually; this is one
+of the problems with this kind of volunteer computing systems: they
+feed on the continuous influx of new users; however, this can be
+automatized and also encouraged, and, as shown, once enough users have
+participated it is sometimes easier to find new ones.
+
+In Figure \ref{fig:minute} check
+also the dot color, which 
 corresponds to the {\em ratio} or, as indicated above, the actual
@@
 stay in the lowest part of the graphic. 
+%
+\begin{figure}[ht!b]
+\centering
+<<entropy, cache=FALSE,echo=FALSE>>=
+max <- max(new.experiment.data$IPs)
+entropy.lm <- lm( new.experiment.data$milliseconds ~ new.experiment.data$entropy )
+exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
+ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
+@ %def entropy data
+%
+\caption{Entropy of the user contribution vs. time to solution in the
+  volunteer computing experiment. The fit to a linear model is significant. \label{fig:entropy}}
+\end{figure}
 
@@
 compression rate) and 1 (no compression).
-%
-\begin{figure}[ht!b]
-\centering
-<<entropy, cache=FALSE,echo=FALSE>>=
-max <- max(new.experiment.data$IPs)
-exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
-ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
-@ %def entropy data
-%
-\caption{Entropy of the user contribution vs. time to solution in the
-  volunteer computing experiment. \label{fig:entropy}}
-\end{figure}
 

### 2016-11-14 EvoPar2017/paper.tex
Adds entropy linear model

@@
 stay in the lowest part of the graphic. 
+%
+\begin{figure}[ht!b]
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/entropy-1}
+
+\end{knitrout}
+%
+\caption{Entropy of the user contribution vs. time to solution in the
+  volunteer computing experiment. \label{fig:entropy}}
+\end{figure}
 
@@
 compression rate) and 1 (no compression).
-%
-\begin{figure}[ht!b]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/entropy-1}
-
-\end{knitrout}
-%
-\caption{Entropy of the user contribution vs. time to solution in the
-  volunteer computing experiment. \label{fig:entropy}}
-\end{figure}
 

### 2016-11-14 EvoPar2017/paper.Rnw
Drawing some early conclusions

@@
 
+The graph that plots the time to solution of every instance of the
+experiment vs. entrop is shown in \ref{fig:entropy}. There is a clear
+trend towards experiments with lower entropy taking more time and
+experiments with higher entropy taking more time, which indeed can be
+statistically fit to a linear model. This dependence on entropy is
+much clearer than the assumed dependence on the number of users, which
+is represented in the graph by dot color, and even the number of PUTs
+or contributions, which is represented by dot size. You can see a
+clearly blue and big dot, meaning more than 40 different IPs
+participating and around 1500 PUTs in the high end of the graph,
+showing that it took the longest to find the solution. On the other
+hand, clearly red and small dots are on the lowest end of the graph,
+indicating that it took a short time to find the solution as long as
+the entropy is high.
+
+This is a very significant result which has many implications on
+volunteer computing experiment design, and is clearly something that
+has to do more with the fact that we are dealing with an evolutionary
+algorithm where creating and maintaining diversity is an important
+part of its exploration component, than with the fact that it is a
+volunteer computing experiment. This is why we claim that volunteer
+computing setups are, in some cases, a better option than synchronous
+and distributed implementations with a similar computing capability,
+because if diversity is kept high, the performance achieved can be
+much better than in similar synchronous and parallel experiments.
+
 %---------------------------------------------------------------

### 2016-11-14 EvoPar2017/paper.Rnw
Starting with conclusions

@@
 algorithm in a volunteer computing environment to improve its
-efficiency. This initial configuration included
-{\em overlaps} or clients running islands that had actually started in
-a previous experiment in the set. If we are interested in measuring
-how many users or their behavior in bulk this is not so important, but
-in this case our focus was on the algorithm so we had to try that all
-clients participating in it started more or less at the same time and
+efficiency and eventually find out what this efficiency depends
+on. The baseline configuration that we had used in other papers  included
+{\em overlaps} or volunteers running islands that had actually started in
+a previous experiment instance in the set. In previous papers, we were
+interested in measuring the number of users or their behavior in bulk
+and this overlap was not so important, but
+in this case our focus was on the evolutionary algorithm so we had to
+make all volunteers in it started more or less at the same time and
 from a random population. In fact, we managed to do so, except in some
@@
 method to detect the beginning of a new run is much more effective and
-will be kept in the future.
+will be kept in the future. But the fact that this happened was due to
+the {\em entropy} as proved in the last section of this paper.
 

### 2016-11-14 EvoPar2017/paper.Rnw
Intro

@@
 volunteer usage patterns and tweaks in the algorithm, such as
-restarting clients when solution has been found, contribute to 
+restarting clients when a solution has been found, contribute to 
 improvements and leveraging of these contributions. We will also try
@@
 protocol itself, there are two factors that contribute to this fact:
-the Javascript virtual machine every browser runs \cite{paulson2005building}
+the JavaScript virtual machine every browser runs \cite{paulson2005building}
 and the simplified standard interface for interacting with servers
@@
 making a JavaScript application interchange information with a server,
-called REST. From the point of view of the programmer, this involves
-relatively common skills and no special libraries, since that
+by using REST. From the point of view of the programmer, this involves
+relatively common skills and no special libraries, since the
 interface is built in the browser, and a simple application that
 responds to those requests on the server side; both involve just a few
-dozens lines of code additionally to whatever business logic your
+dozens lines of code additionally to whatever business logic the
 application has. But, more importantly and from the point of view of
@@
 
-Using this system for creating distributed experiments is called {\em
+Using this approach for creating distributed experiments is called {\em
   volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing

### 2016-11-14 EvoPar2017/paper.Rnw
SoA
The SPACE framework

@@
 
-Several authors have already described system using JavaScript  either
+Several authors have already described systems using JavaScript  either
 for unwitting
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}. Most
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.
+
+Using a peer to peer approach, the SPACE framework \cite{Leclerc2016} distributes fitness 
+evaluations across a heterogeneous pool of cloud compute nodes and volunteer
+peer computers running a browser. In SPACE peers stablis a bi-directional 
+communication with the server using the Socket.io JavaScript library.  
+As an experiment they provide an ASM.js compiled version of the RoboGen 
+software platform, showing that JavaScript can be used in a broad 
+spectrum of applications.
+
+Most
 of these papers focused on testing the system by itself as proofs of

### 2016-11-15 EvoPar2017/paper.Rnw
crowd-proposed symbols
New work added

@@
 spectrum of applications.
+
+Recent works have been using crowdsourcing in order to train robots 
+for human interaction, a representative work of this trend is that of  
+Anetsberger & Bongard in which they propose training simulated robots 
+the grounding of language symbols. They do this by using human 
+observers that issue arbitrary commands to these robots via the web, 
+providing positive or negative reinforcement in response to the
+robot’s resulting actions. For interaction, they use the Twitch 
+video streaming platform where users can issue commands through the
+platform's integrated chat service. Subjects were incentivized to 
+interact with the system by GUI features that provided participants
+with a sense of involvement with the simulation. This kind of systems
+could add additional dynamics, because users can collaborate or 
+even compete with each other when issuing their commands.
 

### 2016-11-15 EvoPar2017/paper.Rnw
bib added

@@
 for human interaction, a representative work of this trend is that of  
-Anetsberger & Bongard in which they propose training simulated robots 
-the grounding of language symbols. They do this by using human 
+Anetsberger & Bongard \cite{anetsbergerrobots} in which they propose 
+training simulated robots  the grounding of language symbols. They do this by using human 
 observers that issue arbitrary commands to these robots via the web, 

### 2016-11-15 EvoPar2017/paper.Rnw
More bib

@@
 
-Using a peer to peer approach, the SPACE framework \cite{Leclerc2016} distributes fitness 
+Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness 
 evaluations across a heterogeneous pool of cloud compute nodes and volunteer

### 2016-11-15 EvoPar2017/paper.Rnw
Addressing reviewer 3 comment
2) This part doesn't make sense:

"In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system which
client is embedded in the browser via JavaScript. We call this system
NodIO."

It's either a general statement, or it's a system called NodIO. It can't
be both. Please tidy up the wording.

@@
 
-In general, a distributed volunteer-based evolutionary computation
-system based on the browser is simply a client-server system
-which client is embedded in the browser via JavaScript. We call this
-system {\sf NodIO}. All parts of the framework are free and available
+In this section we present {\sf NodIO}, a distributed volunteer-based 
+evolutionary computation system immplemented in JavaScript. 
+All parts of the framework are free and available
 with a free license from \url{https://ANONYMOUS-URL}.

### 2016-11-15 EvoPar2017/paper.Rnw
Reviewer 3
1) p2: "...in our previous work" needs a citation
reviewer 3
5) Fig. 1 calls it a "proposed system", but it is not just proposed --
   you have actually built it!
R3:
7) What is the difference between a techno-social system and a
   socio-technical system? The latter seems to be used more commonly (at
   least in the literature I am familiar with). Are they different?

@@
 and maybe most important is approaching volunteer computing as a
-techno-social system \cite{vespignani2009predicting} which integrates
+socio-technical system \cite{vespignani2009predicting} which integrates
 user decisions and behavioral patterns in the system model; this
@@
 % Is it ok that we say that the first line of research is the most important... but then we focus on the second?
-We have approached the first issue in our previous work,
+We have approached the first issue in our previous work \cite{jj-ppsn98anonymous},
 % I have justified to focus, in this case, on the second
@@
 \includegraphics[width=3in]{system.pdf}
-\caption{Description of the proposed system. Clients execute a JavaScript EA
+\caption{Description of the NodIO system. Clients execute a JavaScript EA
   in the browser, which, every 100 generations, sends the best

### 2016-11-15 EvoPar2017/paper.Rnw
R4 Typos
Typos:

Section 4
Paragraph 1. 'Since the point of using volunteer' -> 'Since the point
of using a volunteer'
Paragraph 2. 'to the rest, specially' -> 'to the rest, especially'
Paragraph 5. 'The first thin we will' -> ''The first thing we will'
Paragraph 6. ''which so needs less' -> 'which needs less'

Section 5
Paragraph 1. 'This configuration achieve a' -> 'This configuration
achieved a'

@@
 %
-Since the point of using volunteer computing system is to cut costs,
+Since the point of using a volunteer computing system is to cut costs,
 all experiments were set up using the OpenShift
@@
 is interesting to see how some sets have a number of contributions per
-IP that are slightly superior to the rest, specially so in {\tt
+IP that are slightly superior to the rest, especially so in {\tt
   graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
@@
 %
-The first thin we will have to check is whether we have effectively eliminated these
+The first thing we will have to check is whether we have effectively eliminated these
 overlaps. In order to do that, we will have to establish a threshold
@@
 fact, as it can be seen in the last column, the median number of
-contributions per user is the lowest for {\tt cachereboot}, which so needs less contributions by users to
+contributions per user is the lowest for {\tt cachereboot}, which
+needs less contributions by users to
 find the same solution. Besides, the last row shows that {\em all}
@@
 this client configuration {\tt reboot} and {\tt cachecrash}.  This
-configuration achiever a positive secondary effect: all clients
+configuration achieves a positive secondary effect: all clients
 contribute to fill the cache, having thus an actual effect on the

### 2016-11-15 EvoPar2017/paper.Rnw
Practical commit just to change workplace

@@
 participation and also the interplay of these contributions with the
-dynamics of the algorithm itself, making it more or less efficient. We will show different measures of
+dynamics of the algorithm itself, making it more or less efficient. 
+We will show different measures of
 participation and contribution to the algorithm and how different 
@@
   size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
-  capacity.  
+  its maximum capacity.  
 \item A client that includes the evolutionary algorithm as
@@
   other graphs, some
-  additional links, and information on the experiment. This code runs
+  additional links, and information on the experiment. 
+  This code runs
   an evolutionary algorithm {\em island} that starts with a random
@@
 browser, but the server needs to know when the solution has been
-found, so a subroutine that checks if the solution been found  is also included in the server, as a configuration
+found, so a subroutine that checks if the solution been found  
+is also included in the server, as a configuration
 option. In all cases we are using 50 Traps, a problem
@@
 For every new individual sent by the clients to the cache, several
-pieces of metadata were stored: a time stamp, the client IP, the chromosome and
+pieces of metadata are stored: a time stamp, the client IP, the chromosome and
 fitness value, the cache size in that particular moment, and also if that

### 2016-11-15 EvoPar2017/paper.Rnw
Correcting typing errors and stuff

@@
 for human interaction, a representative work of this trend is that of  
-Anetsberger & Bongard \cite{anetsbergerrobots} in which they propose 
+Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose 
 training simulated robots  the grounding of language symbols. They do this by using human 
@@
 providing positive or negative reinforcement in response to the
-robot’s resulting actions. For interaction, they use the Twitch 
+robot\'s resulting actions. For interaction, they use the Twitch 
 video streaming platform where users can issue commands through the

### 2016-11-15 EvoPar2017/paper.tex
Correcting typing errors and stuff

@@
 participation and also the interplay of these contributions with the
-dynamics of the algorithm itself, making it more or less efficient. We will show different measures of
+dynamics of the algorithm itself, making it more or less efficient. 
+We will show different measures of
 participation and contribution to the algorithm and how different 
 volunteer usage patterns and tweaks in the algorithm, such as
-restarting clients when solution has been found, contribute to 
+restarting clients when a solution has been found, contribute to 
 improvements and leveraging of these contributions. We will also try
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}. Most
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.
+
+Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness 
+evaluations across a heterogeneous pool of cloud compute nodes and volunteer
+peer computers running a browser. In SPACE peers stablis a bi-directional 
+communication with the server using the Socket.io JavaScript library.  
+As an experiment they provide an ASM.js compiled version of the RoboGen 
+software platform, showing that JavaScript can be used in a broad 
+spectrum of applications.
+
+Recent works have been using crowdsourcing in order to train robots 
+for human interaction, a representative work of this trend is that of  
+Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose 
+training simulated robots  the grounding of language symbols. They do this by using human 
+observers that issue arbitrary commands to these robots via the web, 
+providing positive or negative reinforcement in response to the
+robot\'s resulting actions. For interaction, they use the Twitch 
+video streaming platform where users can issue commands through the
+platform's integrated chat service. Subjects were incentivized to 
+interact with the system by GUI features that provided participants
+with a sense of involvement with the simulation. This kind of systems
+could add additional dynamics, because users can collaborate or 
+even compete with each other when issuing their commands.
+
+Most
 of these papers focused on testing the system by itself as proofs of
@@
 \caption{Entropy of the user contribution vs. time to solution in the
-  volunteer computing experiment. \label{fig:entropy}}
+  volunteer computing experiment. The fit to a linear model is significant. \label{fig:entropy}}
 \end{figure}

### 2016-11-15 EvoPar2017/paper.Rnw
Added reference to HTCondor
Closes #133

@@
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \cite{david-seti:home}. However
+processing extraterrestrial signals \cite{david-seti:home} or
+high-throughput queuing system that such as HTCondor \cite{HTCondor}.
+However
 the dual facts of the introduction of JavaScript as a universal language for the
@@
 frameworks such as the one we are using here, and whose first version
-was described in [citation hidden]. 
+was described in [citation hidden]. Systems based on the
+JavaScript/browser combination
+emphasize the ephemerality, ease of use and universality, while
+systems such as HTCondor or BOINC might be more adequate for work that
+require higher availability of volunteer resources, achieving this
+with downloadable clients, in some cases like HTCondor needing
+administration access to the resources.
 %\cite{DBLP:conf/gecco/GuervosG15}.

### 2016-11-15 EvoPar2017/paper.Rnw
some corrections

@@
   reboot}. This size of the cache was not really noticed by some slow clients or too fast runs, so that in the
-last case, labeled {\tt cachecrash}, clients rebooted if the cache
+last case, is labeled {\tt cachecrash}, clients rebooted if the cache
 collapsed by more than a certain length, indicating a cache that had
 been voided and was filling again. Our objective was twofold: to
-encourage engagement by the users and also to increase the involvement
+encourage user engagement and also to increase the involvement
 of every client in the common experiment, eliminating at the same time
@@
 means that one objective has been reached: if clients reboot when they
-detect changes in the cache that imply a new run, overlaps are almost
+detect changes in the cache implying a new run, this way overlaps are almost
 eliminated. Going back to Table \ref{tab:runs}, we can see than the
@@
 increasing the contribution of clients to the common pool. However, we
-should remark that the clients will still stop contributing if they
+should remark that some clients will still stop contributing if they
 get stuck in a particular solution, so the ratio keeps being low
@@
 strategy 30\% of users do not. This might be due to the fact that {\tt
-  reboot} only does so when cache size is less than one, indicating
+  reboot} is done only when the cache size is less than one, indicating
 that this strategy is only partly successful in avoiding overlaps.
 
-In fact, Table \ref{tab:correlation}, which shows
+In fact, Table \ref{tab:correlation}, shows
 the correlation between the number of IPs from one minute to the
@@
 The graph that plots the time to solution of every instance of the
-experiment vs. entrop is shown in \ref{fig:entropy}. There is a clear
+experiment vs. entropy is shown in \ref{fig:entropy}. There is a clear
 trend towards experiments with lower entropy taking more time and

### 2016-11-15 EvoPar2017/paper.Rnw
Finishing first draft

--- 2016-11-15 EvoPar2017/paper.Rnw
@@
 for unwitting
-\cite{unwitting-ec,boldrin2007distributed} or volunteer 
+\cite{unwitting-ec,boldrin2007distributed,apolonia2012enhancing} or volunteer 
 \cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}. In fact, this last paper \cite{pan2015gray} performs
+an analysis of what it calls {\em Gray computing} doing an analysis of
+feasibility, cost-effectiveness, change in the user experience and
+architectural optimization needed, concluding that the computing power
+available can be vast and it that it can be cost-effective to use it.
 
@@
 
-This future work holds many challenges which once again will go in two
-different directions: gathering more users and improving the
+We have also proved in this paper, through an experiment with a
+massive number of users, that in fact the single factor predicting the
+time needed to reach the solution is the compression entropy, meaning
+that it is not so much the raw number of users, but the fact that they
+contribute fit individuals to the pool in such a way that it increases
+its diversity early in the experiment; that is, it is much better a
+few users with different types of computers and browsers contributing
+at the same time that more users with exactly the same computer and
+browser contributing more or less synchronously. This also implies
+that it is obviously better to manage as many users as possible to
+contribute at the same time and at the beginning of the experiment,
+which might not really be possible; however, the reboot strategy used
+in the latest versions really contributed to this early boost in
+diversity. It is interesting also to note that this boost in
+performance due to client contribution diversity is due to the fact
+that we are running an evolutionary algorithm, where diversity is an
+essential component.
+
+These last result has also got implications for architecture
+design. For instance, it might be interesting to combine clients in
+the browser with other kind of clients that contribute mainly to
+diversity. In fact, early experiments with other type and more
+difficult problem point in this direction. Besides, user-interaction
+strategies attracting users to experiment instances at the very
+beginning, like announcing it in social media, might help too to
+achieve higher performance.
+
+This is one of the changes we intend to explore in the future, which
+holds many challenges, such as gathering more users and improving the
 efficiency of the algorithm run in an asynchronous way by the
@@
 be relaxed. Some other policies could also be tested, as well as
-running heterogeneous algorithms in the same way that was done in
+running heterogeneous algorithms in the same way that was done by
+Valdez et al. in 
 \cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other

### 2016-11-15 EvoPar2017/paper.Rnw
Sentence corrected
@Jj please check if this is the idea.

--- 2016-11-15 EvoPar2017/paper.Rnw
@@
 contributing. In this experiment we include an additional measure:
-user-contribution entropy, which is measure by computing the
-compression entropy of the user contributions every for the whole
+user-contribution entropy, which is measured by computing the
+compression entropy of every user contribution for the whole %please check 
 experiment.  This is computed converting the sequence of the number of

### 2016-11-15 EvoPar2017/paper.Rnw
minor changes

@@
 indicating that it took a short time to find the solution as long as
-the entropy is high.
+the entropy was high.
 
@@
 algorithm in a volunteer computing environment to improve its
-efficiency and eventually find out what this efficiency depends
-on. The baseline configuration that we had used in other papers  included
+efficiency and eventually find out on what this efficiency depends
+on. The baseline configuration that we had used in other papers 
+\cite{jj-ppsn98anonymous} included
 {\em overlaps} or volunteers running islands that had actually started in

### 2016-11-15 EvoPar2017/paper.tex
Reviewed abstact and Introduction

--- 2016-11-15 EvoPar2017/paper.tex
@@
 participation and also the interplay of these contributions with the
-dynamics of the algorithm itself, making it more or less efficient. 
+dynamics of the algorithm itself, making it more or less efficient. % efficient enough? Somehow saying "more or less efficient" does not seem "our work is awesome" to me.
 We will show different measures of
-participation and contribution to the algorithm and how different 
+participation and contribution to the algorithm, as well as how different 
 volunteer usage patterns and tweaks in the algorithm, such as
@@
 Using this approach for creating distributed experiments is called {\em
-  volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing
+  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
 \cite{sarmenta2001volunteer} and it dates back, in different shapes
@@
 here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98}.
+computation, as our group has done for a long time \cite{jj-ppsn98}. % Not anonymised
 
@@
 contributions and is able to find
-the solution to the problem in a minimum time and with the least
+the solution to the problem in a minimum time, with the least
 number of contributions.

### 2016-11-15 EvoPar2017/paper.tex
Anonymising refs again

@@
 here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98}. % Not anonymised
+computation, as our group has done for a long time \cite{jj-ppsn98anonymous}.
 
@@
 administration access to the resources.
-%\cite{DBLP:conf/gecco/GuervosG15}.
+%\cite{DBLP:conf/gecco/GuervosG15anonymous}.
 

### 2016-11-15 EvoPar2017/paper.tex
Reviewed the state of the art, fixed some typos. References #132

@@
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \cite{david-seti:home} or
-high-throughput queuing system that such as HTCondor \cite{HTCondor}.
-However
+processing extraterrestrial signals \cite{david-seti:home}, or a 
+high-throughput queuing system such as HTCondor \cite{HTCondor}.
+However,
 the dual facts of the introduction of JavaScript as a universal language for the
-browser and the browser itself as an ubiquitous web and Internet client has
+browser and the browser itself as both an ubiquitous web and Internet client has
 made this combination the most popular for volunteer computing
 frameworks such as the one we are using here, and whose first version
-was described in [citation hidden]. Systems based on the
+was described in \cite{DBLP:conf/gecco/GuervosG15anonymous}. Systems based on the
 JavaScript/browser combination
-emphasize the ephemerality, ease of use and universality, while
-systems such as HTCondor or BOINC might be more adequate for work that
+emphasize the ephemerality, ease of use, and universality, while
+systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
 require higher availability of volunteer resources, achieving this
-with downloadable clients, in some cases like HTCondor needing
+with downloadable clients which in some cases, like HTCondor, need
 administration access to the resources.
-%\cite{DBLP:conf/gecco/GuervosG15anonymous}.
 
@@
 evaluations across a heterogeneous pool of cloud compute nodes and volunteer
-peer computers running a browser. In SPACE peers stablis a bi-directional 
-communication with the server using the Socket.io JavaScript library.  
-As an experiment they provide an ASM.js compiled version of the RoboGen 
+peer computers running a browser. In SPACE peers establish a bi-directional 
+communication with the server using the \texttt{Socket.io} JavaScript library.  
+As an experiment they provide an \texttt{ASM.js} compiled version of the RoboGen 
 software platform, showing that JavaScript can be used in a broad 
@@
 Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose 
-training simulated robots  the grounding of language symbols. They do this by using human 
+training simulated robots for the grounding of language symbols. They do this by using human 
 observers that issue arbitrary commands to these robots via the web, 
@@
 
-The proof of concepts described above do not go further than trying to
+The proofs of concept described above do not go further than trying to
 find out  how many users join the effort and how many the system is
@@
 with single-user single-computer implementations of the same
-algorithm. However, the latest systems try to enphasize the seamless
+algorithm. However, the latest systems try to emphasize the seamless
 integration of peer and volunteer computing systems \cite{leclerc2016seamless}.
@@
 is actually run on the server, using the browser mainly for fitness
-evaluation so no actual contribution to the evaluation of the
+evaluation, and thus no actual contribution to the evaluation of the
 evolutionary algorithm itself is made.

### 2016-11-15 EvoPar2017/paper.tex
Reviewed framework description

@@
 
-The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
+The architecture of the {\sf NodIO} system is divided in two tiers:
+\begin{enumerate}
 \item A REST (representational state transfer) server that responds to
@@
   encoding, in this case chromosomes and related
-  information sent by the clients and responses returned by the
+  information sent by the clients, and responses returned by the
   server.  The server has the capability to
@@
   inserted once into the cache. That means that if a particular IP
-  sends the fitness 99 into the server, it will be accepted only
+  sends a fitness value of 99 into the server, it will be accepted only
   once. The main intention of this policy was to avoid {\em
@@
     old} run of the experiment contributing after a new one had
-  started. This did not completely avoid these contributions: if the
+  started. This however did not completely avoid these contributions: if the
   clients of the {\em old} run increased their new fitness, it would
@@
 
-NodIO needs a fitness function to work with; in this case the classical Trap 
+NodIO needs a fitness function to work with, in this case the classical Trap 
 function \cite{Ackley1987} has been used. 
@@
 completely run in the
-browser, but the server needs to know when the solution has been
-found, so a subroutine that checks if the solution been found  
-is also included in the server, as a configuration
-option. In all cases we are using 50 Traps, a problem
+browser, but the server also includes, as a configuration
+option, a subroutine that checks if the solution has been
+found. In all cases we are using 50 Traps, a problem
 that is difficult enough to need the intervention of several users for
-an extended amount of minutes.
-For every new individual sent by the clients to the cache, several
+an extended amount of minutes. % Why? Do we have results on a single computer or on a cluster to compare?
+For every new individual the clients send to the cache, several
 pieces of metadata are stored: a time stamp, the client IP, the chromosome and
@@
 used for announcements too. We also encourage friends and followers to
-put an original post for announcing it, instead of just sharing, in
-the way allowed by the platform, the original post. During the running
+put a self-authored post for announcing it, instead of just sharing, in
+the way allowed by the platform, our original posts. During the running
 of the experiments, users are engaged, questions answered, and all

### 2016-11-15 EvoPar2017/paper.tex
Adding Twitter example

@@
 used for announcements too. We also encourage friends and followers to
-put a self-authored post for announcing it, instead of just sharing, in
-the way allowed by the platform, our original posts. During the running
+put their own posts for announcing it, instead of just sharing, in
+the way allowed by the platform, our original ones.
+For instance, when using Twitter, to include the URL of the experiment inside a tweet of their own, instead of just sharing it via a re-tweet (RT).
+During the running
 of the experiments, users are engaged, questions answered, and all

### 2016-11-15 EvoPar2017/paper.tex
Couple of comments...

@@
 %
-Since the point of using a volunteer computing system is to cut costs,
+Since the point of using a volunteer computing system is to cut costs, %actually, we talk about the system being efficient, and also the algorithm itself, but I think we don't talk about the costs...
 all experiments were set up using the OpenShift
@@
 equal to nil. In fact, the NodIO system can be deployed to any
-Platform or Infraestructure as a Service 
-system as long as it can run node.js and can store logs on the same
+Platform or Infrastructure as a Service 
+system as long as it can run \texttt{node.js} and can store logs on the same
 place or in some other Log as a Service. Local storage can be used,
-but it is not a requisite and, in fact, Heroku does not allow
+but it is not a requisite and, in fact, Heroku %first time Heroku is mentioned...
+ does not allow
 persistent storage of files in its service. 

### 2016-11-15 EvoPar2017/paper.Rnw
typo spotted
Couple of questions for @JJ related with the conclusion section. Related with issue #127

@@
 
-This is interesting in the sense that in indicates that there might be
+This is interesting in the sense that it indicates that there might be
 something more to performance that just the number of volunteers
@@
 interested in measuring the number of users or their behavior in bulk
-and this overlap was not so important, but
-in this case our focus was on the evolutionary algorithm so we had to
+and this overlap was not so important.
+In this case our focus was on the evolutionary algorithm so we had to
 make all volunteers in it started more or less at the same time and
-from a random population. In fact, we managed to do so, except in some
+from a random population. %Pablo: too long paragraph, added commas and sepparated into two
+In fact, we managed to do so, except in some
 byzantine cases, by making the clients restart when the cache size
@@
 The last configuration managed to find, with a relatively small number
+%Pablo: what is the difference with this configurarion with respect to the others?
 of evaluations, solutions in half the time that in previous sets of
 experiments. Although in the {\tt base} and {\tt graph} the median
+%Pablo: I usually prefer to do self-contained and conclusions, this whole explanation could be moved to discussion as it is too specific. But it is up to you :)
 time and number of contributions was smaller, we cannot disregard the

### 2016-11-15 EvoPar2017/paper.tex
Reviewed results

@@
 far. Every volunteer contributed a certain number of individuals
-(represented by the PUT HTTP request) to the server. The ranked number
+(represented by the PUT HTTP request in Figure \ref{fig:system}) to the server. The ranked number
 of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
@@
   graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
-show a regime in the middle of the graph that shows that some users
+show a regime in the middle of the graph which shows that some users
 are contributing more than in other similar sets. In general, these
@@
 %
-The first thing we will have to check is whether we have effectively eliminated these
+The first thing we will have to check is whether we have effectively eliminated those
 overlaps. In order to do that, we will have to establish a threshold
@@
 is found in just a few evaluations, but that is usually not the
-case. By looking at the logs in the last case we have established that
-minimum as 16 PUTs; in less than 1600 generations no solution to that
-difficult problem will be found. Let us filter then the runs,
+case. By looking at the logs in the last case we have established the
+minimum at 16 PUTs; in less than 1600 generations no solution to that
+difficult problem will be found. % Why?
+ Let us filter then the runs,
 eliminating those that used less than 1600 generations to find the
@@
 results are shown in Table \ref{tab:filtered}. The first change is in
-the column \#Runs, that shows that most of the runs in the initial
+the column \#Runs, which shows that most of the runs in the initial
 configurations found the solution thanks to these carried over

### 2016-11-15 EvoPar2017/paper.tex
Reviewed section 5

@@
 popular Twitter figures and found its way into Reddit. The number of
-PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
-are due to new announcements in social networks, usually; this is one
+PUTs per minute are shown in \ref{fig:minute}. Usually, sudden peaks in them
+are due to new announcements in social networks, usually. This is one
 of the problems with this kind of volunteer computing systems: they
 feed on the continuous influx of new users; however, this can be
-automatized and also encouraged, and, as shown, once enough users have
+automatized and also encouraged, resulting, as shown, in that once enough users have
 participated it is sometimes easier to find new ones.
 
-In Figure \ref{fig:minute} check
-also the dot color, which 
-corresponds to the {\em ratio} or, as indicated above, the actual
+The dot color in Figure \ref{fig:minute} corresponds to the {\em ratio} or, as indicated above, the actual
 number of contributing PUTs, those that have not been discarded due to
@@
 amount of PUTs really increases, they are actually not contributing so
-much, while light dots, corresponding to a ration close to 1, usually
+much, while light dots, corresponding to a ratio close to 1, usually
 stay in the lowest part of the graphic. 
@@
 
-This is interesting in the sense that in indicates that there might be
-something more to performance that just the number of volunteers
-contributing. In this experiment we include an additional measure:
-user-contribution entropy, which is measure by computing the
-compression entropy of the user contributions every for the whole
+This is interesting in the sense that it indicates that there might be
+something more to performance than just the number of volunteers
+contributing. Therefore, in this experiment we include an additional measure:
+user-contribution entropy, which is measured by computing the
+compression entropy of every user contribution for the whole
 experiment.  This is computed converting the sequence of the number of

### 2016-11-15 EvoPar2017/paper.Rnw
Incorporates excellent fixes by @unintendedbear
Also eliminates .tex to avoid confusion. All changes should go to the .Rnw file.

@@
 here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98}.
+computation, as our group has done for a long time \cite{jj-ppsn98anonymous}.
 
@@
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \cite{david-seti:home} or
-high-throughput queuing system that such as HTCondor \cite{HTCondor}.
-However
+processing extraterrestrial signals \cite{david-seti:home}, or a 
+high-throughput queuing system such as HTCondor \cite{HTCondor}.
+However,
 the dual facts of the introduction of JavaScript as a universal language for the
-browser and the browser itself as an ubiquitous web and Internet client has
+browser and the browser itself as both an ubiquitous web and Internet client has
 made this combination the most popular for volunteer computing
 frameworks such as the one we are using here, and whose first version
-was described in [citation hidden]. Systems based on the
+was described in \cite{DBLP:conf/gecco/GuervosG15anonymous}. Systems based on the
 JavaScript/browser combination
-emphasize the ephemerality, ease of use and universality, while
-systems such as HTCondor or BOINC might be more adequate for work that
+emphasize the ephemerality, ease of use, and universality, while
+systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
 require higher availability of volunteer resources, achieving this
-with downloadable clients, in some cases like HTCondor needing
+with downloadable clients which in some cases, like HTCondor, need
 administration access to the resources.
-%\cite{DBLP:conf/gecco/GuervosG15}.
 
@@
 used for announcements too. We also encourage friends and followers to
-put an original post for announcing it, instead of just sharing, in
-the way allowed by the platform, the original post. During the running
+put their own posts for announcing it, instead of just sharing, in
+the way allowed by the platform, our original ones.
+For instance, when using Twitter, to include the URL of the experiment inside a tweet of their own, instead of just sharing it via a re-tweet (RT).
+During the running
 of the experiments, users are engaged, questions answered, and all
@@
   reboot}. This size of the cache was not really noticed by some slow clients or too fast runs, so that in the
-last case, is labeled {\tt cachecrash}, clients rebooted if the cache
+last case, labeled {\tt cachecrash}, clients rebooted if the cache
 collapsed by more than a certain length, indicating a cache that had
 been voided and was filling again. Our objective was twofold: to
-encourage user engagement and also to increase the involvement
+encourage engagement by the users and also to increase the involvement
 of every client in the common experiment, eliminating at the same time
@@
 means that one objective has been reached: if clients reboot when they
-detect changes in the cache implying a new run, this way overlaps are almost
+detect changes in the cache that imply a new run, overlaps are almost
 eliminated. Going back to Table \ref{tab:runs}, we can see than the
@@
 increasing the contribution of clients to the common pool. However, we
-should remark that some clients will still stop contributing if they
+should remark that the clients will still stop contributing if they
 get stuck in a particular solution, so the ratio keeps being low
@@
 strategy 30\% of users do not. This might be due to the fact that {\tt
-  reboot} is done only when the cache size is less than one, indicating
+  reboot} only does so when cache size is less than one, indicating
 that this strategy is only partly successful in avoiding overlaps.
 
-In fact, Table \ref{tab:correlation}, shows
+In fact, Table \ref{tab:correlation}, which shows
 the correlation between the number of IPs from one minute to the
@@
 indicating that it took a short time to find the solution as long as
-the entropy was high.
+the entropy is high.
 
@@
 algorithm in a volunteer computing environment to improve its
-efficiency and eventually find out on what this efficiency depends
-on. The baseline configuration that we had used in other papers 
-\cite{jj-ppsn98anonymous} included
+efficiency and eventually find out what this efficiency depends
+on. The baseline configuration that we had used in other papers  included
 {\em overlaps} or volunteers running islands that had actually started in
@@
 interested in measuring the number of users or their behavior in bulk
-and this overlap was not so important.
-In this case our focus was on the evolutionary algorithm so we had to
+and this overlap was not so important, but
+in this case our focus was on the evolutionary algorithm so we had to
 make all volunteers in it started more or less at the same time and

### 2016-11-15 EvoPar2017/paper.Rnw
Incorporating changes in .tex
From 2b0442adec0f8fab09604fbe139bc219d098ffcf

@@
 This is interesting in the sense that it indicates that there might be
-something more to performance that just the number of volunteers
-contributing. In this experiment we include an additional measure:
+something more to performance than just the number of volunteers
+contributing. Therefore, in this experiment we include an additional measure:
 user-contribution entropy, which is measured by computing the
-compression entropy of every user contribution for the whole %please check 
+compression entropy of every user contribution for the whole
 experiment.  This is computed converting the sequence of the number of

### 2016-11-16 EvoPar2017/paper.Rnw
Adds package
Adds @mariosky's algorithm
And this is the final one.
Closes #131
Closes #132
Closes #127
Fixed typos
and while we're at it, closes #132, closes #127

@@
 \usepackage{caption}
+\usepackage{algorithm,algorithmic}
 
@@
 In this section we present {\sf NodIO}, a distributed volunteer-based 
-evolutionary computation system immplemented in JavaScript. 
+evolutionary computation system implemented in JavaScript. 
 All parts of the framework are free and available
@@
 \end{enumerate}
+%
+\begin{algorithm}
+\caption{Algorithm for each of the clients in a web browser.\label{algorithm}}
+\begin{algorithmic}[1]
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\REQUIRE $ea$, $period$, $population\_size$
+\ENSURE $best$
+\\ 
+\textit{Initialization} :
+\STATE create a random population of $ea.population.size$ =
+$population\_size$
+\STATE generation\_count = 0
+\\ 
+\textit{EA LOOP:}
+\WHILE{$(ea.population.best \ne best\_fitness)$}
+\STATE $ea.generation()$
+\STATE $generation\_count++$
+\IF {$((generation\_count \bmod period) == 0)$}
+\STATE \texttt{GET} $random$
+\STATE ea.population.add(random)
+\STATE \texttt{PUT} ea.population.best
+\STATE \texttt{GET} statistics
+\STATE update GUI with statistics
+\ENDIF
+\ENDWHILE
+\RETURN $ea.population.best$
+\end{algorithmic}
+\end{algorithm}
 
-Figure \ref{fig:system} describes the general system architecture and
-algorithm behavior. Different JavaScript libraries, such as JQuery or {\tt
+Figure \ref{fig:system} describes the general system architecture; the 
+algorithm is described in \ref{algorithm}. Different JavaScript libraries, such as JQuery or {\tt
   Chart.js} have been used to build the user interface elements of the

### 2017-01-30 EvoPar2017/paper.Rnw
Prepares final version

@@
 
-% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
-%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
+\author{Juan-J.~Merelo\inst{1} \and Paloma de las Cuevas\inst{1} \and Pablo
+  Garc\'ia-S\'anchez \inst{2} \and Mario Garc\'ia-Valdez \inst{3}}
 
-% \institute{Dept. of Computer Architecture and Technology, University
-% of Granada, Spain \and
-% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
+\institute{Dept. of Computer Architecture and Technology, University
+of Granada, Spain and CITIC, \email{(jmerelo,palomacd)@ugr.es} 
+\and
+Department of Computer Engineering, University of C\'adiz (Spain), \email{pablo.garciasanchez@uca.es}
+\and
+Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana, \email{amherag,mario@tectijuana.edu.mx}}
 
-\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
-  \and Jet A. Nother \inst{2}}
+%% \author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
+%%   \and Jet A. Nother \inst{2}}
 
-\institute{Invisible University, Frantzia \and
-Institute of Higher Studies, Zootropa}
+%% \institute{Invisible University, Frantzia \and
+%% Institute of Higher Studies, Zootropa}
 
@@
 here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98anonymous}.
+computation, as our group has done for a long time \cite{jj-ppsn98}.
 
@@
 % Is it ok that we say that the first line of research is the most important... but then we focus on the second?
-We have approached the first issue in our previous work \cite{jj-ppsn98anonymous},
+We have approached the first issue in our previous work \cite{jj-ppsn98},
 % I have justified to focus, in this case, on the second
@@
 frameworks such as the one we are using here, and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15anonymous}. Systems based on the
+was described in \cite{DBLP:conf/gecco/GuervosG15}. Systems based on the
 JavaScript/browser combination
@@
 All parts of the framework are free and available
-with a free license from \url{https://ANONYMOUS-URL}.
-%\url{https://github.com/JJ/splash-volunteer}.
+with a free license from \url{https://github.com/JJ/splash-volunteer}.
 
@@
   Chart.js} have been used to build the user interface elements of the
-framework, which should be running in \url{http://ANONYMOUS-URL}, 
-%\url{http://nodio-jmerelo.rhcloud.com}, 
+framework, which should be running in \url{http://nodio-jmerelo.rhcloud.com}, 
 a free resource hosted in the
@@
 running heterogeneous algorithms in the same way that was done in
-\cite{DBLP:journals/grid/ValdezTGVO15anonymous}. Diversity, as in any other
+\cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other
 evolutionary algorithm, seems to be the key, but there are many ways of
@@
 essential feature of the system. That is why this paper, as well as
-the data and processing scripts, are published with a free license in GitHub at
-\url{ANONYMOUS-URL}.
-%\url{https://git.io/gecco-es-15}.
+the data and processing scripts, are published with a free license in GitHub at \url{https://git.io/gecco-es-15}.
 
@@
 
-% This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
-% Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
-% 2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
-% gathering users for the new experiments.
-
-Several projects\\
-Taking\\
-This\\
-Much\\
-Space\\
+This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
+Ministry of Economy and Competitivity). We are also grateful to {\tt @otisdriftwood} for his help
+gathering users for the new experiments.
 

### 2017-04-04 EvoPar2017/paper.Rnw
Add running stuff
Producing PNGs too for poster

@@
 library("RColorBrewer")
-
+opts_chunk$set( dev        = c("pdf", "png") )
 experiment.data <- data.frame(Experiment=character(),
@@
   volunteer computing environments: the importance of entropy} 
+%\titlerunning{A performance assessment of evolutionary algorithms in volunteer computing environments}
 
@@
   Garc\'ia-S\'anchez \inst{2} \and Mario Garc\'ia-Valdez \inst{3}}
-
+%\authorrunning{Merelo, de las Cuevas, García-Sánchez and García-Valdez}
 \institute{Dept. of Computer Architecture and Technology, University

### 2017-04-16 js/reveal.js
Adds presentation for EvoStar 2017

@@ -3,7 +3,7 @@
  * http://lab.hakim.se/reveal-js
  * MIT licensed
  *
- * Copyright (C) 2015 Hakim El Hattab, http://hakim.se
+ * Copyright (C) 2016 Hakim El Hattab, http://hakim.se
  */
 (function( root, factory ) {
 	if( typeof define === 'function' && define.amd ) {
@@ -25,10 +25,14 @@
 
 	var Reveal;
 
+	// The reveal.js version
+	var VERSION = '3.4.1';
+
 	var SLIDES_SELECTOR = '.slides section',
 		HORIZONTAL_SLIDES_SELECTOR = '.slides>section',
 		VERTICAL_SLIDES_SELECTOR = '.slides>section.present>section',
 		HOME_SLIDE_SELECTOR = '.slides>section:first-of-type',
+		UA = navigator.userAgent,
 
 		// Configuration defaults, can be overridden at initialization time
 		config = {
@@ -39,11 +43,11 @@
 			height: 700,
 
 			// Factor of the display size that should remain empty around the content
-			margin: 0.1,
+			margin: 0.04,
 
 			// Bounds for smallest/largest possible scale to apply to content
 			minScale: 0.2,
-			maxScale: 1.5,
+			maxScale: 2.0,
 
 			// Display controls in the bottom right corner
 			controls: true,
@@ -78,6 +82,9 @@
 			// Change the presentation direction to be RTL
 			rtl: false,
 
+			// Randomizes the order of slides each time the presentation loads
+			shuffle: false,
+
 			// Turns fragments on and off globally
 			fragments: true,
 
@@ -85,12 +92,15 @@
 			// i.e. contained within a limited portion of the screen
 			embedded: false,
 
-			// Flags if we should show a help overlay when the questionmark
+			// Flags if we should show a help overlay when the question-mark
 			// key is pressed
 			help: true,
 
 			// Flags if it should be possible to pause the presentation (blackout)
 			pause: true,
+
+			// Flags if speaker notes should be visible to all viewers
+			showNotes: false,
 
 			// Number of milliseconds between automatically proceeding to the
 			// next slide, disabled when set to 0, this value can be overwritten
@@ -99,6 +109,9 @@
 
 			// Stop auto-sliding after user input
 			autoSlideStoppable: true,
+
+			// Use this method for navigation when auto-sliding (defaults to navigateNext)
+			autoSlideMethod: null,
 
[…]

### 2017-04-16 EvoPar2017/paper.Rnw
Adds new figures to avoid deletion

@@
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
 
@@
 library("RColorBrewer")
-opts_chunk$set( dev        = c("pdf", "png") )
+opts_chunk$set( dev        = c("pdf", "png"),
+               fig.width   = 9,
+               fig.height  = 6,
+               dpi     = 200 )
 experiment.data <- data.frame(Experiment=character(),
@@
 \centering
-<<zipf, cache=FALSE,echo=FALSE,fig.height=6>>=
+<<zipf, cache=FALSE,echo=FALSE, fig.width   = 9, fig.height  = 6>>=
 ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@@
 \centering
-<<minutedata, cache=FALSE,echo=FALSE,fig.height=6>>=
+<<minutedata, cache=FALSE,echo=FALSE, fig.width   = 9, fig.height  = 6, fig.dpi     = 100>>=
 ggplot(data=minutes.data, aes(x=minute,y=PUTs,color=ratio))+ scale_x_datetime()+geom_point()+stat_smooth()
@@
 \centering
-<<entropy, cache=FALSE,echo=FALSE>>=
+<<entropy,cache=FALSE,echo=FALSE,fig.width=9,fig.height=6,dpi=200,dev='png'>>=
 max <- max(new.experiment.data$IPs)
